### 2024-01-26 Paper/conference abstract.tex
Updates from Overleaf

\documentclass[12pt]{article}

\usepackage{tgtermes}
\usepackage{epsf}
\usepackage{epstopdf}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage[colorlinks=true,linkcolor=blue,citecolor=blue]{hyperref}
\usepackage{dcolumn}
\usepackage{amsmath, amsthm, amssymb}
\usepackage{mwe}
\usepackage{url}
\usepackage{natbib}
%\usepackage{harvard}
\usepackage{fancyheadings}
\usepackage{longtable}
\usepackage{authblk}
\usepackage{setspace}
%\usepackage[nomarkers]{endfloat}
\usepackage{float}
\usepackage{bbm}
%\usepackage{titling}
\usepackage{subcaption}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{import}
%\usepackage[nomarkers,nofiglist,notablist]{endfloat}

\onehalfspacing
\textwidth 6.5in \oddsidemargin 0in \evensidemargin -0.6in
\textheight 8.5in \topmargin -0.2in

\newcolumntype{L}[1]{>{\raggedright\let\newline\\
		\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\
		\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\
		\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{P}[1]{>{\raggedright\tabularxbackslash}p{#1}}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}

\usepackage{geometry}
 \geometry{,
 left=15mm,
 top=20mm,
 right=15mm,
 bottom=20mm
 }

\newcommand{\xsub}[1]{%
	\mbox{\scriptsize\begin{tabular}{@{}c@{}}#1\end{tabular}}%
}

%\renewcommand{\thetable}{\Roman{table}}

\begin{document}

\large 
\begin{center}
    Does Hospital Leadership Matter? 
    \normalsize \vspace{-1mm}

    Evidence from the Hospital Readmissions Reduction Program
\end{center}

\normalsize

\noindent \underline{Objective:}

An extensive literature has sought to understand the objectives and behaviors of firms that are not classic for-profits. Researchers have speculated a variety of objective functions that could motivate nonprofits such as maximizing prestige, income, or a quality/quantity tradeoff (Steinberg, 1986). An important contribution to this literature is the use of empirical strategies to analyze how firm behavior reveals attributes of the underlying objective function. It remains unclear what type of objective function drives nonprofit hospitals, a common focus in this literature (Erus \& Weisbrod, 2002; Deneffee \& Masson, 2002; Horwitz \& Nichols, 2009). What has not been considered is how differential characteristics within nonprofits drive objectives. For example, the composition of people making decisions on behalf of the hospital likely drives hospital behaviors. There is variation in whether hospitals employ executives with clinical experience or not, which has potential to change how much weight the hospital places on profit vs. quality of care. I contribute to our understanding of hospital behaviors by investigating whether clinical experience on an executive team affects hospital behavior after being financially penalized.

\vspace{2mm}

\noindent \underline{Methods}

Using publicly available tax forms, I construct a novel data set of nonprofit hospitals from 2009-2015 that captures the identities of executives tied to that hospital. I link this data to hospital characteristics in the American Hospital Association survey and Hospital Compare data. With these sources, I limit the sample only to hospitals penalized by the Hospital Readmissions Reduction Program (HRRP) of 2012, which reduced Medicare payments to hospitals with above average readmission rates. I further limit the sample to hospitals with stable executive teams over time and characterize hospitals as having clinical experience on their team or not. I then estimate the difference in readmission and mortality rates between hospitals with and without clinical experience before and after the penalties began, identifying the effect of having a clinical executive on a hospital’s response to the penalty. I assume that before the penalty, clinical and non-clinical team hospitals had similar readmission and mortality trends, and that the composition of hospital executive teams determined prior to the penalties are, conditional on being penalized, exogenous.

\vspace{2mm}

\noindent \underline{Results}

Preliminary evidence suggests that prior to being penalized, non-clinical team hospitals had higher readmission rates relative to clinical team hospitals. Once penalties occurred, the non-clinical team hospitals reduced readmissions at a faster rate, converging with clinical team hospital rates.

\vspace{2mm}

\noindent \underline{Discussion}

This result suggests that non-clinical executive teams respond more drastically to quality-related financial incentives than clinical teams, indicating that non-clinical teams are relatively more profit driven. Further goals include investigation of the mechanisms by which different hospitals reduce readmissions, whether it is through quality improvement or selective patient practices.

\end{document}

### 2024-01-26 Paper/paper.tex
Updates from Overleaf

\documentclass[12pt]{article}

\usepackage{tgtermes}
\usepackage{epsf}
\usepackage{epstopdf}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage[colorlinks=true,linkcolor=blue,citecolor=blue]{hyperref}
\usepackage{dcolumn}
\usepackage{amsmath, amsthm, amssymb}
\usepackage{mwe}
\usepackage{url}
\usepackage{natbib}
%\usepackage{harvard}
\usepackage{fancyheadings}
\usepackage{longtable}
\usepackage{authblk}
\usepackage{setspace}
%\usepackage[nomarkers]{endfloat}
\usepackage{float}
\usepackage{bbm}
%\usepackage{titling}
\usepackage{subcaption}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{import}
%\usepackage[nomarkers,nofiglist,notablist]{endfloat}

\onehalfspacing
\textwidth 6.5in \oddsidemargin 0in \evensidemargin -0.6in
\textheight 8.5in \topmargin -0.2in

\newcolumntype{L}[1]{>{\raggedright\let\newline\\
		\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\
		\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\
		\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{P}[1]{>{\raggedright\tabularxbackslash}p{#1}}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}

\newcommand{\xsub}[1]{%
	\mbox{\scriptsize\begin{tabular}{@{}c@{}}#1\end{tabular}}%
}

%\renewcommand{\thetable}{\Roman{table}}

\begin{document}

	\linespread{1.2}\title{\vspace{-0.5in} Does Hospital Leadership Matter?\\ \large Evidence from the Hospital Readmissions Reduction Program}

	\date{\today}

	\author{\vspace{10mm}Hanna Glenn\footnote{Department of Economics, Emory University, 1602 Fishburne Drive, Atlanta, GA 30322, hanna.glenn@emory.edu.} }

	\maketitle
	%\setlength{\droptitle}{-10pt}

	\vspace{-0.2in}

	\singlespacing\maketitle

	\begin{abstract}
		{\small

		} 
	\end{abstract}

	\vspace{0.1in}

	\noindent Keywords:

	\noindent JEL Codes:

	\onehalfspacing

	\newpage

  The objectives and behaviors of firms that are not classic for-profits are not easily determined, and an extensive literature has sought to understand these types of firms. Nonprofits are thought to care about more than just monetary profit, but also gain utility from furthering a mission and other unknown factors. Researchers have speculated a variety of objective functions that could motivate nonprofits such as maximizing prestige, income, or a quality/quantity tradeoff (Steinberg, 1986). An important contribution to this literature is the use of empirical strategies to analyze how firm behavior reveals attributes of the underlying objective function (add citations).

  Hospitals in particular have been a common focus when seeking to understand not-for-profit firms. This is because, first, there are different hospital ownership types within the industry, giving researchers the ability to compare the behaviors of nonprofits with that of for-profits. Second, nonprofit hospitals make up over 50\% of all hospitals in the US, and staff, on average, 207 beds. For comparison, for-profits make up around 30\% of hospitals and staff 100 beds. Thus, nonprofit behavior in the hospital industry is relevant and directly affects consumers of health care. However, it remains unclear what type of objective function drives nonprofit hospitals (Erus \& Weisbrod, 2002; Deneffee \& Masson, 2002; Horwitz \& Nichols, 2009).

  What has not been considered is how the leadership make-up within nonprofits potentially drive objectives. In large, publicly traded firms, executive characteristics are shown to be correlated with firm performance, indicating that there is more to firm behavior than just for-profit or not-for-profit status. Within nonprofit hospitals, there is variation in the propensity to hire executives with a background in clinical settings, which could reasonably change how much weight the hospital places on profit vs. quality of care in their objective function. In this paper, I contribute to our understanding of hospital behaviors by investigating whether clinical experience on an executive team affects hospital behavior after being financially penalized.

  Using publicly available tax forms, I construct a novel data set of nonprofit hospitals from 2009-2015 which captures the identities of executives tied to that hospital. I link hospitals in this data to hospital level characteristics in the American Hospital Association survey and Hospital Compare data. In order to capture the effect of clinical experience in executives, not just the selection of clinical executives into certain types of hospitals, I specifically investigate hospitals penalized by the Hospital Readmissions Reduction Program (HRRP) of 2012. This program reduced Medicare payments to hospitals with above average readmission rates. I further limit the sample to hospitals with stable executive teams over time, and characterize hospitals as having clinical experience on their team or not. I then estimate the difference in readmission and mortality rates between hospitals with and without clinical experience before and after the penalties began. I assume that before the penalty, clinical and non-clinical team hospitals had similar readmission and mortality trends, and that the composition of hospital executive teams determined prior to the penalties is, conditional on being penalized, exogenous. Under these assumptions, I identify the effect of having a clinical executive on a hospital’s response to the penalty measured in readmission and mortality rates.

    I find that

    I contribute to hospital objectives

    I contribute to HRRP

    I contribute to executives affect on firms

    \section{Background}

    \subsection{Hospital Readmissions Reduction Program}

    Avoidable readmissions are considered a bad outcome for patients. The Hospital Readmissions Reduction Program (HRRP) was enacted as a part of the Affordable Care Act to incentivize hospitals to reduce readmission rates. The goal of the program was to get hospitals to lower readmissions through better care coordination, less initial stay complications, and better post-care instructions. Beginning in October 2012, hospitals with higher readmission rates than the national average (after adjusting for demographic characteristics faced by the hospital) receive a lower reimbursement rate for all Medicare patients they treat.

    \subsection{Nonprofit Hospital Leadership}

	\section{Data}

    \subsection{Names of Hospital Leaders}

    The main data used for analysis is a novel data set of nonprofit hospitals in the US which contains names and select characteristics of executives tied to the hospital from 2009-2015. I gather this data from each hospital's publicly available Tax Form 990, which is required each year for all nonprofits in the US\footnote{something about different types of forms here}. To my knowledge, this is the first large scale gathering of leadership names from these forms. Thus, I will explain each step in the data extraction process. More details can be found in Appendix blank.

    The tax forms for all nonprofits are housed by ProPublica\footnote{https://projects.propublica.org/nonprofits/}. I use the NonProfit Explorer API to extract the Employee Identification Numbers (EIN) of all nonprofits classified as a hospital by National Taxonomy of Exempt Entities (NTEE) code. After filtering out associations and specialty hospitals, there are around 3000 nonprofit hospitals left. Importantly, I need to be able to link EINs to other hospital characteristics in the American Hospital Association annual survey and Hospital Compare data. No official crosswalk between the two identification methods exists, so I must rely on matching via location and name of hospital in both data sets. After limiting to only general, nonprofit hospitals in AHA, I confidently match approximately 1200 EINs to an AHA identification number based on exact name matches within the same state.

    For each of the 1200 EINs, I extract urls to 990 pdfs in each year. I download these locally and use OCR text extraction methods to scrape the relevant section of the pdf, which is titled "Officers, Directors, Trustees, Key Employees, and Highest Compensated Employees". Finally, I use string cleaning methods to identify names and positions for each EIN, year. I also identify titles indicating the person is a physician: md, dr., or do. Using OCR extraction can be unreliable in some instances where the text is handwritten or the font is too small. Therefore, I manually find names, positions, and titles for observations that are missing after the initial extraction.

    My identification strategy relies on hospitals which have stable executive leadership characteristics over time. The ideal sample would be hospitals who have the same people on their team for the entire sample. However, this rarely occurs in reality. Therefore, I proxy for a stable leadership in two ways. First, since my main focus is clinical experience, I limit the sample to hospitals who do not change the presence of a doctor on their leadership team from 2010-2014. Second, I increase the restriction by limiting the sample again to those also who do not change their CEO from 2010-2014. Finally, I aggregate the information up to the hospital level, where an observation captures whether the hospital has clinical experience on their  executive team over the course of the sample.

    \subsection{Other Hospital Characteristics and Outcomes}

    Based on the AHAID-EIN matches I found in the first stages of forming the executive leadership teams of hospitals, I merge the aggregated leadership information to both the American Hospital Association (AHA) survey and Hospital Compare data. From AHA, I gather the number of beds in the hospital and the Medicare number, which I use to merge in the Hospital Compare data. Hospital Compare contains multiple datasets which I use to identify hospitals that were penalized in the Hospital Readmissions Reduction Program (HRRP), along with yearly readmission and mortality rates for the relevant HRRP conditions: heart failure, heart attack, and pneumonia, my outcomes of interest.

    The Hospital Compare excess readmissions files contain hospital level information on the excess readmission measure used by CMS to determine whether a hospital should be penalized. This variable is also condition specific, where a hospital is penalized if any of the three excess readmission ratios are greater than one.

    \section{Model and Empirical Strategy}

    \section{Results}

    \section{Conclusion}

	\newpage
	\appendix

    \section{Data}\label{appendixdata}

    \subsection{Gathering Hospital Leadership Names}

    There is no perfect way to access tax form 990s in bulk over the time period I am considering in this paper. However, using the NonProfit Explorer API seems to be the most straightforward. At the time of writing this, information on using version 2 of the API can be found at \hyperlink{https://projects.propublica.org/nonprofits/api}{https://projects.propublica.org/nonprofits/api}.

    There are over 1.5 million nonprofit entities in the US (cite), making it crucially important to be able to filter by type of entity before analyzing any PDFs. The API allows this by filtering a query based on National Taxonomy of Exempt Entities (NTEE) code. I query only nonprofits categorized as E20 (hospitals), E21 (community health systems), and E22 (general hospitals). The API has a pagination limit of 100, meaning I can only pull information on 100 hospitals at a time. Therefore, I filter the query further to only consider one state at a time. The only state that has more than 100 entities registered is California, and thus I subset the California query even further by names that include the word "hospital" and names that don't. I combine all of these subsets and have information on each nonprofits Employee Identification Number. There are 5,588 EINs total in this list. This acts as a list of entities for which I can pull more information.

    I loop through the list of EINs found in the previous step and query more detailed information from the API on that specific EIN. I save the name, secondary name, state, and zip code, all of which do not vary by year. I also save each year's URL that links to the Tax Form 990 PDF. For the sake of a comprehensive data set, I keep years 2006-2020 (I later limit to 2009-2016 when focusing on the Hospital Readmissions Reduction Program). Thus, I finish this step with a panel data set of EIN characteristics and PDF locations. Importantly, there are multiple types of Tax Form 990s depending on the size of the nonprofit. In many cases, one nonprofit has at least two different forms filed in a given year. I filter out any EIN-years for which there are no PDFs on file. The data on PDF locations contains 4,012 EINs and 61,363 EIN-year-tax forms.

    It is crucial to the analysis to be able to link these nonprofits with other sources of data to recover penalties from HRRP, bed size, and outcomes of interest (definitely need to talk more about outcomes). Therefore, I take a conservative approach to matching EINs to American Hospital Association (AHA) ID, which links fairly easily to Medicare ID numbers, based on name and location. First, I will discuss limitations and cleaning that I do to the AHA data and tax form data separately before doing any matching.

    I download the AHA data from Wharton Research Data Services. I filter only to hospitals in the contiguous US, Alaska, and Hawaii (excluding places like Puerto Rico), classified as nonprofit or state/community, and those that are general acute care. I also filter out any hospitals who weren't present in the data (or change system ID) in 2009-2015, meaning they either closed or were acquired. Due to the survey nature of this data, a hospital name may look slightly different from one year to the next. For example, "Waldo County General Hospital" is also Waldo County General Hospital Maine Health". Further, zip codes may change by one or two digits, making them unreliable to match based on. To deal with this, I first keep only unique AHA ID, name, zip, state, and system name combinations. Then, I convert the data from long to wide so that each AHA ID occurs only once, but may have multiple names, zip codes, or system names associated with it.

    I consider which nonprofit entities are not likely to be hospitals and drop them. There are numerous foundations or auxiliary nonprofits with the purpose of raising funds for the hospital, but do not actually care for patient. I filter out any nonprofit with "foundation" or "auxiliary" in the name. I also filter out various specialty centers that fell into the general hospital category, such as hospice or cancer centers.

    I then proceed matching based on names in multiple layers. I focus on exact string matches, so I remove all spaces and common characters that could cause mismatches such as \&, ', -, and inc. Next, I take each AHA name and look for exact matches in a nonprofit's first or secondary name only for nonprofits in the same state as the AHA hospital. When an exact match is found, I record the link between AHA ID and EIN. In this first layer of matching, 860 hospitals in the AHA data are linked to an EIN, equivalent to 31\% of AHA nonprofit hospitals in the sample.

    In the next layer of matching, I remove common words such as "healthcare", "regional", "hospital", etc. That way if there are subtle differences in names, removing common words may allow for an exact match. Again, I take each AHA hospital name and look for exact matches in the nonprofits within the same state. This adds an additional 90 hospital matches, accounting for a total of 34.5\% of AHA hospitals.

    In some cases, tax forms are associated with a system of hospitals instead of one individual hospital. Thus, I create another variable that captures the names of systems which match. The process of matching is the same, the only difference is that now I am considering system names from the AHA data instead of hospital names. A total of 1,136 AHA hospitals have an EIN match on hospital name, system name, or both. This is roughly 41\% of the sample. These hospitals become the main sample for my analysis. While not capturing a large majority of AHA hospitals that are categorized as nonprofit, I would rather be conservative in selecting hospitals which I am confident that the characteristics and measurement of the independent and dependent variables are accurate.

    (add a paragraph about validity of matches)

    Now that I have a sample of hospitals, the next step is to extract the names of the people in charge of these hospitals from the Tax Form 990 PDFs. In the data set of hospital PDF URLs that I collected earlier, I limit to the hospitals with solid matches described above. I then loop through each EIN, downloading each PDF locally and using tesseract package in R to extract text from the relevant pages of the PDF using OCR text extraction methods. In particular, I loop through each page of the PDF, look for the title associated with leadership names: ``Officers, Directors, Trustees, Key Employees, and Highest Compensated Employees", and save all the text from any pages where this title is found. I save the text to a list of all EIN, years present.

    One tricky aspect of the NonProfit Explorer API is that, only in some cases, if two forms are present for an EIN, year, only the first one (which is typically not the one with the relevant information) is pulled. Therefore, for some hospitals, a couple years will have gaps in text extraction data. I locate EIN, years where this problem is occurring, and a team of RAs locates and downloads the correct forms manually. I extract text from these manually downloaded forms in the same manner as above.

    The form of the text data is a data frame with one column, where each line of text is saved in a different row. I write a text cleaning package that locates names, positions, titles, and indications of resigning. I will now describe this function in three parts: preliminary cleaning of strings, locating names, and locating positions, titles, and key words surrounding resigning.

    Typically on the same page as the names and positions is a list of the highest compensated employees and their compensation. In order to not record extra names, I filter out any rows after the start of this section. I then remove any digits, parentheses and brackets, other punctuation, letters that occur by themselves, two letter ``words" that have no meaning, and excess space between words. I then split up the phrase into individual words, so one phrase with 5 words is broken up into 5 variables.

    Next, I locate first and last names in the data.

\end{document}

### 2024-01-26 Tables/ceoMD_overtime.tex
Updates from Overleaf

\begin{table}[h]
\footnotesize
\caption{Averages for CEO MD and non-MD Hospitals Over Time}
\centering
\begin{tabular}[t]{>{\centering\arraybackslash}p{.9cm}>{\centering\arraybackslash}p{.9cm}>{\centering\arraybackslash}p{.9cm}>{\centering\arraybackslash}p{.9cm}>{\centering\arraybackslash}p{.9cm}>{\centering\arraybackslash}p{.9cm}>{\centering\arraybackslash}p{.9cm}>{\centering\arraybackslash}p{.9cm}>{\centering\arraybackslash}p{.9cm}>{\centering\arraybackslash}p{.9cm}>{\centering\arraybackslash}p{.9cm}>{\centering\arraybackslash}p{.9cm}ccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Beds} & \multicolumn{2}{c}{Penalized} & \multicolumn{2}{c}{HRRP Amount} & \multicolumn{2}{c}{Pneum. Mortality} & \multicolumn{2}{c}{Pneum. Readm.} & \multicolumn{2}{c}{HF Mort.} & \multicolumn{2}{c}{HF Readm.} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-7} \cmidrule(l{3pt}r{3pt}){8-9} \cmidrule(l{3pt}r{3pt}){10-11} \cmidrule(l{3pt}r{3pt}){12-13} \cmidrule(l{3pt}r{3pt}){14-15}
Year & No MD & MD & No MD & MD & No MD & MD & No MD & MD & No MD & MD & No MD & MD & No MD & MD\\
\midrule
2008 & 72 & 16 &  &  &  &  & 10.8 & 12.5 &  &  & 10.9 & 10.5 &  & \\
2009 & 110 & 110 &  &  &  &  & 11.7 & 11.8 & 17.8 & 16.5 & 10.9 & 12.0 & 24.0 & 22.2\\
2010 & 109 & 32 &  &  &  &  & 11.9 & 11.3 & 18.3 & 17.8 & 11.5 & 10.3 & 24.7 & 25.0\\
2011 & 94 & 34 &  &  &  &  & 12.1 & 13.0 & 18.3 & 18.1 & 11.7 & 11.9 & 24.8 & 25.3\\
2012 & 97 & 247 & 0.29 & 0.50 & 51944 & 19601 & 12.2 & 12.2 & 18.3 & 18.3 & 12.0 & 12.8 & 24.5 & 24.6\\
\addlinespace
2013 & 68 & 116 & 0.32 & 0.50 & 65329 & 72226 & 12.1 & 11.9 & 17.4 & 18.1 & 12.1 & 12.2 & 22.9 & 23.5\\
2014 & 68 & 197 & 0.34 & 0.67 & 102671 & 134531 & 12.2 & 12.4 & 17.2 & 18.6 & 12.1 & 11.9 & 22.5 & 22.9\\
2015 & 82 & 100 & 0.40 & 0.50 & 123423 & 204336 & 11.7 & 10.6 & 16.7 & 17.5 & 12.0 & 10.6 & 21.8 & 22.6\\
\bottomrule
\end{tabular}
\end{table}

### 2024-01-26 Slides/data_description.tex

\documentclass[notes,11pt, aspectratio=169]{beamer}

\usepackage{pgfpages}
% These slides also contain speaker notes. You can print just the slides,
% just the notes, or both, depending on the setting below. Comment out the want
% you want.
\setbeameroption{hide notes} % Only slide
%\setbeameroption{show only notes} % Only notes
%\setbeameroption{show notes on second screen=right} % Both

\usepackage{helvet}
\usepackage[default]{lato}
\usepackage{array}

\usepackage{tikz}
\usepackage{verbatim}
\setbeamertemplate{note page}{\pagecolor{yellow!5}\insertnote}
\usetikzlibrary{positioning}
\usetikzlibrary{snakes}
\usetikzlibrary{calc}
\usetikzlibrary{arrows}
\usetikzlibrary{decorations.markings}
\usetikzlibrary{shapes.misc}
\usetikzlibrary{matrix,shapes,arrows,fit,tikzmark}
\usepackage{amsmath}
\usepackage{mathpazo}
\usepackage{hyperref}
\usepackage{lipsum}
\usepackage{multimedia}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{graphicx}
\usepackage{dcolumn}
\usepackage{bbm}
\newcolumntype{d}[0]{D{.}{.}{5}}
\usepackage{subfigure}

\usepackage{changepage}
\usepackage{appendixnumberbeamer}
\newcommand{\beginbackup}{
   \newcounter{framenumbervorappendix}
   \setcounter{framenumbervorappendix}{\value{framenumber}}
   \setbeamertemplate{footline}
   {
     \leavevmode%
     \hline
     box{%
       \begin{beamercolorbox}[wd=\paperwidth,ht=2.25ex,dp=1ex,right]{footlinecolor}%
%         \insertframenumber  \hspace*{2ex} 
       \end{beamercolorbox}}%
     \vskip0pt%
   }
 }
\newcommand{\backupend}{
   \addtocounter{framenumbervorappendix}{-\value{framenumber}}
   \addtocounter{framenumber}{\value{framenumbervorappendix}} 
}

\setbeamertemplate{blocks}[rounded]
\setbeamercolor{block title}{bg=purple, fg=white}
\setbeamercolor{block body}{bg=purple!10}

\usepackage{graphicx}
\usepackage[space]{grffile}
\usepackage{booktabs}

% These are my colors -- there are many like them, but these ones are mine.
\definecolor{sage}{RGB}{102,153,102}
\definecolor{yellow}{RGB}{255,173,1}
\definecolor{purple}{RGB}{153,102,153}

% colors for diagrams
\definecolor{diagramtan}{RGB}{225,190,106}
\definecolor{diagramteal}{RGB}{64, 176, 166}
\definecolor{diagrampurple}{RGB}{170, 131, 239}
\definecolor{diagramred}{RGB}{146, 79, 79}

\hypersetup{
  colorlinks=false,
  linkbordercolor = {white},
  linkcolor = {sage}
}

%% I use a beige off white for my background
\definecolor{MyBackground}{RGB}{255,253,218}

%% Uncomment this if you want to change the background color to something else
%\setbeamercolor{background canvas}{bg=MyBackground}

%% Change the bg color to adjust your transition slide background color!
\newenvironment{transitionframe}{
  \setbeamercolor{background canvas}{bg=white}
  \begin{frame}}{
    \end{frame}
}

\setbeamercolor{frametitle}{fg=sage}
\setbeamercolor{title}{fg=sage}
\setbeamertemplate{footline}[frame number]
\setbeamertemplate{navigation symbols}{} 
\setbeamertemplate{itemize items}{$\rightarrow$}
\setbeamercolor{itemize item}{fg=sage}
\setbeamercolor{itemize subitem}{fg=sage}
\setbeamercolor{enumerate item}{fg=sage}
\setbeamercolor{enumerate subitem}{fg=sage}
\setbeamercolor{button}{bg=MyBackground,fg=purple,}

% If you like road maps, rather than having clutter at the top, have a roadmap show up at the end of each section 
% (and after your introduction)
% Uncomment this is if you want the roadmap!
\AtBeginSection[]
{
   \begin{frame}
       \frametitle{Roadmap of Talk}
       \tableofcontents[currentsection]
   \end{frame}
}
\setbeamercolor{section in toc}{fg=sage}
\setbeamercolor{subsection in toc}{fg=sage}
\setbeamersize{text margin left=1em,text margin right=1em}

\newenvironment{wideitemize}{\itemize\addtolength{\itemsep}{10pt}}{\enditemize}

\usepackage{environ}
\NewEnviron{videoframe}[1]{
  \begin{frame}
    \vspace{-8pt}
    \begin{columns}[onlytextwidth, T] % align columns
      \begin{column}{.58\textwidth}
        \begin{minipage}[t][\textheight][t]
          {\dimexpr\textwidth}
          \vspace{8pt}
          \hspace{4pt} {\Large \sc \textcolor{blue}{#1}}
          \vspace{8pt}

          \BODY
        \end{minipage}
      \end{column}%
      \hfill%
      \begin{column}{.42\textwidth}
        \colorbox{green!20}{\begin{minipage}[t][1.2\textheight][t]
            {\dimexpr\textwidth}
            Face goes here
          \end{minipage}}
      \end{column}%
    \end{columns}
  \end{frame}
}

\title[]{\textcolor{sage}{Description of Hospital Leadership Data}}

\author[]{Hanna Glenn}
\date{\today}

\begin{document}

%%% TIKZ STUFF
\tikzset{   
        every picture/.style={remember picture,baseline},
        every node/.style={anchor=base,align=center,outer sep=1.5pt},
        every path/.style={thick},
        }
\newcommand\marktopleft[1]{%
    \tikz[overlay,remember picture] 
        \node (marker-#1-a) at (-.3em,.3em) {};%
}
\newcommand\markbottomright[2]{%
    \tikz[overlay,remember picture] 
        \node (marker-#1-b) at (0em,0em) {};%
}
\tikzstyle{every picture}+=[remember picture] 
\tikzstyle{mybox} =[draw=black, very thick, rectangle, inner sep=10pt, inner ysep=20pt]
\tikzstyle{fancytitle} =[draw=black,fill=red, text=white]
%%%% END TIKZ STUFF

% Title Slide
\begin{frame}
\maketitle
\end{frame}

\begin{frame}{end goal}
    \textcolor{purple}{the purpose of collecting this data is to get a sense of who is in charge of running nonprofit hospitals in the US}

    \vspace{10mm}

    why?
    \begin{wideitemize}
        \item we don't know a lot about who runs hospitals
        \item we know even less about whether characteristics of these people make a difference 
        \item for this project, I care specifically about the characteristic of clinical experience in executive positions
    \end{wideitemize}
\end{frame}

\section{Tax Form 990}

\subsection{who files?}

\begin{frame}{who files?}
\underline{important acronym:} EIN (Employer Identification Number)

    \centering
    \includegraphics[scale=.7]{Graphics/990_snip_frontpage.PNG}
\end{frame}

\begin{frame}{who files?}
    \centering
    \includegraphics[scale=.7]{Graphics/990_snip_officerspage.PNG}
\end{frame}

\subsection{accessing hospital forms}

\begin{frame}{accessing hospital forms}\label{accessing hospital forms}
    \begin{wideitemize}
        \item completed forms are publicly available on ProPublica's website
        \item through their API, I download a list EINs classified as a hospital
        \begin{itemize}
            \item based on National Taxonomy of Exempt Entities (NTEE) codes: E20, E21, E22
        \end{itemize}
        \item along with EIN, I gather their name, zip, state, and urls to tax form pdfs
        \item years 2009-2015
    \end{wideitemize}

    \vspace{10mm}

    \begin{block}{data point}
        3,925 EINs \hspace{10pt} \hyperlink{why more?}{\beamergotobutton{why more?}}
    \end{block}
\end{frame}

\subsection{matching EIN to AHA hospitals}

\begin{frame}{matching EIN to AHA hospitals}
    I need to be able to link hospitals to other public data sets that capture hospital information
    \vspace{4mm}
    \begin{wideitemize}
        \item I choose to match to the widely used American Hospital Association survey data
        \item only match-able link I have is hospital name
        \item I restrict matches to being in the same state
    \end{wideitemize}
\end{frame}

\begin{frame}{matching EIN to AHA hospitals}
\centering
   \begin{tikzpicture}[node distance=6mm, every node/.style={align=center, font=\large}]

        % Left box
        \node[draw=diagrampurple, line width=1.5pt, minimum width=4cm, minimum height=4cm, text=black] (ahaBox) at (0,0) {%
            \underline{\textbf{\textcolor{black}{AHA Hospitals}}} \\[4pt]
            \begin{tabular}{@{}c@{}}
                {Hosp A} \\
                {Hospital B} \\
                {C Hospital} \\
                {Hosp. of D} \\
                {E} \\
            \end{tabular}
        };

        % Right box
        \node[draw=diagramred, line width=1.5pt, minimum width=4cm, minimum height=4cm, text=black, right=of ahaBox] (einBox) {%
            \underline{\textbf{\textcolor{black}{EIN Hospitals}}} \\[4pt]
            \begin{tabular}{@{}c@{}}
                {Hospital B} \\
                {Hospital D} \\
                {H} \\
                {C Hospital} \\
                {Hospital G} \\
            \end{tabular}
        };

    \end{tikzpicture}
\end{frame}

\begin{frame}{matching EIN to AHA hospitals}
\centering
   \begin{tikzpicture}[node distance=6mm, every node/.style={align=center, font=\large}]

        % Left box
        \node[draw=diagrampurple, line width=1.5pt, minimum width=4cm, minimum height=4cm, text=black] (ahaBox) at (0,0) {%
            \underline{\textbf{\textcolor{black}{AHA Hospitals}}} \\[4pt]
            \begin{tabular}{@{}c@{}}
                \textcolor{diagrampurple}{Hosp A} \\
                {Hospital B} \\
                {C Hospital} \\
                {Hosp. of D} \\
                \textcolor{diagrampurple}{E} \\
            \end{tabular}
        };

        % Right box
        \node[draw=diagramred, line width=1.5pt, minimum width=4cm, minimum height=4cm, text=black, right=of ahaBox] (einBox) {%
            \underline{\textbf{\textcolor{black}{EIN Hospitals}}} \\[4pt]
            \begin{tabular}{@{}c@{}}
                {Hospital B} \\
                {Hospital D} \\
                \textcolor{diagramred}{H} \\
                {C Hospital} \\
                \textcolor{diagramred}{Hospital G} \\
            \end{tabular}
        };

    \end{tikzpicture}
\end{frame}

\begin{frame}{matching EIN to AHA hospitals}
\centering
   \begin{tikzpicture}[node distance=6mm, every node/.style={align=center, font=\large}]

        % Left box
        \node[draw=diagrampurple, line width=1.5pt, minimum width=4cm, minimum height=4cm, text=black] (ahaBox) at (0,0) {%
            \underline{\textbf{\textcolor{black}{AHA Hospitals}}} \\[4pt]
            \begin{tabular}{@{}c@{}}
                \textcolor{diagrampurple}{Hosp A} \\
                \textcolor{diagramtan}{Hospital B} \\
                \textcolor{diagramtan}{C Hospital} \\
                {Hosp. of D} \\
                \textcolor{diagrampurple}{E} \\
            \end{tabular}
        };

        % Right box
        \node[draw=diagramred, line width=1.5pt, minimum width=4cm, minimum height=4cm, text=black, right=of ahaBox] (einBox) {%
            \underline{\textbf{\textcolor{black}{EIN Hospitals}}} \\[4pt]
            \begin{tabular}{@{}c@{}}
                \textcolor{diagramtan}{Hospital B} \\
                {Hospital D} \\
                \textcolor{diagramred}{H} \\
                \textcolor{diagramtan}{C Hospital} \\
                \textcolor{diagramred}{Hospital G} \\
            \end{tabular}
        };

    \end{tikzpicture}

\vspace{7mm}

    \begin{block}{data point}
        I match 1,041 EINs to AHA hospitals using perfect name matching
    \end{block}
\end{frame}

\begin{frame}{matching EIN to AHA hospitals}
\centering
   \begin{tikzpicture}[node distance=6mm, every node/.style={align=center, font=\large}]

        % Left box
        \node[draw=diagrampurple, line width=1.5pt, minimum width=4cm, minimum height=4cm, text=black] (ahaBox) at (0,0) {%
            \underline{\textbf{\textcolor{black}{AHA Hospitals}}} \\[4pt]
            \begin{tabular}{@{}c@{}}
                \textcolor{diagrampurple}{Hosp A} \\
                \textcolor{diagramtan}{Hospital B} \\
                \textcolor{diagramtan}{C Hospital} \\
                \textcolor{diagramteal}{Hosp. of D} \\
                \textcolor{diagrampurple}{E} \\
            \end{tabular}
        };

        % Right box
        \node[draw=diagramred, line width=1.5pt, minimum width=4cm, minimum height=4cm, text=black, right=of ahaBox] (einBox) {%
            \underline{\textbf{\textcolor{black}{EIN Hospitals}}} \\[4pt]
            \begin{tabular}{@{}c@{}}
                \textcolor{diagramtan}{Hospital B} \\
                \textcolor{diagramteal}{Hospital D} \\
                \textcolor{diagramred}{H} \\
                \textcolor{diagramtan}{C Hospital} \\
                \textcolor{diagramred}{Hospital G} \\
            \end{tabular}
        };

    \end{tikzpicture}

    \vspace{7mm}

    \begin{block}{data point}
        manually looking at names, I match 196 more EINs to AHA hospitals
    \end{block}
\end{frame}

\subsection{automated vs. manual collection}

\begin{frame}{automated vs. manual collection}
    I now have a list of EINs I want the pdfs for

    \vspace{4mm}

    \begin{wideitemize}
        \item I automate the download of the pdf URLs for these EINs that I collected from the API
        \item in many cases, the API misses the form even though it exists online
        \item I manually download the forms for which this error exists
    \end{wideitemize}
\end{frame}

\section{Text Extraction and Cleaning}

\subsection{optical character recognition}

\begin{frame}{optical character recognition}
    older pdfs are basically images, which requires using OCR for text recognition

    \vspace{4mm}

    \begin{wideitemize}
        \item the software takes pages and returns a string of text found
    \end{wideitemize}
\end{frame}

\begin{frame}{optical character recognition}
\begin{tikzpicture}
\node (bad) {\includegraphics[scale=.5]{Graphics/990_snip_namesexample.PNG}};
\node (good) [right=of bad] {\includegraphics[scale=.45]{Graphics/r_snip_namesexample.PNG}};

\node (a) [fill=purple!50,draw] at (bad.center) {original pdf};
\node (b) [fill=purple!50,draw] at (good.center) {output from OCR};

\draw [ultra thick,purple,->] (a) to[bend left] (b);
\end{tikzpicture}
\end{frame}

\subsection{cleaning raw text output}

\begin{frame}{cleaning raw text output}
algorithmically locate names and associated positions

\vspace{4mm}

\begin{enumerate}
    \item remove numbers and special characters
    \item record full names where at least one (first or last) is found in census most common names data
    \item note when the name is followed by ``MD", ``DO", ``RN", etc. 
    \item locate the position nearest the name found and record it
\end{enumerate}
\end{frame}

\begin{frame}{cleaning raw text output}
    final names data:
    \vspace{4mm}
    \begin{wideitemize}
        \item each row is a name, hospital, year observation
        \item other variables: position, clinical experience
    \end{wideitemize}
\end{frame}

\section{Clinical Experience Characteristic}

\subsection{link names to NPPES}

\begin{frame}{link names to NPPES}
    next goal: verify and get more information on doctors in the data
    \vspace{4mm}
    \begin{wideitemize}
        \item use names to link to NPPES to get NPI number and specialty
        \item some cases are easy: MD in tax data and one name match to NPPES
        \item most are not that easy and require manual sleuthing
    \end{wideitemize}
\end{frame}

\begin{frame}{why more EINs than nonprofit hospitals?}\label{why more?}
    \begin{wideitemize}
        \item various sources quote the number of nonprofit hospitals to be less than 3,000
        \begin{wideitemize}
            \item (add citations)
        \end{wideitemize}
        \item without any limitations to hospital EINs at this point, my 3,925 EINs include hospital associations, fundraisers, and some large clinics that will later be excluded.
    \end{wideitemize}

    \vspace{10mm}

    \hyperlink{accessing hospital forms}{\beamerreturnbutton{go back}}
\end{frame}

\end{document}

### 2024-01-26 Tables/Yearly_counts_name_data.tex

\begin{table}[h]
\footnotesize
\caption{Number of Hospitals with Relevant Leadership Positions}
\centering
\begin{tabular}[t]{>{\centering\arraybackslash}p{.75cm}>{\centering\arraybackslash}p{.75cm}>{\centering\arraybackslash}p{.75cm}>{\centering\arraybackslash}p{.75cm}>{\centering\arraybackslash}p{.75cm}>{\centering\arraybackslash}p{.75cm}>{\centering\arraybackslash}p{.75cm}>{\centering\arraybackslash}p{.75cm}>{\centering\arraybackslash}p{.75cm}>{\centering\arraybackslash}p{.75cm}>{\centering\arraybackslash}p{.75cm}>{\centering\arraybackslash}p{.75cm}ccc}
\toprule
Year & Total & Any MD & CEO & MD CEO & Pres & MD Pres & CEO or Pres & MD CEO or Pres & CFO & MD CFO & CMO & MD CMO & COO & MD COO\\
\midrule
2008 & 109 & 8 & 70 & 2 & 55 & 4 & 87 & 6 & 64 & 1 & 0 & 0 & 14 & 1\\
2009 & 167 & 25 & 105 & 5 & 96 & 11 & 137 & 13 & 89 & 1 & 13 & 8 & 46 & 6\\
2010 & 251 & 46 & 111 & 5 & 103 & 22 & 159 & 24 & 102 & 3 & 22 & 10 & 44 & 4\\
2011 & 225 & 40 & 130 & 8 & 115 & 17 & 178 & 24 & 115 & 1 & 19 & 12 & 48 & 6\\
2012 & 295 & 61 & 234 & 6 & 201 & 37 & 281 & 39 & 216 & 1 & 43 & 25 & 71 & 6\\
\addlinespace
2013 & 233 & 42 & 184 & 8 & 169 & 21 & 226 & 26 & 169 & 0 & 29 & 17 & 50 & 3\\
2014 & 181 & 34 & 137 & 3 & 123 & 12 & 171 & 14 & 132 & 0 & 23 & 16 & 36 & 2\\
2015 & 291 & 64 & 219 & 9 & 202 & 27 & 277 & 33 & 210 & 0 & 44 & 28 & 61 & 4\\
\bottomrule
\end{tabular}
\end{table}

### 2024-01-26 Tables/cmo_overtime.tex

--- 2024-01-26 Tables/ceoMD_overtime.tex
@@
 \footnotesize
-\caption{Averages for CEO MD and non-MD Hospitals Over Time}
+\caption{Averages for Hospitals with and without CMOs}
 \centering
@@
 \cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-7} \cmidrule(l{3pt}r{3pt}){8-9} \cmidrule(l{3pt}r{3pt}){10-11} \cmidrule(l{3pt}r{3pt}){12-13} \cmidrule(l{3pt}r{3pt}){14-15}
-Year & No MD & MD & No MD & MD & No MD & MD & No MD & MD & No MD & MD & No MD & MD & No MD & MD\\
+Year & No CMO & CMO & No CMO & CMO & No CMO & CMO & No CMO & CMO & No CMO & CMO & No CMO & CMO & No CMO & CMO\\
 \midrule
-2008 & 72 & 16 &  &  &  &  & 10.8 & 12.5 &  &  & 10.9 & 10.5 &  & \\
-2009 & 110 & 110 &  &  &  &  & 11.7 & 11.8 & 17.8 & 16.5 & 10.9 & 12.0 & 24.0 & 22.2\\
-2010 & 109 & 32 &  &  &  &  & 11.9 & 11.3 & 18.3 & 17.8 & 11.5 & 10.3 & 24.7 & 25.0\\
-2011 & 94 & 34 &  &  &  &  & 12.1 & 13.0 & 18.3 & 18.1 & 11.7 & 11.9 & 24.8 & 25.3\\
-2012 & 97 & 247 & 0.29 & 0.50 & 51944 & 19601 & 12.2 & 12.2 & 18.3 & 18.3 & 12.0 & 12.8 & 24.5 & 24.6\\
+2008 & 71 &  &  &  &  &  & 10.9 &  &  &  & 10.9 &  &  & \\
+2009 & 105 & 170 &  &  &  &  & 11.7 & 11.0 & 17.8 & 17.8 & 11.0 & 10.3 & 24.0 & 23.3\\
+2010 & 102 & 179 &  &  &  &  & 11.9 & 11.5 & 18.3 & 18.3 & 11.5 & 11.9 & 24.7 & 24.8\\
+2011 & 87 & 146 &  &  &  &  & 12.1 & 11.6 & 18.3 & 18.1 & 11.7 & 11.8 & 24.8 & 24.9\\
+2012 & 89 & 163 & 0.26 & 0.52 & 56825 & 34802 & 12.2 & 12.1 & 18.2 & 18.5 & 11.9 & 12.2 & 24.5 & 24.6\\
 \addlinespace
-2013 & 68 & 116 & 0.32 & 0.50 & 65329 & 72226 & 12.1 & 11.9 & 17.4 & 18.1 & 12.1 & 12.2 & 22.9 & 23.5\\
-2014 & 68 & 197 & 0.34 & 0.67 & 102671 & 134531 & 12.2 & 12.4 & 17.2 & 18.6 & 12.1 & 11.9 & 22.5 & 22.9\\
-2015 & 82 & 100 & 0.40 & 0.50 & 123423 & 204336 & 11.7 & 10.6 & 16.7 & 17.5 & 12.0 & 10.6 & 21.8 & 22.6\\
+2013 & 68 & 80 & 0.31 & 0.39 & 64428 & 72284 & 12.0 & 12.2 & 17.4 & 17.5 & 12.1 & 12.4 & 22.9 & 23.1\\
+2014 & 74 & 46 & 0.35 & 0.30 & 106902 & 77928 & 12.0 & 12.9 & 17.2 & 17.2 & 12.0 & 12.4 & 22.5 & 22.4\\
+2015 & 83 & 81 & 0.40 & 0.39 & 126182 & 126386 & 11.7 & 11.5 & 16.7 & 16.8 & 12.0 & 11.9 & 21.9 & 21.8\\
 \bottomrule

### 2024-01-26 RScripts/data10_create_leadership_variables.R
update scripts

@@ -136,7 +136,7 @@ executive_data_filled <- executive_data_filled %>%
   ungroup() %>%
   mutate(present_2010_2014 = ifelse(in_2010==1 & in_2011==1 & in_2012==1 & in_2013==1 & in_2014==1, 1, 0)) %>%
   mutate(present_2010_2014 = ifelse(is.na(present_2010_2014), 0 ,present_2010_2014)) 
-  # 79% of observations are in the data
+  # 91% of observations are in the data
 
 # keep a record of those that aren't to see if we can manually fill in some missing info caused by OCR
 executive_data_notin20102014 <- executive_data_filled %>%
@@ -243,6 +243,16 @@ years_of_change <- years_of_change %>%
 executive_data_filled <- executive_data_filled %>%
   left_join(years_of_change, by="ein")
 
+# make a separate data set of ein, name, title, and years in data to be used for physician name matching
+# only keep the ones with no MD changes from 2010-2014
+names_data <- executive_data_filled %>%
+  filter(no_md_changes_2010_2014==1) %>%
+  group_by(ein, name, title) %>%
+  summarize(years = paste(sort(unique(year)), collapse=", ")) %>%
+  ungroup()
+
+saveRDS(names_data, paste0(created_data_path, "names_data.rds"))
+
 # create ein-level summary stats about number of doctors and number of total of executives
 executive_data_filled <- executive_data_filled %>%
   group_by(ein, year) %>%
@@ -260,7 +270,7 @@ saveRDS(ein_leadership_changes_data, paste0(created_data_path, "ein_leadership_c
 
 num <- ein_leadership_changes_data %>%
   distinct(ein)
-  # up to 767 eins after the first batch of manually fixing OCR mess ups 
+  # up to 911 eins after manually fixing some OCR mess ups
 
 
 

### 2024-01-26 RScripts/data11_combine_hosp_level_data.R

@@ -233,7 +233,7 @@ saveRDS(penalized_hospital_data, paste0(created_data_path, "penalized_hospital_d
 observe <- penalized_hospital_data %>%
   filter(no_md_changes_2010_2014==1) %>%
   distinct(ein_hosp)
-
+  # got up to 299 hospitals! 
 
 # Summary Statistics ######
 

### 2024-01-26 RScripts/data3_match_AHA_hospitals.R

@@ -558,3 +558,15 @@ saveRDS(manual_matches, paste0(created_data_path, "manual_matched_eins.rds"))
 
 
 
+# put together one list of all the EINs that I need
+ein_hosp <- manual_matched_eins$ein_hosp
+ein_sys <- manual_matched_eins$ein_sys
+
+manual_eins <- c(ein_hosp, ein_sys)
+
+manual_eins <- as.data.frame(manual_eins) %>%
+  rename(ein = manual_eins)
+
+eins <- rbind(einlist_AHAmatching_1, manual_eins) %>%
+  distinct()
+write.csv(eins, "all_ein_list.csv")

### 2024-01-26 RScripts/data8_clean_rawtext_data.R

@@ -1218,7 +1218,9 @@ cleaned_text <- cleaned_text %>%
   add_row(year = 2010, ein = "131624070", name = "michael breslin", position1 = "cfo") %>%
   add_row(year = 2010, ein = "131624070", name = "marc napp", position1 = "vp", extra = "patient quality") %>%
   add_row(year = 2010, ein = "131624096", name = "kenneth davis", position1 = "ceo", title = "md") %>%
-  add_row(year = 2010, ein = "131624096", name = "wayne keathley", position1 = "president", position2 = "coo") %>%
+  add_row(year = 2010, ein = "131624096", name = "wayne keathley", position1 = "president", position2 = "coo") 
+
+  cleaned_text <- cleaned_text %>%
   add_row(year = 2010, ein = "131624096", name = "donald scnlon", position1 = "vp", extra = "exec") %>%
   add_row(year = 2010, ein = "131624096", name = "jeffrey silberstein", position1 = "vp", extra = "exec") %>%
   add_row(year = 2010, ein = "131624096", name = "jane maksoud", position1 = "vp", extra = "hr") %>%
@@ -2168,7 +2170,9 @@ cleaned_text <- cleaned_text %>%
   add_row(year = 2011, ein = "920162721", name = "don kashevaroff", position1 = "ceo") %>%
   add_row(year = 2011, ein = "920162721", name = "garvin federenko", position1 = "cfo") 
   
-  
+cleaned_text <- cleaned_text %>%
+  distinct()
+
 
 saveRDS(cleaned_text, paste0(created_data_path, "cleaned_text.rds"))
 

### 2024-01-26 RScripts/physician_name_matching.R

@@ -6,85 +6,12 @@ library(tidyr)
 source("paths.R")
 
 # read in names data
-cleaned_text <- read_rds(paste0(created_data_path, "names_data.rds"))
-
-# create common names for the positions that can be grouped
-cleaned_text <- cleaned_text %>%
-  mutate(position1 = ifelse(position1 %in% c("president", "pres"), "president", position1)) %>%
-  mutate(position1 = ifelse(position1 %in% c("ceo", "ceocfo", "chief executive officer"), "ceo", position1)) %>%
-  mutate(position1 = ifelse(position1 %in% c("system executive", "medical director", "executive", "treasurer", "secretary", "vice chairman", "chairman", "vice chair", "chairperson", "chair", "vice-chair", "director", "trustee", "board"), "board", position1)) %>%
-  mutate(position1 = ifelse(position1 %in% c("vice president", "vp"), "vice president", position1)) %>%
-  mutate(position1 = ifelse(position1 %in% c("dermatologist", "anesthesiologist", "surgeon"), "physician", position1)) %>%
-  mutate(position1 = ifelse(position1=="chief financial officer", "cfo", position1)) %>%
-  mutate(position1 = ifelse(position1=="chief medical officer", "cmo", position1)) %>%
-  mutate(position2 = ifelse(position2 %in% c("president", "pres"), "president", position2)) %>%
-  mutate(position2 = ifelse(position2 %in% c("ceo", "ceocfo", "chief executive officer"), "ceo", position2)) %>%
-  mutate(position2 = ifelse(position2 %in% c("system executive", "medical director", "executive", "treasurer", "secretary", "vice chairman", "chairman", "vice chair", "chairperson", "chair", "vice-chair", "director", "trustee", "board"), "board", position2)) %>%
-  mutate(position2 = ifelse(position2 %in% c("vice president", "vp"), "vice president", position2)) %>%
-  mutate(position2 = ifelse(position2 %in% c("dermatologist", "anesthesiologist", "surgeon"), "physician", position2)) %>%
-  mutate(position2 = ifelse(position2=="chief financial officer", "cfo", position2)) %>%
-  mutate(position2 = ifelse(position2=="chief medical officer", "cmo", position2))
-
-# fix year to reflect actual year instead of tax filing year
[…]

### 2024-01-26 RScripts/prelim_regressions.R

@@ -18,7 +18,8 @@ hospital_data <- readRDS(paste0(created_data_path, "penalized_hospital_data(temp
          heartfailure_mortality = rate_heartfailure_mortality,
          heartfailure_readmission = rate_heartfailure_readmission,
          pneum_mortality = rate_pneum_mortality,
-         pneum_readmission = rate_pneum_readmission)
+         pneum_readmission = rate_pneum_readmission) %>%
+  mutate(post_2012 = ifelse(year>=2012, 1, 0))
 
 # create min year penalized variable
 minyr_penalized <- hospital_data %>%
@@ -39,7 +40,7 @@ hospital_data <- hospital_data %>%
 noMDchg_pen_sample <- hospital_data %>%
   filter(no_md_changes_2010_2014==1) %>%
   mutate(has_any_md = ifelse(total_docs>0, 1, 0)) %>%
-  select(ID, year, heartattack_mortality, heartattack_readmission, pneum_mortality, pneum_readmission, heartfailure_mortality, heartfailure_readmission, has_any_md, minyr_pen, yr2009, yr2010, yr2011, yr2012, yr2013, yr2014, yr2015, weightedavg_read, weightedavg_mort) %>%
+  select(ID, year, heartattack_mortality, heartattack_readmission, pneum_mortality, pneum_readmission, heartfailure_mortality, heartfailure_readmission, has_any_md, minyr_pen, yr2009, yr2010, yr2011, yr2012, yr2013, yr2014, yr2015, weightedavg_read, weightedavg_mort, post_2012) %>%
   mutate(post = ifelse(year>=minyr_pen, 1, 0))
 
 # 2. No MD changes, penalized for HF
@@ -47,7 +48,7 @@ noMDchg_hfpen_sample <- hospital_data %>%
   filter(no_md_changes_2010_2014==1) %>%
   filter(ever_pen_hf>0) %>%
   mutate(has_any_md = ifelse(total_docs>0, 1, 0)) %>%
-  select(ID, year, heartattack_mortality, heartattack_readmission, pneum_mortality, pneum_readmission, heartfailure_mortality, heartfailure_readmission, has_any_md, minyr_pen, yr2009, yr2010, yr2011, yr2012, yr2013, yr2014, yr2015, weightedavg_read, weightedavg_mort) %>%
[…]

### 2024-01-26 Paper/paper.tex

@@
 
-    Based on the AHAID-EIN matches I found in the first stages of forming the executive leadership teams of hospitals, I merge the aggregated leadership information to both the American Hospital Association (AHA) survey and Hospital Compare data.
+    Based on the AHA ID-EIN matches I found in the first stages of forming the executive leadership teams of hospitals, I merge the aggregated leadership information to both the American Hospital Association (AHA) survey and Hospital Compare data.
 From AHA, I gather the number of beds in the hospital and the Medicare number, which I use to merge in the Hospital Compare data.
@@
 This variable is also condition specific, where a hospital is penalized if any of the three excess readmission ratios are greater than one.
+Further, Hospital Compare also publishes hospital level readmission and mortality rates for heart failure, heart attack, and pneumonia.
+For my main outcomes of interest, I take an average across conditions and weight it using the total number of patients in that condition.
+For further analysis, I create samples of hospitals who were penalized for specific conditions and investigate the change in readmission and mortality rates for the condition they were penalized for.
 
     \section{Model and Empirical Strategy}
+
+    Hospital that are not traditional for-profits maximize an
 

### 2024-01-29 RScripts/data10_create_leadership_variables.R
add event study graph

@@ -72,11 +72,15 @@ cleaned_text <- cleaned_text %>%
   mutate(position = ifelse(is.na(position) & pos1_pres==1 & pos2_board==1, "board", position)) %>% #4
   mutate(position = ifelse(is.na(position) & pos2_pres==1 & pos1_board==1, "board", position)) %>% #4
   mutate(position = ifelse(is.na(position), "president (board or executive)", position)) #5
+
+cleaned_text <- cleaned_text %>%
+  mutate(ceo = ifelse(position1=="ceo" | position2 == "ceo" | position1 == "chief executive officer" | position2 == "chief executive officer", 1,0)) %>%
+  mutate(ceo = ifelse(is.na(ceo),0,ceo))
   
 executive_data <- cleaned_text %>%
   mutate(vp = ifelse(pos1_vp==1 | pos2_vp==1, 1, 0)) %>%
   filter(position == "executive") %>%
-  select(ein, year, name, position, title, former, vp) %>%
+  select(ein, year, name, position, title, former, vp, ceo) %>%
   filter(!(former %in% c("past", "former", "past interim")))
 
 # loop through eins to fill in any missing years that happen between nonmissing years
@@ -119,7 +123,7 @@ for (i in seq_along(eins)){
 # fill MD
 executive_data_filled <- executive_data_filled %>%
   group_by(ein, id) %>%
-  fill(title, .direction="downup") %>%
+  fill(title, vp, ceo, .direction="downup") %>%
   ungroup()
 
 # goal 2: identify whether changes occur from year to year ####
@@ -139,8 +143,8 @@ executive_data_filled <- executive_data_filled %>%
   # 91% of observations are in the data
 
 # keep a record of those that aren't to see if we can manually fill in some missing info caused by OCR
-executive_data_notin20102014 <- executive_data_filled %>%
-  filter(present_2010_2014==0)
+#executive_data_notin20102014 <- executive_data_filled %>%
+  #filter(present_2010_2014==0)
 
 # saveRDS(executive_data_notin20102014, paste0(created_data_path, "executive_data_notin20102014.rds"))
 # write.csv(executive_data_notin20102014, paste0(created_data_path, "executive_data_notin20102014.csv"))
@@ -243,6 +247,39 @@ years_of_change <- years_of_change %>%
 executive_data_filled <- executive_data_filled %>%
   left_join(years_of_change, by="ein")
 
[…]

### 2024-01-29 RScripts/data11_combine_hosp_level_data.R

@@ -3,6 +3,8 @@ library(dplyr)
 library(tidyr)
 library(stringr)
 library(fabricatr)
+library(knitr)
+library(kableExtra)
 
 # Read in Data ##################
 
@@ -12,8 +14,8 @@ manual_matched_eins <- read_rds(paste0(created_data_path, "manual_matched_eins.r
 # Data from AHA 
 AHA_hosp_names <- read_csv(paste0(created_data_path, "raw data/AHA_hosp_names.csv"))
 # HCRIS data that has HRRP penalty information (and beds)
-# final_HCRIS_v2010 <- read_csv(paste0(created_data_path, "raw data/final_HCRIS_v2010.csv"))
-# final_HCRIS_v1996 <- read_csv(paste0(created_data_path, "raw data/final_HCRIS_v1996.csv"))
+final_HCRIS_v2010 <- read_csv(paste0(created_data_path, "raw data/final_HCRIS_v2010.csv"))
+final_HCRIS_v1996 <- read_csv(paste0(created_data_path, "raw data/final_HCRIS_v1996.csv"))
 # Hospital Compare data that also has penalty information
 hc_readm_2012 <- read_csv(paste0(created_data_path, "raw data/Hospital Compare/vwhqi_readm_reduction_2012.csv"))
 hc_readm_2013 <- read_csv(paste0(created_data_path, "raw data/Hospital Compare/vwhqi_readm_reduction_2013.csv"))
@@ -51,64 +53,64 @@ observe <- hospital_data %>%
 # only keep rows that had values in that data set
 hospital_data <- hospital_data %>%
   filter(!is.na(no_changes_ever))
-  # takes us down to 680 eins
+  # takes us down to 850 eins
 
 # join AHA data to get medicare number
 hospital_data <- hospital_data %>%
-  mutate(ID=as.character(ID)) %>%
+  mutate(ID = as.numeric(ID)) %>%
   left_join(AHA_hosp_names, by=c("ID", "year"="YEAR"))
 hospital_data <- hospital_data %>%
   group_by(ID) %>%
   fill(MCRNUM, .direction="downup") %>%
   ungroup() %>%
   select(ID, ein_hosp, MCRNUM, year, no_changes_ever, no_changes_2010_2014, no_changes_2011_2013, no_small_changes_ever, no_small_changes_2010_2014,
-         no_small_changes_2011_2013, no_md_changes_ever, no_md_changes_2010_2014, no_md_changes_2011_2013, total_execs, total_docs)
[…]

### 2024-01-29 RScripts/data8_clean_rawtext_data.R

@@ -132,6 +132,8 @@ for (i in 1:length(AHA_ein_list_manualmatched)){
   }
 }
 
+cleaned_text <- readRDS(paste0(created_data_path, "cleaned_text.rds"))
+
 # Further clean text data
 cleaned_text <- cleaned_text %>%
   distinct() %>%

### 2024-01-29 RScripts/prelim_regressions.R

@@ -3,6 +3,7 @@ library(dplyr)
 library(fixest)
 library(ggplot2)
 library(ggpubr)
+library(tidyr)
 
 # Read in hospital_data created in "create_hospital_data.R"
 hospital_data <- readRDS(paste0(created_data_path, "penalized_hospital_data(temp).rds")) %>%
@@ -19,7 +20,18 @@ hospital_data <- readRDS(paste0(created_data_path, "penalized_hospital_data(temp
          heartfailure_readmission = rate_heartfailure_readmission,
          pneum_mortality = rate_pneum_mortality,
          pneum_readmission = rate_pneum_readmission) %>%
-  mutate(post_2012 = ifelse(year>=2012, 1, 0))
+  mutate(post_2012 = ifelse(year>=2012, 1, 0)) %>%
+  mutate(post_2011 = ifelse(year>=2011, 1, 0)) %>%
+  mutate(has_any_md = ifelse(total_docs>0, 1, 0)) %>%
+  mutate(did = has_any_md*post_2012) %>%
+  mutate(did_2011 = has_any_md*post_2011) %>%
+  mutate(es2009 = yr2009*has_any_md,
+         es2010 = yr2010*has_any_md,
+         es2011 = yr2011*has_any_md,
+         es2012 = yr2012*has_any_md,
+         es2013 = yr2013*has_any_md,
+         es2014 = yr2014*has_any_md,
+         es2015 = yr2015*has_any_md)
 
 # create min year penalized variable
 minyr_penalized <- hospital_data %>%
@@ -39,8 +51,8 @@ hospital_data <- hospital_data %>%
 # 1. No MD changes, penalized
 noMDchg_pen_sample <- hospital_data %>%
   filter(no_md_changes_2010_2014==1) %>%
-  mutate(has_any_md = ifelse(total_docs>0, 1, 0)) %>%
-  select(ID, year, heartattack_mortality, heartattack_readmission, pneum_mortality, pneum_readmission, heartfailure_mortality, heartfailure_readmission, has_any_md, minyr_pen, yr2009, yr2010, yr2011, yr2012, yr2013, yr2014, yr2015, weightedavg_read, weightedavg_mort, post_2012) %>%
[…]

### 2024-01-29 Paper/paper.tex

@@
 \usepackage{import}
+\usepackage[backend=biber,style=authoryear,
+sorting=ynt,citestyle=authoryear]{biblatex}
+\addbibresource{papercitations.bib}
 %\usepackage[nomarkers,nofiglist,notablist]{endfloat}
@@
 \newtheorem{lemma}[theorem]{Lemma}
+
+\captionsetup{justification=raggedright,singlelinecheck=false}
 
@@
   The objectives and behaviors of firms that are not classic for-profits are not easily determined, and an extensive literature has sought to understand these types of firms.
-Nonprofits are thought to care about more than just monetary profit, but also gain utility from furthering a mission and other unknown factors.
+Nonprofits are thought to gain utility not solely from monetary profit, but also through accomplishing mission-based purposes.
 Researchers have speculated a variety of objective functions that could motivate nonprofits such as maximizing prestige, income, or a quality/quantity tradeoff (Steinberg, 1986).
 An important contribution to this literature is the use of empirical strategies to analyze how firm behavior reveals attributes of the underlying objective function (add citations).
+In this paper, I leverage a large-scale shock to nonprofit hospitals in the US to investigate whether characteristics of the objective function is revealed by the composition of their leadership team.
 
@@
 Second, nonprofit hospitals make up over 50\% of all hospitals in the US, and staff, on average, 207 beds.
-For comparison, for-profits make up around 30\% of hospitals and staff 100 beds.
+For comparison, for-profits make up around 36\% of hospitals and staff 107 beds (\cite{ASPE_2023}).
 Thus, nonprofit behavior in the hospital industry is relevant and directly affects consumers of health care.
@@
 
-  What has not been considered is how the leadership make-up within nonprofits potentially drive objectives.
+  What has not been considered is how the leadership team within nonprofits potentially drives objectives.
 In large, publicly traded firms, executive characteristics are shown to be correlated with firm performance, indicating that there is more to firm behavior than just for-profit or not-for-profit status.
-Within nonprofit hospitals, there is variation in the propensity to hire executives with a background in clinical settings, which could reasonably change how much weight the hospital places on profit vs.
+Within nonprofit hospitals, there is variation in the propensity to hire executives with a background in clinical work, which could reasonably change how much weight the hospital places on profit vs.
 quality of care in their objective function.
-In this paper, I contribute to our understanding of hospital behaviors by investigating whether clinical experience on an executive team affects hospital behavior after being financially penalized.
+I contribute to our understanding of hospital and nonprofit behaviors by investigating whether clinical experience on an executive team affects hospital behavior after being financially penalized.
 
@@
 In order to capture the effect of clinical experience in executives, not just the selection of clinical executives into certain types of hospitals, I specifically investigate hospitals penalized by the Hospital Readmissions Reduction Program (HRRP) of 2012.
-This program reduced Medicare payments to hospitals with above average readmission rates.
+This program reduced Medicare payments to hospitals with above average readmission rates, and is plausibly uncorrelated with previously determined leadership teams.
 I further limit the sample to hospitals with stable executive teams over time, and characterize hospitals as having clinical experience on their team or not.
-I then estimate the difference in readmission and mortality rates between hospitals with and without clinical experience before and after the penalties began.
+I estimate the difference in readmission and mortality rates between hospitals with and without clinical experience before and after the penalties began, conditional on being penalized.
 I assume that before the penalty, clinical and non-clinical team hospitals had similar readmission and mortality trends, and that the composition of hospital executive teams determined prior to the penalties is, conditional on being penalized, exogenous.
@@
 
-    Avoidable readmissions are considered a bad outcome for patients.
-The Hospital Readmissions Reduction Program (HRRP) was enacted as a part of the Affordable Care Act to incentivize hospitals to reduce readmission rates.
+    A readmission is when a patient returns to the hospital within 30 days of being discharged from a previous stay; avoidable readmissions are a bad outcome for patients and increase health care spending.
+The Hospital Readmissions Reduction Program (HRRP) was enacted in 2010 as a part of the Affordable Care Act to incentivize hospitals to reduce readmission rates.
 The goal of the program was to get hospitals to lower readmissions through better care coordination, less initial stay complications, and better post-care instructions.
-Beginning in October 2012, hospitals with higher readmission rates than the national average (after adjusting for demographic characteristics faced by the hospital) receive a lower reimbursement rate for all Medicare patients they treat.
+Beginning in October 2012, hospitals with higher readmission rates than the national average in pneumonia, heart failure, or AMI (after adjusting for demographic characteristics faced by the hospital) receive a lower reimbursement rate for all Medicare patients they treat.
+The penalties are based on each hospital's past performance in readmissions, which means hospitals had incentive to react immediately once details of the program were announced in October of 2011.
 
     \subsection{Nonprofit Hospital Leadership}
+
+    Nonprofits are governed by a board of directors, whose role is to set high level goals and strategies, and provide general oversight.
+The board is not in charge of day-to-day operations, but the general direction of the firm.
+They select members of the executive team to carry out day-to-day operations.
+The executive team usually consists of at least a CEO and CFO.
+Every nonprofit is different in exactly how they structure an executive team.
+Some nonprofits hire Chief Medical Officers, Chief Quality Officers, executive directors, vice presidents of various departments, etc.
+I consider all of these employees as part of the day-to-day management of the hospital and include them in the executive team.
 
@@
 
-    The main data used for analysis is a novel data set of nonprofit hospitals in the US which contains names and select characteristics of executives tied to the hospital from 2009-2015.
-I gather this data from each hospital's publicly available Tax Form 990, which is required each year for all nonprofits in the US\footnote{something about different types of forms here}.
+    To characterize hospitals as having clinical experience on executive team or not, I construct a novel data set of nonprofit hospitals in the US which contains names and select characteristics of executives tied to the hospital from 2009-2015.
+I gather this data from each hospital's publicly available Tax Form 990, which is required each year for all nonprofits in the US.
 To my knowledge, this is the first large scale gathering of leadership names from these forms.
-Thus, I will explain each step in the data extraction process.
-More details can be found in Appendix blank.
+I include a very detailed account of the collection of this data in Appendix \ref{appendixdata}.
 
     The tax forms for all nonprofits are housed by ProPublica\footnote{https://projects.propublica.org/nonprofits/}.
-I use the NonProfit Explorer API to extract the Employee Identification Numbers (EIN) of all nonprofits classified as a hospital by National Taxonomy of Exempt Entities (NTEE) code.
+I use the NonProfit Explorer API to extract the Employee Identification Numbers (EIN) of all nonprofits classified as a hospital according to their National Taxonomy of Exempt Entities (NTEE) code.
 After filtering out associations and specialty hospitals, there are around 3000 nonprofit hospitals left.
-Importantly, I need to be able to link EINs to other hospital characteristics in the American Hospital Association annual survey and Hospital Compare data.
+Importantly, I need to be able to link EINs to other hospital characteristics, including penalty status and readmission/mortality rates, in the American Hospital Association annual survey and Hospital Compare data.
 No official crosswalk between the two identification methods exists, so I must rely on matching via location and name of hospital in both data sets.
@@
     For each of the 1200 EINs, I extract urls to 990 pdfs in each year.
-I download these locally and use OCR text extraction methods to scrape the relevant section of the pdf, which is titled "Officers, Directors, Trustees, Key Employees, and Highest Compensated Employees".
+I download these locally and use OCR text extraction methods to scrape the relevant section of the pdf, which is titled ``Officers, Directors, Trustees, Key Employees, and Highest Compensated Employees".
 Finally, I use string cleaning methods to identify names and positions for each EIN, year.
@@
 
-    Based on the AHA ID-EIN matches I found in the first stages of forming the executive leadership teams of hospitals, I merge the aggregated leadership information to both the American Hospital Association (AHA) survey and Hospital Compare data.
+    I merge this hospital-level data with the American Hospital Association (AHA) survey using the EIN-AHA links I created based on hospital names.
 From AHA, I gather the number of beds in the hospital and the Medicare number, which I use to merge in the Hospital Compare data.
-Hospital Compare contains multiple datasets which I use to identify hospitals that were penalized in the Hospital Readmissions Reduction Program (HRRP), along with yearly readmission and mortality rates for the relevant HRRP conditions: heart failure, heart attack, and pneumonia, my outcomes of interest.
+Hospital Compare contains multiple data sets which I use to identify hospitals that were penalized in the Hospital Readmissions Reduction Program (HRRP), along with yearly readmission and mortality rates for the relevant HRRP conditions: heart failure, heart attack, and pneumonia, my outcomes of interest.
 
@@
 
+    \import{Tables}{noMDchg_sumstats.tex}
+
+    Figure \ref{weighted_read_mort_graph} shows the average of the main outcomes across all penalized hospitals over time.
+While mortality rates did not seem to change much after the penalties began in 2012, readmission rates decreased after the penalty for hospitals with and without clinical experience.
+However, readmission rates for non-clinical hospitals decreased more dramatically than those for clinical hospitals, even dropping below those for clinical teams in 2013 and 2014.
+While suggestive, this motivates further analysis into the differential responses between clinical and non-clinical research teams.
+
+    \begin{figure}[h!]
+        \centering
+        \caption{Rates over time in all penalized hospitals}
+        \includegraphics[scale=.7]{Objects/weighted_read_mort_graph.pdf}
+        \label{weighted_read_mort_graph}
+    \end{figure}
+
     \section{Model and Empirical Strategy}
 
-    Hospital that are not traditional for-profits maximize an
+    Not knowing the true form of nonprofit hospital objective functions, we can assume that they maximize some weighted average of revenue and output. Specifically,
+
+    $$\max\hspace{2mm} \delta R + (1-\delta) u(q,\theta)$$
+
+    \noindent where $R$ is net revenue, $q$ is quantity, and $\theta$ is quality of care.
+While previous research has set out to understand the average $\delta$ for nonprofit hospitals, $\delta$ may also be dependent on the leadership teams within nonprofit hospitals.
+
+    Thus, I directly estimate the effect of having clinical experience on readmission and mortality rates.
+However, looking purely at this correlation is likely a biased picture of the effect of clinical since it likely captures reverse causality: certain types of hospitals may be more likely to hire doctors as executives.
+Therefore, I leverage the HRRP shock, which allows me to identify hospitals similar in quality along readmission rates.
+For all penalized hospitals, I further categorize their type as having clinical experience on their executive team or not.
+This categorization should be time invariant since I cannot rule out changes in leadership teams being correlated with being penalized.
+For this purpose I drop hospitals who had significant changes in their leadership team.
+My preferred specification defines a stable leadership team as one which does not change their propensity to hire a doctor executive.
+
+    For this restricted sample, I estimate the following regression equation:
+
+    $$y_{ht} = (clinical \times post\_penalty)_{ht} + \alpha_{h} + \delta_t + \epsilon_{ht}$$
+
+    \noindent where $y_{ht}$ is one of the readmission or mortality outcome variables discussed in section DATA, $(clinical \times post\_penalty)_{ht}$, is the variable of interest capturing the effect of having clinical experience after the penalties took place, $\alpha_h$ and $\delta_t$ are hospital and time fixed effects, respectively, and $\epsilon$ is an error term.
+
+    I further investigate the assumption that non-clinical and clinical hospitals behaved similarly prior to the HRRP by estimating event study coefficients:
+
+    put event study equation here
 
     \section{Results}
+
+    \import{Tables}{read_results_table.tex}
+
+    \import{Tables}{mort_results_table.tex}
 

### 2024-01-29 RScripts/prelim_regressions.R
remove mortality from first draft of paper

@@ -109,18 +109,21 @@ noMDchg_top_sample <- hospital_data %>%
 
 # Create Graphs for Paper ################
 # 1. weighted average readmissions and mortality on same graph
-weighted_read_mort_graph_data <- noMDchg_pen_sample %>%
+weighted_read_graph_data <- noMDchg_pen_sample %>%
   group_by(has_any_md, year) %>%
-  summarise_at(c("weightedavg_read", "weightedavg_mort"), list(mean), na.rm=T) %>%
+  summarise_at(c("weightedavg_read", "heartattack_readmission", "heartfailure_readmission", "pneum_readmission"), list(mean), na.rm=T) %>%
   filter(year!=2008) %>%
   mutate(has_any_md = as.factor(has_any_md)) %>%
-  pivot_longer(c("weightedavg_read", "weightedavg_mort"), names_to = "outcome", values_to = "rate") %>%
-  mutate(outcome = ifelse(outcome=="weightedavg_mort", "mortality", "readmission")) %>%
-  rename(`clinical experience`=has_any_md)
-weighted_read_mort_graph <- ggplot(weighted_read_mort_graph_data, aes(x=year, y=rate, color=outcome, linetype=`clinical experience`)) + geom_vline(xintercept = 2012, linetype = "dotted") + 
-  geom_line(size=.75) + theme_bw() + 
-  ylim(11,23) + theme(text=element_text(size=18))
-ggsave(plot = weighted_read_mort_graph, filename = paste0(objects_path, "weighted_read_mort_graph.pdf"), width = 9, height = 6, units = "in")
+  pivot_longer(c("weightedavg_read", "heartattack_readmission", "heartfailure_readmission", "pneum_readmission"), names_to = "outcome", values_to = "rate") %>%
+  mutate(outcome = ifelse(outcome=="weightedavg_read", "weighted average", outcome)) %>%
+  mutate(outcome = ifelse(outcome=="heartattack_readmission", "AMI", outcome)) %>%
+  mutate(outcome = ifelse(outcome=="heartfailure_readmission", "heart failure", outcome)) %>%
+  mutate(outcome = ifelse(outcome=="pneum_readmission", "pneumonia", outcome)) %>%
+  rename(`clinical experience`=has_any_md, `category`=outcome)
[…]

### 2024-01-29 Paper/paper.tex

@@
 
-	\section{Data}
+	\section{Data}\label{sec:data}
 
@@
         \centering
-        \caption{Rates over time in all penalized hospitals}
-        \includegraphics[scale=.7]{Objects/weighted_read_mort_graph.pdf}
+        \caption{Readmission rates over time in all penalized hospitals}
+        \includegraphics[scale=.6]{Objects/weighted_read_mort_graph.pdf}
         \label{weighted_read_mort_graph}
@@
 
-    Not knowing the true form of nonprofit hospital objective functions, we can assume that they maximize some weighted average of revenue and output. Specifically,
+    Not knowing the true form of nonprofit hospital objective functions, we can assume that they maximize some weighted average of revenue and quality.
+At a basic level, we can think of nonprofits behaving according to:
 
-    $$\max\hspace{2mm} \delta R + (1-\delta) u(q,\theta)$$
+    $$\max\hspace{2mm} \delta R + (1-\delta) u(\theta)$$
 
-    \noindent where $R$ is net revenue, $q$ is quantity, and $\theta$ is quality of care.
+    \noindent where $R$ is net revenue, $u(.)$ is an increasing function that captures utility from quality of care, and $\theta$ is some quality of care measure.
+In this setting, we care about hospital responding to a penalty by adjusting readmission rates, so $\theta = 1-r$ where $r$ is the readmission rate.
 While previous research has set out to understand the average $\delta$ for nonprofit hospitals, $\delta$ may also be dependent on the leadership teams within nonprofit hospitals.
+Further, introducing the HRRP means that $R$ is directly dependent on $\theta$.
+We can then write the hospital's problem as
 
-    Thus, I directly estimate the effect of having clinical experience on readmission and mortality rates.
-However, looking purely at this correlation is likely a biased picture of the effect of clinical since it likely captures reverse causality: certain types of hospitals may be more likely to hire doctors as executives.
-Therefore, I leverage the HRRP shock, which allows me to identify hospitals similar in quality along readmission rates.
-For all penalized hospitals, I further categorize their type as having clinical experience on their executive team or not.
-This categorization should be time invariant since I cannot rule out changes in leadership teams being correlated with being penalized.
-For this purpose I drop hospitals who had significant changes in their leadership team.
-My preferred specification defines a stable leadership team as one which does not change their propensity to hire a doctor executive.
+    $$\max\hspace{2mm} \delta(L) R(\theta) + (1-\delta(L)) u(\theta)$$
 
-    For this restricted sample, I estimate the following regression equation:
+    \noindent where $L$ captures the composition of the leadership team.
+Therefore, when maximizing this function, $\theta$ is dependent on $L$ through its impact on how the hospital values profit vs.
+quality of care.
 
-    $$y_{ht} = (clinical \times post\_penalty)_{ht} + \alpha_{h} + \delta_t + \epsilon_{ht}$$
+    Thus, I directly estimate the effect of having clinical experience on readmission rates.
+However, looking purely at this correlation is likely a biased picture of the effect of clinical experience since it likely captures reverse causality: certain types of hospitals may be more likely to hire doctors as executives.
+I do two things to address this issue.
+First, as discussed in detail in Section \ref{sec:data}, I only consider hospitals with stable leadership teams over time.
+Second, I leverage the HRRP shock, limiting the sample to those that were penalized, which allows me to identify hospitals similar in quality along readmission rates, and to observe changes in behavior after the shock.
+For the hospitals in my sample, I define their type as having clinical experience on their executive team or not, a time invariant categorization.
 
-    \noindent where $y_{ht}$ is one of the readmission or mortality outcome variables discussed in section DATA, $(clinical \times post\_penalty)_{ht}$, is the variable of interest capturing the effect of having clinical experience after the penalties took place, $\alpha_h$ and $\delta_t$ are hospital and time fixed effects, respectively, and $\epsilon$ is an error term.
+    I estimate the following regression equation:
 
-    I further investigate the assumption that non-clinical and clinical hospitals behaved similarly prior to the HRRP by estimating event study coefficients:
+    $$y_{ht} = \beta (clinical \times post\_penalty)_{ht} + \alpha_{h} + \delta_t + \epsilon_{ht}$$
 
-    put event study equation here
+    \noindent where $y_{ht}$ is one of the outcome variables discussed in Section \ref{sec:data}, $\beta$ is the variable of interest capturing the effect of having clinical experience after the penalties took place, $\alpha_h$ and $\delta_t$ are hospital and time fixed effects, respectively.
+
+    For identification, I rely on the assumption that, prior to being penalized, clinical and non-clinical team hospitals did not differentially change readmission rates for a reason I do not capture here.
+I investigate this assumption by estimating event study coefficients, where the coefficients $\beta_j$ now capture the effect of having clinical experience in a particular year, and $\beta_{2011}$ is normalized to zero:
+
+    $$y_{ht} = \sum_{j=2009}^{2010}\beta_j(clinical_{ht} \times \mathbf{1}\{y=j\}) + \sum_{j=2012}^{2015}\beta_j (clinical_{ht} \times \mathbf{1}\{y=j\}) + \alpha_{h} + \delta_t + \epsilon_{ht}.$$
 
@@
 
-    \import{Tables}{mort_results_table.tex}
+    \begin{figure}
+        \centering
+        \caption{Event Study Estimates}
+        \includegraphics[scale=.5]{Objects/wa_eventstudy.pdf}
+        \label{fig:wa_eventstudy}
+    \end{figure}
 

### 2024-01-29 RScripts/prelim_regressions.R
fix graphs for first draft of paper

@@ -123,7 +123,7 @@ weighted_read_graph_data <- noMDchg_pen_sample %>%
 weighted_read_graph <- ggplot(weighted_read_graph_data, aes(x=year, y=rate, color=category, linetype=`clinical experience`)) + geom_vline(xintercept = 2012, linetype = "dotted") + 
   geom_line(size=.75) + theme_bw() + xlab("\nyear") +
   ylim(15,30) + theme(text=element_text(size=18)) + scale_color_brewer(palette="BrBG")
-ggsave(plot = weighted_read_mort_graph, filename = paste0(objects_path, "weighted_read_mort_graph.pdf"), width = 11, height = 6, units = "in")
+ggsave(plot = weighted_read_graph, filename = paste0(objects_path, "weighted_read_graph.pdf"), width = 11, height = 6, units = "in")
 
 
 # Create tables for paper #########
@@ -188,7 +188,7 @@ read_es <- ggplot(data, aes(vars, Estimate)) +
   ylab("Estimate and 95% CI") +
   scale_color_manual(labels = c("Post", "Pre"), values=c("#E69F00", "#999999"),name="") +
   scale_fill_manual(labels = c("Post", "Pre"), values=c("#E69F00", "#999999"),name="") + ylim(-1,1)
-ggsave(read_es, filename = paste0(objects_path, "read_es_graph.pdf"), height = 11, width = 6, units = "in")
+ggsave(read_es, filename = paste0(objects_path, "read_es_graph.pdf"), width = 11, height = 6, units = "in")
 
 # weighted average mortality
 noMDchg_pen_wa_mort_es <- fixest::feols(weightedavg_mort ~ es2009 + es2010 + es2012 + es2013 + es2014 + es2015 | ID + year, cluster = ~ID,

### 2024-01-31 Paper/paper.tex
Updates from Overleaf

@@
 
+    \begin{center}
+    \large
+    \textbf{This is a preliminary draft. Please do not cite or circulate.}
+	\end{center}
+
+ \vspace{3mm}
+
     \begin{abstract}
 		{\small
+		The objectives that underlie nonprofit firm behaviors are not well understood, but perhaps that is because there are characteristics even within nonprofits that dictate behavior.
+In this paper, I use a novel data set on nonprofit hospital executives in the US from 2009-2015 to investigate whether having clinical experience on a hospital executive team affects hospital behavior.
+In particular, I leverage the Hospital Readmissions Reduction Program of 2012 as a shock to similar quality hospitals, and estimate the difference in response from hospitals with and without clinical experience.
+I find that while all penalized hospitals decreased readmissions after the shock, non-clinical executive team hospital decreased readmissions at a faster rate than those with a clinical executive team.
+The difference equates to approximately 15 additional readmissions for clinical team hospitals.
+I find no differential effect of clinical executive teams on mortality rates.
 
@@
 
-	\vspace{0.1in}
+	\vspace{0.8in}
 
@@
 Researchers have speculated a variety of objective functions that could motivate nonprofits such as maximizing prestige, income, or a quality/quantity tradeoff (Steinberg, 1986).
-An important contribution to this literature is the use of empirical strategies to analyze how firm behavior reveals attributes of the underlying objective function (add citations).
-In this paper, I leverage a large-scale shock to nonprofit hospitals in the US to investigate whether characteristics of the objective function is revealed by the composition of their leadership team.
+An important contribution to this literature is the use of empirical strategies to analyze how firm behavior reveals attributes of the underlying objective function (\cite{sloan2000not}).
+In this paper, I leverage a large-scale shock to nonprofit hospitals in the US to investigate whether characteristics of the objective function are revealed by the composition of their leadership team.
 
@@
 This is because, first, there are different hospital ownership types within the industry, giving researchers the ability to compare the behaviors of nonprofits with that of for-profits.
-Second, nonprofit hospitals make up over 50\% of all hospitals in the US, and staff, on average, 207 beds.
+Second, nonprofit hospitals make up 50\% of all hospitals in the US, and staff, on average, 207 beds.
 For comparison, for-profits make up around 36\% of hospitals and staff 107 beds (\cite{ASPE_2023}).
 Thus, nonprofit behavior in the hospital industry is relevant and directly affects consumers of health care.
-However, it remains unclear what type of objective function drives nonprofit hospitals (Erus \& Weisbrod, 2002; Deneffee \& Masson, 2002; Horwitz \& Nichols, 2009).
+However, it remains unclear what type of objective function drives nonprofit hospitals (\cite{erus2002inferring}; \cite{deneffe2002not}; \cite{horwitz2009hospital}).
 
@@
   Using publicly available tax forms, I construct a novel data set of nonprofit hospitals from 2009-2015 which captures the identities of executives tied to that hospital.
-I link hospitals in this data to hospital level characteristics in the American Hospital Association survey and Hospital Compare data.
-In order to capture the effect of clinical experience in executives, not just the selection of clinical executives into certain types of hospitals, I specifically investigate hospitals penalized by the Hospital Readmissions Reduction Program (HRRP) of 2012.
+I link hospitals in this data to characteristics in the American Hospital Association survey and Hospital Compare data.
+In order to capture the effect of clinical executives, not just their selection into certain types of hospitals, I specifically investigate hospitals penalized by the Hospital Readmissions Reduction Program (HRRP) of 2012.
 This program reduced Medicare payments to hospitals with above average readmission rates, and is plausibly uncorrelated with previously determined leadership teams.
 I further limit the sample to hospitals with stable executive teams over time, and characterize hospitals as having clinical experience on their team or not.
+
   I estimate the difference in readmission and mortality rates between hospitals with and without clinical experience before and after the penalties began, conditional on being penalized.
-I assume that before the penalty, clinical and non-clinical team hospitals had similar readmission and mortality trends, and that the composition of hospital executive teams determined prior to the penalties is, conditional on being penalized, exogenous.
-Under these assumptions, I identify the effect of having a clinical executive on a hospital’s response to the penalty measured in readmission and mortality rates.
+While readmissions directly affects the likelihood of being penalized, mortality rates are an explicit quality measure unrelated to the penalties.
+Including this as an outcome can help distinguish whether hospitals are improving quality or taking measures to only reduce readmissions.
 
-    I find that
+  I assume that before the penalty, clinical and non-clinical team hospitals behaved similarly along quality dimensions, and that the composition of hospital executive teams determined prior to the penalties is, conditional on being penalized, exogenous.
+Under these assumptions, I identify the effect of having a clinical executive on response to the penalty.
+I find that while both types of hospitals, those with clinical executives and those without clinical executives, decrease readmissions after being penalized, non-clinical team hospitals decrease readmissions at a faster rate.
+The difference equates to approximately 15 additional readmission cases per year for clinical teams, and there is no evidence that this result is driven by differences in readmissions prior to penalties.
+I find no differential effect of the program on mortality rates.
 
-    I contribute to hospital objectives
+    This paper contributes to three strands of literature.
+First, it adds to our understanding of what underlies nonprofit hospital behavior.
+This literature is extensive, and a review of it is beyond the scope of this paper.
+To summarize, there are various aspects of nonprofit behavior and objectives for which the evidence of how much it differs from for-profits remains mixed: costs, uncompensated care, technology adoption, and quality (\cite{sloan2000not}; \cite{eggleston2008hospital}; \cite{moscelli2018effect}; \cite{moscone2020public}).
+What underlies this is the need to understand the ways in which nonprofits make choices differently than for-profits.
+To this point, researchers have found that nonprofits do not act as purely profit maximizers, but maximize some combination of profit and output (\cite{deneffe2002not}, \cite{chang2011nonprofit}).
+A paper with a similar methodology to investigate hospital behavior is \citeauthor{chang2011nonprofit} (\citeyear{chang2011nonprofit}), which looks at the effect of a cost shock on likely of shutting down and mix of profitable services.
+They find that nonprofits and for-profits are equally as likely to shut down after the shock, but only nonprofits adjust their mix of profitable services.
+Similarly, I investigate a financial shock to hospitals, but I look at the differential impacts of leadership teams within nonprofits.
+This paper adds to our understanding of inputs into nonprofit hospital objective functions, and whether this is dependent on leadership compositions.
 
-    I contribute to HRRP
+    Second, many researchers have documented a correlation between executive backgrounds and firm performance specifically for CEOs, starting with \citeauthor{bertrand2003managing} (\citeyear{bertrand2003managing}), who show that manager fixed effects are an important driver of many firm behaviors and decisions.
+Female board members are associated with better oversight and more women executives (\cite{matsa2011chipping}; \cite{adams2009women}), but are negatively correlated with firm value (\cite{ahern2012changing}).
+Having more female executives is correlated with female employee wages and corporate strategies (\cite{flabbi2019female}; \cite{matsa2013female}); young male CEOs tend to be more aggressive in mergers and acquisitions while those with military experience are less aggressive (\cite{levi2010deal}; \cite{benmelech2015military}); CEOs with general ability tend to receive higher pay and perform better (\cite{kaplan2012ceo}; \cite{custodio2013generalists}, \cite{adams2018director}; \cite{frydman2019rising}).
+Further, Chief Diversity Officers are not found to have any effect on hiring more diversely in universities (\cite{bradley2022impact}).
+Due to data limitations, these studies focus mostly on large, publicly traded firms in the US.
+However, one study uses data from England to study whether CEOs affect hospital production, and find no association (\cite{janke2019impact}).
+I contribute to this literature first by investigating a context which is not largely studied but is policy relevant, nonprofit US hospitals, and considering a unique executive characteristic pertinent to that context.
+Second, I contribute by addressing leadership team endogeneity concerns through leveraging a shock to hospitals and strategically limiting the sample.
 
-    I contribute to executives affect on firms
+    Finally, I contribute to our understanding of how providers respond to pay-for-performance incentives in health care.
+The most in depth study of how hospitals respond to HRRP is \citeauthor{gupta2021impacts} (\citeyear{gupta2021impacts}).
+The author finds that hospitals decreased readmissions by 5\% and mortality rates by 2\% on average as a result of the program, confirming prior studies (\cite{mellor2017does}; \cite{ziedan2018essays}; \cite{ody2019decreases}; \cite{gupta2021impacts}).
+Importantly, around 40\% of the decrease in readmissions is due to selective patient practices.
+This raises important questions about how to incentivize improvement in quality of care without also incentivizing gamesmanship.
+This paper addresses a necessary first step in considering whether leadership compositions could affect likelihood of gaming, which is to first investigate whether leadership matters to hospital decisions at all.
 
-    \section{Background}
+    \section{Institutional Setting}
 
-    \subsection{Hospital Readmissions Reduction Program}
+    \subsection{Hospital Readmissions Reduction Program}\label{sec:hrrp}
 
+    The Hospital Readmissions Reduction Program (HRRP) was passed as a part of the Affordable Care Act in 2010 as a pay-for-performance incentive for hospitals to increase quality.
+In 2011, the Center for Medicare and Medicaid Services (CMS) released a set of rules that mandated penalties for hospitals with above average readmission rates.
 A readmission is when a patient returns to the hospital within 30 days of being discharged from a previous stay; avoidable readmissions are a bad outcome for patients and increase health care spending.
-The Hospital Readmissions Reduction Program (HRRP) was enacted in 2010 as a part of the Affordable Care Act to incentivize hospitals to reduce readmission rates.
-The goal of the program was to get hospitals to lower readmissions through better care coordination, less initial stay complications, and better post-care instructions.
-Beginning in October 2012, hospitals with higher readmission rates than the national average in pneumonia, heart failure, or AMI (after adjusting for demographic characteristics faced by the hospital) receive a lower reimbursement rate for all Medicare patients they treat.
-The penalties are based on each hospital's past performance in readmissions, which means hospitals had incentive to react immediately once details of the program were announced in October of 2011.
+The goal of HRRP is to lower readmissions through better care coordination, less initial stay complications, and better post-care instructions.
+Beginning in October 2012, hospitals with higher readmission rates than the national average in pneumonia, heart failure, or AMI (after adjusting for demographic characteristics) receive a lower reimbursement rate for Medicare patients.
+In 2015, CMS also included chronic obstructive pulmonary disease, coronary artery bypass graft surgery, and elective primary total hip arthroplasty and/or total knee arthroplasty as conditions which go into the penalty calculation.
+
+    Penalized hospitals pay a flat rate for every Medicare patient, not just those for which they have too high readmissions.
+Further, CMS does not distinguish a necessary readmission vs.
+an avoidable readmission, any repeat hospital visit is included in the penalty calculation.
+Excess readmission rates are calculated using a rolling look-back period of 3 years to determine whether the hospital is penalized.
+Therefore, hospitals had incentive to react immediately once details of the program were announced in October of 2011.
+These penalties are not insignificant; penalized hospitals paid, on average, 4-5\% of revenue.
 
@@
 They select members of the executive team to carry out day-to-day operations.
-The executive team usually consists of at least a CEO and CFO.
+While hospital executives typically do specialize by getting degree(s) in healthcare management or an MBA specific to healthcare, they do not often have hands on clinical experience.
+When there is a hospital executive with clinical experience, they likely decided on that career path after becoming a doctor.
+While some doctors earn additional degrees before stepping into an executive role, this is not a necessary condition to becoming a physician executive.
+An executive team usually consists of at least a CEO and CFO.
 Every nonprofit is different in exactly how they structure an executive team.
 Some nonprofits hire Chief Medical Officers, Chief Quality Officers, executive directors, vice presidents of various departments, etc.
-I consider all of these employees as part of the day-to-day management of the hospital and include them in the executive team.
+I consider all of these employees as part of the day-to-day management of the hospital and include them as a part of the executive team.
+
+    All tax-exempt organizations in the US are required to file a Form 990 with the IRS each year.
+There are different types of forms, but any organization that grosses over \$200,000 must file the most extensive form 990.
+Sections of the form include a statement of revenue, statement of functional expenses, a balance sheet, and, as used in this project, a list of all key employees, executives, and board members.
+In this section, each firm is required to report the name and title, average hours per week, position, and compensation of their board members and executive leaders.
+Since these firms are tax exempt, their financial information is made publicly available through the public housing of these forms.
 
@@
 
-    To characterize hospitals as having clinical experience on executive team or not, I construct a novel data set of nonprofit hospitals in the US which contains names and select characteristics of executives tied to the hospital from 2009-2015.
-I gather this data from each hospital's publicly available Tax Form 990, which is required each year for all nonprofits in the US.
+    To characterize hospitals as having clinical experience on their executive team or not, I construct a novel data set of nonprofit hospitals in the US that contains names and select characteristics of executives tied to the hospital from 2009-2015.
+I gather this data from each hospital's publicly available Tax Form 990, which every sufficiently large nonprofit files each year in the US.
 To my knowledge, this is the first large scale gathering of leadership names from these forms.
-I include a very detailed account of the collection of this data in Appendix \ref{appendixdata}.
 
-    The tax forms for all nonprofits are housed by ProPublica\footnote{https://projects.propublica.org/nonprofits/}.
+    The historical tax forms for all nonprofits are housed by ProPublica\footnote{https://projects.propublica.org/nonprofits/}.
 I use the NonProfit Explorer API to extract the Employee Identification Numbers (EIN) of all nonprofits classified as a hospital according to their National Taxonomy of Exempt Entities (NTEE) code.
-After filtering out associations and specialty hospitals, there are around 3000 nonprofit hospitals left.
-Importantly, I need to be able to link EINs to other hospital characteristics, including penalty status and readmission/mortality rates, in the American Hospital Association annual survey and Hospital Compare data.
-No official crosswalk between the two identification methods exists, so I must rely on matching via location and name of hospital in both data sets.
-After limiting to only general, nonprofit hospitals in AHA, I confidently match approximately 1200 EINs to an AHA identification number based on exact name matches within the same state.
+After filtering out associations and specialty hospitals, there are around 3,000 firms left.
+Importantly, I need to be able to link hospitals in the tax forms to other hospital characteristics, including penalty status and readmission rates, in the American Hospital Association annual survey and Hospital Compare data.
+No official crosswalk between EINs and AHA IDs or Medicare numbers exists, so I must rely on matching by location and name of hospital in both data sets.
+After limiting to only general, nonprofit hospitals in the AHA survey, I confidently match approximately 1200 EINs to an AHA identification number based on exact name matches within the same state.
 
-    For each of the 1200 EINs, I extract urls to 990 pdfs in each year.
-I download these locally and use OCR text extraction methods to scrape the relevant section of the pdf, which is titled ``Officers, Directors, Trustees, Key Employees, and Highest Compensated Employees".
+    For each of the 1200 EINs, I extract web URLs to Form 990 PDFs in each year.
+I download these locally and use optical character recognition (OCR) text extraction methods to scrape the relevant section of the pdf, which is titled ``Officers, Directors, Trustees, Key Employees, and Highest Compensated Employees".
 Finally, I use string cleaning methods to identify names and positions for each EIN, year.
-I also identify titles indicating the person is a physician: md, dr., or do.
-Using OCR extraction can be unreliable in some instances where the text is handwritten or the font is too small.
+I also identify titles indicating a physician: md, dr., or do.
+Using OCR extraction can be unreliable in some instances when the text is handwritten or the font is too small.
 Therefore, I manually find names, positions, and titles for observations that are missing after the initial extraction.
+This yields 850 EINs that are matched to AHA data, and are not missing text information.
 
-    My identification strategy relies on hospitals which have stable executive leadership characteristics over time.
-The ideal sample would be hospitals who have the same people on their team for the entire sample.
+    Due to the lack of consistency in organization of executive teams, I focus mainly on the executive teams as a whole instead of specific positions.
+I drop anyone who is solely a board member, and focus on those who have executive in their title or are labeled as a president or vice president of a specific aspect of the hospital.
+My identification strategy relies on hospitals that have stable executive leadership over time.
+The ideal set of hospitals would be those who have the same executives on their team for the entire sample, as changes could be correlated with the fact that the hospital is penalized.
 However, this rarely occurs in reality.
-Therefore, I proxy for a stable leadership in two ways.
-First, since my main focus is clinical experience, I limit the sample to hospitals who do not change the presence of a doctor on their leadership team from 2010-2014.
-Second, I increase the restriction by limiting the sample again to those also who do not change their CEO from 2010-2014.
+Therefore, I attempt to proxy for a stable leadership team while still having a large enough sample to run an analysis.
+Since my focus is clinical experience, my preferred specification is to limit the sample to hospitals who do not change the presence of a doctor on their leadership team from 2010-2014 (2 years before and after the program began).
+I make this limitation to hospitals in the sample and am left with 524 stable hospitals over time.
 Finally, I aggregate the information up to the hospital level, where an observation captures whether the hospital has clinical experience on their executive team over the course of the sample.
@@
 
-    I merge this hospital-level data with the American Hospital Association (AHA) survey using the EIN-AHA links I created based on hospital names.
-From AHA, I gather the number of beds in the hospital and the Medicare number, which I use to merge in the Hospital Compare data.
-Hospital Compare contains multiple data sets which I use to identify hospitals that were penalized in the Hospital Readmissions Reduction Program (HRRP), along with yearly readmission and mortality rates for the relevant HRRP conditions: heart failure, heart attack, and pneumonia, my outcomes of interest.
+    I merge the leadership information with the AHA survey using the EIN-AHA links I created.
+From AHA, I gather the number of beds in the hospital and the Medicare number, which I use to merge the Hospital Compare data.
+Hospital Compare contains multiple data sets which I use first to limit to hospitals that were penalized in the Hospital Readmissions Reduction Program (HRRP), and also for information on my outcome variables, yearly readmission rates for the relevant HRRP conditions: heart failure, heart attack (AMI), and pneumonia.
 
-    The Hospital Compare excess readmissions files contain hospital level information on the excess readmission measure used by CMS to determine whether a hospital should be penalized.
-This variable is also condition specific, where a hospital is penalized if any of the three excess readmission ratios are greater than one.
-Further, Hospital Compare also publishes hospital level readmission and mortality rates for heart failure, heart attack, and pneumonia.
-For my main outcomes of interest, I take an average across conditions and weight it using the total number of patients in that condition.
-For further analysis, I create samples of hospitals who were penalized for specific conditions and investigate the change in readmission and mortality rates for the condition they were penalized for.
+    The Hospital Compare excess readmissions files contain information on the excess readmission measure used by CMS to determine whether a hospital should be penalized.
+Specifically, CMS calculates the average readmission rate of all hospitals, adjusts for demographic factors, and determines the amount of readmissions a hospital has in excess to of the average.
+This variable is condition specific, where a hospital is penalized if any of the three excess readmission ratios are greater than one\footnote{Additional conditions were added to the penalty calculations in 2015, but I do not consider these since my sample does not extend past 2015}.
+Further, Hospital Compare also publishes hospital level readmission and mortality rates for heart failure, AMI, and pneumonia.
+For my main outcomes of interest, I take an average across conditions and weight it using the total number of patients seen for that condition.
+For further analysis, I create samples of hospitals who were penalized for specific conditions, and investigate the change in readmission and mortality rates for the condition they were penalized in.
 
@@
 
-    Figure \ref{weighted_read_mort_graph} shows the average of the main outcomes across all penalized hospitals over time.
-While mortality rates did not seem to change much after the penalties began in 2012, readmission rates decreased after the penalty for hospitals with and without clinical experience.
-However, readmission rates for non-clinical hospitals decreased more dramatically than those for clinical hospitals, even dropping below those for clinical teams in 2013 and 2014.
-While suggestive, this motivates further analysis into the differential responses between clinical and non-clinical research teams.
+    In Table \ref{tab:sumstats}, I present summary statistics for various hospital characteristics including which condition hospitals are penalized for, and the readmission and mortality outcome variables.
+After limiting to penalized hospitals with stable leadership teams, my sample consists of between 302-310 hospitals from 2009-2015, except in the case of AMI outcomes, which is only 225-264 hospitals due to small patient numbers.
+Fourteen percent of hospitals have someone with clinical experience on their executive team, and the size of the hospitals in the sample varies quite a bit.
+The average number of beds for a hospital in the sample is 186, only slightly lower than the national average for all nonprofits.
+Hospitals in the sample are less likely to be penalized for AMI than either heart failure or pneumonia.
+On average, 21\% of patients with one of these conditions would be readmitted.
+While readmission rates for the specific conditions are similar, the highest likelihood of readmission is for heart failure patients.
+Alternatively, AMI is the most dangerous condition in terms of mortality rates.
 
-    \begin{figure}[h!]
-        \centering
+    Figure \ref{fig:weighted_read_mort_graph} shows average readmission rates for hospitals in the sample across time and across the three different conditions plus the weighted average.
+There are multiple notable points the raw data reveals.
+First, prior to the penalty, hospitals without clinical executives had consistently higher readmission rates than clinical executive hospitals across all conditions.
+While there is a gap between clinical team and non-clinical team readmission rates, the lines seem to move mostly parallel.
+In contrast, 2012 begins a decline in readmission rates over all conditions for all types of hospitals.
+This makes sense and is consistent with previous literature on HRRP, as hospitals have incentive to reduce readmissions in order to reduce future penalties.
+However, non-clinical hospital teams decrease readmissions at a faster rate, quickly converging with the rates of clinical team hospitals.
+While suggestive, this motivates further analysis into the differential responses between clinical and non-clinical leadership teams.
+
+    \begin{figure}[ht!]
         \caption{Readmission rates over time in all penalized hospitals}
-        \includegraphics[scale=.6]{Objects/weighted_read_mort_graph.pdf}
-        \label{weighted_read_mort_graph}
+        \includegraphics[scale=.55]{Objects/weighted_read_graph.pdf}
+        \label{fig:weighted_read_mort_graph}
     \end{figure}
@@
 
-    Not knowing the true form of nonprofit hospital objective functions, we can assume that they maximize some weighted average of revenue and quality.
-At a basic level, we can think of nonprofits behaving according to:
+    Not knowing the true form of nonprofit hospital objective functions, we can assume that they maximize some weighted average of revenue and output.
+Abstracting away from quantity for the purpose of this paper, we can think of firms behaving according to:
 
-    $$\max\hspace{2mm} \delta R + (1-\delta) u(\theta)$$
+    $$\max\hspace{2mm} \delta R + (1-\delta) u(\theta),$$
 
-    \noindent where $R$ is net revenue, $u(.)$ is an increasing function that captures utility from quality of care, and $\theta$ is some quality of care measure.
-In this setting, we care about hospital responding to a penalty by adjusting readmission rates, so $\theta = 1-r$ where $r$ is the readmission rate.
+    \noindent where nonprofits differ from for-profits in that $\delta\neq 1$.
+Then $R$ is net revenue, $u(.)$ is an increasing function that captures utility from quality of care, and $\theta$ is some quality of care measure.
+In this setting, we care about hospital responding to a penalty by adjusting readmission rates, so $\theta = 1-r$, where $r$ is the readmission rate.
 While previous research has set out to understand the average $\delta$ for nonprofit hospitals, $\delta$ may also be dependent on the leadership teams within nonprofit hospitals.
@@
     \noindent where $L$ captures the composition of the leadership team.
-Therefore, when maximizing this function, $\theta$ is dependent on $L$ through its impact on how the hospital values profit vs.
+Therefore, when maximizing this function, the choice of $\theta$ is dependent on $L$ through its impact on how the hospital values profit vs.
 quality of care.
@@
     Thus, I directly estimate the effect of having clinical experience on readmission rates.
-However, looking purely at this correlation is likely a biased picture of the effect of clinical experience since it likely captures reverse causality: certain types of hospitals may be more likely to hire doctors as executives.
+However, looking purely at this correlation is a biased picture of the effect of clinical experience since it likely captures reverse causality: certain types of hospitals may be more likely to hire doctors as executives.
 I do two things to address this issue.
@@
 
-    $$y_{ht} = \beta (clinical \times post\_penalty)_{ht} + \alpha_{h} + \delta_t + \epsilon_{ht}$$
+    \begin{equation}
+    \label{eq:main}
+    y_{ht} = \beta (clinical \times post\_penalty)_{ht} + \alpha_{h} + \delta_t + \epsilon_{ht}
+    \end{equation}
 
     \noindent where $y_{ht}$ is one of the outcome variables discussed in Section \ref{sec:data}, $\beta$ is the variable of interest capturing the effect of having clinical experience after the penalties took place, $\alpha_h$ and $\delta_t$ are hospital and time fixed effects, respectively.
+As discussed in Section \ref{sec:hrrp}, hospitals had incentive to decrease readmissions immediately when the rules were released in October of 2011, before the penalties were actually determined in October of 2012.
+However, I still define $post\_penalty$ as occurring in 2012 or later, the first full year when hospitals would want to respond.
 
@@
 
-    $$y_{ht} = \sum_{j=2009}^{2010}\beta_j(clinical_{ht} \times \mathbf{1}\{y=j\}) + \sum_{j=2012}^{2015}\beta_j (clinical_{ht} \times \mathbf{1}\{y=j\}) + \alpha_{h} + \delta_t + \epsilon_{ht}.$$
+    \begin{equation}
+    \label{eq:es}
+    y_{ht} = \sum_{j=2009}^{2010}\beta_j(clinical_{ht} \times \mathbf{1}\{y=j\}) + \sum_{j=2012}^{2015}\beta_j (clinical_{ht} \times \mathbf{1}\{y=j\}) + \alpha_{h} + \delta_t + \epsilon_{ht}.
+    \end{equation}
 
@@
 
+    \subsection{Readmission Rates}
+
+    I present the estimated coefficient of interest from Equation \ref{eq:main}, $\beta$, in Table \ref{tab:mainresults}.
+In column 1, the outcome is the weighted average of readmission rates.
+The estimated effect of having clinical experience on the hospital's reaction to the penalty is .47, indicating that hospitals with clinical experience on their executive team have a .47 point higher readmission rate than non-clinical team hospitals after being penalized.
+Comparing this to an average readmission rate of 21, this is a difference of approximately 2\%.
+When considering an average total number of patients of 760, this is a difference of 15 readmissions per year.
+Based on previous studies and on Figure \ref{fig:weighted_read_mort_graph}, both types of hospitals decrease readmission rates after the penalties begin, so a positive $\beta$ indicates that non-clinical team hospitals decrease readmissions at a faster rate than clinical team hospitals.
+
     \import{Tables}{read_results_table.tex}
 
-    \begin{figure}
-        \centering
+    In columns 2, 3, and 4, I limit the sample to hospitals penalized specifically for the condition that matches the outcome in the model.
+The coefficients are consistent and similar to that of the weighted average coefficient.
+Therefore, one of the conditions is not driving the entire weighted average.
+While similar in magnitude, the estimate for $\beta$ with the outcome of AMI readmission rate is statistically insignificant.
+Out of all the outcomes, AMI is the most likely to be missing data due to insufficient number of patients, which could explain this.
+
+    Finally, I plot estimated coefficients from Equation \ref{eq:es} in Figure \ref{fig:wa_eventstudy}.
+The result is consistent with clinical team hospitals having higher readmissions post-penalty.
+The magnitude of the effect in the years after the penalty began are still around .5, but range from .26 to .54 from 2012-2014.
+Further, there is no indication that having clinical experience on the executive team causes a reaction in readmissions prior to the penalty.
+
+    \begin{figure}[ht!]
         \caption{Event Study Estimates}
-        \includegraphics[scale=.5]{Objects/wa_eventstudy.pdf}
+        \includegraphics[scale=.5]{Objects/read_es_graph.pdf}
         \label{fig:wa_eventstudy}
@@
 
+    \subsection{Mortality Rates}
+
+    \import{Tables}{mort_results_table.tex}
+
     \section{Conclusion}
 
+    In this paper, I study the effect of having clinical experience within hospital leadership on nonprofit hospital behavior.
+Specifically, I leverage the Hospital Readmissions Reduction Program, which began in 2012, to identify similar quality hospitals along the dimension of readmission rates.
+For these similar hospitals, I estimate the difference in readmissions after being penalized between hospitals with clinical executives and hospitals without clinical executives.
+I find that while the penalty program led both types of hospitals to decrease readmissions after being penalized, those without clinical experience on their executive decreased rates more quickly than non-clinical team hospitals.
+
+    This finding reveals important information about hospital objectives, specifically that the composition of a hospital's leadership team affects the value placed on profit (or in this case, avoiding loss) vs.
+quality of care.
+Therefore, not-for-profit vs.
+for-profit status should not be the only characteristic considered by policymakers when designing incentives to increase quality of care.
+
 	\newpage
-	\appendix
 
-    \section{Data}\label{appendixdata}
+    \printbibliography
 
-    \subsection{Gathering Hospital Leadership Names}
+	% \appendix
 
-    There is no perfect way to access tax form 990s in bulk over the time period I am considering in this paper.
+ %    \section{Data}\label{appendixdata}
+
+ %    \subsection{Gathering Hospital Leadership Names}
+
+ %    There is no perfect way to access tax form 990s in bulk over the time period I am considering in this paper.
 However, using the NonProfit Explorer API seems to be the most straightforward.
@@
 
-    There are over 1.5 million nonprofit entities in the US (cite), making it crucially important to be able to filter by type of entity before analyzing any PDFs.
+ %    There are over 1.5 million nonprofit entities in the US (cite), making it crucially important to be able to filter by type of entity before analyzing any PDFs.
 The API allows this by filtering a query based on National Taxonomy of Exempt Entities (NTEE) code.
@@
 
-    I loop through the list of EINs found in the previous step and query more detailed information from the API on that specific EIN.
+ %    I loop through the list of EINs found in the previous step and query more detailed information from the API on that specific EIN.
 I save the name, secondary name, state, and zip code, all of which do not vary by year.
@@
 
-    It is crucial to the analysis to be able to link these nonprofits with other sources of data to recover penalties from HRRP, bed size, and outcomes of interest (definitely need to talk more about outcomes).
+ %    It is crucial to the analysis to be able to link these nonprofits with other sources of data to recover penalties from HRRP, bed size, and outcomes of interest (definitely need to talk more about outcomes).
 Therefore, I take a conservative approach to matching EINs to American Hospital Association (AHA) ID, which links fairly easily to Medicare ID numbers, based on name and location.
@@
 
-    I download the AHA data from Wharton Research Data Services.
+ %    I download the AHA data from Wharton Research Data Services.
 I filter only to hospitals in the contiguous US, Alaska, and Hawaii (excluding places like Puerto Rico), classified as nonprofit or state/community, and those that are general acute care.
@@
 
-    I consider which nonprofit entities are not likely to be hospitals and drop them.
+ %    I consider which nonprofit entities are not likely to be hospitals and drop them.
 There are numerous foundations or auxiliary nonprofits with the purpose of raising funds for the hospital, but do not actually care for patient.
@@
 
-    I then proceed matching based on names in multiple layers.
+ %    I then proceed matching based on names in multiple layers.
 I focus on exact string matches, so I remove all spaces and common characters that could cause mismatches such as \&, ', -, and inc.
@@
 
-    In the next layer of matching, I remove common words such as "healthcare", "regional", "hospital", etc.
+ %    In the next layer of matching, I remove common words such as "healthcare", "regional", "hospital", etc.
 That way if there are subtle differences in names, removing common words may allow for an exact match.
@@
 
-    In some cases, tax forms are associated with a system of hospitals instead of one individual hospital.
+ %    In some cases, tax forms are associated with a system of hospitals instead of one individual hospital.
 Thus, I create another variable that captures the names of systems which match.
@@
 
-    (add a paragraph about validity of matches)
+ %    (add a paragraph about validity of matches)
 
-    Now that I have a sample of hospitals, the next step is to extract the names of the people in charge of these hospitals from the Tax Form 990 PDFs.
+ %    Now that I have a sample of hospitals, the next step is to extract the names of the people in charge of these hospitals from the Tax Form 990 PDFs.
 In the data set of hospital PDF URLs that I collected earlier, I limit to the hospitals with solid matches described above.
@@
 
-    One tricky aspect of the NonProfit Explorer API is that, only in some cases, if two forms are present for an EIN, year, only the first one (which is typically not the one with the relevant information) is pulled.
+ %    One tricky aspect of the NonProfit Explorer API is that, only in some cases, if two forms are present for an EIN, year, only the first one (which is typically not the one with the relevant information) is pulled.
 Therefore, for some hospitals, a couple years will have gaps in text extraction data.
@@
 
-    The form of the text data is a data frame with one column, where each line of text is saved in a different row.
+ %    The form of the text data is a data frame with one column, where each line of text is saved in a different row.
 I write a text cleaning package that locates names, positions, titles, and indications of resigning.
@@
 
-    Typically on the same page as the names and positions is a list of the highest compensated employees and their compensation.
+ %    Typically on the same page as the names and positions is a list of the highest compensated employees and their compensation.
 In order to not record extra names, I filter out any rows after the start of this section.
@@
 
-    Next, I locate first and last names in the data.
+ %    Next, I locate first and last names in the data.
 

### 2024-04-23 RScripts/HCRIS_v1996.R
update scripts and add distribution map

@@ -32,17 +32,47 @@ hcris.vars = rbind(hcris.vars,c('cash','G000000','00100','0100','numeric'))
 hcris.vars = rbind(hcris.vars,c('tot_pat_rev','G300000','00100','0100','numeric'))
 hcris.vars = rbind(hcris.vars,c('allowance','G300000','00200','0100','numeric'))
 hcris.vars = rbind(hcris.vars,c('net_pat_rev','G300000','00300','0100','numeric'))
+
+# variables added by me
+hcris.vars = rbind(hcris.vars,c('land_purch','A700001','00100','0200','numeric'))
+hcris.vars = rbind(hcris.vars,c('land_assets','A700001','00100','0700','numeric'))
+hcris.vars = rbind(hcris.vars,c('landimpr_purch','A700001','00200','0200','numeric'))
+hcris.vars = rbind(hcris.vars,c('landimpr_assets','A700001','00200','0700','numeric'))
+hcris.vars = rbind(hcris.vars,c('build_purch','A700001','00300','0200','numeric'))
+hcris.vars = rbind(hcris.vars,c('build_assets','A700001','00300','0700','numeric'))
+hcris.vars = rbind(hcris.vars,c('buildimpr_purch','A700001','00400','0200','numeric'))
+hcris.vars = rbind(hcris.vars,c('buildimpr_assets','A700001','00400','0700','numeric'))
+hcris.vars = rbind(hcris.vars,c('fixedequipment_purch','A700001','00500','0200','numeric'))
+hcris.vars = rbind(hcris.vars,c('fixedequipment_assets','A700001','00500','0700','numeric'))
+hcris.vars = rbind(hcris.vars,c('movableequipment_purch','A700001','00600','0200','numeric'))
+hcris.vars = rbind(hcris.vars,c('moveableequipment_assets','A700001','00600','0700','numeric'))
+hcris.vars = rbind(hcris.vars,c('total_purch','A700001','00900','0200','numeric'))
+hcris.vars = rbind(hcris.vars,c('total_assets','A700001','00900','0700','numeric'))
+hcris.vars = rbind(hcris.vars,c('labor_costs','S300003','00600','0100','numeric'))
+hcris.vars = rbind(hcris.vars,c('capital_cost_buildings_old','A000000','00100','0200','numeric'))
[…]

### 2024-04-23 RScripts/HCRIS_v2010.R

@@ -36,17 +36,44 @@ hcris.vars = rbind(hcris.vars,c('cash','G000000','00100','00100','numeric'))
 hcris.vars = rbind(hcris.vars,c('tot_pat_rev','G300000','00100','00100','numeric'))
 hcris.vars = rbind(hcris.vars,c('allowance','G300000','00200','00100','numeric'))
 hcris.vars = rbind(hcris.vars,c('net_pat_rev','G300000','00300','00100','numeric'))
+
+# variables added by me
+hcris.vars = rbind(hcris.vars,c('land_purch','A700001','00100','00200','numeric'))
+hcris.vars = rbind(hcris.vars,c('land_assets','A700001','00100','00700','numeric'))
+hcris.vars = rbind(hcris.vars,c('landimpr_purch','A700001','00200','00200','numeric'))
+hcris.vars = rbind(hcris.vars,c('landimpr_assets','A700001','00200','00700','numeric'))
+hcris.vars = rbind(hcris.vars,c('build_purch','A700001','00300','00200','numeric'))
+hcris.vars = rbind(hcris.vars,c('build_assets','A700001','00300','00700','numeric'))
+hcris.vars = rbind(hcris.vars,c('buildimpr_purch','A700001','00400','00200','numeric'))
+hcris.vars = rbind(hcris.vars,c('buildimpr_assets','A700001','00400','00700','numeric'))
+hcris.vars = rbind(hcris.vars,c('fixedequipment_purch','A700001','00500','00200','numeric'))
+hcris.vars = rbind(hcris.vars,c('fixedequipment_assets','A700001','00500','00700','numeric'))
+hcris.vars = rbind(hcris.vars,c('movableequipment_purch','A700001','00600','00200','numeric'))
+hcris.vars = rbind(hcris.vars,c('moveableequipment_assets','A700001','00600','00700','numeric'))
+hcris.vars = rbind(hcris.vars,c('HIT_purch','A700001','00700','00200','numeric'))
+hcris.vars = rbind(hcris.vars,c('HIT_assets','A700001','00700','00700','numeric'))
+hcris.vars = rbind(hcris.vars,c('total_purch','A700001','01000','00200','numeric'))
[…]

### 2024-04-23 RScripts/analysis.R

@@ -1,0 +1,760 @@
+library(readr)
+library(dplyr)
+library(fixest)
+library(ggplot2)
+library(ggpubr)
+library(tidyr)
+library(synthdid)
+options(knitr.kable.NA=" ")
+
+#### ANALYSIS ##########
+
+# Comparisons I need: 
+# 1. FP - NFP
+# 2. FP - has MD
+# 3. FP - no MD
+# 4. has MD - no MD
+
+# Methodologies:
+# 1. full sample difference in difference
+# 2. Matched difference in difference
+# 3. Synthetic difference in difference
+
+# Outcomes:
+# 1. Readmission rates
+# 2. Mortality Rates
+# 3. Uncompensated Care
+# 4. Case Mix Index
+
+# limit to 2010 to 2014
+hospital_data <- readRDS(paste0(created_data_path, "all_hospital_data.rds")) %>%
+  filter(year>=2010 & year<=2014)
+
+hospital_data$uncomp_care <- abs(hospital_data$uncomp_care)
+hospital_data <- hospital_data %>%
+  mutate(uncomp_care=uncomp_care/1000000) 
+
+# fill ID
+hospital_data <- hospital_data %>%
+  group_by(MCRNUM) %>%
+  fill(ID, .direction="downup") %>%
+  ungroup() %>%
+  filter(!is.na(ID)) 
+
+# Remove hospitals that change from nonprofit to for-profit or vice versa
+hospital_data <- hospital_data %>%
+  mutate(post_2012=ifelse(year>2012, 1, 0),
+         NFP = ifelse(profit_status=="nonprofit", 1, 0),
+         FP = ifelse(profit_status=="forprofit",1,0)) %>%
+  group_by(ID) %>%
+  arrange(year) %>%
+  mutate(lag_NFP = dplyr::lag(NFP)) %>%
+  ungroup() %>%
+  filter(NFP==lag_NFP | year==2010)
+
+comp <- c("FP - NFP", "FP - has MD", "FP - no MD", "has MD - no MD")
+method <- c("full", "matched", "synth")
+
+### OUTCOME: READMISSION ####
+
+readmission_analysis <- lapply(comp, function(x){
+  data <- hospital_data 
+  
+  # create did variable and limit samples accordingly
+  if (x=="FP - NFP"){
+    data <- data %>%
+      mutate(did = post_2012*FP)
+  }
+  if (x=="FP - has MD"){
+    data <- data %>%
+      filter(!(NFP==1 & no_num_md_change_2010_2014==0)) %>%
[…]

### 2024-04-23 RScripts/data10_create_leadership_variables.R

@@ -3,7 +3,6 @@ library(dplyr)
 library(stringr)
 library(tidyr)
 
-source("paths.R")
 
 # read in names data
 cleaned_text <- read_rds(paste0(created_data_path, "cleaned_text.rds"))
@@ -76,11 +75,15 @@ cleaned_text <- cleaned_text %>%
 cleaned_text <- cleaned_text %>%
   mutate(ceo = ifelse(position1=="ceo" | position2 == "ceo" | position1 == "chief executive officer" | position2 == "chief executive officer", 1,0)) %>%
   mutate(ceo = ifelse(is.na(ceo),0,ceo))
+
+cleaned_text <- cleaned_text %>%
+  mutate(cmo = ifelse(position1=="cmo" | position2 == "cmo" | position1 == "chief medical officer" | position2 == "chief medical officer", 1,0)) %>%
+  mutate(cmo = ifelse(is.na(cmo),0,cmo))
   
 executive_data <- cleaned_text %>%
   mutate(vp = ifelse(pos1_vp==1 | pos2_vp==1, 1, 0)) %>%
   filter(position == "executive") %>%
-  select(ein, year, name, position, title, former, vp, ceo) %>%
+  select(ein, year, name, position, title, former, vp, ceo, cmo) %>%
   filter(!(former %in% c("past", "former", "past interim")))
 
 # loop through eins to fill in any missing years that happen between nonmissing years
@@ -91,6 +94,7 @@ executive_data_filled <- executive_data %>% filter(year==0)
 
 for (i in seq_along(eins)){
   data <- executive_data %>%
+    mutate(year=as.numeric(year)) %>%
     filter(ein == eins[[i]]) %>%
     group_by(name, position) %>%
     mutate(id = cur_group_id()) %>%
@@ -123,13 +127,20 @@ for (i in seq_along(eins)){
 # fill MD
 executive_data_filled <- executive_data_filled %>%
   group_by(ein, id) %>%
-  fill(title, vp, ceo, .direction="downup") %>%
+  fill(title, vp, ceo, cmo, .direction="downup") %>%
   ungroup()
+
+# now summarize executive teams
+executive_data <- executive_data_filled %>%
+  group_by(ein, year) %>%
+  mutate(num_execs = n())
+
+rm(executive_data_filled, name_data, data, cleaned_text)
 
 # goal 2: identify whether changes occur from year to year ####
 
[…]

### 2024-04-23 RScripts/data11_combine_hosp_level_data.R

@@ -5,6 +5,7 @@ library(stringr)
 library(fabricatr)
 library(knitr)
 library(kableExtra)
+library(MatchIt)
 
 # Read in Data ##################
 
@@ -12,112 +13,126 @@ library(kableExtra)
 AHA_ein_matches <- read_rds(paste0(created_data_path,"AHA_ein_matches.rds"))
 manual_matched_eins <- read_rds(paste0(created_data_path, "manual_matched_eins.rds"))
 # Data from AHA 
-AHA_hosp_names <- read_csv(paste0(created_data_path, "raw data/AHA_hosp_names.csv"))
+AHA <- read_csv(paste0(raw_data_path, "/allAHAvariables_20092017.csv"))
 # HCRIS data that has HRRP penalty information (and beds)
-final_HCRIS_v2010 <- read_csv(paste0(created_data_path, "raw data/final_HCRIS_v2010.csv"))
-final_HCRIS_v1996 <- read_csv(paste0(created_data_path, "raw data/final_HCRIS_v1996.csv"))
+HCRIS <- read_csv(paste0(raw_data_path, "/final_HCRIS_data.csv"))
 # Hospital Compare data that also has penalty information
-hc_readm_2012 <- read_csv(paste0(created_data_path, "raw data/Hospital Compare/vwhqi_readm_reduction_2012.csv"))
-hc_readm_2013 <- read_csv(paste0(created_data_path, "raw data/Hospital Compare/vwhqi_readm_reduction_2013.csv"))
-hc_readm_2014 <- read_csv(paste0(created_data_path, "raw data/Hospital Compare/vwhqi_readm_reduction_2014.csv"))
-hc_readm_2015 <- read_csv(paste0(created_data_path, "raw data/Hospital Compare/vwhqi_readm_reduction_2015.csv"))
+hc_readm_2012 <- read_csv(paste0(raw_data_path, "/Hospital Compare/vwhqi_readm_reduction_2012.csv"))
+hc_readm_2013 <- read_csv(paste0(raw_data_path, "/Hospital Compare/vwhqi_readm_reduction_2013.csv"))
+hc_readm_2014 <- read_csv(paste0(raw_data_path, "/Hospital Compare/vwhqi_readm_reduction_2014.csv"))
+hc_readm_2015 <- read_csv(paste0(raw_data_path, "/Hospital Compare/vwhqi_readm_reduction_2015.csv"))
 # Data that I created on leadership teams from the tax forms
[…]

### 2024-04-23 RScripts/data3_match_AHA_hospitals.R

@@ -6,16 +6,18 @@ library(tidyr)
 
 # This Script uses hospital names from the Nonprofit Explorer API and names from the AHA survey to match hospitals.
 
-##### ORDER : 3 ##########
-
+# Read in AHA data that includes hospital name and AHAID
 AHA <- read_csv(paste0(created_data_path, "/raw data/AHA_hosp_names.csv"))
+
+# Read in data from tax forms that includes EIN number and hospital name
 tax <- readRDS(paste0(created_data_path, "hospital_pdf_locations.rds"))
 
+# Get rid of symbols in AHA ID
 AHA$MNAME <- iconv(AHA$MNAME,from="ISO-8859-1")
 
 # Get AHA into cleaned data with one row per AHAID ####
 # Limit to nonprofit general hospitals
-AHA = AHA %>%
+AHA <- AHA %>%
   filter(FSTCD<=56 & nchar(MLOCZIP)>4 & (CNTRL==23 | (CNTRL>=12 & CNTRL<=16)) & SERV==10) %>%
   dplyr::rename(name.AHA=MNAME) %>%
   mutate(state=fips(FSTCD, to='Abbreviation')) %>%
@@ -36,7 +38,7 @@ AHA <- AHA %>%
   filter(!str_detect(name.AHA, "cancer"))
 
 
-# take a look at hospitals who have changes in ownership or system name
+# look at hospitals that have changes in ownership or system name
 AHA <- AHA %>%
   group_by(ID) %>%
   mutate(lag_sys=lag(SYSID)) %>%

### 2024-04-23 RScripts/data8_clean_rawtext_data.R

@@ -27,11 +27,11 @@ AHA_ein_list_manualmatched <- readRDS(paste0(created_data_path, "/AHA_ein_list_m
 AHA_ein_data_frame <- data.frame(first_name = character(), last_name=character(), title=character(), position=character())
 
 #### create the other lists as inputs to clean_text() ####
-lastnames <- read_excel(paste0(created_data_path, "raw data/app_c.xlsx")) %>%
+lastnames <- read_excel(paste0(raw_data_path, "/app_c.xlsx")) %>%
   filter(count>7000) %>%
   select(name) %>%
   mutate(name=tolower(name))
-firstnames <- read_csv(paste0(created_data_path, "raw data/yob2000.txt"), col_names = FALSE) %>% 
+firstnames <- read_csv(paste0(raw_data_path, "/yob2000.txt"), col_names = FALSE) %>% 
   filter(X3>50) %>%
   select(X1) %>%
   rename(name=X1) %>%
@@ -132,10 +132,9 @@ for (i in 1:length(AHA_ein_list_manualmatched)){
   }
 }
 
-cleaned_text <- readRDS(paste0(created_data_path, "cleaned_text.rds"))
 
 # Further clean text data
-cleaned_text <- cleaned_text %>%
+cleaned_text <- AHA_ein_data_frame %>%
   distinct() %>%
   filter(first_name!="" & last_name!="") %>%
   mutate(extra = str_remove(extra, "ae eaeeeeeeeaneeees")) %>%

### 2024-04-23 RScripts/data9_create_hosp_outcomes.R

@@ -6,27 +6,27 @@ library(stringr)
 # Read in data from hospital compare
 # focus first and mortality and readmissions data
 for (i in 2008){
-  year <- read_csv(paste0(created_data_path, "raw data/Hospital Compare/dbo_vwhqi_hosp_mortality_xwlk_",i,".csv"))
+  year <- read_csv(paste0(raw_data_path, "/Hospital Compare/dbo_vwhqi_hosp_mortality_xwlk_",i,".csv"))
   assign(paste0("mort",i), year)
 }
 for (i in 2009:2011){
-  year <- read_csv(paste0(created_data_path, "raw data/Hospital Compare/dbo_vwhqi_hosp_mortality_readm_xwlk_",i,".csv"))
+  year <- read_csv(paste0(raw_data_path, "/Hospital Compare/dbo_vwhqi_hosp_mortality_readm_xwlk_",i,".csv"))
   assign(paste0("mort",i), year)
 }
 for (i in 2012){
-  year <- read_csv(paste0(created_data_path, "raw data/Hospital Compare/vwhqi_hosp_mortality_",i,".csv"))
+  year <- read_csv(paste0(raw_data_path, "/Hospital Compare/vwhqi_hosp_mortality_",i,".csv"))
   assign(paste0("mort",i), year)
 }
 for (i in 2013){
-  year <- read_csv(paste0(created_data_path, "raw data/Hospital Compare/dbo_vwhqi_hosp_mortality_readm_xwlk_",i,".csv"))
+  year <- read_csv(paste0(raw_data_path, "/Hospital Compare/dbo_vwhqi_hosp_mortality_readm_xwlk_",i,".csv"))
   assign(paste0("mort",i), year)
 }
 for (i in 2014){
-  year <- read_csv(paste0(created_data_path, "raw data/Hospital Compare/hqi_hosp_readmcompdeath_",i,".csv"))
+  year <- read_csv(paste0(raw_data_path, "/Hospital Compare/hqi_hosp_readmcompdeath_",i,".csv"))
   assign(paste0("mort",i), year)
 }
 for (i in 2015){
-  year <- read_csv(paste0(created_data_path, "raw data/Hospital Compare/hqi_hosp_readmdeath_",i,".csv"))
+  year <- read_csv(paste0(raw_data_path, "/Hospital Compare/hqi_hosp_readmdeath_",i,".csv"))
   assign(paste0("mort",i), year)
 }
 
[…]

### 2024-04-23 RScripts/prelim_regressions.R

@@ -6,7 +6,7 @@ library(ggpubr)
 library(tidyr)
 
 # Read in hospital_data created in "create_hospital_data.R"
-hospital_data <- readRDS(paste0(created_data_path, "penalized_hospital_data(temp).rds")) %>%
+hospital_data <- readRDS(paste0(created_data_path, "all_hospital_data.rds")) %>%
   mutate(yr2009 = ifelse(year==2009,1,0),
          yr2010 = ifelse(year==2010,1,0),
          yr2011 = ifelse(year==2011,1,0),
@@ -22,7 +22,7 @@ hospital_data <- readRDS(paste0(created_data_path, "penalized_hospital_data(temp
          pneum_readmission = rate_pneum_readmission) %>%
   mutate(post_2012 = ifelse(year>=2012, 1, 0)) %>%
   mutate(post_2011 = ifelse(year>=2011, 1, 0)) %>%
-  mutate(has_any_md = ifelse(total_docs>0, 1, 0)) %>%
+  mutate(has_any_md = ifelse(num_doctors>0, 1, 0)) %>%
   mutate(did = has_any_md*post_2012) %>%
   mutate(did_2011 = has_any_md*post_2011) %>%
   mutate(es2009 = yr2009*has_any_md,
@@ -45,12 +45,18 @@ minyr_penalized <- hospital_data %>%
 hospital_data <- hospital_data %>%
   left_join(minyr_penalized, by="ID")
 
+# create sample of no md changes for all hospitals
+noMDchg_all_sample <- hospital_data %>%
+  filter(no_md_change_2010_2014==1) %>%
+  mutate(post = ifelse(year>=minyr_pen, 1, 0))
+
 
 # Create the 7 samples I want ####
 
 # 1. No MD changes, penalized
 noMDchg_pen_sample <- hospital_data %>%
   filter(no_md_changes_2010_2014==1) %>%
+  filter(ever_penalized==1) %>%
   select(ID, year, heartattack_mortality, heartattack_readmission, pneum_mortality, pneum_readmission, heartfailure_mortality, heartfailure_readmission, has_any_md, minyr_pen, yr2009, yr2010, yr2011, yr2012, yr2013, yr2014, yr2015, weightedavg_read, weightedavg_mort, post_2012, did, did_2011,
          es2009, es2010, es2011, es2012, es2013, es2014, es2015) %>%
   mutate(post = ifelse(year>=minyr_pen, 1, 0))
@@ -103,32 +109,142 @@ noMDchg_top_sample <- hospital_data %>%
[…]

### 2024-04-23 RScripts/summary_stats.R

@@ -3,173 +3,681 @@ library(dplyr)
 library(tidyr)
 library(knitr)
 library(kableExtra)
+library(weights)
+library(ggplot2)
+library(maps)
+library(mapdata)
+library(cdlTools)
+library(tibble)
 
-# Read in all data
-names_data <- read_rds(paste0(created_data_path, "names_data.rds"))
-#AHA_ein_matches <- read_rds(paste0(created_data_path,"AHA_ein_matches.rds"))
+# read in hospital data
+hospital_data <- read_rds(paste0(created_data_path, "all_hospital_data.rds"))
+
+# put uncompensated care in millions
+hospital_data$uncomp_care <- abs(hospital_data$uncomp_care)
+hospital_data <- hospital_data %>%
+  mutate(uncomp_care=uncomp_care/1000000) 
 
 
-  
+# List of tables:
+# Table 1: mean, SD, min, max for the whole sample
+# Table 2: means for different sub-samples
+
+### TABLE 1: mean, sd, min, max for the whole sample
+stats <- hospital_data %>% 
+  summarise_at(c("Number Beds"="beds",
+                 "Penalized for AMI"="ever_pen_ha", "Penalized for HF"="ever_pen_hf",
+                 "Penalized for Pneumonia"="ever_pen_pnem",
+                 "Weighted Avg. Readmission Rate"="weightedavg_read",
+                 "HF Readmission Rate" = "rate_heartfailure_readmission",
+                 "AMI Readmission Rate" = "rate_heartattack_readmission",
+                 "Pneum. Readmission Rate" = "rate_pneum_readmission",
+                 "Weighted Avg. Mortality Rate"="weightedavg_mort",
+                 "HF Mortality Rate" = "rate_heartfailure_mortality",
+                 "AMI Mortality Rate" = "rate_heartattack_mortality",
+                 "Pneum. Mortality Rate" = "rate_pneum_mortality",
+                 "Num. Patients HF" = "patnum_heartfailure_readmission",
+                 "Num. Patients AMI" = "patnum_heartattack_readmission",
+                 "Num. Patients Pneumonia" = "patnum_pneum_readmission",
+                 "Uncomp. Care (Millions)"="uncomp_care",
+                 "Case Mix Index"="cmiv"), 
+               list(m=mean,sd=sd, min=min, max=max), na.rm=TRUE) %>%
[…]

### 2024-04-23 RScripts/sumstats_profitstatus.R

@@ -1,0 +1,73 @@
+library(readr)
+library(dplyr)
+library(tidyr)
+library(knitr)
+library(kableExtra)
+library(weights)
+library(ggplot2)
+
+# create graphs exploring differences in outcomes between nonprofit and for-profit ownership status
+
+hospital_data <- read_rds(paste0(created_data_path, "all_hospital_data.rds")) %>%
+  filter(year!=2016 & year!=2017)
+
+# create "ever has MD" variable
+hospital_data <- hospital_data %>%
+  group_by(ID) %>%
+  mutate(ever_has_md = ifelse(sum(has_any_md, na.rm=T)>0,1,0)) %>%
+  ungroup() 
+
+# create variable for "has leadership info"
+hospital_data <- hospital_data %>%
+  group_by(ID) %>%
+  mutate(has_leader_info = ifelse(sum(num_execs,na.rm=T)>0,1,0)) %>%
+  ungroup()
+
+# compare weighted avg readmissions for: for profit, NFP w/ leadership info, NFP w/out leadership info
+hospital_data <- hospital_data %>%
+  mutate(profit_leadership_status = ifelse(profit_status=="forprofit","forprofit",NA)) %>%
+  mutate(profit_leadership_status = ifelse(profit_status=="nonprofit" & has_leader_info==1,"nonprofit w/ leader info",profit_leadership_status)) %>%
+  mutate(profit_leadership_status = ifelse(profit_status=="nonprofit" & has_leader_info==0,"nonprofit w/out leader info",profit_leadership_status)) 
+
+sum <- hospital_data %>%
+  group_by(year,profit_leadership_status) %>%
+  summarise_at(c("weightedavg_read"), list(mean), na.rm=T) %>%
+  filter(year!=2016 & year!=2017)
+
+ggplot(sum, aes(x=year, y=weightedavg_read, color=profit_leadership_status)) + geom_line()
+
+# compare weighted avg readmissions for: for profit, NFP w/ stable team, NFP w/out stable team
+hospital_data <- hospital_data %>%
+  group_by(MCRNUM) %>%
+  fill(no_md_change_2010_2014, .direction="downup") %>%
+  ungroup() %>%
+  mutate(profit_stability_status = ifelse(profit_status=="forprofit","forprofit",NA)) %>%
[…]

### 2024-04-23 Paper/paper.tex
Updates from Overleaf

@@
 \usepackage{url}
-\usepackage{natbib}
 %\usepackage{harvard}
@@
 
-	\linespread{1.2}\title{\vspace{-0.5in} Does Hospital Leadership Matter?\\ \large Evidence from the Hospital Readmissions Reduction Program}
+	\linespread{1.2}\title{\vspace{-0.5in} Does Hospital Leadership Matter?\\ \large Evidence from Pay-for-Performance Programs}
 
@@
 		{\small
-		The objectives that underlie nonprofit firm behaviors are not well understood, but perhaps that is because there are characteristics even within nonprofits that dictate behavior.
-In this paper, I use a novel data set on nonprofit hospital executives in the US from 2009-2015 to investigate whether having clinical experience on a hospital executive team affects hospital behavior.
-In particular, I leverage the Hospital Readmissions Reduction Program of 2012 as a shock to similar quality hospitals, and estimate the difference in response from hospitals with and without clinical experience.
-I find that while all penalized hospitals decreased readmissions after the shock, non-clinical executive team hospital decreased readmissions at a faster rate than those with a clinical executive team.
+        Nonprofit firms have diverse goals and missions that make understanding their true objectives and behaviors difficult.
+A setting where this is particularly relevant is that of nonprofit hospitals, which make up the majority of hospitals in the US, yet there is mixed evidence on what drives their behavior.
+In some ways they seem to act like for-profits, while in other ways they may care more about prestige or quantity.
+In this paper, I investigate whether the leadership team composition of nonprofit hospitals is correlated with more for-profit behavior.
+I use a novel data set on nonprofit hospital executives in the US from 2009-2015 to investigate whether having clinical experience on a hospital executive team affects hospital behavior.
+In particular, I leverage several pay-for-performance initiatives in the US in 2012, and estimate the difference in response between for-profits, nonprofits, and clinical nonprofits.
+I find that while readmissions decreased on average for all types of hospitals, non-clinical executive team hospital decreased readmissions at a faster rate than those with a clinical executive team, and behaved more similarly to for-profit hospitals.
 The difference equates to approximately 15 additional readmissions for clinical team hospitals.
-I find no differential effect of clinical executive teams on mortality rates.
 
@@
   The objectives and behaviors of firms that are not classic for-profits are not easily determined, and an extensive literature has sought to understand these types of firms.
-Nonprofits are thought to gain utility not solely from monetary profit, but also through accomplishing mission-based purposes.
-Researchers have speculated a variety of objective functions that could motivate nonprofits such as maximizing prestige, income, or a quality/quantity tradeoff (Steinberg, 1986).
+Nonprofits are thought to gain utility not solely from monetary profit, but also through accomplishing mission-based goals.
+Researchers have speculated a variety of objective functions that could motivate nonprofits such as maximizing prestige, income, or a quality/quantity trade-off (\cite{steinberg1986revealed}).
 An important contribution to this literature is the use of empirical strategies to analyze how firm behavior reveals attributes of the underlying objective function (\cite{sloan2000not}).
-In this paper, I leverage a large-scale shock to nonprofit hospitals in the US to investigate whether characteristics of the objective function are revealed by the composition of their leadership team.
+In this paper, I leverage a large-scale shock to hospitals in the US to investigate whether differences in leadership composition reveal characteristics of hospital objective functions, namely the weight placed on profit vs.
+utility gained from benefiting society.
 
-  Hospitals in particular have been a common focus when seeking to understand not-for-profit firms.
+  Hospitals in particular have been a common focus when seeking to understand nonprofit firms.
 This is because, first, there are different hospital ownership types within the industry, giving researchers the ability to compare the behaviors of nonprofits with that of for-profits.
@@
 However, it remains unclear what type of objective function drives nonprofit hospitals (\cite{erus2002inferring}; \cite{deneffe2002not}; \cite{horwitz2009hospital}).
+In particular, the extent to which nonprofit hospitals gain utility from bettering society (e.g.
+increased charity care, high quality of care, etc.) is unknown.
 
   What has not been considered is how the leadership team within nonprofits potentially drives objectives.
-In large, publicly traded firms, executive characteristics are shown to be correlated with firm performance, indicating that there is more to firm behavior than just for-profit or not-for-profit status.
-Within nonprofit hospitals, there is variation in the propensity to hire executives with a background in clinical work, which could reasonably change how much weight the hospital places on profit vs.
-quality of care in their objective function.
+In large, publicly traded firms, executive characteristics are shown to be correlated with firm performance, indicating that there is more to firm behavior than for-profit or not-for-profit status (\cite{bertrand2003managing}; \cite{matsa2013female}; \cite{ahern2012changing}).
+Within nonprofit hospitals, there is variation in the propensity to hire executives with a clinical background, which could reasonably change how much weight the hospital places on profit vs.
+quality of care when making decisions.
 I contribute to our understanding of hospital and nonprofit behaviors by investigating whether clinical experience on an executive team affects hospital behavior after being financially penalized.
 
-  Using publicly available tax forms, I construct a novel data set of nonprofit hospitals from 2009-2015 which captures the identities of executives tied to that hospital.
+  Using publicly available tax forms, I construct a novel data set of nonprofit hospital executives from 2009-2015.
 I link hospitals in this data to characteristics in the American Hospital Association survey and Hospital Compare data.
-In order to capture the effect of clinical executives, not just their selection into certain types of hospitals, I specifically investigate hospitals penalized by the Hospital Readmissions Reduction Program (HRRP) of 2012.
-This program reduced Medicare payments to hospitals with above average readmission rates, and is plausibly uncorrelated with previously determined leadership teams.
-I further limit the sample to hospitals with stable executive teams over time, and characterize hospitals as having clinical experience on their team or not.
+In order to capture the effect of clinical executives, not just their selection into certain types of hospitals, I consider the sample of hospitals whose leadership team remains stable over time, and characterize them as having clinical executive experience on their team or not, a time-invariant categorization.
+I then leverage the Hospital Readmissions Reduction Program (HRRP) of 2012.
+This program reduced Medicare payments to hospitals with above average readmission rates, and is a shock that is plausibly uncorrelated with previously determined leadership teams.
 
-  I estimate the difference in readmission and mortality rates between hospitals with and without clinical experience before and after the penalties began, conditional on being penalized.
+  I estimate the difference in readmission and mortality rates between hospitals with and without clinical experience before and after the penalties began.
 While readmissions directly affects the likelihood of being penalized, mortality rates are an explicit quality measure unrelated to the penalties.
-Including this as an outcome can help distinguish whether hospitals are improving quality or taking measures to only reduce readmissions.
+Including this as an outcome can help distinguish whether hospitals are improving quality or only taking measures to reduce readmissions.
+Not all hospitals were penalized, but all hospitals (with few exceptions) could have been penalized if readmission rates were high enough.
+Thus, I present estimates first for all hospitals.
+However, these estimates are biased if I am simply capturing the effect of clinical experience on already being a good quality hospital.
+Thus, I present two further specifications: one limiting to only hospitals who were eventually penalized, and another where I limit the sample further to hospitals which have an observably similar match.
 
-  I assume that before the penalty, clinical and non-clinical team hospitals behaved similarly along quality dimensions, and that the composition of hospital executive teams determined prior to the penalties is, conditional on being penalized, exogenous.
-Under these assumptions, I identify the effect of having a clinical executive on response to the penalty.
+  I assume that before the penalty, clinical and non-clinical team hospitals behaved similarly along quality dimensions, and that the composition of hospital executive teams determined prior to the penalties is exogenous.
+Under these assumptions, I identify the effect of having a clinical executive on hospital response to the penalty.
 I find that while both types of hospitals, those with clinical executives and those without clinical executives, decrease readmissions after being penalized, non-clinical team hospitals decrease readmissions at a faster rate.
 The difference equates to approximately 15 additional readmission cases per year for clinical teams, and there is no evidence that this result is driven by differences in readmissions prior to penalties.
-I find no differential effect of the program on mortality rates.
+However, I find no differential effect of the program on mortality rates.
 
@@
 To this point, researchers have found that nonprofits do not act as purely profit maximizers, but maximize some combination of profit and output (\cite{deneffe2002not}, \cite{chang2011nonprofit}).
-A paper with a similar methodology to investigate hospital behavior is \citeauthor{chang2011nonprofit} (\citeyear{chang2011nonprofit}), which looks at the effect of a cost shock on likely of shutting down and mix of profitable services.
+A paper with an approach similar to mine which investigates hospital behavior is \citeauthor{chang2011nonprofit} (\citeyear{chang2011nonprofit}), which looks at the effect of a cost shock on likely of shutting down and mix of profitable services.
 They find that nonprofits and for-profits are equally as likely to shut down after the shock, but only nonprofits adjust their mix of profitable services.
@@
 From AHA, I gather the number of beds in the hospital and the Medicare number, which I use to merge the Hospital Compare data.
-Hospital Compare contains multiple data sets which I use first to limit to hospitals that were penalized in the Hospital Readmissions Reduction Program (HRRP), and also for information on my outcome variables, yearly readmission rates for the relevant HRRP conditions: heart failure, heart attack (AMI), and pneumonia.
+Hospital Compare contains multiple data sets which I use first to limit to hospitals that were penalized in the Hospital Readmissions Reduction Program (HRRP), and also for information on outcome variables: readmission and mortality rates for the relevant HRRP conditions: heart failure, heart attack (AMI), and pneumonia.
 
@@
 
-    \import{Tables}{noMDchg_sumstats.tex}
+    \subsection{Summary Statistics}
 
-    In Table \ref{tab:sumstats}, I present summary statistics for various hospital characteristics including which condition hospitals are penalized for, and the readmission and mortality outcome variables.
+    \import{Tables}{overall_sumstats.tex}
+
+    \import{Tables}{sample_sumstats.tex}
+
+    In Table \ref{tab1:sumstats}, I present summary statistics for various hospital characteristics including which condition hospitals are penalized for, and the readmission and mortality outcome variables.
 After limiting to penalized hospitals with stable leadership teams, my sample consists of between 302-310 hospitals from 2009-2015, except in the case of AMI outcomes, which is only 225-264 hospitals due to small patient numbers.
@@
 
-    Figure \ref{fig:weighted_read_mort_graph} shows average readmission rates for hospitals in the sample across time and across the three different conditions plus the weighted average.
+    Figure \ref{fig:weighted_read_mort_graph} shows average readmission and mortality rates across time and across the three different conditions, along with the weighted average.
 There are multiple notable points the raw data reveals.
@@
 While suggestive, this motivates further analysis into the differential responses between clinical and non-clinical leadership teams.
+On the right panel, average mortality rates over time reveal no clear change due to the penalty, and no consistent pattern in the difference between clinical and non-clinical hospital teams.
 
     \begin{figure}[ht!]
-        \caption{Readmission rates over time in all penalized hospitals}
-        \includegraphics[scale=.55]{Objects/weighted_read_graph.pdf}
+        \caption{Outcomes over time in all penalized hospitals}
+        \includegraphics[scale=.5]{Objects/weighted_read_mort_graph.pdf}
         \label{fig:weighted_read_mort_graph}
     \end{figure}
+
+    \import{Tables}{hosp_correlations.tex}
 
@@
 
-    $$\max\hspace{2mm} \delta R + (1-\delta) u(\theta),$$
+    $$\max_{\theta}\hspace{2mm} \delta R + (1-\delta) u(\theta),$$
 
@@
 
-    $$\max\hspace{2mm} \delta(L) R(\theta) + (1-\delta(L)) u(\theta)$$
+    $$\max_{\theta}\hspace{2mm} \delta(L) R(\theta) + (1-\delta(L)) u(\theta)$$
 
@@
 Therefore, when maximizing this function, the choice of $\theta$ is dependent on $L$ through its impact on how the hospital values profit vs.
-quality of care.
+quality of care:
+
+    $$\theta = f(\delta(L)).$$
 
     Thus, I directly estimate the effect of having clinical experience on readmission rates.
+Additionally, I estimate the effect on mortality rates to capture any spillover effects of improving readmissions into another dimension of quality.
 However, looking purely at this correlation is a biased picture of the effect of clinical experience since it likely captures reverse causality: certain types of hospitals may be more likely to hire doctors as executives.
@@
 First, as discussed in detail in Section \ref{sec:data}, I only consider hospitals with stable leadership teams over time.
-Second, I leverage the HRRP shock, limiting the sample to those that were penalized, which allows me to identify hospitals similar in quality along readmission rates, and to observe changes in behavior after the shock.
+Second, I leverage the HRRP shock, limiting the sample to those that were penalized, which allows me to identify hospitals similar in quality, and to observe changes in behavior after the shock.
 For the hospitals in my sample, I define their type as having clinical experience on their executive team or not, a time invariant categorization.
@@
 As discussed in Section \ref{sec:hrrp}, hospitals had incentive to decrease readmissions immediately when the rules were released in October of 2011, before the penalties were actually determined in October of 2012.
-However, I still define $post\_penalty$ as occurring in 2012 or later, the first full year when hospitals would want to respond.
+However, I still define $post\_penalty$ as occurring in 2012 or later, the first full year when hospitals could respond.
 
-    For identification, I rely on the assumption that, prior to being penalized, clinical and non-clinical team hospitals did not differentially change readmission rates for a reason I do not capture here.
+    For identification, I rely on the assumption that, prior to being penalized, clinical and non-clinical team hospitals did not differentially change readmission or mortality rates for a reason I do not capture here.
 I investigate this assumption by estimating event study coefficients, where the coefficients $\beta_j$ now capture the effect of having clinical experience in a particular year, and $\beta_{2011}$ is normalized to zero:
@@
 
-    \subsection{Readmission Rates}
+    \import{Tables}{MD_noMD_readmort_synthdid.tex}
+
+    \import{Tables}{forprofit_synth.tex}
+
+    \import{Tables}{MD_noMD_uncompCMI_synth.tex}
+
+    \import{Tables}{forprofit_uncompCMI_synth.tex}
 
@@
 
-    \import{Tables}{read_results_table.tex}
-
     In columns 2, 3, and 4, I limit the sample to hospitals penalized specifically for the condition that matches the outcome in the model.
@@
 
+    \begin{figure}[ht!]
+        \caption{Event Study Estimates, Readmission}
+        \includegraphics[scale=.5]{Objects/read_es_graph.pdf}
+        \label{fig:wa_eventstudy}
+    \end{figure}
+
     Finally, I plot estimated coefficients from Equation \ref{eq:es} in Figure \ref{fig:wa_eventstudy}.
@@
 
+    Estimates for the coefficient of interest from Equation \ref{eq:main} with mortality rates are presented in Table \ref{tab:mort_results}.
+Column 1 presents my estimate of the effect of having a clinical experience team on the weighted average of mortality.
+Columns 2, 3, and 4 present estimates where the sample is limited to hospitals who are penalized for the outcome being considered.
+The coefficients are all positive, ranging from .13 to .31, but none of them are statistically significant.
+This is consistent with prior literature that finds hospitals do not change mortality rates after HRRP, and this result reveals that holds across different leadership teams.
+
+    I investigate whether mortality rates seem to be driven by a pre-trend, or whether the effect of clinical experience on mortality rates differs across time after the penalty occurs.
+Event study estimates from Equation \ref{eq:es} for the weighted average mortality model is shown in Figure \ref{fig:wa_eventstudy_mort}.
+These estimates reveal that hospital teams with clinical experience do not differentially change mortality rates compared to non-clinical experience teams before or after the penalties begin.
+
     \begin{figure}[ht!]
-        \caption{Event Study Estimates}
-        \includegraphics[scale=.5]{Objects/read_es_graph.pdf}
-        \label{fig:wa_eventstudy}
+        \caption{Event Study Estimates, Mortality}
+        \includegraphics[scale=.5]{Objects/mort_es_graph.pdf}
+        \label{fig:wa_eventstudy_mort}
     \end{figure}
-
-    \subsection{Mortality Rates}
-
-    \import{Tables}{mort_results_table.tex}
 
@@
 For these similar hospitals, I estimate the difference in readmissions after being penalized between hospitals with clinical executives and hospitals without clinical executives.
-I find that while the penalty program led both types of hospitals to decrease readmissions after being penalized, those without clinical experience on their executive decreased rates more quickly than non-clinical team hospitals.
+I find that while the penalty program led both types of hospitals to decrease readmissions after being penalized, those without clinical experience on their executive team decreased rates more quickly than non-clinical team hospitals.
 
@@
 quality of care.
-Therefore, not-for-profit vs.
-for-profit status should not be the only characteristic considered by policymakers when designing incentives to increase quality of care.
+Therefore, policymakers should take into consideration the composition of the leadership team when designing pay-for-performance incentives.
 
@@
 
-	% \appendix
+\appendix
 
- %    \section{Data}\label{appendixdata}
+ \section{Data}\label{appendixdata}
 
- %    \subsection{Gathering Hospital Leadership Names}
+\subsection{Gathering Hospital Leadership Names}
 
- %    There is no perfect way to access tax form 990s in bulk over the time period I am considering in this paper.
+There is no perfect way to access tax form 990s in bulk over the time period I am considering in this paper.
 However, using the NonProfit Explorer API seems to be the most straightforward.
@@
 
- %    There are over 1.5 million nonprofit entities in the US (cite), making it crucially important to be able to filter by type of entity before analyzing any PDFs.
+There are over 1.5 million nonprofit entities in the US (cite), making it crucially important to be able to filter by type of entity before analyzing any PDFs.
 The API allows this by filtering a query based on National Taxonomy of Exempt Entities (NTEE) code.
@@
 
- %    I loop through the list of EINs found in the previous step and query more detailed information from the API on that specific EIN.
+I loop through the list of EINs found in the previous step and query more detailed information from the API on that specific EIN.
 I save the name, secondary name, state, and zip code, all of which do not vary by year.
@@
 
- %    It is crucial to the analysis to be able to link these nonprofits with other sources of data to recover penalties from HRRP, bed size, and outcomes of interest (definitely need to talk more about outcomes).
+It is crucial to the analysis to be able to link these nonprofits with other sources of data to recover penalties from HRRP, bed size, and outcomes of interest (definitely need to talk more about outcomes).
 Therefore, I take a conservative approach to matching EINs to American Hospital Association (AHA) ID, which links fairly easily to Medicare ID numbers, based on name and location.
@@
 
- %    I download the AHA data from Wharton Research Data Services.
+I download the AHA data from Wharton Research Data Services.
 I filter only to hospitals in the contiguous US, Alaska, and Hawaii (excluding places like Puerto Rico), classified as nonprofit or state/community, and those that are general acute care.
@@
 
- %    I consider which nonprofit entities are not likely to be hospitals and drop them.
+I consider which nonprofit entities are not likely to be hospitals and drop them.
 There are numerous foundations or auxiliary nonprofits with the purpose of raising funds for the hospital, but do not actually care for patient.
@@
 
- %    I then proceed matching based on names in multiple layers.
+I then proceed matching based on names in multiple layers.
 I focus on exact string matches, so I remove all spaces and common characters that could cause mismatches such as \&, ', -, and inc.
@@
 
- %    In the next layer of matching, I remove common words such as "healthcare", "regional", "hospital", etc.
+In the next layer of matching, I remove common words such as "healthcare", "regional", "hospital", etc.
 That way if there are subtle differences in names, removing common words may allow for an exact match.
@@
 
- %    In some cases, tax forms are associated with a system of hospitals instead of one individual hospital.
+In some cases, tax forms are associated with a system of hospitals instead of one individual hospital.
 Thus, I create another variable that captures the names of systems which match.
@@
 
- %    (add a paragraph about validity of matches)
+(add a paragraph about validity of matches)
 
- %    Now that I have a sample of hospitals, the next step is to extract the names of the people in charge of these hospitals from the Tax Form 990 PDFs.
+ Now that I have a sample of hospitals, the next step is to extract the names of the people in charge of these hospitals from the Tax Form 990 PDFs.
 In the data set of hospital PDF URLs that I collected earlier, I limit to the hospitals with solid matches described above.
@@
 
- %    One tricky aspect of the NonProfit Explorer API is that, only in some cases, if two forms are present for an EIN, year, only the first one (which is typically not the one with the relevant information) is pulled.
+One tricky aspect of the NonProfit Explorer API is that, only in some cases, if two forms are present for an EIN, year, only the first one (which is typically not the one with the relevant information) is pulled.
 Therefore, for some hospitals, a couple years will have gaps in text extraction data.
@@
 
- %    The form of the text data is a data frame with one column, where each line of text is saved in a different row.
+The form of the text data is a data frame with one column, where each line of text is saved in a different row.
 I write a text cleaning package that locates names, positions, titles, and indications of resigning.
@@
 
- %    Typically on the same page as the names and positions is a list of the highest compensated employees and their compensation.
+Typically on the same page as the names and positions is a list of the highest compensated employees and their compensation.
 In order to not record extra names, I filter out any rows after the start of this section.
@@
 
- %    Next, I locate first and last names in the data.
+Next, I locate first and last names in the data.
+
+\section{Full Sample Regression Results}
+
+\import{Tables}{MD_noMD_readmort_fullsample.tex}
+
+\import{Tables}{forprofit_fullsample.tex}
+
+\import{Tables}{MD_noMD_uncompCMI_fullsample.tex}
+
+\import{Tables}{forprofit_uncompCMI_fullsample.tex}
+
+\section{Matched Sample Regression Results}
+
+\import{Tables}{MD_noMD_readmort_matchsample.tex}
+
+\import{Tables}{forprofit_matchsample.tex}
+
+\import{Tables}{MD_noMD_uncompCMI_matchsample.tex}
+
+\import{Tables}{forprofit_uncompCMI_matchsample.tex}
 

### 2024-04-23 Slides/readinggroupMarch2024.tex
Updates from Overleaf

--- 2024-01-26 Slides/data_description.tex
@@
 \usepackage{array}
+
+\usepackage[backend=biber,style=authoryear,
+sorting=ynt,citestyle=authoryear]{biblatex}
+\addbibresource{Paper/papercitations.bib}
 
@@
 \usepackage{subfigure}
+\usepackage{import}
 
@@
 \setbeamercolor{enumerate subitem}{fg=sage}
-\setbeamercolor{button}{bg=MyBackground,fg=purple,}
+\setbeamercolor{button}{bg=sage,fg=white}
 
@@
 
-\title[]{\textcolor{sage}{Description of Hospital Leadership Data}}
+\title[]{\textcolor{sage}{Does Hospital Leadership Matter? Evidence from the Hospital Readmissions Reduction Program}}
+\subtitle[]{A JMP update}
 
@@
 
-\begin{frame}{end goal}
-    \textcolor{purple}{the purpose of collecting this data is to get a sense of who is in charge of running nonprofit hospitals in the US}
+\begin{frame}{Motivation: Not-for-profit Firms}
+    What do not-for-profit objective functions look like?
+    \begin{itemize}
+        \item Maximizing prestige, income, quality/quantity tradeoff?  (\cite{steinberg1986revealed})
+    \end{itemize}
+
+    \vspace{3mm}
+
+    Important aspect of this literature: using empirical strategies to analyze how firm behavior reveals attributes of objective function (\cite{sloan2000not})
+
+    \vspace{10mm} \pause
+
+    \begin{block}{This Paper}
+    I investigate whether executive leadership team composition reveals anything about nonprofit hospital objective functions/behaviors
+    \end{block}
+\end{frame}
+
+\begin{frame}{Motivation: Nonprofit Hospitals}
+    Relevant 
+    \begin{itemize}
+        \item 50\% of all US hospitals
+        \item Staff 207 beds on average
+    \end{itemize}
+
+    \vspace{5mm}
+
+    Unclear what objective function looks like 
+    \begin{itemize}
+        \item Mixed evidence whether costs, uncompensated care, technology adoption, quality differ from for-profit (\cite{sloan2000not}; \cite{eggleston2008hospital}; \cite{moscelli2018effect}; \cite{moscone2020public})
+        \item Maximize some combination of profit and output (\cite{deneffe2002not})
+        \item Nonprofit hospitals more likely to adjust patient mix after financial shock (\cite{chang2011nonprofit})
+    \end{itemize}
+\end{frame}
+
+\begin{frame}{Executive Leadership Teams}
+    What hasn't been considered is how characteristics even within nonprofits affect individual hospital objectives
+    \begin{wideitemize}
+        \item The people actually running the hospital are plausibly important
+        \item Characteristic important to hospitals is clinical experience 
+    \end{wideitemize}
+\end{frame}
+
+\begin{frame}{This Paper}
+    \begin{block}{}
+        I leverage a large-scale policy affecting hospitals to investigate whether clinical experience on a hospital executive team affects hospital response to potential penalties
+    \end{block}
+\end{frame}
+
+\begin{frame}{About the Policy}
+    Hospital Readmissions Reduction Program: passed 2010, took effect 2012
+
+    \vspace{5mm}
+
+    \begin{wideitemize}
+        \item Goal: get hospitals to lower readmissions
+        \item Hospitals w/ above average readmission rates during look-back period penalized with lower Medicare reimbursements
+        \item Conditions: AMI, pneumonia, heart failure
+        \item \textbf{All} hospitals had incentive to lower readmissions
+    \end{wideitemize}
+
+    \vspace{5mm}\pause
+
+    I mainly use the timing of the policy as variation, but get data on hospital penalties for heterogeneity analysis from Hospital Compare
+\end{frame}
+
+\begin{frame}{About the Clinical Experience Variable}\label{clinexp}
+    I scrape publicly available Tax Form 990s to get names and titles of all executives for each hospital from 2009-2015 \hyperlink{execnames}{\beamerbutton{text extraction}}
 
@@
 
-    why?
+    Two things I need from this:
     \begin{wideitemize}
-        \item we don't know a lot about who runs hospitals
-        \item we know even less about whether characteristics of these people make a difference 
-        \item for this project, I care specifically about the characteristic of clinical experience in executive positions
+        \item Characterize stability of leadership teams over time (limit sample)
+        \item Characterize whether any executives are MDs 
+        \begin{itemize}
+            \item This ends up being a time-invariant hospital categorization due to the above sample limitation
+        \end{itemize}
     \end{wideitemize}
+
+    \vspace{5mm}
+
+    sample: 580 hospitals, 13\% with clinical experience
 \end{frame}
 
-\section{Tax Form 990}
+\begin{frame}{About the Outcomes}
+    Readmission rates:
+    \begin{itemize}
+        \item Targeted outcome of the policy, direct incentives for hospitals to lower 
+        \item I create average readmission rate across conditions, weighted by number of patients with each condition
+    \end{itemize}
 
-\subsection{who files?}
+    \vspace{5mm}
 
-\begin{frame}{who files?}
-\underline{important acronym:} EIN (Employer Identification Number)
+    Mortality rates:
+    \begin{itemize}
+        \item Not the targeted outcome, has been used to investigate whether hospitals are actually improving quality or just adjusting readmissions to avoid penalties
+        \item I create average mortality rate across conditions, weighted by number of patients with each condition
+    \end{itemize}
 
-    \centering
-    \includegraphics[scale=.7]{Graphics/990_snip_frontpage.PNG}
+    \vspace{10mm}
+
+    I get hospital level readmission and mortality rates from the Hospital Compare Data
 \end{frame}
 
-\begin{frame}{who files?}
+\begin{frame}{Summary Statistics}\label{sumstatsgraph}
+    \begin{figure}
         \centering
-    \includegraphics[scale=.7]{Graphics/990_snip_officerspage.PNG}
+        \includegraphics[width=\linewidth]{Objects/weighted_read_mort_graph.pdf}
+    \end{figure}
+
+    \hyperlink{sumstats}{\beamerbutton{stats}}
 \end{frame}
 
-\subsection{accessing hospital forms}
+\begin{frame}{Estimation Strategy}
 
-\begin{frame}{accessing hospital forms}\label{accessing hospital forms}
+    Compare hospitals with and without clinical experience before and after penalties began:
+
+    \vspace{3mm}
+
+    \begin{equation*}
+    y_{ht} = \alpha_{h} + \delta_t + \beta (clinical_h \times post\_penalty_{ht}) + \epsilon_{ht}
+    \end{equation*}
+
+    \vspace{10mm}\pause
+
+    Identifying Assumptions:
+    \begin{itemize}
+        \item Before the penalty, clinical and non-clinical team hospitals trended similarly along quality dimensions
+        \item Previously determined leadership team is exogenous to the financial shock
+    \end{itemize}
+
+    \vspace{5mm}
+
+    Threat to identification would have to be a time-varying hospital characteristic correlated with both clinical experience and outcomes
+\end{frame}
+
+\begin{frame}{All Hospital Response to Policy}\footnotesize
+    \import{Tables}{policy_read_table.tex}
+    \import{Tables}{policy_mort_table.tex}
+\end{frame}
+
+\begin{frame}{Differential Response in Readmissions}
+   \small \import{Tables}{read_results_table.tex}
+\end{frame}
+
+\begin{frame}{Differential Response in Mortality}
+    \small \import{Tables}{mort_results_table.tex}
+\end{frame}
+
+\begin{frame}{Conclusion}
 \begin{wideitemize}
-        \item completed forms are publicly available on ProPublica's website
-        \item through their API, I download a list EINs classified as a hospital
-        \begin{itemize}
-            \item based on National Taxonomy of Exempt Entities (NTEE) codes: E20, E21, E22
-        \end{itemize}
-        \item along with EIN, I gather their name, zip, state, and urls to tax form pdfs
-        \item years 2009-2015
+    \item While non-clinical teams have higher readmission rates before the penalty, they decrease readmissions at a faster rate after the penalties, though the effect is small
+    \item No difference in mortality response
 \end{wideitemize}
@@
 
-    \begin{block}{data point}
-        3,925 EINs \hspace{10pt} \hyperlink{why more?}{\beamergotobutton{why more?}}
-    \end{block}
+Now the question is why? 
+\begin{itemize}
+    \item Are non-clinical teams adjusting patient mix? 
+\end{itemize}
 \end{frame}
 
-\subsection{matching EIN to AHA hospitals}
-
-\begin{frame}{matching EIN to AHA hospitals}
-    I need to be able to link hospitals to other public data sets that capture hospital information
-    \vspace{4mm}
-    \begin{wideitemize}
-        \item I choose to match to the widely used American Hospital Association survey data
-        \item only match-able link I have is hospital name
-        \item I restrict matches to being in the same state
-    \end{wideitemize}
-\end{frame}
-
-\begin{frame}{matching EIN to AHA hospitals}
-\centering
-   \begin{tikzpicture}[node distance=6mm, every node/.style={align=center, font=\large}]
-
-        % Left box
-        \node[draw=diagrampurple, line width=1.5pt, minimum width=4cm, minimum height=4cm, text=black] (ahaBox) at (0,0) {%
-            \underline{\textbf{\textcolor{black}{AHA Hospitals}}} \\[4pt]
-            \begin{tabular}{@{}c@{}}
-                {Hosp A} \\
-                {Hospital B} \\
-                {C Hospital} \\
-                {Hosp. of D} \\
-                {E} \\
-            \end{tabular}
-        };
-
-        % Right box
-        \node[draw=diagramred, line width=1.5pt, minimum width=4cm, minimum height=4cm, text=black, right=of ahaBox] (einBox) {%
-            \underline{\textbf{\textcolor{black}{EIN Hospitals}}} \\[4pt]
-            \begin{tabular}{@{}c@{}}
-                {Hospital B} \\
-                {Hospital D} \\
-                {H} \\
-                {C Hospital} \\
-                {Hospital G} \\
-            \end{tabular}
-        };
-
-    \end{tikzpicture}
-\end{frame}
-
-\begin{frame}{matching EIN to AHA hospitals}
-\centering
-   \begin{tikzpicture}[node distance=6mm, every node/.style={align=center, font=\large}]
-
-        % Left box
-        \node[draw=diagrampurple, line width=1.5pt, minimum width=4cm, minimum height=4cm, text=black] (ahaBox) at (0,0) {%
-            \underline{\textbf{\textcolor{black}{AHA Hospitals}}} \\[4pt]
-            \begin{tabular}{@{}c@{}}
-                \textcolor{diagrampurple}{Hosp A} \\
-                {Hospital B} \\
-                {C Hospital} \\
-                {Hosp. of D} \\
-                \textcolor{diagrampurple}{E} \\
-            \end{tabular}
-        };
-
-        % Right box
-        \node[draw=diagramred, line width=1.5pt, minimum width=4cm, minimum height=4cm, text=black, right=of ahaBox] (einBox) {%
-            \underline{\textbf{\textcolor{black}{EIN Hospitals}}} \\[4pt]
-            \begin{tabular}{@{}c@{}}
-                {Hospital B} \\
-                {Hospital D} \\
-                \textcolor{diagramred}{H} \\
-                {C Hospital} \\
-                \textcolor{diagramred}{Hospital G} \\
-            \end{tabular}
-        };
-
-    \end{tikzpicture}
-\end{frame}
-
-\begin{frame}{matching EIN to AHA hospitals}
-\centering
-   \begin{tikzpicture}[node distance=6mm, every node/.style={align=center, font=\large}]
-
-        % Left box
-        \node[draw=diagrampurple, line width=1.5pt, minimum width=4cm, minimum height=4cm, text=black] (ahaBox) at (0,0) {%
-            \underline{\textbf{\textcolor{black}{AHA Hospitals}}} \\[4pt]
-            \begin{tabular}{@{}c@{}}
-                \textcolor{diagrampurple}{Hosp A} \\
-                \textcolor{diagramtan}{Hospital B} \\
-                \textcolor{diagramtan}{C Hospital} \\
-                {Hosp. of D} \\
-                \textcolor{diagrampurple}{E} \\
-            \end{tabular}
-        };
-
-        % Right box
-        \node[draw=diagramred, line width=1.5pt, minimum width=4cm, minimum height=4cm, text=black, right=of ahaBox] (einBox) {%
-            \underline{\textbf{\textcolor{black}{EIN Hospitals}}} \\[4pt]
-            \begin{tabular}{@{}c@{}}
-                \textcolor{diagramtan}{Hospital B} \\
-                {Hospital D} \\
-                \textcolor{diagramred}{H} \\
-                \textcolor{diagramtan}{C Hospital} \\
-                \textcolor{diagramred}{Hospital G} \\
-            \end{tabular}
-        };
-
-    \end{tikzpicture}
-
-\vspace{7mm}
-
-    \begin{block}{data point}
-        I match 1,041 EINs to AHA hospitals using perfect name matching
-    \end{block}
-\end{frame}
-
-\begin{frame}{matching EIN to AHA hospitals}
-\centering
-   \begin{tikzpicture}[node distance=6mm, every node/.style={align=center, font=\large}]
-
-        % Left box
-        \node[draw=diagrampurple, line width=1.5pt, minimum width=4cm, minimum height=4cm, text=black] (ahaBox) at (0,0) {%
-            \underline{\textbf{\textcolor{black}{AHA Hospitals}}} \\[4pt]
-            \begin{tabular}{@{}c@{}}
-                \textcolor{diagrampurple}{Hosp A} \\
-                \textcolor{diagramtan}{Hospital B} \\
-                \textcolor{diagramtan}{C Hospital} \\
-                \textcolor{diagramteal}{Hosp. of D} \\
-                \textcolor{diagrampurple}{E} \\
-            \end{tabular}
-        };
-
-        % Right box
-        \node[draw=diagramred, line width=1.5pt, minimum width=4cm, minimum height=4cm, text=black, right=of ahaBox] (einBox) {%
-            \underline{\textbf{\textcolor{black}{EIN Hospitals}}} \\[4pt]
-            \begin{tabular}{@{}c@{}}
-                \textcolor{diagramtan}{Hospital B} \\
-                \textcolor{diagramteal}{Hospital D} \\
-                \textcolor{diagramred}{H} \\
-                \textcolor{diagramtan}{C Hospital} \\
-                \textcolor{diagramred}{Hospital G} \\
-            \end{tabular}
-        };
-
-    \end{tikzpicture}
-
-    \vspace{7mm}
-
-    \begin{block}{data point}
-        manually looking at names, I match 196 more EINs to AHA hospitals
-    \end{block}
-\end{frame}
-
-\subsection{automated vs. manual collection}
-
-\begin{frame}{automated vs. manual collection}
-    I now have a list of EINs I want the pdfs for
-
-    \vspace{4mm}
-
-    \begin{wideitemize}
-        \item I automate the download of the pdf URLs for these EINs that I collected from the API
-        \item in many cases, the API misses the form even though it exists online
-        \item I manually download the forms for which this error exists
-    \end{wideitemize}
-\end{frame}
-
-\section{Text Extraction and Cleaning}
-
-\subsection{optical character recognition}
-
-\begin{frame}{optical character recognition}
-    older pdfs are basically images, which requires using OCR for text recognition
-
-    \vspace{4mm}
-
-    \begin{wideitemize}
-        \item the software takes pages and returns a string of text found
-    \end{wideitemize}
-\end{frame}
-
-\begin{frame}{optical character recognition}
+\begin{frame}[noframenumbering]{}\label{execnames}
+\hyperlink{clinexp}{\beamerbutton{back}}
 \begin{tikzpicture}
@@
 \node (a) [fill=purple!50,draw] at (bad.center) {original pdf};
-\node (b) [fill=purple!50,draw] at (good.center) {output from OCR};
+\node (b) [fill=purple!50,draw] at (good.center) {output from text extraction};
 
@@
 
-\subsection{cleaning raw text output}
+\begin{frame}[noframenumbering]{}\label{sumstats}
+    \scriptsize \import{Tables}{sumstats_nooutcomes.tex}
 
-\begin{frame}{cleaning raw text output}
-algorithmically locate names and associated positions
-
-\vspace{4mm}
-
-\begin{enumerate}
-    \item remove numbers and special characters
-    \item record full names where at least one (first or last) is found in census most common names data
-    \item note when the name is followed by ``MD", ``DO", ``RN", etc. 
-    \item locate the position nearest the name found and record it
-\end{enumerate}
-\end{frame}
-
-\begin{frame}{cleaning raw text output}
-    final names data:
-    \vspace{4mm}
-    \begin{wideitemize}
-        \item each row is a name, hospital, year observation
-        \item other variables: position, clinical experience
-    \end{wideitemize}
-\end{frame}
-
-\section{Clinical Experience Characteristic}
-
-\subsection{link names to NPPES}
-
-\begin{frame}{link names to NPPES}
-    next goal: verify and get more information on doctors in the data
-    \vspace{4mm}
-    \begin{wideitemize}
-        \item use names to link to NPPES to get NPI number and specialty
-        \item some cases are easy: MD in tax data and one name match to NPPES
-        \item most are not that easy and require manual sleuthing
-    \end{wideitemize}
-\end{frame}
-
-\begin{frame}{why more EINs than nonprofit hospitals?}\label{why more?}
-    \begin{wideitemize}
-        \item various sources quote the number of nonprofit hospitals to be less than 3,000
-        \begin{wideitemize}
-            \item (add citations)
-        \end{wideitemize}
-        \item without any limitations to hospital EINs at this point, my 3,925 EINs include hospital associations, fundraisers, and some large clinics that will later be excluded.
-    \end{wideitemize}
-
-    \vspace{10mm}
-
-    \hyperlink{accessing hospital forms}{\beamerreturnbutton{go back}}
+    \hyperlink{sumstatsgraph}{\beamerbutton{back}}
 \end{frame}

### 2024-04-23 Tables/MD_noMD_readmort_fullsample.tex
Updates from Overleaf

\begin{table}[htbp]
   \caption{\label{tab:MD_noMD_fullsample} Comparison of MD vs. non-MD Nonprofit Readmissions and Mortality Regression Results}
   \bigskip
   \centering
   \begin{tabular}{lcc}
      \toprule
                                        & Weighted Avg. Readmission Rate & Weighted Avg. Mortality Rate\\  
                                        & (1)                            & (2)\\  
      \midrule 
      NP w/ MD Executive x Post Penalty & -0.40$^{***}$                  & -0.19\\   
                                        & (0.14)                         & (0.15)\\   
      \textbf{Comparison Group:}        &                                & \\  
      NP w/out MD Executive             & $\checkmark$                   & $\checkmark$\\   
       \\
      Observations                      & 2,471                          & 2,462\\  
      \bottomrule
   \end{tabular}
\end{table}

### 2024-04-23 Tables/MD_noMD_readmort_synthdid.tex
Updates from Overleaf

\begin{table}[ht!]

\caption{Comparison of MD vs. non-MD Nonprofit Readmissions and Mortality Synthetic DiD Results}
\centering
\begin{tabular}[t]{lcc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{1}{c}{Weighted Avg. Readmission Rate} & \multicolumn{1}{c}{Weighted Avg. Mortality Rate} \\
 & (1) & (2)\\
\midrule
NP w/ MD Executive x Post Penalty & -0.28$^{*}$ & -0.19\\
 & (0.12) & (0.11)\\
\textbf{Comparison Group:} &  & \\
NP w/out MD Executive & $\checkmark$ & $\checkmark$\\
\\
Observations & 2,335 & 2,325\\
\bottomrule
\end{tabular}
\end{table}

### 2024-04-23 Tables/MD_noMD_uncompCMI_fullsample.tex
Updates from Overleaf

\begin{table}[htbp]
   \caption{\label{tab:MD_noMD_uncomp_CMI_fullsample} Comparison of MD vs. non-MD Nonprofit Uncompensated Care and CMI Regression Results}
   \bigskip
   \centering
   \begin{tabular}{lcc}
      \toprule
                                        & Uncomp. Care (Millions) & Case Mix Index\\  
                                        & (1)                     & (2)\\  
      \midrule 
      NP w/ MD Executive x Post Penalty & -4.46                   & 0.03$^{***}$\\   
                                        & (10.39)                 & (0.01)\\   
      \textbf{Comparison Group:}        &                         & \\  
      NP w/out MD Executive             & $\checkmark$            & $\checkmark$\\   
       \\
      Observations                      & 1,440                   & 1,654\\  
      \bottomrule
   \end{tabular}
\end{table}

### 2024-04-23 Tables/MD_noMD_uncompCMI_synth.tex
Updates from Overleaf

\begin{table}[ht!]

\caption{\label{tab:MD_noMD_uncompCMI_synth}Comparison of MD vs. non-MD Nonprofit Uncompensated Care and CMI Synthetic DiD Results}
\centering
\begin{tabular}[t]{lcc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{1}{c}{Uncomp. Care (Millions)} & \multicolumn{1}{c}{Case Mix Index} \\
\cmidrule(l{3pt}r{3pt}){2-2} \cmidrule(l{3pt}r{3pt}){3-3}
 & (1) & (2)\\
\midrule
NP w/ MD Executive x Post Penalty & 3.12 & 0.03$^{**}$\\
 & (3.04) & (0.01)\\
\textbf{Comparison Group:} &  & \\
NP w/out MD Executive & $\checkmark$ & $\checkmark$\\
Observations & 625 & 1635\\
\bottomrule
\end{tabular}
\end{table}

### 2024-04-23 Tables/forprofit_uncompCMI_synth.tex
Updates from Overleaf

\begin{table}[ht!]

\caption{\label{tab:forprofit_uncompCMI_synth} Comparison to For-Profit Uncompensated Care and CMI Synthetic DiD Results}
\centering
\begin{tabular}[t]{lcccccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{3}{c}{Uncomp. Care (Millions)} & \multicolumn{3}{c}{Case Mix Index} \\
 & (1) & (2) & (3) & (1) & (2) & (3)\\
\midrule
For Profit x Post Penalty & 10.53$^{***}$ & 11.69$^{***}$ & 7.6$^{**}$ & 0 & 0.03$^{**}$ & 0\\
 & (2.36) & (2.33) & (2.94) & (0.01) & (0.01) & (0.01)\\
\textbf{Comparison Group:} &  &  &  &  &  & \\
All Nonprofits & $\checkmark$ &  &  & $\checkmark$ &  & \\
Nonprofits w/ MD &  & $\checkmark$ &  &  & $\checkmark$ & \\
\addlinespace
Nonprofits w/out MD &  &  & $\checkmark$ &  &  & $\checkmark$\\
Observations & 4945 & 1330 & 1815 & 12345 & 2895 & 4080\\
\bottomrule
\end{tabular}
\end{table}

### 2024-04-23 Tables/forprofit_synth.tex

\begin{table}[ht!]

\caption{\label{tab:forprofit_synth} Comparison to For-Profit Readmissions and Mortality Synthetic DiD Results}
\centering
\begin{tabular}[t]{lcccccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{3}{c}{Weighted Avg. Readmission Rate} & \multicolumn{3}{c}{Weighted Avg. Mortality Rate} \\
 & (1) & (2) & (3) & (1) & (2) & (3)\\
\midrule
For Profit x Post Penalty & -0.07 & -0.32$^{**}$ & -0.03 & -0.01 & -0.17 & 0.02\\
 & (0.04) & (0.12) & (0.06) & (0.04) & (0.11) & (0.05)\\
\textbf{Comparison Group:} &  &  &  &  &  & \\
All Nonprofits & $\checkmark$ &  &  & $\checkmark$ &  & \\
Nonprofits w/ MD &  & $\checkmark$ &  &  & $\checkmark$ & \\
Nonprofits w/out MD &  &  & $\checkmark$ &  &  & $\checkmark$\\
\\
Observations & 15,595 & 2,900 & 4,705 & 15,515 & 2,895 & 4,690\\
\bottomrule
\end{tabular}
\end{table}

### 2024-04-25 RScripts/analysis.R
add tables for ever MD and 2012 MD without limiting sample to stable teams

@@ -5,9 +5,8 @@ library(ggplot2)
 library(ggpubr)
 library(tidyr)
 library(synthdid)
+library(kableExtra)
 options(knitr.kable.NA=" ")
-
-#### ANALYSIS ##########
 
 # Comparisons I need: 
 # 1. FP - NFP
@@ -23,12 +22,23 @@ options(knitr.kable.NA=" ")
 # Outcomes:
 # 1. Readmission rates
 # 2. Mortality Rates
-# 3. Uncompensated Care
-# 4. Case Mix Index
+# 3. Uncompensated Care and Case Mix Index
+
+# Sample limitations:
+# 1. dropping hospitals with leadership changes
+# 2. not dropping hospitals with leadership changes
 
 # limit to 2010 to 2014
 hospital_data <- readRDS(paste0(created_data_path, "all_hospital_data.rds")) %>%
   filter(year>=2010 & year<=2014)
+
+# drop hospitals with less than 10 beds
+hospital_data <- hospital_data %>%
+  group_by(ID) %>%
+  mutate(maxbeds=max(beds, na.rm=T)) %>%
+  ungroup() %>%
+  filter(maxbeds>=10) %>%
+  select(-maxbeds)
 
 hospital_data$uncomp_care <- abs(hospital_data$uncomp_care)
 hospital_data <- hospital_data %>%
@@ -54,6 +64,8 @@ hospital_data <- hospital_data %>%
 
 comp <- c("FP - NFP", "FP - has MD", "FP - no MD", "has MD - no MD")
 method <- c("full", "matched", "synth")
+
+### ANALYSIS WHERE I DROP HOSPITALS WITH LEADERSHIP CHANGES####
 
 ### OUTCOME: READMISSION ####
 
@@ -399,104 +411,112 @@ uncomp_cmi_analysis <- lapply(comp, function(x){
 ## FULL SAMPLE TABLES ####
 
 # make a table for weighted average readmissions and mortality comparing clinical NFP to non-clinical NFP for full sample and matched
-etable(readmission_analysis[[4]][[1]][["wa_did"]], mortality_analysis[[4]][[1]][["wa_did"]], tex=TRUE,
-       dict = c("weightedavg_read"="Weighted Avg. Readmission Rate", "weightedavg_mort"="Weighted Avg. Mortality Rate",
-                "did"="NP w/ MD Executive x Post Penalty","ID"="Hospital", "year"="Year"),
[…]

### 2024-04-25 RScripts/summary_stats.R

@@ -18,10 +18,69 @@ hospital_data$uncomp_care <- abs(hospital_data$uncomp_care)
 hospital_data <- hospital_data %>%
   mutate(uncomp_care=uncomp_care/1000000) 
 
+# drop hospitals with less than 10 beds
+hospital_data <- hospital_data %>%
+  group_by(ID) %>%
+  mutate(maxbeds=max(beds, na.rm=T)) %>%
+  ungroup() %>%
+  filter(maxbeds>=10) %>%
+  select(-maxbeds)
+
+
+# List of Figures:
+# 1. Graph outcomes over time for forprofit and nonprofit
+# 2. Graph outcomes over time for MD and non-MD
+
 
 # List of tables:
 # Table 1: mean, SD, min, max for the whole sample
 # Table 2: means for different sub-samples
+  
+
+
+## FIGURE 1: outcomes for forprofit and nonprofit over time ####
+forprofit_avgs <- hospital_data %>%
+  filter(profit_status=="forprofit") %>%
+  group_by(year) %>%
+  summarise_at(c("Weighted Avg. Readmission Rate"="weightedavg_read",
+                 "Weighted Avg. Mortality Rate"="weightedavg_mort"), list(mean), na.rm=T) %>%
+  mutate(Group="For Profit")
+nonprofit_avgs <- hospital_data %>%
+  filter(profit_status=="nonprofit") %>%
+  group_by(year) %>%
+  summarise_at(c("Weighted Avg. Readmission Rate"="weightedavg_read",
+                 "Weighted Avg. Mortality Rate"="weightedavg_mort"), list(mean), na.rm=T) %>%
+  mutate(Group="All Nonprofits")
+NP_md_avgs <- hospital_data %>%
+  filter(ever_has_md==1 & no_num_md_change_2010_2014==1) %>%
+  group_by(year) %>%
+  summarise_at(c("Weighted Avg. Readmission Rate"="weightedavg_read",
+                 "Weighted Avg. Mortality Rate"="weightedavg_mort"), list(mean), na.rm=T) %>%
+  mutate(Group="Nonprofit w/ MD")
+NP_nomd_avgs <- hospital_data %>%
+  filter(ever_has_md==0 & no_num_md_change_2010_2014==1) %>%
+  group_by(year) %>%
+  summarise_at(c("Weighted Avg. Readmission Rate"="weightedavg_read",
[…]

### 2024-04-25 Paper/paper.tex
fix latex documents

@@
     \begin{figure}[ht!]
+    \centering
         \caption{Outcomes over time in all penalized hospitals}
-        \includegraphics[scale=.5]{Objects/weighted_read_mort_graph.pdf}
+        \includegraphics[width=\textwidth]{Objects/weighted_read_mort_graph.pdf}
         \label{fig:weighted_read_mort_graph}
@@
 
+    \section{Endogenous Executive Team Changes?}
+
+    \import{Tables}{change_analysis.tex}
+
     \section{Results}
 
-    \import{Tables}{MD_noMD_readmort_synthdid.tex}
+    \import{Tables}{MD_noMD_readmort_synth.tex}
 
-    \import{Tables}{forprofit_synth.tex}
+    \import{Tables}{forprofit_readmort_synth.tex}
 
-    \import{Tables}{MD_noMD_uncompCMI_synth.tex}
+    %\import{Tables}{MD_noMD_uncompCMI_synth.tex}
 
-    \import{Tables}{forprofit_uncompCMI_synth.tex}
+    %\import{Tables}{forprofit_uncompCMI_synth.tex}
 
@@
 
+\includegraphics[width=\textwidth]{Objects/has_doc_avg_map.pdf}
+
 \section{Full Sample Regression Results}
@@
 
-\import{Tables}{forprofit_fullsample.tex}
+\import{Tables}{forprofit_readmort_fullsample.tex}
 
@@
 
-\import{Tables}{forprofit_matchsample.tex}
+\import{Tables}{forprofit_readmort_matchsample.tex}
 

### 2024-04-30 Paper/paper.tex
Updates from Overleaf

@@
 		{\small
-        Nonprofit firms have diverse goals and missions that make understanding their true objectives and behaviors difficult.
-A setting where this is particularly relevant is that of nonprofit hospitals, which make up the majority of hospitals in the US, yet there is mixed evidence on what drives their behavior.
-In some ways they seem to act like for-profits, while in other ways they may care more about prestige or quantity.
-In this paper, I investigate whether the leadership team composition of nonprofit hospitals is correlated with more for-profit behavior.
-I use a novel data set on nonprofit hospital executives in the US from 2009-2015 to investigate whether having clinical experience on a hospital executive team affects hospital behavior.
-In particular, I leverage several pay-for-performance initiatives in the US in 2012, and estimate the difference in response between for-profits, nonprofits, and clinical nonprofits.
-I find that while readmissions decreased on average for all types of hospitals, non-clinical executive team hospital decreased readmissions at a faster rate than those with a clinical executive team, and behaved more similarly to for-profit hospitals.
-The difference equates to approximately 15 additional readmissions for clinical team hospitals.
-
+        In many industries, including hospitals, nonprofits make up the majority of firms in the market.
+Understanding the objectives and behaviors of these firms is important when creating incentives.
+However, nonprofit firms have diverse goals and missions that make understanding their objectives difficult.
+For hospitals in particular, there is mixed evidence on whether nonprofit hospitals differ meaningfully from for-profit hospitals.
+In this paper, I investigate whether leadership team composition of nonprofit hospitals affects how similarly they behaves to for-profits.
+I use a novel data set on nonprofit hospital executives in the US from 2009-2015 to investigate whether the existence of clinical training on a hospital executive team affects hospital behavior.
+In particular, I leverage several pay-for-performance initiatives in the US in 2012, and estimate the difference in response between for-profits, nonprofits, and nonprofits with clinically trained executives.
+I find that while readmissions decreased on average for all types of hospitals, the response was larger for for-profits and nonprofits without clinically trained executives.
+This is consistent with theoretical predictions under the assumption that nonprofits with clinically trained executives put higher weight on societal outcomes than those without clinically trained executives.
 		} 
@@
   The objectives and behaviors of firms that are not classic for-profits are not easily determined, and an extensive literature has sought to understand these types of firms.
-Nonprofits are thought to gain utility not solely from monetary profit, but also through accomplishing mission-based goals.
+Nonprofits are thought to gain utility not solely from monetary profit, but also through accomplishing societal benefit.
 Researchers have speculated a variety of objective functions that could motivate nonprofits such as maximizing prestige, income, or a quality/quantity trade-off (\cite{steinberg1986revealed}).
-An important contribution to this literature is the use of empirical strategies to analyze how firm behavior reveals attributes of the underlying objective function (\cite{sloan2000not}).
-In this paper, I leverage a large-scale shock to hospitals in the US to investigate whether differences in leadership composition reveal characteristics of hospital objective functions, namely the weight placed on profit vs.
-utility gained from benefiting society.
+Empirical contributions to this literature center around investigating behaviors that differ between for-profit, private nonprofit, and government firms (\cite{sloan2000not}).
+In this paper, I investigate whether differences in observed behavior between for-profit and nonprofit hospitals vary differentially by nonprofit leadership team composition.
 
-  Hospitals in particular have been a common focus when seeking to understand nonprofit firms.
+  Hospitals have been a common focus when seeking to understand nonprofit firms.
 This is because, first, there are different hospital ownership types within the industry, giving researchers the ability to compare the behaviors of nonprofits with that of for-profits.
-Second, nonprofit hospitals make up 50\% of all hospitals in the US, and staff, on average, 207 beds.
+Second, private nonprofit hospitals make up 50\% of all hospitals in the US, and staff, on average, 207 beds.
 For comparison, for-profits make up around 36\% of hospitals and staff 107 beds (\cite{ASPE_2023}).
-Thus, nonprofit behavior in the hospital industry is relevant and directly affects consumers of health care.
-However, it remains unclear what type of objective function drives nonprofit hospitals (\cite{erus2002inferring}; \cite{deneffe2002not}; \cite{horwitz2009hospital}).
-In particular, the extent to which nonprofit hospitals gain utility from bettering society (e.g.
-increased charity care, high quality of care, etc.) is unknown.
+Thus, nonprofit behavior in the hospital industry is relevant, and directly affects consumers of health care.
+However, it remains unclear what objective function drives nonprofit hospital behavior, and whether nonprofit hospitals are meaningfully different from for-profits (\cite{sloan2000not}; \cite{erus2002inferring}; \cite{deneffe2002not}; \cite{horwitz2009hospital}).
 
-  What has not been considered is how the leadership team within nonprofits potentially drives objectives.
-In large, publicly traded firms, executive characteristics are shown to be correlated with firm performance, indicating that there is more to firm behavior than for-profit or not-for-profit status (\cite{bertrand2003managing}; \cite{matsa2013female}; \cite{ahern2012changing}).
-Within nonprofit hospitals, there is variation in the propensity to hire executives with a clinical background, which could reasonably change how much weight the hospital places on profit vs.
-quality of care when making decisions.
-I contribute to our understanding of hospital and nonprofit behaviors by investigating whether clinical experience on an executive team affects hospital behavior after being financially penalized.
+  While the characterization of being for-profit or nonprofit is certainly important, there could be other characteristics that drive objectives and behavior that are not perfectly correlated with ownership status.
+Particularly, firm leadership varies widely across firms and potentially drives differences in objectives even within nonprofits.
+In large, publicly traded firms, executive characteristics are shown to be correlated with firm performance, indicating that there is more to firm behavior than for-profit or nonprofit status (\cite{bertrand2003managing}; \cite{matsa2013female}; \cite{ahern2012changing}).
+If hospital objectives differ in the weights placed on profit vs.
+societal benefit, then ownership status certainly plays a role in these weights, but characteristics of those managing the hospital could also play an important role.
+I contribute to our understanding of hospital and nonprofit behaviors by investigating whether occupational background, whether executives have clinical training, affects response to pay-for-performance incentives.
 
@@
 I link hospitals in this data to characteristics in the American Hospital Association survey and Hospital Compare data.
-In order to capture the effect of clinical executives, not just their selection into certain types of hospitals, I consider the sample of hospitals whose leadership team remains stable over time, and characterize them as having clinical executive experience on their team or not, a time-invariant categorization.
-I then leverage the Hospital Readmissions Reduction Program (HRRP) of 2012.
-This program reduced Medicare payments to hospitals with above average readmission rates, and is a shock that is plausibly uncorrelated with previously determined leadership teams.
+I leverage various pay-for-performance initiatives enacted in 2012 to investigate differential hospital responses between for-profits, nonprofits without clinically trained executives, and nonprofits with clinically trained executives.
+While executive teams change over time, I find no evidence that they change in response to the programs.
+This allows me to investigate the role of particular executives in two ways.
+First, I analyze the differential response to incentives of nonprofits who ever hire a clinically trained executive, as this might be a signal of underlying preferences.
+Second, I analyze the differential response of nonprofits who have a clinically trained executive in 2012 compared to those who have a clinically trained executive at other times, but not in 2012.
+These two analyses aim to disentangle the effect of the actual leadership team from the underlying preferences of the hospital.
 
-  I estimate the difference in readmission and mortality rates between hospitals with and without clinical experience before and after the penalties began.
-While readmissions directly affects the likelihood of being penalized, mortality rates are an explicit quality measure unrelated to the penalties.
-Including this as an outcome can help distinguish whether hospitals are improving quality or only taking measures to reduce readmissions.
-Not all hospitals were penalized, but all hospitals (with few exceptions) could have been penalized if readmission rates were high enough.
-Thus, I present estimates first for all hospitals.
-However, these estimates are biased if I am simply capturing the effect of clinical experience on already being a good quality hospital.
-Thus, I present two further specifications: one limiting to only hospitals who were eventually penalized, and another where I limit the sample further to hospitals which have an observably similar match.
+  I estimate how the interaction between hospital type and post program enactment affects readmission and mortality rates among different groups.
+Under the assumption that hospitals would have continued to behave similarly in the absence of the programs, and that leadership teams do not change endogenously with the programs, I identify the differential response to pay-for-performance incentives by different hospital types.
+Among the hospitals in the data, hospitals with clinically trained executives are the minority, and differ on average from other hospitals along various dimensions.
+Thus, my preferred specification is synthetic difference-in-differences, to analyze a group of comparison hospitals with similar characteristics to treatment hospitals.\footnote{I also present results for the full sample, and with a pre-analysis matching procedure and find qualitatively similar results.}
 
-  I assume that before the penalty, clinical and non-clinical team hospitals behaved similarly along quality dimensions, and that the composition of hospital executive teams determined prior to the penalties is exogenous.
-Under these assumptions, I identify the effect of having a clinical executive on hospital response to the penalty.
-I find that while both types of hospitals, those with clinical executives and those without clinical executives, decrease readmissions after being penalized, non-clinical team hospitals decrease readmissions at a faster rate.
-The difference equates to approximately 15 additional readmission cases per year for clinical teams, and there is no evidence that this result is driven by differences in readmissions prior to penalties.
-However, I find no differential effect of the program on mortality rates.
+   While all hospital types decrease readmission rates after the programs are enacted, for-profits and nonprofits without clinically trained executives decrease readmissions more than nonprofits without clinically trained executives.
+That is, nonprofits without clinically trained executives respond to incentives more similarly to for-profits than other nonprofits.
+To investigate the mechanisms by which hospitals decrease readmissions, I also conduct the analyses with uncompensated care and case mix index as outcomes.
+(results of this analysis here)
 
@@
 To this point, researchers have found that nonprofits do not act as purely profit maximizers, but maximize some combination of profit and output (\cite{deneffe2002not}, \cite{chang2011nonprofit}).
-A paper with an approach similar to mine which investigates hospital behavior is \citeauthor{chang2011nonprofit} (\citeyear{chang2011nonprofit}), which looks at the effect of a cost shock on likely of shutting down and mix of profitable services.
+A paper with an approach similar to mine which investigates hospital behavior is \citeauthor{chang2011nonprofit} (\citeyear{chang2011nonprofit}), which looks at the effect of a cost shock on likelihood of shutting down and mix of profitable services.
 They find that nonprofits and for-profits are equally as likely to shut down after the shock, but only nonprofits adjust their mix of profitable services.
-Similarly, I investigate a financial shock to hospitals, but I look at the differential impacts of leadership teams within nonprofits.
-This paper adds to our understanding of inputs into nonprofit hospital objective functions, and whether this is dependent on leadership compositions.
+Similarly, I investigate a change in incentives to hospitals, but I look at the differential impacts of leadership teams within nonprofits.
+This paper adds to our understanding of inputs into nonprofit hospital objective functions.
 
-    Second, many researchers have documented a correlation between executive backgrounds and firm performance specifically for CEOs, starting with \citeauthor{bertrand2003managing} (\citeyear{bertrand2003managing}), who show that manager fixed effects are an important driver of many firm behaviors and decisions.
+    Second, many researchers have documented a correlation between executive backgrounds, specifically for CEOs, and firm performance, starting with \citeauthor{bertrand2003managing} (\citeyear{bertrand2003managing}), who show that manager fixed effects are an important driver of many firm behaviors and decisions.
 Female board members are associated with better oversight and more women executives (\cite{matsa2011chipping}; \cite{adams2009women}), but are negatively correlated with firm value (\cite{ahern2012changing}).
-Having more female executives is correlated with female employee wages and corporate strategies (\cite{flabbi2019female}; \cite{matsa2013female}); young male CEOs tend to be more aggressive in mergers and acquisitions while those with military experience are less aggressive (\cite{levi2010deal}; \cite{benmelech2015military}); CEOs with general ability tend to receive higher pay and perform better (\cite{kaplan2012ceo}; \cite{custodio2013generalists}, \cite{adams2018director}; \cite{frydman2019rising}).
+Having more female executives is correlated with choices of female employee wages and corporate strategies (\cite{flabbi2019female}; \cite{matsa2013female}); young male CEOs tend to be more aggressive in mergers and acquisitions, while those with military experience are less aggressive (\cite{levi2010deal}; \cite{benmelech2015military}); CEOs with general ability tend to receive higher pay and perform better (\cite{kaplan2012ceo}; \cite{custodio2013generalists}, \cite{adams2018director}; \cite{frydman2019rising}).
 Further, Chief Diversity Officers are not found to have any effect on hiring more diversely in universities (\cite{bradley2022impact}).
-Due to data limitations, these studies focus mostly on large, publicly traded firms in the US.
+Due to data limitations, many of these studies focus on large, publicly traded firms in the US.
 However, one study uses data from England to study whether CEOs affect hospital production, and find no association (\cite{janke2019impact}).
+\textcolor{red}{(add paper about Chile hospital executives)}.
 I contribute to this literature first by investigating a context which is not largely studied but is policy relevant, nonprofit US hospitals, and considering a unique executive characteristic pertinent to that context.
-Second, I contribute by addressing leadership team endogeneity concerns through leveraging a shock to hospitals and strategically limiting the sample.
 
@@
 This raises important questions about how to incentivize improvement in quality of care without also incentivizing gamesmanship.
-This paper addresses a necessary first step in considering whether leadership compositions could affect likelihood of gaming, which is to first investigate whether leadership matters to hospital decisions at all.
+This paper addresses a necessary first step in considering whether leadership compositions could affect likelihood of gaming.
+\textcolor{red}{(add papers about other program too)}
 
@@
     The Hospital Readmissions Reduction Program (HRRP) was passed as a part of the Affordable Care Act in 2010 as a pay-for-performance incentive for hospitals to increase quality.
-In 2011, the Center for Medicare and Medicaid Services (CMS) released a set of rules that mandated penalties for hospitals with above average readmission rates.
+In October 2011, the Center for Medicare and Medicaid Services (CMS) released a set of rules that mandated penalties for hospitals with above average readmission rates.
 A readmission is when a patient returns to the hospital within 30 days of being discharged from a previous stay; avoidable readmissions are a bad outcome for patients and increase health care spending.
@@
 In 2015, CMS also included chronic obstructive pulmonary disease, coronary artery bypass graft surgery, and elective primary total hip arthroplasty and/or total knee arthroplasty as conditions which go into the penalty calculation.
+\textcolor{red}{(add citations??)}
 
-    Penalized hospitals pay a flat rate for every Medicare patient, not just those for which they have too high readmissions.
-Further, CMS does not distinguish a necessary readmission vs.
-an avoidable readmission, any repeat hospital visit is included in the penalty calculation.
+    Penalized hospitals pay penalties in the form of a fixed rate for every Medicare patient regardless of the condition.
+Further, CMS does not distinguish a necessary readmission from an avoidable readmission; any repeat hospital visit is included in the penalty calculation.
 Excess readmission rates are calculated using a rolling look-back period of 3 years to determine whether the hospital is penalized.
@@
 These penalties are not insignificant; penalized hospitals paid, on average, 4-5\% of revenue.
+
+    \textcolor{red}{(is the program still going?)}
+
+    \subsection{Hospital Value-Based Purchasing Program}
 
@@
 They select members of the executive team to carry out day-to-day operations.
-While hospital executives typically do specialize by getting degree(s) in healthcare management or an MBA specific to healthcare, they do not often have hands on clinical experience.
-When there is a hospital executive with clinical experience, they likely decided on that career path after becoming a doctor.
+While hospital executives typically do specialize by getting degree(s) in healthcare management or an MBA specific to healthcare, they do not often have clinical training.
 While some doctors earn additional degrees before stepping into an executive role, this is not a necessary condition to becoming a physician executive.
-An executive team usually consists of at least a CEO and CFO.
-Every nonprofit is different in exactly how they structure an executive team.
+An executive team usually consists of at least a CEO and CFO, but every nonprofit is different in exactly how they structure their executive team.
 Some nonprofits hire Chief Medical Officers, Chief Quality Officers, executive directors, vice presidents of various departments, etc.
@@
 
-    All tax-exempt organizations in the US are required to file a Form 990 with the IRS each year.
-There are different types of forms, but any organization that grosses over \$200,000 must file the most extensive form 990.
-Sections of the form include a statement of revenue, statement of functional expenses, a balance sheet, and, as used in this project, a list of all key employees, executives, and board members.
+    All tax-exempt organizations in the US are required to file a Form 990 with Internal Revenue Services (IRS) each year.
+There are different types of forms, but any organization grossing over \$200,000 must file the most extensive Form 990.
+Sections of this form include a statement of revenue, statement of functional expenses, a balance sheet, and, as used in this project, a list of all key employees, executives, and board members.
 In this section, each firm is required to report the name and title, average hours per week, position, and compensation of their board members and executive leaders.
-Since these firms are tax exempt, their financial information is made publicly available through the public housing of these forms.
+Since these firms are tax exempt, their information in these forms is made publicly available.
 
@@
 
-    To characterize hospitals as having clinical experience on their executive team or not, I construct a novel data set of nonprofit hospitals in the US that contains names and select characteristics of executives tied to the hospital from 2009-2015.
+    To characterize hospitals as having clinically trained executives or not, I construct a novel data set of nonprofit hospitals in the US that contains names and select characteristics of executives tied to the hospital from 2009-2015.
 I gather this data from each hospital's publicly available Tax Form 990, which every sufficiently large nonprofit files each year in the US.
@@
 
-    The historical tax forms for all nonprofits are housed by ProPublica\footnote{https://projects.propublica.org/nonprofits/}.
-I use the NonProfit Explorer API to extract the Employee Identification Numbers (EIN) of all nonprofits classified as a hospital according to their National Taxonomy of Exempt Entities (NTEE) code.
+    The historical tax forms for all nonprofits are housed by ProPublica.\footnote{https://projects.propublica.org/nonprofits/} I use the NonProfit Explorer API to extract the Employee Identification Numbers (EIN) of all nonprofits classified as a hospital according to their National Taxonomy of Exempt Entities (NTEE) code.
 After filtering out associations and specialty hospitals, there are around 3,000 firms left.
-Importantly, I need to be able to link hospitals in the tax forms to other hospital characteristics, including penalty status and readmission rates, in the American Hospital Association annual survey and Hospital Compare data.
+Importantly, I need to be able to link hospitals in the tax forms to other hospital characteristics from publicly availble sources.
 No official crosswalk between EINs and AHA IDs or Medicare numbers exists, so I must rely on matching by location and name of hospital in both data sets.
-After limiting to only general, nonprofit hospitals in the AHA survey, I confidently match approximately 1200 EINs to an AHA identification number based on exact name matches within the same state.
+After limiting to only general, nonprofit hospitals that I can link to the AHA survey, I confidently match approximately 1200 EINs to an AHA identification number based on exact name matches within the same state.
 
-    For each of the 1200 EINs, I extract web URLs to Form 990 PDFs in each year.
+    For each of the 1200 EINs, I then extract web URLs to Form 990 PDFs in each year.
 I download these locally and use optical character recognition (OCR) text extraction methods to scrape the relevant section of the pdf, which is titled ``Officers, Directors, Trustees, Key Employees, and Highest Compensated Employees".
 Finally, I use string cleaning methods to identify names and positions for each EIN, year.
-I also identify titles indicating a physician: md, dr., or do.
+I also identify titles indicating a physician: MD, Dr., or DO.
 Using OCR extraction can be unreliable in some instances when the text is handwritten or the font is too small.
@@
 I drop anyone who is solely a board member, and focus on those who have executive in their title or are labeled as a president or vice president of a specific aspect of the hospital.
-My identification strategy relies on hospitals that have stable executive leadership over time.
-The ideal set of hospitals would be those who have the same executives on their team for the entire sample, as changes could be correlated with the fact that the hospital is penalized.
-However, this rarely occurs in reality.
-Therefore, I attempt to proxy for a stable leadership team while still having a large enough sample to run an analysis.
-Since my focus is clinical experience, my preferred specification is to limit the sample to hospitals who do not change the presence of a doctor on their leadership team from 2010-2014 (2 years before and after the program began).
-I make this limitation to hospitals in the sample and am left with 524 stable hospitals over time.
-Finally, I aggregate the information up to the hospital level, where an observation captures whether the hospital has clinical experience on their executive team over the course of the sample.
+I create various hospital-level characterizations based on their executive teams, such as the existence of a clinically trained executive, the number of clinically trained executives, the number of total executives, and whether the hospital has a Chief Medical Officer.
 
@@
 
-    I merge the leadership information with the AHA survey using the EIN-AHA links I created.
-From AHA, I gather the number of beds in the hospital and the Medicare number, which I use to merge the Hospital Compare data.
-Hospital Compare contains multiple data sets which I use first to limit to hospitals that were penalized in the Hospital Readmissions Reduction Program (HRRP), and also for information on outcome variables: readmission and mortality rates for the relevant HRRP conditions: heart failure, heart attack (AMI), and pneumonia.
+    I include hospital information from the AHA survey on the number of beds in the hospital and the Medicare number, which I use to merge the Hospital Compare and Healthcare Cost Reporting Information System (HCRIS) data.
+Further, I use the AHA categorization of for-profit or nonprofit.
+Hospital Compare contains information on penalties from HRRP and readmission and mortality rates for the relevant HRRP conditions: heart failure, heart attack (AMI), and pneumonia.
+From HCRIS, I gather information on penalties from the HVBP program, operating expenses, and uncompensated care.
+I also link to the publicly available CMS file containing each hospital's case mix index.
 
@@
 Further, Hospital Compare also publishes hospital level readmission and mortality rates for heart failure, AMI, and pneumonia.
-For my main outcomes of interest, I take an average across conditions and weight it using the total number of patients seen for that condition.
-For further analysis, I create samples of hospitals who were penalized for specific conditions, and investigate the change in readmission and mortality rates for the condition they were penalized in.
+For my main outcomes of interest, I take an average across conditions, weighted by the total number of patients seen for that condition.
 
@@
 
-    In Table \ref{tab1:sumstats}, I present summary statistics for various hospital characteristics including which condition hospitals are penalized for, and the readmission and mortality outcome variables.
-After limiting to penalized hospitals with stable leadership teams, my sample consists of between 302-310 hospitals from 2009-2015, except in the case of AMI outcomes, which is only 225-264 hospitals due to small patient numbers.
-Fourteen percent of hospitals have someone with clinical experience on their executive team, and the size of the hospitals in the sample varies quite a bit.
-The average number of beds for a hospital in the sample is 186, only slightly lower than the national average for all nonprofits.
-Hospitals in the sample are less likely to be penalized for AMI than either heart failure or pneumonia.
-On average, 21\% of patients with one of these conditions would be readmitted.
+    In Table \ref{sumstats}, I present summary statistics for hospital variables.
+There are 3,766 total hospitals in the sample, with 150 beds on average.
+Hospitals in the sample are less likely to be penalized for AMI than either heart failure or pneumonia under the HRRP program.
+On average, 20\% of patients with one of these conditions would be readmitted.
 While readmission rates for the specific conditions are similar, the highest likelihood of readmission is for heart failure patients.
@@
 
-    Figure \ref{fig:weighted_read_mort_graph} shows average readmission and mortality rates across time and across the three different conditions, along with the weighted average.
-There are multiple notable points the raw data reveals.
-First, prior to the penalty, hospitals without clinical executives had consistently higher readmission rates than clinical executive hospitals across all conditions.
-While there is a gap between clinical team and non-clinical team readmission rates, the lines seem to move mostly parallel.
-In contrast, 2012 begins a decline in readmission rates over all conditions for all types of hospitals.
-This makes sense and is consistent with previous literature on HRRP, as hospitals have incentive to reduce readmissions in order to reduce future penalties.
-However, non-clinical hospital teams decrease readmissions at a faster rate, quickly converging with the rates of clinical team hospitals.
-While suggestive, this motivates further analysis into the differential responses between clinical and non-clinical leadership teams.
-On the right panel, average mortality rates over time reveal no clear change due to the penalty, and no consistent pattern in the difference between clinical and non-clinical hospital teams.
+    %\import{Tables}{overall_sumstats.tex}
+
+    I present a table of means for each hospital type (for-profit, all nonprofits, nonprofits with a clinically trained executive, and nonprofits without a clinically trained executive) in Table \ref{tab:sumstats_samples}.
+Eighty-five percent of hospital in the sample are classified as nonprofits.
+Importantly, the two sub-samples of nonprofits do not add up to the total number of nonprofits because of the limited ability to extract leadership text for all nonprofits.
+There are key differences in nonprofits with and without clinically trained executives: those with MDs are larger, more likely to be penalized, have higher uncompensated care, and more complex patients.
+
+    %\import{Tables}{sample_sumstats.tex}
+
+    I plot readmission and mortality rates across time for different hospital types in Figure \ref{fig:weighted_read_mort_graph}.
+While this is purely descriptive evidence, the trends in outcomes point to more similar behavior between for-profits and nonprofits without clinically trained executives in comparison to nonprofits with a clinically trained executive.
 
@@
 
+    Finally, I present a table of correlations between having at least one clinically trained executive and other time invariant hospital characteristics.
+There may be a concern that this particular characteristic of executive teams is simply a proxy for another quality, such as being an academic medical center.
+Unsurprisingly, the most correlated characteristic with having a clinically trained executive is having a CMO on the executive team.
+However, even this correlation is only .4, meaning there is still variation in having clinically trained executives even when controlling for having a CMO.
+
     \import{Tables}{hosp_correlations.tex}
 
-    \section{Model and Empirical Strategy}
+    \section{Model and Empirical Strategy}\label{sec:model}
 
-    Not knowing the true form of nonprofit hospital objective functions, we can assume that they maximize some weighted average of revenue and output.
-Abstracting away from quantity for the purpose of this paper, we can think of firms behaving according to:
+    \subsection{First Period}
 
-    $$\max_{\theta}\hspace{2mm} \delta R + (1-\delta) u(\theta),$$
+    Not knowing the true form of hospital objective functions, we can consider them maximizing some weighted average of profit and societal (or patient) benefit.
+Abstracting away from quantity and price as functions of quality for the purpose of this paper, I present a two-stage model of firm behavior that captures the change of quality due to pay-for-performance incentives.
+In the first period, prior to program enactment, hospitals do not expect any penalties in the future, and choose quality according to:
 
-    \noindent where nonprofits differ from for-profits in that $\delta\neq 1$.
-Then $R$ is net revenue, $u(.)$ is an increasing function that captures utility from quality of care, and $\theta$ is some quality of care measure.
-In this setting, we care about hospital responding to a penalty by adjusting readmission rates, so $\theta = 1-r$, where $r$ is the readmission rate.
-While previous research has set out to understand the average $\delta$ for nonprofit hospitals, $\delta$ may also be dependent on the leadership teams within nonprofit hospitals.
-Further, introducing the HRRP means that $R$ is directly dependent on $\theta$.
-We can then write the hospital's problem as
+    $$\max_{\theta}\hspace{2mm}\alpha\left[R - c_{\theta}(\theta) \right] + (1-\alpha) u(\theta),$$
 
-    $$\max_{\theta}\hspace{2mm} \delta(L) R(\theta) + (1-\delta(L)) u(\theta)$$
+    \noindent where $R$ is net revenue, $c_{\theta}(.)$ is an increasing cost function specific to quality $\theta$, $u(.)$ is extra utility, which is increasing and concave is $\theta$, and $\alpha\in[0,1]$ is a hospital's type, indicating how much they care about profit vs.
+societal benefit.
+There is an implicit stay open condition throughout the model.
+Taking the first order condition yields
 
-    \noindent where $L$ captures the composition of the leadership team.
-Therefore, when maximizing this function, the choice of $\theta$ is dependent on $L$ through its impact on how the hospital values profit vs.
-quality of care:
+    $$(1-\alpha)u'(\theta) = \alpha c_{\theta}'(\theta),$$
 
-    $$\theta = f(\delta(L)).$$
+    \noindent marginal benefit of increasing quality equals marginal cost. Solving for $u'(\theta)$ and differentiating with respect to $\alpha$ yields
 
-    Thus, I directly estimate the effect of having clinical experience on readmission rates.
-Additionally, I estimate the effect on mortality rates to capture any spillover effects of improving readmissions into another dimension of quality.
-However, looking purely at this correlation is a biased picture of the effect of clinical experience since it likely captures reverse causality: certain types of hospitals may be more likely to hire doctors as executives.
-I do two things to address this issue.
-First, as discussed in detail in Section \ref{sec:data}, I only consider hospitals with stable leadership teams over time.
-Second, I leverage the HRRP shock, limiting the sample to those that were penalized, which allows me to identify hospitals similar in quality, and to observe changes in behavior after the shock.
-For the hospitals in my sample, I define their type as having clinical experience on their executive team or not, a time invariant categorization.
+    $$\frac{du'(\theta)}{d\alpha} = c_{\theta)}'(\theta)\frac{1}{(1-\alpha)^2} > 0.$$
 
+    Thus, by the Implicit Function Theorem,
+
+    $$\frac{d\theta}{d\alpha} = \frac{du'(\theta)/d\alpha}{du'(\theta)/d\theta} < 0.$$
+
+    That is, the more weight the hospital places on profit, the lower quality of care they will provide.
+
+    \subsection{Second Period}
+
+    Hospitals determine period 1 quality $\theta_1$ according to $\theta(\alpha)$ described above.
+Then, in period two, hospitals are penalized according to $\theta_1$, and they choose a new quality, $\theta_2$, which will determine penalties in the future.
+The penalty $(k)$ comes in the form of a negative shock to revenue for hospitals with quality below a certain threshold.
+That is,
+
+    $$R_2 = \begin{cases}
+        R - k, \text{  if  } \theta_1<\overline{\theta}\\
+        R, \text{  if  } \theta_1\geq \overline{\theta}.
+    \end{cases}$$
+
+    The penalty $k$ is large and exceeds the maximum cost of quality.
+The penalty threshold, $\overline{\theta}$, is exogenous.
+Since $k>max\{c_{\theta}(\theta)\}$, all hospitals will choose $\theta_2$ to not be penalized in the future.
+Thus,
+
+    $$\theta_2 = \begin{cases}
+        \overline{\theta}, \text{  if  } \theta_1<\overline{\theta}\\
+        \theta_1, \text{  if  } \theta_1\geq \overline{\theta}.
+    \end{cases}$$
+
+    Taking the change in theta from period 1 to period 2,
+
+    $$\Delta\theta = \theta_2-\theta_1 = \begin{cases}
+        \overline{\theta}-\theta_1, \text{  if  } \theta_1<\overline{\theta}\\
+        0, \text{  if  } \theta_1\geq \overline{\theta}.
+    \end{cases}$$
+
+    Differentiating with respect to $\alpha$, we get
+
+    $$\frac{d\Delta\theta}{d\alpha} = \begin{cases}
+        -\frac{d\theta_1}{d\alpha}, \text{  if  } \theta_1<\overline{\theta}\\
+        0, \text{  if  } \theta_1\geq \overline{\theta}.
+    \end{cases}$$
+
+    I show in period one that $\frac{d\theta}{d\alpha}<0$.
+Therefore, $\frac{d\Delta\theta}{d\alpha}>=0$.
+That is, the response to the program incentives is weakly increasing in $\alpha$, and there is a threshold for which increasing $\alpha$ (a hospital with more weight on profit), increases the response to the penalty.
+
+    The theoretical prediction is that change in quality depends on how much weight the hospital places on profit vs.
+societal benefit.
+However, $\alpha$ for each hospital is unobserved.
+Thus, I empirically investigate whether the response to the program is correlated with hospital ownership or clinical training on the executive team in order to gain understanding about $\alpha$.
+
+    \subsection{Estimation}
+
+    First, I test empirically whether nonprofits respond to program incentives differently from for-profits.
+While nonprofit hospital objectives are largely unknown, for-profits have revealed their objective in their ownership classification.
+That is, $\alpha=1$.
+To learn something about $\alpha$ for nonprofit firms with different leadership compositions, I compare their program response to that of for-profits.
 I estimate the following regression equation:
@@
     \begin{equation}
-    \label{eq:main}
-    y_{ht} = \beta (clinical \times post\_penalty)_{ht} + \alpha_{h} + \delta_t + \epsilon_{ht}
+    \label{eq:forprofit}
+    y_{ht} = \beta (forprofit_h \times post\_penalty_t) + \alpha_{h} + \delta_t + \epsilon_{ht}
     \end{equation}
 
-    \noindent where $y_{ht}$ is one of the outcome variables discussed in Section \ref{sec:data}, $\beta$ is the variable of interest capturing the effect of having clinical experience after the penalties took place, $\alpha_h$ and $\delta_t$ are hospital and time fixed effects, respectively.
+    \noindent where $y_{ht}$ is one of the outcome variables discussed in Section \ref{sec:data}, $\beta$ is the coefficient of interest capturing the effect of being for-profit after the penalties took place, and $\alpha_h$ and $\delta_t$ are hospital and time fixed effects, respectively.
+For each outcome, I estimate three models where the comparison group differs by hospital type: all nonprofits, nonprofits with a clinically trained executive, and nonprofits without a clinically trained executive.
+
+    An important distinction here is whether hospital time is a time-invariant characterization.
+For nonprofits and for-profits, switching is relatively rare, and I limit to those who never switch.
+However, hospital executive teams are dynamic, and whether or not a hospital employs a clinically trained executive in each period can change.
+Under the assumption that these changes are not correlated with the program enactments (see Section \ref{sec:endog}), these changes can be advantageous in understanding hospital objectives.
+There are two ways that having a clinically trained executive might affect outcomes: first, a hospital that hires a clinically trained executive is signalling their underlying $\alpha$.
+Second, hiring a clinically trained executive who carries out day-to-day operations in a specific way actually changes the underlying $\alpha$ of the hospital.
+Thus, I use two types of variation in clinically trained executives to disentangle the two: a variable for whether the hospital ever hires a clinically trained executive, and a variable for whether the hospital has a clinically trained executive in 2012 when the programs took place.
+
+    Similarly, I test empirically whether having a clinically trained executive affects hospital response to program incentives.
+Thus, I limit the sample to only hospitals for which I have leadership team information, and estimate:
+
+    \begin{equation}
+    \label{eq:clinical}
+    y_{ht} = \beta (clinical_h \times post\_penalty_t) + \alpha_{h} + \delta_t + \epsilon_{ht}.
+    \end{equation}
+
     As discussed in Section \ref{sec:hrrp}, hospitals had incentive to decrease readmissions immediately when the rules were released in October of 2011, before the penalties were actually determined in October of 2012.
-However, I still define $post\_penalty$ as occurring in 2012 or later, the first full year when hospitals could respond.
+However, I still define $post\_penalty$ as occurring in 2012 or later, the first full of hospital response.
 
-    For identification, I rely on the assumption that, prior to being penalized, clinical and non-clinical team hospitals did not differentially change readmission or mortality rates for a reason I do not capture here.
-I investigate this assumption by estimating event study coefficients, where the coefficients $\beta_j$ now capture the effect of having clinical experience in a particular year, and $\beta_{2011}$ is normalized to zero:
+    To identify the causal effect of hospital type on hospital program response, I rely on several assumptions.
+First, I assume that executive teams are not determined endogenously when the programs were enacted.
+I investigate the validity of this assumption in Section \ref{sec:endog}.
+Second, I assume that no other unobserved changes occurred in 2012 that differentially impact the outcomes of different hospital types.
+Finally, I assume that, absent the program enactments, different hospital types would have had parallel trends in outcomes.
+While not directly testable, I investigate the presence of pre-trends using an event study specification:
 
@@
     \label{eq:es}
-    y_{ht} = \sum_{j=2009}^{2010}\beta_j(clinical_{ht} \times \mathbf{1}\{y=j\}) + \sum_{j=2012}^{2015}\beta_j (clinical_{ht} \times \mathbf{1}\{y=j\}) + \alpha_{h} + \delta_t + \epsilon_{ht}.
+    y_{ht} = \sum_{j=2009}^{2010}\beta_j(hospitaltype_{h} \times \mathbf{1}\{y=j\}) + \sum_{j=2012}^{2015}\beta_j (hospitaltype_{h} \times \mathbf{1}\{y=j\}) + \alpha_{h} + \delta_t + \epsilon_{ht},
     \end{equation}
 
-    \section{Endogenous Executive Team Changes?}
+    \noindent  where the coefficients $\beta_j$ now capture the effect of having being a specific hospital type in a particular year, and $\beta_{2011}$ is normalized to zero.
+\textcolor{red}{(figure out how to shift this to synthetic difference in difference specification and talk about the differences across hospital types)}
+
+    \section{Results}
+
+    \subsection{Changes in Executive Team}\label{sec:endog}
+
+    Before presenting the main results of the paper, I investigate whether executive team changes are occurring endogenously due to the programs.
+If hospitals change their propensity to hire a clinically trained executive because they expect to be penalized or want to avoid penalties, then there is selection in the independent variable.
+First, I analyze whether changes in clinically trained executives are concentrated around the program enactment, 2011-2012.
+Second, I analyze whether hospitals who end up getting penalized are more likely to change clinically trained executives on their team than non-penalized hospitals.
+That is, are hospitals that expect to be more affected by the initial penalties more likely to hire or fire clinically trained executives?
+I consider changes along having any clinically trained executive as well as changes in the number of clinically trained executives.
+I estimate the following regression equations:
+
+    \begin{equation}
+    change_{ht} = \sum_{j=2011}^{2014}\beta_j\mathbf{1}\{t=j\} + \alpha_h + \epsilon_{ht},
+    \end{equation}
+
+    \begin{equation}
+    change_{ht} = \beta(everpenalized_{h} \times post\_2011)+ \alpha_h + \delta_t + \epsilon_{ht}.
+    \end{equation}
+
+    The estimates from this analysis are presented in Table \ref{tab:change_analysis}.
+Changes in any clinically trained executive are less likely to occur in 2012-2014 relative to 2010, indicating that executives are actually becoming more stable after the programs.
+Further, hospitals who end up being penalized are not more likely to change their propensity to hire any clinically trained executive.
+Results are similar when considering changes in the number of clinically trained executives.
+Thus, I conclude that changes in executive teams are not endogenously occurring with the change in incentives due to the programs.
 
@@
 
-    \section{Results}
+     \subsection{Readmissions and Mortality}
+
+     First, I present results for the differential program response of for-profit firms compared to the different sub-samples of nonprofit firms in Table \ref{tab:forprofit_synth}.
+Looking first at the left panel, for-profits change their readmission rates after the program similarly to all nonprofits taken together, as well as nonprofits without clinically trained executives.
+However, for-profits have a .3ppt larger decrease in readmissions than nonprofits with clinically trained executives.
+When considering the right panel, there is no statistical difference in change in mortality rates between different hospital types.
+
+     \import{Tables}{forprofit_readmort_synth.tex}
+
+     Next, I directly estimate the difference in response by nonprofits with and without clinically trained executives.
+These results are presented in Table \ref{tab:MD_noMD_synth}.
+Nonprofits without clinically trained executives have a larger decrease in readmissions than those with clinically trained executive by .3ppts, while there is no difference in change in mortality rates.
+
+     These findings are consistent with the theoretical prediction presented in Section \ref{sec:model} if having a clinically trained executive means the hospital places more weight on societal benefit than for-profits and other nonprofits do.
 
@@
 
-    \import{Tables}{forprofit_readmort_synth.tex}
-
-    %\import{Tables}{MD_noMD_uncompCMI_synth.tex}
-
-    %\import{Tables}{forprofit_uncompCMI_synth.tex}
-
-    I present the estimated coefficient of interest from Equation \ref{eq:main}, $\beta$, in Table \ref{tab:mainresults}.
-In column 1, the outcome is the weighted average of readmission rates.
-The estimated effect of having clinical experience on the hospital's reaction to the penalty is .47, indicating that hospitals with clinical experience on their executive team have a .47 point higher readmission rate than non-clinical team hospitals after being penalized.
-Comparing this to an average readmission rate of 21, this is a difference of approximately 2\%.
-When considering an average total number of patients of 760, this is a difference of 15 readmissions per year.
-Based on previous studies and on Figure \ref{fig:weighted_read_mort_graph}, both types of hospitals decrease readmission rates after the penalties begin, so a positive $\beta$ indicates that non-clinical team hospitals decrease readmissions at a faster rate than clinical team hospitals.
-
-    In columns 2, 3, and 4, I limit the sample to hospitals penalized specifically for the condition that matches the outcome in the model.
-The coefficients are consistent and similar to that of the weighted average coefficient.
-Therefore, one of the conditions is not driving the entire weighted average.
-While similar in magnitude, the estimate for $\beta$ with the outcome of AMI readmission rate is statistically insignificant.
-Out of all the outcomes, AMI is the most likely to be missing data due to insufficient number of patients, which could explain this.
+    Finally, I plot estimated coefficients from Equation \ref{eq:es} in Figure \ref{fig:wa_eventstudy}.
+The result is consistent with clinical team hospitals having higher readmissions post-penalty.
+The magnitude of the effect in the years after the penalty began are still around .5, but range from .26 to .54 from 2012-2014.
+Further, there is no indication that having clinical experience on the executive team causes a reaction in readmissions prior to the penalty.
 
@@
 
-    Finally, I plot estimated coefficients from Equation \ref{eq:es} in Figure \ref{fig:wa_eventstudy}.
-The result is consistent with clinical team hospitals having higher readmissions post-penalty.
-The magnitude of the effect in the years after the penalty began are still around .5, but range from .26 to .54 from 2012-2014.
-Further, there is no indication that having clinical experience on the executive team causes a reaction in readmissions prior to the penalty.
+    \subsection{Uncompensated Care and Case Mix Index}
 
-    Estimates for the coefficient of interest from Equation \ref{eq:main} with mortality rates are presented in Table \ref{tab:mort_results}.
-Column 1 presents my estimate of the effect of having a clinical experience team on the weighted average of mortality.
-Columns 2, 3, and 4 present estimates where the sample is limited to hospitals who are penalized for the outcome being considered.
-The coefficients are all positive, ranging from .13 to .31, but none of them are statistically significant.
-This is consistent with prior literature that finds hospitals do not change mortality rates after HRRP, and this result reveals that holds across different leadership teams.
+    As a secondary analysis, I present estimates of differential response of potential mechanisms, changing input into another societal benefit measure: uncompensated care, and the complexity of patients seen in the hospital: case mix index.
+First, I investigate whether for-profits change these differentially than types of nonprofits in Table \ref{tab:forprofit_uncompCMI_synth}.
+
+    \import{Tables}{forprofit_uncompCMI_synth.tex}
+
+    Next, I estimate the difference in response for nonprofits with and without clinically trained executives, shown in Table \ref{tab:MD_noMD_uncompCMI_synth}.
+
+    \import{Tables}{MD_noMD_uncompCMI_synth.tex}
 
@@
 
-    In this paper, I study the effect of having clinical experience within hospital leadership on nonprofit hospital behavior.
-Specifically, I leverage the Hospital Readmissions Reduction Program, which began in 2012, to identify similar quality hospitals along the dimension of readmission rates.
-For these similar hospitals, I estimate the difference in readmissions after being penalized between hospitals with clinical executives and hospitals without clinical executives.
-I find that while the penalty program led both types of hospitals to decrease readmissions after being penalized, those without clinical experience on their executive team decreased rates more quickly than non-clinical team hospitals.
-
-    This finding reveals important information about hospital objectives, specifically that the composition of a hospital's leadership team affects the value placed on profit (or in this case, avoiding loss) vs.
-quality of care.
-Therefore, policymakers should take into consideration the composition of the leadership team when designing pay-for-performance incentives.
+    In this paper, I study differential objective between different firm types.
+While the differences in for-profit and nonprofit hospitals have been studied in the past, it is unclear whether certain nonprofit characteristics make them act more like for-profits.
+In order to better understand the role of leadership in objectives, I estimate the differential response to a change in incentives among for-profits, nonprofits, and nonprofits with and without a clinically trained executive executives.
+I find that nonprofits without clinically trained executives respond to incentives much more similarly to for-profits than to nonprofits without clinically trained executives.
 

### 2024-04-30 Tables/MD_noMD_readmort_decomp_fullsample.tex
add decomposition tables

\begin{table}[htbp]
   \caption{\label{tab:MD_noMD_readmort_decomp_fullsample} Decomposition of Readmission and Mortality Results for Nonprofits}
   \bigskip
   \centering
   \begin{tabular}{lcccc}
      \toprule
       & \multicolumn{2}{c}{Weighted Avg. Readmission Rate} & \multicolumn{2}{c}{Weighted Avg. Mortality Rate}\\
                                       & (1)           & (2)           & (3)           & (4)\\  
      \midrule 
      Ever MD Executive x Post Penalty & -0.03         &               & -0.02         &   \\   
                                       & (0.07)        &               & (0.07)        &   \\   
      MD Executive in 2012             &               & -0.27$^{***}$ &               & -0.11\\   
                                       &               & (0.10)        &               & (0.10)\\   
      \textbf{Comparison Group:}       &               &               &               & \\  
      Never MD                         & $\checkmark$  &               & $\checkmark$  &  \\  
      Has MD not in 2012               &               & $\checkmark$  &               & $\checkmark$\\   
       \\
      Observations                     & 4,095         & 1,883         & 4,083         & 1,880\\  
      \bottomrule
   \end{tabular}
\end{table}

### 2024-04-30 Tables/MD_noMD_readmort_decomp_synth.tex
add decomposition tables

\begin{table}[ht!]

\caption{\label{tab:MD_noMD_readmort_decomp_synth}Decomposition of Readmissions and Mortality for Nonprofits Synthetic DiD Results}
\centering
\begin{tabular}[t]{lcccl}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Weighted Avg. Readmission Rate} & \multicolumn{2}{c}{Weighted Avg. Mortality Rate} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
 & (1) & (2) & (1) & (2)\\
\midrule
Ever MD Executive x Post 2012 & 0 &  & -0.06 & \\
 & (0.07) &  & (0.06) & \\
2012 MD Executive x Post 2012 &  & -0.19$^{*}$ &  & -0.1\\
 &  & (0.09) &  & (0.09)\\
 &  &  &  \vphantom{1} & \\
\addlinespace
\textbf{Comparison Group:} &  &  &  & \\
Never MD & $\checkmark$ &  & $\checkmark$ & \\
Has MD not in 2012 &  & $\checkmark$ &  & $\checkmark$\\
 &  &  &  & \\
Observations & 3925 & 1845 & 3910 & 1840\\
\bottomrule
\end{tabular}
\end{table}

### 2024-04-30 Tables/MD_noMD_uncompCMI_decomp_synth.tex
add decomposition tables

\begin{table}[ht!]

\caption{\label{tab:MD_noMD_uncompCMI_decomp_synth}Decomposition of Uncompensated Care and CMI for Nonprofits Synthetic DiD Results}
\centering
\begin{tabular}[t]{lcccl}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Uncompensated Care} & \multicolumn{2}{c}{Case Mix Index} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
 & (1) & (2) & (1) & (2)\\
\midrule
Ever MD Executive x Post 2012 & 0.01 &  & 0.01$^{*}$ & \\
 & (0.01) &  & (0) & \\
2012 MD Executive x Post 2012 &  & -0.03 &  & 0.01\\
 &  & (0.03) &  & (0.01)\\
 &  &  &  \vphantom{1} & \\
\addlinespace
\textbf{Comparison Group:} &  &  &  & \\
Never MD & $\checkmark$ &  & $\checkmark$ & \\
Has MD not in 2012 &  & $\checkmark$ &  & $\checkmark$\\
 &  &  &  & \\
Observations & 1195 & 635 & 3015 & 1595\\
\bottomrule
\end{tabular}
\end{table}

### 2024-04-30 Paper/paper.tex
Updates from Overleaf

@@
 
+    \import{Tables}{stable_sample_sumstats.tex}
+
     In Table \ref{sumstats}, I present summary statistics for hospital variables.
@@
     \label{eq:forprofit}
-    y_{ht} = \beta (forprofit_h \times post\_penalty_t) + \alpha_{h} + \delta_t + \epsilon_{ht}
+    y_{ht} = \beta (for\_profit_h \times post\_penalty_t) + \alpha_{h} + \delta_t + \epsilon_{ht}
     \end{equation}
@@
 Finally, I assume that, absent the program enactments, different hospital types would have had parallel trends in outcomes.
-While not directly testable, I investigate the presence of pre-trends using an event study specification:
-
-    \begin{equation}
-    \label{eq:es}
-    y_{ht} = \sum_{j=2009}^{2010}\beta_j(hospitaltype_{h} \times \mathbf{1}\{y=j\}) + \sum_{j=2012}^{2015}\beta_j (hospitaltype_{h} \times \mathbf{1}\{y=j\}) + \alpha_{h} + \delta_t + \epsilon_{ht},
-    \end{equation}
-
-    \noindent  where the coefficients $\beta_j$ now capture the effect of having being a specific hospital type in a particular year, and $\beta_{2011}$ is normalized to zero.
-\textcolor{red}{(figure out how to shift this to synthetic difference in difference specification and talk about the differences across hospital types)}
 
@@
     \begin{equation}
-    change_{ht} = \beta(everpenalized_{h} \times post\_2011)+ \alpha_h + \delta_t + \epsilon_{ht}.
+    change_{ht} = \beta(ever\_penalized_{h} \times post\_2011)+ \alpha_h + \delta_t + \epsilon_{ht}.
     \end{equation}
@@
 
-     \subsection{Readmissions and Mortality}
+     \subsection{Readmission and Mortality Rates}
 
@@
 
-    Finally, I plot estimated coefficients from Equation \ref{eq:es} in Figure \ref{fig:wa_eventstudy}.
-The result is consistent with clinical team hospitals having higher readmissions post-penalty.
-The magnitude of the effect in the years after the penalty began are still around .5, but range from .26 to .54 from 2012-2014.
-Further, there is no indication that having clinical experience on the executive team causes a reaction in readmissions prior to the penalty.
-
-    \begin{figure}[ht!]
-        \caption{Event Study Estimates, Readmission}
-        \includegraphics[scale=.5]{Objects/read_es_graph.pdf}
-        \label{fig:wa_eventstudy}
-    \end{figure}
+    \subsubsection{Decomposition}
 
@@
     \import{Tables}{MD_noMD_uncompCMI_synth.tex}
-
-    I investigate whether mortality rates seem to be driven by a pre-trend, or whether the effect of clinical experience on mortality rates differs across time after the penalty occurs.
-Event study estimates from Equation \ref{eq:es} for the weighted average mortality model is shown in Figure \ref{fig:wa_eventstudy_mort}.
-These estimates reveal that hospital teams with clinical experience do not differentially change mortality rates compared to non-clinical experience teams before or after the penalties begin.
-
-    \begin{figure}[ht!]
-        \caption{Event Study Estimates, Mortality}
-        \includegraphics[scale=.5]{Objects/mort_es_graph.pdf}
-        \label{fig:wa_eventstudy_mort}
-    \end{figure}
 

### 2024-04-30 Paper/paper.tex
Updates from Overleaf

@@
 
-    Before presenting the main results of the paper, I investigate whether executive team changes are occurring endogenously due to the programs.
-If hospitals change their propensity to hire a clinically trained executive because they expect to be penalized or want to avoid penalties, then there is selection in the independent variable.
-First, I analyze whether changes in clinically trained executives are concentrated around the program enactment, 2011-2012.
-Second, I analyze whether hospitals who end up getting penalized are more likely to change clinically trained executives on their team than non-penalized hospitals.
-That is, are hospitals that expect to be more affected by the initial penalties more likely to hire or fire clinically trained executives?
+    Before presenting the main findings, I investigate whether executive team changes occur endogenously due to the programs.
+If hospitals change their propensity to hire a clinically trained executive because they expect to be penalized or want to gain extra payments, then there is selection in the independent variable.
+First, I analyze whether changes in clinically trained executives are concentrated around the enactment of both HRRP and HVBP, 2011-2012.
+Second, I analyze whether hospitals who end up getting penalized through HRRP are more likely to change clinically trained executives on their team than non-penalized hospitals, and similarly for hospitals who end up getting extra payments for high quality of care through HVBP.
+That is, are hospitals that expect to be more affected by the programs more likely to hire or fire clinically trained executives?
 I consider changes along having any clinically trained executive as well as changes in the number of clinically trained executives.
@@
 
+    \import{Tables}{MD_noMD_readmort_decomp_synth.tex}
+
     \subsection{Uncompensated Care and Case Mix Index}
@@
     \import{Tables}{MD_noMD_uncompCMI_synth.tex}
+
+    \subsubsection{Decomposition}
+
+    \import{Tables}{MD_noMD_uncompCMI_decomp_synth.tex}
 

### 2024-04-30 Tables/MD_noMD_readmort_decomp_synth.tex
Updates from Overleaf

@@
  &  & (0.09) &  & (0.09)\\
- &  &  &  \vphantom{1} & \\
+ &  &  & & \\
 \addlinespace

### 2024-05-01 Tables/MD_noMD_readmort_fullsample.tex
Add Table files
make tables/figures look better and more consistent

@@
-\begin{table}[htbp]
-   \caption{\label{tab:MD_noMD_fullsample} Comparison of MD vs. non-MD Nonprofit Readmissions and Mortality Regression Results}
+\begin{table}[ht!]
+   \caption{\label{tab:MD_noMD_readmort_fullsample} Non-MD vs. MD Executive Readmission and Mortality Regression Results}
    \bigskip
@@
       \midrule 
-      NP w/ MD Executive x Post Penalty & -0.40$^{***}$                  & -0.19\\   
+      NP w/out MD Exec x Post Programs & -0.40$^{***}$                  & -0.19\\   
                                        & (0.14)                         & (0.15)\\   
       \textbf{Comparison Group:}       &                                & \\  
-      NP w/out MD Executive             & $\checkmark$                   & $\checkmark$\\   
+      NP w/ MD Exec                    & $\checkmark$                   & $\checkmark$\\   
        \\
@@
    \end{tabular}
+
+   \par \raggedright 
+   Results from Equation \ref{eq:clinical} with readmission and mortality rates as outcome variables.\\
+   Standard errors are clustered at the hospital level.\\
+   Signif. codes: 0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1
 \end{table}

### 2024-05-01 Tables/MD_noMD_readmort_synth.tex
add tables and update descriptive graph to include for-profits
make tables/figures look better and more consistent

--- 2024-04-23 Tables/MD_noMD_readmort_synthdid.tex
@@
 
-\caption{Comparison of MD vs. non-MD Nonprofit Readmissions and Mortality Synthetic DiD Results}
+\caption{\label{tab:MD_noMD_readmort_synth}Non-MD vs. MD Executive Readmission and Mortality Synthetic DiD Results}
 \centering
@@
 \midrule
-NP w/ MD Executive x Post Penalty & -0.28$^{*}$ & -0.19\\
- & (0.12) & (0.11)\\
+NP w/out MD Exec x Post Program & -0.28$^{*}$ & -0.19\\
+ & (0.13) & (0.11)\\
 \textbf{Comparison Group:} &  & \\
-NP w/out MD Executive & $\checkmark$ & $\checkmark$\\
-\\
+NP w/ MD Exec & $\checkmark$ & $\checkmark$\\
+ &  & \\
+\addlinespace
 Observations & 2,335 & 2,325\\
 \bottomrule
+\multicolumn{3}{l}{\textsuperscript{} Results from Equation with readmission and mortality rates as}\\
+\multicolumn{3}{l}{outcome variables.}\\
+\multicolumn{3}{l}{\textsuperscript{} Standard errors are clustered at the hospital level.}\\
+\multicolumn{3}{l}{\textsuperscript{} Signif. codes: 0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1}\\
 \end{tabular}

### 2024-05-01 Tables/MD_noMD_uncompCMI_fullsample.tex
Add Table files
update tables
make tables/figures look better and more consistent

@@
-\begin{table}[htbp]
-   \caption{\label{tab:MD_noMD_uncomp_CMI_fullsample} Comparison of MD vs. non-MD Nonprofit Uncompensated Care and CMI Regression Results}
+\begin{table}[ht!]
+   \caption{\label{tab:MD_noMD_uncompCMI_fullsample} Non-MD vs. MD Executive Uncompensated Care and CMI Regression Results}
    \bigskip
@@
       \toprule
-                                        & Uncomp. Care (Millions) & Case Mix Index\\  
+                                       & Uncompensated Care & Case Mix Index\\  
                                        & (1)                & (2)\\  
       \midrule 
-      NP w/ MD Executive x Post Penalty & -4.46                   & 0.03$^{***}$\\   
-                                        & (10.39)                 & (0.01)\\   
+      NP w/out MD Exec x Post Programs & 0.01               & 0.03$^{***}$\\   
+                                       & (0.01)             & (0.01)\\   
       \textbf{Comparison Group:}       &                    & \\  
-      NP w/out MD Executive             & $\checkmark$            & $\checkmark$\\   
+      NP w/ MD Executive               & $\checkmark$       & $\checkmark$\\   
        \\
@@
    \end{tabular}
+
+   \par \raggedright 
+   Results from Equation \ref{eq:forprofit} with readmission and mortality rates as outcome variables.\\
+   Standard errors are clustered at the hospital level.\\
+   Signif. codes: 0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1\\
+   Uncompensated care measured as percent of total operating expenses.
 \end{table}

### 2024-05-01 Tables/MD_noMD_uncompCMI_synth.tex
Add Table files
make tables/figures look better and more consistent

@@
 
-\caption{\label{tab:MD_noMD_uncompCMI_synth}Comparison of MD vs. non-MD Nonprofit Uncompensated Care and CMI Synthetic DiD Results}
+\caption{\label{tab:MD_noMD_uncompCMI_synth}Non-MD vs. MD Executive Uncompensated Care and CMI Synthetic DiD Results}
 \centering
@@
 \toprule
-\multicolumn{1}{c}{ } & \multicolumn{1}{c}{Uncomp. Care (Millions)} & \multicolumn{1}{c}{Case Mix Index} \\
-\cmidrule(l{3pt}r{3pt}){2-2} \cmidrule(l{3pt}r{3pt}){3-3}
+\multicolumn{1}{c}{ } & \multicolumn{1}{c}{Uncompensated Care} & \multicolumn{1}{c}{Case Mix Index} \\
  & (1) & (2)\\
 \midrule
-NP w/ MD Executive x Post Penalty & 3.12 & 0.03$^{**}$\\
- & (3.04) & (0.01)\\
+NP w/out MD Exec x Post Programs & 0.02$^{*}$ & 0.03$^{***}$\\
+ & (0.01) & (0.01)\\
 \textbf{Comparison Group:} &  & \\
-NP w/out MD Executive & $\checkmark$ & $\checkmark$\\
-Observations & 625 & 1635\\
+NP w/ MD Exec & $\checkmark$ & $\checkmark$\\
+ &  & \\
+\addlinespace
+Observations & 15,585 & 2,900\\
 \bottomrule
+\multicolumn{3}{l}{\textsuperscript{} Results from Equation with readmission and mortality rates as outcome}\\
+\multicolumn{3}{l}{variables.}\\
+\multicolumn{3}{l}{\textsuperscript{} Standard errors are clustered at the hospital level.}\\
+\multicolumn{3}{l}{\textsuperscript{} Signif. codes: 0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1}\\
+\multicolumn{3}{l}{\textsuperscript{} Uncompensated care measured as percent of total operating expenses.}\\
 \end{tabular}

### 2024-05-01 Tables/forprofit_readmort_synth.tex
add tables and update descriptive graph to include for-profits
make tables/figures look better and more consistent

--- 2024-04-23 Tables/forprofit_uncompCMI_synth.tex
@@
 
-\caption{\label{tab:forprofit_uncompCMI_synth} Comparison to For-Profit Uncompensated Care and CMI Synthetic DiD Results}
+\caption{\label{tab:forprofit_readmort_synth}Comparison to For-Profit Readmission and Mortality Synthetic DiD Results}
 \centering
-\begin{tabular}[t]{lcccccc}
+\begin{tabular}[t]{l>{\centering\arraybackslash}p{.2textwidth}>{\centering\arraybackslash}p{.2textwidth}>{\centering\arraybackslash}p{.2textwidth}>{\centering\arraybackslash}p{.2textwidth}>{\centering\arraybackslash}p{.2textwidth}>{\centering\arraybackslash}p{.2textwidth}}
 \toprule
-\multicolumn{1}{c}{ } & \multicolumn{3}{c}{Uncomp. Care (Millions)} & \multicolumn{3}{c}{Case Mix Index} \\
+\multicolumn{1}{c}{ } & \multicolumn{3}{c}{Weighted Avg. Readmission Rate} & \multicolumn{3}{c}{Weighted Avg. Mortality Rate} \\
  & (1) & (2) & (3) & (1) & (2) & (3)\\
 \midrule
-For Profit x Post Penalty & 10.53$^{***}$ & 11.69$^{***}$ & 7.6$^{**}$ & 0 & 0.03$^{**}$ & 0\\
- & (2.36) & (2.33) & (2.94) & (0.01) & (0.01) & (0.01)\\
+For Profit x Post Programs & -0.07 & -0.32$^{*}$ & -0.03 & -0.01 & -0.17 & 0.02\\
+ & (0.04) & (0.12) & (0.06) & (0.04) & (0.11) & (0.06)\\
 \textbf{Comparison Group:} &  &  &  &  &  & \\
@@
 Nonprofits w/out MD &  &  & $\checkmark$ &  &  & $\checkmark$\\
-Observations & 4945 & 1330 & 1815 & 12345 & 2895 & 4080\\
+ &  &  &  &  &  & \\
+Observations & 15,585 & 2,900 & 4,705 & 15,505 & 2,895 & 4,690\\
 \bottomrule
+\multicolumn{7}{l}{\textsuperscript{} Results from Equation with readmission and mortality rates as outcome variables.}\\
+\multicolumn{7}{l}{\textsuperscript{} Standard errors are clustered at the hospital level.}\\
+\multicolumn{7}{l}{\textsuperscript{} Signif. codes: 0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1}\\
 \end{tabular}

### 2024-05-01 Tables/forprofit_uncompCMI_synth.tex
Add Table files
make tables/figures look better and more consistent

@@
 
-\caption{\label{tab:forprofit_uncompCMI_synth} Comparison to For-Profit Uncompensated Care and CMI Synthetic DiD Results}
+\caption{\label{tab:forprofit_uncompCMI_synth}Comparison to For-Profit Uncompensated Care and CMI Synthetic DiD Results}
 \centering
-\begin{tabular}[t]{lcccccc}
+\begin{tabular}[t]{l>{\centering\arraybackslash}p{.2textwidth}>{\centering\arraybackslash}p{.2textwidth}>{\centering\arraybackslash}p{.2textwidth}>{\centering\arraybackslash}p{.2textwidth}>{\centering\arraybackslash}p{.2textwidth}>{\centering\arraybackslash}p{.2textwidth}}
 \toprule
-\multicolumn{1}{c}{ } & \multicolumn{3}{c}{Uncomp. Care (Millions)} & \multicolumn{3}{c}{Case Mix Index} \\
+\multicolumn{1}{c}{ } & \multicolumn{3}{c}{Uncompensated Care} & \multicolumn{3}{c}{Case Mix Index} \\
+\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-7}
  & (1) & (2) & (3) & (1) & (2) & (3)\\
 \midrule
-For Profit x Post Penalty & 10.53$^{***}$ & 11.69$^{***}$ & 7.6$^{**}$ & 0 & 0.03$^{**}$ & 0\\
- & (2.36) & (2.33) & (2.94) & (0.01) & (0.01) & (0.01)\\
+For-Profit x Post Programs & 0.06$^{***}$ & 0.08$^{***}$ & 0.06$^{***}$ & 0 & 0.02$^{**}$ & 0\\
+ & (0.01) & (0.01) & (0.01) & (0.01) & (0.01) & (0.01)\\
 \textbf{Comparison Group:} &  &  &  &  &  & \\
@@
 Nonprofits w/out MD &  &  & $\checkmark$ &  &  & $\checkmark$\\
-Observations & 4945 & 1330 & 1815 & 12345 & 2895 & 4080\\
+ &  &  &  &  &  & \\
+Observations & 4935 & 1330 & 1815 & 12320 & 2880 & 4065\\
 \bottomrule
+\multicolumn{7}{l}{\textsuperscript{} Results from Equation with readmission and mortality rates as outcome variables.}\\
+\multicolumn{7}{l}{\textsuperscript{} Standard errors are clustered at the hospital level.}\\
+\multicolumn{7}{l}{\textsuperscript{} Signif. codes: 0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1}\\
+\multicolumn{7}{l}{\textsuperscript{} Uncompensated care measured as percent of total operating expenses.}\\
 \end{tabular}

### 2024-05-01 Paper/paper.tex
Updates from Overleaf

@@
 
-\captionsetup{justification=raggedright,singlelinecheck=false}
+\captionsetup{justification=center,singlelinecheck=false}
 
@@
 
-	\linespread{1.2}\title{\vspace{-0.5in} Does Hospital Leadership Matter?\\ \large Evidence from Pay-for-Performance Programs}
+	\linespread{1.2}\title{\vspace{-0.5in} Does Hospital Leadership Matter?\\ \large Evidence from Pay-for-Performance Incentives}
 
@@
 
-    First, I test empirically whether nonprofits respond to program incentives differently from for-profits.
+    First, I test empirically whether different sub-samples of nonprofits respond to program incentives differently from for-profits.
 While nonprofit hospital objectives are largely unknown, for-profits have revealed their objective in their ownership classification.
 That is, $\alpha=1$.
-To learn something about $\alpha$ for nonprofit firms with different leadership compositions, I compare their program response to that of for-profits.
+To learn something about $\alpha$ for nonprofit firms with different executive team compositions, I compare their program response to that of for-profits.
 I estimate the following regression equation:
@@
     \label{eq:forprofit}
-    y_{ht} = \beta (for\_profit_h \times post\_penalty_t) + \alpha_{h} + \delta_t + \epsilon_{ht}
+    y_{ht} = \beta (for\_profit_h \times post\_programs_t) + \alpha_{h} + \delta_t + \epsilon_{ht}
     \end{equation}
@@
 For each outcome, I estimate three models where the comparison group differs by hospital type: all nonprofits, nonprofits with a clinically trained executive, and nonprofits without a clinically trained executive.
-
-    An important distinction here is whether hospital time is a time-invariant characterization.
-For nonprofits and for-profits, switching is relatively rare, and I limit to those who never switch.
-However, hospital executive teams are dynamic, and whether or not a hospital employs a clinically trained executive in each period can change.
-Under the assumption that these changes are not correlated with the program enactments (see Section \ref{sec:endog}), these changes can be advantageous in understanding hospital objectives.
-There are two ways that having a clinically trained executive might affect outcomes: first, a hospital that hires a clinically trained executive is signalling their underlying $\alpha$.
-Second, hiring a clinically trained executive who carries out day-to-day operations in a specific way actually changes the underlying $\alpha$ of the hospital.
-Thus, I use two types of variation in clinically trained executives to disentangle the two: a variable for whether the hospital ever hires a clinically trained executive, and a variable for whether the hospital has a clinically trained executive in 2012 when the programs took place.
-
-    Similarly, I test empirically whether having a clinically trained executive affects hospital response to program incentives.
-Thus, I limit the sample to only hospitals for which I have leadership team information, and estimate:
+I also directly estimate the difference in response between hospital with and without a clinically trained executive:
 
@@
     \label{eq:clinical}
-    y_{ht} = \beta (clinical_h \times post\_penalty_t) + \alpha_{h} + \delta_t + \epsilon_{ht}.
+    y_{ht} = \beta (no\_MD\_exec_h \times post\_programs_t) + \alpha_{h} + \delta_t + \epsilon_{ht}
     \end{equation}
+
+    An important distinction here is whether the sub-categorization of nonprofit hospitals by executive team is time-invariant.
+For nonprofits and for-profits, switching overall ownership type is rare, and I limit to those who never switch.
+However, hospital executive teams are dynamic, and whether or not a hospital employs a clinically trained executive in each period may change.
+Therefore, I first present effects of having a clinically trained executive among stable executive teams.
+That is, among hospitals that don't change their propensity to hire a clinically trained executive over time.
+
+    However, under the assumption that these changes are not correlated with the program enactments (see Section \ref{sec:endog}), using the changes can actually be advantageous in better understanding underlying hospital motives.
+There are two reasons we might see having a clinically trained executive affecting outcomes: first, a clinically trained executive is a signal of the underlying weight placed on profit, $\alpha$.
+Second, hiring a clinically trained executive who carries out day-to-day operations in a specific way actually changes the underlying $\alpha$.
+Thus, I use two types of variation in clinically trained executives to disentangle the two: a variable for whether the hospital ever hires a clinically trained executive, and a variable for whether the hospital has a clinically trained executive in 2012 when the programs took place.
+I use the same empirical specification with these two variables as the independent variation in executive teams.
 
@@
 Second, I analyze whether hospitals who end up getting penalized through HRRP are more likely to change clinically trained executives on their team than non-penalized hospitals, and similarly for hospitals who end up getting extra payments for high quality of care through HVBP.
-That is, are hospitals that expect to be more affected by the programs more likely to hire or fire clinically trained executives?
+That is, are hospitals that expect to be affected by the programs more likely to hire or fire clinically trained executives?
 I consider changes along having any clinically trained executive as well as changes in the number of clinically trained executives.
@@
     \begin{equation}
-    change_{ht} = \beta(ever\_penalized_{h} \times post\_2011)+ \alpha_h + \delta_t + \epsilon_{ht}.
+    change_{ht} = \beta(program\_exposed_{h} \times post\_2011)+ \alpha_h + \delta_t + \epsilon_{ht}.
     \end{equation}
 
+    The variable $program\_exposed_{h}$ is an indicator for whether the hospital eventually was penalized under HRRP or received payments under HVBP.
 The estimates from this analysis are presented in Table \ref{tab:change_analysis}.
-Changes in any clinically trained executive are less likely to occur in 2012-2014 relative to 2010, indicating that executives are actually becoming more stable after the programs.
-Further, hospitals who end up being penalized are not more likely to change their propensity to hire any clinically trained executive.
+Changes in any clinically trained executive are less likely to occur in 2012-2014 relative to 2010, indicating that executives are actually becoming more stable over time along the dimension of hiring clinically trained executives.
+Further, hospitals who end up being penalized through HRRP or receiving payments through HVBP are not more likely to change their propensity to hire any clinically trained executive.
 Results are similar when considering changes in the number of clinically trained executives.
-Thus, I conclude that changes in executive teams are not endogenously occurring with the change in incentives due to the programs.
+Thus, I conclude that changes in executive teams are likely not endogenously occurring with the incentive changes due to the programs.
 
@@
 
-     First, I present results for the differential program response of for-profit firms compared to the different sub-samples of nonprofit firms in Table \ref{tab:forprofit_synth}.
-Looking first at the left panel, for-profits change their readmission rates after the program similarly to all nonprofits taken together, as well as nonprofits without clinically trained executives.
-However, for-profits have a .3ppt larger decrease in readmissions than nonprofits with clinically trained executives.
-When considering the right panel, there is no statistical difference in change in mortality rates between different hospital types.
+     First, I consider the differential program response of for-profit firms compared to the different sub-samples of nonprofit firms.
+These results are presented in Table \ref{tab:forprofit_synth}.
+Models in the left panel estimate the effect of being for-profit after the programs were enacted, compared to nonprofit sub-samples, on readmission rates.
+There is no statistically meaningful difference between the response of for-profits and nonprofits as a whole, or nonprofits without a clinically trained executive.
+However, for-profits lower their readmissions by .3ppts compared to nonprofits without clinically trained executives.
+Models in the right panel estimate the effect of being for-profit after the programs on mortality rates.
+Here, there is no statistical difference in mortality rate response between different hospital types.
 
@@
      Next, I directly estimate the difference in response by nonprofits with and without clinically trained executives.
-These results are presented in Table \ref{tab:MD_noMD_synth}.
+These results are presented in Table \ref{tab:MD_noMD_synth}, where the left has models with outcome readmission rates and the right has models with outcome mortality rates.
 Nonprofits without clinically trained executives have a larger decrease in readmissions than those with clinically trained executive by .3ppts, while there is no difference in change in mortality rates.
-
-     These findings are consistent with the theoretical prediction presented in Section \ref{sec:model} if having a clinically trained executive means the hospital places more weight on societal benefit than for-profits and other nonprofits do.
+Tables \ref{tab:forprofit_synth} and \ref{tab:MD_noMD_synth} taken together show that, in terms of readmission rates, nonprofits without clinically trained executives behave more similarly to for-profits than they do other nonprofits.
+Under the assumption that having a clinically trained executive on the leadership team indicates a higher preference for social utility, these findings are consistent with the theoretical prediction presented in Section \ref{sec:model}.
 
@@
 
+    However, it is not clear whether these results are due to clinically trained executives being a signal of underlying hospital objectives, or that the leaders themselves make a difference.
+Therefore, I present a decomposition of the results using two models in Table \ref{tab:MD_noMD_readmort_decomp_synth}, where the variation in executives is either whether the hospital ever has a clinically trained executive, or whether the hospital has a clinically trained executive in 2012 when the programs were enacted.
+The left panel on readmission rates shows that the difference is response is fully driven by having a clinically trained executive in 2012 compared to having a clinically trained trained executive at another point in time.
+Again, there is nothing going on with mortality.
+
     \import{Tables}{MD_noMD_readmort_decomp_synth.tex}
@@
 
-    As a secondary analysis, I present estimates of differential response of potential mechanisms, changing input into another societal benefit measure: uncompensated care, and the complexity of patients seen in the hospital: case mix index.
+    As a secondary analysis, I present estimates of differential response of potential mechanisms.
+The first is changing input into another societal benefit measure: uncompensated care, and the second is complexity of patients: case mix index.
 First, I investigate whether for-profits change these differentially than types of nonprofits in Table \ref{tab:forprofit_uncompCMI_synth}.

### 2024-05-01 Tables/MD_noMD_readmort_synth.tex
Updates from Overleaf

--- 2024-04-23 Tables/MD_noMD_readmort_synthdid.tex
@@
 
-\caption{Comparison of MD vs. non-MD Nonprofit Readmissions and Mortality Synthetic DiD Results}
+\caption{\label{tab:MD_noMD_synth}Comparison of MD vs. non-MD Nonprofit Readmissions and Mortality Synthetic DiD Results}
 \centering
@@
 \multicolumn{1}{c}{ } & \multicolumn{1}{c}{Weighted Avg. Readmission Rate} & \multicolumn{1}{c}{Weighted Avg. Mortality Rate} \\
+\cmidrule(l{3pt}r{3pt}){2-2} \cmidrule(l{3pt}r{3pt}){3-3}
  & (1) & (2)\\
 \midrule
-NP w/ MD Executive x Post Penalty & -0.28$^{*}$ & -0.19\\
+NP w/out MD Exec x Post Penalty & -0.28$^{*}$ & -0.19\\
  & (0.12) & (0.11)\\
 \textbf{Comparison Group:} &  & \\
-NP w/out MD Executive & $\checkmark$ & $\checkmark$\\
-\\
-Observations & 2,335 & 2,325\\
+NP w/ MD Executive & $\checkmark$ & $\checkmark$\\
+Observations & 2335 & 2325\\
 \bottomrule

### 2024-05-01 Tables/MD_noMD_readmort_decomp_fullsample.tex
Update tables AGAIN

@@
-\begin{table}[htbp]
-   \caption{\label{tab:MD_noMD_readmort_decomp_fullsample} Decomposition of Readmission and Mortality Results for Nonprofits}
+\begin{table}[ht!]
+   \caption{\label{tab:MD_noMD_readmort_decomp_fullsample} Decomposition of Readmission and Mortality Regression Results}
    \bigskip
@@
       \midrule 
-      Ever MD Executive x Post Penalty & -0.03         &               & -0.02         &   \\   
+      Ever MD Exec x Post Programs & -0.03         &               & -0.02         &   \\   
                                    & (0.07)        &               & (0.07)        &   \\   
-      MD Executive in 2012             &               & -0.27$^{***}$ &               & -0.11\\   
+      MD Exec in 2012              &               & -0.27$^{***}$ &               & -0.11\\   
                                    &               & (0.10)        &               & (0.10)\\   
@@
    \end{tabular}
+
+   \par \raggedright 
+   Results from Equation \ref{eq:decomp} with readmission and mortality rates as outcome variables.\\
+   Standard errors are clustered at the hospital level.\\
+   Signif. codes: 0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1
 \end{table}

### 2024-05-01 Tables/MD_noMD_readmort_decomp_synth.tex
Update tables AGAIN

@@
 
-\caption{\label{tab:MD_noMD_readmort_decomp_synth}Decomposition of Readmissions and Mortality for Nonprofits Synthetic DiD Results}
+\caption{\label{tab:MD_noMD_readmort_decomp_synth}Decomposition of Readmissions and Mortality Synthetic DiD Results}
 \centering
-\begin{tabular}[t]{lcccl}
+\begin{tabular}[t]{lcccc}
 \toprule
 \multicolumn{1}{c}{ } & \multicolumn{2}{c}{Weighted Avg. Readmission Rate} & \multicolumn{2}{c}{Weighted Avg. Mortality Rate} \\
-\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
  & (1) & (2) & (1) & (2)\\
 \midrule
-Ever MD Executive x Post 2012 & 0 &  & -0.06 & \\
- & (0.07) &  & (0.06) & \\
-2012 MD Executive x Post 2012 &  & -0.19$^{*}$ &  & -0.1\\
- &  & (0.09) &  & (0.09)\\
+Ever MD Exec x Post Programs & 0 &  & -0.06 & \\
+ & (0.06) &  & (0.06) & \\
+2012 MD Exec x Post Programs &  & -0.19$^{*}$ &  & -0.1\\
+ &  & (0.09) &  & (0.08)\\
  &  &  &  & \\
@@
 \bottomrule
+\multicolumn{5}{l}{\textsuperscript{} Results from Equation with readmission and mortality rates as}\\
+\multicolumn{5}{l}{outcome variables.}\\
+\multicolumn{5}{l}{\textsuperscript{} Standard errors are clustered at the hospital level.}\\
+\multicolumn{5}{l}{\textsuperscript{} Signif. codes: 0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1}\\
 \end{tabular}

### 2024-05-01 Tables/MD_noMD_uncompCMI_decomp_synth.tex
Update tables AGAIN

@@
 
-\caption{\label{tab:MD_noMD_uncompCMI_decomp_synth}Decomposition of Uncompensated Care and CMI for Nonprofits Synthetic DiD Results}
+\caption{\label{tab:MD_noMD_uncompCMI_decomp_synth}Decomposition of Uncompensated Care and CMI Synthetic DiD Results}
 \centering
-\begin{tabular}[t]{lcccl}
+\begin{tabular}[t]{lcccc}
 \toprule
 \multicolumn{1}{c}{ } & \multicolumn{2}{c}{Uncompensated Care} & \multicolumn{2}{c}{Case Mix Index} \\
-\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
  & (1) & (2) & (1) & (2)\\
 \midrule
-Ever MD Executive x Post 2012 & 0.01 &  & 0.01$^{*}$ & \\
+Ever MD Exec x Post 2012 & 0.01 &  & 0.01$^{*}$ & \\
  & (0.01) &  & (0) & \\
-2012 MD Executive x Post 2012 &  & -0.03 &  & 0.01\\
+2012 MD Exec x Post 2012 &  & -0.03 &  & 0.01\\
  &  & (0.03) &  & (0.01)\\
- &  &  &  \vphantom{1} & \\
+ &  &  &  & \\
 \addlinespace
@@
 \bottomrule
+\multicolumn{5}{l}{\textsuperscript{} Results from Equation with readmission and mortality rates as}\\
+\multicolumn{5}{l}{outcome variables.}\\
+\multicolumn{5}{l}{\textsuperscript{} Standard errors are clustered at the hospital level.}\\
+\multicolumn{5}{l}{\textsuperscript{} Signif. codes: 0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1}\\
+\multicolumn{5}{l}{\textsuperscript{} Uncompensated care measured as percent of total operating expenses.}\\
 \end{tabular}

### 2024-05-01 Paper/paper.tex

@@
 
-\captionsetup{justification=center,singlelinecheck=false}
+\captionsetup{justification=centering,singlelinecheck=false}
 
@@
 		{\small
-        In many industries, including hospitals, nonprofits make up the majority of firms in the market.
+        In many industries, nonprofits make up the majority of firms in the market.
 Understanding the objectives and behaviors of these firms is important when creating incentives.
-However, nonprofit firms have diverse goals and missions that make understanding their objectives difficult.
+However, nonprofit firms have goals and missions that make understanding their objectives difficult.
 For hospitals in particular, there is mixed evidence on whether nonprofit hospitals differ meaningfully from for-profit hospitals.
-In this paper, I investigate whether leadership team composition of nonprofit hospitals affects how similarly they behaves to for-profits.
-I use a novel data set on nonprofit hospital executives in the US from 2009-2015 to investigate whether the existence of clinical training on a hospital executive team affects hospital behavior.
-In particular, I leverage several pay-for-performance initiatives in the US in 2012, and estimate the difference in response between for-profits, nonprofits, and nonprofits with clinically trained executives.
+In this paper, I investigate whether leadership team composition of nonprofit hospitals affects how similarly they behave to for-profits.
+I use a novel data set on nonprofit hospital executives in the US from 2009-2015 to investigate whether the presence of clinically trained executives affects hospital behavior.
+In particular, I leverage two pay-for-performance initiatives in the US enacted in 2012, and estimate the difference in response between for-profits, nonprofits, and nonprofits with clinically trained executives.
 I find that while readmissions decreased on average for all types of hospitals, the response was larger for for-profits and nonprofits without clinically trained executives.
-This is consistent with theoretical predictions under the assumption that nonprofits with clinically trained executives put higher weight on societal outcomes than those without clinically trained executives.
+This is consistent with theoretical predictions under the assumption that nonprofits with clinically trained executives put greater weight on societal outcomes than those without clinically trained executives.
 		} 
@@
 Further, Chief Diversity Officers are not found to have any effect on hiring more diversely in universities (\cite{bradley2022impact}).
+
     Due to data limitations, many of these studies focus on large, publicly traded firms in the US.
-However, one study uses data from England to study whether CEOs affect hospital production, and find no association (\cite{janke2019impact}).
-\textcolor{red}{(add paper about Chile hospital executives)}.
-I contribute to this literature first by investigating a context which is not largely studied but is policy relevant, nonprofit US hospitals, and considering a unique executive characteristic pertinent to that context.
+However, two studies focus on hospital performance in other countries.
+\citeauthor{janke2019impact} (\citeyear{janke2019impact}) uses data from England to study whether CEOs affect hospital production, and find no association (\cite{janke2019impact}).
+However, \citeauthor{otero2022managers} (\citeyear{otero2022managers}) investigates the role of CEOs in public hospitals and Chile, and finds an 8\% decrease in mortality rates for hospitals with top managers (\cite{otero2022managers}).
+I contribute to this literature by looking at the executive team as a whole rather than just CEOs, investigating a context which is not largely studied but is policy relevant, nonprofit US hospitals, and considering a unique executive characteristic pertinent to that context.
 
     Finally, I contribute to our understanding of how providers respond to pay-for-performance incentives in health care.
-The most in depth study of how hospitals respond to HRRP is \citeauthor{gupta2021impacts} (\citeyear{gupta2021impacts}).
-The author finds that hospitals decreased readmissions by 5\% and mortality rates by 2\% on average as a result of the program, confirming prior studies (\cite{mellor2017does}; \cite{ziedan2018essays}; \cite{ody2019decreases}; \cite{gupta2021impacts}).
+The most in depth study of how hospitals respond to HRRP is \citeauthor{gupta2021impacts} (\citeyear{gupta2021impacts}), who finds that hospitals decreased readmissions by 5\% and mortality rates by 2\% on average as a result of the program, confirming prior studies (\cite{mellor2017does}; \cite{ziedan2018essays}; \cite{ody2019decreases}; \cite{gupta2021impacts}).
 Importantly, around 40\% of the decrease in readmissions is due to selective patient practices.
 This raises important questions about how to incentivize improvement in quality of care without also incentivizing gamesmanship.
-This paper addresses a necessary first step in considering whether leadership compositions could affect likelihood of gaming.
-\textcolor{red}{(add papers about other program too)}
+Research on HVBP generally finds that the program had no effect on underlying hospital quality (\cite{us2015hospital}; \cite{norton2018moneyball}; \cite{friedson2019so}).
+This paper contributes to our understanding of how characteristics of hospitals affect response to pay-for-performance incentives.
 
-    \section{Institutional Setting}
+    \section{Setting}
 
-    \subsection{Hospital Readmissions Reduction Program}\label{sec:hrrp}
+    \subsection{Nonprofit Hospital Leadership}
+
+    Nonprofits are governed by a board of directors, whose role is to set high level goals and strategies, and provide general oversight.
+The board is not in charge of day-to-day operations, but the broad direction of the firm.
+They select executive team members to carry out day-to-day operations.
+While hospital executives typically do specialize by getting degree(s) in healthcare management or an MBA specific to healthcare, they do not often have clinical training.
+While some doctors earn additional degrees before stepping into an executive role, this is not a necessary condition to becoming a physician executive.
+An executive team usually consists of at least a CEO and CFO, but every nonprofit is different in exactly how they structure their executive team.
+Some nonprofits hire Chief Medical Officers, Chief Quality Officers, executive directors, vice presidents of various departments, etc.
+I consider all of these employees as part of the day-to-day management of the hospital and include them as a part of the executive team.
+
+    All tax-exempt organizations in the US are required to file a Form 990 with Internal Revenue Services (IRS) each year.
+There are different types of forms, but any organization grossing over \$200,000 must file the most extensive Form 990.
+Sections of this form include a statement of revenue, statement of functional expenses, a balance sheet, and, as used in this project, a list of all key employees, executives, and board members.
+In this section, each firm is required to report the name and title, average hours per week, position, and compensation of their board members and executive leaders.
+Since these firms are tax exempt, their information in these forms is made publicly available.
+
+    \subsection{Pay-for-Performance Programs}\label{sec:hrrp}
 
@@
 
-    \subsection{Hospital Value-Based Purchasing Program}
-
-    \subsection{Nonprofit Hospital Leadership}
-
-    Nonprofits are governed by a board of directors, whose role is to set high level goals and strategies, and provide general oversight.
-The board is not in charge of day-to-day operations, but the general direction of the firm.
-They select members of the executive team to carry out day-to-day operations.
-While hospital executives typically do specialize by getting degree(s) in healthcare management or an MBA specific to healthcare, they do not often have clinical training.
-While some doctors earn additional degrees before stepping into an executive role, this is not a necessary condition to becoming a physician executive.
-An executive team usually consists of at least a CEO and CFO, but every nonprofit is different in exactly how they structure their executive team.
-Some nonprofits hire Chief Medical Officers, Chief Quality Officers, executive directors, vice presidents of various departments, etc.
-I consider all of these employees as part of the day-to-day management of the hospital and include them as a part of the executive team.
-
-    All tax-exempt organizations in the US are required to file a Form 990 with Internal Revenue Services (IRS) each year.
-There are different types of forms, but any organization grossing over \$200,000 must file the most extensive Form 990.
-Sections of this form include a statement of revenue, statement of functional expenses, a balance sheet, and, as used in this project, a list of all key employees, executives, and board members.
-In this section, each firm is required to report the name and title, average hours per week, position, and compensation of their board members and executive leaders.
-Since these firms are tax exempt, their information in these forms is made publicly available.
-
 	\section{Data}\label{sec:data}
@@
 
-    To characterize hospitals as having clinically trained executives or not, I construct a novel data set of nonprofit hospitals in the US that contains names and select characteristics of executives tied to the hospital from 2009-2015.
-I gather this data from each hospital's publicly available Tax Form 990, which every sufficiently large nonprofit files each year in the US.
-To my knowledge, this is the first large scale gathering of leadership names from these forms.
+    To characterize hospitals as having clinically trained executives or not, I construct a novel data set of nonprofit hospitals in the US that contains names and select characteristics of all executives tied to the hospital from 2009-2015.
+I gather this data from publicly available Tax Form 990s, which every sufficiently large nonprofit files each year in the US.
+To my knowledge, this is the first large scale gathering of executive names from these forms.
 
@@
 After filtering out associations and specialty hospitals, there are around 3,000 firms left.
-Importantly, I need to be able to link hospitals in the tax forms to other hospital characteristics from publicly availble sources.
+Importantly, I need to be able to link hospitals in the tax forms to other hospital characteristics from publicly available sources.
 No official crosswalk between EINs and AHA IDs or Medicare numbers exists, so I must rely on matching by location and name of hospital in both data sets.
@@
 
+    \begin{equation}
+    \label{eq:decomp}
+    y_{ht} = \beta (no\_MD\_exec_h \times post\_programs_t) + \alpha_{h} + \delta_t + \epsilon_{ht}
+    \end{equation}
+
     As discussed in Section \ref{sec:hrrp}, hospitals had incentive to decrease readmissions immediately when the rules were released in October of 2011, before the penalties were actually determined in October of 2012.
@@
 
-    \begin{equation}
+    \begin{equation}\label{eq:change1}
     change_{ht} = \sum_{j=2011}^{2014}\beta_j\mathbf{1}\{t=j\} + \alpha_h + \epsilon_{ht},
@@
 
-    \begin{equation}
+    \begin{equation}\label{eq:change2}
     change_{ht} = \beta(program\_exposed_{h} \times post\_2011)+ \alpha_h + \delta_t + \epsilon_{ht}.

### 2024-05-01 Tables/forprofit_readmort_synth.tex
Update tables AGAIN

@@
 \centering
-\begin{tabular}[t]{l>{\centering\arraybackslash}p{.2textwidth}>{\centering\arraybackslash}p{.2textwidth}>{\centering\arraybackslash}p{.2textwidth}>{\centering\arraybackslash}p{.2textwidth}>{\centering\arraybackslash}p{.2textwidth}>{\centering\arraybackslash}p{.2textwidth}}
+\begin{tabular}[t]{lcccccc}
 \toprule
@@
 \midrule
-For Profit x Post Programs & -0.07 & -0.32$^{*}$ & -0.03 & -0.01 & -0.17 & 0.02\\
- & (0.04) & (0.12) & (0.06) & (0.04) & (0.11) & (0.06)\\
+For-Profit x Post Programs & -0.07 & -0.32$^{**}$ & -0.03 & -0.01 & -0.17 & 0.02\\
+ & (0.04) & (0.11) & (0.06) & (0.04) & (0.11) & (0.06)\\
 \textbf{Comparison Group:} &  &  &  &  &  & \\
@@
 Nonprofits w/ MD &  & $\checkmark$ &  &  & $\checkmark$ & \\
-\addlinespace
 Nonprofits w/out MD &  &  & $\checkmark$ &  &  & $\checkmark$\\

### 2024-05-01 Paper/paper.tex
Updates from Overleaf

@@
 
+    \subsection{Leadership Correlation with Other Hospital Characteristics}
+
     Finally, I present a table of correlations between having at least one clinically trained executive and other time invariant hospital characteristics.
@@
     \import{Tables}{hosp_correlations.tex}
-
-    \section{Model and Empirical Strategy}\label{sec:model}
-
-    \subsection{First Period}
-
-    Not knowing the true form of hospital objective functions, we can consider them maximizing some weighted average of profit and societal (or patient) benefit.
-Abstracting away from quantity and price as functions of quality for the purpose of this paper, I present a two-stage model of firm behavior that captures the change of quality due to pay-for-performance incentives.
-In the first period, prior to program enactment, hospitals do not expect any penalties in the future, and choose quality according to:
-
-    $$\max_{\theta}\hspace{2mm}\alpha\left[R - c_{\theta}(\theta) \right] + (1-\alpha) u(\theta),$$
-
-    \noindent where $R$ is net revenue, $c_{\theta}(.)$ is an increasing cost function specific to quality $\theta$, $u(.)$ is extra utility, which is increasing and concave is $\theta$, and $\alpha\in[0,1]$ is a hospital's type, indicating how much they care about profit vs.
-societal benefit.
-There is an implicit stay open condition throughout the model.
-Taking the first order condition yields
-
-    $$(1-\alpha)u'(\theta) = \alpha c_{\theta}'(\theta),$$
-
-    \noindent marginal benefit of increasing quality equals marginal cost. Solving for $u'(\theta)$ and differentiating with respect to $\alpha$ yields
-
-    $$\frac{du'(\theta)}{d\alpha} = c_{\theta)}'(\theta)\frac{1}{(1-\alpha)^2} > 0.$$
-
-    Thus, by the Implicit Function Theorem,
-
-    $$\frac{d\theta}{d\alpha} = \frac{du'(\theta)/d\alpha}{du'(\theta)/d\theta} < 0.$$
-
-    That is, the more weight the hospital places on profit, the lower quality of care they will provide.
-
-    \subsection{Second Period}
-
-    Hospitals determine period 1 quality $\theta_1$ according to $\theta(\alpha)$ described above.
-Then, in period two, hospitals are penalized according to $\theta_1$, and they choose a new quality, $\theta_2$, which will determine penalties in the future.
-The penalty $(k)$ comes in the form of a negative shock to revenue for hospitals with quality below a certain threshold.
-That is,
-
-    $$R_2 = \begin{cases}
-        R - k, \text{  if  } \theta_1<\overline{\theta}\\
-        R, \text{  if  } \theta_1\geq \overline{\theta}.
-    \end{cases}$$
-
-    The penalty $k$ is large and exceeds the maximum cost of quality.
-The penalty threshold, $\overline{\theta}$, is exogenous.
-Since $k>max\{c_{\theta}(\theta)\}$, all hospitals will choose $\theta_2$ to not be penalized in the future.
-Thus,
-
-    $$\theta_2 = \begin{cases}
-        \overline{\theta}, \text{  if  } \theta_1<\overline{\theta}\\
-        \theta_1, \text{  if  } \theta_1\geq \overline{\theta}.
-    \end{cases}$$
-
-    Taking the change in theta from period 1 to period 2,
-
-    $$\Delta\theta = \theta_2-\theta_1 = \begin{cases}
-        \overline{\theta}-\theta_1, \text{  if  } \theta_1<\overline{\theta}\\
-        0, \text{  if  } \theta_1\geq \overline{\theta}.
-    \end{cases}$$
-
-    Differentiating with respect to $\alpha$, we get
-
-    $$\frac{d\Delta\theta}{d\alpha} = \begin{cases}
-        -\frac{d\theta_1}{d\alpha}, \text{  if  } \theta_1<\overline{\theta}\\
-        0, \text{  if  } \theta_1\geq \overline{\theta}.
-    \end{cases}$$
-
-    I show in period one that $\frac{d\theta}{d\alpha}<0$.
-Therefore, $\frac{d\Delta\theta}{d\alpha}>=0$.
-That is, the response to the program incentives is weakly increasing in $\alpha$, and there is a threshold for which increasing $\alpha$ (a hospital with more weight on profit), increases the response to the penalty.
-
-    The theoretical prediction is that change in quality depends on how much weight the hospital places on profit vs.
-societal benefit.
-However, $\alpha$ for each hospital is unobserved.
-Thus, I empirically investigate whether the response to the program is correlated with hospital ownership or clinical training on the executive team in order to gain understanding about $\alpha$.
-
-    \subsection{Estimation}
-
-    First, I test empirically whether different sub-samples of nonprofits respond to program incentives differently from for-profits.
-While nonprofit hospital objectives are largely unknown, for-profits have revealed their objective in their ownership classification.
-That is, $\alpha=1$.
-To learn something about $\alpha$ for nonprofit firms with different executive team compositions, I compare their program response to that of for-profits.
-I estimate the following regression equation:
-
-    \begin{equation}
-    \label{eq:forprofit}
-    y_{ht} = \beta (for\_profit_h \times post\_programs_t) + \alpha_{h} + \delta_t + \epsilon_{ht}
-    \end{equation}
-
-    \noindent where $y_{ht}$ is one of the outcome variables discussed in Section \ref{sec:data}, $\beta$ is the coefficient of interest capturing the effect of being for-profit after the penalties took place, and $\alpha_h$ and $\delta_t$ are hospital and time fixed effects, respectively.
-For each outcome, I estimate three models where the comparison group differs by hospital type: all nonprofits, nonprofits with a clinically trained executive, and nonprofits without a clinically trained executive.
-I also directly estimate the difference in response between hospital with and without a clinically trained executive:
-
-    \begin{equation}
-    \label{eq:clinical}
-    y_{ht} = \beta (no\_MD\_exec_h \times post\_programs_t) + \alpha_{h} + \delta_t + \epsilon_{ht}
-    \end{equation}
-
-    An important distinction here is whether the sub-categorization of nonprofit hospitals by executive team is time-invariant.
-For nonprofits and for-profits, switching overall ownership type is rare, and I limit to those who never switch.
-However, hospital executive teams are dynamic, and whether or not a hospital employs a clinically trained executive in each period may change.
-Therefore, I first present effects of having a clinically trained executive among stable executive teams.
-That is, among hospitals that don't change their propensity to hire a clinically trained executive over time.
-
-    However, under the assumption that these changes are not correlated with the program enactments (see Section \ref{sec:endog}), using the changes can actually be advantageous in better understanding underlying hospital motives.
-There are two reasons we might see having a clinically trained executive affecting outcomes: first, a clinically trained executive is a signal of the underlying weight placed on profit, $\alpha$.
-Second, hiring a clinically trained executive who carries out day-to-day operations in a specific way actually changes the underlying $\alpha$.
-Thus, I use two types of variation in clinically trained executives to disentangle the two: a variable for whether the hospital ever hires a clinically trained executive, and a variable for whether the hospital has a clinically trained executive in 2012 when the programs took place.
-I use the same empirical specification with these two variables as the independent variation in executive teams.
-
-    \begin{equation}
-    \label{eq:decomp}
-    y_{ht} = \beta (no\_MD\_exec_h \times post\_programs_t) + \alpha_{h} + \delta_t + \epsilon_{ht}
-    \end{equation}
-
-    As discussed in Section \ref{sec:hrrp}, hospitals had incentive to decrease readmissions immediately when the rules were released in October of 2011, before the penalties were actually determined in October of 2012.
-However, I still define $post\_penalty$ as occurring in 2012 or later, the first full of hospital response.
-
-    To identify the causal effect of hospital type on hospital program response, I rely on several assumptions.
-First, I assume that executive teams are not determined endogenously when the programs were enacted.
-I investigate the validity of this assumption in Section \ref{sec:endog}.
-Second, I assume that no other unobserved changes occurred in 2012 that differentially impact the outcomes of different hospital types.
-Finally, I assume that, absent the program enactments, different hospital types would have had parallel trends in outcomes.
-
-    \section{Results}
 
@@
 
+    \section{Model and Empirical Strategy}\label{sec:model}
+
+    I model hospital behavior as maximizing some weighted average of profit and societal (or patient) benefit.
+Abstracting away from quantity and price as functions of quality for the purpose of this paper, I present a two-period model of hospital behavior that captures the change in quality that occurs after pay-for-performance incentives are introduced.
+
+    \subsection{Pre-Program Incentives}
+
+    In the first period, prior to program enactment, hospitals do not expect any penalties in the future, and choose quality according to:
+
+    $$\max_{\theta}\hspace{2mm}\alpha\left[R - c_{\theta}(\theta) \right] + (1-\alpha) u(\theta),$$
+
+    \noindent where $R$ is net revenue, $c_{\theta}(.)$ is an increasing cost function specific to quality $\theta$, $u(.)$ is extra utility gained from quality, which is increasing and concave is $\theta$, and $\alpha\in[0,1]$ captures how much the hospital cares about profit vs.
+societal benefit.
+There is an implicit stay open condition throughout the model.
+Taking the first order condition yields
+
+    $$(1-\alpha)u'(\theta) = \alpha c_{\theta}'(\theta),$$
+
+    \noindent marginal benefit equals marginal cost of increasing quality. Solving for $u'(\theta)$ and differentiating with respect to $\alpha$ yields
+
+    $$\frac{du'(\theta)}{d\alpha} = c_{\theta}'(\theta)\frac{1}{(1-\alpha)^2} > 0.$$
+
+    \noindent Thus, by the Implicit Function Theorem,
+
+    $$\frac{d\theta}{d\alpha} = \frac{du'(\theta)/d\alpha}{du'(\theta)/d\theta} < 0.$$
+
+    \noindent That is, the more weight the hospital places on profit, the lower quality of care they will provide.
+
+    \subsection{Post-Program Incentives}
+
+    For simplicity, I model pay-for-performance incentives as a simplified version of how they were applied in HRRP, as penalties for low quality hospitals.
+However, these results still hold when including incentive payments as in HVBP.
+
+    Hospitals determine period 1 quality, $\theta_1$, according to the optimal $\theta(\alpha)$ described above.
+Then, in period two, hospitals are penalized or not depending on $\theta_1$, and they choose a new quality, $\theta_2$, which will determine penalties in the future.
+The penalty $(k)$ comes in the form of a negative shock to revenue for hospitals with quality below a certain exogenous threshold, $\overline{\theta}$.
+That is, new revenue is:
+
+    $$R_2 = \begin{cases}
+        R - k, \text{  if  } \theta_1<\overline{\theta}\\
+        R, \text{  if  } \theta_1\geq \overline{\theta}.
+    \end{cases}$$
+
+    The penalty $k$ is large and exceeds the maximum cost of quality.
+Therefore, all hospitals will choose $\theta_2$ to not be penalized in the future.
+Hospitals who were penalized will increase quality just to the point of not being penalized, and hospitals who were not penalized will not change their quality.
+Thus,
+
+    $$\theta_2 = \begin{cases}
+        \overline{\theta}, \text{  if  } \theta_1<\overline{\theta}\\
+        \theta_1, \text{  if  } \theta_1\geq \overline{\theta}.
+    \end{cases}$$
+
+    \noindent Taking the change in theta from period 1 to period 2, and then differentiating with respect to $\alpha$, yields:
+
+    $$\Delta\theta = \theta_2-\theta_1 = \begin{cases}
+        \overline{\theta}-\theta_1, \text{  if  } \theta_1<\overline{\theta}\\
+        0, \text{  if  } \theta_1\geq \overline{\theta}.
+    \end{cases}$$
+
+    $$\frac{d\Delta\theta}{d\alpha} = \begin{cases}
+        -\frac{d\theta_1}{d\alpha}, \text{  if  } \theta_1<\overline{\theta}\\
+        0, \text{  if  } \theta_1\geq \overline{\theta}.
+    \end{cases}$$
+
+    I show in period one that $\frac{d\theta}{d\alpha}<0$.
+Therefore, $\frac{d\Delta\theta}{d\alpha}\geq0$.
+That is, the response to the program incentives is weakly increasing in $\alpha$, and there is a threshold for which increasing $\alpha$ (a hospital with more weight on profit), increases the response to the penalty.
+Intuitively, the programs change hospital incentives in that quality now enters directly into the profit function.
+Therefore, hospitals who place more weight on profit will have a greater response to the program incentives.
+
+    The theoretical prediction is that change in quality depends on how much weight the hospital places on profit vs.
+societal benefit.
+However, $\alpha$s are not observed in reality.
+Thus, I empirically investigate whether the response to the program is correlated with hospital ownership or clinical training on the executive team in order to gain understanding about how these things point to an underlying $\alpha$.
+
+    \subsection{Estimation}
+
+    First, I test empirically whether different sub-samples of nonprofits respond to program incentives differently than for-profits.
+While nonprofit hospital objectives are largely unknown, for-profits have revealed their objective in their ownership classification.
+That is, $\alpha=1$ (or is very close).
+The goal in this analysis is to learn something about the underlying $\alpha$ for nonprofit firms, and determine if executive team composition is a characteristic that points to differing $\alpha$s even among nonprofits.
+Thus, I compare their program response to that of for-profits.
+I estimate the following regression equation:
+
+    \begin{equation}
+    \label{eq:forprofit}
+    y_{ht} = \beta (\text{for-profit}_h \times \text{post programs}_t) + \gamma_{h} + \delta_t + \epsilon_{ht}
+    \end{equation}
+
+    \noindent where $y_{ht}$ is one of the outcome variables discussed in Section \ref{sec:data}, $\beta$ is the coefficient of interest capturing the effect of being for-profit after the programs were introduced, and $\gamma_h$ and $\delta_t$ are hospital and time fixed effects, respectively.
+For each outcome, I estimate three models where the comparison group differs by hospital type: all nonprofits, nonprofits with a clinically trained executive, and nonprofits without a clinically trained executive.
+I also directly estimate the difference in response of hospitals with and without a clinically trained executive:
+
+    \begin{equation}
+    \label{eq:clinical}
+    y_{ht} = \beta (\text{no MD exec}_h \times \text{post programs}_t) + \gamma_{h} + \delta_t + \epsilon_{ht}
+    \end{equation}
+
+    An important distinction is whether the classification of nonprofits as those with or without clinically trained executives is time-varying.
+For overall ownership status, switching is rare, and I limit to those who never switch.
+However, hospital executive teams are dynamic, and whether or not a hospital employs a clinically trained executive in each period may change.
+Therefore, I first present effects of having a clinically trained executive among stable executive teams, i.e., hospitals with no change in their propensity to hire a clinically trained executive.
+
+    Additionally, under the assumption that these changes are not correlated with the program enactments (see Section \ref{sec:endog}), using the changes can actually be advantageous in better understanding underlying hospital motives.
+There are two reasons we might see having a clinically trained executive affecting outcomes: first, a clinically trained executive could be a signal of the underlying weight placed on profit, $\alpha$.
+Second, hiring a clinically trained executive who carries out day-to-day operations in a distinct way effectively changes $\alpha$.
+Thus, I use two types of variation in clinically trained executives to disentangle the two: a variable for whether the hospital ever hires a clinically trained executive, and a variable for whether the hospital has a clinically trained executive in 2012 when the programs took place.
+I use the same empirical specification with these two variables as the independent variation in executive teams, and limit the sample to the relevant comparison group:
+
+    \begin{equation}
+    \label{eq:decomp}
+    y_{ht} = \beta (\text{never/no 2012 exec}_h \times \text{post programs}_t) + \gamma_{h} + \delta_t + \epsilon_{ht}
+    \end{equation}
+
+    As discussed in Section \ref{sec:hrrp}, hospitals had incentive to decrease readmissions immediately when the HRRP rules were released in October of 2011, before the penalties were actually determined in October of 2012.
+However, I still define post penalty as occurring in 2012 or later, the first full of hospital response.
+
+    To identify the causal effect of hospital type on hospital program response, I rely on several assumptions.
+First, I assume that executive teams are not determined endogenously when the programs were enacted.
+I investigate the validity of this assumption in Section \ref{sec:endog}.
+Second, I assume that no other unobserved changes occurred in 2012 that are correlated with both hospital performance and type.
+
+    Finally, I assume that, absent the program enactments, different hospital types would have had parallel trends in outcomes.
+While event studies show no difference in pre-trends (see appendix blank), Table \ref{tab:sumstats_samples} shows that hospitals with a clinically trained executive are different than other hospitals in multiple characteristics, and the number of hospitals with clinically trained executives is relatively small.
+These together may cause concern that there are systematic differences between hospital types that could be driving the results.
+Therefore, my main specification uses the synthetic difference-in-differences method established in \citeauthor{arkhangelsky2021synthetic} (\citeyear{arkhangelsky2021synthetic}).
+This method creates unit weights that more heavily count comparison hospitals that have similar trends in outcomes prior to the programs.
+It also weighs time periods that balance pre-program with post-program for the comparison hospitals.
+Then, a simple two way fixed effects estimator is used with these weights.
+This is a more robust average treatment effect that relies less on a strong parallel trends assumption (\cite{arkhangelsky2021synthetic}).
+I present the synthetic diff-in-diff estimates as the main results, and they are largely similar to the standard two way fixed effects results in Appendix \ref{app:fullsample}.
+
+    \section{Results}
+
      \subsection{Readmission and Mortality Rates}
@@
      First, I consider the differential program response of for-profit firms compared to the different sub-samples of nonprofit firms.
-These results are presented in Table \ref{tab:forprofit_synth}.
+These results are presented in Table \ref{tab:forprofit_readmort_synth}.
 Models in the left panel estimate the effect of being for-profit after the programs were enacted, compared to nonprofit sub-samples, on readmission rates.
-There is no statistically meaningful difference between the response of for-profits and nonprofits as a whole, or nonprofits without a clinically trained executive.
+There is no difference between the response of for-profits and nonprofits as a whole, or nonprofits without a clinically trained executive.
 However, for-profits lower their readmissions by .3ppts compared to nonprofits without clinically trained executives.
@@
      Next, I directly estimate the difference in response by nonprofits with and without clinically trained executives.
-These results are presented in Table \ref{tab:MD_noMD_synth}, where the left has models with outcome readmission rates and the right has models with outcome mortality rates.
+These results are presented in Table \ref{tab:MD_noMD_readmort_synth}, where the left has models with outcome readmission rates and the right has models with outcome mortality rates.
 Nonprofits without clinically trained executives have a larger decrease in readmissions than those with clinically trained executive by .3ppts, while there is no difference in change in mortality rates.
-Tables \ref{tab:forprofit_synth} and \ref{tab:MD_noMD_synth} taken together show that, in terms of readmission rates, nonprofits without clinically trained executives behave more similarly to for-profits than they do other nonprofits.
+Tables \ref{tab:forprofit_readmort_synth} and \ref{tab:MD_noMD_readmort_synth} taken together show that, in terms of readmission rates, nonprofits without clinically trained executives behave more similarly to for-profits than they do other nonprofits.
 Under the assumption that having a clinically trained executive on the leadership team indicates a higher preference for social utility, these findings are consistent with the theoretical prediction presented in Section \ref{sec:model}.
@@
 
-    However, it is not clear whether these results are due to clinically trained executives being a signal of underlying hospital objectives, or that the leaders themselves make a difference.
+    However, it is not clear whether these results are due to clinically trained executives being a signal of underlying hospital objectives, or that the leaders themselves make a difference in outcomes.
 Therefore, I present a decomposition of the results using two models in Table \ref{tab:MD_noMD_readmort_decomp_synth}, where the variation in executives is either whether the hospital ever has a clinically trained executive, or whether the hospital has a clinically trained executive in 2012 when the programs were enacted.
@@
 
-\section{Full Sample Regression Results}
+\section{Full Sample Regression Results}\label{app:fullsample}
 
@@
 
-\section{Matched Sample Regression Results}
-
-\import{Tables}{MD_noMD_readmort_matchsample.tex}
-
-\import{Tables}{forprofit_readmort_matchsample.tex}
-
-\import{Tables}{MD_noMD_uncompCMI_matchsample.tex}
-
-\import{Tables}{forprofit_uncompCMI_matchsample.tex}
-
 \end{document}

### 2024-05-01 Tables/MD_noMD_readmort_decomp_synth.tex
Updates from Overleaf

@@
 \midrule
-Ever MD Exec x Post Programs & 0 &  & -0.06 & \\
+Never MD Exec x Post Programs & 0 &  & -0.06 & \\
  & (0.06) &  & (0.06) & \\
-2012 MD Exec x Post Programs &  & -0.19$^{*}$ &  & -0.1\\
- &  & (0.09) &  & (0.08)\\
+Ever MD (not in 2012) &  & -0.19$^{*}$ &  & -0.1\\
+ \hspace{3mm} x Post Programs &  & (0.09) &  & (0.08)\\
  &  &  &  & \\
@@
 \textbf{Comparison Group:} &  &  &  & \\
-Never MD & $\checkmark$ &  & $\checkmark$ & \\
-Has MD not in 2012 &  & $\checkmark$ &  & $\checkmark$\\
+Ever MD Exec & $\checkmark$ &  & $\checkmark$ & \\
+Has MD Exec 2012 &  & $\checkmark$ &  & $\checkmark$\\
  &  &  &  & \\
-Observations & 3925 & 1845 & 3910 & 1840\\
+Observations & 3,925 & 1,845 & 3,910 & 1,840\\
 \bottomrule
-\multicolumn{5}{l}{\textsuperscript{} Results from Equation with readmission and mortality rates as}\\
-\multicolumn{5}{l}{outcome variables.}\\
+\multicolumn{5}{l}{\textsuperscript{} Results from Equation \ref{eq:decomp} with readmission and mortality rates as outcome variables.}\\
 \multicolumn{5}{l}{\textsuperscript{} Standard errors are clustered at the hospital level.}\\

### 2024-05-01 Tables/MD_noMD_readmort_synth.tex
Updates from Overleaf

--- 2024-05-01 Tables/MD_noMD_readmort_synth.tex
@@
 \midrule
-NP w/out MD Exec x Post Program & -0.28$^{*}$ & -0.19\\
+No MD Exec x Post Programs & -0.28$^{*}$ & -0.19\\
  & (0.13) & (0.11)\\
@@
 \bottomrule
-\multicolumn{3}{l}{\textsuperscript{} Results from Equation with readmission and mortality rates as}\\
-\multicolumn{3}{l}{outcome variables.}\\
+\multicolumn{3}{l}{\textsuperscript{} Results from Equation \ref{eq:clinical} with readmission and mortality rates as outcome variables.}\\
 \multicolumn{3}{l}{\textsuperscript{} Standard errors are clustered at the hospital level.}\\

### 2024-05-01 Tables/MD_noMD_uncompCMI_decomp_synth.tex
Updates from Overleaf

@@
 \midrule
-Ever MD Exec x Post 2012 & 0.01 &  & 0.01$^{*}$ & \\
+Never MD Exec x Post Programs & 0.01 &  & 0.01$^{*}$ & \\
  & (0.01) &  & (0) & \\
-2012 MD Exec x Post 2012 &  & -0.03 &  & 0.01\\
- &  & (0.03) &  & (0.01)\\
+Ever MD (not in 2012) &  & -0.03 &  & 0.01\\
+ \hspace{3mm} x Post Programs&  & (0.03) &  & (0.01)\\
  &  &  &  & \\
@@
 \textbf{Comparison Group:} &  &  &  & \\
-Never MD & $\checkmark$ &  & $\checkmark$ & \\
-Has MD not in 2012 &  & $\checkmark$ &  & $\checkmark$\\
+Ever MD Exec & $\checkmark$ &  & $\checkmark$ & \\
+Has MD Exec 2012 &  & $\checkmark$ &  & $\checkmark$\\
  &  &  &  & \\
-Observations & 1195 & 635 & 3015 & 1595\\
+Observations & 1,195 & 635 & 3,015 & 1,595\\
 \bottomrule
-\multicolumn{5}{l}{\textsuperscript{} Results from Equation with readmission and mortality rates as}\\
-\multicolumn{5}{l}{outcome variables.}\\
+\multicolumn{5}{l}{\textsuperscript{} Results from Equation with readmission and mortality rates as outcome variables.}\\
 \multicolumn{5}{l}{\textsuperscript{} Standard errors are clustered at the hospital level.}\\

### 2024-05-01 Tables/forprofit_uncompCMI_synth.tex
Update tables AGAIN
Updates from Overleaf

@@
 \centering
-\begin{tabular}[t]{l>{\centering\arraybackslash}p{.2textwidth}>{\centering\arraybackslash}p{.2textwidth}>{\centering\arraybackslash}p{.2textwidth}>{\centering\arraybackslash}p{.2textwidth}>{\centering\arraybackslash}p{.2textwidth}>{\centering\arraybackslash}p{.2textwidth}}
+\begin{tabular}[t]{lcccccc}
 \toprule
 \multicolumn{1}{c}{ } & \multicolumn{3}{c}{Uncompensated Care} & \multicolumn{3}{c}{Case Mix Index} \\
-\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-7}
  & (1) & (2) & (3) & (1) & (2) & (3)\\
@@
 Nonprofits w/ MD &  & $\checkmark$ &  &  & $\checkmark$ & \\
-\addlinespace
 Nonprofits w/out MD &  &  & $\checkmark$ &  &  & $\checkmark$\\
  &  &  &  &  &  & \\
-Observations & 4935 & 1330 & 1815 & 12320 & 2880 & 4065\\
+Observations & 4,935 & 1,330 & 1,815 & 12,320 & 2,880 & 4,065\\
 \bottomrule
-\multicolumn{7}{l}{\textsuperscript{} Results from Equation with readmission and mortality rates as outcome variables.}\\
+\multicolumn{7}{l}{\textsuperscript{} Results from Equation with uncompensated care and CMI as outcome variables.}\\
 \multicolumn{7}{l}{\textsuperscript{} Standard errors are clustered at the hospital level.}\\

### 2024-05-06 Tables/MD_noMD_uncompCMI_synth.tex
Updates from Overleaf
fix sample of nonprofits and add synthetic difference in differences graphs

@@
 \midrule
-NP w/out MD Exec x Post Programs & 0.02$^{*}$ & 0.03$^{***}$\\
+NFP w/out MD Exec x Post Programs & 0.02$^{*}$ & 0.03$^{**}$\\
  & (0.01) & (0.01)\\
 \textbf{Comparison Group:} &  & \\
-NP w/ MD Exec & $\checkmark$ & $\checkmark$\\
+NFP w/ MD Exec & $\checkmark$ & $\checkmark$\\
  &  & \\
 \addlinespace
-Observations & 15,585 & 2,900\\
+Observations & 15,610 & 2,900\\
 \bottomrule
-\multicolumn{3}{l}{\textsuperscript{} Results from Equation with readmission and mortality rates as outcome}\\
-\multicolumn{3}{l}{variables.}\\
+\multicolumn{3}{l}{\textsuperscript{} Results from Equation with uncompensated care and CMI as outcome variables.}\\
 \multicolumn{3}{l}{\textsuperscript{} Standard errors are clustered at the hospital level.}\\

### 2024-05-06 Tables/forprofit_readmort_synth.tex
fix sample of nonprofits and add synthetic difference in differences graphs

@@
 \multicolumn{1}{c}{ } & \multicolumn{3}{c}{Weighted Avg. Readmission Rate} & \multicolumn{3}{c}{Weighted Avg. Mortality Rate} \\
- & (1) & (2) & (3) & (1) & (2) & (3)\\
+ & (1) & (2) & (3) & (4) & (5) & (6)\\
 \midrule
@@
 \textbf{Comparison Group:} &  &  &  &  &  & \\
-All Nonprofits & $\checkmark$ &  &  & $\checkmark$ &  & \\
-Nonprofits w/ MD &  & $\checkmark$ &  &  & $\checkmark$ & \\
-Nonprofits w/out MD &  &  & $\checkmark$ &  &  & $\checkmark$\\
+In-Sample NFPs & $\checkmark$ &  &  & $\checkmark$ &  & \\
+NFPs w/ MD &  & $\checkmark$ &  &  & $\checkmark$ & \\
+\addlinespace
+NFPs w/out MD &  &  & $\checkmark$ &  &  & $\checkmark$\\
  &  &  &  &  &  & \\
-Observations & 15,585 & 2,900 & 4,705 & 15,505 & 2,895 & 4,690\\
+Observations & 15,610 & 2,900 & 4,705 & 15,530 & 2,895 & 4,690\\
 \bottomrule

### 2024-05-06 Tables/forprofit_uncompCMI_synth.tex
fix sample of nonprofits and add synthetic difference in differences graphs

@@
 \multicolumn{1}{c}{ } & \multicolumn{3}{c}{Uncompensated Care} & \multicolumn{3}{c}{Case Mix Index} \\
- & (1) & (2) & (3) & (1) & (2) & (3)\\
+\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-7}
+ & (1) & (2) & (3) & (4) & (5) & (6)\\
 \midrule
-For-Profit x Post Programs & 0.06$^{***}$ & 0.08$^{***}$ & 0.06$^{***}$ & 0 & 0.02$^{**}$ & 0\\
+For-Profit x Post Programs & 0.06$^{***}$ & 0.08$^{***}$ & 0.06$^{***}$ & 0 & 0.02$^{*}$ & 0\\
  & (0.01) & (0.01) & (0.01) & (0.01) & (0.01) & (0.01)\\
 \textbf{Comparison Group:} &  &  &  &  &  & \\
-All Nonprofits & $\checkmark$ &  &  & $\checkmark$ &  & \\
-Nonprofits w/ MD &  & $\checkmark$ &  &  & $\checkmark$ & \\
-Nonprofits w/out MD &  &  & $\checkmark$ &  &  & $\checkmark$\\
+In-Sample NFPs & $\checkmark$ &  &  & $\checkmark$ &  & \\
+NFPs w/ MD &  & $\checkmark$ &  &  & $\checkmark$ & \\
+\addlinespace
+NFPs w/out MD &  &  & $\checkmark$ &  &  & $\checkmark$\\
  &  &  &  &  &  & \\
-Observations & 4,935 & 1,330 & 1,815 & 12,320 & 2,880 & 4,065\\
+Observations & 4930 & 1330 & 1815 & 12320 & 2880 & 4065\\
 \bottomrule

### 2024-05-06 Paper/paper.tex
Updates from Overleaf

@@
 
-    \begin{center}
-    \large
-    \textbf{This is a preliminary draft. Please do not cite or circulate.}
-	\end{center}
-
  \vspace{3mm}
@@
 		{\small
-        In many industries, nonprofits make up the majority of firms in the market.
+        In many industries, not-for-profit (NFP) make up the majority of firms in the market.
 Understanding the objectives and behaviors of these firms is important when creating incentives.
-However, nonprofit firms have goals and missions that make understanding their objectives difficult.
-For hospitals in particular, there is mixed evidence on whether nonprofit hospitals differ meaningfully from for-profit hospitals.
+However, NFP firms have goals and missions that make understanding their objectives difficult.
+For hospitals in particular, there is mixed evidence on whether NFP hospitals differ meaningfully from for-profit (FP) hospitals.
 In this paper, I investigate whether leadership team composition of nonprofit hospitals affects how similarly they behave to for-profits.
-I use a novel data set on nonprofit hospital executives in the US from 2009-2015 to investigate whether the presence of clinically trained executives affects hospital behavior.
-In particular, I leverage two pay-for-performance initiatives in the US enacted in 2012, and estimate the difference in response between for-profits, nonprofits, and nonprofits with clinically trained executives.
-I find that while readmissions decreased on average for all types of hospitals, the response was larger for for-profits and nonprofits without clinically trained executives.
-This is consistent with theoretical predictions under the assumption that nonprofits with clinically trained executives put greater weight on societal outcomes than those without clinically trained executives.
+I use a novel data set on NFP hospital executives in the US (2009-2015) to investigate whether the presence of clinically trained executives affects hospital behavior.
+In particular, I leverage two pay-for-performance initiatives in the US enacted in 2012, and estimate the difference in readmission and mortality rates afterwards between FP, NFP, and NFP with and without clinically trained executives.
+I find that NFPs without clinically trained executives behave more like FPs than other NFPs.
+This is consistent with theoretical predictions under the assumption that NFP hospitals with clinically trained executives value societal outcomes more than those without clinically trained executives.
 		} 
@@
   The objectives and behaviors of firms that are not classic for-profits are not easily determined, and an extensive literature has sought to understand these types of firms.
-Nonprofits are thought to gain utility not solely from monetary profit, but also through accomplishing societal benefit.
-Researchers have speculated a variety of objective functions that could motivate nonprofits such as maximizing prestige, income, or a quality/quantity trade-off (\cite{steinberg1986revealed}).
-Empirical contributions to this literature center around investigating behaviors that differ between for-profit, private nonprofit, and government firms (\cite{sloan2000not}).
-In this paper, I investigate whether differences in observed behavior between for-profit and nonprofit hospitals vary differentially by nonprofit leadership team composition.
+Not-for-profits (NFP) are thought to gain utility not solely from monetary profit, but also through accomplishing societal benefit.
+Researchers have speculated a variety of objective functions that could motivate NFP firms such as maximizing prestige, income, or a quality/quantity trade-off (\cite{steinberg1986revealed}).
+Empirical contributions to this literature center around investigating behaviors that differ between for-profit (FP), private NFP, and government firms (\cite{sloan2000not}).
+In this paper, I investigate whether differences in observed behavior between for-profit and not-for-profit hospitals vary differentially by leadership team composition.
 
-  Hospitals have been a common focus when seeking to understand nonprofit firms.
-This is because, first, there are different hospital ownership types within the industry, giving researchers the ability to compare the behaviors of nonprofits with that of for-profits.
-Second, private nonprofit hospitals make up 50\% of all hospitals in the US, and staff, on average, 207 beds.
-For comparison, for-profits make up around 36\% of hospitals and staff 107 beds (\cite{ASPE_2023}).
-Thus, nonprofit behavior in the hospital industry is relevant, and directly affects consumers of health care.
-However, it remains unclear what objective function drives nonprofit hospital behavior, and whether nonprofit hospitals are meaningfully different from for-profits (\cite{sloan2000not}; \cite{erus2002inferring}; \cite{deneffe2002not}; \cite{horwitz2009hospital}).
+  Hospitals have been a common focus when seeking to understand NFP firms.
+This is because, first, there are different hospital ownership types within the industry, giving researchers the ability to compare the behaviors of NFPs and FPs.
+Second, private NFP hospitals make up 50\% of all hospitals in the US, and staff, on average, 207 beds.
+For comparison, FPs make up around 36\% of hospitals and staff 107 beds (\cite{ASPE_2023}).
+Thus, NFP behavior in the hospital industry is relevant, and directly affects consumers of health care.
+However, it remains unclear what objective function drives NFP hospital behavior, and whether NFP hospitals are meaningfully different from FP hospitals (\cite{sloan2000not}; \cite{erus2002inferring}; \cite{deneffe2002not}; \cite{horwitz2009hospital}).
 
-  While the characterization of being for-profit or nonprofit is certainly important, there could be other characteristics that drive objectives and behavior that are not perfectly correlated with ownership status.
-Particularly, firm leadership varies widely across firms and potentially drives differences in objectives even within nonprofits.
-In large, publicly traded firms, executive characteristics are shown to be correlated with firm performance, indicating that there is more to firm behavior than for-profit or nonprofit status (\cite{bertrand2003managing}; \cite{matsa2013female}; \cite{ahern2012changing}).
-If hospital objectives differ in the weights placed on profit vs.
-societal benefit, then ownership status certainly plays a role in these weights, but characteristics of those managing the hospital could also play an important role.
-I contribute to our understanding of hospital and nonprofit behaviors by investigating whether occupational background, whether executives have clinical training, affects response to pay-for-performance incentives.
+  While the characterization of being for-profit or not-for-profit is certainly important, there could be other characteristics that affect hospital behavior that are not perfectly correlated with individual ownership status.
+Researchers have focused in particular on explaining the differences in NFP behavior by the market they exist again (citations).
+I argue that an observable and economically meaningful factor is the composition of leadership teams.
+Firm leadership varies widely across firms and potentially drives differences in objectives even within NFPs.
+In large, publicly traded firms, executive characteristics are shown to be correlated with firm performance, indicating that there is more to firm behavior than ownership status (\cite{bertrand2003managing}; \cite{matsa2013female}; \cite{ahern2012changing}).
+If hospital objectives differ in how they value monetary profit vs.
+gains to society, then ownership status certainly plays a role in these valuations, but the question remains as to whether management also plays an important role.
+I contribute to our understanding of hospital and NFP behaviors by investigating whether occupational background, specifically clinical training in executives, affects hospital response to pay-for-performance incentives.
 
-  Using publicly available tax forms, I construct a novel data set of nonprofit hospital executives from 2009-2015.
-I link hospitals in this data to characteristics in the American Hospital Association survey and Hospital Compare data.
-I leverage various pay-for-performance initiatives enacted in 2012 to investigate differential hospital responses between for-profits, nonprofits without clinically trained executives, and nonprofits with clinically trained executives.
-While executive teams change over time, I find no evidence that they change in response to the programs.
-This allows me to investigate the role of particular executives in two ways.
-First, I analyze the differential response to incentives of nonprofits who ever hire a clinically trained executive, as this might be a signal of underlying preferences.
-Second, I analyze the differential response of nonprofits who have a clinically trained executive in 2012 compared to those who have a clinically trained executive at other times, but not in 2012.
-These two analyses aim to disentangle the effect of the actual leadership team from the underlying preferences of the hospital.
+  Using publicly available tax forms, I construct a novel data set of NFP hospital executives from 2009-2015.
+I link hospitals in this data to characteristics in the American Hospital Association (AHA) survey and Hospital Compare data.
+I leverage various pay-for-performance initiatives in 2012 to investigate differential hospital responses between for-profits, NFPs without clinically trained executives, and NFPs with clinically trained executives.
 
@@
 Among the hospitals in the data, hospitals with clinically trained executives are the minority, and differ on average from other hospitals along various dimensions.
-Thus, my preferred specification is synthetic difference-in-differences, to analyze a group of comparison hospitals with similar characteristics to treatment hospitals.\footnote{I also present results for the full sample, and with a pre-analysis matching procedure and find qualitatively similar results.}
+Thus, my preferred specification is synthetic difference-in-differences, to analyze a group of comparison hospitals with similar characteristics to treatment hospitals.\footnote{I also present results for the full sample, and find qualitatively similar results.}
 
-   While all hospital types decrease readmission rates after the programs are enacted, for-profits and nonprofits without clinically trained executives decrease readmissions more than nonprofits without clinically trained executives.
-That is, nonprofits without clinically trained executives respond to incentives more similarly to for-profits than other nonprofits.
-To investigate the mechanisms by which hospitals decrease readmissions, I also conduct the analyses with uncompensated care and case mix index as outcomes.
+   While all hospital types decrease readmission rates after the programs are enacted, for-profits and not-for-profits without clinically trained executives decrease readmissions more than not-for-profits with clinically trained executives.
+That is, NFPs without clinically trained executives respond to incentives more similarly to FPs than other NFPs.
+I then consider whether this effect is due to clinically trained executives being a signal of underlying hospital preferences, or changing hospital preferences in a way that then affects behavior.
+I decompose the two by leveraging hospitals who change their propensity to hire a clinically trained executive at some point in the sample, which I find is not occurring endogenously with the change in incentives.
+The results of this analysis suggest that the difference in response is driven by changing hospital preferences, not as a signal of underlying preferences.
+While readmissions and mortality were targeted directly by the programs, I also consider uncompensated care and case mix index as outcomes, since re-allocation of resources and patient selectivity are potential mechanisms to changing hospital outcomes.
 (results of this analysis here)
@@
     This paper contributes to three strands of literature.
-First, it adds to our understanding of what underlies nonprofit hospital behavior.
-This literature is extensive, and a review of it is beyond the scope of this paper.
-To summarize, there are various aspects of nonprofit behavior and objectives for which the evidence of how much it differs from for-profits remains mixed: costs, uncompensated care, technology adoption, and quality (\cite{sloan2000not}; \cite{eggleston2008hospital}; \cite{moscelli2018effect}; \cite{moscone2020public}).
-What underlies this is the need to understand the ways in which nonprofits make choices differently than for-profits.
-To this point, researchers have found that nonprofits do not act as purely profit maximizers, but maximize some combination of profit and output (\cite{deneffe2002not}, \cite{chang2011nonprofit}).
+First, it adds to our understanding of what underlies NFP hospital behavior.
+There is mixed evidence on whether NFP behavior differs from FP behavior: costs, uncompensated care, technology adoption, and quality (\cite{sloan2000not}; \cite{eggleston2008hospital}; \cite{moscelli2018effect}; \cite{moscone2020public}).
+Underlying hospital are hospital objectives.
+To this point, researchers have found that NFPs do not act as purely profit maximizers, but maximize some combination of profit and output (\cite{deneffe2002not}, \cite{chang2011nonprofit}).
 A paper with an approach similar to mine which investigates hospital behavior is \citeauthor{chang2011nonprofit} (\citeyear{chang2011nonprofit}), which looks at the effect of a cost shock on likelihood of shutting down and mix of profitable services.
-They find that nonprofits and for-profits are equally as likely to shut down after the shock, but only nonprofits adjust their mix of profitable services.
+They find that NFPs and FPs are equally as likely to shut down after the shock, but only NFPs adjust their mix of profitable services.
 Similarly, I investigate a change in incentives to hospitals, but I look at the differential impacts of leadership teams within nonprofits.
@@
 Female board members are associated with better oversight and more women executives (\cite{matsa2011chipping}; \cite{adams2009women}), but are negatively correlated with firm value (\cite{ahern2012changing}).
-Having more female executives is correlated with choices of female employee wages and corporate strategies (\cite{flabbi2019female}; \cite{matsa2013female}); young male CEOs tend to be more aggressive in mergers and acquisitions, while those with military experience are less aggressive (\cite{levi2010deal}; \cite{benmelech2015military}); CEOs with general ability tend to receive higher pay and perform better (\cite{kaplan2012ceo}; \cite{custodio2013generalists}, \cite{adams2018director}; \cite{frydman2019rising}).
+Having more female executives is correlated with female employee wages and corporate strategies (\cite{flabbi2019female}; \cite{matsa2013female}); young male CEOs tend to be more aggressive in mergers and acquisitions, while those with military experience are less aggressive (\cite{levi2010deal}; \cite{benmelech2015military}); CEOs with general ability tend to receive higher pay and perform better (\cite{kaplan2012ceo}; \cite{custodio2013generalists}; \cite{adams2018director}; \cite{frydman2019rising}).
 Further, Chief Diversity Officers are not found to have any effect on hiring more diversely in universities (\cite{bradley2022impact}).
@@
     Due to data limitations, many of these studies focus on large, publicly traded firms in the US.
-However, two studies focus on hospital performance in other countries.
+\citeauthor{brickley2010board} (\citeyear{brickley2010board}) has the only study to my knowledge answering this question in the US NFP context.
+Using exogenous variation in expected Medicare profits, the authors find that having an internal board of directors increases CEO compensation, and having physician board members decreases public donations (\cite{brickley2010board}).
+Additionally, two studies focus on hospital performance in other countries.
 \citeauthor{janke2019impact} (\citeyear{janke2019impact}) uses data from England to study whether CEOs affect hospital production, and find no association (\cite{janke2019impact}).
 However, \citeauthor{otero2022managers} (\citeyear{otero2022managers}) investigates the role of CEOs in public hospitals and Chile, and finds an 8\% decrease in mortality rates for hospitals with top managers (\cite{otero2022managers}).
-I contribute to this literature by looking at the executive team as a whole rather than just CEOs, investigating a context which is not largely studied but is policy relevant, nonprofit US hospitals, and considering a unique executive characteristic pertinent to that context.
+I contribute to this literature by considering the executive team as a whole rather than just CEOs, by investigating a context which is not largely studied but is policy relevant, NFP US hospitals, and by leveraging a unique executive characteristic pertinent to that context.
 
@@
 The most in depth study of how hospitals respond to HRRP is \citeauthor{gupta2021impacts} (\citeyear{gupta2021impacts}), who finds that hospitals decreased readmissions by 5\% and mortality rates by 2\% on average as a result of the program, confirming prior studies (\cite{mellor2017does}; \cite{ziedan2018essays}; \cite{ody2019decreases}; \cite{gupta2021impacts}).
-Importantly, around 40\% of the decrease in readmissions is due to selective patient practices.
-This raises important questions about how to incentivize improvement in quality of care without also incentivizing gamesmanship.
+Around 40\% of the decrease in readmissions is due to selective patient practices.
 Research on HVBP generally finds that the program had no effect on underlying hospital quality (\cite{us2015hospital}; \cite{norton2018moneyball}; \cite{friedson2019so}).
@@
 
-    Nonprofits are governed by a board of directors, whose role is to set high level goals and strategies, and provide general oversight.
-The board is not in charge of day-to-day operations, but the broad direction of the firm.
-They select executive team members to carry out day-to-day operations.
-While hospital executives typically do specialize by getting degree(s) in healthcare management or an MBA specific to healthcare, they do not often have clinical training.
+    NFPs are governed by a board of directors, whose role is to set broad goals and strategies, and provide general oversight.
+The board selects executive team members to carry out day-to-day operations.
+While hospital executives typically specialize by getting degree(s) in healthcare management or an MBA specific to healthcare, they do not often have clinical training.
 While some doctors earn additional degrees before stepping into an executive role, this is not a necessary condition to becoming a physician executive.
-An executive team usually consists of at least a CEO and CFO, but every nonprofit is different in exactly how they structure their executive team.
-Some nonprofits hire Chief Medical Officers, Chief Quality Officers, executive directors, vice presidents of various departments, etc.
+An executive team usually consists of at least a CEO and CFO, but every NFP is different in exactly how they structure their executive team.
+Some NFPs hire Chief Medical Officers, Chief Quality Officers, executive directors, vice presidents of various departments, etc.
 I consider all of these employees as part of the day-to-day management of the hospital and include them as a part of the executive team.

### 2024-05-06 Paper/paper.tex
Updates from Overleaf

@@
     \label{eq:forprofit}
-    y_{ht} = \beta (\text{for-profit}_h \times \text{post programs}_t) + \gamma_{h} + \delta_t + \epsilon_{ht}
+    y_{ht} = \beta (\text{For-Profit}_h \times \text{Post Programs}_t) + \gamma_{h} + \delta_t + \epsilon_{ht}
     \end{equation}
@@
     \label{eq:clinical}
-    y_{ht} = \beta (\text{no MD exec}_h \times \text{post programs}_t) + \gamma_{h} + \delta_t + \epsilon_{ht}
+    y_{ht} = \beta (\text{No MD Exec}_h \times \text{Post Programs}_t) + \gamma_{h} + \delta_t + \epsilon_{ht}
     \end{equation}
@@
     \label{eq:decomp}
-    y_{ht} = \beta (\text{never/no 2012 exec}_h \times \text{post programs}_t) + \gamma_{h} + \delta_t + \epsilon_{ht}
+    y_{ht} = \beta (\text{Never/No 2012 Exec}_h \times \text{Post Programs}_t) + \gamma_{h} + \delta_t + \epsilon_{ht}
     \end{equation}
@@
 
-     \import{Tables}{forprofit_readmort_synth.tex}
+     \begin{figure}
+         \centering
+         \caption{Readmission Results}
+         \includegraphics[width=\textwidth]{Objects/read_synth_plot.pdf}
+         \label{fig:read_synth_plot}
+     \end{figure}
 
@@
 
-     \import{Tables}{MD_noMD_readmort_synth.tex}
+      \begin{figure}
+         \centering
+         \caption{Mortality Results}
+         \includegraphics[width=\textwidth]{Objects/mort_synth_plot.pdf}
+         \label{fig:read_synth_plot}
+     \end{figure}
 
@@
 
-\section{Full Sample Regression Results}\label{app:fullsample}
+\section{Full Sample Regression and Event Study Results}\label{app:fullsample}
 
@@
 
+\section{Synthetic Diff-in-Diff Result Tables}
+
+\import{Tables}{forprofit_readmort_synth.tex}
+
+\import{Tables}{MD_noMD_readmort_synth.tex}
+
 \end{document}

### 2024-05-14 Paper/paper.tex
Updates from Overleaf

@@
 %\usepackage[nomarkers,nofiglist,notablist]{endfloat}
+\usepackage{subcaption}
+\usepackage{caption}
 
@@
 		{\small
-        In many industries, not-for-profit (NFP) make up the majority of firms in the market.
-Understanding the objectives and behaviors of these firms is important when creating incentives.
+        In many industries, not-for-profits (NFP) make up the majority of firms in the market.
+Understanding the objectives and behaviors of these firms is important when changing incentives.
 However, NFP firms have goals and missions that make understanding their objectives difficult.
 For hospitals in particular, there is mixed evidence on whether NFP hospitals differ meaningfully from for-profit (FP) hospitals.
-In this paper, I investigate whether leadership team composition of nonprofit hospitals affects how similarly they behave to for-profits.
+In this paper, I investigate whether leadership team composition of not-for-profit hospitals affects how similarly they behave to for-profits.
 I use a novel data set on NFP hospital executives in the US (2009-2015) to investigate whether the presence of clinically trained executives affects hospital behavior.
@@
 Underlying hospital are hospital objectives.
-To this point, researchers have found that NFPs do not act as purely profit maximizers, but maximize some combination of profit and output (\cite{deneffe2002not}, \cite{chang2011nonprofit}).
-A paper with an approach similar to mine which investigates hospital behavior is \citeauthor{chang2011nonprofit} (\citeyear{chang2011nonprofit}), which looks at the effect of a cost shock on likelihood of shutting down and mix of profitable services.
+To this point, researchers have found that NFPs do not act as purely profit maximizers, but maximize some combination of profit and output (\cite{deneffe2002not}, \cite{chang2011not-for-profit}).
+A paper with an approach similar to mine which investigates hospital behavior is \citeauthor{chang2011not-for-profit} (\citeyear{chang2011not-for-profit}), which looks at the effect of a cost shock on likelihood of shutting down and mix of profitable services.
 They find that NFPs and FPs are equally as likely to shut down after the shock, but only NFPs adjust their mix of profitable services.
-Similarly, I investigate a change in incentives to hospitals, but I look at the differential impacts of leadership teams within nonprofits.
-This paper adds to our understanding of inputs into nonprofit hospital objective functions.
+Similarly, I investigate a change in incentives to hospitals, but I look at the differential impacts of leadership teams within not-for-profits.
+This paper adds to our understanding of inputs into not-for-profit hospital objective functions.
 
@@
 
-    \subsection{Nonprofit Hospital Leadership}
+    \subsection{Not-for-Profit Hospital Leadership}
+
+    Not-for-profit firms differ from for-profits mainly in that for-profit firms distribute revenue to stakeholders, while not-for-profits reinvest profit back into the firm.
+Another important distinction is that NFPs are, in general, driven by a mission to better society.
+Additionally, NFPs in the US receive tax benefits for this designation.
 
     NFPs are governed by a board of directors, whose role is to set broad goals and strategies, and provide general oversight.
-The board selects executive team members to carry out day-to-day operations.
-While hospital executives typically specialize by getting degree(s) in healthcare management or an MBA specific to healthcare, they do not often have clinical training.
+The board selects executives, the highest level of management of a firm, to carry out day-to-day operations.
+An executive team usually consists of at least a CEO and CFO, but every NFP is different in exactly how they structure their executive team.
+Many hospital executives specialize in health care settings by getting degree(s) in health care management, or an MBA specific to health care.
+There are specific positions that are often filled by someone with clinical training, such as a Chief Medical Officer, and doctors can even find themselves in other top managerial positions as well, such as CEO or president.
 While some doctors earn additional degrees before stepping into an executive role, this is not a necessary condition to becoming a physician executive.
-An executive team usually consists of at least a CEO and CFO, but every NFP is different in exactly how they structure their executive team.
-Some NFPs hire Chief Medical Officers, Chief Quality Officers, executive directors, vice presidents of various departments, etc.
-I consider all of these employees as part of the day-to-day management of the hospital and include them as a part of the executive team.
+
+    Anecdotal evidence is mixed on the benefits of hiring physician executives.
+On one hand, physicians bring a unique combination of clinical expertise and being comfortable with administration, and they can bring insights that improve hospital outcomes (\cite{Stajduhar_2023}, \cite{Ahmed_2022}).
+Alternatively, physicians may not always be adequately trained to enter into high level management positions, and therefore are not equipped to lead a firm (\cite{HarvardBusinessReview2018}).
 
@@
 Sections of this form include a statement of revenue, statement of functional expenses, a balance sheet, and, as used in this project, a list of all key employees, executives, and board members.
-In this section, each firm is required to report the name and title, average hours per week, position, and compensation of their board members and executive leaders.
-Since these firms are tax exempt, their information in these forms is made publicly available.
+Each firm is required to report the name and title, average hours per week, position, and compensation of their board members and executive leaders.
 
-    \subsection{Pay-for-Performance Programs}\label{sec:hrrp}
+    \subsection{Pay-for-Performance Policies}\label{sec:hrrp}
 
-    The Hospital Readmissions Reduction Program (HRRP) was passed as a part of the Affordable Care Act in 2010 as a pay-for-performance incentive for hospitals to increase quality.
-In October 2011, the Center for Medicare and Medicaid Services (CMS) released a set of rules that mandated penalties for hospitals with above average readmission rates.
+    Two programs were passed as part of the Affordable Care Act that focused on pay-for-performance incentives for hospitals, the Hospital Readmissions Reduction Program (HRRP), and the Hospital Value Based Purchasing Program (HVBP).
+The HRRP focused on penalizing hospitals with poor quality in terms of readmission rates, and the HVBP fpcused on rewarding hospitals with high quality in terms of something.
+
+    In October 2011, the Center for Medicare and Medicaid Services (CMS) released a set of rules under HRRP that mandated penalties for hospitals with above average readmission rates.
 A readmission is when a patient returns to the hospital within 30 days of being discharged from a previous stay; avoidable readmissions are a bad outcome for patients and increase health care spending.
 The goal of HRRP is to lower readmissions through better care coordination, less initial stay complications, and better post-care instructions.
-Beginning in October 2012, hospitals with higher readmission rates than the national average in pneumonia, heart failure, or AMI (after adjusting for demographic characteristics) receive a lower reimbursement rate for Medicare patients.
-In 2015, CMS also included chronic obstructive pulmonary disease, coronary artery bypass graft surgery, and elective primary total hip arthroplasty and/or total knee arthroplasty as conditions which go into the penalty calculation.
-\textcolor{red}{(add citations??)}
+Beginning in October 2012, hospitals with higher readmission rates than the national average in pneumonia, heart failure, or AMI (after adjusting for demographic characteristics) receive a fixed lower reimbursement rate for all Medicare patients seen in their hospital.
+In 2015, CMS also included chronic obstructive pulmonary disease, coronary artery bypass graft surgery, and elective primary total hip arthroplasty and/or total knee arthroplasty as conditions which go into the penalty calculation (\cite{CMS}).
 
     Penalized hospitals pay penalties in the form of a fixed rate for every Medicare patient regardless of the condition.
+The rate began at 1\% and increased to 3\% in 2015.
 Further, CMS does not distinguish a necessary readmission from an avoidable readmission; any repeat hospital visit is included in the penalty calculation.
@@
 
-    \textcolor{red}{(is the program still going?)}
+    The HVBP Program instead rewards hospitals with high quality or shown improvement in quality.
+Specifically, CMS deducts Medicare payments by 2\%, collects this sum, and divides it among the rewarded hospitals.
+Several quality and cost measures in categories safety, efficiency and cost reductions, clinical outcomes, and community engagement, are combined to create a single score metric for each hospital.
+Hospitals are then compared to each other and score points for being above average quality and for showing improvement (\cite{CMS_2023}).
 
@@
 
-    \subsection{Names of Hospital Leaders}
+    One reason that hospital executive teams are understudied is because of the inaccessibility of the information.
+I construct a novel data set of not-for-profit hospitals in the US that contains names, titles, and positions of all executives tied to the hospital from 2009-2015.
+I gather this data from publicly available Tax Form 990s, which every sufficiently large not-for-profit files each year in the US.
+To my knowledge, this is the first large scale gathering of executive names from these forms.\footnote{\citeauthor{brickley2010board} (\citeyear{brickley2010board}) collects compensation information for a select number of hospitals.}
 
-    To characterize hospitals as having clinically trained executives or not, I construct a novel data set of nonprofit hospitals in the US that contains names and select characteristics of all executives tied to the hospital from 2009-2015.
-I gather this data from publicly available Tax Form 990s, which every sufficiently large nonprofit files each year in the US.
-To my knowledge, this is the first large scale gathering of executive names from these forms.
-
-    The historical tax forms for all nonprofits are housed by ProPublica.\footnote{https://projects.propublica.org/nonprofits/} I use the NonProfit Explorer API to extract the Employee Identification Numbers (EIN) of all nonprofits classified as a hospital according to their National Taxonomy of Exempt Entities (NTEE) code.
-After filtering out associations and specialty hospitals, there are around 3,000 firms left.
-Importantly, I need to be able to link hospitals in the tax forms to other hospital characteristics from publicly available sources.
-No official crosswalk between EINs and AHA IDs or Medicare numbers exists, so I must rely on matching by location and name of hospital in both data sets.
-After limiting to only general, nonprofit hospitals that I can link to the AHA survey, I confidently match approximately 1200 EINs to an AHA identification number based on exact name matches within the same state.
+    The historical tax forms for all not-for-profits are housed by ProPublica.\footnote{https://projects.propublica.org/not-for-profits/} I use the not-for-profit Explorer API to extract Employee Identification Numbers (EIN) of not-for-profits classified as a hospital according to their National Taxonomy of Exempt Entities (NTEE) code.
+After filtering out associations and specialty hospitals, there are approximately 3,000 not-for-profit EINs.
+Importantly, linking EINs to other hospital information is crucial for the analysis.
+I rely on matching by location and name of hospital in the tax forms and AHA data.
+I confidently match 1200 EINs to an AHA identification number based on exact name matches within the same state.
+I compare the matched hospitals to non-matched hospitals in Appendix \ref{app:matched}, and find that the matched sample is not qualitatively different than the non-matched hospitals.
 
     For each of the 1200 EINs, I then extract web URLs to Form 990 PDFs in each year.
-I download these locally and use optical character recognition (OCR) text extraction methods to scrape the relevant section of the pdf, which is titled ``Officers, Directors, Trustees, Key Employees, and Highest Compensated Employees".
+I download these locally and use optical character recognition (OCR) text extraction methods to scrape the relevant section of the PDF, which is titled ``Officers, Directors, Trustees, Key Employees, and Highest Compensated Employees".
 Finally, I use string cleaning methods to identify names and positions for each EIN, year.
-I also identify titles indicating a physician: MD, Dr., or DO.
-Using OCR extraction can be unreliable in some instances when the text is handwritten or the font is too small.
-Therefore, I manually find names, positions, and titles for observations that are missing after the initial extraction.
-This yields 850 EINs that are matched to AHA data, and are not missing text information.
+I also identify titles indicating clinical training: MD, Dr., or DO.
+Using OCR extraction can be unreliable in some instances if the text is handwritten or small.
+Therefore, I manually record names, positions, and titles for observations that are missing after the initial extraction.
+This yields 852 not-for-profit hospitals that are matched to AHA data, and are not missing text information.
 
     Due to the lack of consistency in organization of executive teams, I focus mainly on the executive teams as a whole instead of specific positions.
-I drop anyone who is solely a board member, and focus on those who have executive in their title or are labeled as a president or vice president of a specific aspect of the hospital.
+I drop anyone who is solely a board member, and focus on those who have executive in their title or are labeled as a president or vice president of a department.
 I create various hospital-level characterizations based on their executive teams, such as the existence of a clinically trained executive, the number of clinically trained executives, the number of total executives, and whether the hospital has a Chief Medical Officer.
 
-    \subsection{Other Hospital Characteristics and Outcomes}
+    I merge the created sample of not-for-profits with all for-profits in the AHA survey.
+The main outcomes I consider in this paper are readmission and mortality rates, which are directly targeted in the pay-for-performance policies, and are from the CMS Hospital Compare data.
+I combine readmission and mortality rates for pneumonia, AMI, and heart failure (the relevant HRRP conditions) into a weighted average based on the number of patients in each condition.
+I also include as outcomes uncompensated care and case mix index (a measure of patient complexity).
+While not directly targeted by the policies, uncompensated care is a mechanism by which a hospital might re-allocate resources by taking on less charitable cases, and case mix index reflects potential gamesmanship in the type of patients admitted to the hospital.
+Uncompensated care comes from the Healthcare Cost Information Reporting Systems (HCRIS) data, and case mix index is from CMS public use files.
 
-    I include hospital information from the AHA survey on the number of beds in the hospital and the Medicare number, which I use to merge the Hospital Compare and Healthcare Cost Reporting Information System (HCRIS) data.
-Further, I use the AHA categorization of for-profit or nonprofit.
-Hospital Compare contains information on penalties from HRRP and readmission and mortality rates for the relevant HRRP conditions: heart failure, heart attack (AMI), and pneumonia.
-From HCRIS, I gather information on penalties from the HVBP program, operating expenses, and uncompensated care.
-I also link to the publicly available CMS file containing each hospital's case mix index.
+    Additionally, I include several relevant hospital characteristics.
+Beds, whether the hospital is an academic medical center, owns subsidiaries, is physician owned or system affiliated, come from the AHA survey.
+Whether the hospital was penalized under HRRP or received payments under HVBP is found in the HCRIS data.
+Finally, I limit to years 2010-2014.
 
-    The Hospital Compare excess readmissions files contain information on the excess readmission measure used by CMS to determine whether a hospital should be penalized.
-Specifically, CMS calculates the average readmission rate of all hospitals, adjusts for demographic factors, and determines the amount of readmissions a hospital has in excess to of the average.
-This variable is condition specific, where a hospital is penalized if any of the three excess readmission ratios are greater than one\footnote{Additional conditions were added to the penalty calculations in 2015, but I do not consider these since my sample does not extend past 2015}.
-Further, Hospital Compare also publishes hospital level readmission and mortality rates for heart failure, AMI, and pneumonia.
-For my main outcomes of interest, I take an average across conditions, weighted by the total number of patients seen for that condition.
-
-    \subsection{Summary Statistics}
-
-    \import{Tables}{overall_sumstats.tex}
-
-    \import{Tables}{sample_sumstats.tex}
-
-    \import{Tables}{stable_sample_sumstats.tex}
-
-    In Table \ref{sumstats}, I present summary statistics for hospital variables.
+    % In Table \ref{sumstats}, I present summary statistics for hospital variables.
 There are 3,766 total hospitals in the sample, with 150 beds on average.
@@
 
-    I present a table of means for each hospital type (for-profit, all nonprofits, nonprofits with a clinically trained executive, and nonprofits without a clinically trained executive) in Table \ref{tab:sumstats_samples}.
-Eighty-five percent of hospital in the sample are classified as nonprofits.
-Importantly, the two sub-samples of nonprofits do not add up to the total number of nonprofits because of the limited ability to extract leadership text for all nonprofits.
-There are key differences in nonprofits with and without clinically trained executives: those with MDs are larger, more likely to be penalized, have higher uncompensated care, and more complex patients.
+    I present a table of means for each hospital type (for-profit, in-sample not-for-profits, and not-for-profits with and without a clinically trained executive at some point in time) in Table \ref{tab:sumstats_samples}.
+Half of the hospitals in the sample are classified as not-for-profit with leadership information.
+There are key differences in not-for-profits with and without clinically trained executives: those with MDs are larger, more likely to be penalized, have higher uncompensated care, and more complex patients.
+I discuss the implications of these differences for identification in Section \ref{sec:model}.
 
-    %\import{Tables}{sample_sumstats.tex}
+    \import{Tables}{sample_sumstats.tex}
 
     I plot readmission and mortality rates across time for different hospital types in Figure \ref{fig:weighted_read_mort_graph}.
-While this is purely descriptive evidence, the trends in outcomes point to more similar behavior between for-profits and nonprofits without clinically trained executives in comparison to nonprofits with a clinically trained executive.
+For-profit hospitals have the highest readmission rates prior to 2012.
+Not-for-profits with MDs have readmission rates that fall between for-profits and not-for-profits without MDs.
+While the trend in readmission rates behaves similarly across hospital types prior to the programs taking affect, there are differential trends after the programs.
+NFPs with an MD have the lowest mortality rates before the programs, but differentially continue with an increasing trend in mortality after the programs.
+While this is purely descriptive evidence, the trends in outcomes point to more similar behavior between for-profits and NFPs without clinically trained executives in comparison to NFPs with a clinically trained executive.
 
@@
     \centering
-        \caption{Outcomes over time in all penalized hospitals}
+        \caption{Outcomes Over Time}
         \includegraphics[width=\textwidth]{Objects/weighted_read_mort_graph.pdf}
@@
 
-    \subsection{Leadership Correlation with Other Hospital Characteristics}
+    \section{Model and Empirical Strategy}\label{sec:model}
 
-    Finally, I present a table of correlations between having at least one clinically trained executive and other time invariant hospital characteristics.
-There may be a concern that this particular characteristic of executive teams is simply a proxy for another quality, such as being an academic medical center.
-Unsurprisingly, the most correlated characteristic with having a clinically trained executive is having a CMO on the executive team.
-However, even this correlation is only .4, meaning there is still variation in having clinically trained executives even when controlling for having a CMO.
+    I model hospital behavior in two worlds: one where quality does not directly affect profit, and one where it does.
+Pay-for-performance incentives essentially incorporate quality of care into a hospital's profit function.
+Intuitively, a hospital that places more weight on profit (for-profit type) will only care about quality when it enters into the profit function, and therefore will respond more drastically than a hospital who already cared about quality before the incentives.
+Therefore, hospital response to pay-for-performance incentives reveals something about their underlying preference on profit vs.
+societal outcomes.
 
-    \import{Tables}{hosp_correlations.tex}
+    Formally, I model the hospital objective function as a weighted average of profit and societal (or patient) outcomes.
+For simplicity, I abstract away from quality having a direct effect of quantity and price, though these are possible and don't change the important findings here.
+Without any financial incentives to increase quality, this looks like:
+
+    $$\max_{\theta}\hspace{2mm}\alpha\left[R - c_{\theta}(\theta) \right] + (1-\alpha) u(\theta),$$
+
+    \noindent where $R$ is net revenue, $c_{\theta}(.)$ is an increasing cost function specific to quality $\theta$, $u(.)$ is extra utility gained from quality, which is increasing and concave in $\theta$, and $\alpha\in[0,1]$ captures how much the hospital cares about profit vs.
+societal benefit.
+There is an implicit stay open condition throughout the model.
+Taking the first order condition yields
+
+    $$(1-\alpha)u'(\theta) = \alpha c_{\theta}'(\theta),$$
+
+    \noindent marginal benefit equals marginal cost of increasing quality. Solving for $u'(\theta)$ and differentiating with respect to $\alpha$ yields
+
+    $$\frac{du'(\theta)}{d\alpha} = \frac{1}{(1-\alpha)^2}c_{\theta}'(\theta) > 0.$$
+
+    \noindent Thus, by the Implicit Function Theorem,
+
+    $$\frac{d\theta}{d\alpha} = \frac{du'(\theta)/d\alpha}{du'(\theta)/d\theta} < 0.$$
+
+    \noindent That is, the more weight on profit, the lower quality of care they will provide.
+
+    In a world with pay-for-performance incentives, which can either look like benefits from high quality (such as HVBP) or penalties for low quality (as in HRRP), quality directly affects revenue.
+Thus, $R(\theta)$ is an increasing function of $\theta$, and I assume that the marginal financial benefit of increasing quality is greater than the marginal cost ($R'(\theta)\geq c_{\theta}'(\theta))$.
+Thus, the new first order condition yields:
+
+    $$u'(\theta) = \frac{\alpha}{1-\alpha} \left[c_{\theta}'(\theta)-R(\theta)\right] \leq 0.$$
+
+    \noindent Taking the derivative of this with respect to $\alpha$,
+
+    $$\frac{du'(\theta)}{d\alpha} = \frac{1}{(1-\alpha)^2}c_{\theta}'(\theta).$$
+
+    Thus,
+
+    $$\frac{d\theta}{d\alpha} = \frac{du'(\theta)/d\alpha}{du'(\theta)/d\theta}\geq0.$$
+
+    Therefore, in a world with financial incentives on quality, quality is increasing with more weight placed on profit.
+The purpose of this paper is to investigate the response of different types of hospitals when moving from a world closer to this stylized world without pay-for-performance to the stylized world with pay-for-performance.
+Hence, I combine these results into one response function that depends on $\alpha$, where $\theta_2$ is quality under pay-for-performance incentives and $\theta_1$ is quality with no financial incentive to quality:
+
+    \begin{align*}
+        \frac{d\Delta\theta}{d\alpha}&=\frac{d(\theta_2-\theta_1)}{d\alpha}\\
+        &=\frac{d\theta_2}{d\alpha}-\frac{d\theta_1}{d\alpha}\\
+        &\geq 0.
+    \end{align*}
+
+    The theoretical prediction is that change in quality depends on how much weight the hospital places on profit vs.
+societal benefit.
+Particularly, hospitals with more weight on profit respond more than hospitals with more weight placed on societal benefit.
+In reality, $\alpha$ is unknown for many hospitals.
+What is observed is ownership status, where for-profit hospitals have revealed that they have a high $\alpha$.
+However, not-for-profit behavior on average does not bring much clarity about their objectives (see Section for the literature review).
+Thus, I compare the response to incentive change between for-profits and not-for-profits with different leadership teams to bring clarity on how these things point to an underlying $\alpha$.
+
+    \subsection{Estimation}
+
+    In light of the theoretical predictions, I first test empirically whether different sub-samples of not-for-profits respond to program incentives differently than for-profits.
+I estimate the following regression equation:
+
+    \begin{equation}
+    \label{eq:forprofit}
+    y_{ht} = \beta (\text{For-Profit}_h \times \text{Post Programs}_t) + \gamma_{h} + \delta_t + \epsilon_{ht}
+    \end{equation}
+
+    \noindent where $y_{ht}$ is one of the outcome variables discussed in Section \ref{sec:data}, $\beta$ is the coefficient of interest capturing the effect of being for-profit after the programs were introduced, and $\gamma_h$ and $\delta_t$ are hospital and time fixed effects, respectively.
+For each outcome, I estimate three models where the comparison group differs by hospital type: all NFPs, NFPs with a clinically trained executive, and NFPs without a clinically trained executive.
+I also directly estimate the difference in response of NFPs with and without a clinically trained executive:
+
+    \begin{equation}
+    \label{eq:clinical}
+    y_{ht} = \beta (\text{No MD Exec}_h \times \text{Post Programs}_t) + \gamma_{h} + \delta_t + \epsilon_{ht}
+    \end{equation}
+
+    An important distinction is whether the classification of NFPs by leadership team is time-varying.
+For overall ownership status, switching is rare, and I limit to those who never switch.
+However, hospital executive teams are dynamic, and whether or not a hospital employs a clinically trained executive in each period may change.
+Therefore, I first present effects of having a clinically trained executive among stable executive teams, i.e., NFPs with no change in their propensity to hire a clinically trained executive.
+
+    As discussed in Section \ref{sec:hrrp}, hospitals had incentive to decrease readmissions immediately when the HRRP rules were released in October of 2011, before the penalties were actually determined in October of 2012.
+However, I still define post penalty as occurring in 2012 or later, the first full of hospital response.
+
+    To identify the causal effect of hospital type on hospital program response, I rely on several assumptions.
+First, I assume that executive teams are not determined endogenously when the programs were enacted.
+I investigate the validity of this assumption in Section \ref{sec:endog}.
+Second, I assume that no other unobserved changes occurred in 2012 that are correlated with both hospital performance and type.
+
+    Finally, I assume that, absent the program enactments, different hospital types would have had parallel trends in outcomes.
+While event studies show no difference in pre-trends (see appendix blank), Table \ref{tab:sumstats_samples} shows that hospitals with a clinically trained executive are different than other hospitals in multiple characteristics, and the number of hospitals with clinically trained executives is relatively small.
+These together may cause concern that there are systematic differences between hospital types that could be driving the results.
+Therefore, my main specification uses the synthetic difference-in-differences method established in \citeauthor{arkhangelsky2021synthetic} (\citeyear{arkhangelsky2021synthetic}).
+This method creates unit weights that more heavily count comparison hospitals that have similar trends in outcomes prior to the programs.
+It also weighs time periods that balance pre-program with post-program for the comparison hospitals.
+Then, a simple two way fixed effects estimator is used with these weights.
+This is a more robust average treatment effect that relies less on a strong parallel trends assumption (\cite{arkhangelsky2021synthetic}).
+I present the synthetic diff-in-diff estimates as the main results, and they are largely similar to the standard two way fixed effects results in Appendix \ref{app:fullsample}.
+
+    \subsection{Identification}
+
+    While ultimately I employ a synthetic difference-in-differences method, the conceptual identification is identical to a standard difference-in-differences.
+
+    \subsection{Decomposition}
+
+    Additionally, under the assumption that these changes are not correlated with the program enactments (see Section \ref{sec:endog}), using the changes can actually be advantageous in better understanding underlying hospital motives.
+There are two reasons we might see having a clinically trained executive affecting outcomes: first, a clinically trained executive could be a signal of the underlying weight placed on profit, $\alpha$.
+Second, hiring a clinically trained executive who carries out day-to-day operations in a distinct way effectively changes $\alpha$.
+Thus, I use two types of variation in clinically trained executives to disentangle the two: a variable for whether the hospital ever hires a clinically trained executive, and a variable for whether the hospital has a clinically trained executive in 2012 when the programs took place.
+I use the same empirical specification with these two variables as the independent variation in executive teams, and limit the sample to the relevant comparison group:
+
+    \begin{equation}
+    \label{eq:decomp}
+    y_{ht} = \beta (\text{Never/No 2012 Exec}_h \times \text{Post Programs}_t) + \gamma_{h} + \delta_t + \epsilon_{ht}
+    \end{equation}
 
@@
 
-    \section{Model and Empirical Strategy}\label{sec:model}
-
-    I model hospital behavior as maximizing some weighted average of profit and societal (or patient) benefit.
-Abstracting away from quantity and price as functions of quality for the purpose of this paper, I present a two-period model of hospital behavior that captures the change in quality that occurs after pay-for-performance incentives are introduced.
-
-    \subsection{Pre-Program Incentives}
-
-    In the first period, prior to program enactment, hospitals do not expect any penalties in the future, and choose quality according to:
-
-    $$\max_{\theta}\hspace{2mm}\alpha\left[R - c_{\theta}(\theta) \right] + (1-\alpha) u(\theta),$$
-
-    \noindent where $R$ is net revenue, $c_{\theta}(.)$ is an increasing cost function specific to quality $\theta$, $u(.)$ is extra utility gained from quality, which is increasing and concave is $\theta$, and $\alpha\in[0,1]$ captures how much the hospital cares about profit vs.
-societal benefit.
-There is an implicit stay open condition throughout the model.
-Taking the first order condition yields
-
-    $$(1-\alpha)u'(\theta) = \alpha c_{\theta}'(\theta),$$
-
-    \noindent marginal benefit equals marginal cost of increasing quality. Solving for $u'(\theta)$ and differentiating with respect to $\alpha$ yields
-
-    $$\frac{du'(\theta)}{d\alpha} = c_{\theta}'(\theta)\frac{1}{(1-\alpha)^2} > 0.$$
-
-    \noindent Thus, by the Implicit Function Theorem,
-
-    $$\frac{d\theta}{d\alpha} = \frac{du'(\theta)/d\alpha}{du'(\theta)/d\theta} < 0.$$
-
-    \noindent That is, the more weight the hospital places on profit, the lower quality of care they will provide.
-
-    \subsection{Post-Program Incentives}
-
-    For simplicity, I model pay-for-performance incentives as a simplified version of how they were applied in HRRP, as penalties for low quality hospitals.
-However, these results still hold when including incentive payments as in HVBP.
-
-    Hospitals determine period 1 quality, $\theta_1$, according to the optimal $\theta(\alpha)$ described above.
-Then, in period two, hospitals are penalized or not depending on $\theta_1$, and they choose a new quality, $\theta_2$, which will determine penalties in the future.
-The penalty $(k)$ comes in the form of a negative shock to revenue for hospitals with quality below a certain exogenous threshold, $\overline{\theta}$.
-That is, new revenue is:
-
-    $$R_2 = \begin{cases}
-        R - k, \text{  if  } \theta_1<\overline{\theta}\\
-        R, \text{  if  } \theta_1\geq \overline{\theta}.
-    \end{cases}$$
-
-    The penalty $k$ is large and exceeds the maximum cost of quality.
-Therefore, all hospitals will choose $\theta_2$ to not be penalized in the future.
-Hospitals who were penalized will increase quality just to the point of not being penalized, and hospitals who were not penalized will not change their quality.
-Thus,
-
-    $$\theta_2 = \begin{cases}
-        \overline{\theta}, \text{  if  } \theta_1<\overline{\theta}\\
-        \theta_1, \text{  if  } \theta_1\geq \overline{\theta}.
-    \end{cases}$$
-
-    \noindent Taking the change in theta from period 1 to period 2, and then differentiating with respect to $\alpha$, yields:
-
-    $$\Delta\theta = \theta_2-\theta_1 = \begin{cases}
-        \overline{\theta}-\theta_1, \text{  if  } \theta_1<\overline{\theta}\\
-        0, \text{  if  } \theta_1\geq \overline{\theta}.
-    \end{cases}$$
-
-    $$\frac{d\Delta\theta}{d\alpha} = \begin{cases}
-        -\frac{d\theta_1}{d\alpha}, \text{  if  } \theta_1<\overline{\theta}\\
-        0, \text{  if  } \theta_1\geq \overline{\theta}.
-    \end{cases}$$
-
-    I show in period one that $\frac{d\theta}{d\alpha}<0$.
-Therefore, $\frac{d\Delta\theta}{d\alpha}\geq0$.
-That is, the response to the program incentives is weakly increasing in $\alpha$, and there is a threshold for which increasing $\alpha$ (a hospital with more weight on profit), increases the response to the penalty.
-Intuitively, the programs change hospital incentives in that quality now enters directly into the profit function.
-Therefore, hospitals who place more weight on profit will have a greater response to the program incentives.
-
-    The theoretical prediction is that change in quality depends on how much weight the hospital places on profit vs.
-societal benefit.
-However, $\alpha$s are not observed in reality.
-Thus, I empirically investigate whether the response to the program is correlated with hospital ownership or clinical training on the executive team in order to gain understanding about how these things point to an underlying $\alpha$.
-
-    \subsection{Estimation}
-
-    First, I test empirically whether different sub-samples of nonprofits respond to program incentives differently than for-profits.
-While nonprofit hospital objectives are largely unknown, for-profits have revealed their objective in their ownership classification.
-That is, $\alpha=1$ (or is very close).
-The goal in this analysis is to learn something about the underlying $\alpha$ for nonprofit firms, and determine if executive team composition is a characteristic that points to differing $\alpha$s even among nonprofits.
-Thus, I compare their program response to that of for-profits.
-I estimate the following regression equation:
-
-    \begin{equation}
-    \label{eq:forprofit}
-    y_{ht} = \beta (\text{For-Profit}_h \times \text{Post Programs}_t) + \gamma_{h} + \delta_t + \epsilon_{ht}
-    \end{equation}
-
-    \noindent where $y_{ht}$ is one of the outcome variables discussed in Section \ref{sec:data}, $\beta$ is the coefficient of interest capturing the effect of being for-profit after the programs were introduced, and $\gamma_h$ and $\delta_t$ are hospital and time fixed effects, respectively.
-For each outcome, I estimate three models where the comparison group differs by hospital type: all nonprofits, nonprofits with a clinically trained executive, and nonprofits without a clinically trained executive.
-I also directly estimate the difference in response of hospitals with and without a clinically trained executive:
-
-    \begin{equation}
-    \label{eq:clinical}
-    y_{ht} = \beta (\text{No MD Exec}_h \times \text{Post Programs}_t) + \gamma_{h} + \delta_t + \epsilon_{ht}
-    \end{equation}
-
-    An important distinction is whether the classification of nonprofits as those with or without clinically trained executives is time-varying.
-For overall ownership status, switching is rare, and I limit to those who never switch.
-However, hospital executive teams are dynamic, and whether or not a hospital employs a clinically trained executive in each period may change.
-Therefore, I first present effects of having a clinically trained executive among stable executive teams, i.e., hospitals with no change in their propensity to hire a clinically trained executive.
-
-    Additionally, under the assumption that these changes are not correlated with the program enactments (see Section \ref{sec:endog}), using the changes can actually be advantageous in better understanding underlying hospital motives.
-There are two reasons we might see having a clinically trained executive affecting outcomes: first, a clinically trained executive could be a signal of the underlying weight placed on profit, $\alpha$.
-Second, hiring a clinically trained executive who carries out day-to-day operations in a distinct way effectively changes $\alpha$.
-Thus, I use two types of variation in clinically trained executives to disentangle the two: a variable for whether the hospital ever hires a clinically trained executive, and a variable for whether the hospital has a clinically trained executive in 2012 when the programs took place.
-I use the same empirical specification with these two variables as the independent variation in executive teams, and limit the sample to the relevant comparison group:
-
-    \begin{equation}
-    \label{eq:decomp}
-    y_{ht} = \beta (\text{Never/No 2012 Exec}_h \times \text{Post Programs}_t) + \gamma_{h} + \delta_t + \epsilon_{ht}
-    \end{equation}
-
-    As discussed in Section \ref{sec:hrrp}, hospitals had incentive to decrease readmissions immediately when the HRRP rules were released in October of 2011, before the penalties were actually determined in October of 2012.
-However, I still define post penalty as occurring in 2012 or later, the first full of hospital response.
-
-    To identify the causal effect of hospital type on hospital program response, I rely on several assumptions.
-First, I assume that executive teams are not determined endogenously when the programs were enacted.
-I investigate the validity of this assumption in Section \ref{sec:endog}.
-Second, I assume that no other unobserved changes occurred in 2012 that are correlated with both hospital performance and type.
-
-    Finally, I assume that, absent the program enactments, different hospital types would have had parallel trends in outcomes.
-While event studies show no difference in pre-trends (see appendix blank), Table \ref{tab:sumstats_samples} shows that hospitals with a clinically trained executive are different than other hospitals in multiple characteristics, and the number of hospitals with clinically trained executives is relatively small.
-These together may cause concern that there are systematic differences between hospital types that could be driving the results.
-Therefore, my main specification uses the synthetic difference-in-differences method established in \citeauthor{arkhangelsky2021synthetic} (\citeyear{arkhangelsky2021synthetic}).
-This method creates unit weights that more heavily count comparison hospitals that have similar trends in outcomes prior to the programs.
-It also weighs time periods that balance pre-program with post-program for the comparison hospitals.
-Then, a simple two way fixed effects estimator is used with these weights.
-This is a more robust average treatment effect that relies less on a strong parallel trends assumption (\cite{arkhangelsky2021synthetic}).
-I present the synthetic diff-in-diff estimates as the main results, and they are largely similar to the standard two way fixed effects results in Appendix \ref{app:fullsample}.
-
     \section{Results}
@@
 
-     First, I consider the differential program response of for-profit firms compared to the different sub-samples of nonprofit firms.
-These results are presented in Table \ref{tab:forprofit_readmort_synth}.
-Models in the left panel estimate the effect of being for-profit after the programs were enacted, compared to nonprofit sub-samples, on readmission rates.
-There is no difference between the response of for-profits and nonprofits as a whole, or nonprofits without a clinically trained executive.
-However, for-profits lower their readmissions by .3ppts compared to nonprofits without clinically trained executives.
-Models in the right panel estimate the effect of being for-profit after the programs on mortality rates.
-Here, there is no statistical difference in mortality rate response between different hospital types.
+     I consider the differential program response in readmission rates of for-profit firms compared to the different sub-samples of not-for-profit firms (specification \ref{eq:forprofit}).
+First, I compare for-profits to all not-for-profits in the sample in Figure \ref{fig:read_synth_plota}, which shows that on average, there is no difference in how for-profits and not-for-profits change their readmission rates in response to the incentive changes.
+Figures \ref{fig:read_synth_plota} and \ref{fig:read_synth_plotb} compare for-profit readmission rates to NFPs with a clinically trained executive and without a clinically trained executive, respectively.
+Figure \ref{fig:read_synth_plota} shows that for-profit hospitals decrease readmission rates more than not-for-profits with an MD executive after pay-for-performance incentives.
+However, while for-profits have higher readmission rates than NFPs without a clinically trained executive, there is no difference in their change in readmission rates after the incentives are enacted, shown in Figure \ref{fig:read_synth_plotb}.
 
      \begin{figure}
+     \caption{Readmission Rate Synthetic Difference in Differences Results}
      \centering
-         \caption{Readmission Results}
-         \includegraphics[width=\textwidth]{Objects/read_synth_plot.pdf}
+          \begin{subfigure}[b]{0.45\textwidth}
+         \centering
+         \caption{For-Profit and All NFP}
+         \includegraphics[width=\textwidth]{Objects/read_fp_nfp_synth_graph.pdf}
+         \label{fig:read_synth_plota}
+     \end{subfigure}%
+     \hfill
+     \begin{subfigure}[b]{0.45\textwidth}
+         \centering
+         \caption{For-Profit and NFP w/ MD Exec}
+         \includegraphics[width=\textwidth]{Objects/read_fp_md_synth_graph.pdf}
+         \label{fig:read_synth_plotb}
+     \end{subfigure}%
+     \hfill
+     \begin{subfigure}[b]{0.45\textwidth}
+         \centering
+         \caption{For-Profit and NFP w/out MD Exec}
+         \includegraphics[width=\textwidth]{Objects/read_fp_nomd_synth_graph.pdf}
+         \label{fig:read_synth_plotc}
+     \end{subfigure}
+     \hfill
+     \begin{subfigure}[b]{0.45\textwidth}
+         \centering
+         \caption{NFP w/ and w/out MD Exec}
+         \includegraphics[width=\textwidth]{Objects/read_md_nomd_synth_graph.pdf}
+         \label{fig:read_synth_plotd}
+     \end{subfigure}
         \label{fig:read_synth_plot}
@@
 
-     Next, I directly estimate the difference in response by nonprofits with and without clinically trained executives.
-These results are presented in Table \ref{tab:MD_noMD_readmort_synth}, where the left has models with outcome readmission rates and the right has models with outcome mortality rates.
-Nonprofits without clinically trained executives have a larger decrease in readmissions than those with clinically trained executive by .3ppts, while there is no difference in change in mortality rates.
-Tables \ref{tab:forprofit_readmort_synth} and \ref{tab:MD_noMD_readmort_synth} taken together show that, in terms of readmission rates, nonprofits without clinically trained executives behave more similarly to for-profits than they do other nonprofits.
-Under the assumption that having a clinically trained executive on the leadership team indicates a higher preference for social utility, these findings are consistent with the theoretical prediction presented in Section \ref{sec:model}.
+    Finally, in Figure \ref{fig:read_synth_plotd}, I present estimates of Equation \ref{eq:clinical}, directly comparing NFPs with and without clinical executives.
+NFPs without a clinically trained executive lowered readmission rates by .3ppts more than NFPs with a clinically trained executive.
+Relative to the average readmission rates in 2010, this is approximately a 1.5\% decrease, or 11 less readmissions per year.
+
+    Next, I present estimated differences in mortality rates between different hospital types in Figure \ref{fig:mort_synth_plot}.
+Figure \ref{fig:mort_synth_plota} compares for-profit and all NFP mortality rates after the pay-for-performance incentives.
+These results show that there is no statistical difference in mortality rates between for-profits and not-for-profits after the programs.
+However, when considering NFPs with different leadership team compositions, shown in in Figures \ref{fig:mort_synth_plotb} and \ref{fig:mort_synth_plotc}, there seems to be a difference in behavior.
+While mortality rates are significantly more noisy than readmission rates, the coefficient on the difference in mortality rates for for-profit compared to NFP with an MD is -.17, a magnitude similar to that of readmission rates.
+However, for-profits and NFPs without a clinical executive behave similarly.
+
+    I also directly estimate the difference in mortality rates among NFPs with and without clinical experience in Figure \ref{fig:mort_synth_plotd}.
+Again, the difference is not statistically significant, but this is likely due to the noisiness in mortality since the results is similar to that of readmission rates.
+NFP hospital without a clinically trained executive decrease mortality rates more than NFPs without a clinically trained executive due to new pay-for-performance incentives.
 
      \begin{figure}
+     \caption{Mortality Rate Synthetic Difference in Differences Results}
      \centering
-         \caption{Mortality Results}
-         \includegraphics[width=\textwidth]{Objects/mort_synth_plot.pdf}
-         \label{fig:read_synth_plot}
+          \begin{subfigure}[b]{0.45\textwidth}
+         \centering
+         \caption{For-Profit and All NFP}
+         \includegraphics[width=\textwidth]{Objects/mort_fp_nfp_synth_graph.pdf}
+         \label{fig:mort_synth_plota}
+     \end{subfigure}%
+     \hfill
+     \begin{subfigure}[b]{0.45\textwidth}
+         \centering
+         \caption{For-Profit and NFP w/ MD Exec}
+         \includegraphics[width=\textwidth]{Objects/mort_fp_md_synth_graph.pdf}
+         \label{fig:mort_synth_plotb}
+     \end{subfigure}%
+     \hfill
+     \begin{subfigure}[b]{0.45\textwidth}
+         \centering
+         \caption{For-Profit and NFP w/out MD Exec}
+         \includegraphics[width=\textwidth]{Objects/mort_fp_nomd_synth_graph.pdf}
+         \label{fig:mort_synth_plotc}
+     \end{subfigure}
+     \hfill
+     \begin{subfigure}[b]{0.45\textwidth}
+         \centering
+         \caption{NFP w/ and w/out MD Exec}
+         \includegraphics[width=\textwidth]{Objects/mort_md_nomd_synth_graph.pdf}
+         \label{fig:mort_synth_plotd}
+     \end{subfigure}
+        \label{fig:mort_synth_plot}
     \end{figure}
@@
 
-    However, it is not clear whether these results are due to clinically trained executives being a signal of underlying hospital objectives, or that the leaders themselves make a difference in outcomes.
-Therefore, I present a decomposition of the results using two models in Table \ref{tab:MD_noMD_readmort_decomp_synth}, where the variation in executives is either whether the hospital ever has a clinically trained executive, or whether the hospital has a clinically trained executive in 2012 when the programs were enacted.
-The left panel on readmission rates shows that the difference is response is fully driven by having a clinically trained executive in 2012 compared to having a clinically trained trained executive at another point in time.
-Again, there is nothing going on with mortality.
+    While it is telling that executive team composition affects behavior after incentives change, more clarity is needed about the root cause of the differences.
+It is not clear whether these results are due to clinically trained executives being a signal of underlying hospital objectives, or the leaders themselves making a difference in outcomes.
+Therefore, I present the results of specification \ref{eq:decomp} in Table \ref{tab:MD_noMD_readmort_decomp_synth}, where the variation in executives is either existence of a hospital executive at any point in time, or whether the hospital has a clinically trained executive in 2012 when the programs were enacted.
+The left panel on readmission rates shows that the difference in response is fully driven by having a clinically trained executive in 2012 compared to having a clinically trained executive at another point in time.
+The right panel shows similar magnitudes of coefficients with more noise, as in the main specification.
 
@@
 The first is changing input into another societal benefit measure: uncompensated care, and the second is complexity of patients: case mix index.
-First, I investigate whether for-profits change these differentially than types of nonprofits in Table \ref{tab:forprofit_uncompCMI_synth}.
+First, I investigate whether for-profits change these differentially than types of not-for-profits in Table \ref{tab:forprofit_uncompCMI_synth}.
 
@@
 
-    Next, I estimate the difference in response for nonprofits with and without clinically trained executives, shown in Table \ref{tab:MD_noMD_uncompCMI_synth}.
+    Next, I estimate the difference in response for not-for-profits with and without clinically trained executives, shown in Table \ref{tab:MD_noMD_uncompCMI_synth}.
 
@@
     In this paper, I study differential objective between different firm types.
-While the differences in for-profit and nonprofit hospitals have been studied in the past, it is unclear whether certain nonprofit characteristics make them act more like for-profits.
-In order to better understand the role of leadership in objectives, I estimate the differential response to a change in incentives among for-profits, nonprofits, and nonprofits with and without a clinically trained executive executives.
-I find that nonprofits without clinically trained executives respond to incentives much more similarly to for-profits than to nonprofits without clinically trained executives.
+While the differences in for-profit and not-for-profit hospitals have been studied in the past, it is unclear whether certain not-for-profit characteristics make them act more like for-profits.
+In order to better understand the role of leadership in objectives, I estimate the differential response to a change in incentives among for-profits, not-for-profits, and not-for-profits with and without a clinically trained executive executives.
+I find that not-for-profits without clinically trained executives respond to incentives much more similarly to for-profits than to not-for-profits without clinically trained executives.
 
@@
 There is no perfect way to access tax form 990s in bulk over the time period I am considering in this paper.
-However, using the NonProfit Explorer API seems to be the most straightforward.
-At the time of writing this, information on using version 2 of the API can be found at \hyperlink{https://projects.propublica.org/nonprofits/api}{https://projects.propublica.org/nonprofits/api}.
+However, using the not-for-profit Explorer API seems to be the most straightforward.
+At the time of writing this, information on using version 2 of the API can be found at \hyperlink{https://projects.propublica.org/not-for-profits/api}{https://projects.propublica.org/not-for-profits/api}.
 
-There are over 1.5 million nonprofit entities in the US (cite), making it crucially important to be able to filter by type of entity before analyzing any PDFs.
+There are over 1.5 million not-for-profit entities in the US (cite), making it crucially important to be able to filter by type of entity before analyzing any PDFs.
 The API allows this by filtering a query based on National Taxonomy of Exempt Entities (NTEE) code.
-I query only nonprofits categorized as E20 (hospitals), E21 (community health systems), and E22 (general hospitals).
+I query only not-for-profits categorized as E20 (hospitals), E21 (community health systems), and E22 (general hospitals).
 The API has a pagination limit of 100, meaning I can only pull information on 100 hospitals at a time.
@@
 The only state that has more than 100 entities registered is California, and thus I subset the California query even further by names that include the word "hospital" and names that don't.
-I combine all of these subsets and have information on each nonprofits Employee Identification Number.
+I combine all of these subsets and have information on each not-for-profits Employee Identification Number.
 There are 5,588 EINs total in this list.
@@
 Thus, I finish this step with a panel data set of EIN characteristics and PDF locations.
-Importantly, there are multiple types of Tax Form 990s depending on the size of the nonprofit.
-In many cases, one nonprofit has at least two different forms filed in a given year.
+Importantly, there are multiple types of Tax Form 990s depending on the size of the not-for-profit.
+In many cases, one not-for-profit has at least two different forms filed in a given year.
 I filter out any EIN-years for which there are no PDFs on file.
@@
 
-It is crucial to the analysis to be able to link these nonprofits with other sources of data to recover penalties from HRRP, bed size, and outcomes of interest (definitely need to talk more about outcomes).
+It is crucial to the analysis to be able to link these not-for-profits with other sources of data to recover penalties from HRRP, bed size, and outcomes of interest (definitely need to talk more about outcomes).
 Therefore, I take a conservative approach to matching EINs to American Hospital Association (AHA) ID, which links fairly easily to Medicare ID numbers, based on name and location.
@@
 I download the AHA data from Wharton Research Data Services.
-I filter only to hospitals in the contiguous US, Alaska, and Hawaii (excluding places like Puerto Rico), classified as nonprofit or state/community, and those that are general acute care.
+I filter only to hospitals in the contiguous US, Alaska, and Hawaii (excluding places like Puerto Rico), classified as not-for-profit or state/community, and those that are general acute care.
 I also filter out any hospitals who weren't present in the data (or change system ID) in 2009-2015, meaning they either closed or were acquired.
@@
 
-I consider which nonprofit entities are not likely to be hospitals and drop them.
-There are numerous foundations or auxiliary nonprofits with the purpose of raising funds for the hospital, but do not actually care for patient.
-I filter out any nonprofit with "foundation" or "auxiliary" in the name.
+I consider which not-for-profit entities are not likely to be hospitals and drop them.
+There are numerous foundations or auxiliary not-for-profits with the purpose of raising funds for the hospital, but do not actually care for patient.
+I filter out any not-for-profit with "foundation" or "auxiliary" in the name.
 I also filter out various specialty centers that fell into the general hospital category, such as hospice or cancer centers.
@@
 I focus on exact string matches, so I remove all spaces and common characters that could cause mismatches such as \&, ', -, and inc.
-Next, I take each AHA name and look for exact matches in a nonprofit's first or secondary name only for nonprofits in the same state as the AHA hospital.
+Next, I take each AHA name and look for exact matches in a not-for-profit's first or secondary name only for not-for-profits in the same state as the AHA hospital.
 When an exact match is found, I record the link between AHA ID and EIN.
-In this first layer of matching, 860 hospitals in the AHA data are linked to an EIN, equivalent to 31\% of AHA nonprofit hospitals in the sample.
+In this first layer of matching, 860 hospitals in the AHA data are linked to an EIN, equivalent to 31\% of AHA not-for-profit hospitals in the sample.
 
@@
 That way if there are subtle differences in names, removing common words may allow for an exact match.
-Again, I take each AHA hospital name and look for exact matches in the nonprofits within the same state.
+Again, I take each AHA hospital name and look for exact matches in the not-for-profits within the same state.
 This adds an additional 90 hospital matches, accounting for a total of 34.5\% of AHA hospitals.
@@
 These hospitals become the main sample for my analysis.
-While not capturing a large majority of AHA hospitals that are categorized as nonprofit, I would rather be conservative in selecting hospitals which I am confident that the characteristics and measurement of the independent and dependent variables are accurate.
+While not capturing a large majority of AHA hospitals that are categorized as not-for-profit, I would rather be conservative in selecting hospitals which I am confident that the characteristics and measurement of the independent and dependent variables are accurate.
 
@@
 
-One tricky aspect of the NonProfit Explorer API is that, only in some cases, if two forms are present for an EIN, year, only the first one (which is typically not the one with the relevant information) is pulled.
+One tricky aspect of the not-for-profit Explorer API is that, only in some cases, if two forms are present for an EIN, year, only the first one (which is typically not the one with the relevant information) is pulled.
 Therefore, for some hospitals, a couple years will have gaps in text extraction data.
@@
 \includegraphics[width=\textwidth]{Objects/has_doc_avg_map.pdf}
+
+\section{Hospitals Matched to AHA vs. Not}\label{app:matched}
 

### 2024-05-15 Paper/paper.tex
Updates from Overleaf

@@
         In many industries, not-for-profits (NFP) make up the majority of firms in the market.
-Understanding the objectives and behaviors of these firms is important when changing incentives.
-However, NFP firms have goals and missions that make understanding their objectives difficult.
+Understanding the objectives of these firms is important as they respond to incentives, but NFP firms have underlying goals that make understanding their objectives difficult.
 For hospitals in particular, there is mixed evidence on whether NFP hospitals differ meaningfully from for-profit (FP) hospitals.
-In this paper, I investigate whether leadership team composition of not-for-profit hospitals affects how similarly they behave to for-profits.
+In this paper, I investigate whether leadership team composition of NFP hospitals affects how similarly they behave to for-profits.
 I use a novel data set on NFP hospital executives in the US (2009-2015) to investigate whether the presence of clinically trained executives affects hospital behavior.
-In particular, I leverage two pay-for-performance initiatives in the US enacted in 2012, and estimate the difference in readmission and mortality rates afterwards between FP, NFP, and NFP with and without clinically trained executives.
+In particular, I leverage two pay-for-performance initiatives in the US enacted in 2012, and measure the  readmission and mortality rate responses between FP, NFP, and NFP with and without clinically trained executives.
 I find that NFPs without clinically trained executives behave more like FPs than other NFPs.
@@
 
-    One reason that hospital executive teams are understudied is because of the inaccessibility of the information.
-I construct a novel data set of not-for-profit hospitals in the US that contains names, titles, and positions of all executives tied to the hospital from 2009-2015.
+    Hospital executive teams are understudied partly because of the inaccessibility of granular information.
+I construct a novel data set of not-for-profit hospitals in the US that contains names, titles, and positions of all board members and executives tied to the hospital from 2009-2015.
 I gather this data from publicly available Tax Form 990s, which every sufficiently large not-for-profit files each year in the US.
@@
 
-    The historical tax forms for all not-for-profits are housed by ProPublica.\footnote{https://projects.propublica.org/not-for-profits/} I use the not-for-profit Explorer API to extract Employee Identification Numbers (EIN) of not-for-profits classified as a hospital according to their National Taxonomy of Exempt Entities (NTEE) code.
+    The historical tax forms for all not-for-profits are housed by ProPublica.\footnote{https://projects.propublica.org/not-for-profits/} I use the NonProfit Explorer API to extract Employee Identification Numbers (EIN) of not-for-profits classified as a hospital according to their National Taxonomy of Exempt Entities (NTEE) code.
 After filtering out associations and specialty hospitals, there are approximately 3,000 not-for-profit EINs.
@@
 I confidently match 1200 EINs to an AHA identification number based on exact name matches within the same state.
-I compare the matched hospitals to non-matched hospitals in Appendix \ref{app:matched}, and find that the matched sample is not qualitatively different than the non-matched hospitals.
+I compare the matched hospitals to non-matched hospitals in Appendix \ref{app:matched}.
 
@@
 
-    Due to the lack of consistency in organization of executive teams, I focus mainly on the executive teams as a whole instead of specific positions.
+    Due to the lack of consistency in the organization of executive teams, I focus mainly on the executive teams as a whole instead of specific positions.
 I drop anyone who is solely a board member, and focus on those who have executive in their title or are labeled as a president or vice president of a department.
 I create various hospital-level characterizations based on their executive teams, such as the existence of a clinically trained executive, the number of clinically trained executives, the number of total executives, and whether the hospital has a Chief Medical Officer.
+Summary statistics for these variables are shown in Table
 
@@
         \caption{Outcomes Over Time}
-        \includegraphics[width=\textwidth]{Objects/weighted_read_mort_graph.pdf}
+        \includegraphics[width=\textwidth]{Objects/weighted_read_mort_adjusted_graph.pdf}
         \label{fig:weighted_read_mort_graph}
+        \caption*{\footnotesize{Averages are adjusted for patient complexity.}}
     \end{figure}
@@
 
-     \subsection{Readmission and Mortality Rates}
-
      I consider the differential program response in readmission rates of for-profit firms compared to the different sub-samples of not-for-profit firms (specification \ref{eq:forprofit}).
@@
 
-    \subsubsection{Decomposition}
+    \subsection{Decomposition}
 
@@
     \import{Tables}{MD_noMD_readmort_decomp_synth.tex}
-
-    \subsection{Uncompensated Care and Case Mix Index}
-
-    As a secondary analysis, I present estimates of differential response of potential mechanisms.
-The first is changing input into another societal benefit measure: uncompensated care, and the second is complexity of patients: case mix index.
-First, I investigate whether for-profits change these differentially than types of not-for-profits in Table \ref{tab:forprofit_uncompCMI_synth}.
-
-    \import{Tables}{forprofit_uncompCMI_synth.tex}
-
-    Next, I estimate the difference in response for not-for-profits with and without clinically trained executives, shown in Table \ref{tab:MD_noMD_uncompCMI_synth}.
-
-    \import{Tables}{MD_noMD_uncompCMI_synth.tex}
-
-    \subsubsection{Decomposition}
-
-    \import{Tables}{MD_noMD_uncompCMI_decomp_synth.tex}
 

### 2024-05-17 Paper/paper.tex

@@
 I create various hospital-level characterizations based on their executive teams, such as the existence of a clinically trained executive, the number of clinically trained executives, the number of total executives, and whether the hospital has a Chief Medical Officer.
-Summary statistics for these variables are shown in Table
+Summary statistics for these variables are shown in Table \ref{leader_sumstats}.
+
+    \import{Tables}{leader_sumstats.tex}
 
     I merge the created sample of not-for-profits with all for-profits in the AHA survey.
-The main outcomes I consider in this paper are readmission and mortality rates, which are directly targeted in the pay-for-performance policies, and are from the CMS Hospital Compare data.
+The outcomes I consider are readmission and mortality rates, which are directly targeted in the pay-for-performance policies, and are from the CMS Hospital Compare data.
 I combine readmission and mortality rates for pneumonia, AMI, and heart failure (the relevant HRRP conditions) into a weighted average based on the number of patients in each condition.
-I also include as outcomes uncompensated care and case mix index (a measure of patient complexity).
-While not directly targeted by the policies, uncompensated care is a mechanism by which a hospital might re-allocate resources by taking on less charitable cases, and case mix index reflects potential gamesmanship in the type of patients admitted to the hospital.
-Uncompensated care comes from the Healthcare Cost Information Reporting Systems (HCRIS) data, and case mix index is from CMS public use files.
-
 Additionally, I include several relevant hospital characteristics.
-Beds, whether the hospital is an academic medical center, owns subsidiaries, is physician owned or system affiliated, come from the AHA survey.
+Number of beds, whether the hospital is an academic medical center, is physician owned or is system affiliated, all come from the AHA survey.
+Uncompensated care comes from HCRIS, and case mix index, which measures patient complexity, comes from CMS.
 Whether the hospital was penalized under HRRP or received payments under HVBP is found in the HCRIS data.
@@
 
-    I present a table of means for each hospital type (for-profit, in-sample not-for-profits, and not-for-profits with and without a clinically trained executive at some point in time) in Table \ref{tab:sumstats_samples}.
-Half of the hospitals in the sample are classified as not-for-profit with leadership information.
+    I present a table of means for each hospital type (for-profit, in-sample not-for-profits, and not-for-profits with and without a clinically trained executive) in Table \ref{tab:sumstats_samples}.
+Slightly over half of the hospitals are classified as not-for-profit with leadership information.
 There are key differences in not-for-profits with and without clinically trained executives: those with MDs are larger, more likely to be penalized, have higher uncompensated care, and more complex patients.
@@
 
-    I plot readmission and mortality rates across time for different hospital types in Figure \ref{fig:weighted_read_mort_graph}.
+    I plot patient-complexity adjusted readmission and mortality rates across time for different hospital types in Figure \ref{fig:weighted_read_mort_graph}.
 For-profit hospitals have the highest readmission rates prior to 2012.
-Not-for-profits with MDs have readmission rates that fall between for-profits and not-for-profits without MDs.
 While the trend in readmission rates behaves similarly across hospital types prior to the programs taking affect, there are differential trends after the programs.
-NFPs with an MD have the lowest mortality rates before the programs, but differentially continue with an increasing trend in mortality after the programs.
+NFPs without cilnically trained executives decrease readmissions more similarly to FPs than other NFPs.
+In mortality rates, NFPs with a clinically trained executive have the lowest rates before the programs, but differentially continue with an increasing trend in mortality after the programs.
 While this is purely descriptive evidence, the trends in outcomes point to more similar behavior between for-profits and NFPs without clinically trained executives in comparison to NFPs with a clinically trained executive.
@@
         \label{fig:weighted_read_mort_graph}
-        \caption*{\footnotesize{Averages are adjusted for patient complexity.}}
     \end{figure}
@@
 
-    I model hospital behavior in two worlds: one where quality does not directly affect profit, and one where it does.
-Pay-for-performance incentives essentially incorporate quality of care into a hospital's profit function.
-Intuitively, a hospital that places more weight on profit (for-profit type) will only care about quality when it enters into the profit function, and therefore will respond more drastically than a hospital who already cared about quality before the incentives.
+    I model hospital behavior in two simplified scenarios: where quality does not directly affect profit, and where it does.
+Pay-for-performance incentives essentially incorporate performance directly into a hospital's profit function.
+Here, I think of performance as quality of care, which is inversely related to readmission and mortality rates.
+While the literature speculates about the true form of NFP objective functions, we can generally think of hospitals choosing some linear combination of profit and societal benefit.
+Intuitively, a hospital that places more weight on profit (for-profit type) will only care about quality when it enters into the profit function, and therefore will respond more drastically to quality entering the profit function than a hospital who already cared about quality before the incentives.
 Therefore, hospital response to pay-for-performance incentives reveals something about their underlying preference on profit vs.
@@
     Formally, I model the hospital objective function as a weighted average of profit and societal (or patient) outcomes.
-For simplicity, I abstract away from quality having a direct effect of quantity and price, though these are possible and don't change the important findings here.
-Without any financial incentives to increase quality, this looks like:
+For simplicity, I abstract away from quality having a direct effect of quantity and price, which does not drastically change the main results.
+Without any financial incentives to increase quality, hospital quality is decided by:
 
@@
 
-    \noindent That is, the more weight on profit, the lower quality of care they will provide.
+    \noindent That is, the more weight on profit, the lower quality of care chosen by the hospital.
 
@@
     Therefore, in a world with financial incentives on quality, quality is increasing with more weight placed on profit.
-The purpose of this paper is to investigate the response of different types of hospitals when moving from a world closer to this stylized world without pay-for-performance to the stylized world with pay-for-performance.
-Hence, I combine these results into one response function that depends on $\alpha$, where $\theta_2$ is quality under pay-for-performance incentives and $\theta_1$ is quality with no financial incentive to quality:
+The purpose of this paper is to investigate the response of hospitals with different $\alpha$ when a policy changes hospital incentives using pay-for-performance.
+Hence, I combine the results found in each scenario into one response function that depends on $\alpha$, where $\theta_2$ is quality under pay-for-performance incentives and $\theta_1$ is quality with no financial incentive to quality.
+Thus,
 
@@
 Particularly, hospitals with more weight on profit respond more than hospitals with more weight placed on societal benefit.
-In reality, $\alpha$ is unknown for many hospitals.
-What is observed is ownership status, where for-profit hospitals have revealed that they have a high $\alpha$.
-However, not-for-profit behavior on average does not bring much clarity about their objectives (see Section for the literature review).
-Thus, I compare the response to incentive change between for-profits and not-for-profits with different leadership teams to bring clarity on how these things point to an underlying $\alpha$.
+In reality, $\alpha$ is not revealed by the hospital, except in the case of for-profits, who have revealed that they have a high $\alpha$ by their ownership status.
+However, not-for-profit behavior on average does not bring much clarity about their objectives.
+An observable and economically relevant characteristic of a hospital that could affect $\alpha$ is composition of their leadership team.
+Thus, I compare the response to incentive change between for-profits and not-for-profits with different leadership teams to bring clarity on how this might point to underlying $\alpha$.
 
@@
     \noindent where $y_{ht}$ is one of the outcome variables discussed in Section \ref{sec:data}, $\beta$ is the coefficient of interest capturing the effect of being for-profit after the programs were introduced, and $\gamma_h$ and $\delta_t$ are hospital and time fixed effects, respectively.
-For each outcome, I estimate three models where the comparison group differs by hospital type: all NFPs, NFPs with a clinically trained executive, and NFPs without a clinically trained executive.
+For each outcome, I estimate three models with differing comparison groups: all NFPs, NFPs with a clinically trained executive, and NFPs without a clinically trained executive.
 I also directly estimate the difference in response of NFPs with and without a clinically trained executive:
@@
 However, hospital executive teams are dynamic, and whether or not a hospital employs a clinically trained executive in each period may change.
-Therefore, I first present effects of having a clinically trained executive among stable executive teams, i.e., NFPs with no change in their propensity to hire a clinically trained executive.
+Therefore, for simplicity, I first present effects of having a clinically trained executive among stable executive teams, i.e., NFPs with no change in their propensity to hire a clinically trained executive.
 
-    As discussed in Section \ref{sec:hrrp}, hospitals had incentive to decrease readmissions immediately when the HRRP rules were released in October of 2011, before the penalties were actually determined in October of 2012.
-However, I still define post penalty as occurring in 2012 or later, the first full of hospital response.
-
-    To identify the causal effect of hospital type on hospital program response, I rely on several assumptions.
-First, I assume that executive teams are not determined endogenously when the programs were enacted.
-I investigate the validity of this assumption in Section \ref{sec:endog}.
-Second, I assume that no other unobserved changes occurred in 2012 that are correlated with both hospital performance and type.
-
-    Finally, I assume that, absent the program enactments, different hospital types would have had parallel trends in outcomes.
-While event studies show no difference in pre-trends (see appendix blank), Table \ref{tab:sumstats_samples} shows that hospitals with a clinically trained executive are different than other hospitals in multiple characteristics, and the number of hospitals with clinically trained executives is relatively small.
-These together may cause concern that there are systematic differences between hospital types that could be driving the results.
-Therefore, my main specification uses the synthetic difference-in-differences method established in \citeauthor{arkhangelsky2021synthetic} (\citeyear{arkhangelsky2021synthetic}).
-This method creates unit weights that more heavily count comparison hospitals that have similar trends in outcomes prior to the programs.
-It also weighs time periods that balance pre-program with post-program for the comparison hospitals.
-Then, a simple two way fixed effects estimator is used with these weights.
-This is a more robust average treatment effect that relies less on a strong parallel trends assumption (\cite{arkhangelsky2021synthetic}).
-I present the synthetic diff-in-diff estimates as the main results, and they are largely similar to the standard two way fixed effects results in Appendix \ref{app:fullsample}.
-
-    \subsection{Identification}
-
-    While ultimately I employ a synthetic difference-in-differences method, the conceptual identification is identical to a standard difference-in-differences.
-
-    \subsection{Decomposition}
-
-    Additionally, under the assumption that these changes are not correlated with the program enactments (see Section \ref{sec:endog}), using the changes can actually be advantageous in better understanding underlying hospital motives.
-There are two reasons we might see having a clinically trained executive affecting outcomes: first, a clinically trained executive could be a signal of the underlying weight placed on profit, $\alpha$.
+    Additionally, under certain assumptions that I discuss in Section \ref{sec:identification}, using executive team changes can be advantageous in better understanding underlying hospital motives.
+There are two mechanisms to having a clinically trained executives affecting outcomes: first, a clinically trained executive could be a signal of the underlying weight placed on profit, $\alpha$.
 Second, hiring a clinically trained executive who carries out day-to-day operations in a distinct way effectively changes $\alpha$.
@@
 
-    \subsection{Changes in Executive Team}\label{sec:endog}
+    \subsection{Identification}\label{sec:identification}
 
-    Before presenting the main findings, I investigate whether executive team changes occur endogenously due to the programs.
-If hospitals change their propensity to hire a clinically trained executive because they expect to be penalized or want to gain extra payments, then there is selection in the independent variable.
+    To identify the causal effect of hospital type on hospital program response, I rely on several assumptions.
+First, I assume that, absent the program enactments, different hospital types would have had parallel trends in outcomes.
+While I see no evidence of pre-trends, Table \ref{tab:sumstats_samples} does show that hospitals with a clinically trained executive are different than other hospitals in multiple characteristics.
+Therefore, there may be concern that there are systematic differences between hospital types that could be driving the results.
+Therefore, my main specification uses the synthetic difference-in-differences method established in \citeauthor{arkhangelsky2021synthetic} (\citeyear{arkhangelsky2021synthetic}).
+This method creates unit weights that more heavily count comparison hospitals that have similar trends in outcomes prior to the programs.
+It also weighs time periods that balance pre-program with post-program for the comparison hospitals.
+Then, a simple two way fixed effects estimator is used with these weights.
+This is a more robust average treatment effect that relies less on a strong parallel trends assumption (\cite{arkhangelsky2021synthetic}).
+I present the synthetic diff-in-diff estimates as the main results, and they are largely similar to the standard two way fixed effects results in Appendix \ref{app:fullsample}.
+Further, while ultimately I employ a synthetic difference-in-differences method, the conceptual identification is identical to a standard difference-in-differences.
+
+     Second, I assume that executive teams are not determined endogenously with the programs.
+In the main specification, I limit to executive teams which do not change their propensity to hire an MD over time.
+Thus, I rely on this assumption so that I am not conditioning on an outcome.
+In supplementary analyses, I use executive team changes to disentangle signalling vs.
+managing effects, which is only identified if leadership changes are uncorrelated with the policies.
+
+     I assess the validity of this assumption in two ways.
 First, I analyze whether changes in clinically trained executives are concentrated around the enactment of both HRRP and HVBP, 2011-2012.
@@
     \begin{equation}\label{eq:change1}
-    change_{ht} = \sum_{j=2011}^{2014}\beta_j\mathbf{1}\{t=j\} + \alpha_h + \epsilon_{ht},
+    \text{change}_{ht} = \sum_{j=2011}^{2014}\beta_j\mathbf{1}\{t=j\} + \alpha_h + \epsilon_{ht},
     \end{equation}
@@
     \begin{equation}\label{eq:change2}
-    change_{ht} = \beta(program\_exposed_{h} \times post\_2011)+ \alpha_h + \delta_t + \epsilon_{ht}.
+    \text{change}_{ht} = \beta(\text{program\_exposed}_{h} \times \text{Post Programs}_t)+ \alpha_h + \delta_t + \epsilon_{ht}.
     \end{equation}
 
-    The variable $program\_exposed_{h}$ is an indicator for whether the hospital eventually was penalized under HRRP or received payments under HVBP.
+    The variable $\text{program\_exposed}_{h}$ is an indicator for whether the hospital eventually was penalized under HRRP or received payments under HVBP.
 The estimates from this analysis are presented in Table \ref{tab:change_analysis}.
@@
 Results are similar when considering changes in the number of clinically trained executives.
-Thus, I conclude that changes in executive teams are likely not endogenously occurring with the incentive changes due to the programs.
+Thus, it seems unlikely that endogenous team formation is biasing the estimates.
 
      \import{Tables}{change_analysis.tex}
+
+    Finally, I assume that no other unobserved changes occurred in 2012 that are correlated with both hospital performance and type.
 
@@
 Figures \ref{fig:read_synth_plota} and \ref{fig:read_synth_plotb} compare for-profit readmission rates to NFPs with a clinically trained executive and without a clinically trained executive, respectively.
-Figure \ref{fig:read_synth_plota} shows that for-profit hospitals decrease readmission rates more than not-for-profits with an MD executive after pay-for-performance incentives.
+Figure \ref{fig:read_synth_plotb} shows that for-profit hospitals decrease readmission rates more than not-for-profits with an MD executive after pay-for-performance incentives.
 However, while for-profits have higher readmission rates than NFPs without a clinically trained executive, there is no difference in their change in readmission rates after the incentives are enacted, shown in Figure \ref{fig:read_synth_plotb}.
@@
      \end{subfigure}%
+     \vspace{5mm}
      \hfill
@@
 
+\import{Tables}{NFP_sample_comparison.tex}
+
 \section{Full Sample Regression and Event Study Results}\label{app:fullsample}

### 2024-05-28 Paper/paper.tex

@@
 		{\small
-        In many industries, not-for-profits (NFP) make up the majority of firms in the market.
-Understanding the objectives of these firms is important as they respond to incentives, but NFP firms have underlying goals that make understanding their objectives difficult.
+        In many industries, not-for-profits (NFP) are the dominant firm type.
+Understanding the objectives of these firms is important, but NFP firms have underlying goals that make understanding their objectives difficult.
 For hospitals in particular, there is mixed evidence on whether NFP hospitals differ meaningfully from for-profit (FP) hospitals.
@@
 I use a novel data set on NFP hospital executives in the US (2009-2015) to investigate whether the presence of clinically trained executives affects hospital behavior.
-In particular, I leverage two pay-for-performance initiatives in the US enacted in 2012, and measure the  readmission and mortality rate responses between FP, NFP, and NFP with and without clinically trained executives.
+In particular, I leverage pay-for-performance initiatives in the US enacted in 2012, and measure the  readmission and mortality rate responses between FP, NFP, and NFP with and without clinically trained executives.
 I find that NFPs without clinically trained executives behave more like FPs than other NFPs.
@@
 Not-for-profits (NFP) are thought to gain utility not solely from monetary profit, but also through accomplishing societal benefit.
-Researchers have speculated a variety of objective functions that could motivate NFP firms such as maximizing prestige, income, or a quality/quantity trade-off (\cite{steinberg1986revealed}).
+Researchers have speculated a variety of objective functions that could motivate NFP firms such as maximizing profit (for-profit in disguise), prestige, income, or a quality/quantity trade-off (\cite{steinberg1986revealed}).
 Empirical contributions to this literature center around investigating behaviors that differ between for-profit (FP), private NFP, and government firms (\cite{sloan2000not}).
@@
 
-  While the characterization of being for-profit or not-for-profit is certainly important, there could be other characteristics that affect hospital behavior that are not perfectly correlated with individual ownership status.
-Researchers have focused in particular on explaining the differences in NFP behavior by the market they exist again (citations).
-I argue that an observable and economically meaningful factor is the composition of leadership teams.
-Firm leadership varies widely across firms and potentially drives differences in objectives even within NFPs.
+  While ownership is certainly relevant, there could be other characteristics correlated with hospital behavior independent of ownership type.
+Researchers have focused in particular on explaining variation in NFP behavior by the market they exist in (citations).
+However, I argue that an observable and economically meaningful factor is the composition top-level hospital managers.
 In large, publicly traded firms, executive characteristics are shown to be correlated with firm performance, indicating that there is more to firm behavior than ownership status (\cite{bertrand2003managing}; \cite{matsa2013female}; \cite{ahern2012changing}).
-If hospital objectives differ in how they value monetary profit vs.
-gains to society, then ownership status certainly plays a role in these valuations, but the question remains as to whether management also plays an important role.
+If hospitals differ in how they value monetary profit vs.
+societal benefit, then ownership status certainly plays a role in these valuations, but the question remains as to whether management also plays an important role.
 I contribute to our understanding of hospital and NFP behaviors by investigating whether occupational background, specifically clinical training in executives, affects hospital response to pay-for-performance incentives.
@@
 I link hospitals in this data to characteristics in the American Hospital Association (AHA) survey and Hospital Compare data.
-I leverage various pay-for-performance initiatives in 2012 to investigate differential hospital responses between for-profits, NFPs without clinically trained executives, and NFPs with clinically trained executives.
+I leverage various pay-for-performance initiatives in 2012 to investigate differential hospital responses between for-profits, and NFPs with and without without clinically trained executives.
 
-  I estimate how the interaction between hospital type and post program enactment affects readmission and mortality rates among different groups.
+  I estimate how the interaction between hospital type and post program enactment affects readmission and mortality rates.
 Under the assumption that hospitals would have continued to behave similarly in the absence of the programs, and that leadership teams do not change endogenously with the programs, I identify the differential response to pay-for-performance incentives by different hospital types.
 Among the hospitals in the data, hospitals with clinically trained executives are the minority, and differ on average from other hospitals along various dimensions.
-Thus, my preferred specification is synthetic difference-in-differences, to analyze a group of comparison hospitals with similar characteristics to treatment hospitals.\footnote{I also present results for the full sample, and find qualitatively similar results.}
+Thus, I ultimately employ synthetic difference-in-differences to analyze a group of comparison hospitals with similar characteristics to treatment hospitals.\footnote{I also present results for the full sample, and find qualitatively similar results.}
 
-   While all hospital types decrease readmission rates after the programs are enacted, for-profits and not-for-profits without clinically trained executives decrease readmissions more than not-for-profits with clinically trained executives.
+   While all hospital types decrease readmission rates after the programs are enacted, NFPs with clinically trained executives respond less than both FPs and NFPs without clinically trained executives.
 That is, NFPs without clinically trained executives respond to incentives more similarly to FPs than other NFPs.
-I then consider whether this effect is due to clinically trained executives being a signal of underlying hospital preferences, or changing hospital preferences in a way that then affects behavior.
-I decompose the two by leveraging hospitals who change their propensity to hire a clinically trained executive at some point in the sample, which I find is not occurring endogenously with the change in incentives.
-The results of this analysis suggest that the difference in response is driven by changing hospital preferences, not as a signal of underlying preferences.
-While readmissions and mortality were targeted directly by the programs, I also consider uncompensated care and case mix index as outcomes, since re-allocation of resources and patient selectivity are potential mechanisms to changing hospital outcomes.
-(results of this analysis here)
+Using changes in executive teams, I decompose this effect into signalling (clinical executives signal underlying hospital preferences), or managing (clinical executives manage the hospital differently) effects.
+I find that the entire difference in response is driven by a managing effect.
 
     This paper contributes to three strands of literature.
-First, it adds to our understanding of what underlies NFP hospital behavior.
+First, it adds to our understanding of NFP hospital objective functions.
 There is mixed evidence on whether NFP behavior differs from FP behavior: costs, uncompensated care, technology adoption, and quality (\cite{sloan2000not}; \cite{eggleston2008hospital}; \cite{moscelli2018effect}; \cite{moscone2020public}).
-Underlying hospital are hospital objectives.
-To this point, researchers have found that NFPs do not act as purely profit maximizers, but maximize some combination of profit and output (\cite{deneffe2002not}, \cite{chang2011not-for-profit}).
-A paper with an approach similar to mine which investigates hospital behavior is \citeauthor{chang2011not-for-profit} (\citeyear{chang2011not-for-profit}), which looks at the effect of a cost shock on likelihood of shutting down and mix of profitable services.
+To this point, researchers have found that NFPs do not act as purely profit maximizers, but maximize some combination of profit and output (\cite{deneffe2002not}, \cite{chang2011nonprofit}).
+A paper with an approach similar to mine which investigates hospital behavior is \citeauthor{chang2011nonprofit} (\citeyear{chang2011nonprofit}), which looks at the effect of a cost shock on likelihood of shutting down and mix of profitable services.
 They find that NFPs and FPs are equally as likely to shut down after the shock, but only NFPs adjust their mix of profitable services.
@@
     Not-for-profit firms differ from for-profits mainly in that for-profit firms distribute revenue to stakeholders, while not-for-profits reinvest profit back into the firm.
-Another important distinction is that NFPs are, in general, driven by a mission to better society.
-Additionally, NFPs in the US receive tax benefits for this designation.
+Another important distinction is that NFPs are, in general, driven by a mission to better society, and therefore receive tax benefits.
 
-    NFPs are governed by a board of directors, whose role is to set broad goals and strategies, and provide general oversight.
+    Hospitals are governed by a board of directors, whose role is to set broad goals and strategies, and provide general oversight.
 The board selects executives, the highest level of management of a firm, to carry out day-to-day operations.
-An executive team usually consists of at least a CEO and CFO, but every NFP is different in exactly how they structure their executive team.
+An executive team usually consists of at least a CEO and CFO, but every firm is different in exactly how they structure their executive team.
 Many hospital executives specialize in health care settings by getting degree(s) in health care management, or an MBA specific to health care.
@@
     Two programs were passed as part of the Affordable Care Act that focused on pay-for-performance incentives for hospitals, the Hospital Readmissions Reduction Program (HRRP), and the Hospital Value Based Purchasing Program (HVBP).
-The HRRP focused on penalizing hospitals with poor quality in terms of readmission rates, and the HVBP fpcused on rewarding hospitals with high quality in terms of something.
+HRRP focused on penalizing hospitals with poor quality in terms of readmission rates, and HVBP focused on rewarding hospitals with high quality or show improvement in quality.
 
@@
 
-    The HVBP Program instead rewards hospitals with high quality or shown improvement in quality.
-Specifically, CMS deducts Medicare payments by 2\%, collects this sum, and divides it among the rewarded hospitals.
-Several quality and cost measures in categories safety, efficiency and cost reductions, clinical outcomes, and community engagement, are combined to create a single score metric for each hospital.
-Hospitals are then compared to each other and score points for being above average quality and for showing improvement (\cite{CMS_2023}).
+    The HVBP Program instead rewards hospitals with high quality or improvement in quality.
+Specifically, CMS deducts Medicare payments by 2\% from all eligible hospitals, collects this sum, and divides it among the rewarded hospitals.
+Several quality and cost measures surrounding safety, efficiency, cost reductions, clinical outcomes, and community engagement, are combined to create a single score metric for each hospital.
+Hospitals are then compared to the average and are rewarded for being above average quality and for showing improvement (\cite{CMS_2023}).
 
@@
 
-    Hospital executive teams are understudied partly because of the inaccessibility of granular information.
+    Hospital executive teams are understudied partly because of the inaccessibility of information on them.
 I construct a novel data set of not-for-profit hospitals in the US that contains names, titles, and positions of all board members and executives tied to the hospital from 2009-2015.
@@
 
-    The historical tax forms for all not-for-profits are housed by ProPublica.\footnote{https://projects.propublica.org/not-for-profits/} I use the NonProfit Explorer API to extract Employee Identification Numbers (EIN) of not-for-profits classified as a hospital according to their National Taxonomy of Exempt Entities (NTEE) code.
-After filtering out associations and specialty hospitals, there are approximately 3,000 not-for-profit EINs.
+    The historical tax forms for all not-for-profits are housed by ProPublica.\footnote{https://projects.propublica.org/not-for-profits/} I use the NonProfit Explorer API to extract Employee Identification Numbers (EIN) of NFPs classified as a hospital according to their National Taxonomy of Exempt Entities (NTEE) code.
+After filtering out associations and specialty hospitals, there are approximately 3,000 NFP EINs.
 Importantly, linking EINs to other hospital information is crucial for the analysis.
@@
 
-    Due to the lack of consistency in the organization of executive teams, I focus mainly on the executive teams as a whole instead of specific positions.
+    Due to the lack of consistency in the organization of executive teams, I focus mainly on executive teams as a whole instead of specific positions.
 I drop anyone who is solely a board member, and focus on those who have executive in their title or are labeled as a president or vice president of a department.
-I create various hospital-level characterizations based on their executive teams, such as the existence of a clinically trained executive, the number of clinically trained executives, the number of total executives, and whether the hospital has a Chief Medical Officer.
+I create various hospital-level characterizations based on their executive team, such as the existence of a clinically trained executive, the number of clinically trained executives, the number of total executives, and whether the hospital has a Chief Medical Officer.
+Additionally, I examine changes in executive teams to determine whether the hospital changes its propensity to hire a clinically trained executive.
 Summary statistics for these variables are shown in Table \ref{leader_sumstats}.
+On average, NFP hospitals in my sample have 4 executives and .4 clinically trained executives.
+Sixty-one percent of the sample never changes propensity to hire a clinically trained executive, and 33\% of NFPs in the sample have a Chief Medical Officer.
 
@@
     I merge the created sample of not-for-profits with all for-profits in the AHA survey.
-The outcomes I consider are readmission and mortality rates, which are directly targeted in the pay-for-performance policies, and are from the CMS Hospital Compare data.
+The outcomes I consider are readmission and mortality rates, which are directly targeted in the pay-for-performance policies, and are found in the CMS Hospital Compare data.
 I combine readmission and mortality rates for pneumonia, AMI, and heart failure (the relevant HRRP conditions) into a weighted average based on the number of patients in each condition.
 Additionally, I include several relevant hospital characteristics.
-Number of beds, whether the hospital is an academic medical center, is physician owned or is system affiliated, all come from the AHA survey.
+Number of beds, whether the hospital is an academic medical center, is physician owned, or is system affiliated, all come from the AHA survey.
 Uncompensated care comes from HCRIS, and case mix index, which measures patient complexity, comes from CMS.
@@
     I present a table of means for each hospital type (for-profit, in-sample not-for-profits, and not-for-profits with and without a clinically trained executive) in Table \ref{tab:sumstats_samples}.
-Slightly over half of the hospitals are classified as not-for-profit with leadership information.
-There are key differences in not-for-profits with and without clinically trained executives: those with MDs are larger, more likely to be penalized, have higher uncompensated care, and more complex patients.
+Approximately 850 NFPs have leadership information, with 384 of them hiring an MD executive at some point in the sample.
+I compare these NFPs with the 740 for-profits from the AHA Survey.
+For-profits and NFPs are roughly the same bed size, but NFPs see more patients in each of the three relevant conditions.
+For-profits are more likely to be penalized, but also more likely to receive HVBP incentives (potentially for improvement in quality).
+For-profits have higher readmission rates, and higher mortality rates in all categories except heart failure.
+There are also key differences in not-for-profits with and without clinically trained executives.
+Those with MDs are larger, more likely to be penalized, have higher uncompensated care, and more complex patients.
 I discuss the implications of these differences for identification in Section \ref{sec:model}.
@@
 
-    I plot patient-complexity adjusted readmission and mortality rates across time for different hospital types in Figure \ref{fig:weighted_read_mort_graph}.
+    I plot weighted average readmission and mortality rates across time for different hospital types in Figure \ref{fig:weighted_read_mort_graph}, where I adjust both rates by baseline patient complexity.
+Pay-for-performance incentive changes are shown as a dotted line between 2011 and 2012.
 For-profit hospitals have the highest readmission rates prior to 2012.
-While the trend in readmission rates behaves similarly across hospital types prior to the programs taking affect, there are differential trends after the programs.
-NFPs without cilnically trained executives decrease readmissions more similarly to FPs than other NFPs.
+While the readmission rate trend behaves similarly across hospital types prior to the programs, there are notable trend differences after the programs.
+NFPs with clinically trained executives show a more mitigated decrease in readmissions compared to the other types.
 In mortality rates, NFPs with a clinically trained executive have the lowest rates before the programs, but differentially continue with an increasing trend in mortality after the programs.
-While this is purely descriptive evidence, the trends in outcomes point to more similar behavior between for-profits and NFPs without clinically trained executives in comparison to NFPs with a clinically trained executive.
+While this is purely descriptive evidence, the trends in outcomes point to more similar behavior between for-profits and NFPs without a clinically trained executive in comparison to NFPs with a clinically trained executive.
 
@@
 Pay-for-performance incentives essentially incorporate performance directly into a hospital's profit function.
-Here, I think of performance as quality of care, which is inversely related to readmission and mortality rates.
+Here, I model performance as quality of care, which is inversely related to readmission and mortality rates.
 While the literature speculates about the true form of NFP objective functions, we can generally think of hospitals choosing some linear combination of profit and societal benefit.
-Intuitively, a hospital that places more weight on profit (for-profit type) will only care about quality when it enters into the profit function, and therefore will respond more drastically to quality entering the profit function than a hospital who already cared about quality before the incentives.
+Intuitively, a hospital that places only weight on profit (for-profit type) will only care about quality when it enters into the profit function, and therefore will respond more drastically to pay-for-performance than a hospital who already cared about quality through societal benefit before the incentives.
 Therefore, hospital response to pay-for-performance incentives reveals something about their underlying preference on profit vs.
@@
 
-    Formally, I model the hospital objective function as a weighted average of profit and societal (or patient) outcomes.
-For simplicity, I abstract away from quality having a direct effect of quantity and price, which does not drastically change the main results.
+    To formalize this intuition, I specify the hospital objective function as a weighted average of profit and societal (or patient) outcomes and derive characteristics of the chosen quality.
+For simplicity, I abstract away from quality having a direct effect of quantity and price.
 Without any financial incentives to increase quality, hospital quality is decided by:
@@
 
-    \noindent where $R$ is net revenue, $c_{\theta}(.)$ is an increasing cost function specific to quality $\theta$, $u(.)$ is extra utility gained from quality, which is increasing and concave in $\theta$, and $\alpha\in[0,1]$ captures how much the hospital cares about profit vs.
+    \noindent where $R$ is net revenue, $c_{\theta}(.)$ is an increasing cost function specific to quality $\theta$, $u(.)$ is extra utility gained from quality, which is increasing and concave in $\theta$, and $\alpha\in[0,1]$ captures the weight placed on profit vs.
 societal benefit.
@@
     Therefore, in a world with financial incentives on quality, quality is increasing with more weight placed on profit.
-The purpose of this paper is to investigate the response of hospitals with different $\alpha$ when a policy changes hospital incentives using pay-for-performance.
-Hence, I combine the results found in each scenario into one response function that depends on $\alpha$, where $\theta_2$ is quality under pay-for-performance incentives and $\theta_1$ is quality with no financial incentive to quality.
-Thus,
+I combine the results found in each scenario into one response function that depends on $\alpha$, where $\theta_2$ is quality under pay-for-performance incentives and $\theta_1$ is quality with no financial incentive to quality.
+That is,
 
@@
 
-    The theoretical prediction is that change in quality depends on how much weight the hospital places on profit vs.
+    Hence, this model predicts that change in quality depends on how much weight the hospital places on profit vs.
 societal benefit.
 Particularly, hospitals with more weight on profit respond more than hospitals with more weight placed on societal benefit.
-In reality, $\alpha$ is not revealed by the hospital, except in the case of for-profits, who have revealed that they have a high $\alpha$ by their ownership status.
+In reality, $\alpha$ is unknown, except in the case of for-profits, who have revealed that they have a high $\alpha$ by their ownership status.
 However, not-for-profit behavior on average does not bring much clarity about their objectives.
 An observable and economically relevant characteristic of a hospital that could affect $\alpha$ is composition of their leadership team.
-Thus, I compare the response to incentive change between for-profits and not-for-profits with different leadership teams to bring clarity on how this might point to underlying $\alpha$.
+Thus, I compare the response of for-profits and not-for-profits with different leadership teams to bring clarity on how this might point to underlying $\alpha$.
 
@@
 
-    In light of the theoretical predictions, I first test empirically whether different sub-samples of not-for-profits respond to program incentives differently than for-profits.
-I estimate the following regression equation:
+    In light of the theoretical predictions, I first test empirically whether NFPs respond to program incentives differently than for-profits. I estimate the following regression equation:
 
@@
 However, hospital executive teams are dynamic, and whether or not a hospital employs a clinically trained executive in each period may change.
-Therefore, for simplicity, I first present effects of having a clinically trained executive among stable executive teams, i.e., NFPs with no change in their propensity to hire a clinically trained executive.
+Therefore, I first present effects of having a clinically trained executive among stable executive teams, i.e., NFPs with no change in their propensity to hire a clinically trained executive.
 
@@
 First, I assume that, absent the program enactments, different hospital types would have had parallel trends in outcomes.
-While I see no evidence of pre-trends, Table \ref{tab:sumstats_samples} does show that hospitals with a clinically trained executive are different than other hospitals in multiple characteristics.
-Therefore, there may be concern that there are systematic differences between hospital types that could be driving the results.
+While I see no evidence of pre-trends, Table \ref{tab:sumstats_samples} suggests that hospitals with a clinically trained executive are different than other hospitals along multiple characteristics.
+Therefore, there may be concern that these differences between hospital types could be driving the results.
 Therefore, my main specification uses the synthetic difference-in-differences method established in \citeauthor{arkhangelsky2021synthetic} (\citeyear{arkhangelsky2021synthetic}).
-This method creates unit weights that more heavily count comparison hospitals that have similar trends in outcomes prior to the programs.
-It also weighs time periods that balance pre-program with post-program for the comparison hospitals.
-Then, a simple two way fixed effects estimator is used with these weights.
-This is a more robust average treatment effect that relies less on a strong parallel trends assumption (\cite{arkhangelsky2021synthetic}).
+This method creates unit weights that more heavily count comparison hospitals that are similar to treatment hospitals prior to the programs.
+It also weighs time periods that balance pre-program and post-program for the comparison hospitals.
+Then, a simple two way fixed effects estimator is used with these weights, yielding a more robust average treatment effect that relies less on a strong parallel trends assumption (\cite{arkhangelsky2021synthetic}).
 I present the synthetic diff-in-diff estimates as the main results, and they are largely similar to the standard two way fixed effects results in Appendix \ref{app:fullsample}.
-Further, while ultimately I employ a synthetic difference-in-differences method, the conceptual identification is identical to a standard difference-in-differences.
+Further, while ultimately I employ a synthetic difference-in-differences method, the conceptual identification is the same as a standard difference-in-differences.
 
      Second, I assume that executive teams are not determined endogenously with the programs.
-In the main specification, I limit to executive teams which do not change their propensity to hire an MD over time.
-Thus, I rely on this assumption so that I am not conditioning on an outcome.
-In supplementary analyses, I use executive team changes to disentangle signalling vs.
+In the main specification, I limit the sample to hospitals with executive teams that do not change their propensity to hire a clinical executive over time.
+Thus, I rely on this sample selection not being correlated with the program enactment.
+In supplementary analyses, I leverage executive team changes to disentangle signalling vs.
 managing effects, which is only identified if leadership changes are uncorrelated with the policies.
@@
      I assess the validity of this assumption in two ways.
-First, I analyze whether changes in clinically trained executives are concentrated around the enactment of both HRRP and HVBP, 2011-2012.
-Second, I analyze whether hospitals who end up getting penalized through HRRP are more likely to change clinically trained executives on their team than non-penalized hospitals, and similarly for hospitals who end up getting extra payments for high quality of care through HVBP.
+First, I analyze whether changes in clinically trained executives occur more often in time periods of program enactment.
+Second, I analyze whether hospitals who end up getting penalized or receiving payments are more likely to make changes in clinical executive members than those who are not.
 That is, are hospitals that expect to be affected by the programs more likely to hire or fire clinically trained executives?
@@
 The estimates from this analysis are presented in Table \ref{tab:change_analysis}.
-Changes in any clinically trained executive are less likely to occur in 2012-2014 relative to 2010, indicating that executives are actually becoming more stable over time along the dimension of hiring clinically trained executives.
+Changes in any clinically trained executive are less likely to occur in 2012-2014 relative to 2010, indicating that executive teams are actually becoming more stable over time along the dimension of hiring clinically trained executives.
 Further, hospitals who end up being penalized through HRRP or receiving payments through HVBP are not more likely to change their propensity to hire any clinically trained executive.
@@
 
-     I consider the differential program response in readmission rates of for-profit firms compared to the different sub-samples of not-for-profit firms (specification \ref{eq:forprofit}).
-First, I compare for-profits to all not-for-profits in the sample in Figure \ref{fig:read_synth_plota}, which shows that on average, there is no difference in how for-profits and not-for-profits change their readmission rates in response to the incentive changes.
-Figures \ref{fig:read_synth_plota} and \ref{fig:read_synth_plotb} compare for-profit readmission rates to NFPs with a clinically trained executive and without a clinically trained executive, respectively.
-Figure \ref{fig:read_synth_plotb} shows that for-profit hospitals decrease readmission rates more than not-for-profits with an MD executive after pay-for-performance incentives.
-However, while for-profits have higher readmission rates than NFPs without a clinically trained executive, there is no difference in their change in readmission rates after the incentives are enacted, shown in Figure \ref{fig:read_synth_plotb}.
+     The estimated difference in readmission rate trends for all hospital types in response to the programs, along with ATE and standard errors, are shown in Figure \ref{fig:read_synth_plot}.
+I first discuss the differential program response of for-profit firms compared to the different NFP sub-samples (Equation \ref{eq:forprofit}), shown in panels \ref{fig:read_synth_plota}-\ref{fig:read_synth_plotc}.
+
+     Panel \ref{fig:read_synth_plota} shows that on average, there is no difference in how for-profits and not-for-profits change their readmission rates in response to the incentive changes.
+However, when NFPs are subset into whether they have a clinical executive or not, panel \ref{fig:read_synth_plotb} shows that there is a differential response to the program between for-profits and NFPs with a clinical executive.
+In particular, for-profits decrease readmission rates by .3 more than NFPs with a clinical executive, an additional 3 patient readmission decrease relative to the average number of patients.
+While numerically small, this difference in response is non-existent between for-profits and NFPs without a clinical executive (panel \ref{fig:read_synth_plotc}), indicating some, if minor, difference in objectives resulting from having a clinical executive.
 
@@
 
-    Finally, in Figure \ref{fig:read_synth_plotd}, I present estimates of Equation \ref{eq:clinical}, directly comparing NFPs with and without clinical executives.
-NFPs without a clinically trained executive lowered readmission rates by .3ppts more than NFPs with a clinically trained executive.
-Relative to the average readmission rates in 2010, this is approximately a 1.5\% decrease, or 11 less readmissions per year.
+    Next, in panel \ref{fig:read_synth_plotd}, I present estimates of Equation \ref{eq:clinical}, directly comparing NFPs with and without a clinical executive.
+NFPs without a clinically trained executive lowered readmission rates by .24 ppts more than NFPs with a clinically trained executive.
 
-    Next, I present estimated differences in mortality rates between different hospital types in Figure \ref{fig:mort_synth_plot}.
-Figure \ref{fig:mort_synth_plota} compares for-profit and all NFP mortality rates after the pay-for-performance incentives.
-These results show that there is no statistical difference in mortality rates between for-profits and not-for-profits after the programs.
-However, when considering NFPs with different leadership team compositions, shown in in Figures \ref{fig:mort_synth_plotb} and \ref{fig:mort_synth_plotc}, there seems to be a difference in behavior.
-While mortality rates are significantly more noisy than readmission rates, the coefficient on the difference in mortality rates for for-profit compared to NFP with an MD is -.17, a magnitude similar to that of readmission rates.
-However, for-profits and NFPs without a clinical executive behave similarly.
-
-    I also directly estimate the difference in mortality rates among NFPs with and without clinical experience in Figure \ref{fig:mort_synth_plotd}.
-Again, the difference is not statistically significant, but this is likely due to the noisiness in mortality since the results is similar to that of readmission rates.
-NFP hospital without a clinically trained executive decrease mortality rates more than NFPs without a clinically trained executive due to new pay-for-performance incentives.
+    I now shift to discussing differential responses of mortality rates.
+I present estimated differences in mortality rates between different hospital types in Figure \ref{fig:mort_synth_plot}.
+Panel \ref{fig:mort_synth_plota} compares for-profit and all NFP mortality rates after the pay-for-performance incentives.
+These results show that there is no difference, both statistically and in magnitude, between mortality rates of for-profits and not-for-profits after the programs.
+This result holds when the comparison group is limited to NFPs without a clinical executive (panel \ref{fig:mort_synth_plotc}).
+However, when comparing for-profits to NFPs with a clinical executive, the magnitude of the difference in mortality trends jumps to -.11.
+While still statistically insignificant, I argue that this still indicates a difference in behavior between the hospital types, particularly due to the documented noisiness of mortality measures (add citations).
 
@@
 
+    I also directly estimate the difference in mortality rates among NFPs with and without clinical experience in panel \ref{fig:mort_synth_plotd}.
+Again, the difference is not statistically significant, but the coefficient compared to the true nulls in panels \ref{fig:mort_synth_plota} and \ref{fig:mort_synth_plotc} indicates NFPs without a clinical executive potentially decreasing mortality more aggressively than NFPs with a clinical executive.
+
     \subsection{Decomposition}
 
-    While it is telling that executive team composition affects behavior after incentives change, more clarity is needed about the root cause of the differences.
-It is not clear whether these results are due to clinically trained executives being a signal of underlying hospital objectives, or the leaders themselves making a difference in outcomes.
-Therefore, I present the results of specification \ref{eq:decomp} in Table \ref{tab:MD_noMD_readmort_decomp_synth}, where the variation in executives is either existence of a hospital executive at any point in time, or whether the hospital has a clinically trained executive in 2012 when the programs were enacted.
-The left panel on readmission rates shows that the difference in response is fully driven by having a clinically trained executive in 2012 compared to having a clinically trained executive at another point in time.
-The right panel shows similar magnitudes of coefficients with more noise, as in the main specification.
+    In the above analysis, I show that NFPs with a clinical executive seem to respond to incentives differently than both for-profits and NFPs without a clinical executive, suggestive evidence of differential objectives correlated with leadership teams.
+However, the analysis to this point does not speak to the mechanism by which clinical executives lead to differential behavior.
+In this section, I present two theories by which a clinical executive leads to differential behavior, and present descriptive evidence of these empirically.
+
+    First, having a clinical executive could be a signal of underlying hospital objectives.
+That is, a hospital who hires a clinical executive is reveals their underlying objectives.
+Alternatively, hiring a clinical executive could actually change the objectives of the hospital for the time they are actively managing the hospital.
+To disentangle the two, I present the results of specification \ref{eq:decomp} in Table \ref{tab:MD_noMD_readmort_decomp_synth}, where the variation in executives is either existence of a hospital executive at any point in time, or whether the hospital has a clinically trained executive in 2012 when the programs were enacted.
+Columns (1) and (2) show that the difference in readmission response is driven by having a clinically trained executive in 2012, compared to having a clinically trained executive at another point in time.
+Columns (3) and (4) support the same theory given the magnitudes of the coefficients, but are still statistically insignificant due to the noisiness of mortality.
+The results suggest that managers make a difference in behavior and objectives, rather than being a signal of already existing objectives.
 
@@
 
-    In this paper, I study differential objective between different firm types.
-While the differences in for-profit and not-for-profit hospitals have been studied in the past, it is unclear whether certain not-for-profit characteristics make them act more like for-profits.
-In order to better understand the role of leadership in objectives, I estimate the differential response to a change in incentives among for-profits, not-for-profits, and not-for-profits with and without a clinically trained executive executives.
-I find that not-for-profits without clinically trained executives respond to incentives much more similarly to for-profits than to not-for-profits without clinically trained executives.
+    Existing literature suggests that not-for-profit hospitals may or may not be meaningfully different from for-profits.
+Along some dimensions, they seem to have meaningful dimensions, but we still do not understand underlying objectives of not-for-profit hospitals, which ultimately affect cost and quality of health care.
+One dimension of NFP hospitals that has not been studied is how manager composition affects objectives and behaviors of the firm.
+
+    In this paper, I study differential response to a change in incentives between for-profit and NFP hospitals with and without a clinical executive.
+I do this by collecting a novel data set on hospital executives from Tax Form 990s and merging this to existing hospital data.
+I then estimate differential responses in readmission and mortality rates after several pay-for-performance initiatives are enacted in the US in 2012 using a synthetic difference in differences methodology.
+
+     I find that not-for-profits with a clinical executive respond differently than both for-profits and NFPs without a clinical executive.
+Specifically, all firm types decrease readmissions, but NFPs with a clinical executive are less aggressive in their responsiveness than the other firm types.
+A similar result holds for mortality rates.
+This finding is consistent with the theoretical prediction that firms who place more weight on societal benefit than on profit will have a more mitigated response to policies affecting the profit function.
+Additionally, under the assumption that executive team changes are not correlated with the program, I find that the differential response is driven by the actual management of the clinical executive, not by a clinical executive being a signal of underlying objectives.
+These findings suggest that policymakers should take into account more than just ownership status when evaluating how hospitals will respond to incentives.
 
@@
 
-There are over 1.5 million not-for-profit entities in the US (cite), making it crucially important to be able to filter by type of entity before analyzing any PDFs.
+There are over 1.5 million not-for-profit entities in the US, making it crucially important to be able to filter by type of entity before analyzing any PDFs.
 The API allows this by filtering a query based on National Taxonomy of Exempt Entities (NTEE) code.
@@
 I save the name, secondary name, state, and zip code, all of which do not vary by year.
-I also save each year's URL that links to the Tax Form 990 PDF.
-For the sake of a comprehensive data set, I keep years 2006-2020 (I later limit to 2009-2016 when focusing on the Hospital Readmissions Reduction Program).
+I also save each year's URL link to the Tax Form 990 PDF.
+For the sake of a comprehensive data set, I keep years 2006-2020 (I later limit to 2010-2014 when focusing on pay-for-performance initiatives in 2012).
 Thus, I finish this step with a panel data set of EIN characteristics and PDF locations.
@@
 
-It is crucial to the analysis to be able to link these not-for-profits with other sources of data to recover penalties from HRRP, bed size, and outcomes of interest (definitely need to talk more about outcomes).
-Therefore, I take a conservative approach to matching EINs to American Hospital Association (AHA) ID, which links fairly easily to Medicare ID numbers, based on name and location.
-First, I will discuss limitations and cleaning that I do to the AHA data and tax form data separately before doing any matching.
+It is crucial to the analysis to be able to link these not-for-profits with other sources of data to recover penalties from HRRP, payments from HVBP, bed size, and outcomes of interest.
+The ideal hospital data set to match to is American Hospital Association (AHA) Survey, which contains hospital characteristics and Medicare ID number.
+However, an EIN to AHA ID crosswalk does not exist.
+Therefore, I take a conservative approach to matching EINs to American Hospital Association (AHA) ID based on hospital name and location.
+First, I will discuss limitations and cleaning of the AHA data and tax form data.
 
-I download the AHA data from Wharton Research Data Services.
-I filter only to hospitals in the contiguous US, Alaska, and Hawaii (excluding places like Puerto Rico), classified as not-for-profit or state/community, and those that are general acute care.
+In the AHA data, I filter only to hospitals in the contiguous US, Alaska, and Hawaii (excluding places like Puerto Rico), classified as not-for-profit or state/community, and those that are general acute care.
 I also filter out any hospitals who weren't present in the data (or change system ID) in 2009-2015, meaning they either closed or were acquired.
 Due to the survey nature of this data, a hospital name may look slightly different from one year to the next.
-For example, "Waldo County General Hospital" is also Waldo County General Hospital Maine Health".
+For example, ``Waldo County General Hospital" is also ``Waldo County General Hospital Maine Health".
 Further, zip codes may change by one or two digits, making them unreliable to match based on.
@@
 I consider which not-for-profit entities are not likely to be hospitals and drop them.
-There are numerous foundations or auxiliary not-for-profits with the purpose of raising funds for the hospital, but do not actually care for patient.
+There are numerous foundations or auxiliary firms with the purpose of raising funds for the hospital, but do not provide services to patients.
 I filter out any not-for-profit with "foundation" or "auxiliary" in the name.
@@
 I focus on exact string matches, so I remove all spaces and common characters that could cause mismatches such as \&, ', -, and inc.
-Next, I take each AHA name and look for exact matches in a not-for-profit's first or secondary name only for not-for-profits in the same state as the AHA hospital.
+Next, I take each AHA name and look for exact matches in a not-for-profit's first or secondary name for not-for-profits in the same state as the AHA hospital.
 When an exact match is found, I record the link between AHA ID and EIN.
@@
 This adds an additional 90 hospital matches, accounting for a total of 34.5\% of AHA hospitals.
+Finally, I manually search through unmatched EINs to identify any matches.
+From google searches, I identify an additional 300 EIN-AHA ID matches.
 
-In some cases, tax forms are associated with a system of hospitals instead of one individual hospital.
-Thus, I create another variable that captures the names of systems which match.
-The process of matching is the same, the only difference is that now I am considering system names from the AHA data instead of hospital names.
-A total of 1,136 AHA hospitals have an EIN match on hospital name, system name, or both.
-This is roughly 41\% of the sample.
-These hospitals become the main sample for my analysis.
-While not capturing a large majority of AHA hospitals that are categorized as not-for-profit, I would rather be conservative in selecting hospitals which I am confident that the characteristics and measurement of the independent and dependent variables are accurate.
-
-(add a paragraph about validity of matches)
-
- Now that I have a sample of hospitals, the next step is to extract the names of the people in charge of these hospitals from the Tax Form 990 PDFs.
+I then extract the names of board members and executives from the Tax Form 990 PDFs of matched hospitals.
 In the data set of hospital PDF URLs that I collected earlier, I limit to the hospitals with solid matches described above.
-I then loop through each EIN, downloading each PDF locally and using tesseract package in R to extract text from the relevant pages of the PDF using OCR text extraction methods.
+I then loop through each EIN, downloading PDFs locally and using the tesseract package in R to extract text from the relevant pages of the PDF using OCR text extraction methods.
 In particular, I loop through each page of the PDF, look for the title associated with leadership names: ``Officers, Directors, Trustees, Key Employees, and Highest Compensated Employees", and save all the text from any pages where this title is found.
@@
 
-One tricky aspect of the not-for-profit Explorer API is that, only in some cases, if two forms are present for an EIN, year, only the first one (which is typically not the one with the relevant information) is pulled.
+One tricky aspect of the NonProfit Explorer API is that, only in some cases, if two forms are present for an EIN, year, only the first one (which is typically not the one with the relevant information) is pulled.
 Therefore, for some hospitals, a couple years will have gaps in text extraction data.
@@
 The form of the text data is a data frame with one column, where each line of text is saved in a different row.
-I write a text cleaning package that locates names, positions, titles, and indications of resigning.
-I will now describe this function in three parts: preliminary cleaning of strings, locating names, and locating positions, titles, and key words surrounding resigning.
-
 Typically on the same page as the names and positions is a list of the highest compensated employees and their compensation.
@@
 I then split up the phrase into individual words, so one phrase with 5 words is broken up into 5 variables.
+I write a text cleaning function that locates names, positions, titles, and indications of resigning.
+I flag name rows using the Census name list data for the year 2000.
+The columns with the most flags are then identified as name columns.
+I then extract any text that indicates a doctor title and link it to the name located the closest to it.
+Similarly, I extract text of all potential positions such as CEO, CFO, CMO, president, board member, etc., and link it to the name most closely located to it.
 
-Next, I locate first and last names in the data.
+I then create a name-level data set that only includes executives.
+That is, I remove all board members from the data.
+I then create hospital level indicators based on the names, titles and positions associated with the hospital: the number of MD executives, the number of total executives, and whether the hospital employs a CMO.
+In Figure \ref{fig:state_doc}, I show the percent of hospitals with a clinical executive in each state.
+This figure verifies that clinical executives are not concentrated in a small region geographically.
 
+\begin{figure}[ht!]
+    \centering
+    \caption{Percent of Hospitals with Clinical Executives by State}
     \includegraphics[width=\textwidth]{Objects/has_doc_avg_map.pdf}
+    \label{fig:state_doc}
+\end{figure}
 
@@
 
-\import{Tables}{MD_noMD_uncompCMI_fullsample.tex}
-
-\import{Tables}{forprofit_uncompCMI_fullsample.tex}
-
 \section{Synthetic Diff-in-Diff Result Tables}

### 2024-05-30 Tables/MD_noMD_readmort_condition_synth.tex
add tables for specific conditions and dsecriptive stats about leadership team changes

\begin{table}[ht!]

\caption{\label{tab:MD_noMD_readmort_condition_synth}Non-MD vs. MD Executive Readmission and Mortality Synthetic DiD Results, All Conditions}
\centering
\begin{tabular}[t]{lcccccc}
\toprule
\multicolumn{1}{c}{\underline{ }} & \multicolumn{3}{c}{\underline{Readmission Rates}} & \multicolumn{3}{c}{\underline{Mortality Rates}} \\
 & AMI & HF & Pneum. & AMI & HF & Pneum.\\
\midrule
NFP w/out MD Exec x Post Program & -0.19 & -0.18 & -0.4$^{*}$ & 0.09 & -0.1 & -0.3\\
 & (0.19) & (0.2) & (0.17) & (0.2) & (0.15) & (0.19)\\
\textbf{Comparison Group:} &  &  &  &  &  & \\
NFP w/ MD Exec & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
 &  &  &  &  &  & \\
\addlinespace
Observations & 1,135 & 1,625 & 1,625 & 1,295 & 1,625 & 1,625\\
\bottomrule
\multicolumn{7}{l}{\textsuperscript{} Results from Equation with condition specific readmission and mortality rates as outcome variables.}\\
\multicolumn{7}{l}{\textsuperscript{} Standard errors are clustered at the hospital level.}\\
\multicolumn{7}{l}{\textsuperscript{} Signif. codes: 0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1}\\
\end{tabular}
\end{table}

### 2024-05-30 Tables/md_readmort_condition_fullsample.tex
add tables for specific conditions and dsecriptive stats about leadership team changes

\begin{table}[ht!]
   \caption{\label{tab:forprofit_mort_conditions_fullsample} Comparison to For-Profit Mortality Regression Results, All Conditions}
   \bigskip
   \centering
   \begin{tabular}{lcccccc}
      \toprule
                                 & Heart Attack & Heart Failure & Pneumonia     & Heart Attack & Heart Failure & Pneumonia\\  
       & \multicolumn{3}{c}{Readmission Rates} & \multicolumn{3}{c}{Mortality Rates} \\ \cmidrule(lr){2-4} \cmidrule(lr){5-7}
                                 & (1)          & (2)           & (3)           & (4)          & (5)           & (6)\\  
      \midrule 
      For-Profit x Post Programs & -0.20        & -0.20         & -0.48$^{***}$ & -0.08        & -0.03         & -0.33\\   
                                 & (0.20)       & (0.20)        & (0.15)        & (0.21)       & (0.18)        & (0.23)\\   
      \textbf{Comparison Group:} &  &  &  &  &  &  &  &  & \\
      In-Sample NFPs & $\checkmark$ &  &  & $\checkmark$ &  &  & $\checkmark$ &  & \\
      NFP w/ MD &  & $\checkmark$ &  &  & $\checkmark$ &  &  & $\checkmark$ & \\
      NFPs w/out MD &  &  & $\checkmark$ &  &  & $\checkmark$ &  &  & $\checkmark$\\
       \\
      Observations               & 1,293        & 2,342         & 2,467         & 1,514        & 2,288         & 2,459\\  
      \bottomrule
   \end{tabular}

   \par \raggedright 
   Results from Equation \ref{eq:forprofit} for mortality rates by condition.\\
   Standard errors are clustered at the hospital level.\\
   Signif. codes: 0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1
\end{table}

### 2024-05-30 Paper/paper.tex
Updates from Overleaf

@@
 In this paper, I investigate whether leadership team composition of NFP hospitals affects how similarly they behave to for-profits.
-I use a novel data set on NFP hospital executives in the US (2009-2015) to investigate whether the presence of clinically trained executives affects hospital behavior.
+I use a novel data set on NFP hospital executives in the US (2010-2014) to investigate whether the presence of clinically trained executives affects hospital behavior.
 In particular, I leverage pay-for-performance initiatives in the US enacted in 2012, and measure the  readmission and mortality rate responses between FP, NFP, and NFP with and without clinically trained executives.
 I find that NFPs without clinically trained executives behave more like FPs than other NFPs.
-This is consistent with theoretical predictions under the assumption that NFP hospitals with clinically trained executives value societal outcomes more than those without clinically trained executives.
+This is consistent with theoretical predictions if NFP hospitals with clinically trained executives value societal outcomes more than those without clinically trained executives.
 		} 
@@
 
-    \subsection{Not-for-Profit Hospital Leadership}
+    \subsection{Not-for-Profit Hospital Executives}
 
-    Not-for-profit firms differ from for-profits mainly in that for-profit firms distribute revenue to stakeholders, while not-for-profits reinvest profit back into the firm.
-Another important distinction is that NFPs are, in general, driven by a mission to better society, and therefore receive tax benefits.
+    Not-for-profit firms in the US differ from for-profits mainly in that for-profit firms distribute revenue to stakeholders, while not-for-profits reinvest profit back into the firm.
+An important distinction between the two is that NFPs are, in general, driven by goals to better society, and therefore receive tax benefits from the government.
 
@@
 The board selects executives, the highest level of management of a firm, to carry out day-to-day operations.
-An executive team usually consists of at least a CEO and CFO, but every firm is different in exactly how they structure their executive team.
-Many hospital executives specialize in health care settings by getting degree(s) in health care management, or an MBA specific to health care.
-There are specific positions that are often filled by someone with clinical training, such as a Chief Medical Officer, and doctors can even find themselves in other top managerial positions as well, such as CEO or president.
+An executive team usually consists of at least a Chief Executive Officer (CEO) and Chief Financial Officer (CFO), but there is variation in how firms organize executive teams.
+It is not uncommon for hospital executives to specialize in health care settings by earning a degree in health care management, or an MBA specific to health care.
+There are executive positions that are often filled by someone with clinical training, such as a Chief Medical Officer (CMO), and doctors can even find themselves in other top managerial positions as well, such as CEO or president.
 While some doctors earn additional degrees before stepping into an executive role, this is not a necessary condition to becoming a physician executive.
 
-    Anecdotal evidence is mixed on the benefits of hiring physician executives.
-On one hand, physicians bring a unique combination of clinical expertise and being comfortable with administration, and they can bring insights that improve hospital outcomes (\cite{Stajduhar_2023}, \cite{Ahmed_2022}).
-Alternatively, physicians may not always be adequately trained to enter into high level management positions, and therefore are not equipped to lead a firm (\cite{HarvardBusinessReview2018}).
+    Hiring physician executives can be controversial.
+On one hand, physicians bring a unique combination of clinical and administration expertise, and they have insights that can potentially improve hospital outcomes (\cite{Stajduhar_2023}, \cite{Ahmed_2022}).
+Alternatively, physicians may not always be adequately trained to enter into high level management positions, and therefore worsen hospital outcomes (\cite{HarvardBusinessReview2018}).
 
@@
 Sections of this form include a statement of revenue, statement of functional expenses, a balance sheet, and, as used in this project, a list of all key employees, executives, and board members.
-Each firm is required to report the name and title, average hours per week, position, and compensation of their board members and executive leaders.
+Each firm is required to report the name and title, average hours per week, position, and compensation of their board members and executives.
 
@@
 
-    Two programs were passed as part of the Affordable Care Act that focused on pay-for-performance incentives for hospitals, the Hospital Readmissions Reduction Program (HRRP), and the Hospital Value Based Purchasing Program (HVBP).
-HRRP focused on penalizing hospitals with poor quality in terms of readmission rates, and HVBP focused on rewarding hospitals with high quality or show improvement in quality.
+    Two programs were passed as part of the Affordable Care Act that focused on pay-for-performance incentives for hospitals: the Hospital Readmissions Reduction Program (HRRP), and the Hospital Value Based Purchasing Program (HVBP).
+HRRP focused on penalizing hospitals with poor quality, and HVBP focused on rewarding hospitals with high quality or show improvement in quality.
 
-    In October 2011, the Center for Medicare and Medicaid Services (CMS) released a set of rules under HRRP that mandated penalties for hospitals with above average readmission rates.
+    In October 2011, the Center for Medicare and Medicaid Services (CMS) released a set of rules under HRRP mandating penalties for hospitals with above average readmission rates.
 A readmission is when a patient returns to the hospital within 30 days of being discharged from a previous stay; avoidable readmissions are a bad outcome for patients and increase health care spending.
@@
 
-    Penalized hospitals pay penalties in the form of a fixed rate for every Medicare patient regardless of the condition.
-The rate began at 1\% and increased to 3\% in 2015.
+    Penalties are given in the form of a fixed rate reduction of 1-3\% for every Medicare patient regardless of the condition.
 Further, CMS does not distinguish a necessary readmission from an avoidable readmission; any repeat hospital visit is included in the penalty calculation.
@@
 
-    The HVBP Program instead rewards hospitals with high quality or improvement in quality.
+    The HVBP Program instead rewards hospitals with high quality or significant improvement in quality.
 Specifically, CMS deducts Medicare payments by 2\% from all eligible hospitals, collects this sum, and divides it among the rewarded hospitals.
-Several quality and cost measures surrounding safety, efficiency, cost reductions, clinical outcomes, and community engagement, are combined to create a single score metric for each hospital.
-Hospitals are then compared to the average and are rewarded for being above average quality and for showing improvement (\cite{CMS_2023}).
+Several quality and cost measures surrounding safety, efficiency, cost reductions, clinical outcomes, and community engagement are combined to create a single score metric for each hospital.
+Hospitals are then compared to the average and are rewarded for being above average quality or for showing improvement (\cite{CMS_2023}).
 
@@
 
-    Hospital executive teams are understudied partly because of the inaccessibility of information on them.
+    Hospital executive teams are understudied partly due to the inaccessibility of granular data.
 I construct a novel data set of not-for-profit hospitals in the US that contains names, titles, and positions of all board members and executives tied to the hospital from 2009-2015.
 I gather this data from publicly available Tax Form 990s, which every sufficiently large not-for-profit files each year in the US.
-To my knowledge, this is the first large scale gathering of executive names from these forms.\footnote{\citeauthor{brickley2010board} (\citeyear{brickley2010board}) collects compensation information for a select number of hospitals.}
+To my knowledge, this is the first large scale gathering of executive names from these forms.\footnote{\citeauthor{brickley2010board} (\citeyear{brickley2010board}) collects compensation data for a select number of hospitals.}
 
-    The historical tax forms for all not-for-profits are housed by ProPublica.\footnote{https://projects.propublica.org/not-for-profits/} I use the NonProfit Explorer API to extract Employee Identification Numbers (EIN) of NFPs classified as a hospital according to their National Taxonomy of Exempt Entities (NTEE) code.
+    The historical tax forms for all not-for-profits are housed by ProPublica.\footnote{https://projects.propublica.org/not-for-profits/} I use the NonProfit Explorer API to extract Employee Identification Numbers (EIN) of NFPs with a hospital National Taxonomy of Exempt Entities (NTEE) code.
 After filtering out associations and specialty hospitals, there are approximately 3,000 NFP EINs.
-Importantly, linking EINs to other hospital information is crucial for the analysis.
-I rely on matching by location and name of hospital in the tax forms and AHA data.
+Linking these EINs to other hospital information is crucial for the analysis.
+I therefore rely on matching by location and name of hospital in both the tax forms and AHA data.
 I confidently match 1200 EINs to an AHA identification number based on exact name matches within the same state.
-I compare the matched hospitals to non-matched hospitals in Appendix \ref{app:matched}.
+I assess the observable differences between matched and non-matched hospitals in Appendix \ref{app:matched}, and conclude that the samples are overall similar apart from an under-sampling of hospitals that belong to systems.
 
-    For each of the 1200 EINs, I then extract web URLs to Form 990 PDFs in each year.
-I download these locally and use optical character recognition (OCR) text extraction methods to scrape the relevant section of the PDF, which is titled ``Officers, Directors, Trustees, Key Employees, and Highest Compensated Employees".
-Finally, I use string cleaning methods to identify names and positions for each EIN, year.
-I also identify titles indicating clinical training: MD, Dr., or DO.
-Using OCR extraction can be unreliable in some instances if the text is handwritten or small.
-Therefore, I manually record names, positions, and titles for observations that are missing after the initial extraction.
+    For each of the matched EINs, I extract web URLs to Form 990 PDFs for each year.
+I use an algorithm to download these locally and use optical character recognition (OCR) text extraction methods to scrape the relevant section of the PDF, which is titled ``Officers, Directors, Trustees, Key Employees, and Highest Compensated Employees".
+Finally, I use string cleaning methods to identify all names and positions under each EIN, year.
+With each name, I extract titles indicating clinical training: MD, Dr., or DO.
+Using OCR extraction can be unreliable in cases where the text on the form is handwritten or small.
+Therefore, for observations that are missing after the initial extraction, I manually record names, positions, and titles.
 This yields 852 not-for-profit hospitals that are matched to AHA data, and are not missing text information.
 
-    Due to the lack of consistency in the organization of executive teams, I focus mainly on executive teams as a whole instead of specific positions.
-I drop anyone who is solely a board member, and focus on those who have executive in their title or are labeled as a president or vice president of a department.
+    For the purposes of this paper, I focus mainly on executive teams as a whole instead of specific positions.
+The main reason for this is because the organization of teams varies so drastically across hospitals that a specific position, even if it exists in two different hospitals, can be meaningfully different.
+Additionally, the more granular the information I extract, the more likely there will be a mistake in the OCR extraction.
+Thus, to preserve correctness of the independent variable of interest, I focus on whether there is any MD executive at the hospital.
+
+    I drop anyone who is solely a board member, and focus on those who have executive in their title, or are labeled as a president or vice president of a department.
 I create various hospital-level characterizations based on their executive team, such as the existence of a clinically trained executive, the number of clinically trained executives, the number of total executives, and whether the hospital has a Chief Medical Officer.
-Additionally, I examine changes in executive teams to determine whether the hospital changes its propensity to hire a clinically trained executive.
+Additionally, I examine changes in propensity to hire a clinical executive.
 Summary statistics for these variables are shown in Table \ref{leader_sumstats}.
 On average, NFP hospitals in my sample have 4 executives and .4 clinically trained executives.
-Sixty-one percent of the sample never changes propensity to hire a clinically trained executive, and 33\% of NFPs in the sample have a Chief Medical Officer.
+Sixty-one percent of the sample never changes propensity to hire a clinically trained executive, and 33\% of NFPs in the sample have a Chief Medical Officer at some point.
 
@@
 
-    I merge the created sample of not-for-profits with all for-profits in the AHA survey.
+    I combine the created sample of not-for-profits with all for-profits in the AHA survey.
 The outcomes I consider are readmission and mortality rates, which are directly targeted in the pay-for-performance policies, and are found in the CMS Hospital Compare data.
-I combine readmission and mortality rates for pneumonia, AMI, and heart failure (the relevant HRRP conditions) into a weighted average based on the number of patients in each condition.
-Additionally, I include several relevant hospital characteristics.
-Number of beds, whether the hospital is an academic medical center, is physician owned, or is system affiliated, all come from the AHA survey.
-Uncompensated care comes from HCRIS, and case mix index, which measures patient complexity, comes from CMS.
+I combine rates for pneumonia, AMI, and heart failure (the relevant HRRP conditions) into a weighted average based on the number of patients in each condition.
+Additionally, I include several relevant hospital characteristics: number of beds, whether the hospital is an academic medical center, is physician owned, or is system affiliated, all come from the AHA survey.
+Case mix index, a measure of overall patient complexity, comes from CMS.
 Whether the hospital was penalized under HRRP or received payments under HVBP is found in the HCRIS data.
@@
 
-    I present a table of means for each hospital type (for-profit, in-sample not-for-profits, and not-for-profits with and without a clinically trained executive) in Table \ref{tab:sumstats_samples}.
+    There are several samples of hospitals that are relevant for the analysis.
+First, there are all for-profits and all not-for-profits.
+Further, I subset the sample of NFPs into whether they have a clinical executive at some point in the sample.
+Finally, I characterize a NFP leadership team as being stable or not.
+A stable team means the hospital did not change their propensity to hire a clinical executive over time.
+I present a table of means for each of these sub-samples in Table \ref{tab:sumstats_samples}.
 Approximately 850 NFPs have leadership information, with 384 of them hiring an MD executive at some point in the sample.
-I compare these NFPs with the 740 for-profits from the AHA Survey.
+I compare these NFPs with the 708 for-profits from the AHA Survey.
 For-profits and NFPs are roughly the same bed size, but NFPs see more patients in each of the three relevant conditions.
@@
 Those with MDs are larger, more likely to be penalized, have higher uncompensated care, and more complex patients.
+This holds for stable leadership teams as well, though the number of NFPs with a clinical executive drops significantly.
 I discuss the implications of these differences for identification in Section \ref{sec:model}.
@@
 
-    I model hospital behavior in two simplified scenarios: where quality does not directly affect profit, and where it does.
+    I model hospital behavior under two simplified worlds: one where quality does not directly affect profit, and one where it does.
 Pay-for-performance incentives essentially incorporate performance directly into a hospital's profit function.
-Here, I model performance as quality of care, which is inversely related to readmission and mortality rates.
-While the literature speculates about the true form of NFP objective functions, we can generally think of hospitals choosing some linear combination of profit and societal benefit.
+I directly model performance as quality of care, which is inversely related to readmission and mortality rates.
+While the literature speculates about the true form of NFP objective functions, we can generally think of hospitals choosing some linear combination of profit and societal benefit, without knowing how much weight the hospital places on each.
 Intuitively, a hospital that places only weight on profit (for-profit type) will only care about quality when it enters into the profit function, and therefore will respond more drastically to pay-for-performance than a hospital who already cared about quality through societal benefit before the incentives.
-Therefore, hospital response to pay-for-performance incentives reveals something about their underlying preference on profit vs.
+Therefore, hospital responses to pay-for-performance incentives reveal information about their underlying preference on profit vs.
 societal outcomes.
 
-    To formalize this intuition, I specify the hospital objective function as a weighted average of profit and societal (or patient) outcomes and derive characteristics of the chosen quality.
+    To formalize this intuition, I specify the hospital objective function as a weighted average of profit and societal (or patient) outcomes and derive characteristics of the optimal quality.
 For simplicity, I abstract away from quality having a direct effect of quantity and price.
@@
 
-    \noindent marginal benefit equals marginal cost of increasing quality. Solving for $u'(\theta)$ and differentiating with respect to $\alpha$ yields
+    \noindent where marginal benefit equals marginal cost of increasing quality. Solving for $u'(\theta)$ and differentiating with respect to $\alpha$:
 
@@
 
-    Thus,
+    Therefore, again using the Implicit Function Theorem,
 
@@
 
-    Therefore, in a world with financial incentives on quality, quality is increasing with more weight placed on profit.
+    That is, all else equal, in a world with financial incentives on quality, quality is increasing with more weight placed on profit.
 I combine the results found in each scenario into one response function that depends on $\alpha$, where $\theta_2$ is quality under pay-for-performance incentives and $\theta_1$ is quality with no financial incentive to quality.
@@
 However, not-for-profit behavior on average does not bring much clarity about their objectives.
-An observable and economically relevant characteristic of a hospital that could affect $\alpha$ is composition of their leadership team.
+An observable and economically relevant characteristic of a hospital that could affect $\alpha$ is composition of their executive leadership team.
 Thus, I compare the response of for-profits and not-for-profits with different leadership teams to bring clarity on how this might point to underlying $\alpha$.
@@
 
-    In light of the theoretical predictions, I first test empirically whether NFPs respond to program incentives differently than for-profits. I estimate the following regression equation:
+    In light of the theoretical predictions, I test empirically whether different hospital types respond to pay-for-performance incentives differently.
+I present here linear models with a two-way fixed effects specification.
+Ultimately, I employ synthetic difference-in-differences, which is estimating the same model with weights that more heavily factor similar trending control groups and time periods.
+However, for clarity, I present the unweighted specification:
 
@@
     \label{eq:forprofit}
-    y_{ht} = \beta (\text{For-Profit}_h \times \text{Post Programs}_t) + \gamma_{h} + \delta_t + \epsilon_{ht}
+    y_{ht} = \beta \text{ treat}_{ht} + \gamma_{h} + \delta_t + \epsilon_{ht},
     \end{equation}
 
-    \noindent where $y_{ht}$ is one of the outcome variables discussed in Section \ref{sec:data}, $\beta$ is the coefficient of interest capturing the effect of being for-profit after the programs were introduced, and $\gamma_h$ and $\delta_t$ are hospital and time fixed effects, respectively.
-For each outcome, I estimate three models with differing comparison groups: all NFPs, NFPs with a clinically trained executive, and NFPs without a clinically trained executive.
-I also directly estimate the difference in response of NFPs with and without a clinically trained executive:
-
-    \begin{equation}
-    \label{eq:clinical}
-    y_{ht} = \beta (\text{No MD Exec}_h \times \text{Post Programs}_t) + \gamma_{h} + \delta_t + \epsilon_{ht}
-    \end{equation}
+    \noindent where $y_{ht}$ is one of the outcome variables discussed in Section \ref{sec:data}, and $\gamma_h$ and $\delta_t$ are hospital and time fixed effects, respectively.
+The independent variable of interest, treat$_{ht}$, captures the effect of being a certain hospital type after incentives changed in 2012.
+That is, I interact an indicator for hospital type with an indicator for year post 2012 for various hospital types.
+First, I define treat so that for-profit hospitals are being compared to subsets of not-for-profits (by limiting the sample to the relevant treatment and comparison group).
+Then, I limit to only not-for-profits and define treat as those without a clinical executive.
+Specifically, I define treat$_{ht}$ and limit the sample in four ways, shown in Figure \ref{fig:spec}.
 
@@
 
+\begin{figure}[ht!]
+    \begin{center}
+    \caption{\label{fig:spec}Treated and Comparison Groups}
+ \begin{tabular}{| m{18em} |}
+ \hline
+ Combination 1:\\ [0.5ex]
+ \hline\hline 
+ \vspace{2mm}
+ Treat Group: \hspace{23.5mm} For-Profit \\
+ \vspace{2mm} 
+ Comparison Group: \hspace{11.5mm} Stable NFPs  \\
+ [1ex]
+ \hline
+ \end{tabular}
+\hfil   %<---
+ \begin{tabular}{|m{18em}|}
+ \hline
+ Combination 2:\\ [0.5ex]
+ \hline\hline
+ \vspace{2mm}
+ Treat Group: \hspace{13.5mm} For-Profit \\
+ \vspace{2mm}
+ Comparison Group: \hspace{1.5mm} Stable NFP w/out MD  \\
+ [1ex]
+ \hline
+ \end{tabular}
+
+\medskip
+
+\vspace{2mm}
+
+  \begin{tabular}{|m{18em}|}
+ \hline
+ Combination 3:\\ [0.5ex]
+ \hline\hline
+ \vspace{2mm}
+ Treat Group: \hspace{17mm} For-Profit \\
+ \vspace{2mm}
+ Comparison Group: \hspace{5mm} Stable NFPs w/ MD  \\
+ [1ex]
+ \hline
+ \end{tabular}
+\hfil   %<---
+  \begin{tabular}{|m{18em}|}
+ \hline
+ Combination 4:\\ [0.5ex]
+ \hline\hline
+ \vspace{2mm}
+ Treat Group:  \hspace{13mm} Stable NFPs w/out MD \\
+ \vspace{2mm}
+ Comparison Group:  \hspace{1mm} Stable NFPs w/ MD  \\
+ [1ex]
+ \hline
+ \end{tabular}
+ \end{center}
+ \end{figure}
+
     Additionally, under certain assumptions that I discuss in Section \ref{sec:identification}, using executive team changes can be advantageous in better understanding underlying hospital motives.
@@
 Second, hiring a clinically trained executive who carries out day-to-day operations in a distinct way effectively changes $\alpha$.
-Thus, I use two types of variation in clinically trained executives to disentangle the two: a variable for whether the hospital ever hires a clinically trained executive, and a variable for whether the hospital has a clinically trained executive in 2012 when the programs took place.
-I use the same empirical specification with these two variables as the independent variation in executive teams, and limit the sample to the relevant comparison group:
+Thus, I also define treat and limit the comparison group in such a way that decomposes the two.
+Without limiting to stable teams, I consider hospitals that ever hire a clinically trained executive vs.
+hospitals that have a clinically trained executive in 2012 when the programs took place.
+These variations of Equation \ref{eq:forprofit} are shown in Figure \ref{fig:decomp_spec}.
 
-    \begin{equation}
-    \label{eq:decomp}
-    y_{ht} = \beta (\text{Never/No 2012 Exec}_h \times \text{Post Programs}_t) + \gamma_{h} + \delta_t + \epsilon_{ht}
-    \end{equation}
+    \begin{figure}[ht!]
+    \begin{center}
+    \caption{\label{fig:decomp_spec}Decomposition Model Specification Details}
+
+ \begin{tabular}{| m{18em} |}
+ \hline
+ Decomposition 1:\\ [0.5ex]
+ \hline\hline 
+ \vspace{2mm}
+ Treat Group:  \hspace{15mm} NFP w/out MD \\
+ \vspace{2mm}
+ Comparison Group: \hspace{3mm} NFPs w/ MD  \\
+ [1ex]
+ \hline
+ \end{tabular}
+\hfil   %<---
+ \begin{tabular}{|m{18em}|}
+ \hline
+ Decomposition 2:\\ [0.5ex]
+ \hline\hline
+ \vspace{2mm}
+ Treat Group: \hspace{11mm} No MD Exec 2012 \\
+ \vspace{2mm}
+ Comparison Group:  NFPs w/ MD (not 2012)  \\
+ [1ex]
+ \hline
+ \end{tabular}
+
+\end{center}
+ \end{figure}
 
@@
     To identify the causal effect of hospital type on hospital program response, I rely on several assumptions.
-First, I assume that, absent the program enactments, different hospital types would have had parallel trends in outcomes.
+First, I assume that, absent the program enactments, treat and comparison group types would have behaved similarly in outcomes.
 While I see no evidence of pre-trends, Table \ref{tab:sumstats_samples} suggests that hospitals with a clinically trained executive are different than other hospitals along multiple characteristics.
-Therefore, there may be concern that these differences between hospital types could be driving the results.
+There may be concern that these differences between hospital types could be driving the results.
 Therefore, my main specification uses the synthetic difference-in-differences method established in \citeauthor{arkhangelsky2021synthetic} (\citeyear{arkhangelsky2021synthetic}).
-This method creates unit weights that more heavily count comparison hospitals that are similar to treatment hospitals prior to the programs.
-It also weighs time periods that balance pre-program and post-program for the comparison hospitals.
+This method creates unit weights that more heavily factor comparison hospitals that are similar to the treat group hospitals prior to the programs.
+It also weighs time periods that balance pre-programs and post-programs for the comparison hospitals.
 Then, a simple two way fixed effects estimator is used with these weights, yielding a more robust average treatment effect that relies less on a strong parallel trends assumption (\cite{arkhangelsky2021synthetic}).
-I present the synthetic diff-in-diff estimates as the main results, and they are largely similar to the standard two way fixed effects results in Appendix \ref{app:fullsample}.
-Further, while ultimately I employ a synthetic difference-in-differences method, the conceptual identification is the same as a standard difference-in-differences.
+I present the synthetic diff-in-diff estimates as the main results, and they are largely similar to the standard two way fixed effects results with no weights, shown in Appendix \ref{app:fullsample}.
+Further, the conceptual identification is the same as a standard difference-in-differences.
 
@@
 That is, are hospitals that expect to be affected by the programs more likely to hire or fire clinically trained executives?
-I consider changes along having any clinically trained executive as well as changes in the number of clinically trained executives.
-I estimate the following regression equations:
+I consider changes in having any clinically trained executive, as well as changes in the number of clinically trained executives.
+I estimate the following linear equations:
 
@@
     \begin{equation}\label{eq:change2}
-    \text{change}_{ht} = \beta(\text{program\_exposed}_{h} \times \text{Post Programs}_t)+ \alpha_h + \delta_t + \epsilon_{ht}.
+    \text{change}_{ht} = \beta(\text{Program Exposed}_{h} \times \text{Post Programs}_t)+ \alpha_h + \delta_t + \epsilon_{ht}.
     \end{equation}
 
-    The variable $\text{program\_exposed}_{h}$ is an indicator for whether the hospital eventually was penalized under HRRP or received payments under HVBP.
+    The variable $\text{Program Exposed}_{h}$ is an indicator for whether the hospital eventually was penalized under HRRP or received payments under HVBP.
 The estimates from this analysis are presented in Table \ref{tab:change_analysis}.
@@
 
-     The estimated difference in readmission rate trends for all hospital types in response to the programs, along with ATE and standard errors, are shown in Figure \ref{fig:read_synth_plot}.
-I first discuss the differential program response of for-profit firms compared to the different NFP sub-samples (Equation \ref{eq:forprofit}), shown in panels \ref{fig:read_synth_plota}-\ref{fig:read_synth_plotc}.
+     The estimated difference in readmission rate trends for all hospital types in response to the programs are shown in Figure \ref{fig:read_synth_plot}.
+Each panel corresponds to a particular definition of the treat group of hospitals and limitation of comparison group hospitals.
 
-     Panel \ref{fig:read_synth_plota} shows that on average, there is no difference in how for-profits and not-for-profits change their readmission rates in response to the incentive changes.
-However, when NFPs are subset into whether they have a clinical executive or not, panel \ref{fig:read_synth_plotb} shows that there is a differential response to the program between for-profits and NFPs with a clinical executive.
+     Panel \ref{fig:read_synth_plota} compares readmission rate responses among for-profits and all NFPs, and shows that on average, there is no difference in how these types of hospitals change readmission rates in response to the incentive changes.
+However, when the comparison group is limited to NFPs with a clinical executive, panel \ref{fig:read_synth_plotb} shows that there is a differential response to the program between these hospital types.
 In particular, for-profits decrease readmission rates by .3 more than NFPs with a clinical executive, an additional 3 patient readmission decrease relative to the average number of patients.
-While numerically small, this difference in response is non-existent between for-profits and NFPs without a clinical executive (panel \ref{fig:read_synth_plotc}), indicating some, if minor, difference in objectives resulting from having a clinical executive.
+While numerically small, this difference in response is non-existent between for-profits and NFPs without a clinical executive (panel \ref{fig:read_synth_plotc}), indicating some, if minor, differences in objectives resulting from having a clinical executive.
 
@@
 
-    Next, in panel \ref{fig:read_synth_plotd}, I present estimates of Equation \ref{eq:clinical}, directly comparing NFPs with and without a clinical executive.
+    Next, in panel \ref{fig:read_synth_plotd}, I directly compare NFPs with and without a clinical executive.
 NFPs without a clinically trained executive lowered readmission rates by .24 ppts more than NFPs with a clinically trained executive.
@@
 However, when comparing for-profits to NFPs with a clinical executive, the magnitude of the difference in mortality trends jumps to -.11.
-While still statistically insignificant, I argue that this still indicates a difference in behavior between the hospital types, particularly due to the documented noisiness of mortality measures (add citations).
+While still statistically insignificant, I argue that this still indicates a difference in behavior between the hospital types, particularly due to the documented noisiness of mortality measures (\cite{mackenzie2016measuring}).
 
@@
     I also directly estimate the difference in mortality rates among NFPs with and without clinical experience in panel \ref{fig:mort_synth_plotd}.
-Again, the difference is not statistically significant, but the coefficient compared to the true nulls in panels \ref{fig:mort_synth_plota} and \ref{fig:mort_synth_plotc} indicates NFPs without a clinical executive potentially decreasing mortality more aggressively than NFPs with a clinical executive.
+Again, the difference is not statistically significant, but the coefficient compared to the true nulls in panels \ref{fig:mort_synth_plota} and \ref{fig:mort_synth_plotc} indicates NFPs without a clinical executive potentially decrease mortality more aggressively than NFPs with a clinical executive.
 
@@
 However, the analysis to this point does not speak to the mechanism by which clinical executives lead to differential behavior.
-In this section, I present two theories by which a clinical executive leads to differential behavior, and present descriptive evidence of these empirically.
+In this section, I present two theories by which a clinical executive leads to differential behavior, and present descriptive evidence that one dominates the other.
 
     First, having a clinical executive could be a signal of underlying hospital objectives.
-That is, a hospital who hires a clinical executive is reveals their underlying objectives.
-Alternatively, hiring a clinical executive could actually change the objectives of the hospital for the time they are actively managing the hospital.
-To disentangle the two, I present the results of specification \ref{eq:decomp} in Table \ref{tab:MD_noMD_readmort_decomp_synth}, where the variation in executives is either existence of a hospital executive at any point in time, or whether the hospital has a clinically trained executive in 2012 when the programs were enacted.
+Alternatively, hiring a clinical executive could actually change the objectives of the hospital for the time the clinical executive is actively managing the hospital.
+To disentangle the two, I use variation in the timing of having a clinical executive, with either existence of a hospital executive at any point in time, or whether the hospital has a clinically trained executive in 2012 when the programs were enacted.
+The results are shown in Table \ref{tab:MD_noMD_readmort_decomp_synth}.
 Columns (1) and (2) show that the difference in readmission response is driven by having a clinically trained executive in 2012, compared to having a clinically trained executive at another point in time.
@@
 
-    Existing literature suggests that not-for-profit hospitals may or may not be meaningfully different from for-profits.
-Along some dimensions, they seem to have meaningful dimensions, but we still do not understand underlying objectives of not-for-profit hospitals, which ultimately affect cost and quality of health care.
-One dimension of NFP hospitals that has not been studied is how manager composition affects objectives and behaviors of the firm.
+    Existing literature is inconclusive as to whether not-for-profit (NFP) hospitals differ meaningfully from for-profit hospitals.
+NFP objectives have direct implications for patients, yet are largely unknown.
+An observable and economically meaningful characteristic of NFP hospitals that may affect underlying objectives is the composition of the hospital's top management.
+Particularly, whether any executives at the hospital have a background in practicing medicine.
 
-    In this paper, I study differential response to a change in incentives between for-profit and NFP hospitals with and without a clinical executive.
+    In this paper, I study differential responses to a change in incentives between for-profit and NFP hospitals with and without a clinical executive.
 I do this by collecting a novel data set on hospital executives from Tax Form 990s and merging this to existing hospital data.
-I then estimate differential responses in readmission and mortality rates after several pay-for-performance initiatives are enacted in the US in 2012 using a synthetic difference in differences methodology.
+I then use a synthetic difference-in-differences methodology to estimate differential responses in readmission and mortality rates after several pay-for-performance initiatives are enacted in the US in 2012.
 
-     I find that not-for-profits with a clinical executive respond differently than both for-profits and NFPs without a clinical executive.
+     When comparing for-profits to all not-for-profits, I find no difference in behavior after the policies take effect, which aligns with suggestions that NFPs are for-profits in disguise.
+However, when I compare for-profits to NFPs with clinical executives, the two types of hospitals respond differently to the incentives.
 Specifically, all firm types decrease readmissions, but NFPs with a clinical executive are less aggressive in their responsiveness than the other firm types.
 A similar result holds for mortality rates.
-This finding is consistent with the theoretical prediction that firms who place more weight on societal benefit than on profit will have a more mitigated response to policies affecting the profit function.
+This finding is consistent with the theoretical prediction that firms who place more weight on societal benefit than on profit will have a more mitigated response to policies that affect profit.
 Additionally, under the assumption that executive team changes are not correlated with the program, I find that the differential response is driven by the actual management of the clinical executive, not by a clinical executive being a signal of underlying objectives.
-These findings suggest that policymakers should take into account more than just ownership status when evaluating how hospitals will respond to incentives.
+
+     These findings are informative for policymakers predicting how hospitals will responsd to a change in incentives.
+Specifically, they suggest that policymakers should take into account more than just ownership status when evaluating how hospitals will respond to policy changes.
 
@@
 
-There is no perfect way to access tax form 990s in bulk over the time period I am considering in this paper.
-However, using the not-for-profit Explorer API seems to be the most straightforward.
+I use the Nonprofit Explorer API to access the archive of NFP tax form 990s.
 At the time of writing this, information on using version 2 of the API can be found at \hyperlink{https://projects.propublica.org/not-for-profits/api}{https://projects.propublica.org/not-for-profits/api}.
@@
 
-It is crucial to the analysis to be able to link these not-for-profits with other sources of data to recover penalties from HRRP, payments from HVBP, bed size, and outcomes of interest.
+It is necessary to link these not-for-profits with other sources of data to recover penalties from HRRP, payments from HVBP, bed size, and outcomes of interest.
 The ideal hospital data set to match to is American Hospital Association (AHA) Survey, which contains hospital characteristics and Medicare ID number.
@@
 
-\section{Hospitals Matched to AHA vs. Not}\label{app:matched}
+\section{Matched vs. Unmatched Not-for-Profits}\label{app:matched}
+
+The sample of all NFP hospitals that I use for analysis consists of 852 hospitals.
+In the AHA data, there are 2,258 additional NFP hospitals that either were not matched to tax forms, did not exist in the tax form data, or did not have executive names in the tax forms.
+While it would be ideal to analyze all NFPs, a sample size of 852 firms is relatively large in the hospital executive literature.
+
+In this section, I compare the in-sample NFPs to the out-of-sample NFPs along observable measures.
+I present means and standard deviations of these measures in Table \ref{tab:nfp_sample_compare}.
+The in-sample NFPs are slightly larger in terms of beds and patients seen.
+They are also slightly more likely to be penalized or receive payments under the pay-for-performance incentives.
+However, average readmission and mortality rates are very similar, as well as case mix index.
+The largest difference in the two samples is that I under-represent NFP hospitals in systems.
+This is due to the nature of the tax form 990s, where systems often file one tax form for the entire system, making it difficult to discern the true managers of a specific hospital.
+Because of this, I drop hospitals from the sample who only have leadership information from a system tax form.
 
@@
 
-\section{Full Sample Regression and Event Study Results}\label{app:fullsample}
+\section{TWFE Regression and Event Study Results}\label{app:fullsample}
+
+While I ultimately employ a synthetic difference-in-differences framework, I also present the coefficient estimates and event study tables under a typical TWFE model specification.
+That is, there is no weighting of the control group or time periods.
+Table \ref{tab:forprofit_readmort_fullsample} shows coefficient estimates for the three specifications under Equation \ref{eq:forprofit}.
+Columns (1)-(3) show coefficient estimates for the weighted average readmission rate outcome, and columns (4)-(6) show coefficient estimates from the weighted average mortality outcome.
+Generally, these results are qualitatively identical to the synthetic difference-in-difference specification results.
+There is a statistically significant difference in readmission rate response between for-profits and NFPs with a clinical executive.
+Specifically, for-profits decrease their readmission rates by .4ppts more than NFPs with a clinical executive.
+There is no difference (statistically or in magnitude) between readmission rate responses of for-profits compared to all NFPs and NFPs without clincal executives.
+
+Similarly to the synthetic diff-in-diff estimates, there is no statistical difference in mortality between any of the hospital types, but the magnitude difference between for-profits and NFPs with clinical executives is much larger.
+This oculd be due to the noisiness of mortality as a measure.
+
+\import{Tables}{forprofit_readmort_fullsample.tex}
+
+Further, I present coefficient estimates from Equation \ref{eq:clinical} in Table \ref{tab:MD_noMD_readmort_fullsample}, directly comparing NFPs with and without clinical executives for both readmission and mortality rates.
+Again, these results are very similar to the main results of the paper, with only slightly larger magnitudes.
+NFPs without a clinical executive respond to the pay-for-performance incentives by decreasing readmission and mortality rates more aggressively than NFPs with clinical executives.
 
@@
 
-\import{Tables}{forprofit_readmort_fullsample.tex}
+Finally, I present these results in typical event study form in Figure \ref{fig:es_plot}, which show no evidence of significant pre-trends driving the findings.
 
-\section{Synthetic Diff-in-Diff Result Tables}
+\begin{figure}[ht!]
+     \caption{TWFE Event Study Results}
+     \centering
+          \begin{subfigure}[b]{0.45\textwidth}
+         \centering
+         \caption{}
+         \includegraphics[width=\textwidth]{Objects/read_forprofit_es_graph.pdf}
+         \label{fig:es_plota}
+     \end{subfigure}%
+     \hfill
+     \begin{subfigure}[b]{0.45\textwidth}
+         \centering
+         \caption{}
+         \includegraphics[width=\textwidth]{Objects/read_MD_es_graph.pdf}
+         \label{fig:es_plotb}
+     \end{subfigure}%
+     \vspace{5mm}
+     \hfill
+     \begin{subfigure}[b]{0.45\textwidth}
+         \centering
+         \caption{}
+         \includegraphics[width=\textwidth]{Objects/mort_forprofit_es_graph.pdf}
+         \label{fig:es_plotc}
+     \end{subfigure}
+     \hfill
+     \begin{subfigure}[b]{0.45\textwidth}
+         \centering
+         \caption{}
+         \includegraphics[width=\textwidth]{Objects/mort_MD_es_graph.pdf}
+         \label{fig:es_plotd}
+     \end{subfigure}
+        \label{fig:es_plot}
+    \end{figure}
+
+\section{Synthetic Difference-in-Difference Coefficient Tables}
+
+While I present the graphical representation of results in the main text, I also present a table of coefficients and standard errors for readmission and mortality rates in Tables \ref{tab:forprofit_readmort_synth} and \ref{tab:MD_noMD_readmort_synth}.
 

### 2024-05-30 Tables/MD_noMD_readmort_synth.tex
fix sample of nonprofits and add synthetic difference in differences graphs
Updates from Overleaf

@@
 \midrule
-No MD Exec x Post Programs & -0.28$^{*}$ & -0.19\\
- & (0.13) & (0.11)\\
+NFP w/out MD Exec x Post Program & -0.28$^{*}$ & -0.19\\
+ & (0.12) & (0.12)\\
 \textbf{Comparison Group:} &  & \\
-NP w/ MD Exec & $\checkmark$ & $\checkmark$\\
+NFP w/ MD Exec & $\checkmark$ & $\checkmark$\\
  &  & \\
@@
 \bottomrule
-\multicolumn{3}{l}{\textsuperscript{} Results from Equation \ref{eq:clinical} with readmission and mortality rates as outcome variables.}\\
+\multicolumn{3}{l}{\textsuperscript{} Results from Equation with readmission and mortality rates as outcome variables.}\\
 \multicolumn{3}{l}{\textsuperscript{} Standard errors are clustered at the hospital level.}\\

### 2024-05-31 Paper/paper.tex
Updates from Overleaf

@@
   While ownership is certainly relevant, there could be other characteristics correlated with hospital behavior independent of ownership type.
-Researchers have focused in particular on explaining variation in NFP behavior by the market they exist in (citations).
-However, I argue that an observable and economically meaningful factor is the composition top-level hospital managers.
+Researchers have focused in particular on explaining variation in NFP behavior using variation in the market they exist in (\cite{horwitz2009hospital}).
+However, I argue that an observable and economically meaningful factor is the composition of top-level hospital managers.
 In large, publicly traded firms, executive characteristics are shown to be correlated with firm performance, indicating that there is more to firm behavior than ownership status (\cite{bertrand2003managing}; \cite{matsa2013female}; \cite{ahern2012changing}).
@@
   I estimate how the interaction between hospital type and post program enactment affects readmission and mortality rates.
-Under the assumption that hospitals would have continued to behave similarly in the absence of the programs, and that leadership teams do not change endogenously with the programs, I identify the differential response to pay-for-performance incentives by different hospital types.
-Among the hospitals in the data, hospitals with clinically trained executives are the minority, and differ on average from other hospitals along various dimensions.
-Thus, I ultimately employ synthetic difference-in-differences to analyze a group of comparison hospitals with similar characteristics to treatment hospitals.\footnote{I also present results for the full sample, and find qualitatively similar results.}
+Under the assumption that hospitals would have continued to behave similarly in the absence of the programs, and that leadership teams do not change endogenously with the programs, I identify the differential response to pay-for-performance incentives among hospital types.
+That is, I define a group of treatment and comparison hospitals, and estimate how outcomes differ between the two groups after the policy changes.
+I ultimately employ synthetic difference-in-differences to analyze a group of comparison hospitals with similar characteristics to treatment hospitals.\footnote{I also present results for the full sample, and find qualitatively similar results.}
 
@@
 That is, NFPs without clinically trained executives respond to incentives more similarly to FPs than other NFPs.
-Using changes in executive teams, I decompose this effect into signalling (clinical executives signal underlying hospital preferences), or managing (clinical executives manage the hospital differently) effects.
+Leveraging changes in executive teams, I decompose this effect into signalling (clinical executives signal underlying hospital preferences), or managing (clinical executives manage the hospital differently) effects.
 I find that the entire difference in response is driven by a managing effect.
@@
     In light of the theoretical predictions, I test empirically whether different hospital types respond to pay-for-performance incentives differently.
-I present here linear models with a two-way fixed effects specification.
+I present here a two-way fixed effects specification, where the key variable is an interaction between a certain hospital type and time periods after program enactments.
 Ultimately, I employ synthetic difference-in-differences, which is estimating the same model with weights that more heavily factor similar trending control groups and time periods.
@@
 managing effects, which is only identified if leadership changes are uncorrelated with the policies.
+In Appendix \ref{app:changes}, I present various descriptive statistics of executive team changes over time, showing that hiring and firing of MD executives do not change much over time, the rate of clinical executives to total executives remains stable over time, and the distribution of timing of clinical executive changes.
 
-     I assess the validity of this assumption in two ways.
+     I assess the validity of the exogenous changes assumption in two ways.
 First, I analyze whether changes in clinically trained executives occur more often in time periods of program enactment.
@@
 
+\section{Executive Team Changes}\label{app:changes}
+
+Here, I present descriptive statistics of executive team changes.
+In Figure \ref{fig:change_graph}, I show the average rate of MD executives to total executives, the percent of hospitals who fire an MD, and the percent of hospitals who hire an MD.
+The fraction of MD executives remains stable over time at approximately .22.
+Similarly, the percent of hospitals who hire an MD over time remains stable at approximately 21\%.
+There is a small increase in the percent of hospitals who fire an MD in 2011, but even with this increase firing is not common and stable, ranging from 12\% to 13\%.
+
+\begin{figure}[ht!]
+    \centering
+    \caption{Executive Team Changes Over Time}
+    \includegraphics[width=\textwidth]{Objects/change_graph.pdf}
+    \label{fig:change_graph}
+\end{figure}
+
+Additionally, I present the number of hospitals who change their propensity to hire a clinically trained executive at different times in Table \ref{tab:change_timing}.
+I break the timing up into pre-2012, 2012, and post-2012.
+I show the number of hospitals who have MDs in different time combinations.
+The number of hospitals that only have an MD executive in 2012, but at no other time, is only 6.
+The majority of hospitals (187) who ever have a clinical executive have one in every time period.
+
+\import{Tables}{change_num_table_2012.tex}
+
+\section{Results By Condition}
+
+\import{Tables}{fp_read_condition_synth.tex}
+
+\import{Tables}{fp_mort_condition_synth.tex}
+
+\import{Tables}{MD_noMD_readmort_condition_synth.tex}
+
 \section{TWFE Regression and Event Study Results}\label{app:fullsample}
@@
 
-Further, I present coefficient estimates from Equation \ref{eq:clinical} in Table \ref{tab:MD_noMD_readmort_fullsample}, directly comparing NFPs with and without clinical executives for both readmission and mortality rates.
+Further, I present coefficient estimates in Table \ref{tab:MD_noMD_readmort_fullsample} directly comparing NFPs with and without clinical executives for both readmission and mortality rates.
 Again, these results are very similar to the main results of the paper, with only slightly larger magnitudes.

### 2024-06-10 Paper/paper.tex
Updates from Overleaf

@@
 		{\small
-        In many industries, not-for-profits (NFP) are the dominant firm type.
-Understanding the objectives of these firms is important, but NFP firms have underlying goals that make understanding their objectives difficult.
-For hospitals in particular, there is mixed evidence on whether NFP hospitals differ meaningfully from for-profit (FP) hospitals.
+        There is mixed evidence on whether not-for-profit hospitals differ meaningfully from for-profit (FP) hospitals, but what has not been considered is the effect leadership teams have on hospital behavior.
 In this paper, I investigate whether leadership team composition of NFP hospitals affects how similarly they behave to for-profits.
 I use a novel data set on NFP hospital executives in the US (2010-2014) to investigate whether the presence of clinically trained executives affects hospital behavior.
-In particular, I leverage pay-for-performance initiatives in the US enacted in 2012, and measure the  readmission and mortality rate responses between FP, NFP, and NFP with and without clinically trained executives.
-I find that NFPs without clinically trained executives behave more like FPs than other NFPs.
-This is consistent with theoretical predictions if NFP hospitals with clinically trained executives value societal outcomes more than those without clinically trained executives.
+In particular, I leverage pay-for-performance initiatives in the US enacted in 2012, and measure the  readmission and mortality rate responses between FP and NFP with and without clinically trained executives.
+I find that for-profits nad NFPs without clinical executives respond more aggressively to financial incentives on quality, consistent with theoretical predictions of firms that value profit relatively more than having high quality.
 		} 
@@
 
-  The objectives and behaviors of firms that are not classic for-profits are not easily determined, and an extensive literature has sought to understand these types of firms.
-Not-for-profits (NFP) are thought to gain utility not solely from monetary profit, but also through accomplishing societal benefit.
-Researchers have speculated a variety of objective functions that could motivate NFP firms such as maximizing profit (for-profit in disguise), prestige, income, or a quality/quantity trade-off (\cite{steinberg1986revealed}).
+  The objectives and behaviors of not-for-profit firms are challenging to determine, and an extensive literature has sought to understand them.
+Researchers have posited a variety of objective functions that could motivate NFP firms such as maximizing profit (for-profit in disguise), prestige, income, or a quality/quantity trade-off (\cite{steinberg1986revealed}).
 Empirical contributions to this literature center around investigating behaviors that differ between for-profit (FP), private NFP, and government firms (\cite{sloan2000not}).
+However, the literature has neglected to consider other relevant characteristics of hospitals that could affect objectives, such as the hospital's leadership team.
 In this paper, I investigate whether differences in observed behavior between for-profit and not-for-profit hospitals vary differentially by leadership team composition.
@@
 Second, private NFP hospitals make up 50\% of all hospitals in the US, and staff, on average, 207 beds.
-For comparison, FPs make up around 36\% of hospitals and staff 107 beds (\cite{ASPE_2023}).
+For comparison, FPs make up 36\% of hospitals and staff 107 beds (\cite{ASPE_2023}).
 Thus, NFP behavior in the hospital industry is relevant, and directly affects consumers of health care.
@@
   While ownership is certainly relevant, there could be other characteristics correlated with hospital behavior independent of ownership type.
-Researchers have focused in particular on explaining variation in NFP behavior using variation in the market they exist in (\cite{horwitz2009hospital}).
+Researchers have focused in particular on explaining variation in NFP behavior using variation in the market they exist in (\cite{duggan2002hospital}; \cite{horwitz2009hospital}).
 However, I argue that an observable and economically meaningful factor is the composition of top-level hospital managers.
-In large, publicly traded firms, executive characteristics are shown to be correlated with firm performance, indicating that there is more to firm behavior than ownership status (\cite{bertrand2003managing}; \cite{matsa2013female}; \cite{ahern2012changing}).
-If hospitals differ in how they value monetary profit vs.
-societal benefit, then ownership status certainly plays a role in these valuations, but the question remains as to whether management also plays an important role.
-I contribute to our understanding of hospital and NFP behaviors by investigating whether occupational background, specifically clinical training in executives, affects hospital response to pay-for-performance incentives.
+In large publicly traded firms, executive characteristics are shown to be correlated with firm performance and governance, indicating that there is more to firm behavior than ownership status (\cite{bertrand2003managing}; \cite{ahern2012changing}\cite{matsa2013female}).
+
+  A policy relevant aspect of hospital behavior in the US has been the shift towards pay-for-performance incentives.
+While quality may or may not directly matter to hospitals, pay-for-performance incentives are such that quality of care directly impacts profit.
+For hospitals that already valued quality before pay-for-performance incentives, the change in incentives may not change quality much.
+However, if the hospital is more profit driven, the policy change should translate to a larger change in quality.
+Ownership status certainly plays a role in the valuations of profit vs.
+quality, but the question remains as to whether management also plays an important role.
+I contribute to our understanding of hospital and NFP behaviors by investigating whether occupational background of executives, specifically clinical training, affects hospital response to pay-for-performance incentives.
 
@@
 
-  I estimate how the interaction between hospital type and post program enactment affects readmission and mortality rates.
+  I estimate how the interaction between hospital type and pay-for-performance incentives affects readmission and mortality rates.
 Under the assumption that hospitals would have continued to behave similarly in the absence of the programs, and that leadership teams do not change endogenously with the programs, I identify the differential response to pay-for-performance incentives among hospital types.
 That is, I define a group of treatment and comparison hospitals, and estimate how outcomes differ between the two groups after the policy changes.
-I ultimately employ synthetic difference-in-differences to analyze a group of comparison hospitals with similar characteristics to treatment hospitals.\footnote{I also present results for the full sample, and find qualitatively similar results.}
+I ultimately employ synthetic difference-in-differences to analyze a group of comparison hospitals with similar characteristics to treatment hospitals.\footnote{I also present results for the full sample and find qualitatively similar results.}
 
-   While all hospital types decrease readmission rates after the programs are enacted, NFPs with clinically trained executives respond less than both FPs and NFPs without clinically trained executives.
-That is, NFPs without clinically trained executives respond to incentives more similarly to FPs than other NFPs.
+   While quality generally improves after pay-for-performance incentives for all hospital types, the most aggressive response is from for-profits and NFPs without clinical executives, who decrease readmission and mortality rates more so than NFPs with clinical executives.
+That is, the objectives of NFPs without clinically trained executives seem to align more with those of for-profits: placing more weight on profit than societal outcomes.
 Leveraging changes in executive teams, I decompose this effect into signalling (clinical executives signal underlying hospital preferences), or managing (clinical executives manage the hospital differently) effects.
@@
 First, it adds to our understanding of NFP hospital objective functions.
-There is mixed evidence on whether NFP behavior differs from FP behavior: costs, uncompensated care, technology adoption, and quality (\cite{sloan2000not}; \cite{eggleston2008hospital}; \cite{moscelli2018effect}; \cite{moscone2020public}).
+There is mixed evidence on whether NFP behavior differs from FP behavior in terms of costs, uncompensated care, technology adoption, and quality (\cite{sloan2000not}; \cite{eggleston2008hospital}; \cite{moscelli2018effect}; \cite{moscone2020public}).
 To this point, researchers have found that NFPs do not act as purely profit maximizers, but maximize some combination of profit and output (\cite{deneffe2002not}, \cite{chang2011nonprofit}).
-A paper with an approach similar to mine which investigates hospital behavior is \citeauthor{chang2011nonprofit} (\citeyear{chang2011nonprofit}), which looks at the effect of a cost shock on likelihood of shutting down and mix of profitable services.
-They find that NFPs and FPs are equally as likely to shut down after the shock, but only NFPs adjust their mix of profitable services.
-Similarly, I investigate a change in incentives to hospitals, but I look at the differential impacts of leadership teams within not-for-profits.
 This paper adds to our understanding of inputs into not-for-profit hospital objective functions.
@@
     Not-for-profit firms in the US differ from for-profits mainly in that for-profit firms distribute revenue to stakeholders, while not-for-profits reinvest profit back into the firm.
-An important distinction between the two is that NFPs are, in general, driven by goals to better society, and therefore receive tax benefits from the government.
+Additionally, NFP firms are generally expected to set goals to better society due to government tax breaks received.
 
@@
 It is not uncommon for hospital executives to specialize in health care settings by earning a degree in health care management, or an MBA specific to health care.
-There are executive positions that are often filled by someone with clinical training, such as a Chief Medical Officer (CMO), and doctors can even find themselves in other top managerial positions as well, such as CEO or president.
+There are executive positions that are often filled by someone with clinical training, such as a Chief Medical Officer (CMO), and medical doctors pursue other top managerial positions as well, such as CEO or president.
 While some doctors earn additional degrees before stepping into an executive role, this is not a necessary condition to becoming a physician executive.
@@
 Alternatively, physicians may not always be adequately trained to enter into high level management positions, and therefore worsen hospital outcomes (\cite{HarvardBusinessReview2018}).
-
-    All tax-exempt organizations in the US are required to file a Form 990 with Internal Revenue Services (IRS) each year.
-There are different types of forms, but any organization grossing over \$200,000 must file the most extensive Form 990.
-Sections of this form include a statement of revenue, statement of functional expenses, a balance sheet, and, as used in this project, a list of all key employees, executives, and board members.
-Each firm is required to report the name and title, average hours per week, position, and compensation of their board members and executives.
 
@@
     In October 2011, the Center for Medicare and Medicaid Services (CMS) released a set of rules under HRRP mandating penalties for hospitals with above average readmission rates.
-A readmission is when a patient returns to the hospital within 30 days of being discharged from a previous stay; avoidable readmissions are a bad outcome for patients and increase health care spending.
 The goal of HRRP is to lower readmissions through better care coordination, less initial stay complications, and better post-care instructions.
@@
 Therefore, hospitals had incentive to react immediately once details of the program were announced in October of 2011.
-These penalties are not insignificant; penalized hospitals paid, on average, 4-5\% of revenue.
 
@@
 To my knowledge, this is the first large scale gathering of executive names from these forms.\footnote{\citeauthor{brickley2010board} (\citeyear{brickley2010board}) collects compensation data for a select number of hospitals.}
+
+    All tax-exempt organizations in the US are required to file a Form 990 with Internal Revenue Services (IRS) each year.
+There are different types of forms, but any organization grossing over \$200,000 must file the most extensive Form 990.
+Sections of this form include a statement of revenue, statement of functional expenses, a balance sheet, and, as used in this project, a list of all key employees, executives, and board members.
+Each firm is required to report the name and title, average hours per week, position, and compensation of their board members and executives.
 
@@
 Therefore, for observations that are missing after the initial extraction, I manually record names, positions, and titles.
-This yields 852 not-for-profit hospitals that are matched to AHA data, and are not missing text information.
+As I limit the sample to hospitals who have text information in every year, I limit the years to 2010-2014 to retain more hospitals.
+This yields 852 not-for-profit hospitals that are matched to AHA data, and with complete text information in each year.
 
@@
 The main reason for this is because the organization of teams varies so drastically across hospitals that a specific position, even if it exists in two different hospitals, can be meaningfully different.
-Additionally, the more granular the information I extract, the more likely there will be a mistake in the OCR extraction.
-Thus, to preserve correctness of the independent variable of interest, I focus on whether there is any MD executive at the hospital.
-
 I drop anyone who is solely a board member, and focus on those who have executive in their title, or are labeled as a president or vice president of a department.
@@
 
-    I combine the created sample of not-for-profits with all for-profits in the AHA survey.
+    I combine the resulting sample of not-for-profits with all for-profits in the AHA survey.
 The outcomes I consider are readmission and mortality rates, which are directly targeted in the pay-for-performance policies, and are found in the CMS Hospital Compare data.
 I combine rates for pneumonia, AMI, and heart failure (the relevant HRRP conditions) into a weighted average based on the number of patients in each condition.
-Additionally, I include several relevant hospital characteristics: number of beds, whether the hospital is an academic medical center, is physician owned, or is system affiliated, all come from the AHA survey.
-Case mix index, a measure of overall patient complexity, comes from CMS.
+Additionally, I include several relevant hospital characteristics from the AHA survey: number of beds, and whether the hospital is an academic medical center, is physician owned, or is system affiliated.
+Case mix index, a measure of overall patient complexity, comes from the CMS Case Mix Index files.
 Whether the hospital was penalized under HRRP or received payments under HVBP is found in the HCRIS data.
-Finally, I limit to years 2010-2014.
 
@@
     %\import{Tables}{overall_sumstats.tex}
+
+    In Table \ref{tab:sumstats_samples}, I present a table of means of relevant variables for several hospital samples used throughout the analysis.
+First, there are 708 for-profit hospitals in the sample, shown in the left-most column.
+Notably,
 
@@
 
-    \section{Model and Empirical Strategy}\label{sec:model}
+    \section{Modelling Hospital Response to Pay-for-Performance}\label{sec:model}
 
@@
 
-    \subsection{Estimation}
+    \section{Estimating Differences in Hospital Response}
+
+    \subsection{Method}
 
@@
 
-    Additionally, under certain assumptions that I discuss in Section \ref{sec:identification}, using executive team changes can be advantageous in better understanding underlying hospital motives.
-There are two mechanisms to having a clinically trained executives affecting outcomes: first, a clinically trained executive could be a signal of the underlying weight placed on profit, $\alpha$.
-Second, hiring a clinically trained executive who carries out day-to-day operations in a distinct way effectively changes $\alpha$.
-Thus, I also define treat and limit the comparison group in such a way that decomposes the two.
-Without limiting to stable teams, I consider hospitals that ever hire a clinically trained executive vs.
-hospitals that have a clinically trained executive in 2012 when the programs took place.
-These variations of Equation \ref{eq:forprofit} are shown in Figure \ref{fig:decomp_spec}.
-
-    \begin{figure}[ht!]
-    \begin{center}
-    \caption{\label{fig:decomp_spec}Decomposition Model Specification Details}
-
- \begin{tabular}{| m{18em} |}
- \hline
- Decomposition 1:\\ [0.5ex]
- \hline\hline 
- \vspace{2mm}
- Treat Group:  \hspace{15mm} NFP w/out MD \\
- \vspace{2mm}
- Comparison Group: \hspace{3mm} NFPs w/ MD  \\
- [1ex]
- \hline
- \end{tabular}
-\hfil   %<---
- \begin{tabular}{|m{18em}|}
- \hline
- Decomposition 2:\\ [0.5ex]
- \hline\hline
- \vspace{2mm}
- Treat Group: \hspace{11mm} No MD Exec 2012 \\
- \vspace{2mm}
- Comparison Group:  NFPs w/ MD (not 2012)  \\
- [1ex]
- \hline
- \end{tabular}
-
-\end{center}
- \end{figure}
-
-    \subsection{Identification}\label{sec:identification}
+ \subsection{Identification}
 
@@
 
-     Second, I assume that executive teams are not determined endogenously with the programs.
-In the main specification, I limit the sample to hospitals with executive teams that do not change their propensity to hire a clinical executive over time.
-Thus, I rely on this sample selection not being correlated with the program enactment.
-In supplementary analyses, I leverage executive team changes to disentangle signalling vs.
-managing effects, which is only identified if leadership changes are uncorrelated with the policies.
-In Appendix \ref{app:changes}, I present various descriptive statistics of executive team changes over time, showing that hiring and firing of MD executives do not change much over time, the rate of clinical executives to total executives remains stable over time, and the distribution of timing of clinical executive changes.
-
-     I assess the validity of the exogenous changes assumption in two ways.
-First, I analyze whether changes in clinically trained executives occur more often in time periods of program enactment.
-Second, I analyze whether hospitals who end up getting penalized or receiving payments are more likely to make changes in clinical executive members than those who are not.
-That is, are hospitals that expect to be affected by the programs more likely to hire or fire clinically trained executives?
-I consider changes in having any clinically trained executive, as well as changes in the number of clinically trained executives.
-I estimate the following linear equations:
-
-    \begin{equation}\label{eq:change1}
-    \text{change}_{ht} = \sum_{j=2011}^{2014}\beta_j\mathbf{1}\{t=j\} + \alpha_h + \epsilon_{ht},
-    \end{equation}
-
-    \begin{equation}\label{eq:change2}
-    \text{change}_{ht} = \beta(\text{Program Exposed}_{h} \times \text{Post Programs}_t)+ \alpha_h + \delta_t + \epsilon_{ht}.
-    \end{equation}
-
-    The variable $\text{Program Exposed}_{h}$ is an indicator for whether the hospital eventually was penalized under HRRP or received payments under HVBP.
-The estimates from this analysis are presented in Table \ref{tab:change_analysis}.
-Changes in any clinically trained executive are less likely to occur in 2012-2014 relative to 2010, indicating that executive teams are actually becoming more stable over time along the dimension of hiring clinically trained executives.
-Further, hospitals who end up being penalized through HRRP or receiving payments through HVBP are not more likely to change their propensity to hire any clinically trained executive.
-Results are similar when considering changes in the number of clinically trained executives.
-Thus, it seems unlikely that endogenous team formation is biasing the estimates.
-
-     \import{Tables}{change_analysis.tex}
-
-    Finally, I assume that no other unobserved changes occurred in 2012 that are correlated with both hospital performance and type.
-
-    \section{Results}
+ \subsection{Results}
 
@@
 Again, the difference is not statistically significant, but the coefficient compared to the true nulls in panels \ref{fig:mort_synth_plota} and \ref{fig:mort_synth_plotc} indicates NFPs without a clinical executive potentially decrease mortality more aggressively than NFPs with a clinical executive.
+In Appendix \ref{app:condition}, I show that these results are mainly driven by pneumonia patients.
 
-    \subsection{Decomposition}
+\section{Signalling vs. Managing Decomposition}
+
+\subsection{Method}
+
+Additionally, under certain assumptions that I discuss in Section \ref{sec:identification}, using executive team changes can be advantageous in better understanding underlying hospital motives.
+There are two mechanisms to having a clinically trained executives affecting outcomes: first, a clinically trained executive could be a signal of the underlying weight placed on profit, $\alpha$.
+Second, hiring a clinically trained executive who carries out day-to-day operations in a distinct way effectively changes $\alpha$.
+Thus, I also define treat and limit the comparison group in such a way that decomposes the two.
+Without limiting to stable teams, I consider hospitals that ever hire a clinically trained executive vs.
+hospitals that have a clinically trained executive in 2012 when the programs took place.
+These variations of Equation \ref{eq:forprofit} are shown in Figure \ref{fig:decomp_spec}.
+
+\begin{figure}[ht!]
+\begin{center}
+\caption{\label{fig:decomp_spec}Decomposition Model Specification Details}
+
+ \begin{tabular}{| m{18em} |}
+ \hline
+ Decomposition 1:\\ [0.5ex]
+ \hline\hline 
+ \vspace{2mm}
+ Treat Group:  \hspace{15mm} NFP w/out MD \\
+ \vspace{2mm}
+ Comparison Group: \hspace{3mm} NFPs w/ MD  \\
+ [1ex]
+ \hline
+ \end{tabular}
+\hfil   %<---
+ \begin{tabular}{|m{18em}|}
+ \hline
+ Decomposition 2:\\ [0.5ex]
+ \hline\hline
+ \vspace{2mm}
+ Treat Group: \hspace{11mm} No MD Exec 2012 \\
+ \vspace{2mm}
+ Comparison Group:  NFPs w/ MD (not 2012)  \\
+ [1ex]
+ \hline
+ \end{tabular}
+
+\end{center}
+ \end{figure}
+
+    \subsection{Identification}\label{sec:identification}
+
+    To identify the causal effect of hospital type on hospital program response, I rely on several assumptions.
+First, I assume that, absent the program enactments, treat and comparison group types would have behaved similarly in outcomes.
+While I see no evidence of pre-trends, Table \ref{tab:sumstats_samples} suggests that hospitals with a clinically trained executive are different than other hospitals along multiple characteristics.
+There may be concern that these differences between hospital types could be driving the results.
+Therefore, my main specification uses the synthetic difference-in-differences method established in \citeauthor{arkhangelsky2021synthetic} (\citeyear{arkhangelsky2021synthetic}).
+This method creates unit weights that more heavily factor comparison hospitals that are similar to the treat group hospitals prior to the programs.
+It also weighs time periods that balance pre-programs and post-programs for the comparison hospitals.
+Then, a simple two way fixed effects estimator is used with these weights, yielding a more robust average treatment effect that relies less on a strong parallel trends assumption (\cite{arkhangelsky2021synthetic}).
+I present the synthetic diff-in-diff estimates as the main results, and they are largely similar to the standard two way fixed effects results with no weights, shown in Appendix \ref{app:fullsample}.
+Further, the conceptual identification is the same as a standard difference-in-differences.
+
+     Second, I assume that executive teams are not determined endogenously with the programs.
+In the main specification, I limit the sample to hospitals with executive teams that do not change their propensity to hire a clinical executive over time.
+Thus, I rely on this sample selection not being correlated with the program enactment.
+In supplementary analyses, I leverage executive team changes to disentangle signalling vs.
+managing effects, which is only identified if leadership changes are uncorrelated with the policies.
+In Appendix \ref{app:changes}, I present various descriptive statistics of executive team changes over time, showing that hiring and firing of MD executives do not change much over time, the rate of clinical executives to total executives remains stable over time, and the distribution of timing of clinical executive changes.
+
+     I assess the validity of the exogenous changes assumption in two ways.
+First, I analyze whether changes in clinically trained executives occur more often in time periods of program enactment.
+Second, I analyze whether hospitals who end up getting penalized or receiving payments are more likely to make changes in clinical executive members than those who are not.
+That is, are hospitals that expect to be affected by the programs more likely to hire or fire clinically trained executives?
+I consider changes in having any clinically trained executive, as well as changes in the number of clinically trained executives.
+I estimate the following linear equations:
+
+    \begin{equation}\label{eq:change1}
+    \text{change}_{ht} = \sum_{j=2011}^{2014}\beta_j\mathbf{1}\{t=j\} + \alpha_h + \epsilon_{ht},
+    \end{equation}
+
+    \begin{equation}\label{eq:change2}
+    \text{change}_{ht} = \beta(\text{Program Exposed}_{h} \times \text{Post Programs}_t)+ \alpha_h + \delta_t + \epsilon_{ht}.
+    \end{equation}
+
+    The variable $\text{Program Exposed}_{h}$ is an indicator for whether the hospital eventually was penalized under HRRP or received payments under HVBP.
+The estimates from this analysis are presented in Table \ref{tab:change_analysis}.
+Changes in any clinically trained executive are less likely to occur in 2012-2014 relative to 2010, indicating that executive teams are actually becoming more stable over time along the dimension of hiring clinically trained executives.
+Further, hospitals who end up being penalized through HRRP or receiving payments through HVBP are not more likely to change their propensity to hire any clinically trained executive.
+Results are similar when considering changes in the number of clinically trained executives.
+Thus, it seems unlikely that endogenous team formation is biasing the estimates.
+
+     \import{Tables}{change_analysis.tex}
+
+    Finally, I assume that no other unobserved changes occurred in 2012 that are correlated with both hospital performance and type.
+
+    \subsection{Results}
 
@@
 
-\section{Results By Condition}
+\section{Results By Condition}\label{app:condition}
+
+The main results of the paper consider the weighted average readmission or mortality rate as the main outcomes.
+Here, I present results for each condition individually.
+Table \ref{tab:fp_read_condition_synth.tex}, I show the effect of being for-profit on AMI, heart failure, and pneumonia readmission rates compared to differing NFP comparison groups.
+The estimates indicate that the weighted average is mainly driven by pneumonia patients, since AMI readmission rate responses differ between FP and all NFP, where no other condition does.
 
@@
 
+Table \ref{tab:fp_mort_condition_synth.tex} shows a very similar result for mortality rates.
+AMI responses differ between FP and all NFP or NFP without clinical executives, but the weighted average is driven mainly by pneumonia patients.
+
 \import{Tables}{fp_mort_condition_synth.tex}
+
+The differences between NFPs with and without clinical executives are also driven mainly by pneumonia patients, shown in Table \ref{tab:MD_noMD_readmort_condition_synth}.
 

### 2024-07-23 Tables/MD_noMD_readmort_decomp_synth.tex
Updates from Overleaf
update objects

--- 2024-05-01 Tables/MD_noMD_readmort_decomp_synth.tex
@@
 \midrule
-Ever MD Exec x Post Programs & 0 &  & -0.06 & \\
- & (0.06) &  & (0.06) & \\
-2012 MD Exec x Post Programs &  & -0.19$^{*}$ &  & -0.1\\
- &  & (0.09) &  & (0.08)\\
+Ever MD Exec x Post Programs & -0.01 &  & 0.03 & \\
+ & (0.07) &  & (0.06) & \\
+2012 MD Exec x Post Programs &  & 0.17 &  & 0.12\\
+ &  & (0.1) &  & (0.09)\\
  &  &  &  & \\
@@
  &  &  &  & \\
-Observations & 3925 & 1845 & 3910 & 1840\\
+Observations & 3005 & 1595 & 3005 & 1595\\
 \bottomrule

### 2024-07-23 Paper/paper2.tex
Updates from Overleaf

--- 2024-06-10 Paper/paper.tex
@@
 		{\small
-        There is mixed evidence on whether not-for-profit hospitals differ meaningfully from for-profit (FP) hospitals, but what has not been considered is the effect leadership teams have on hospital behavior.
-In this paper, I investigate whether leadership team composition of NFP hospitals affects how similarly they behave to for-profits.
-I use a novel data set on NFP hospital executives in the US (2010-2014) to investigate whether the presence of clinically trained executives affects hospital behavior.
-In particular, I leverage pay-for-performance initiatives in the US enacted in 2012, and measure the  readmission and mortality rate responses between FP and NFP with and without clinically trained executives.
-I find that for-profits nad NFPs without clinical executives respond more aggressively to financial incentives on quality, consistent with theoretical predictions of firms that value profit relatively more than having high quality.
+         Firm leaders are shown to be correlated with firm decisions and performance.
+In this paper, I investigate this topic in the context of hospital executive teams.
+Specifically, I empirically the effect of clinically trained executives on hospital behavior in response to an incentive shock.
+In particular, I leverage policies changing quality incentives for hospitals in the US in 2012, and estimate the difference in readmission and mortality rate responses for nonprofit hospitals with and without clinically trained executives.
+I find that a lack of clinical executives leads hospitals to improve quality more drastically after the incentive change.
+Further, I provide evidence that the more drastic response is due to non-clinical executives being more profit driven, and that clinical executives are not just a signal of underlying hospital objectives, but manage the hospital differently.
 		} 
@@
 
-  The objectives and behaviors of not-for-profit firms are challenging to determine, and an extensive literature has sought to understand them.
-Researchers have posited a variety of objective functions that could motivate NFP firms such as maximizing profit (for-profit in disguise), prestige, income, or a quality/quantity trade-off (\cite{steinberg1986revealed}).
-Empirical contributions to this literature center around investigating behaviors that differ between for-profit (FP), private NFP, and government firms (\cite{sloan2000not}).
-However, the literature has neglected to consider other relevant characteristics of hospitals that could affect objectives, such as the hospital's leadership team.
-In this paper, I investigate whether differences in observed behavior between for-profit and not-for-profit hospitals vary differentially by leadership team composition.
+    Various characteristics of firm leadership such as gender, age, and background have been shown to be correlated with firm decisions and performance.
+I investigate this topic in the context of hospital executives with differing backgrounds: some with training as a medical doctor and some without.
+Hiring physician executives can be controversial.
+On one hand, physicians bring a unique combination of clinical and administration expertise, and they have insights that can potentially improve hospital outcomes (\cite{Stajduhar_2023}, \cite{Ahmed_2022}).
+Alternatively, physicians may not always be adequately trained to enter into high level management positions, and therefore worsen hospital outcomes (\cite{HarvardBusinessReview2018}).
+In this paper, I investigate how having clinically trained executives in leadership affects hospital response to pay-for-performance incentives.
 
-  Hospitals have been a common focus when seeking to understand NFP firms.
-This is because, first, there are different hospital ownership types within the industry, giving researchers the ability to compare the behaviors of NFPs and FPs.
-Second, private NFP hospitals make up 50\% of all hospitals in the US, and staff, on average, 207 beds.
-For comparison, FPs make up 36\% of hospitals and staff 107 beds (\cite{ASPE_2023}).
-Thus, NFP behavior in the hospital industry is relevant, and directly affects consumers of health care.
-However, it remains unclear what objective function drives NFP hospital behavior, and whether NFP hospitals are meaningfully different from FP hospitals (\cite{sloan2000not}; \cite{erus2002inferring}; \cite{deneffe2002not}; \cite{horwitz2009hospital}).
+    I scrape executive team names and titles from publicly available Tax Form 990s, filled out by nonprofit firms in the US.
+Combining this with publicly available hospital characteristics, I construct a novel data set of nonprofit hospitals from 2010-2014 containing information on both executive team characteristics and quality metrics.
+Nonprofit behavior in the hospital industry directly affects health care consumers, as private nonprofit hospitals make up 50\% of all hospitals in the US, and staff, on average, 207 beds.
+For comparison, for-profits make up 36\% of hospitals and staff 107 beds (\cite{ASPE_2023}).
 
-  While ownership is certainly relevant, there could be other characteristics correlated with hospital behavior independent of ownership type.
-Researchers have focused in particular on explaining variation in NFP behavior using variation in the market they exist in (\cite{duggan2002hospital}; \cite{horwitz2009hospital}).
-However, I argue that an observable and economically meaningful factor is the composition of top-level hospital managers.
-In large publicly traded firms, executive characteristics are shown to be correlated with firm performance and governance, indicating that there is more to firm behavior than ownership status (\cite{bertrand2003managing}; \cite{ahern2012changing}\cite{matsa2013female}).
+    The selection of leaders makes it difficult to interpret any direct comparison of differing executive teams as causal.
+Therefore, I leverage an exogenous shock to hospital incentives, pay-for-performance policies enacted in the US in 2012, and compare the change in readmission and mortality rates of executive teams with and without clinical executives.
+Under the assumption that the two types of teams would have had similar trends in readmission and mortality rates absent the incentive change, I find that a lack of clinically trained executives leads a hospital to respond more drastically to pay-for-performance incentives.
 
-  A policy relevant aspect of hospital behavior in the US has been the shift towards pay-for-performance incentives.
-While quality may or may not directly matter to hospitals, pay-for-performance incentives are such that quality of care directly impacts profit.
-For hospitals that already valued quality before pay-for-performance incentives, the change in incentives may not change quality much.
-However, if the hospital is more profit driven, the policy change should translate to a larger change in quality.
-Ownership status certainly plays a role in the valuations of profit vs.
-quality, but the question remains as to whether management also plays an important role.
-I contribute to our understanding of hospital and NFP behaviors by investigating whether occupational background of executives, specifically clinical training, affects hospital response to pay-for-performance incentives.
-
-  Using publicly available tax forms, I construct a novel data set of NFP hospital executives from 2009-2015.
-I link hospitals in this data to characteristics in the American Hospital Association (AHA) survey and Hospital Compare data.
-I leverage various pay-for-performance initiatives in 2012 to investigate differential hospital responses between for-profits, and NFPs with and without without clinically trained executives.
-
-  I estimate how the interaction between hospital type and pay-for-performance incentives affects readmission and mortality rates.
-Under the assumption that hospitals would have continued to behave similarly in the absence of the programs, and that leadership teams do not change endogenously with the programs, I identify the differential response to pay-for-performance incentives among hospital types.
-That is, I define a group of treatment and comparison hospitals, and estimate how outcomes differ between the two groups after the policy changes.
-I ultimately employ synthetic difference-in-differences to analyze a group of comparison hospitals with similar characteristics to treatment hospitals.\footnote{I also present results for the full sample and find qualitatively similar results.}
-
-   While quality generally improves after pay-for-performance incentives for all hospital types, the most aggressive response is from for-profits and NFPs without clinical executives, who decrease readmission and mortality rates more so than NFPs with clinical executives.
-That is, the objectives of NFPs without clinically trained executives seem to align more with those of for-profits: placing more weight on profit than societal outcomes.
-Leveraging changes in executive teams, I decompose this effect into signalling (clinical executives signal underlying hospital preferences), or managing (clinical executives manage the hospital differently) effects.
+    Having found that hospital leadership does, in fact, matter to hospital behavior, I investigate several mechanisms that might contribute to this difference.
+First, \textcolor{red}{the more drastic decrease in readmissions by nonclinical teams is not driven by decreasing average patient complexity.
+There is no difference in patient case mix between clinical and nonclinical executive teams}.
+Further, I comparing each type of leadership team with for-profit hospitals reveals that the absence of clinical executives leads a nonprofit hospital to respond more similarly to a for-profit hospital.
+That is, having clinical training on a leadership team leads a hospital to act less profit-driven than their general training counterparts.
+Finally, I show that changes in a hospital's propensity to hire a clinical executive are not correlated with the pay-for-performance policies.
+Thus, I exploit such changes in leadership teams to decompose the effect of clinical executives into signaling (clinical executives signal underlying hospital preferences), or managing (clinical executives manage the hospital differently).
 I find that the entire difference in response is driven by a managing effect.
 
-    This paper contributes to three strands of literature.
-First, it adds to our understanding of NFP hospital objective functions.
-There is mixed evidence on whether NFP behavior differs from FP behavior in terms of costs, uncompensated care, technology adoption, and quality (\cite{sloan2000not}; \cite{eggleston2008hospital}; \cite{moscelli2018effect}; \cite{moscone2020public}).
-To this point, researchers have found that NFPs do not act as purely profit maximizers, but maximize some combination of profit and output (\cite{deneffe2002not}, \cite{chang2011nonprofit}).
-This paper adds to our understanding of inputs into not-for-profit hospital objective functions.
-
-    Second, many researchers have documented a correlation between executive backgrounds, specifically for CEOs, and firm performance, starting with \citeauthor{bertrand2003managing} (\citeyear{bertrand2003managing}), who show that manager fixed effects are an important driver of many firm behaviors and decisions.
+    Many researchers have documented a correlation between executive characteristics, specifically for CEOs, and firm performance, starting with \citeauthor{bertrand2003managing} (\citeyear{bertrand2003managing}), who show that manager fixed effects are an important driver of many firm behaviors and decisions.
 Female board members are associated with better oversight and more women executives (\cite{matsa2011chipping}; \cite{adams2009women}), but are negatively correlated with firm value (\cite{ahern2012changing}).
@@
     Due to data limitations, many of these studies focus on large, publicly traded firms in the US.
-\citeauthor{brickley2010board} (\citeyear{brickley2010board}) has the only study to my knowledge answering this question in the US NFP context.
+\citeauthor{brickley2010board} (\citeyear{brickley2010board}) has the only study to my knowledge answering this question in the US nonprofit context.
 Using exogenous variation in expected Medicare profits, the authors find that having an internal board of directors increases CEO compensation, and having physician board members decreases public donations (\cite{brickley2010board}).
@@
 \citeauthor{janke2019impact} (\citeyear{janke2019impact}) uses data from England to study whether CEOs affect hospital production, and find no association (\cite{janke2019impact}).
-However, \citeauthor{otero2022managers} (\citeyear{otero2022managers}) investigates the role of CEOs in public hospitals and Chile, and finds an 8\% decrease in mortality rates for hospitals with top managers (\cite{otero2022managers}).
-I contribute to this literature by considering the executive team as a whole rather than just CEOs, by investigating a context which is not largely studied but is policy relevant, NFP US hospitals, and by leveraging a unique executive characteristic pertinent to that context.
+However, \citeauthor{otero2022managers} (\citeyear{otero2022managers}) investigates the role of CEOs in public hospitals in Chile, and finds an 8\% decrease in mortality rates for hospitals with top managers.
+Further, a major contributor of this quality improvement is the exodus of older physicians as CEOs (\cite{otero2022managers}).
+I contribute to our knowledge of how executive team characteristics affect firm behaviors by (1) considering the important setting of nonprofit hospitals, which serve the majority of patients in the US, (2) considering a unique executive characteristic which is plausibly relevant, and (3) observing firm behavior in the presence of an exogenous incentive change.
 
-    Finally, I contribute to our understanding of how providers respond to pay-for-performance incentives in health care.
-The most in depth study of how hospitals respond to HRRP is \citeauthor{gupta2021impacts} (\citeyear{gupta2021impacts}), who finds that hospitals decreased readmissions by 5\% and mortality rates by 2\% on average as a result of the program, confirming prior studies (\cite{mellor2017does}; \cite{ziedan2018essays}; \cite{ody2019decreases}; \cite{gupta2021impacts}).
+    Additionally, I contribute to our understanding of how providers respond to pay-for-performance incentives in health care.
+The most in depth study of how hospitals respond to the Hospital Readmissions Reduction Program, one of the two large pay-for-performance initiatives, is \citeauthor{gupta2021impacts} (\citeyear{gupta2021impacts}).
+This study finds that hospitals decreased readmissions by 5\% and mortality rates by 2\% on average as a result of the program, confirming prior studies (\cite{mellor2017does}; \cite{ziedan2018essays}; \cite{ody2019decreases}; \cite{gupta2021impacts}).
 Around 40\% of the decrease in readmissions is due to selective patient practices.
-Research on HVBP generally finds that the program had no effect on underlying hospital quality (\cite{us2015hospital}; \cite{norton2018moneyball}; \cite{friedson2019so}).
+Further, market concentration and hospital systems play a large role in how hospitals respond to the program (\cite{kunz2024assessing}).
+Research on the other large pay-for-performance initiative, the Hospital Value Based Purchasing Program, generally finds that the program had no effect on underlying hospital quality (\cite{us2015hospital}; \cite{norton2018moneyball}; \cite{friedson2019so}).
 This paper contributes to our understanding of how characteristics of hospitals affect response to pay-for-performance incentives.
 
-    \section{Setting}
+    \section{Setting and Data}
 
-    \subsection{Not-for-Profit Hospital Executives}
+    \subsection{Nonprofit Hospital Executives}
 
-    Not-for-profit firms in the US differ from for-profits mainly in that for-profit firms distribute revenue to stakeholders, while not-for-profits reinvest profit back into the firm.
-Additionally, NFP firms are generally expected to set goals to better society due to government tax breaks received.
-
+    Nonprofit firms in the US differ from for-profits mainly in that for-profit firms distribute revenue to stakeholders, while nonprofits reinvest profit back into the firm.
 Hospitals are governed by a board of directors, whose role is to set broad goals and strategies, and provide general oversight.
@@
 An executive team usually consists of at least a Chief Executive Officer (CEO) and Chief Financial Officer (CFO), but there is variation in how firms organize executive teams.
-It is not uncommon for hospital executives to specialize in health care settings by earning a degree in health care management, or an MBA specific to health care.
-There are executive positions that are often filled by someone with clinical training, such as a Chief Medical Officer (CMO), and medical doctors pursue other top managerial positions as well, such as CEO or president.
+Some hospital executives to specialize in health care settings by earning a degree in health care management, or an MBA specific to health care.
+There are executive positions that are often filled by someone with clinical training, such as a Chief Medical Officer (CMO), but medical doctors can pursue other top managerial positions as well such as CEO or president.
 While some doctors earn additional degrees before stepping into an executive role, this is not a necessary condition to becoming a physician executive.
 
-    Hiring physician executives can be controversial.
-On one hand, physicians bring a unique combination of clinical and administration expertise, and they have insights that can potentially improve hospital outcomes (\cite{Stajduhar_2023}, \cite{Ahmed_2022}).
-Alternatively, physicians may not always be adequately trained to enter into high level management positions, and therefore worsen hospital outcomes (\cite{HarvardBusinessReview2018}).
+    Hospital executive teams are understudied partly due to the inaccessibility of granular data.
+I construct a novel data set of not-for-profit hospitals in the US that contains names, titles, and positions of all board members and executives tied to the hospital from 2009-2015.
+I gather this data from publicly available Tax Form 990s, which every sufficiently large nonprofit files each year in the US.
+To my knowledge, this is the first large scale gathering of executive names from these forms.\footnote{\citeauthor{brickley2010board} (\citeyear{brickley2010board}) collects compensation data for a select number of hospitals.}
 
-    \subsection{Pay-for-Performance Policies}\label{sec:hrrp}
+    All tax-exempt organizations in the US are required to file a Form 990 with Internal Revenue Services (IRS) each year.
+There are different types of forms, but any organization grossing over \$200,000 must file the most extensive Form 990.
+Sections of this form include a statement of revenue, statement of functional expenses, a balance sheet, and, as used in this project, a list of all key employees, executives, and board members.
+Each firm is required to report the name and title, average hours per week, position, and compensation of their board members and executives.
+
+    The historical tax forms for all not-for-profits are housed by ProPublica.\footnote{https://projects.propublica.org/not-for-profits/} I use the NonProfit Explorer API to extract Employee Identification Numbers (EIN) of NFPs with a hospital National Taxonomy of Exempt Entities (NTEE) code.
+After filtering out associations and specialty hospitals, there are approximately 3,000 NFP EINs.
+Linking these EINs to other hospital information is crucial for the analysis.
+I therefore rely on matching by location and name of hospital in both the tax forms and AHA data.
+In the AHA data, I limit to general hospitals.
+Then, I confidently match 1200 EINs to an AHA identification number based on exact name matches within the same state.
+I assess the observable differences between matched and non-matched hospitals in Appendix \ref{app:matched}, and conclude that the samples are overall similar apart from an under-sampling of hospitals that belong to systems.
+
+    For each of the matched EINs, I extract web URLs to Form 990 PDFs for each year.
+I use an algorithm to download these locally and use optical character recognition (OCR) text extraction methods to scrape the relevant section of the PDF, which is titled ``Officers, Directors, Trustees, Key Employees, and Highest Compensated Employees".
+Finally, I use string cleaning methods to identify all names and positions under each EIN, year.
+With each name, I extract titles indicating clinical training: MD, Dr., or DO.
+Using OCR extraction can be unreliable in cases where the text on the form is handwritten or small.
+Therefore, for observations that are missing after the initial extraction, I manually record names, positions, and titles.
+As I limit the sample to hospitals who have text information in every year, I limit the years to 2010-2014 to retain more hospitals.
+This yields 852 not-for-profit hospitals that are matched to AHA data, and with complete text information in each year.
+
+    Finally, I match each name with the National Plan and Provider Enumeration System (NPPES) database of all registered physicians.
+The goal is to match each physician with an NPI number in order to merge publicly available data on physician specialty, gender, and tenure.
+For any person who has a doctor title and uniquely matches the NPPES data, I record their NPI number.
+For anyone who has a doctor title and no matches, no doctor title and matches, or a doctor title and multiple matches, I use information on location and hospital name to manually match the NPI.
+
+    For the purposes of this paper, I focus on executive teams as a whole instead of specific positions.
+The main reason for this is because the organization of teams varies so drastically across hospitals that a specific position, even if it exists in two different hospitals, can be meaningfully different.
+I drop anyone who is solely a board member, and focus on those who have executive in their title, or are labeled as a president or vice president of a department.
+I create various hospital-level characterizations based on their executive team, such as the existence of a clinically trained executive, the number of clinically trained executives, the number of total executives, and whether the hospital has a Chief Medical Officer.
+For my main sample, I limit to hospitals who do not change their propensity to hire a clinical executive for the duration of the sample.
+Summary statistics for these variables are shown in Table \ref{leader_sumstats}.
+On average, NFP hospitals in my sample have 4 executives and .4 clinically trained executives.
+Sixty-one percent of the sample never changes propensity to hire a clinically trained executive, and 33\% of NFPs in the sample have a Chief Medical Officer at some point.
+
+     \import{Tables}{leader_sumstats.tex}
+
+    \subsection{Pay-for-Performance Policies Targeting Readmission and Mortality Rates}\label{sec:hrrp}
 
@@
 
-	\section{Data}\label{sec:data}
-
-    Hospital executive teams are understudied partly due to the inaccessibility of granular data.
-I construct a novel data set of not-for-profit hospitals in the US that contains names, titles, and positions of all board members and executives tied to the hospital from 2009-2015.
-I gather this data from publicly available Tax Form 990s, which every sufficiently large not-for-profit files each year in the US.
-To my knowledge, this is the first large scale gathering of executive names from these forms.\footnote{\citeauthor{brickley2010board} (\citeyear{brickley2010board}) collects compensation data for a select number of hospitals.}
-
-    All tax-exempt organizations in the US are required to file a Form 990 with Internal Revenue Services (IRS) each year.
-There are different types of forms, but any organization grossing over \$200,000 must file the most extensive Form 990.
-Sections of this form include a statement of revenue, statement of functional expenses, a balance sheet, and, as used in this project, a list of all key employees, executives, and board members.
-Each firm is required to report the name and title, average hours per week, position, and compensation of their board members and executives.
-
-    The historical tax forms for all not-for-profits are housed by ProPublica.\footnote{https://projects.propublica.org/not-for-profits/} I use the NonProfit Explorer API to extract Employee Identification Numbers (EIN) of NFPs with a hospital National Taxonomy of Exempt Entities (NTEE) code.
-After filtering out associations and specialty hospitals, there are approximately 3,000 NFP EINs.
-Linking these EINs to other hospital information is crucial for the analysis.
-I therefore rely on matching by location and name of hospital in both the tax forms and AHA data.
-I confidently match 1200 EINs to an AHA identification number based on exact name matches within the same state.
-I assess the observable differences between matched and non-matched hospitals in Appendix \ref{app:matched}, and conclude that the samples are overall similar apart from an under-sampling of hospitals that belong to systems.
-
-    For each of the matched EINs, I extract web URLs to Form 990 PDFs for each year.
-I use an algorithm to download these locally and use optical character recognition (OCR) text extraction methods to scrape the relevant section of the PDF, which is titled ``Officers, Directors, Trustees, Key Employees, and Highest Compensated Employees".
-Finally, I use string cleaning methods to identify all names and positions under each EIN, year.
-With each name, I extract titles indicating clinical training: MD, Dr., or DO.
-Using OCR extraction can be unreliable in cases where the text on the form is handwritten or small.
-Therefore, for observations that are missing after the initial extraction, I manually record names, positions, and titles.
-As I limit the sample to hospitals who have text information in every year, I limit the years to 2010-2014 to retain more hospitals.
-This yields 852 not-for-profit hospitals that are matched to AHA data, and with complete text information in each year.
-
-    For the purposes of this paper, I focus mainly on executive teams as a whole instead of specific positions.
-The main reason for this is because the organization of teams varies so drastically across hospitals that a specific position, even if it exists in two different hospitals, can be meaningfully different.
-I drop anyone who is solely a board member, and focus on those who have executive in their title, or are labeled as a president or vice president of a department.
-I create various hospital-level characterizations based on their executive team, such as the existence of a clinically trained executive, the number of clinically trained executives, the number of total executives, and whether the hospital has a Chief Medical Officer.
-Additionally, I examine changes in propensity to hire a clinical executive.
-Summary statistics for these variables are shown in Table \ref{leader_sumstats}.
-On average, NFP hospitals in my sample have 4 executives and .4 clinically trained executives.
-Sixty-one percent of the sample never changes propensity to hire a clinically trained executive, and 33\% of NFPs in the sample have a Chief Medical Officer at some point.
-
-    \import{Tables}{leader_sumstats.tex}
-
-    I combine the resulting sample of not-for-profits with all for-profits in the AHA survey.
-The outcomes I consider are readmission and mortality rates, which are directly targeted in the pay-for-performance policies, and are found in the CMS Hospital Compare data.
+    Therefore, I focus on outcomes of readmission and mortality rates, which are publicly available at the hospital-condition level in CMS Hospital Compare.
 I combine rates for pneumonia, AMI, and heart failure (the relevant HRRP conditions) into a weighted average based on the number of patients in each condition.
 Additionally, I include several relevant hospital characteristics from the AHA survey: number of beds, and whether the hospital is an academic medical center, is physician owned, or is system affiliated.
-Case mix index, a measure of overall patient complexity, comes from the CMS Case Mix Index files.
+Additionally, case mix index, a measure of overall patient complexity, comes from the CMS Case Mix Index files.
 Whether the hospital was penalized under HRRP or received payments under HVBP is found in the HCRIS data.
 
-    % In Table \ref{sumstats}, I present summary statistics for hospital variables.
-There are 3,766 total hospitals in the sample, with 150 beds on average.
-Hospitals in the sample are less likely to be penalized for AMI than either heart failure or pneumonia under the HRRP program.
-On average, 20\% of patients with one of these conditions would be readmitted.
-While readmission rates for the specific conditions are similar, the highest likelihood of readmission is for heart failure patients.
-Alternatively, AMI is the most dangerous condition in terms of mortality rates.
-
-    %\import{Tables}{overall_sumstats.tex}
+	\subsection{Summary Statistics}\label{sec:data}
 
     In Table \ref{tab:sumstats_samples}, I present a table of means of relevant variables for several hospital samples used throughout the analysis.
-First, there are 708 for-profit hospitals in the sample, shown in the left-most column.
-Notably,
-
-    There are several samples of hospitals that are relevant for the analysis.
-First, there are all for-profits and all not-for-profits.
-Further, I subset the sample of NFPs into whether they have a clinical executive at some point in the sample.
-Finally, I characterize a NFP leadership team as being stable or not.
-A stable team means the hospital did not change their propensity to hire a clinical executive over time.
-I present a table of means for each of these sub-samples in Table \ref{tab:sumstats_samples}.
-Approximately 850 NFPs have leadership information, with 384 of them hiring an MD executive at some point in the sample.
-I compare these NFPs with the 708 for-profits from the AHA Survey.
-For-profits and NFPs are roughly the same bed size, but NFPs see more patients in each of the three relevant conditions.
-For-profits are more likely to be penalized, but also more likely to receive HVBP incentives (potentially for improvement in quality).
-For-profits have higher readmission rates, and higher mortality rates in all categories except heart failure.
-There are also key differences in not-for-profits with and without clinically trained executives.
-Those with MDs are larger, more likely to be penalized, have higher uncompensated care, and more complex patients.
-This holds for stable leadership teams as well, though the number of NFPs with a clinical executive drops significantly.
-I discuss the implications of these differences for identification in Section \ref{sec:model}.
+First, there are 708 for-profit hospitals in the sample, shown in column (1).
+For-profit hospitals in the sample have 155 beds on average, 112-276 patients depending on the condition, and are highly likely to be affiliated with a system.
+Thirty six percent of for-profits were penalized for all three conditions at some point.
 
     \import{Tables}{sample_sumstats.tex}
+
+    Columns 2-4 show variable means for not-for-profits in the sample, with all NFPs in column 2, NFPs that ever have a clinically trained executive in column 3, and NFPs that never have a clinically trained executive in column 4.
+NFPs are much less likely to be physician owned or system affiliated compared to FPs.
+They do, however, have similar bed size and case mix index to FPs.
+Twenty-two percent of all NFPs are penalized for all conditions.
+There are several differences among NFPs with and without clinically trained executives.
+Those with a clinically trained executive are larger in terms of beds and patients seen, are more likely to be an academic medical center, and see slightly more complex patients.
+Further, they are more likely to be penalized for every condition.
+
+    Finally, columns 5-6 limit the sample of not-for-profits further using changes in executive teams.
+The NFPs included in this sample do not change their propensity to hire a clinically trained executive, they either always have the same number of clinically trained executives or always have no clinically trained executives.
+The differences in NFPs with and without clinically trained executives remains similar with this sample limitation.
+I discuss the implications of these differences for identification in Section \ref{sec:model}.
 
@@
 While the readmission rate trend behaves similarly across hospital types prior to the programs, there are notable trend differences after the programs.
-NFPs with clinically trained executives show a more mitigated decrease in readmissions compared to the other types.
+For-profits decrease readmissions drastically.
+NFP trends depend on the presence of an executive MD, where those without a clinically trained executive respond drastically, more similarly to for-profits, than those with a clinically trained executive.
 In mortality rates, NFPs with a clinically trained executive have the lowest rates before the programs, but differentially continue with an increasing trend in mortality after the programs.
@@
 
-    \section{Modelling Hospital Response to Pay-for-Performance}\label{sec:model}
+    \section{Effect of Executive Clinical Training on Hospital Response}\label{sec:clinical}
 
-    I model hospital behavior under two simplified worlds: one where quality does not directly affect profit, and one where it does.
-Pay-for-performance incentives essentially incorporate performance directly into a hospital's profit function.
-I directly model performance as quality of care, which is inversely related to readmission and mortality rates.
-While the literature speculates about the true form of NFP objective functions, we can generally think of hospitals choosing some linear combination of profit and societal benefit, without knowing how much weight the hospital places on each.
-Intuitively, a hospital that places only weight on profit (for-profit type) will only care about quality when it enters into the profit function, and therefore will respond more drastically to pay-for-performance than a hospital who already cared about quality through societal benefit before the incentives.
-Therefore, hospital responses to pay-for-performance incentives reveal information about their underlying preference on profit vs.
-societal outcomes.
-
-    To formalize this intuition, I specify the hospital objective function as a weighted average of profit and societal (or patient) outcomes and derive characteristics of the optimal quality.
-For simplicity, I abstract away from quality having a direct effect of quantity and price.
-Without any financial incentives to increase quality, hospital quality is decided by:
-
-    $$\max_{\theta}\hspace{2mm}\alpha\left[R - c_{\theta}(\theta) \right] + (1-\alpha) u(\theta),$$
-
-    \noindent where $R$ is net revenue, $c_{\theta}(.)$ is an increasing cost function specific to quality $\theta$, $u(.)$ is extra utility gained from quality, which is increasing and concave in $\theta$, and $\alpha\in[0,1]$ captures the weight placed on profit vs.
-societal benefit.
-There is an implicit stay open condition throughout the model.
-Taking the first order condition yields
-
-    $$(1-\alpha)u'(\theta) = \alpha c_{\theta}'(\theta),$$
-
-    \noindent where marginal benefit equals marginal cost of increasing quality. Solving for $u'(\theta)$ and differentiating with respect to $\alpha$:
-
-    $$\frac{du'(\theta)}{d\alpha} = \frac{1}{(1-\alpha)^2}c_{\theta}'(\theta) > 0.$$
-
-    \noindent Thus, by the Implicit Function Theorem,
-
-    $$\frac{d\theta}{d\alpha} = \frac{du'(\theta)/d\alpha}{du'(\theta)/d\theta} < 0.$$
-
-    \noindent That is, the more weight on profit, the lower quality of care chosen by the hospital.
-
-    In a world with pay-for-performance incentives, which can either look like benefits from high quality (such as HVBP) or penalties for low quality (as in HRRP), quality directly affects revenue.
-Thus, $R(\theta)$ is an increasing function of $\theta$, and I assume that the marginal financial benefit of increasing quality is greater than the marginal cost ($R'(\theta)\geq c_{\theta}'(\theta))$.
-Thus, the new first order condition yields:
-
-    $$u'(\theta) = \frac{\alpha}{1-\alpha} \left[c_{\theta}'(\theta)-R(\theta)\right] \leq 0.$$
-
-    \noindent Taking the derivative of this with respect to $\alpha$,
-
-    $$\frac{du'(\theta)}{d\alpha} = \frac{1}{(1-\alpha)^2}c_{\theta}'(\theta).$$
-
-    Therefore, again using the Implicit Function Theorem,
-
-    $$\frac{d\theta}{d\alpha} = \frac{du'(\theta)/d\alpha}{du'(\theta)/d\theta}\geq0.$$
-
-    That is, all else equal, in a world with financial incentives on quality, quality is increasing with more weight placed on profit.
-I combine the results found in each scenario into one response function that depends on $\alpha$, where $\theta_2$ is quality under pay-for-performance incentives and $\theta_1$ is quality with no financial incentive to quality.
-That is,
-
-    \begin{align*}
-        \frac{d\Delta\theta}{d\alpha}&=\frac{d(\theta_2-\theta_1)}{d\alpha}\\
-        &=\frac{d\theta_2}{d\alpha}-\frac{d\theta_1}{d\alpha}\\
-        &\geq 0.
-    \end{align*}
-
-    Hence, this model predicts that change in quality depends on how much weight the hospital places on profit vs.
-societal benefit.
-Particularly, hospitals with more weight on profit respond more than hospitals with more weight placed on societal benefit.
-In reality, $\alpha$ is unknown, except in the case of for-profits, who have revealed that they have a high $\alpha$ by their ownership status.
-However, not-for-profit behavior on average does not bring much clarity about their objectives.
-An observable and economically relevant characteristic of a hospital that could affect $\alpha$ is composition of their executive leadership team.
-Thus, I compare the response of for-profits and not-for-profits with different leadership teams to bring clarity on how this might point to underlying $\alpha$.
-
-    \section{Estimating Differences in Hospital Response}
-
-    \subsection{Method}
-
-    In light of the theoretical predictions, I test empirically whether different hospital types respond to pay-for-performance incentives differently.
-I present here a two-way fixed effects specification, where the key variable is an interaction between a certain hospital type and time periods after program enactments.
-Ultimately, I employ synthetic difference-in-differences, which is estimating the same model with weights that more heavily factor similar trending control groups and time periods.
+    My first goal is to estimate the effect of having clinical executive training on how hospitals change readmission and mortality rates in response to pay-for-performance incentives.
+Ultimately, I employ synthetic difference-in-differences, weighting more heavily non-clinical hospitals with readmission, mortality, and baseline case mix index that better match the clinical team trends.
 However, for clarity, I present the unweighted specification:
@@
     \begin{equation}
-    \label{eq:forprofit}
-    y_{ht} = \beta \text{ treat}_{ht} + \gamma_{h} + \delta_t + \epsilon_{ht},
+    \label{eq:clinical}
+    y_{ht} = \beta \text{ (clinical training x post 2012)}_{ht} + \gamma_{h} + \delta_t + \epsilon_{ht},
     \end{equation}
 
-    \noindent where $y_{ht}$ is one of the outcome variables discussed in Section \ref{sec:data}, and $\gamma_h$ and $\delta_t$ are hospital and time fixed effects, respectively.
-The independent variable of interest, treat$_{ht}$, captures the effect of being a certain hospital type after incentives changed in 2012.
-That is, I interact an indicator for hospital type with an indicator for year post 2012 for various hospital types.
-First, I define treat so that for-profit hospitals are being compared to subsets of not-for-profits (by limiting the sample to the relevant treatment and comparison group).
-Then, I limit to only not-for-profits and define treat as those without a clinical executive.
-Specifically, I define treat$_{ht}$ and limit the sample in four ways, shown in Figure \ref{fig:spec}.
+    \noindent where $y_{ht}$ is one of the outcome variables discussed in Section \ref{sec:hrrp}, and $\gamma_h$ and $\delta_t$ are hospital and time fixed effects, respectively.
 
-    An important distinction is whether the classification of NFPs by leadership team is time-varying.
-For overall ownership status, switching is rare, and I limit to those who never switch.
-However, hospital executive teams are dynamic, and whether or not a hospital employs a clinically trained executive in each period may change.
-Therefore, I first present effects of having a clinically trained executive among stable executive teams, i.e., NFPs with no change in their propensity to hire a clinically trained executive.
-
-\begin{figure}[ht!]
-    \begin{center}
-    \caption{\label{fig:spec}Treated and Comparison Groups}
- \begin{tabular}{| m{18em} |}
- \hline
- Combination 1:\\ [0.5ex]
- \hline\hline 
- \vspace{2mm}
- Treat Group: \hspace{23.5mm} For-Profit \\
- \vspace{2mm} 
- Comparison Group: \hspace{11.5mm} Stable NFPs  \\
- [1ex]
- \hline
- \end{tabular}
-\hfil   %<---
- \begin{tabular}{|m{18em}|}
- \hline
- Combination 2:\\ [0.5ex]
- \hline\hline
- \vspace{2mm}
- Treat Group: \hspace{13.5mm} For-Profit \\
- \vspace{2mm}
- Comparison Group: \hspace{1.5mm} Stable NFP w/out MD  \\
- [1ex]
- \hline
- \end{tabular}
-
-\medskip
-
-\vspace{2mm}
-
-  \begin{tabular}{|m{18em}|}
- \hline
- Combination 3:\\ [0.5ex]
- \hline\hline
- \vspace{2mm}
- Treat Group: \hspace{17mm} For-Profit \\
- \vspace{2mm}
- Comparison Group: \hspace{5mm} Stable NFPs w/ MD  \\
- [1ex]
- \hline
- \end{tabular}
-\hfil   %<---
-  \begin{tabular}{|m{18em}|}
- \hline
- Combination 4:\\ [0.5ex]
- \hline\hline
- \vspace{2mm}
- Treat Group:  \hspace{13mm} Stable NFPs w/out MD \\
- \vspace{2mm}
- Comparison Group:  \hspace{1mm} Stable NFPs w/ MD  \\
- [1ex]
- \hline
- \end{tabular}
- \end{center}
- \end{figure}
-
- \subsection{Identification}
-
- To identify the causal effect of hospital type on hospital program response, I rely on several assumptions.
-First, I assume that, absent the program enactments, treat and comparison group types would have behaved similarly in outcomes.
+     I rely on several assumptions to identify this effect.
+First, I assume that, absent the program enactments, hospitals with different leadership team characterization types would have behaved similarly in outcomes.
 While I see no evidence of pre-trends, Table \ref{tab:sumstats_samples} suggests that hospitals with a clinically trained executive are different than other hospitals along multiple characteristics.
@@
 
- The estimated difference in readmission rate trends for all hospital types in response to the programs are shown in Figure \ref{fig:read_synth_plot}.
-Each panel corresponds to a particular definition of the treat group of hospitals and limitation of comparison group hospitals.
+    The estimated difference in readmission rate trends for all hospital types in response to the programs are shown in Figure \ref{fig:clinicalsynthdid}.
+In panel \ref{fig:read_synth_clinical}, I compare readmission rate responses among nonprofit hospitals with and without clinical training on their executive team.
+Estimate $\beta$ is represented in the difference in slope after 2012 when pay-for-performance incentives began.
+Nonprofits without a clinically trained executive lowered readmission rates by .24 ppts more than those with a clinically trained executive.
 
-     Panel \ref{fig:read_synth_plota} compares readmission rate responses among for-profits and all NFPs, and shows that on average, there is no difference in how these types of hospitals change readmission rates in response to the incentive changes.
-However, when the comparison group is limited to NFPs with a clinical executive, panel \ref{fig:read_synth_plotb} shows that there is a differential response to the program between these hospital types.
-In particular, for-profits decrease readmission rates by .3 more than NFPs with a clinical executive, an additional 3 patient readmission decrease relative to the average number of patients.
-While numerically small, this difference in response is non-existent between for-profits and NFPs without a clinical executive (panel \ref{fig:read_synth_plotc}), indicating some, if minor, differences in objectives resulting from having a clinical executive.
+    I also directly estimate the difference in mortality rates among NFPs with and without clinical experience in panel \ref{fig:mort_synth_clinical}.
+Again, the difference is not statistically significant, but the coefficient compared to the true nulls in panels \ref{fig:mort_synth_clinical} and \ref{fig:mort_synth_plotc} indicates NFPs without a clinical executive potentially decrease mortality more aggressively than NFPs with a clinical executive.
+In Appendix \ref{app:condition}, I show that these results are mainly driven by pneumonia patients.
 
      \begin{figure}
-     \caption{Readmission Rate Synthetic Difference in Differences Results}
+     \caption{Effect of Clinical Experience on Readmission and Mortality Rates}
      \centering
@@
          \centering
-         \caption{For-Profit and All NFP}
-         \includegraphics[width=\textwidth]{Objects/read_fp_nfp_synth_graph.pdf}
-         \label{fig:read_synth_plota}
-     \end{subfigure}%
+         \caption{Readmission Rate}
+         \includegraphics[width=\textwidth]{Objects/read_md_nomd_synth_graph.pdf}
+         \label{fig:read_synth_clinical}
+     \end{subfigure}
      \hfill
+     \begin{subfigure}[b]{0.45\textwidth}
+         \centering
+         \caption{Mortality Rate}
+         \includegraphics[width=\textwidth]{Objects/mort_md_nomd_synth_graph.pdf}
+         \label{fig:mort_synth_clinical}
+     \end{subfigure}
+        \label{fig:clinicalsynthdid}
+    \end{figure}
+
+    \section{Are Clinical Executives less Profit Driven?}\label{sec:forprofit}
+
+    A reason for the difference in hospital behavior with the presence of clinically trained executives could be how profit-driven a physician vs.
+a general leader is.
+To demonstrate this, consider hospital behavior under the enactment of pay-for-performance policies, where hospitals face a trade-off between profit and societal benefit.
+Pay-for-performance incentives essentially incorporate quality directly into a hospital's profit function (\cite{dranove2011health}), whereas a fee-for-service system only incentivizes quantity.
+A hospital that places all value on profit will only care about quality when it enters into the profit function, and therefore will respond more drastically to pay-for-performance than a hospital who already cared about quality through societal benefit before the incentives.\footnote{I present the theoretical model and mathematical results in Appendix \ref{sec:model}.} Therefore, observing hospital responses to pay-for-performance incentives reveals information about their underlying preferences.
+
+    I compare readmission and mortality rates of each nonprofit leadership team type to for-profit hospitals.
+Since for-profit hospitals have revealed through their ownership status that they are largely profit-driven, comparing different executive team behavior to for-profit behavior helps reveal how profit driven certain leadership teams are.
+Thus, I estimate
+
+    \begin{equation}
+    \label{eq:forprofit}
+    y_{ht} = \beta \text{ (for-profit x post 2012)}_{ht} + \gamma_{h} + \delta_t + \epsilon_{ht},
+    \end{equation}
+
+    \noindent for two subsets of data.
+First, I limit only to for-profits and nonprofits with clinical leadership.
+Second, I limit only to for-profits and nonprofits without clinical leadership.
+Thus, $\beta$ represents the effect of being for-profit compared to being nonprofit with a certain leadership team type.
+
+    The identification assumptions I make here are largely similar to those in the main analysis, only now incorporate for-profit hospitals.
+First, I assume that, absent the change in incentives, for-profit and the relevant comparison group would have had parallel trends in outcomes.
+While this is not directly testable, both the synthetic difference in difference and event studies indicate parallel trends prior to the incentive changes.
+Second, I assume that the stability of nonprofit leadership teams is not correlated with the 2012 policy changes.
+I explore this assumption in Section \ref{sec:identification}.
+I also assume that there is no anticipation of the policy changes prior to 2012, which is reasonable as the rules of the policies were announced in October of 2011.
+Finally, I assume no other changes correlated with both hospital type (for-profit, nonprofit with and without clinically trained executives) and quality occurred in conjunction with the 2012 pay-for-performance policies.
+
+    The estimates and graphical representation of the effect of being for-profit is shown in Figure \ref{fig:read_synth_plot}.
+In panel \ref{fig:read_synth_plotb}, the average treatment effect represents the effect of being for-profit compared to nonprofit with clinical executives on readmission rates after pay-for-performance changes.
+While readmission rates trend similarly prior to the incentive change, for-profits decrease readmission rates by .3 more than nonprofits with clinically trained executives.
+That is, for-profits respond more aggressively to the incentive change than nonprofits with clinical leadership.
+However, panel \ref{fig:read_synth_plotc} shows no such difference when comparing for-profits to nonprofits without clinically trained executives.
+Not only is the estimate not statistically significant, but the magnitude indicates a precise zero.
+That is, a lack of clinically trained executives leads nonprofits to act like for-profits in terms of readmission rates when responding to this incentive change.
+
+ \begin{figure}[ht!]
+     \caption{Readmission Rate Synthetic Difference in Differences Results}
+     \centering
      \begin{subfigure}[b]{0.45\textwidth}
@@
      \end{subfigure}
-     \hfill
-     \begin{subfigure}[b]{0.45\textwidth}
-         \centering
-         \caption{NFP w/ and w/out MD Exec}
-         \includegraphics[width=\textwidth]{Objects/read_md_nomd_synth_graph.pdf}
-         \label{fig:read_synth_plotd}
-     \end{subfigure}
         \label{fig:read_synth_plot}
@@
 
-    Next, in panel \ref{fig:read_synth_plotd}, I directly compare NFPs with and without a clinical executive.
-NFPs without a clinically trained executive lowered readmission rates by .24 ppts more than NFPs with a clinically trained executive.
+    Next, I present estimated differences in mortality rates between different hospital types in Figure \ref{fig:mort_synth_plot}.
+In panel \ref{fig:mort_synth_plotb}, the estimated difference is between for-profits and nonprofits with clinically trained executives.
+The magnitude of the difference in mortality rates is .11, where only for-profits seem to be reacting to the incentive change.
+While still statistically insignificant, I argue that this still indicates a difference in behavior between the hospital types, particularly due to the documented noisiness of mortality measures (\cite{mackenzie2016measuring}).
+However, the estimated difference in mortality between for-profits and nonprofits without clinically trained executives is certainly zero.
 
-    I now shift to discussing differential responses of mortality rates.
-I present estimated differences in mortality rates between different hospital types in Figure \ref{fig:mort_synth_plot}.
-Panel \ref{fig:mort_synth_plota} compares for-profit and all NFP mortality rates after the pay-for-performance incentives.
-These results show that there is no difference, both statistically and in magnitude, between mortality rates of for-profits and not-for-profits after the programs.
-This result holds when the comparison group is limited to NFPs without a clinical executive (panel \ref{fig:mort_synth_plotc}).
-However, when comparing for-profits to NFPs with a clinical executive, the magnitude of the difference in mortality trends jumps to -.11.
-While still statistically insignificant, I argue that this still indicates a difference in behavior between the hospital types, particularly due to the documented noisiness of mortality measures (\cite{mackenzie2016measuring}).
-
-     \begin{figure}
+    \begin{figure}[ht!]
      \caption{Mortality Rate Synthetic Difference in Differences Results}
      \centering
-          \begin{subfigure}[b]{0.45\textwidth}
-         \centering
-         \caption{For-Profit and All NFP}
-         \includegraphics[width=\textwidth]{Objects/mort_fp_nfp_synth_graph.pdf}
-         \label{fig:mort_synth_plota}
-     \end{subfigure}%
-     \hfill
      \begin{subfigure}[b]{0.45\textwidth}
@@
      \end{subfigure}
-     \hfill
-     \begin{subfigure}[b]{0.45\textwidth}
-         \centering
-         \caption{NFP w/ and w/out MD Exec}
-         \includegraphics[width=\textwidth]{Objects/mort_md_nomd_synth_graph.pdf}
-         \label{fig:mort_synth_plotd}
-     \end{subfigure}
         \label{fig:mort_synth_plot}
@@
 
-    I also directly estimate the difference in mortality rates among NFPs with and without clinical experience in panel \ref{fig:mort_synth_plotd}.
-Again, the difference is not statistically significant, but the coefficient compared to the true nulls in panels \ref{fig:mort_synth_plota} and \ref{fig:mort_synth_plotc} indicates NFPs without a clinical executive potentially decrease mortality more aggressively than NFPs with a clinical executive.
-In Appendix \ref{app:condition}, I show that these results are mainly driven by pneumonia patients.
+\section{Signaling vs. Managing Decomposition}
 
-\section{Signalling vs. Managing Decomposition}
+    To this point, I have limited the sample of nonprofit hospitals to those that don't change their propensity to hire a clinically trained executive over the sample period.
+However, under the assumption that hiring or firing these types of executives is not correlated with the pay-for-performance policy changes, these changes can be leveraged to understand further how executive characteristics affect hospital behavior.
+Particularly, a clinically trained executive could be a signal of the underlying objectives of the hospital, or hiring a clinically trained executive who carries out day-to-day operations in a distinct way effectively changes the objectives of the hospital.
 
-\subsection{Method}
+    To disentangle the two, I carefully define the ``treat" and comparison group of hospitals in the following estimating equation.
+For clarity, I present the specification details in Figure \ref{fig:decomp_spec}.
+If clinically trained leaders are simply signals of underlying objectives, it shouldn't matter when hospitals have hired a clinical executive.
+For example, take two hypothetical hospitals.
+Hospital A has a clinically trained executive from 2010-2011.
+Hospital B has a clinical executive from 2010-2013.
+If clinically trained executives are a signal of underlying objectives, these two hospitals should respond similarly to pay-for-performance objectives despite only one of them having a clinical executive in 2012.
+However, if clinically trained leaders manage the hospital in a unique way, these two hospitals would respond differently since only one of them has clinical leadership in 2012 when the policy changes occurred.
 
-Additionally, under certain assumptions that I discuss in Section \ref{sec:identification}, using executive team changes can be advantageous in better understanding underlying hospital motives.
-There are two mechanisms to having a clinically trained executives affecting outcomes: first, a clinically trained executive could be a signal of the underlying weight placed on profit, $\alpha$.
-Second, hiring a clinically trained executive who carries out day-to-day operations in a distinct way effectively changes $\alpha$.
-Thus, I also define treat and limit the comparison group in such a way that decomposes the two.
-Without limiting to stable teams, I consider hospitals that ever hire a clinically trained executive vs.
-hospitals that have a clinically trained executive in 2012 when the programs took place.
-These variations of Equation \ref{eq:forprofit} are shown in Figure \ref{fig:decomp_spec}.
+    \begin{equation}
+    \label{eq:decomp}
+    y_{ht} = \beta \text{ (treat x post 2012)}_{ht} + \gamma_{h} + \delta_t + \epsilon_{ht},
+    \end{equation}
 
@@
  \hline
- Decomposition 1:\\ [0.5ex]
+ Signaling:\\ [0.5ex]
  \hline\hline 
  \vspace{2mm}
- Treat Group:  \hspace{15mm} NFP w/out MD \\
+ Treat Group:  \hspace{15mm} Never MD \\
  \vspace{2mm}
- Comparison Group: \hspace{3mm} NFPs w/ MD  \\
+ Comparison Group: \hspace{3mm} Ever MD  \\
  [1ex]
@@
  \hline
- Decomposition 2:\\ [0.5ex]
+ Managing:\\ [0.5ex]
  \hline\hline
  \vspace{2mm}
- Treat Group: \hspace{11mm} No MD Exec 2012 \\
+ Treat Group: \hspace{11mm} No MD in 2012 \\
  \vspace{2mm}
- Comparison Group:  NFPs w/ MD (not 2012)  \\
+ Comparison Group:  Ever MD (not in 2012)  \\
  [1ex]
@@
 
-    \subsection{Identification}\label{sec:identification}
-
-    To identify the causal effect of hospital type on hospital program response, I rely on several assumptions.
-First, I assume that, absent the program enactments, treat and comparison group types would have behaved similarly in outcomes.
-While I see no evidence of pre-trends, Table \ref{tab:sumstats_samples} suggests that hospitals with a clinically trained executive are different than other hospitals along multiple characteristics.
-There may be concern that these differences between hospital types could be driving the results.
-Therefore, my main specification uses the synthetic difference-in-differences method established in \citeauthor{arkhangelsky2021synthetic} (\citeyear{arkhangelsky2021synthetic}).
-This method creates unit weights that more heavily factor comparison hospitals that are similar to the treat group hospitals prior to the programs.
-It also weighs time periods that balance pre-programs and post-programs for the comparison hospitals.
-Then, a simple two way fixed effects estimator is used with these weights, yielding a more robust average treatment effect that relies less on a strong parallel trends assumption (\cite{arkhangelsky2021synthetic}).
-I present the synthetic diff-in-diff estimates as the main results, and they are largely similar to the standard two way fixed effects results with no weights, shown in Appendix \ref{app:fullsample}.
-Further, the conceptual identification is the same as a standard difference-in-differences.
-
-     Second, I assume that executive teams are not determined endogenously with the programs.
-In the main specification, I limit the sample to hospitals with executive teams that do not change their propensity to hire a clinical executive over time.
-Thus, I rely on this sample selection not being correlated with the program enactment.
-In supplementary analyses, I leverage executive team changes to disentangle signalling vs.
-managing effects, which is only identified if leadership changes are uncorrelated with the policies.
-In Appendix \ref{app:changes}, I present various descriptive statistics of executive team changes over time, showing that hiring and firing of MD executives do not change much over time, the rate of clinical executives to total executives remains stable over time, and the distribution of timing of clinical executive changes.
+    I have discussed the assumptions of parallel trends, no anticipation, and no other confounding events in both Sections \ref{sec:clinical} and \ref{sec:forprofit}.
+These assumptions are also necessary here.
+In those Sections, I have discussed that stability of leadership team must not be correlated with the 2012 policy changes.
+Up to this point, that assumption was important because of the way I selected the sample.
+However, this decomposition depends directly on leveraging such leadership changes.
+Thus, I now explore this assumption in detail.
 
@@
 Thus, it seems unlikely that endogenous team formation is biasing the estimates.
+In Appendix \ref{app:changes}, I present various descriptive statistics of executive team changes over time, showing that hiring and firing of MD executives do not change much over time, the rate of clinical executives to total executives remains stable over time, and the distribution of timing of clinical executive changes.
 
@@
 
-    Finally, I assume that no other unobserved changes occurred in 2012 that are correlated with both hospital performance and type.
-
-    \subsection{Results}
-
-    In the above analysis, I show that NFPs with a clinical executive seem to respond to incentives differently than both for-profits and NFPs without a clinical executive, suggestive evidence of differential objectives correlated with leadership teams.
-However, the analysis to this point does not speak to the mechanism by which clinical executives lead to differential behavior.
-In this section, I present two theories by which a clinical executive leads to differential behavior, and present descriptive evidence that one dominates the other.
-
-    First, having a clinical executive could be a signal of underlying hospital objectives.
-Alternatively, hiring a clinical executive could actually change the objectives of the hospital for the time the clinical executive is actively managing the hospital.
-To disentangle the two, I use variation in the timing of having a clinical executive, with either existence of a hospital executive at any point in time, or whether the hospital has a clinically trained executive in 2012 when the programs were enacted.
-The results are shown in Table \ref{tab:MD_noMD_readmort_decomp_synth}.
+    The estimates of each decomposition specification are shown in Table \ref{tab:MD_noMD_readmort_decomp_synth}.
 Columns (1) and (2) show that the difference in readmission response is driven by having a clinically trained executive in 2012, compared to having a clinically trained executive at another point in time.
@@
 
-     These findings are informative for policymakers predicting how hospitals will responsd to a change in incentives.
+     These findings are informative for policymakers predicting how hospitals will respond to a change in incentives.
 Specifically, they suggest that policymakers should take into account more than just ownership status when evaluating how hospitals will respond to policy changes.
@@
 \end{figure}
+
+\section{Modelling Hospital Response to Pay-for-Performance}\label{sec:model}
+
+    I model hospital behavior under two simplified worlds: one where quality does not directly affect profit, and one where it does.
+Pay-for-performance incentives essentially incorporate performance directly into a hospital's profit function (\cite{dranove2011health}).
+I directly model performance as quality of care, which is inversely related to readmission and mortality rates.
+While the literature speculates about the true form of NFP objective functions, we can generally think of hospitals choosing some linear combination of profit and societal benefit, without knowing how much weight the hospital places on each.
+Intuitively, a hospital that places only weight on profit (for-profit type) will only care about quality when it enters into the profit function, and therefore will respond more drastically to pay-for-performance than a hospital who already cared about quality through societal benefit before the incentives.
+Therefore, hospital responses to pay-for-performance incentives reveal information about their underlying preference on profit vs.
+societal outcomes.
+
+    To formalize this intuition, I specify the hospital objective function as a weighted average of profit and societal (or patient) outcomes and derive characteristics of the optimal quality.
+For simplicity, I abstract away from quality having a direct effect of quantity and price.
+Without any financial incentives to increase quality, hospital quality is decided by:
+
+    $$\max_{\theta}\hspace{2mm}\alpha\left[R - c_{\theta}(\theta) \right] + (1-\alpha) u(\theta),$$
+
+    \noindent where $R$ is net revenue, $c_{\theta}(.)$ is an increasing cost function specific to quality $\theta$, $u(.)$ is extra utility gained from quality, which is increasing and concave in $\theta$, and $\alpha\in[0,1]$ captures the weight placed on profit vs.
+societal benefit.
+There is an implicit stay-open condition throughout the model.
+Taking the first order condition yields
+
+    $$(1-\alpha)u'(\theta) = \alpha c_{\theta}'(\theta),$$
+
+    \noindent where marginal benefit equals marginal cost of increasing quality. Solving for $u'(\theta)$ and differentiating with respect to $\alpha$:
+
+    $$\frac{du'(\theta)}{d\alpha} = \frac{1}{(1-\alpha)^2}c_{\theta}'(\theta) > 0.$$
+
+    \noindent Thus, by the Implicit Function Theorem,
+
+    $$\frac{d\theta}{d\alpha} = \frac{du'(\theta)/d\alpha}{du'(\theta)/d\theta} < 0.$$
+
+    \noindent That is, the more weight on profit, the lower quality of care chosen by the hospital.
+
+    In a world with pay-for-performance incentives, which can either look like benefits from high quality (such as HVBP) or penalties for low quality (as in HRRP), quality directly affects revenue.
+Thus, $R(\theta)$ is an increasing function of $\theta$, and I assume that the marginal financial benefit of increasing quality is greater than the marginal cost ($R'(\theta)\geq c_{\theta}'(\theta))$.
+Thus, the new first order condition yields:
+
+    $$u'(\theta) = \frac{\alpha}{1-\alpha} \left[c_{\theta}'(\theta)-R(\theta)\right] \leq 0.$$
+
+    \noindent Taking the derivative of this with respect to $\alpha$,
+
+    $$\frac{du'(\theta)}{d\alpha} = \frac{1}{(1-\alpha)^2}c_{\theta}'(\theta).$$
+
+    Therefore, again using the Implicit Function Theorem,
+
+    $$\frac{d\theta}{d\alpha} = \frac{du'(\theta)/d\alpha}{du'(\theta)/d\theta}\geq0.$$
+
+    That is, all else equal, in a world with financial incentives on quality, quality is increasing with more weight placed on profit.
+I combine the results found in each scenario into one response function that depends on $\alpha$, where $\theta_2$ is quality under pay-for-performance incentives and $\theta_1$ is quality with no financial incentive to quality.
+That is,
+
+    \begin{align*}
+        \frac{d\Delta\theta}{d\alpha}&=\frac{d(\theta_2-\theta_1)}{d\alpha}\\
+        &=\frac{d\theta_2}{d\alpha}-\frac{d\theta_1}{d\alpha}\\
+        &\geq 0.
+    \end{align*}
+
+    Hence, this model predicts that change in quality depends on how much weight the hospital places on profit vs.
+societal benefit.
+Particularly, hospitals with more weight on profit respond more than hospitals with more weight placed on societal benefit.
+In reality, $\alpha$ is unknown, except in the case of for-profits, who have revealed that they have a high $\alpha$ by their ownership status.
+However, not-for-profit behavior on average does not bring much clarity about their objectives.
+An observable and economically relevant characteristic of a hospital that could affect $\alpha$ is composition of their executive leadership team.
+Thus, I compare the response of for-profits and not-for-profits with different leadership teams to bring clarity on how this might point to underlying $\alpha$.
 

### 2024-07-23 Slides/Europe2024conf.tex
Updates from Overleaf

\documentclass[notes,11pt, aspectratio=169]{beamer}

% { BEGIN_FOLD ---------- PREAMBLE -------------

\usepackage{pgfpages}
% These slides also contain speaker notes. You can print just the slides,
% just the notes, or both, depending on the setting below. Comment out the want
% you want.
\setbeameroption{hide notes} % Only slide
%\setbeameroption{show only notes} % Only notes
%\setbeameroption{show notes on second screen=right} % Both

\usepackage{helvet}
\usepackage[default]{lato}
\usepackage{array}

\usepackage[backend=biber,style=authoryear,
sorting=ynt,citestyle=authoryear]{biblatex}
\addbibresource{Paper/papercitations.bib}

\usepackage{tikz}
\usepackage{verbatim}
\setbeamertemplate{note page}{\pagecolor{yellow!5}\insertnote}
\usetikzlibrary{positioning}
\usetikzlibrary{snakes}
\usetikzlibrary{calc}
\usetikzlibrary{arrows}
\usetikzlibrary{decorations.markings}
\usetikzlibrary{shapes.misc}
\usetikzlibrary{matrix,shapes,arrows,fit,tikzmark}
\usepackage{amsmath}
\usepackage{mathpazo}
\usepackage{hyperref}
\usepackage{lipsum}
\usepackage{multimedia}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{dcolumn}
\usepackage{bbm}
\newcolumntype{d}[0]{D{.}{.}{5}}
\usepackage{subfigure}
\usepackage{import}
\usepackage{xcolor}

\usepackage{changepage}
\usepackage{appendixnumberbeamer}
\newcommand{\beginbackup}{
   \newcounter{framenumbervorappendix}
   \setcounter{framenumbervorappendix}{\value{framenumber}}
   \setbeamertemplate{footline}
   {
     \leavevmode%
     \hline
     box{%
       \begin{beamercolorbox}[wd=\paperwidth,ht=2.25ex,dp=1ex,right]{footlinecolor}%
%         \insertframenumber  \hspace*{2ex} 
       \end{beamercolorbox}}%
     \vskip0pt%
   }
 }
\newcommand{\backupend}{
   \addtocounter{framenumbervorappendix}{-\value{framenumber}}
   \addtocounter{framenumber}{\value{framenumbervorappendix}} 
}

\newcommand{\btVFill}{\vskip0pt plus 1filll}

\setbeamertemplate{blocks}[rounded]
\setbeamercolor{block title}{bg=purple, fg=white}
\setbeamercolor{block body}{bg=purple!10}

\usepackage{graphicx}
\usepackage[space]{grffile}
\usepackage{booktabs}

% These are my colors -- there are many like them, but these ones are mine.
\definecolor{sage}{RGB}{102,153,102}
\definecolor{yellow}{RGB}{255,173,1}
\definecolor{purple}{RGB}{153,102,153}

% colors for diagrams
\definecolor{diagramtan}{RGB}{225,190,106}
\definecolor{diagramteal}{RGB}{64, 176, 166}
\definecolor{diagrampurple}{RGB}{170, 131, 239}
\definecolor{diagramred}{RGB}{146, 79, 79}

\hypersetup{
  colorlinks=false,
  linkbordercolor = {white},
  linkcolor = {sage}
}

%% I use a beige off white for my background
\definecolor{MyBackground}{RGB}{255,253,218}

%% Uncomment this if you want to change the background color to something else
%\setbeamercolor{background canvas}{bg=MyBackground}

%% Change the bg color to adjust your transition slide background color!
\newenvironment{transitionframe}{
  \setbeamercolor{background canvas}{bg=white}
  \begin{frame}}{
    \end{frame}
}

\setbeamercolor{frametitle}{fg=sage}
\setbeamercolor{title}{fg=sage}
\setbeamertemplate{footline}[frame number]
\setbeamertemplate{navigation symbols}{} 
\setbeamertemplate{itemize items}{$\rightarrow$}
\setbeamertemplate{itemize subitem}{$\bullet$}
\setbeamercolor{itemize item}{fg=sage}
\setbeamercolor{itemize subitem}{fg=sage}
\setbeamercolor{enumerate item}{fg=sage}
\setbeamercolor{enumerate subitem}{fg=sage}
\setbeamercolor{button}{bg=sage,fg=white}

% If you like road maps, rather than having clutter at the top, have a roadmap show up at the end of each section 
% (and after your introduction)
% Uncomment this is if you want the roadmap!
\AtBeginSection[]
{
   \begin{frame}
       \frametitle{Roadmap of Talk}
       \tableofcontents[currentsection]
   \end{frame}
}
\setbeamercolor{section in toc}{fg=sage}
\setbeamercolor{subsection in toc}{fg=sage}
\setbeamersize{text margin left=1em,text margin right=1em}

\newenvironment{wideitemize}{\itemize\addtolength{\itemsep}{10pt}}{\enditemize}

\usepackage{environ}
\NewEnviron{videoframe}[1]{
  \begin{frame}
    \vspace{-8pt}
    \begin{columns}[onlytextwidth, T] % align columns
      \begin{column}{.58\textwidth}
        \begin{minipage}[t][\textheight][t]
          {\dimexpr\textwidth}
          \vspace{8pt}
          \hspace{4pt} {\Large \sc \textcolor{blue}{#1}}
          \vspace{8pt}

          \BODY
        \end{minipage}
      \end{column}%
      \hfill%
      \begin{column}{.42\textwidth}
        \colorbox{green!20}{\begin{minipage}[t][1.2\textheight][t]
            {\dimexpr\textwidth}
            Face goes here
          \end{minipage}}
      \end{column}%
    \end{columns}
  \end{frame}
}

% } ---------- PREAMBLE -------------

\title[]{\textcolor{sage}{Does Hospital Leadership Matter? \\Evidence from Pay-for-Performance Incentives}}

\author[]{Hanna Glenn\\ \textit{Emory University}}
\date{\today}

\begin{document}

%%% { TIKZ STUFF
\tikzset{   
        every picture/.style={remember picture,baseline},
        every node/.style={anchor=base,align=center,outer sep=1.5pt},
        every path/.style={thick},
        }
\newcommand\marktopleft[1]{%
    \tikz[overlay,remember picture] 
        \node (marker-#1-a) at (-.3em,.3em) {};%
}
\newcommand\markbottomright[2]{%
    \tikz[overlay,remember picture] 
        \node (marker-#1-b) at (0em,0em) {};%
}
\tikzstyle{every picture}+=[remember picture] 
\tikzstyle{mybox} =[draw=black, very thick, rectangle, inner sep=10pt, inner ysep=20pt]
\tikzstyle{fancytitle} =[draw=black,fill=red, text=white]
%%%% } ------------ TIKZ STUFF ---------

% Title Slide
\begin{frame}
\maketitle
\end{frame}

\begin{frame}{Leader Characteristics Influence Firms}

    \vspace{5mm}

    \begin{wideitemize}
        \item Female executives correlated with female wages (\cite{flabbi2019female})
        \item Young male CEOs more aggressive in mergers (\cite{levi2010deal})
        \item Military CEOs less aggressive (\cite{benmelech2015military})
        \item General ability CEOs are at better performing firms than specialized ability (\cite{custodio2013generalists})
    \end{wideitemize}
    \btVFill \pause

    \begin{block}{Gap}
        \begin{itemize}
            \item What we know is typically in large, publicly traded firms
            \item Limited evidence in not-for-profit (NFP) settings with different incentives
        \end{itemize}
    \end{block}

\end{frame}

\begin{frame}{Not-for-Profit Behavior}

    \vspace{12mm}
    NFP objectives in general are unclear

    \vspace{8mm}
    \begin{wideitemize}
        \item Focus in research on understanding NFP hospitals: profit driven or not?
        \item NFP hospitals in the US: private but designated as NFP by taxes
        \begin{itemize}
            \item Make up majority of hospitals in the US
            \item Behavior has direct impact on patient population
        \end{itemize}
    \end{wideitemize}

    \btVFill \pause

    \begin{block}{This Paper}
        Examine whether executive team characteristics reveal NFP hospital behavior
        \begin{itemize}
            \item leverage exogenous shock to incentives: pay-for-performance
        \end{itemize}
    \end{block}

\end{frame}

\begin{frame}{Hospital Choice of Quality}
\vspace{5mm}
$$\max_{\theta}\hspace{2mm}\alpha\pi(\theta) + (1-\alpha) u(\theta)$$
\begin{itemize}
        \item $\theta$: quality
        \item $\pi(\theta)$: profit function 
        \item $u(\theta)$: patient/society benefit
        \item $\alpha\in\left[0,1\right]$: weight placed on profit function 
    \end{itemize}
    \btVFill

\end{frame}

\begin{frame}[noframenumbering]{Hospital Choice of Quality}
\vspace{5mm}
$$\max_{\theta}\hspace{2mm}\alpha\pi(\theta) + (1-\alpha) u(\theta)$$

    \vspace{3mm}

The US health care system is still largely fee-for-service
\begin{itemize}
    \item providers do not financially benefit from quality, only from quantity
    \item $\pi(\theta)$ is decreasing in $\theta$
\end{itemize}
    \begin{block}{}
        Quality is decreasing in $\alpha$ 
    \end{block}

\vspace{3mm}\pause

With pay-for-performance:
\begin{itemize}
    \item Hospitals are rewarded for high quality or penalized for low quality
    \item $\pi(\theta)$ is an increasing function with $\pi(\theta)\geq 0$
\end{itemize}
    \begin{block}{}
        Quality is increasing in $\alpha$ 
    \end{block}

    \btVFill
\end{frame}

\begin{frame}{Response to P4P}
\vspace{10mm}
    \begin{align*}
        \Delta\theta &= \theta_{P4P}-\theta\\
        \\
        \frac{d\Delta\theta}{d\alpha}&=\frac{d(\theta_{P4P}-\theta)}{d\alpha}\\
        &=\frac{d\theta_{P4P}}{d\alpha}-\frac{d\theta}{d\alpha}\\
        &\geq 0.
    \end{align*}
    \btVFill\pause

    \begin{block}{}
    Hospital response to P4P is increasing as more weight is placed on profit
    \end{block}
\end{frame}

\begin{frame}{This Paper}
\vspace{5mm}
\begin{wideitemize}
    \item Objectives reflected in $\alpha$: does leadership team help reveal these objectives?
    \item Debated characteristic of hospital leaders: clinical experience
    \begin{itemize}
        \item better understanding of patients
        \item lack of managerial experience (\cite{otero2022managers})
        \item could affect weight placed on profit vs. patients
    \end{itemize}
\end{wideitemize}

\btVFill\pause

\begin{block}{}
    I compare response to P4P between NFP hospitals with and without clinical executives
    \begin{itemize}
        \item I find that NFP hospitals without clinical executives respond more aggressively to P4P
    \end{itemize}\pause
\end{block}
\begin{block}{Mechanisms}
    \begin{itemize}
        \item Are clinical executives less profit driven?
        \item Are clinical executives a signal of underlying objectives or do they manage the hospital differently? 
    \end{itemize}
\end{block}

\end{frame}

\begin{frame}{Clinical Executives}
NFP executives are understudied partly because of difficulty identifying them

\vspace{4mm}

Scrape executive teams from public Tax Form 990s

\vspace{4mm}

    \centering
    \includegraphics[width=.8\textwidth]{Graphics/990_snip_frontpage.PNG}
\end{frame}

\begin{frame}{Clinical Executives}
Section about executives and board members:
\begin{itemize}
                \item Use OCR text extraction algorithms to get names, positions, titles
                \item Limit to executives
            \end{itemize}

            \vspace{3mm}

\centering
\includegraphics[width=.75\textwidth]{Graphics/990_snip_namesexample.PNG}

\end{frame}

\begin{frame}{Clinical Executives}\label{clinexecs}
    \vspace{10mm}
    \begin{wideitemize}
            \item Characterize NFP hospitals as having any executives trained as an MD
            \item Focus on stable teams
            \begin{itemize}
                \item don't change propensity to hire an MD over time
            \end{itemize}
            \item 521 NFP hospitals w/ stable teams 
            \begin{itemize}
                \item 10\% have MD exec
            \end{itemize}
    \end{wideitemize}
    \btVFill
    \hyperlink{990s}{\beamerbutton{getting data}} \hyperlink{execstats}{\beamerbutton{stats}}
\end{frame}

\begin{frame}{2012 Pay-for-Performance Policies}\label{policies}
    \vspace{7mm}

    Hospital Readmissions Reduction Program:
    \begin{itemize}
        \item Penalty for high readmissions for given conditions: heart attack, heart failure, pneumonia

    \end{itemize}

    \vspace{5mm}

    Hospital Value Based Purchasing Program:
    \begin{itemize}
        \item Rewarded based on quality compared to other hospitals and whether quality improves over time
    \end{itemize}

    \btVFill

    \hyperlink{policies}{\beamerbutton{policies}}

    \vspace{2mm}

    \begin{block}{}
        Exogenous shocks to hospitals: predetermined leadership teams are not correlated with the incentive change
    \end{block}
\end{frame}

\begin{frame}{Estimating the Effect of Clinical Executives on Quality after P4P}
\begin{equation*}
    y_{ht} = \beta \text{ treat}_{ht} + \gamma_{h} + \delta_t + \epsilon_{ht}
\end{equation*}

\vspace{3mm}

\begin{wideitemize}
    \item h: NFP hospitals in the US
    \begin{itemize}
        \item limit to general hospitals w/ stable executive teams
    \end{itemize}
    \item t: year $\in$ 2010-2014
    \item $y_{ht}$: Readmission and mortality rates
    \begin{itemize}
        \item I create a weighted average based on number of patients in each relevant condition
        \item data: government collection of hospital quality statistics (Hospital Compare)
    \end{itemize} 
    \item treat$_{ht}$: Has clinical executive x Post P4P
\end{wideitemize}
\btVFill
\begin{block}{}
    Ultimately employ a synthetic diff-in-diff
\end{block}
\end{frame}

\begin{frame}{Effect of Leadership on Readmission and Mortality}
\begin{columns}
    \begin{column}{0.5\textwidth}
        \centering
        Readmission

        \vspace{1mm}
        \includegraphics[width=.8\textwidth]{Objects/read_md_nomd_synth_graph.pdf}
    \end{column}
        \begin{column}{0.5\textwidth}
        \centering
        Mortality

        \vspace{1mm}
        \includegraphics[width=.8\textwidth]{Objects/mort_md_nomd_synth_graph.pdf}
    \end{column}
\end{columns}
\end{frame}

\begin{frame}{Are non-clinical execs more profit driven?}
\begin{block}{}
    Compare each type of NFP w/ for-profit
\end{block}
\begin{equation*}
    y_{ht} = \beta \text{ treat}_{ht} + \gamma_{h} + \delta_t + \epsilon_{ht}
\end{equation*}

\begin{figure}[ht!]
    \begin{center}
 \begin{tabular}{| m{18em} |}
 \hline
 Main Analysis:\\ [0.5ex]
 \hline\hline 
 \vspace{2mm}
 Treat: \hspace{20mm} NFP w/out MD \\
 \vspace{2mm} 
 Comparison: \hspace{8mm} NFP w/ MD  \\
 [1ex]
 \hline
 \end{tabular}
\hfil   %<---

\medskip

\vspace{2mm}

  \begin{tabular}{|m{18em}|}
 \hline
 Comparison 1:\\ [0.5ex]
 \hline\hline
 \vspace{2mm}
 Treat: \hspace{20mm} For-Profit \\
 \vspace{2mm}
 Comparison: \hspace{8mm} NFP w/ MD  \\
 [1ex]
 \hline
 \end{tabular}
\hfil   %<---
  \begin{tabular}{|m{18em}|}
 \hline
 Comparison 2:\\ [0.5ex]
 \hline\hline
 \vspace{2mm}
 Treat:  \hspace{20mm} For-Profit \\
 \vspace{2mm}
 Comparison:  \hspace{8mm} NFP w/out MD  \\
 [1ex]
 \hline
 \end{tabular}
 \end{center}
 \end{figure}

\end{frame}

\begin{frame}{Compare to FP: Readmissions}
\begin{columns}
    \begin{column}{0.5\textwidth}
        \centering
        FP vs. NFP w/ MD Exec\\\vspace{2mm}
        \includegraphics[width=.75\textwidth]{Objects/read_fp_md_synth_graph.pdf}
    \end{column}
        \begin{column}{0.5\textwidth}
        \centering
        FP vs. NFP w/out MD Exec

        \vspace{2mm}
        \includegraphics[width=.75\textwidth]{Objects/read_fp_nomd_synth_graph.pdf}
    \end{column}
\end{columns}
\end{frame}

\begin{frame}{Compare to FP: Mortality}
\begin{columns}
    \begin{column}{0.5\textwidth}
        \centering
        FP vs. NFP w/ MD Exec\\\vspace{2mm}
        \includegraphics[width=.75\textwidth]{Objects/mort_fp_md_synth_graph.pdf}
    \end{column}
        \begin{column}{0.5\textwidth}
        \centering
        FP vs. NFP w/out MD Exec

        \vspace{2mm}
        \includegraphics[width=.75\textwidth]{Objects/mort_fp_nomd_synth_graph.pdf}
    \end{column}
\end{columns}
\end{frame}

\begin{frame}{Signalling vs. Managing?}\label{sigman}
    \vspace{7mm}

   Are clinical executives signal of objectives or do they change objectives?

   \vspace{7mm}

\begin{block}{}
    Include hospitals with change in propensity to hire executive MD
    \begin{itemize}
        \item leverage timing of changes
        \item requires changes not endogenous with P4P incentives \hyperlink{execchanges}{\beamerbutton{change analysis}}
    \end{itemize}
\end{block}
\btVFill\pause

\begin{figure}[ht!]
    \begin{center}
 \begin{tabular}{| m{18em} |}
 \hline
 Signalling:\\ [0.5ex]
 \hline\hline 
 \vspace{2mm}
 Treat: \hspace{20mm} Never MD Exec \\
 \vspace{2mm} 
 Comparison: \hspace{8mm} Ever MD Exec  \\
 [1ex]
 \hline
 \end{tabular}
\hfil   %<---
 \begin{tabular}{|m{18em}|}
 \hline
 Managing:\\ [0.5ex]
 \hline\hline
 \vspace{2mm}
 Treat: \hspace{20mm} MD Exec (not 2012) \\
 \vspace{2mm}
 Comparison: \hspace{8mm} MD Exec 2012  \\
 [1ex]
 \hline
 \end{tabular}

 \end{center}
 \end{figure}
\end{frame}

\begin{frame}{Signalling vs. Managing?}

\footnotesize
\begin{table}[ht!]
\centering
\begin{tabular}[t]{lcccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{\underline{Weighted Avg. Readmission Rate}} & \multicolumn{2}{c}{\underline{Weighted Avg. Mortality Rate}} \\
\addlinespace
 & (1) & (2) & (3) & (4)\\
\midrule
Never MD Exec x Post Programs & 0 &  & -0.06 & \\
 & (0.06) &  & (0.06) & \\
Ever MD (not in 2012) &  & -0.19$^{*}$ &  & -0.1\\
 \hspace{3mm} x Post Programs &  & (0.09) &  & (0.08)\\
 &  &  &  & \\
\addlinespace
\textbf{Comparison Group:} &  &  &  & \\
Ever MD Exec & $\checkmark$ &  & $\checkmark$ & \\
Has MD Exec 2012 &  & $\checkmark$ &  & $\checkmark$\\
 &  &  &  & \\
Observations & 3,925 & 1,845 & 3,910 & 1,840\\
\bottomrule
\end{tabular}
\end{table}

\end{frame}

\begin{frame}{Conclusion}
I examine whether having clinical experience on a leadership team affects how the hospital responds to a change in incentives

\vspace{5mm}

\begin{enumerate}
    \item Hospitals w/out clinical executives respond more aggressively to P4P incentives \pause

    \vspace{3mm}

    \item Hospitals w/out clinical executive act more like for-profits \pause

    \vspace{3mm}

    \item Seems to be driven by differences in management, not a signal of underlying objectives
\end{enumerate}
\end{frame}

\begin{frame}{}
    \Large
    \centering

    \textcolor{sage}{Thank you!}

    \vspace{5mm}

    \textcolor{sage}{hanna.glenn@emory.edu}

\end{frame}

\begin{frame}[noframenumbering]{Tax Form 990s: Collection}\label{990s}

\vspace{5mm}
    \begin{wideitemize}
        \item Forms stored at ProPublica
        \item Use API to collect EIN numbers from NFPs classified as hospitals 
        \item For each EIN, download pdf locally using URLs and API
        \item Manually collect those that get skipped
    \end{wideitemize}
    \btVFill

    \hyperlink{clinexecs}{\beamerbutton{back}}
\end{frame}

\begin{frame}[noframenumbering]{Tax Form 990s: Matching}
\vspace{5mm}
    Need to match to AHA data to get other hospital characteristics
    \vspace{3mm}
    \begin{wideitemize}
        \item Crosswalks from EIN to AHA ID or Medicare number do not exist
        \item Best I can do is match based on name of hospital and location
        \item I match based on exact name matches in the same state (800 matches)
        \item Then I match based on rough name matches in the same state (300 matches)
        \item Manually match as many as I can (100 matches)
    \end{wideitemize}
    \btVFill

    \hyperlink{clinexecs}{\beamerbutton{back}}
\end{frame}

\begin{frame}[noframenumbering]{Tax Form 990s: Text Extraction and Cleaning}
\vspace{5mm}
    \begin{wideitemize}
        \item Extract text from all pages with executives and key directors
        \item find name locations based on common appearances of Census first and last names
        \item Locate position and title that appear closest to the name
    \end{wideitemize}

    \btVFill

    \hyperlink{clinexecs}{\beamerbutton{back}}
\end{frame}

\begin{frame}[noframenumbering]{Executive Team Statistics}\label{execstats}

\vspace{10mm}
    \begin{table}[ht!]
    \centering
    \begin{tabular}[t]{lcccc}
    \toprule
    Variable & Mean & Std. Dev. & Min & Max\\
    \midrule
    Number Executives & 3.91 & 2.75 & 1 & 27\\
    Number MD Executives & 0.40 & 0.73 & 0 & 6\\
    \hspace{5mm}Percent MD Execs & 0.09 & 0.17 & 0 & 1\\
    No Change in MD Hiring Propensity & 0.61 & 0.49 & 0 & 1\\
    \bottomrule
    \end{tabular}
    \end{table}

    \btVFill

    \hyperlink{clinexecs}{\beamerbutton{back}}
\end{frame}

\begin{frame}[noframenumbering]{Endogenous Executive Changes?}\label{execchanges}
\small
    \begin{table}[ht!]
   \centering
   \begin{tabular}{lcccccc}
      \toprule
       & \multicolumn{3}{c}{Change in Any MD} & \multicolumn{3}{c}{Change in Num. MDs}\\
                                      & (1)    & (2)    & (3)           & (4)    & (5)    & (6)\\  
      \midrule 
      Ever Penalized HRRP x Post 2011 & 0.01   &        &               & 0.02   &        &   \\   
                                      & (0.02) &        &               & (0.02) &        &   \\   
      Ever Payment HVBP x Post 2011   &        & 0.00   &               &        & 0.02   &   \\   
                                      &        & (0.02) &               &        & (0.02) &   \\   
      2011                            &        &        & -0.02         &        &        & -0.02\\   
                                      &        &        & (0.02)        &        &        & (0.02)\\   
      2012                            &        &        & -0.05$^{**}$  &        &        & -0.04$^{*}$\\   
                                      &        &        & (0.02)        &        &        & (0.02)\\   
      2013                            &        &        & -0.04$^{**}$  &        &        & -0.03\\   
                                      &        &        & (0.02)        &        &        & (0.02)\\   
      2014                            &        &        & -0.05$^{***}$ &        &        & -0.03\\   
                                      &        &        & (0.02)        &        &        & (0.02)\\   
       \\
      Observations                    & 3,808  & 3,808  & 3,808         & 3,808  & 3,808  & 3,808\\  
      \bottomrule
   \end{tabular}
\end{table}

\hyperlink{sigman}{\beamerbutton{back}}

\end{frame}

\begin{frame}[noframenumbering]{2012 Pay-for-Performance Policies}\label{policies}
    \vspace{3mm}

    Hospital Readmissions Reduction Program:
    \begin{itemize}
        \item Penalty for high readmissions for given conditions: heart attack, heart failure, pneumonia
        \item If penalized, pay 1-3\% reimbursement for \textit{every} Medicare patient
        \begin{itemize}
            \item 19\% of the US population is on Medicare
            \item 21\% of health care is paid by Medicare
        \end{itemize}

    \end{itemize}

    \vspace{5mm}

    Hospital Value Based Purchasing Program:
    \begin{itemize}
        \item Rewarded based on quality compared to other hospitals and whether quality improves over time
        \item All hospitals pay in, hospitals above average receive the payment
    \end{itemize}
    \btVFill

    \begin{block}{}
        Exogenous shocks to hospitals: predetermined leadership teams are not correlated with the incentive change \hyperlink{policies}{\beamerbutton{back}}
    \end{block}
\end{frame}

\begin{frame}[noframenumbering]{Summary Stats: Hospital Characteristics}
    \begin{table}[ht!]
\centering
\begin{tabular}[t]{lcccc}
\toprule
Variable & Mean & Std. Dev. & Min & Max\\
\midrule
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{Hospital Characteristics}}\\
\hspace{1em}Number Beds & 162.15 & 211.79 & 10 & 9,900\\
\hspace{1em}Num. Patients AMI & 143.23 & 198.32 & 1 & 1,667\\
\hspace{1em}Num. Patients HF & 325.50 & 337.01 & 1 & 2,547\\
\hspace{1em}Num. Patients Pneumonia & 283.11 & 230.93 & 1 & 2,032\\
\hspace{1em}Academic Med. Center & 0.25 & 0.43 & 0 & 1\\
\hspace{1em}Physician Owned & 0.08 & 0.27 & 0 & 1\\
\hspace{1em}System Affiliated & 0.57 & 0.50 & 0 & 1\\
\hspace{1em}Case Mix Index & 1.44 & 0.26 & 0.58 & 3.75\\
\bottomrule
\end{tabular}
\end{table}
\end{frame}

\begin{frame}[noframenumbering]{Summary Stats: Penalty/Payment}
    \begin{table}[ht!]
\centering
\begin{tabular}[t]{lcccc}
\toprule
Variable & Mean & Std. Dev. & Min & Max\\
\midrule
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{Penalty/Payment Variables}}\\
\hspace{1em}Ever Received HVBP Incentive & 0.77 & 0.42 & 0.00 & 1.00\\
\hspace{1em}Penalized for AMI & 0.04 & 0.19 & 0.00 & 1.00\\
\hspace{1em}Penalized for HF & 0.05 & 0.22 & 0.00 & 1.00\\
\hspace{1em}Penalized for Pneumonia & 0.06 & 0.24 & 0.00 & 1.00\\
\hspace{1em}Penalized for AMI + HF & 0.05 & 0.23 & 0.00 & 1.00\\
\hspace{1em}Penalized for AMI + Pneumonia & 0.15 & 0.36 & 0.00 & 1.00\\
\hspace{1em}Penalized for HF + Pneumonia & 0.15 & 0.36 & 0.00 & 1.00\\
\hspace{1em}Penalized for All Conditions & 0.30 & 0.46 & 0.00 & 1.00\\
\bottomrule
\end{tabular}
\end{table}
\end{frame}

\begin{frame}[noframenumbering]{Summary Stats: Outcomes}
    \begin{table}[ht!]
\centering
\begin{tabular}[t]{lcccc}
\toprule
Variable & Mean & Std. Dev. & Min & Max\\
\midrule
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{Readmission Outcome Variables}}\\
\hspace{1em}Weighted Avg. Readmission Rate & 20.76 & 1.86 & 15.10 & 29.09\\
\hspace{1em}AMI Readmission Rate & 19.14 & 1.53 & 14.50 & 26.30\\
\hspace{1em}HF Readmission Rate & 24.15 & 2.08 & 16.60 & 33.60\\
\hspace{1em}Pneum. Readmission Rate & 18.15 & 1.61 & 13.60 & 24.80\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{Mortality Outcome Variables}}\\
\hspace{1em}Weighted Avg. Mortality Rate & 12.37 & 1.42 & 7.23 & 20.25\\
\hspace{1em}AMI Mortality Rate & 15.60 & 1.60 & 10.10 & 22.70\\
\hspace{1em}HF Mortality Rate & 11.66 & 1.58 & 6.00 & 18.00\\
\hspace{1em}Pneum. Mortality Rate & 12.01 & 1.88 & 6.40 & 22.10\\
\\
Num. Hospitals & 1560.00 &  &  & \\
\bottomrule
\end{tabular}
\end{table}
\end{frame}

\begin{frame}[noframenumbering]{Sub-sample Summary Stats: Hospital Characteristics}
    \begin{table}[ht!]
    \small
\centering
\begin{tabular}[t]{lcccccl}
\toprule
 & (1) & (2) & (3) & (4) & (5) & (6)\\
\addlinespace[0.3em]
\multicolumn{2}{c}{ } & \multicolumn{3}{c}{Not-for-Profit} & \multicolumn{2}{c}{Stable Not-for-Profit} \\
\cmidrule(l{3pt}r{3pt}){3-5} \cmidrule(l{3pt}r{3pt}){6-7}
Variable & For-Profit & All & Ever MD & Never MD & Ever MD & Never MD\\
\midrule
\addlinespace[0.3em]
\multicolumn{7}{l}{\textbf{Hospital Characteristics}}\\
\hspace{1em}Number Beds & 155 & 149 & 232 & 115 & 216 & 115\\
\hspace{1em}Num. Patients AMI & 112 & 138 & 228 & 109 & 209 & 109\\
\hspace{1em}Num. Patients HF & 276 & 299 & 475 & 266 & 453 & 265\\
\hspace{1em}Num. Patients Pneum. & 234 & 258 & 394 & 256 & 390 & 256\\
\hspace{1em}Academic Med. Center & 0.17 & 0.27 & 0.40 & 0.24 & 0.48 & 0.24\\
\hspace{1em}Physician Owned & 0.19 & 0.01 & 0.01 & 0.01 & 0.03 & 0.01\\
\hspace{1em}System Affiliated & 0.90 & 0.60 & 0.60 & 0.43 & 0.44 & 0.28\\
\hspace{1em}Case Mix Index & 1.44 & 1.44 & 1.50 & 1.37 & 1.52 & 1.37\\
\addlinespace[0.3em]
Num. Hospitals & 708 & 852 & 384 & 468 & 56 & 465\\
\bottomrule
\end{tabular}
\end{table}
\end{frame}

\begin{frame}[noframenumbering]{Sub-sample Summary Stats: Penalty/Payment}
    \begin{table}[ht!]
    \footnotesize
\centering
\begin{tabular}[t]{lcccccl}
\toprule
 & (1) & (2) & (3) & (4) & (5) & (6)\\
\addlinespace[0.3em]
\multicolumn{2}{c}{ } & \multicolumn{3}{c}{Not-for-Profit} & \multicolumn{2}{c}{Stable Not-for-Profit} \\
\cmidrule(l{3pt}r{3pt}){3-5} \cmidrule(l{3pt}r{3pt}){6-7}
Variable & For-Profit & All & Ever MD & Never MD & Ever MD & Never MD\\
\midrule
\addlinespace[0.3em]
\multicolumn{7}{l}{\textbf{Penalty/Payment Variables}}\\
\hspace{1em}Ever Received HVBP & 0.87 & 0.62 & 0.81 & 0.59 & 0.71 & 0.59\\
\hspace{1em}Penalized for AMI & 0.03 & 0.03 & 0.05 & 0.03 & 0.04 & 0.03\\
\hspace{1em}Penalized for HF & 0.06 & 0.06 & 0.04 & 0.05 & 0.05 & 0.05\\
\hspace{1em}Penalized for Pneumonia & 0.06 & 0.05 & 0.09 & 0.04 & 0.09 & 0.05\\
\hspace{1em}Penalized for AMI + HF & 0.06 & 0.04 & 0.06 & 0.04 & 0.05 & 0.04\\
\hspace{1em}Penalized for AMI + Pneum. & 0.04 & 0.05 & 0.07 & 0.05 & 0.07 & 0.05\\
\hspace{1em}Penalized for HF + Pneum. & 0.20 & 0.13 & 0.12 & 0.12 & 0.11 & 0.12\\
\hspace{1em}Penalized for All Conditions & 0.36 & 0.22 & 0.31 & 0.21 & 0.30 & 0.21\\
\addlinespace[0.3em]
Num. Hospitals & 708 & 852 & 384 & 468 & 56 & 465\\
\bottomrule
\end{tabular}
\end{table}
\end{frame}

\begin{frame}[noframenumbering]{Sub-sample Summary Stats: Outcomes}
    \begin{table}[ht!]
    \footnotesize
\centering
\begin{tabular}[t]{lcccccl}
\toprule
 & (1) & (2) & (3) & (4) & (5) & (6)\\
\addlinespace[0.3em]
\multicolumn{2}{c}{ } & \multicolumn{3}{c}{Not-for-Profit} & \multicolumn{2}{c}{Stable Not-for-Profit} \\
\cmidrule(l{3pt}r{3pt}){3-5} \cmidrule(l{3pt}r{3pt}){6-7}
Variable & For-Profit & All & Ever MD & Never MD & Ever MD & Never MD\\
\midrule
\addlinespace[0.3em]
\multicolumn{7}{l}{\textbf{Readmission Outcome Variables}}\\
\hspace{1em}Weighted Avg. Readmission & 21.09 & 20.47 & 20.63 & 20.44 & 20.79 & 20.44\\
\hspace{1em}AMI Readmission Rate & 19.19 & 19.13 & 19.01 & 19.23 & 18.90 & 19.23\\
\hspace{1em}HF Readmission Rate & 24.42 & 23.99 & 23.81 & 24.08 & 23.93 & 24.08\\
\hspace{1em}Pneum. Readmission Rate & 18.30 & 18 & 18.10 & 18 & 18.26 & 18\\
\addlinespace[0.3em]
\multicolumn{7}{l}{\textbf{Mortality Outcome Variables}}\\
\hspace{1em}Weighted Avg. Mortality Rate & 12.38 & 12.29 & 12.36 & 12.36 & 12.34 & 12.36\\
\hspace{1em}AMI Mortality Rate & 15.82 & 15.46 & 15.29 & 15.61 & 15.17 & 15.61\\
\hspace{1em}HF Mortality Rate & 11.52 & 11.67 & 11.71 & 11.83 & 11.57 & 11.82\\
\hspace{1em}Pneum. Mortality Rate & 12.08 & 11.98 & 11.83 & 12.07 & 12 & 12.06\\
\addlinespace[0.3em]
Num. Hospitals & 708 & 852 & 384 & 468 & 56 & 465\\
\bottomrule
\end{tabular}
\end{table}
\end{frame}

\end{document}

### 2024-07-23 Paper/paper.tex

@@
 		{\small
-        There is mixed evidence on whether not-for-profit hospitals differ meaningfully from for-profit (FP) hospitals, but what has not been considered is the effect leadership teams have on hospital behavior.
-In this paper, I investigate whether leadership team composition of NFP hospitals affects how similarly they behave to for-profits.
-I use a novel data set on NFP hospital executives in the US (2010-2014) to investigate whether the presence of clinically trained executives affects hospital behavior.
-In particular, I leverage pay-for-performance initiatives in the US enacted in 2012, and measure the  readmission and mortality rate responses between FP and NFP with and without clinically trained executives.
-I find that for-profits nad NFPs without clinical executives respond more aggressively to financial incentives on quality, consistent with theoretical predictions of firms that value profit relatively more than having high quality.
+        In highly specialized careers, it is not uncommon to have leaders with specific training, such as lawyers becoming judges, athletes becoming coaches, or doctors becoming hospital leaders.
+However, there is little evidence on how specialized training vs.
+general leadership training affects firm behavior.
+In this paper, I investigate empirically the effect of clinically trained executives on hospital behavior.
+In particular, I leverage policies changing quality incentives for hospitals in the US in 2012, and estimate the difference in readmission and mortality rate responses for nonprofit hospitals with and without clinically trained executives.
+I find that a lack of clinical executives leads hospitals to improve quality more drastically after the incentive change.
+Further, I provide evidence that the more drastic response is due to non-clinical executives being more profit driven, and that clinical executives are not just a signal of underlying hospital objectives, but manage the hospital differently.
 		} 
@@
 	\newpage
+
+    There are many industries where some leaders have highly specialized training specific to that industry.
+For example, lawyers can be promoted to judges, athletes can be promoted to coaches, or doctors can be promoted to hospital leaders.
+However, there is little evidence on how specialized vs.
+general training in leadership affects firm behaviors.
+In this paper, I investigate how having clinically trained executives in leadership affects hospital behavior.
 
@@
 
-  Using publicly available tax forms, I construct a novel data set of NFP hospital executives from 2009-2015.
+  Using publicly available tax forms, I construct a novel data set of NFP hospital executives from 2009-2015, identifying which executives have clinical training.
 I link hospitals in this data to characteristics in the American Hospital Association (AHA) survey and Hospital Compare data.
@@
 I therefore rely on matching by location and name of hospital in both the tax forms and AHA data.
-I confidently match 1200 EINs to an AHA identification number based on exact name matches within the same state.
+In the AHA data, I limit to general hospitals.
+Then, I confidently match 1200 EINs to an AHA identification number based on exact name matches within the same state.
 I assess the observable differences between matched and non-matched hospitals in Appendix \ref{app:matched}, and conclude that the samples are overall similar apart from an under-sampling of hospitals that belong to systems.
@@
     In Table \ref{tab:sumstats_samples}, I present a table of means of relevant variables for several hospital samples used throughout the analysis.
-First, there are 708 for-profit hospitals in the sample, shown in the left-most column.
-Notably,
-
-    There are several samples of hospitals that are relevant for the analysis.
-First, there are all for-profits and all not-for-profits.
-Further, I subset the sample of NFPs into whether they have a clinical executive at some point in the sample.
-Finally, I characterize a NFP leadership team as being stable or not.
-A stable team means the hospital did not change their propensity to hire a clinical executive over time.
-I present a table of means for each of these sub-samples in Table \ref{tab:sumstats_samples}.
-Approximately 850 NFPs have leadership information, with 384 of them hiring an MD executive at some point in the sample.
-I compare these NFPs with the 708 for-profits from the AHA Survey.
-For-profits and NFPs are roughly the same bed size, but NFPs see more patients in each of the three relevant conditions.
-For-profits are more likely to be penalized, but also more likely to receive HVBP incentives (potentially for improvement in quality).
-For-profits have higher readmission rates, and higher mortality rates in all categories except heart failure.
-There are also key differences in not-for-profits with and without clinically trained executives.
-Those with MDs are larger, more likely to be penalized, have higher uncompensated care, and more complex patients.
-This holds for stable leadership teams as well, though the number of NFPs with a clinical executive drops significantly.
-I discuss the implications of these differences for identification in Section \ref{sec:model}.
+First, there are 708 for-profit hospitals in the sample, shown in column (1).
+For-profit hospitals in the sample have 155 beds on average, 112-276 patients depending on the condition, and are highly likely to be affiliated with a system.
+Thirty six percent of for-profits were penalized for all three conditions at some point.
 
     \import{Tables}{sample_sumstats.tex}
+
+    Columns 2-4 show variable means for not-for-profits in the sample, with all NFPs in column 2, NFPs that ever have a clinically trained executive in column 3, and NFPs that never have a clinically trained executive in column 4.
+NFPs are much less likely to be physician owned or system affiliated compared to FPs.
+They do, however, have similar bed size and case mix index to FPs.
+Twenty-two percent of all NFPs are penalized for all conditions.
+There are several differences among NFPs with and without clinically trained executives.
+Those with a clinically trained executive are larger in terms of beds and patients seen, are more likely to be an academic medical center, and see slightly more complex patients.
+Further, they are more likely to be penalized for every condition.
+
+    Finally, columns 5-6 limit the sample of not-for-profits further using changes in executive teams.
+The NFPs included in this sample do not change their propensity to hire a clinically trained executive, they either always have the same number of clinically trained executives or always have no clinically trained executives.
+The differences in NFPs with and without clinically trained executives remains similar with this sample limitation.
+I discuss the implications of these differences for identification in Section \ref{sec:model}.
 
@@
     I model hospital behavior under two simplified worlds: one where quality does not directly affect profit, and one where it does.
-Pay-for-performance incentives essentially incorporate performance directly into a hospital's profit function.
+Pay-for-performance incentives essentially incorporate performance directly into a hospital's profit function (\cite{dranove2011health}).
 I directly model performance as quality of care, which is inversely related to readmission and mortality rates.
@@
 societal benefit.
-There is an implicit stay open condition throughout the model.
+There is an implicit stay-open condition throughout the model.
 Taking the first order condition yields
@@
 I present here a two-way fixed effects specification, where the key variable is an interaction between a certain hospital type and time periods after program enactments.
-Ultimately, I employ synthetic difference-in-differences, which is estimating the same model with weights that more heavily factor similar trending control groups and time periods.
+Ultimately, I employ synthetic difference-in-differences, which is effectively estimating the same model with weights that more heavily factor similar trending control groups and time periods.
 However, for clarity, I present the unweighted specification:
@@
 Specifically, I define treat$_{ht}$ and limit the sample in four ways, shown in Figure \ref{fig:spec}.
-
-    An important distinction is whether the classification of NFPs by leadership team is time-varying.
-For overall ownership status, switching is rare, and I limit to those who never switch.
-However, hospital executive teams are dynamic, and whether or not a hospital employs a clinically trained executive in each period may change.
-Therefore, I first present effects of having a clinically trained executive among stable executive teams, i.e., NFPs with no change in their propensity to hire a clinically trained executive.
 

### 2024-07-29 Paper/paper2.tex

@@
 
-    Various characteristics of firm leadership such as gender, age, and background have been shown to be correlated with firm decisions and performance.
+    Across firm types and industries, firm leader characteristics such as gender, age, and background have been shown to be correlated with behavior and performance.
 I investigate this topic in the context of hospital executives with differing backgrounds: some with training as a medical doctor and some without.
-Hiring physician executives can be controversial.
-On one hand, physicians bring a unique combination of clinical and administration expertise, and they have insights that can potentially improve hospital outcomes (\cite{Stajduhar_2023}, \cite{Ahmed_2022}).
-Alternatively, physicians may not always be adequately trained to enter into high level management positions, and therefore worsen hospital outcomes (\cite{HarvardBusinessReview2018}).
-In this paper, I investigate how having clinically trained executives in leadership affects hospital response to pay-for-performance incentives.
+Physicians bring a unique combination of clinical and administration expertise, and they have insights that can potentially improve hospital outcomes (\cite{Stajduhar_2023}, \cite{Ahmed_2022}).
+However, hiring physician executives can also be controversial, as physicians may not always be adequately trained to enter into high level management positions (\cite{HarvardBusinessReview2018}).
 
+    In this paper, I investigate how having clinically trained executives in leadership affects quality of care in hospitals.
 I scrape executive team names and titles from publicly available Tax Form 990s, filled out by nonprofit firms in the US.
@@
 
-    The selection of leaders makes it difficult to interpret any direct comparison of differing executive teams as causal.
+    The non-random selection of leaders makes it difficult to interpret any direct comparison of differing executive teams as causal.
 Therefore, I leverage an exogenous shock to hospital incentives, pay-for-performance policies enacted in the US in 2012, and compare the change in readmission and mortality rates of executive teams with and without clinical executives.
 Under the assumption that the two types of teams would have had similar trends in readmission and mortality rates absent the incentive change, I find that a lack of clinically trained executives leads a hospital to respond more drastically to pay-for-performance incentives.
+That is, while both types of hospital leadership teams decrease readmission rates after the policy change, those without clinical executives decrease readmissions at a faster rate than those with clinical executives.
 
     Having found that hospital leadership does, in fact, matter to hospital behavior, I investigate several mechanisms that might contribute to this difference.
-First, \textcolor{red}{the more drastic decrease in readmissions by nonclinical teams is not driven by decreasing average patient complexity.
-There is no difference in patient case mix between clinical and nonclinical executive teams}.
-Further, I comparing each type of leadership team with for-profit hospitals reveals that the absence of clinical executives leads a nonprofit hospital to respond more similarly to a for-profit hospital.
+First, the more drastic decrease in readmissions by nonclinical teams is not driven by decreasing average patient complexity.
+Further, comparing each type of leadership team with for-profit hospitals reveals that the absence of clinical executives leads a nonprofit hospital to respond more similarly to a for-profit hospital.
 That is, having clinical training on a leadership team leads a hospital to act less profit-driven than their general training counterparts.
@@
     Due to data limitations, many of these studies focus on large, publicly traded firms in the US.
-\citeauthor{brickley2010board} (\citeyear{brickley2010board}) has the only study to my knowledge answering this question in the US nonprofit context.
+\citeauthor{brickley2010board} (\citeyear{brickley2010board}) have the only study to my knowledge answering this question in the US nonprofit context.
 Using exogenous variation in expected Medicare profits, the authors find that having an internal board of directors increases CEO compensation, and having physician board members decreases public donations (\cite{brickley2010board}).
 Additionally, two studies focus on hospital performance in other countries.
-\citeauthor{janke2019impact} (\citeyear{janke2019impact}) uses data from England to study whether CEOs affect hospital production, and find no association (\cite{janke2019impact}).
-However, \citeauthor{otero2022managers} (\citeyear{otero2022managers}) investigates the role of CEOs in public hospitals in Chile, and finds an 8\% decrease in mortality rates for hospitals with top managers.
+\citeauthor{janke2019impact} (\citeyear{janke2019impact}) use data from England to study whether CEOs affect hospital production, and find no association (\cite{janke2019impact}).
+However, \citeauthor{otero2022managers} (\citeyear{otero2022managers}) investigate the role of CEOs in public hospitals in Chile, and finds an 8\% decrease in mortality rates for hospitals with top managers.
 Further, a major contributor of this quality improvement is the exodus of older physicians as CEOs (\cite{otero2022managers}).
@@
 For anyone who has a doctor title and no matches, no doctor title and matches, or a doctor title and multiple matches, I use information on location and hospital name to manually match the NPI.
+I present a table of means for various characteristics of doctor executives in the data in Table \ref{doc_sumstats}.
+In total, there are 813 doctor executives.
+On average, they are 52 years old, and 11\% of them are female.
+While it is more likely they are a Chief Medical Officer than a Chief Executive Officer, still less than a third are CMOs.
+The most common specialty is internal medicine.
+
+    \import{Tables}{doc_exec_stats_tab.tex}
 
@@
 For my main sample, I limit to hospitals who do not change their propensity to hire a clinical executive for the duration of the sample.
-Summary statistics for these variables are shown in Table \ref{leader_sumstats}.
-On average, NFP hospitals in my sample have 4 executives and .4 clinically trained executives.
-Sixty-one percent of the sample never changes propensity to hire a clinically trained executive, and 33\% of NFPs in the sample have a Chief Medical Officer at some point.
-
-     \import{Tables}{leader_sumstats.tex}
 
@@
 
-    In Table \ref{tab:sumstats_samples}, I present a table of means of relevant variables for several hospital samples used throughout the analysis.
-First, there are 708 for-profit hospitals in the sample, shown in column (1).
-For-profit hospitals in the sample have 155 beds on average, 112-276 patients depending on the condition, and are highly likely to be affiliated with a system.
-Thirty six percent of for-profits were penalized for all three conditions at some point.
+    In Table \ref{tab:sumstats_samples_stable}, I present a table of means of relevant variables for several hospital samples used throughout the analysis.
+First, in column (1), are means for hospitals who ever have a clinical executive at any point in the sample.
+In column (2) are means for hospitals who have a clinical executive for the entire sample period, and column (3) shows means for hospitals who never have a clinical executive in the sample period.
 
-    \import{Tables}{sample_sumstats.tex}
+    \import{Tables}{stable_sample_sumstats.tex}
 
-    Columns 2-4 show variable means for not-for-profits in the sample, with all NFPs in column 2, NFPs that ever have a clinically trained executive in column 3, and NFPs that never have a clinically trained executive in column 4.
-NFPs are much less likely to be physician owned or system affiliated compared to FPs.
-They do, however, have similar bed size and case mix index to FPs.
-Twenty-two percent of all NFPs are penalized for all conditions.
-There are several differences among NFPs with and without clinically trained executives.
-Those with a clinically trained executive are larger in terms of beds and patients seen, are more likely to be an academic medical center, and see slightly more complex patients.
-Further, they are more likely to be penalized for every condition.
+    Hospitals that never have a clinical executive are smaller, less likely to be an academic medical center, and less likely to be affiliated with a system.
+Further, they have smaller executive teams on average by 1-2 people, and are less likely to have anyone acting as a CMO.\footnote{Though, they may title the position differently.} Finally, in regards to exposure to pay-for-performance programs, hospitals who never have clinical executives are less likely to receive benefits from the HVBP Program, but seem more similar to hospitals with clinical leadership in terms of penalties from HRRP.
 
-    Finally, columns 5-6 limit the sample of not-for-profits further using changes in executive teams.
-The NFPs included in this sample do not change their propensity to hire a clinically trained executive, they either always have the same number of clinically trained executives or always have no clinically trained executives.
-The differences in NFPs with and without clinically trained executives remains similar with this sample limitation.
-I discuss the implications of these differences for identification in Section \ref{sec:model}.
-
-    I plot weighted average readmission and mortality rates across time for different hospital types in Figure \ref{fig:weighted_read_mort_graph}, where I adjust both rates by baseline patient complexity.
-Pay-for-performance incentive changes are shown as a dotted line between 2011 and 2012.
-For-profit hospitals have the highest readmission rates prior to 2012.
-While the readmission rate trend behaves similarly across hospital types prior to the programs, there are notable trend differences after the programs.
-For-profits decrease readmissions drastically.
-NFP trends depend on the presence of an executive MD, where those without a clinically trained executive respond drastically, more similarly to for-profits, than those with a clinically trained executive.
-In mortality rates, NFPs with a clinically trained executive have the lowest rates before the programs, but differentially continue with an increasing trend in mortality after the programs.
-While this is purely descriptive evidence, the trends in outcomes point to more similar behavior between for-profits and NFPs without a clinically trained executive in comparison to NFPs with a clinically trained executive.
+    Next, I plot weighted average readmission, mortality rates, and patient case mix index across time for different hospital types in Figure \ref{fig:outcomes_graph}.
+Pay-for-performance incentive changes are shown as a dotted line in 2012.
+Average readmission rates trend similarly prior to 2012, and all hospital type exhibit a decrease in readmission after 2012.
+However, hospitals who consistently have a clinical executive show a slower rate of decrease in raw readmissions.
+When considering mortality, there is not a drastic change in 2012 for any hospital type, but hospitals who always have a clinical executive continue on an increasing trend in mortality where the other hospitals types remain more flat.
+Finally, there do not seem to be any differential trends in case mix index prior to or after 2012.
 
@@
         \caption{Outcomes Over Time}
-        \includegraphics[width=\textwidth]{Objects/weighted_read_mort_adjusted_graph.pdf}
-        \label{fig:weighted_read_mort_graph}
+        \includegraphics[width=.8\textwidth]{Objects/outcomes_graph.pdf}
+        \label{fig:outcomes_graph}
     \end{figure}
@@
 
-    My first goal is to estimate the effect of having clinical executive training on how hospitals change readmission and mortality rates in response to pay-for-performance incentives.
-Ultimately, I employ synthetic difference-in-differences, weighting more heavily non-clinical hospitals with readmission, mortality, and baseline case mix index that better match the clinical team trends.
-However, for clarity, I present the unweighted specification:
+    First, I test empirically whether having clinical experience matters in how hospitals respond to the pay-for-performance policies effective in 2012.
+For a clear comparison, I limit the sample to hospitals who always have a clinical executive or never have a clinical executive.
+That is, these hospitals do not change their propensity to hire a clinical executive during the sample period.
+Therefore, an important assumption that I make is that this sample selection is not correlated with the program enactment.
+I investigate this assumption in Section \ref{sec:identification}, and find no evidence that hospitals are more or less likely to change their propensity to hire a clinical executive because of the pay-for-performance programs.
+
+    I estimate a two way fixed effects specification, where the key variable is the interaction between having clinical executives and post 2012, when the programs began.
+Importantly, I am not estimating the effect of being penalized.
+All hospitals had incentive to improve quality as they are being compared to other hospitals who are also decreasing quality.
+Thus, I simply estimate the effect of being in a time period after the incentive changed for all hospitals.
+Ultimately, because there are differences in hospitals with different leadership types, I employ a synthetic difference-in-differences estimation strategy (\cite{arkhangelsky2021synthetic}).
+That is, hospitals with similar trends in readmission, mortality, and baseline case mix before the programs are more heavily weighted, and time periods that balance pre and post-program outcomes for hospitals without clinical executives.
+Below is the unweighted specification for clarity, where the identification assumptions are conceptually the same for the two specifications:
 
@@
 
-     I rely on several assumptions to identify this effect.
-First, I assume that, absent the program enactments, hospitals with different leadership team characterization types would have behaved similarly in outcomes.
-While I see no evidence of pre-trends, Table \ref{tab:sumstats_samples} suggests that hospitals with a clinically trained executive are different than other hospitals along multiple characteristics.
-There may be concern that these differences between hospital types could be driving the results.
-Therefore, my main specification uses the synthetic difference-in-differences method established in \citeauthor{arkhangelsky2021synthetic} (\citeyear{arkhangelsky2021synthetic}).
-This method creates unit weights that more heavily factor comparison hospitals that are similar to the treat group hospitals prior to the programs.
-It also weighs time periods that balance pre-programs and post-programs for the comparison hospitals.
-Then, a simple two way fixed effects estimator is used with these weights, yielding a more robust average treatment effect that relies less on a strong parallel trends assumption (\cite{arkhangelsky2021synthetic}).
-I present the synthetic diff-in-diff estimates as the main results, and they are largely similar to the standard two way fixed effects results with no weights, shown in Appendix \ref{app:fullsample}.
-Further, the conceptual identification is the same as a standard difference-in-differences.
+    In the unweighted two way fixed effects specification, we have reason to believe that the parallel trends assumption may not hold due to observed differences in characteristics other than the leadership team, as discussed in Section \ref{sec:data}.
+Once incorporating the weights used in the synthetic differences in differences method, we ensure parallel trends prior to 2012.
+Thus, I assume that under the given clinical and nonclinical hospital composition, trends would have continued to be parallel absent the incentive change in 2012.
+A component of this assumption is the implicit assumption that no other event occurred in 2012 correlated with both the outcomes and composition of leadership team.
+\textcolor{red}{say something about ACA here?
+and need to add anticipation assumption}
 
-    \subsection{Results}
-
-    The estimated difference in readmission rate trends for all hospital types in response to the programs are shown in Figure \ref{fig:clinicalsynthdid}.
-In panel \ref{fig:read_synth_clinical}, I compare readmission rate responses among nonprofit hospitals with and without clinical training on their executive team.
+    The estimated difference in readmission and mortality rate trends between hospitals with and without clinical executives are shown in Figure \ref{fig:clinicalsynthdid}.
 Estimate $\beta$ is represented in the difference in slope after 2012 when pay-for-performance incentives began.
-Nonprofits without a clinically trained executive lowered readmission rates by .24 ppts more than those with a clinically trained executive.
-
-    I also directly estimate the difference in mortality rates among NFPs with and without clinical experience in panel \ref{fig:mort_synth_clinical}.
-Again, the difference is not statistically significant, but the coefficient compared to the true nulls in panels \ref{fig:mort_synth_clinical} and \ref{fig:mort_synth_plotc} indicates NFPs without a clinical executive potentially decrease mortality more aggressively than NFPs with a clinical executive.
-In Appendix \ref{app:condition}, I show that these results are mainly driven by pneumonia patients.
+Panel \ref{read_synth_clinical} shows the estimated difference in readmission rates, and reveals that nonclinical teams decrease readmissions by .3 more than clinical teams.
+This is not a huge magnitude given an average readmission rate of 21, but reveals a meaningful difference in behavior given that both hospital types significantly decrease readmissions.
+In panel \ref{fig:mort_synth_clinical}, I show the estimated difference in mortality.
+While the magnitude of the difference is similar to that for readmission rates, mortality is a noisy measure, and I cannot rule out no difference in mortality response.
+In Appendix \ref{app:condition}, I show that these results are primarily driven by pneumonia patients.
 
@@
 
-    \section{Are Clinical Executives less Profit Driven?}\label{sec:forprofit}
+    \section{Selective Patient Practices}
 
-    A reason for the difference in hospital behavior with the presence of clinically trained executives could be how profit-driven a physician vs.
-a general leader is.
+    A method hospitals have been shown to employ in order to decrease readmission rates is being more selective with which patients to admit to the hospital.
+One hypothesis is that executive teams without any clinicians may be more willing to participate in such practices, whereas doctors prioritize care for all patients.
+This would explain why non-clinical hospitals are able to decrease readmissions more drastically.
+
+    To test this, I employ case mix index as the outcome in specification \ref{eq:clinical}, to determine whether one type of hospital differentially changes patient composition after pay-for-performance incentives, where the same identification assumptions hold.
+I present the estimated difference in Figure \ref{fig:main_cmi_clinical}, which shows no differential response to the programs in terms of the complexity of patients admitted to the hospital.
+\textcolor{red}{sentence about the implications of this?}
+
+    \begin{figure}[ht!]
+        \centering
+        \caption{Effect of Clinical Executive on Cases Mix Index}
+        \includegraphics[width=0.5\textwidth]{Objects/cmi_md_nomd_synth_graph.pdf}
+        \label{fig:main_cmi_clinical}
+    \end{figure}
+
+    \section{Does Specialty Matter?}
+
+    \section{Are Clinical Executives Less Profit Driven?}\label{sec:forprofit}
+
+    Another hypothesis that could explain the differential response due to clinical leadership is how profit-driven a clinician is compared to other training backgrounds such as an MBA.
 To demonstrate this, consider hospital behavior under the enactment of pay-for-performance policies, where hospitals face a trade-off between profit and societal benefit.
@@
 
-    I compare readmission and mortality rates of each nonprofit leadership team type to for-profit hospitals.
-Since for-profit hospitals have revealed through their ownership status that they are largely profit-driven, comparing different executive team behavior to for-profit behavior helps reveal how profit driven certain leadership teams are.
+    I compare readmission and mortality rates of always clinical executives and never clinical executives to for-profit hospitals.
+While ideally I would also have leadership information for for-profit hospitals, the tax forms only have nonprofit hospitals.
+However, on average, for-profit hospitals have revealed through their ownership status that they heavily weigh profit.
+Thus, comparing different executive team behavior to for-profit behavior helps reveal how profit driven certain leadership teams are.
 Thus, I estimate
@@
 Thus, $\beta$ represents the effect of being for-profit compared to being nonprofit with a certain leadership team type.
+I show summary statistics of relevant variables for for-profit hospitals in Appendix \textcolor{red}{add table}.
+Again, I employ a synthetic difference in differences strategy to make the parallel trends assumption more plausible.
 
     The identification assumptions I make here are largely similar to those in the main analysis, only now incorporate for-profit hospitals.
-First, I assume that, absent the change in incentives, for-profit and the relevant comparison group would have had parallel trends in outcomes.
-While this is not directly testable, both the synthetic difference in difference and event studies indicate parallel trends prior to the incentive changes.
+First, I assume that, absent the change in incentives, for-profits and the relevant comparison group would have had parallel trends in outcomes.
 Second, I assume that the stability of nonprofit leadership teams is not correlated with the 2012 policy changes.
-I explore this assumption in Section \ref{sec:identification}.
+I explore this assumption in Section \ref{sec:sig_man}.
 I also assume that there is no anticipation of the policy changes prior to 2012, which is reasonable as the rules of the policies were announced in October of 2011.
@@
  \begin{figure}[ht!]
-     \caption{Readmission Rate Synthetic Difference in Differences Results}
+     \caption{Comparison to For-Profit: Readmission}
      \centering
@@
          \centering
-         \caption{For-Profit and NFP w/ MD Exec}
-         \includegraphics[width=\textwidth]{Objects/read_fp_md_synth_graph.pdf}
+         \caption{For-Profit and Always Clinical Exec}
+         \includegraphics[width=\textwidth]{Objects/fp_read_md_synth_graph.pdf}
          \label{fig:read_synth_plotb}
@@
          \centering
-         \caption{For-Profit and NFP w/out MD Exec}
-         \includegraphics[width=\textwidth]{Objects/read_fp_nomd_synth_graph.pdf}
+         \caption{For-Profit and Never Clinical Exec}
+         \includegraphics[width=\textwidth]{Objects/fp_read_nomd_synth_graph.pdf}
          \label{fig:read_synth_plotc}
@@
 In panel \ref{fig:mort_synth_plotb}, the estimated difference is between for-profits and nonprofits with clinically trained executives.
-The magnitude of the difference in mortality rates is .11, where only for-profits seem to be reacting to the incentive change.
-While still statistically insignificant, I argue that this still indicates a difference in behavior between the hospital types, particularly due to the documented noisiness of mortality measures (\cite{mackenzie2016measuring}).
+The magnitude of the difference in mortality rates is .08, where clinical team hospitals continue on an increasing trend and for-profits remain stable in mortality rate.
+While I cannot rule out a lack of differential response, the documented noisiness of mortality measures (\cite{mackenzie2016measuring}) may mean that the true effect is different from zero.
 However, the estimated difference in mortality between for-profits and nonprofits without clinically trained executives is certainly zero.
@@
          \caption{For-Profit and NFP w/ MD Exec}
-         \includegraphics[width=\textwidth]{Objects/mort_fp_md_synth_graph.pdf}
+         \includegraphics[width=\textwidth]{Objects/fp_mort_md_synth_graph.pdf}
          \label{fig:mort_synth_plotb}
@@
          \caption{For-Profit and NFP w/out MD Exec}
-         \includegraphics[width=\textwidth]{Objects/mort_fp_nomd_synth_graph.pdf}
+         \includegraphics[width=\textwidth]{Objects/fp_mort_nomd_synth_graph.pdf}
          \label{fig:mort_synth_plotc}
@@
 
-\section{Signaling vs. Managing Decomposition}
+\section{Signaling vs. Managing Decomposition}\label{sec:sig_man}
 
-    To this point, I have limited the sample of nonprofit hospitals to those that don't change their propensity to hire a clinically trained executive over the sample period.
-However, under the assumption that hiring or firing these types of executives is not correlated with the pay-for-performance policy changes, these changes can be leveraged to understand further how executive characteristics affect hospital behavior.
-Particularly, a clinically trained executive could be a signal of the underlying objectives of the hospital, or hiring a clinically trained executive who carries out day-to-day operations in a distinct way effectively changes the objectives of the hospital.
+    Given that having clinical leaders affects behavior, an important distinction is whether the existing clinical leaders \textit{reveal} the underlying objectives of the hospital, or \textit{manage} the hospital differently.
+Up to this point, I have limited the sample of nonprofit hospitals to those that don't change their propensity to hire a clinically trained executive over the sample period, which combines signalling and managing.
+However, under the assumption that hiring or firing these types of executives is not correlated with the pay-for-performance policy changes, such changes can be leveraged to decompose the effect into signaling vs.
+managing.
 
@@
 
+    \subsection{Are Leadership Changes Correlated with Pay-for-Performance Policies?} \label{sec:changes}
+
     I have discussed the assumptions of parallel trends, no anticipation, and no other confounding events in both Sections \ref{sec:clinical} and \ref{sec:forprofit}.
 These assumptions are also necessary here.
-In those Sections, I have discussed that stability of leadership team must not be correlated with the 2012 policy changes.
+Further, this decomposition analysis depends heavily on leadership team changes not endogenously occurring due to pay-for-performance policies.
 Up to this point, that assumption was important because of the way I selected the sample.
@@
 Thus, I now explore this assumption in detail.
+First, I present raw means over time that represent the fraction of hospitals who hire a clinical executive, fire a clinical executive, or have any change in their propensity to have a clinical executive in Figure \ref{fig:change_means}.
+There do not seem to be any drastic changes in 2011 or 2012 when the policies were becoming relevant to hospitals.
 
-     I assess the validity of the exogenous changes assumption in two ways.
-First, I analyze whether changes in clinically trained executives occur more often in time periods of program enactment.
-Second, I analyze whether hospitals who end up getting penalized or receiving payments are more likely to make changes in clinical executive members than those who are not.
-That is, are hospitals that expect to be affected by the programs more likely to hire or fire clinically trained executives?
-I consider changes in having any clinically trained executive, as well as changes in the number of clinically trained executives.
-I estimate the following linear equations:
+    \begin{figure}[ht!]
+    \centering
+    \caption{Leadership Change Means Over Time}
+    \includegraphics[width=0.8\textwidth]{Objects/change_means.pdf}
+    \label{fig:change_means}
+     \end{figure}
 
-    \begin{equation}\label{eq:change1}
-    \text{change}_{ht} = \sum_{j=2011}^{2014}\beta_j\mathbf{1}\{t=j\} + \alpha_h + \epsilon_{ht},
+    Further, I assess whether this assumption is reasonable by analyzing how executive team changes are occurring over time for certain groups of hospitals. I estimate:
+
+    \begin{equation}\label{eq:change2}
+    \text{change}_{ht} = \sum_{j=2011}^{2014}\beta_j(\mathbf{1}\{t=j\}\times \text{Program Exposed})_{ht} + \alpha_h + \epsilon_{ht},
     \end{equation}
 
-    \begin{equation}\label{eq:change2}
-    \text{change}_{ht} = \beta(\text{Program Exposed}_{h} \times \text{Post Programs}_t)+ \alpha_h + \delta_t + \epsilon_{ht}.
-    \end{equation}
-
-    The variable $\text{Program Exposed}_{h}$ is an indicator for whether the hospital eventually was penalized under HRRP or received payments under HVBP.
-The estimates from this analysis are presented in Table \ref{tab:change_analysis}.
-Changes in any clinically trained executive are less likely to occur in 2012-2014 relative to 2010, indicating that executive teams are actually becoming more stable over time along the dimension of hiring clinically trained executives.
-Further, hospitals who end up being penalized through HRRP or receiving payments through HVBP are not more likely to change their propensity to hire any clinically trained executive.
-Results are similar when considering changes in the number of clinically trained executives.
+    where the variable $\text{Program Exposed}_{h}$ has three definitions.
+First, an indicator for whether the hospital was ever penalized under HRRP.
+Second, an indicator for whether the hospital ever received payments under HVBP.
+And finally, I leave out the Program Exposed variable to simply see how changes are correlated with given years for all hospitals.
+The outcome change$_{ht}$ is an indicator for whether the hospital changes the number of clinical executives in a given year.
+The estimates from this analysis are presented in Figure \ref{fig:change_analysis}.
+All estimates are not statistically different from zero, indicating that expectations of program exposure do not lead to changes in propensity to hire a clinical executive.
 Thus, it seems unlikely that endogenous team formation is biasing the estimates.
@@
 
-     \import{Tables}{change_analysis.tex}
+     \begin{figure}[ht!]
+         \centering
+         \caption{Leadership Analysis Results}
+         \includegraphics[width=0.8\textwidth]{Objects/change_analysis_plot.pdf}
+         \label{fig:change_analysis}
+     \end{figure}
+
+    \subsection{Decomposition Results}
 
     The estimates of each decomposition specification are shown in Table \ref{tab:MD_noMD_readmort_decomp_synth}.
-Columns (1) and (2) show that the difference in readmission response is driven by having a clinically trained executive in 2012, compared to having a clinically trained executive at another point in time.
-Columns (3) and (4) support the same theory given the magnitudes of the coefficients, but are still statistically insignificant due to the noisiness of mortality.
-The results suggest that managers make a difference in behavior and objectives, rather than being a signal of already existing objectives.
+The decomposition for readmission rates is in Columns (1) and (2), and for mortality is in Columns (3) and (4).
+The first row compares ever clinical executive hospitals to never clinical executive hospitals, and the second row compares hospitals with a clinical executive in 2012 to hospitals who have a clinical executive in a year other than 2012.
+While there is no statistically significant estimate, the coefficients of the managing effect are very similar to the full effect found in Section \ref{sec:clinical}.
+The coefficients for the signaling effect are both statistically insignificant and close to zero in magnitude.
+Thus, I can rule out that clinical executives are simply a signal of underlying hospital objectives.
 
@@
 
-    Existing literature is inconclusive as to whether not-for-profit (NFP) hospitals differ meaningfully from for-profit hospitals.
-NFP objectives have direct implications for patients, yet are largely unknown.
-An observable and economically meaningful characteristic of NFP hospitals that may affect underlying objectives is the composition of the hospital's top management.
-Particularly, whether any executives at the hospital have a background in practicing medicine.
+    While previous literature has documented correlations between executive characteristics and firm behaviors/performance, research is still limited on this topic in the health care sector.
+Particularly, does having executive leaders with clinical training trickle down into hospital behaviors/outcomes?
+In this paper, I investigate the differential response of hospital executive teams to a policy change that financially incentivizes quality.
 
-    In this paper, I study differential responses to a change in incentives between for-profit and NFP hospitals with and without a clinical executive.
-I do this by collecting a novel data set on hospital executives from Tax Form 990s and merging this to existing hospital data.
-I then use a synthetic difference-in-differences methodology to estimate differential responses in readmission and mortality rates after several pay-for-performance initiatives are enacted in the US in 2012.
+    I collect data on the names of hospital executives from 2010-2014 from publicly available Tax Form 990s, which are completed by all nonprofits in the US.
+I merge this data with existing hospital and physician characteristics, and aggregate to the hospital level.
+Thus, I end up with a data set of nonprofit US hospitals that captures the proportion of clinical executives, specialties of clinical executives, characteristics, and quality metric outcomes.
 
-     When comparing for-profits to all not-for-profits, I find no difference in behavior after the policies take effect, which aligns with suggestions that NFPs are for-profits in disguise.
-However, when I compare for-profits to NFPs with clinical executives, the two types of hospitals respond differently to the incentives.
-Specifically, all firm types decrease readmissions, but NFPs with a clinical executive are less aggressive in their responsiveness than the other firm types.
-A similar result holds for mortality rates.
-This finding is consistent with the theoretical prediction that firms who place more weight on societal benefit than on profit will have a more mitigated response to policies that affect profit.
-Additionally, under the assumption that executive team changes are not correlated with the program, I find that the differential response is driven by the actual management of the clinical executive, not by a clinical executive being a signal of underlying objectives.
+    In 2012, two pay-for-performance policies went into effect that either penalized hospitals with high readmission rates, or gave payments to hospitals with low readmission/mortality rates.
+These policies were relevant to hospital finances, and gave tangible incentive for hospitals to increase quality.
+While I don't observe and can't causally estimate why certain hospitals have clinical leadership and others don't, I can observe how clinical executives may respond to an exogenous shock differently than non-clinical executives conditional on teams being previously formed.
 
-     These findings are informative for policymakers predicting how hospitals will respond to a change in incentives.
-Specifically, they suggest that policymakers should take into account more than just ownership status when evaluating how hospitals will respond to policy changes.
+    First, I estimate the effect of always having clinical leadership vs.
+never having clinical leadership on readmission and mortality rates after 2012 using a synthetic difference in differences estimation strategy.
+Conditional on parallel trends holding, I find that having clinical experience mitigates the response to the incentive change.
+That is, while all hospitals reduce readmissions after the policies take place, non-clinical teams reduce readmissions at a fast rate than clinical teams.
+Next, I investigate several hypotheses regarding this differential response.
+This result is not driven by non-clinical teams adjusting their patient complexity.
+However, comparing nonclinical teams to for-profit hospitals reveals that it could be underlying preferences for profit vs.
+patient benefit that drive the difference.
+Finally, I show that changes in leadership teams are not correlated with the policies, and leverage such changes to show that clinical leaders manage the hospital differently and are not just a revealed preference of underlying objectives.
+
+    \textcolor{red}{discuss implications of these findings}
 

### 2024-08-01 Paper/paper2.tex
Updates from Overleaf

@@
 		{\small
-         Firm leaders are shown to be correlated with firm decisions and performance.
+         One determinant of firm behaviors and performance is the type of management leading the firm.
+Characteristics such as age, gender, and background are correlated with employee wages, decisions of mergers and acquisitions, and other firm outcomes.
 In this paper, I investigate this topic in the context of hospital executive teams.
-Specifically, I empirically the effect of clinically trained executives on hospital behavior in response to an incentive shock.
-In particular, I leverage policies changing quality incentives for hospitals in the US in 2012, and estimate the difference in readmission and mortality rate responses for nonprofit hospitals with and without clinically trained executives.
-I find that a lack of clinical executives leads hospitals to improve quality more drastically after the incentive change.
-Further, I provide evidence that the more drastic response is due to non-clinical executives being more profit driven, and that clinical executives are not just a signal of underlying hospital objectives, but manage the hospital differently.
+Specifically, I estimate the effect of clinically trained executives on hospital quality.
+I leverage pay-for-performance policies that changed quality incentives for hospitals in the US in 2012, and estimate the difference in readmission and mortality rates for nonprofit hospitals with and without clinically trained executives.
+I find that clinically trained executives do not improve quality as much as executive teams without clinically trained executives.
+This difference is not driven by a change in patient complexity, but may be driven by preferences towards profit vs.
+patient welfare.
+Further, clinically trained executives are not just a signal of underlying hospital incentives, but manage the hospital differently.
 		} 
@@
 
-    Across firm types and industries, firm leader characteristics such as gender, age, and background have been shown to be correlated with behavior and performance.
-I investigate this topic in the context of hospital executives with differing backgrounds: some with training as a medical doctor and some without.
-Physicians bring a unique combination of clinical and administration expertise, and they have insights that can potentially improve hospital outcomes (\cite{Stajduhar_2023}, \cite{Ahmed_2022}).
-However, hiring physician executives can also be controversial, as physicians may not always be adequately trained to enter into high level management positions (\cite{HarvardBusinessReview2018}).
+    Broadly, firm leader characteristics such as gender, age, and background have been shown to be correlated with financial performance and behaviors such as employee diversity, wages, aggressiveness in mergers and acquisitions, and other corporate strategies.
+I consider this topic in the context of hospital quality under executives with differing backgrounds: some with training as a medical doctor and some without.
+Physicians can bring a unique combination of clinical and administration expertise, and they have insights that can potentially benefit their employees and lead to better quality (\cite{Stajduhar_2023}, \cite{Ahmed_2022}).
+However, hiring physician executives can also be controversial, as physicians are not always adequately trained in business management, which can lead to poor business practices and lower quality (\cite{HarvardBusinessReview2018}).
 
-    In this paper, I investigate how having clinically trained executives in leadership affects quality of care in hospitals.
+    In this paper, I estimate how having clinically trained executives in leadership affects quality of care in hospitals.
 I scrape executive team names and titles from publicly available Tax Form 990s, filled out by nonprofit firms in the US.
 Combining this with publicly available hospital characteristics, I construct a novel data set of nonprofit hospitals from 2010-2014 containing information on both executive team characteristics and quality metrics.
-Nonprofit behavior in the hospital industry directly affects health care consumers, as private nonprofit hospitals make up 50\% of all hospitals in the US, and staff, on average, 207 beds.
+While limited to nonprofit hospitals due to data availability, this sample is a very important driver of health care in the US.
+Private nonprofit hospitals make up 50\% of all hospitals in the US, and staff, on average, 207 beds.
 For comparison, for-profits make up 36\% of hospitals and staff 107 beds (\cite{ASPE_2023}).
+Therefore, the behavior of these hospitals directly affects health care consumers in the US.
 
@@
 Therefore, I leverage an exogenous shock to hospital incentives, pay-for-performance policies enacted in the US in 2012, and compare the change in readmission and mortality rates of executive teams with and without clinical executives.
-Under the assumption that the two types of teams would have had similar trends in readmission and mortality rates absent the incentive change, I find that a lack of clinically trained executives leads a hospital to respond more drastically to pay-for-performance incentives.
-That is, while both types of hospital leadership teams decrease readmission rates after the policy change, those without clinical executives decrease readmissions at a faster rate than those with clinical executives.
+Under the assumption that the two types of teams would have had similar trends in readmission and mortality rates absent the incentive change, I find that executive teams with clinical executives on average do not improve quality as much as executive teams without clinical executives.
+While both types of hospital leadership teams decrease readmission rates after the policy change, those without clinical executives decrease readmissions at a faster rate than those with clinical executives.
 
     Having found that hospital leadership does, in fact, matter to hospital behavior, I investigate several mechanisms that might contribute to this difference.
-First, the more drastic decrease in readmissions by nonclinical teams is not driven by decreasing average patient complexity.
-Further, comparing each type of leadership team with for-profit hospitals reveals that the absence of clinical executives leads a nonprofit hospital to respond more similarly to a for-profit hospital.
+First, the more drastic decrease in readmissions by non-clinical teams is not driven by decreasing average patient complexity.
+Further, comparing each type of leadership team with for-profit hospitals reveals that the absence of clinical executives leads a nonprofit hospital to respond more similarly to a for-profit hospital than nonprofits with clinical executives.
 That is, having clinical training on a leadership team leads a hospital to act less profit-driven than their general training counterparts.
@@
 
-    Many researchers have documented a correlation between executive characteristics, specifically for CEOs, and firm performance, starting with \citeauthor{bertrand2003managing} (\citeyear{bertrand2003managing}), who show that manager fixed effects are an important driver of many firm behaviors and decisions.
+    Many researchers have documented a correlation between executive characteristics and firm performance, starting with \citeauthor{bertrand2003managing} (\citeyear{bertrand2003managing}), who show that manager fixed effects are an important driver of many firm behaviors and decisions.
 Female board members are associated with better oversight and more women executives (\cite{matsa2011chipping}; \cite{adams2009women}), but are negatively correlated with firm value (\cite{ahern2012changing}).
 Having more female executives is correlated with female employee wages and corporate strategies (\cite{flabbi2019female}; \cite{matsa2013female}); young male CEOs tend to be more aggressive in mergers and acquisitions, while those with military experience are less aggressive (\cite{levi2010deal}; \cite{benmelech2015military}); CEOs with general ability tend to receive higher pay and perform better (\cite{kaplan2012ceo}; \cite{custodio2013generalists}; \cite{adams2018director}; \cite{frydman2019rising}).
-Further, Chief Diversity Officers are not found to have any effect on hiring more diversely in universities (\cite{bradley2022impact}).
+Finally, Chief Diversity Officers are not found to have any effect on hiring more diversely in universities (\cite{bradley2022impact}).
 
     Due to data limitations, many of these studies focus on large, publicly traded firms in the US.
-\citeauthor{brickley2010board} (\citeyear{brickley2010board}) have the only study to my knowledge answering this question in the US nonprofit context.
+\citeauthor{brickley2010board} (\citeyear{brickley2010board}) have the only study to my knowledge in the US nonprofit context considering hospital board of directors.
 Using exogenous variation in expected Medicare profits, the authors find that having an internal board of directors increases CEO compensation, and having physician board members decreases public donations (\cite{brickley2010board}).
@@
 Further, a major contributor of this quality improvement is the exodus of older physicians as CEOs (\cite{otero2022managers}).
-I contribute to our knowledge of how executive team characteristics affect firm behaviors by (1) considering the important setting of nonprofit hospitals, which serve the majority of patients in the US, (2) considering a unique executive characteristic which is plausibly relevant, and (3) observing firm behavior in the presence of an exogenous incentive change.
+I contribute to our knowledge of how executive team characteristics affect firm behaviors by (1) considering the important setting of nonprofit hospitals, which serve the majority of patients in the US, (2) considering a unique and relevant executive characteristic, and (3) observing firm behavior in the presence of an exogenous incentive change.
 
@@
 An executive team usually consists of at least a Chief Executive Officer (CEO) and Chief Financial Officer (CFO), but there is variation in how firms organize executive teams.
-Some hospital executives to specialize in health care settings by earning a degree in health care management, or an MBA specific to health care.
+Some hospital executives specialize in health care settings by earning a degree in health care management, or an MBA specific to health care.
 There are executive positions that are often filled by someone with clinical training, such as a Chief Medical Officer (CMO), but medical doctors can pursue other top managerial positions as well such as CEO or president.
@@
     Hospital executive teams are understudied partly due to the inaccessibility of granular data.
-I construct a novel data set of not-for-profit hospitals in the US that contains names, titles, and positions of all board members and executives tied to the hospital from 2009-2015.
+I construct a novel data set of nonprofit hospitals in the US which contains names, titles, and positions of all board members and executives tied to the hospital from 2009-2015.
 I gather this data from publicly available Tax Form 990s, which every sufficiently large nonprofit files each year in the US.
@@
 
-    The historical tax forms for all not-for-profits are housed by ProPublica.\footnote{https://projects.propublica.org/not-for-profits/} I use the NonProfit Explorer API to extract Employee Identification Numbers (EIN) of NFPs with a hospital National Taxonomy of Exempt Entities (NTEE) code.
-After filtering out associations and specialty hospitals, there are approximately 3,000 NFP EINs.
-Linking these EINs to other hospital information is crucial for the analysis.
-I therefore rely on matching by location and name of hospital in both the tax forms and AHA data.
+    The historical tax forms are housed by ProPublica.\footnote{https://projects.propublica.org/not-for-profits/} I use the NonProfit Explorer API to extract Employee Identification Numbers (EIN) of firms with a hospital National Taxonomy of Exempt Entities (NTEE) code.
+After filtering out associations and specialty hospitals, there are approximately 3,000 EINs.
+Linking these EINs to other hospital information is crucial to the analysis, but no crosswalk from EIN to other hospital identifiers is publicly available.
+I therefore rely on matching by location and name of hospital in both the tax forms and American Hospital Association (AHA) data.
 In the AHA data, I limit to general hospitals.
-Then, I confidently match 1200 EINs to an AHA identification number based on exact name matches within the same state.
-I assess the observable differences between matched and non-matched hospitals in Appendix \ref{app:matched}, and conclude that the samples are overall similar apart from an under-sampling of hospitals that belong to systems.
+Then, I confidently match 1200 EINs to an AHA identification number based on exact name matches within the same state.\footnote{I assess the observable differences between matched and non-matched hospitals in Appendix \ref{app:matched}, and conclude that the samples are overall similar apart from an under-sampling of hospitals that belong to systems.}
 
@@
 As I limit the sample to hospitals who have text information in every year, I limit the years to 2010-2014 to retain more hospitals.
-This yields 852 not-for-profit hospitals that are matched to AHA data, and with complete text information in each year.
+This yields 852 nonprofit hospitals that are matched to AHA data, and with complete text information in each year.
 
-    Finally, I match each name with the National Plan and Provider Enumeration System (NPPES) database of all registered physicians.
-The goal is to match each physician with an NPI number in order to merge publicly available data on physician specialty, gender, and tenure.
-For any person who has a doctor title and uniquely matches the NPPES data, I record their NPI number.
-For anyone who has a doctor title and no matches, no doctor title and matches, or a doctor title and multiple matches, I use information on location and hospital name to manually match the NPI.
+    Finally, I seek to (1) verify clinical training of those who claim to be doctors, and (2) obtain additional information about clinically trained executives.
+Thus, I match each name with the National Plan and Provider Enumeration System (NPPES) database of all registered physicians.
+For any person who has a doctor title and uniquely matches the NPPES data, I record their National Provider Identifier (NPI) number.
+For anyone who has a doctor title and no matches, no doctor title and matches, or a doctor title and multiple matches, I use information on location and hospital name to manually match NPI numbers of those who did have clinical training.
 I present a table of means for various characteristics of doctor executives in the data in Table \ref{doc_sumstats}.
@@
     For the purposes of this paper, I focus on executive teams as a whole instead of specific positions.
-The main reason for this is because the organization of teams varies so drastically across hospitals that a specific position, even if it exists in two different hospitals, can be meaningfully different.
+The main reason for this is because the organization of teams varies so drastically across hospitals.
+It is not uncommon for two hospitals to have the same position but the tasks performed are drastically different, or to have two distinct positions that essentially perform the same tasks.
 I drop anyone who is solely a board member, and focus on those who have executive in their title, or are labeled as a president or vice president of a department.
 I create various hospital-level characterizations based on their executive team, such as the existence of a clinically trained executive, the number of clinically trained executives, the number of total executives, and whether the hospital has a Chief Medical Officer.
-For my main sample, I limit to hospitals who do not change their propensity to hire a clinical executive for the duration of the sample.
+For the main analysis, I limit to hospitals who either always have a clinical executive or never have a clinical executive over the course of the sample in order to have a clean comparison.
 
@@
 
-    Therefore, I focus on outcomes of readmission and mortality rates, which are publicly available at the hospital-condition level in CMS Hospital Compare.
+    I focus on outcomes of readmission and mortality rates, which are publicly available at the hospital-condition level in CMS Hospital Compare.
 I combine rates for pneumonia, AMI, and heart failure (the relevant HRRP conditions) into a weighted average based on the number of patients in each condition.
@@
 
-    In Table \ref{tab:sumstats_samples_stable}, I present a table of means of relevant variables for several hospital samples used throughout the analysis.
+    The hospitals in my sample are limited to general nonprofit hospitals with executive team information in each year (2010-2014) and a link from EIN to AHA ID.
+In Table \ref{tab:sumstats_samples_stable}, I present means of relevant variables for several hospital samples used throughout the paper.
 First, in column (1), are means for hospitals who ever have a clinical executive at any point in the sample.
 In column (2) are means for hospitals who have a clinical executive for the entire sample period, and column (3) shows means for hospitals who never have a clinical executive in the sample period.
+The main analysis compares hospitals in columns (2) and (3), and supplemental analyses leverage hospitals in column (1).
 
@@
     Next, I plot weighted average readmission, mortality rates, and patient case mix index across time for different hospital types in Figure \ref{fig:outcomes_graph}.
-Pay-for-performance incentive changes are shown as a dotted line in 2012.
+Pay-for-performance incentive change timing is shown as a dotted line in 2012.
 Average readmission rates trend similarly prior to 2012, and all hospital type exhibit a decrease in readmission after 2012.
@@
 When considering mortality, there is not a drastic change in 2012 for any hospital type, but hospitals who always have a clinical executive continue on an increasing trend in mortality where the other hospitals types remain more flat.
-Finally, there do not seem to be any differential trends in case mix index prior to or after 2012.
+Finally, hospitals that never have clinical executives consistently have less complex patients, but there is no notable change in raw case mix index over time.
 
@@
 
-    \section{Effect of Executive Clinical Training on Hospital Response}\label{sec:clinical}
+    \section{Effect of Executive Clinical Training on Hospital Quality}\label{sec:clinical}
 
-    First, I test empirically whether having clinical experience matters in how hospitals respond to the pay-for-performance policies effective in 2012.
-For a clear comparison, I limit the sample to hospitals who always have a clinical executive or never have a clinical executive.
+    First, I test empirically whether having clinical training matters in how hospitals change readmission and mortality rates in response to the pay-for-performance policies effective in 2012.
+For a clean comparison, I limit the sample to hospitals who always have clinical executives or never have clinical executives.
 That is, these hospitals do not change their propensity to hire a clinical executive during the sample period.
 Therefore, an important assumption that I make is that this sample selection is not correlated with the program enactment.
-I investigate this assumption in Section \ref{sec:identification}, and find no evidence that hospitals are more or less likely to change their propensity to hire a clinical executive because of the pay-for-performance programs.
+I investigate this assumption in Section \ref{sec:changes}, and find no evidence that hospitals are more or less likely to change their propensity to hire a clinical executive because of the pay-for-performance programs.
 
-    I estimate a two way fixed effects specification, where the key variable is the interaction between having clinical executives and post 2012, when the programs began.
+    I estimate a two-way-fixed-effects specification, where the key independent variable is the interaction between having clinically trained executives and post 2012, when the programs began.
 Importantly, I am not estimating the effect of being penalized.
-All hospitals had incentive to improve quality as they are being compared to other hospitals who are also decreasing quality.
-Thus, I simply estimate the effect of being in a time period after the incentive changed for all hospitals.
-Ultimately, because there are differences in hospitals with different leadership types, I employ a synthetic difference-in-differences estimation strategy (\cite{arkhangelsky2021synthetic}).
-That is, hospitals with similar trends in readmission, mortality, and baseline case mix before the programs are more heavily weighted, and time periods that balance pre and post-program outcomes for hospitals without clinical executives.
-Below is the unweighted specification for clarity, where the identification assumptions are conceptually the same for the two specifications:
+All hospitals had incentive to improve quality in response to the policy change, as penalties and payments are determined by comparing quality among hospitals.
+Thus, I leverage the timing of the policy change, rather than the policies themselves.
+However, because of potential differences in quality trends between clinical and non-clinical hospital executive teams even before the policies, I employ a synthetic difference-in-differences estimation strategy (\cite{arkhangelsky2021synthetic}).
+That is, hospitals in the control group with similar trends in readmission, mortality, and baseline case mix before the programs are more heavily weighted, as well as time periods that balance pre- and post-program outcomes for hospitals without clinical executives.
+Below is the unweighted specification for clarity, where the identification assumptions are conceptually the same for synthetic difference-in-differences:
 
@@
 
-    \noindent where $y_{ht}$ is one of the outcome variables discussed in Section \ref{sec:hrrp}, and $\gamma_h$ and $\delta_t$ are hospital and time fixed effects, respectively.
+    \noindent where $y_{ht}$ is one of the outcome variables capturing readmission or mortality rates discussed in Section \ref{sec:hrrp}, and $\gamma_h$ and $\delta_t$ are hospital and time fixed effects, respectively.
 
-    In the unweighted two way fixed effects specification, we have reason to believe that the parallel trends assumption may not hold due to observed differences in characteristics other than the leadership team, as discussed in Section \ref{sec:data}.
-Once incorporating the weights used in the synthetic differences in differences method, we ensure parallel trends prior to 2012.
-Thus, I assume that under the given clinical and nonclinical hospital composition, trends would have continued to be parallel absent the incentive change in 2012.
+    In the unweighted two-way-fixed-effects specification, we have reason to believe that the parallel trends assumption may not hold due to observed differences in characteristics other than the leadership team, as discussed in Section \ref{sec:data}.
+Once incorporating the weights used in the synthetic differences in differences method, we ensure parallel trends in the outcome variables prior to 2012.
+Thus, I assume that under the weighted clinical and non-clinical hospital composition, trends would have continued to be parallel absent the incentive change in 2012.
 A component of this assumption is the implicit assumption that no other event occurred in 2012 correlated with both the outcomes and composition of leadership team.
-\textcolor{red}{say something about ACA here?
-and need to add anticipation assumption}
+Finally, I assume that there is no anticipation of the policy change before 2012.
+The rules of the policies were announced in October of 2011, so 2012 was the first full year that hospitals could meaningfully respond to the policies.
 
-    The estimated difference in readmission and mortality rate trends between hospitals with and without clinical executives are shown in Figure \ref{fig:clinicalsynthdid}.
+    The estimated difference in readmission and mortality rates between hospitals with and without clinical executives are shown in Figure \ref{fig:clinicalsynthdid}.
 Estimate $\beta$ is represented in the difference in slope after 2012 when pay-for-performance incentives began.
-Panel \ref{read_synth_clinical} shows the estimated difference in readmission rates, and reveals that nonclinical teams decrease readmissions by .3 more than clinical teams.
-This is not a huge magnitude given an average readmission rate of 21, but reveals a meaningful difference in behavior given that both hospital types significantly decrease readmissions.
+Panel \ref{fig:read_synth_clinical} shows the estimated difference in readmission rates, and reveals that non-clinical teams decrease readmissions by .3 (standard error .13) more than clinical teams.
+This is a relatively small magnitude given an average readmission rate of 21, but reveals a meaningful difference in behavior given that both hospital types significantly decrease readmissions.
 In panel \ref{fig:mort_synth_clinical}, I show the estimated difference in mortality.
@@
 
-     \begin{figure}
+     \begin{figure}[ht!]
      \caption{Effect of Clinical Experience on Readmission and Mortality Rates}
@@
 
+    Additionally, I estimate the intensive margin of clinical leadership: whether the fraction of clinical executives on the team affects quality.
+I bin the fraction of clinical executives into less than a quarter or more than a quarter and less than one half.\footnote{Less than .5\% of hospitals have greater than 50\% clinical executives on the team.} Then, I estimate:
+
+    \begin{equation}
+    \label{eq:clinical_continuous_onefourth}
+    y_{ht} = \beta \left(\mathbbm{1}\{0 < \rho \leq .25\}\text{ x post 2012}\right)_{ht} + \gamma_{h} + \delta_t + \epsilon_{ht}, \text{ and}
+    \end{equation}
+    \begin{equation}
+    \label{eq:clinical_continuous_onehalf}
+    y_{ht} = \beta \left(\mathbbm{1}\{.25 < \rho \leq .5\}\text{ x post 2012}\right)_{ht} + \gamma_{h} + \delta_t + \epsilon_{ht},
+    \end{equation}
+
+    \noindent where $\rho$ is the fraction of clinical executives on hospital $h$'s executive team, and $\gamma$ and $\delta$ are fixed effects as defined in Equation \ref{eq:clinical}.
+
+    I show the results of this estimation with readmission rates as the outcome in Figure \ref{fig:cont_clinicalsynthdid}.
+In panel \ref{fig:onefourth_read_synth_clinical}, the estimation compares executive teams with no clinical experience to those with up to 25\% of the team having clinical experience.
+The average treatment effect is .04 and not statistically significant.
+Thus, there is no significant difference in how these types of teams respond to the incentive change.
+In panel \ref{fig:onehalf_read_synth_clinical}, I compare hospitals with no clinically trained executives over the course of the sample to hospitals with over a quarter of their team having clinical training.
+While these hospitals have parallel pre-trends, hospitals without clinically trained executives decrease readmissions by .64 (standard error .33) more than hospitals with clinically trained executives.
+Thus, the readmission rate response is fully driven by having a higher proportion of clinical executives relative to no clinical executives.
+
+    \begin{figure}[ht!]
+     \caption{Effect of Clinical Experience on Readmission Rates, Binned Treatment}
+     \centering
+     \begin{subfigure}[b]{0.45\textwidth}
+         \centering
+         \caption{$0 < \rho \leq .25$}
+         \includegraphics[width=\textwidth]{Objects/cont_onefourthread_md_nomd_synth_graph.pdf}
+         \label{fig:onefourth_read_synth_clinical}
+     \end{subfigure}
+     \hfill
+     \begin{subfigure}[b]{0.45\textwidth}
+         \centering
+         \caption{$.25 < \rho \leq .5$}
+         \includegraphics[width=\textwidth]{Objects/cont_onehalfread_md_nomd_synth_graph.pdf}
+         \label{fig:onehalf_read_synth_clinical}
+     \end{subfigure}
+        \label{fig:cont_clinicalsynthdid}
+    \end{figure}
+
+    Any potential difference in mortality rate response also seems to be driven by executive teams with over a quarter of the executives having clinical experience, as shown in Figure \ref{fig:cont_clinicalsynthdid_mort}.
+However, the standard errors of this estimation are high, potentially due to the noisiness in the mortality measure.
+The magnitude of the estimate in panel $\ref{fig:onefourth_mort_synth_clinical}$ is .07, so I can rule out differential responses by teams with less than a quarter of clinical executives.
+Taking the readmission and mortality rate results together, the effect of having clinical leadership on quality after an incentive change depends on having a sufficient proportion of the team with clinical training.
+
+    \begin{figure}[ht!]
+     \caption{Effect of Clinical Experience on Mortality Rates, Binned Treatment}
+     \centering
+     \begin{subfigure}[b]{0.45\textwidth}
+         \centering
+         \caption{$0 < \rho \leq .25$}
+         \includegraphics[width=\textwidth]{Objects/cont_onefourthmort_md_nomd_synth_graph.pdf}
+         \label{fig:onefourth_mort_synth_clinical}
+     \end{subfigure}
+     \hfill
+     \begin{subfigure}[b]{0.45\textwidth}
+         \centering
+         \caption{$.25 < \rho \leq .5$}
+         \includegraphics[width=\textwidth]{Objects/cont_onehalfmort_md_nomd_synth_graph.pdf}
+         \label{fig:onehalf_mort_synth_clinical}
+     \end{subfigure}
+        \label{fig:cont_clinicalsynthdid_mort}
+    \end{figure}
+
     \section{Selective Patient Practices}
 
-    A method hospitals have been shown to employ in order to decrease readmission rates is being more selective with which patients to admit to the hospital.
-One hypothesis is that executive teams without any clinicians may be more willing to participate in such practices, whereas doctors prioritize care for all patients.
-This would explain why non-clinical hospitals are able to decrease readmissions more drastically.
+    A method hospitals have been shown to employ in order to decrease readmission rates is being more selective with which patients to admit to the hospital (\cite{gupta2021impacts}).
+One hypothesis behind a more drastic quality decrease by non-clinical teams is that they may be more willing to participate in such practices, while executives with clinical training prioritize care for all types of patients.
 
-    To test this, I employ case mix index as the outcome in specification \ref{eq:clinical}, to determine whether one type of hospital differentially changes patient composition after pay-for-performance incentives, where the same identification assumptions hold.
-I present the estimated difference in Figure \ref{fig:main_cmi_clinical}, which shows no differential response to the programs in terms of the complexity of patients admitted to the hospital.
-\textcolor{red}{sentence about the implications of this?}
+    To test this, I employ case mix index as the outcome in specification \ref{eq:clinical} to determine whether one type of hospital differentially changes patient composition after pay-for-performance incentives, where the same identification assumptions hold.
+I present the estimated difference in Figure \ref{fig:main_cmi_clinical}, comparing hospitals that never have a clinical executive to hospitals that always have a clinical executive.
+Neither type of hospital drastically changes case mix index after the incentive change, and there is no estimated differential response in complexity of patients seen in the hospital.
+Thus, the estimated difference in readmission rate response between clinical and non-clinical executives is not due to differentially changing case mix.
 
@@
 
-    \section{Does Specialty Matter?}
+    \section{Internal Medicine vs. Other Specialties}
+
+    While most doctors still take a hippocratic oath to care for patients with integrity and compassion, doctors with training in certain specialties may be more equipped to improve care for patients with the specific conditions relevant in this context.
+Specifically, a pulmonologist typically treats pneumonia, and a cardiologist typically treats heart failure and heart attack patients, which are the three conditions considered in the HRRP penalties.
+Both of these specialties are under the umbrella of internal medicine.
+My physician specialty information does not get more granular than internal medicine as a specialty.
+Thus, I estimate the difference in response between
 
@@
 
-    Another hypothesis that could explain the differential response due to clinical leadership is how profit-driven a clinician is compared to other training backgrounds such as an MBA.
+    Another hypothesis that could explain the differential response due to clinical leadership is how profit-driven a clinician is compared to other training backgrounds.
 To demonstrate this, consider hospital behavior under the enactment of pay-for-performance policies, where hospitals face a trade-off between profit and societal benefit.
@@
 
-    I compare readmission and mortality rates of always clinical executives and never clinical executives to for-profit hospitals.
-While ideally I would also have leadership information for for-profit hospitals, the tax forms only have nonprofit hospitals.
-However, on average, for-profit hospitals have revealed through their ownership status that they heavily weigh profit.
+    I merge all general for-profit hospitals in the AHA data to the sample of nonprofit hospitals.
+Then, I compare readmission and mortality rates of always clinical executives and never clinical executives to for-profit hospitals after the incentive change.
+Ideally, I would also have leadership heterogeneity information for for-profit hospitals, but this data does not exist publicly.
+However, on average, for-profit hospitals have revealed through their ownership status that they more heavily weigh profit.
 Thus, comparing different executive team behavior to for-profit behavior helps reveal how profit driven certain leadership teams are.
@@
     \label{eq:forprofit}
-    y_{ht} = \beta \text{ (for-profit x post 2012)}_{ht} + \gamma_{h} + \delta_t + \epsilon_{ht},
+    y_{ht} = \beta \text{ (for-profit x post 2012)}_{ht} + \gamma_{h} + \delta_t + \epsilon_{ht}
     \end{equation}
@@
     \noindent for two subsets of data.
-First, I limit only to for-profits and nonprofits with clinical leadership.
-Second, I limit only to for-profits and nonprofits without clinical leadership.
+First, I limit only to for-profits and nonprofits with clinically trained executives.
+Second, I limit to for-profits and nonprofits without clinically trained executives.
 Thus, $\beta$ represents the effect of being for-profit compared to being nonprofit with a certain leadership team type.
-I show summary statistics of relevant variables for for-profit hospitals in Appendix \textcolor{red}{add table}.
-Again, I employ a synthetic difference in differences strategy to make the parallel trends assumption more plausible.
+I show summary statistics of relevant variables for for-profit hospitals in Appendix \ref{app:sumstats}.
+Again, I employ a synthetic difference-in-differences strategy to rely less heavily on a strong parallel trends assumption for different types of hospitals.
 
-    The identification assumptions I make here are largely similar to those in the main analysis, only now incorporate for-profit hospitals.
+    The identification assumptions I make here are largely similar to those in Section \ref{sec:clinical}, only now incorporate for-profit hospitals.
 First, I assume that, absent the change in incentives, for-profits and the relevant comparison group would have had parallel trends in outcomes.
@@
 Finally, I assume no other changes correlated with both hospital type (for-profit, nonprofit with and without clinically trained executives) and quality occurred in conjunction with the 2012 pay-for-performance policies.
-
-    The estimates and graphical representation of the effect of being for-profit is shown in Figure \ref{fig:read_synth_plot}.
-In panel \ref{fig:read_synth_plotb}, the average treatment effect represents the effect of being for-profit compared to nonprofit with clinical executives on readmission rates after pay-for-performance changes.
-While readmission rates trend similarly prior to the incentive change, for-profits decrease readmission rates by .3 more than nonprofits with clinically trained executives.
-That is, for-profits respond more aggressively to the incentive change than nonprofits with clinical leadership.
-However, panel \ref{fig:read_synth_plotc} shows no such difference when comparing for-profits to nonprofits without clinically trained executives.
-Not only is the estimate not statistically significant, but the magnitude indicates a precise zero.
-That is, a lack of clinically trained executives leads nonprofits to act like for-profits in terms of readmission rates when responding to this incentive change.
 
@@
 
+    The estimates and graphical representation of the effect of being for-profit is shown in Figure \ref{fig:read_synth_plot}.
+In panel \ref{fig:read_synth_plotb}, the average treatment effect shows the effect of being for-profit compared to nonprofit with clinical executives on readmission rates after pay-for-performance changes.
+While readmission rates trend similarly prior to the incentive change, for-profits decrease readmission rates by .3 more than nonprofits with clinically trained executives.
+That is, for-profits respond more aggressively to the incentive change than nonprofits with clinical leadership.
+This difference is nearly identical to the difference between clinical and non-clinical teams shown in Section \ref{sec:clinical}.
+So, unsurprisingly, panel \ref{fig:read_synth_plotc} shows no such difference when comparing for-profits to nonprofits without clinically trained executives.
+Not only is the estimate not statistically significant, but the magnitude indicates a precise zero.
+That is, a lack of clinically trained executives leads nonprofits to act like for-profits in terms of readmission rates when responding to this incentive change.
+
     Next, I present estimated differences in mortality rates between different hospital types in Figure \ref{fig:mort_synth_plot}.
@@
 While I cannot rule out a lack of differential response, the documented noisiness of mortality measures (\cite{mackenzie2016measuring}) may mean that the true effect is different from zero.
-However, the estimated difference in mortality between for-profits and nonprofits without clinically trained executives is certainly zero.
+However, the estimated difference in mortality between for-profits and nonprofits without clinically trained executives is certainly zero with a amgnitude of .03, shown in panel \ref{fig:mort_synth_plotc}.
 
     \begin{figure}[ht!]
-     \caption{Mortality Rate Synthetic Difference in Differences Results}
+     \caption{Comparison to For-Profit: Mortality}
      \centering
@@
          \centering
-         \caption{For-Profit and NFP w/ MD Exec}
+         \caption{For-Profit and Always Clinical Exec}
          \includegraphics[width=\textwidth]{Objects/fp_mort_md_synth_graph.pdf}
@@
          \centering
-         \caption{For-Profit and NFP w/out MD Exec}
+         \caption{For-Profit and Never Clinical Exec}
          \includegraphics[width=\textwidth]{Objects/fp_mort_nomd_synth_graph.pdf}
@@
 
-    Given that having clinical leaders affects behavior, an important distinction is whether the existing clinical leaders \textit{reveal} the underlying objectives of the hospital, or \textit{manage} the hospital differently.
-Up to this point, I have limited the sample of nonprofit hospitals to those that don't change their propensity to hire a clinically trained executive over the sample period, which combines signalling and managing.
-However, under the assumption that hiring or firing these types of executives is not correlated with the pay-for-performance policy changes, such changes can be leveraged to decompose the effect into signaling vs.
-managing.
+    Given that having clinically trained executives affects behavior, an important distinction is whether the existing clinical leaders \textit{reveal} the underlying objectives of the hospital, or \textit{manage} the hospital differently.
+Up to this point, I have limited the sample of nonprofit hospitals to those that don't change their propensity to hire a clinically trained executive over the sample period, which combines any signaling and managing effects.
+However, under the assumption that hiring or firing these types of executives is not correlated with the pay-for-performance policy changes, such changes can be leveraged to decompose the effect into signaling versus managing.
+Therefore, I first investigate leadership changes in response to the policy change.
+
+    \subsection{Leadership Change Analysis} \label{sec:changes}
+
+    I have discussed the assumptions of parallel trends, no anticipation, and no other confounding events in both Sections \ref{sec:clinical} and \ref{sec:forprofit}.
+These assumptions are also necessary here.
+Further, this decomposition analysis depends heavily on leadership team changes not endogenously occurring due to pay-for-performance policies.
+Up to this point, that assumption was important because of the way I selected the sample.
+However, this decomposition depends directly on leveraging such leadership changes.
+Thus, I now explore this assumption in detail.
+First, I present raw means over time that represent the fraction of hospitals who hire a clinical executive, fire a clinical executive, or have any change in their propensity to have a clinical executive in Figure \ref{fig:change_means}.
+There do not seem to be any drastic changes in 2011 or 2012 when the policies were becoming relevant to hospitals.
+
+    \begin{figure}[ht!]
+    \centering
+    \caption{Leadership Change Means Over Time}
+    \includegraphics[width=0.8\textwidth]{Objects/change_means.pdf}
+    \label{fig:change_means}
+     \end{figure}
+
+    Next, I assess whether this assumption is reasonable by analyzing how executive team changes are occurring over time for certain groups of hospitals. I estimate:
+
+    \begin{equation}\label{eq:change2}
+    \text{change}_{ht} = \sum_{j=2011}^{2014}\beta_j(\mathbf{1}\{t=j\}\times \text{Program Exposed})_{ht} + \alpha_h + \epsilon_{ht},
+    \end{equation}
+
+    where the variable $\text{Program Exposed}_{h}$ has three definitions.
+First, an indicator for whether the hospital was ever penalized under HRRP.
+Second, an indicator for whether the hospital ever received payments under HVBP.
+And finally, I leave out the Program Exposed variable to simply see how changes are correlated with given years for all hospitals.
+The outcome change$_{ht}$ is an indicator for whether the hospital changes the number of clinical executives in a given year.
+The estimates from this analysis are presented in Figure \ref{fig:change_analysis}.
+All estimates are not statistically different from zero, indicating that expectations of program exposure do not lead to changes in propensity to hire a clinical executive.
+Thus, it seems unlikely that endogenous team formation is biasing the estimates.
+
+     \begin{figure}[ht!]
+         \centering
+         \caption{Leadership Analysis Results}
+         \includegraphics[width=0.8\textwidth]{Objects/change_analysis_plot.pdf}
+         \label{fig:change_analysis}
+     \end{figure}
+
+     \subsection{Estimation}
 
@@
     \end{equation}
+
+    Thus, in one specification I compare hospitals who ever have a clinically trained executive to hospitals who never have a clinically trained executive.
+This captures the signaling effect.
+In another specification, I remove all hospitals who never had a clinically trained executive, and compare hospitals who had a clinically trained executive in 2012 to those who had a clinically trained executive in any year except 2012, capturing a managing effect.
 
@@
 
-    \subsection{Are Leadership Changes Correlated with Pay-for-Performance Policies?} \label{sec:changes}
-
-    I have discussed the assumptions of parallel trends, no anticipation, and no other confounding events in both Sections \ref{sec:clinical} and \ref{sec:forprofit}.
-These assumptions are also necessary here.
-Further, this decomposition analysis depends heavily on leadership team changes not endogenously occurring due to pay-for-performance policies.
-Up to this point, that assumption was important because of the way I selected the sample.
-However, this decomposition depends directly on leveraging such leadership changes.
-Thus, I now explore this assumption in detail.
-First, I present raw means over time that represent the fraction of hospitals who hire a clinical executive, fire a clinical executive, or have any change in their propensity to have a clinical executive in Figure \ref{fig:change_means}.
-There do not seem to be any drastic changes in 2011 or 2012 when the policies were becoming relevant to hospitals.
-
-    \begin{figure}[ht!]
-    \centering
-    \caption{Leadership Change Means Over Time}
-    \includegraphics[width=0.8\textwidth]{Objects/change_means.pdf}
-    \label{fig:change_means}
-     \end{figure}
-
-    Further, I assess whether this assumption is reasonable by analyzing how executive team changes are occurring over time for certain groups of hospitals. I estimate:
-
-    \begin{equation}\label{eq:change2}
-    \text{change}_{ht} = \sum_{j=2011}^{2014}\beta_j(\mathbf{1}\{t=j\}\times \text{Program Exposed})_{ht} + \alpha_h + \epsilon_{ht},
-    \end{equation}
-
-    where the variable $\text{Program Exposed}_{h}$ has three definitions.
-First, an indicator for whether the hospital was ever penalized under HRRP.
-Second, an indicator for whether the hospital ever received payments under HVBP.
-And finally, I leave out the Program Exposed variable to simply see how changes are correlated with given years for all hospitals.
-The outcome change$_{ht}$ is an indicator for whether the hospital changes the number of clinical executives in a given year.
-The estimates from this analysis are presented in Figure \ref{fig:change_analysis}.
-All estimates are not statistically different from zero, indicating that expectations of program exposure do not lead to changes in propensity to hire a clinical executive.
-Thus, it seems unlikely that endogenous team formation is biasing the estimates.
-In Appendix \ref{app:changes}, I present various descriptive statistics of executive team changes over time, showing that hiring and firing of MD executives do not change much over time, the rate of clinical executives to total executives remains stable over time, and the distribution of timing of clinical executive changes.
-
-     \begin{figure}[ht!]
-         \centering
-         \caption{Leadership Analysis Results}
-         \includegraphics[width=0.8\textwidth]{Objects/change_analysis_plot.pdf}
-         \label{fig:change_analysis}
-     \end{figure}
-
-    \subsection{Decomposition Results}
-
     The estimates of each decomposition specification are shown in Table \ref{tab:MD_noMD_readmort_decomp_synth}.
-The decomposition for readmission rates is in Columns (1) and (2), and for mortality is in Columns (3) and (4).
+The readmission rate estimates are in columns (1) and (2), and mortality rate estimates are in columns (3) and (4).
 The first row compares ever clinical executive hospitals to never clinical executive hospitals, and the second row compares hospitals with a clinical executive in 2012 to hospitals who have a clinical executive in a year other than 2012.
@@
     While previous literature has documented correlations between executive characteristics and firm behaviors/performance, research is still limited on this topic in the health care sector.
-Particularly, does having executive leaders with clinical training trickle down into hospital behaviors/outcomes?
-In this paper, I investigate the differential response of hospital executive teams to a policy change that financially incentivizes quality.
-
-    I collect data on the names of hospital executives from 2010-2014 from publicly available Tax Form 990s, which are completed by all nonprofits in the US.
+This paper investigates whether having executive leaders with clinical training impacts hospital behaviors and outcomes, particularly in response to a policy change that financially incentivizes quality.I collect data on the names of hospital executives from 2010-2014 from publicly available Tax Form 990s, which are completed by all nonprofits in the US.
 I merge this data with existing hospital and physician characteristics, and aggregate to the hospital level.
-Thus, I end up with a data set of nonprofit US hospitals that captures the proportion of clinical executives, specialties of clinical executives, characteristics, and quality metric outcomes.
+Thus, I utilize a unique data set of nonprofit US hospitals that captures the proportion of clinical executives, specialties of clinical executives, hospital characteristics, and hospital quality metric outcomes.
 
@@
 These policies were relevant to hospital finances, and gave tangible incentive for hospitals to increase quality.
-While I don't observe and can't causally estimate why certain hospitals have clinical leadership and others don't, I can observe how clinical executives may respond to an exogenous shock differently than non-clinical executives conditional on teams being previously formed.
-
-    First, I estimate the effect of always having clinical leadership vs.
-never having clinical leadership on readmission and mortality rates after 2012 using a synthetic difference in differences estimation strategy.
+Thus, I leverage this exogenous policy change and estimate how clinically trained executives affect readmission and mortality rates in response to the change using a synthetic difference-in-differences strategy.
 Conditional on parallel trends holding, I find that having clinical experience mitigates the response to the incentive change.
 That is, while all hospitals reduce readmissions after the policies take place, non-clinical teams reduce readmissions at a fast rate than clinical teams.
-Next, I investigate several hypotheses regarding this differential response.
-This result is not driven by non-clinical teams adjusting their patient complexity.
-However, comparing nonclinical teams to for-profit hospitals reveals that it could be underlying preferences for profit vs.
+
+    I investigate several mechanisms by which this difference could occur.
+First, this result is not driven by non-clinical teams adjusting their patient complexity.
+However, comparing non-clinical teams to for-profit hospitals reveals that it could be underlying preferences for profit vs.
 patient benefit that drive the difference.
-Finally, I show that changes in leadership teams are not correlated with the policies, and leverage such changes to show that clinical leaders manage the hospital differently and are not just a revealed preference of underlying objectives.
+Finally, I leverage exogenous changes in executive teams to show that clinically trained executives are not simply a signal of underlying preferences, but manage the hospital differently.
 
-    \textcolor{red}{discuss implications of these findings}
+    These findings have significant policy implications, as it informs the way different types of hospital leadership respond to policy changes.
+While I do not directly estimate the effect of hiring a clinically trained executive, I do show evidence that clinically trained executives on average place higher value on patient care.
+Thus, another way to incentivize quality is to encourage clinical leadership.
 
@@
 
-\section{Modelling Hospital Response to Pay-for-Performance}\label{sec:model}
+\section{Additional Summary Statistics}\label{app:sumstats}
+
+\import{Tables}{forprofit_sample_sumstats.tex}
+
+\subsection{Matched vs. Unmatched Not-for-Profits}\label{app:matched}
+
+The sample of all NFP hospitals that I use for analysis consists of 852 hospitals.
+In the AHA data, there are 2,258 additional NFP hospitals that either were not matched to tax forms, did not exist in the tax form data, or did not have executive names in the tax forms.
+While it would be ideal to analyze all NFPs, a sample size of 852 firms is relatively large in the hospital executive literature.
+
+In this section, I compare the in-sample NFPs to the out-of-sample NFPs along observable measures.
+I present means and standard deviations of these measures in Table \ref{tab:nfp_sample_compare}.
+The in-sample NFPs are slightly larger in terms of beds and patients seen.
+They are also slightly more likely to be penalized or receive payments under the pay-for-performance incentives.
+However, average readmission and mortality rates are very similar, as well as case mix index.
+The largest difference in the two samples is that I under-represent NFP hospitals in systems.
+This is due to the nature of the tax form 990s, where systems often file one tax form for the entire system, making it difficult to discern the true managers of a specific hospital.
+Because of this, I drop hospitals from the sample who only have leadership information from a system tax form.
+
+\import{Tables}{NFP_sample_comparison.tex}
+
+\subsection{Executive Team Changes}\label{app:changes}
+
+I present the number of hospitals who change their propensity to hire a clinically trained executive at different times in Table \ref{tab:change_timing}.
+I break the timing up into pre-2012, 2012, and post-2012.
+I show the number of hospitals who have MDs in different time combinations.
+The number of hospitals that only have an MD executive in 2012, but at no other time, is only 6.
+The majority of hospitals (187) who ever have a clinical executive have one in every time period.
+
+\import{Tables}{change_num_table_2012.tex}
+
+\section{Modeling Hospital Response to Pay-for-Performance}\label{sec:model}
 
@@
 
-\section{Matched vs. Unmatched Not-for-Profits}\label{app:matched}
-
-The sample of all NFP hospitals that I use for analysis consists of 852 hospitals.
-In the AHA data, there are 2,258 additional NFP hospitals that either were not matched to tax forms, did not exist in the tax form data, or did not have executive names in the tax forms.
-While it would be ideal to analyze all NFPs, a sample size of 852 firms is relatively large in the hospital executive literature.
-
-In this section, I compare the in-sample NFPs to the out-of-sample NFPs along observable measures.
-I present means and standard deviations of these measures in Table \ref{tab:nfp_sample_compare}.
-The in-sample NFPs are slightly larger in terms of beds and patients seen.
-They are also slightly more likely to be penalized or receive payments under the pay-for-performance incentives.
-However, average readmission and mortality rates are very similar, as well as case mix index.
-The largest difference in the two samples is that I under-represent NFP hospitals in systems.
-This is due to the nature of the tax form 990s, where systems often file one tax form for the entire system, making it difficult to discern the true managers of a specific hospital.
-Because of this, I drop hospitals from the sample who only have leadership information from a system tax form.
-
-\import{Tables}{NFP_sample_comparison.tex}
-
-\section{Executive Team Changes}\label{app:changes}
-
-Here, I present descriptive statistics of executive team changes.
-In Figure \ref{fig:change_graph}, I show the average rate of MD executives to total executives, the percent of hospitals who fire an MD, and the percent of hospitals who hire an MD.
-The fraction of MD executives remains stable over time at approximately .22.
-Similarly, the percent of hospitals who hire an MD over time remains stable at approximately 21\%.
-There is a small increase in the percent of hospitals who fire an MD in 2011, but even with this increase firing is not common and stable, ranging from 12\% to 13\%.
-
-\begin{figure}[ht!]
-    \centering
-    \caption{Executive Team Changes Over Time}
-    \includegraphics[width=\textwidth]{Objects/change_graph.pdf}
-    \label{fig:change_graph}
-\end{figure}
-
-Additionally, I present the number of hospitals who change their propensity to hire a clinically trained executive at different times in Table \ref{tab:change_timing}.
-I break the timing up into pre-2012, 2012, and post-2012.
-I show the number of hospitals who have MDs in different time combinations.
-The number of hospitals that only have an MD executive in 2012, but at no other time, is only 6.
-The majority of hospitals (187) who ever have a clinical executive have one in every time period.
-
-\import{Tables}{change_num_table_2012.tex}
-
 \section{Results By Condition}\label{app:condition}
-
-The main results of the paper consider the weighted average readmission or mortality rate as the main outcomes.
-Here, I present results for each condition individually.
-Table \ref{tab:fp_read_condition_synth.tex}, I show the effect of being for-profit on AMI, heart failure, and pneumonia readmission rates compared to differing NFP comparison groups.
-The estimates indicate that the weighted average is mainly driven by pneumonia patients, since AMI readmission rate responses differ between FP and all NFP, where no other condition does.
-
-\import{Tables}{fp_read_condition_synth.tex}
-
-Table \ref{tab:fp_mort_condition_synth.tex} shows a very similar result for mortality rates.
-AMI responses differ between FP and all NFP or NFP without clinical executives, but the weighted average is driven mainly by pneumonia patients.
-
-\import{Tables}{fp_mort_condition_synth.tex}
-
-The differences between NFPs with and without clinical executives are also driven mainly by pneumonia patients, shown in Table \ref{tab:MD_noMD_readmort_condition_synth}.
 

### 2024-08-01 Paper/paper2.tex
Updates from Overleaf

@@
     \begin{figure}[ht!]
-     \caption{Effect of Clinical Experience on Readmission Rates, Binned Treatment}
+     \caption{Effect of Clinical Training on Readmission Rates, Binned Treatment}
      \centering
@@
     \begin{figure}[ht!]
-     \caption{Effect of Clinical Experience on Mortality Rates, Binned Treatment}
+     \caption{Effect of Clinical Training on Mortality Rates, Binned Treatment}
      \centering
@@
 
-    While most doctors still take a hippocratic oath to care for patients with integrity and compassion, doctors with training in certain specialties may be more equipped to improve care for patients with the specific conditions relevant in this context.
+    While most doctors still take a hippocratic oath to provide care with integrity and compassion, doctors with training in certain specialties may be more equipped to improve care for patients with the particular conditions relevant in this context.
 Specifically, a pulmonologist typically treats pneumonia, and a cardiologist typically treats heart failure and heart attack patients, which are the three conditions considered in the HRRP penalties.
 Both of these specialties are under the umbrella of internal medicine.
-My physician specialty information does not get more granular than internal medicine as a specialty.
-Thus, I estimate the difference in response between
+The physician specialty information in my data does not get more granular than internal medicine as a specialty.
+Thus, I estimate the difference in response between internal medicine physicians and other specialty physicians.
+
+    \begin{figure}[ht!]
+     \caption{Effect of Internal Medicine Training on Quality}
+     \centering
+     \begin{subfigure}[b]{0.45\textwidth}
+         \centering
+         \caption{Readmission Rates}
+         \includegraphics[width=\textwidth]{Objects/read_specialty_synth_graph.pdf}
+         \label{fig:read_spec}
+     \end{subfigure}
+     \hfill
+     \begin{subfigure}[b]{0.45\textwidth}
+         \centering
+         \caption{Mortality Rates}
+         \includegraphics[width=\textwidth]{Objects/mort_specialty_synth_graph.pdf}
+         \label{fig:mort_spec}
+     \end{subfigure}
+        \label{fig:specialty}
+    \end{figure}
 

### 2024-08-01 Paper/paper2.tex
Updates from Overleaf

@@
 Specifically, a pulmonologist typically treats pneumonia, and a cardiologist typically treats heart failure and heart attack patients, which are the three conditions considered in the HRRP penalties.
-Both of these specialties are under the umbrella of internal medicine.
-The physician specialty information in my data does not get more granular than internal medicine as a specialty.
-Thus, I estimate the difference in response between internal medicine physicians and other specialty physicians.
+Both of these specialties are under the umbrella of internal medicine, which is the most granular specialty type in my data.
+
+    I limit the sample of hospitals to those who always have some type of clinical executive over the sample period.
+Then, I estimate the difference in readmission and mortality rates after pay-for-performance among hospitals who do and do not have an internal medicine clinical executive.
+I lose a significant fraction of the sample with this limitation: there are 60 hospitals that consistently have an internal medicine executive or only a different specialty executive.
+The summary statistics for these samples are in Appendix
 
@@
 
+\import{Tables}{specialty_sumstats.tex}
+
 \subsection{Matched vs. Unmatched Not-for-Profits}\label{app:matched}

### 2024-08-12 Paper/paper2.tex
Updates from Overleaf

@@
 
-	\linespread{1.2}\title{\vspace{-0.5in} Does Hospital Leadership Matter?\\ \large Evidence from Pay-for-Performance Incentives}
+	\linespread{1.2}\title{\vspace{-0.5in} Does Hospital Leadership Matter?\\ \large Evidence from Pay-for-Performance}
 
@@
 		{\small
-         One determinant of firm behaviors and performance is the type of management leading the firm.
-Characteristics such as age, gender, and background are correlated with employee wages, decisions of mergers and acquisitions, and other firm outcomes.
-In this paper, I investigate this topic in the context of hospital executive teams.
-Specifically, I estimate the effect of clinically trained executives on hospital quality.
-I leverage pay-for-performance policies that changed quality incentives for hospitals in the US in 2012, and estimate the difference in readmission and mortality rates for nonprofit hospitals with and without clinically trained executives.
-I find that clinically trained executives do not improve quality as much as executive teams without clinically trained executives.
+         Managers are often seen as drivers of firm decisions, strategies, and performance, and manager heterogeneity has been linked to employee wages, decisions of mergers and acquisitions, and financial performance.
+A significant goal in US health care is to improve quality while decreasing spending, and yet it is unclear whether hospital managers might contribute to this goal.
+In this paper, I estimate the effect of clinically trained executives on hospital quality.
+I leverage pay-for-performance initiatives that changed financial incentives on quality of care for hospitals in the US in 2012, and estimate the difference in readmission and mortality rates for nonprofit hospitals with and without clinically trained executives.
+I find that hospitals without clinically trained executives respond more to financial incentives than those without clinically trained executives.
 This difference is not driven by a change in patient complexity, but may be driven by preferences towards profit vs.
 patient welfare.
-Further, clinically trained executives are not just a signal of underlying hospital incentives, but manage the hospital differently.
 		} 
 	\end{abstract}
-
-	\vspace{0.8in}
-
-	\noindent Keywords:
-
-	\noindent JEL Codes:
 
@@
 
-    Broadly, firm leader characteristics such as gender, age, and background have been shown to be correlated with financial performance and behaviors such as employee diversity, wages, aggressiveness in mergers and acquisitions, and other corporate strategies.
-I consider this topic in the context of hospital quality under executives with differing backgrounds: some with training as a medical doctor and some without.
+    \section{Introduction}
+
+    Executives are considered crucial drivers of a company's practices and performance, as they often set broad visions and implement strategies to achieve them.
+Executive heterogeneity has been linked to various decisions and outcomes of firms: characteristics such as gender, age, and background are correlated with financial performance and behaviors such as employee diversity, wages, aggressiveness in mergers and acquisitions, and other corporate strategies.\footnote{\cite{bertrand2003managing}; \cite{matsa2011chipping}; \cite{matsa2011chipping}; \cite{adams2009women}; \cite{ahern2012changing}; \cite{flabbi2019female}; \cite{benmelech2015military}; \cite{custodio2013generalists}; \cite{frydman2019rising}} However, there is still much to learn about how firm behaviors and objectives can be shaped by differences in leadership.
+
+    In this paper, I study how executive background affects firm behavior in the context of US hospitals.
+In the US, health care makes up almost 20\% of spending (\cite{CMS_National_Health_Expenditure}), and a major goal of policymakers is to improve the value of health care.
+Yet, little is known about how hospital managers affect health care value, and it is unclear whether existing findings on executive characteristics can be extrapolated to the health care context.
+This is because existing literature is generally limited to large, publicly traded firms, whereas a majority of US hospitals are nonprofit.
+Relatedly, health care executives have a particularly unique trade-off between profit and patient welfare, which varies from the types of firms studied in existing literature.
+
+    A characteristic relevant to this context is whether or not hospital executives have clinical training.
 Physicians can bring a unique combination of clinical and administration expertise, and they have insights that can potentially benefit their employees and lead to better quality (\cite{Stajduhar_2023}, \cite{Ahmed_2022}).
-However, hiring physician executives can also be controversial, as physicians are not always adequately trained in business management, which can lead to poor business practices and lower quality (\cite{HarvardBusinessReview2018}).
-
-    In this paper, I estimate how having clinically trained executives in leadership affects quality of care in hospitals.
-I scrape executive team names and titles from publicly available Tax Form 990s, filled out by nonprofit firms in the US.
+However, hiring physician executives is controversial, as physicians are not always trained in business management, which can lead to poor practices and lower quality (\cite{HarvardBusinessReview2018}).
+Thus, I estimate how having clinically trained executives in hospital leadership affects quality of care.
+I scrape executive names and titles from publicly available Tax Form 990s, filled out by nonprofit firms in the US.
 Combining this with publicly available hospital characteristics, I construct a novel data set of nonprofit hospitals from 2010-2014 containing information on both executive team characteristics and quality metrics.
-While limited to nonprofit hospitals due to data availability, this sample is a very important driver of health care in the US.
+While limited to nonprofits due to data availability, this sample of hospitals is a very important driver of health care in the US.
 Private nonprofit hospitals make up 50\% of all hospitals in the US, and staff, on average, 207 beds.
-For comparison, for-profits make up 36\% of hospitals and staff 107 beds (\cite{ASPE_2023}).
-Therefore, the behavior of these hospitals directly affects health care consumers in the US.
+For comparison, for-profits make up 36\% of hospitals and staff 107 beds on average (\cite{ASPE_2023}).
 
     The non-random selection of leaders makes it difficult to interpret any direct comparison of differing executive teams as causal.
-Therefore, I leverage an exogenous shock to hospital incentives, pay-for-performance policies enacted in the US in 2012, and compare the change in readmission and mortality rates of executive teams with and without clinical executives.
-Under the assumption that the two types of teams would have had similar trends in readmission and mortality rates absent the incentive change, I find that executive teams with clinical executives on average do not improve quality as much as executive teams without clinical executives.
-While both types of hospital leadership teams decrease readmission rates after the policy change, those without clinical executives decrease readmissions at a faster rate than those with clinical executives.
+Therefore, I leverage pay-for-performance policies enacted in the US as an exogenous shock to hospital incentives, and compare the change in readmission and mortality rates of hospitals with and without clinically trained executives over the entire sample period.
+Under the assumption that the two types of teams would have had similar trends in readmission and mortality rates absent the incentive change, I find that hospitals without clinically trained trained executives respond more to financial incentives on quality than hospitals with clinical executives.
+That is, while both types of hospital leadership teams decrease readmission rates after the policy change, those without clinical executives decrease readmissions at a faster rate than those with clinical executives.
+Interestingly, this result is driven by hospitals with between one quarter and one half of their executive team having clinical training.
+Further, executives with clinical training in fields specific to the policy change respond more than other clinically trained executives.
 
-    Having found that hospital leadership does, in fact, matter to hospital behavior, I investigate several mechanisms that might contribute to this difference.
-First, the more drastic decrease in readmissions by non-clinical teams is not driven by decreasing average patient complexity.
-Further, comparing each type of leadership team with for-profit hospitals reveals that the absence of clinical executives leads a nonprofit hospital to respond more similarly to a for-profit hospital than nonprofits with clinical executives.
-That is, having clinical training on a leadership team leads a hospital to act less profit-driven than their general training counterparts.
-Finally, I show that changes in a hospital's propensity to hire a clinical executive are not correlated with the pay-for-performance policies.
-Thus, I exploit such changes in leadership teams to decompose the effect of clinical executives into signaling (clinical executives signal underlying hospital preferences), or managing (clinical executives manage the hospital differently).
-I find that the entire difference in response is driven by a managing effect.
+    One potential reason for this differential response is that hospitals without clinically trained executives are more willing to participate in selective patient practices.
+To investigate this, I use patient case mix index as an outcome and estimate the difference in case mix index after the financial incentive change.
+I find no evidence that non-clinical executive teams differentially change patient complexity.
+
+    Next, I investigate whether this differential response could be due to differences in objective functions.
+I consider hospital objectives as a function of profit and patient welfare, and allow weight placed on each to vary with executive team type.
+Hospitals that place more weight on profit will naturally respond more drastically to a policy that adds quality directly into the profit function.
+Thus, a potential reason for the difference in response is that non-clinical executive teams are more profit driven.
+I investigate this empirically by comparing both types of hospital leadership teams to for-profit hospitals, who have revealed that they are largely profit driven by their ownership type.
+This analysis reveals that non-clinical executive teams respond to the financial incentives similarly to for-profits, and hospitals with clinically trained executives act differently than for-profits.
+
+    Finally, the question still remains whether clinically trained executives are a signal of underlying hospital preferences, or if they actively manage the hospital differently than their non-clinically trained counterparts.
+I analyze this by leveraging changes in hospital propensity to hire clinical executives to decompose the full effect of clinical leadership into a signaling effect (clinical executives signal underlying hospital preferences), or managing effect (clinical executives manage the hospital differently).
+First, I show that hospital executive changes are not correlated with the change to financial incentives.
+For the signaling effect, I compare hospitals who ever have a clinically trained executive to hospitals who never have a clinically trained executive.
+For the managing effect, I limit to hospitals who ever have a clinically trained executive and compare those who have a clinically trained executive when the policies were announced to those who have clinically trained executives at other times.
+I find that the difference in response is driven by a managing effect.
 
@@
 
+    I use several data sets to analyze the relationship between hospital executives with clinical training and response to financial incentives on quality.
+I construct a novel data set of hospital executives by scraping publicly available tax forms, and merge this data to publicly available data sets containing hospital characteristics: the American Hospital Association Survey (AHA), the Centers for Medicare and Medicaid Services (CMS) Hospital Compare data and Case Mix Index files, and Healthcare Cost Report Information System (HCRIS) data.
+Ultimately, I employ this hospital-level data spanning from 2010-2014, which contains executive team characteristics, mortality and readmission rates as measures of quality, and other characteristics.
+A detailed description of the construction of this data can be found in Appendix \ref{appendixdata}.
+
     \subsection{Nonprofit Hospital Executives}
 
-    Nonprofit firms in the US differ from for-profits mainly in that for-profit firms distribute revenue to stakeholders, while nonprofits reinvest profit back into the firm.
-Hospitals are governed by a board of directors, whose role is to set broad goals and strategies, and provide general oversight.
+    Nonprofit hospitals, as with most nonprofits, are typically governed by a board of directors, whose role is to set broad goals and strategies, and provide general oversight.
 The board selects executives, the highest level of management of a firm, to carry out day-to-day operations.
@@
 Some hospital executives specialize in health care settings by earning a degree in health care management, or an MBA specific to health care.
-There are executive positions that are often filled by someone with clinical training, such as a Chief Medical Officer (CMO), but medical doctors can pursue other top managerial positions as well such as CEO or president.
+There are executive positions that are often filled by someone with clinical training, such as a Chief Medical Officer (CMO), but medical doctors can also fill other executive positions.
 While some doctors earn additional degrees before stepping into an executive role, this is not a necessary condition to becoming a physician executive.
@@
 I gather this data from publicly available Tax Form 990s, which every sufficiently large nonprofit files each year in the US.
-To my knowledge, this is the first large scale gathering of executive names from these forms.\footnote{\citeauthor{brickley2010board} (\citeyear{brickley2010board}) collects compensation data for a select number of hospitals.}
-
-    All tax-exempt organizations in the US are required to file a Form 990 with Internal Revenue Services (IRS) each year.
+To my knowledge, this is the first large scale gathering of executive names from these forms.\footnote{\citeauthor{brickley2010board} (\citeyear{brickley2010board}) collects compensation data for a select number of hospitals.} All tax-exempt organizations in the US are required to file a Form 990 with Internal Revenue Services (IRS) each year.
 There are different types of forms, but any organization grossing over \$200,000 must file the most extensive Form 990.
 Sections of this form include a statement of revenue, statement of functional expenses, a balance sheet, and, as used in this project, a list of all key employees, executives, and board members.
-Each firm is required to report the name and title, average hours per week, position, and compensation of their board members and executives.
 
-    The historical tax forms are housed by ProPublica.\footnote{https://projects.propublica.org/not-for-profits/} I use the NonProfit Explorer API to extract Employee Identification Numbers (EIN) of firms with a hospital National Taxonomy of Exempt Entities (NTEE) code.
-After filtering out associations and specialty hospitals, there are approximately 3,000 EINs.
-Linking these EINs to other hospital information is crucial to the analysis, but no crosswalk from EIN to other hospital identifiers is publicly available.
-I therefore rely on matching by location and name of hospital in both the tax forms and American Hospital Association (AHA) data.
-In the AHA data, I limit to general hospitals.
-Then, I confidently match 1200 EINs to an AHA identification number based on exact name matches within the same state.\footnote{I assess the observable differences between matched and non-matched hospitals in Appendix \ref{app:matched}, and conclude that the samples are overall similar apart from an under-sampling of hospitals that belong to systems.}
-
-    For each of the matched EINs, I extract web URLs to Form 990 PDFs for each year.
-I use an algorithm to download these locally and use optical character recognition (OCR) text extraction methods to scrape the relevant section of the PDF, which is titled ``Officers, Directors, Trustees, Key Employees, and Highest Compensated Employees".
-Finally, I use string cleaning methods to identify all names and positions under each EIN, year.
-With each name, I extract titles indicating clinical training: MD, Dr., or DO.
-Using OCR extraction can be unreliable in cases where the text on the form is handwritten or small.
-Therefore, for observations that are missing after the initial extraction, I manually record names, positions, and titles.
-As I limit the sample to hospitals who have text information in every year, I limit the years to 2010-2014 to retain more hospitals.
+    For each hospital in the tax form data, I seek matches in the AHA data based on hospital name and location.\footnote{I assess the observable differences between matched and non-matched hospitals in Appendix \ref{app:matched}, and conclude that the samples are overall similar apart from an under-sampling of hospitals that belong to systems.} For all hospitals that can be matched to a general hospital in the AHA data, I use optimal character recognition (OCR) methods to extract the names, titles, and positions of each listed executive from its Tax Form 990.
+While board members are an important aspect of hospital organization, I do not include them in this analysis as they are beyond the scope of this paper.
+While I extract text spanning 2009-2015, I retain a much larger sample of hospitals by limiting to hospitals that have text information in 2010-2014.
 This yields 852 nonprofit hospitals that are matched to AHA data, and with complete text information in each year.
 
-    Finally, I seek to (1) verify clinical training of those who claim to be doctors, and (2) obtain additional information about clinically trained executives.
+    Finally, I seek to (1) verify clinical training of those who are identified as doctors, and (2) obtain additional information about clinically trained executives.
 Thus, I match each name with the National Plan and Provider Enumeration System (NPPES) database of all registered physicians.
-For any person who has a doctor title and uniquely matches the NPPES data, I record their National Provider Identifier (NPI) number.
-For anyone who has a doctor title and no matches, no doctor title and matches, or a doctor title and multiple matches, I use information on location and hospital name to manually match NPI numbers of those who did have clinical training.
 I present a table of means for various characteristics of doctor executives in the data in Table \ref{doc_sumstats}.
-In total, there are 813 doctor executives.
-On average, they are 52 years old, and 11\% of them are female.
+In total, there are 813 clinically trained executives.
+On average, they are 52 years old, and 11\% of clinically trained executives are female.
 While it is more likely they are a Chief Medical Officer than a Chief Executive Officer, still less than a third are CMOs.
@@
 It is not uncommon for two hospitals to have the same position but the tasks performed are drastically different, or to have two distinct positions that essentially perform the same tasks.
-I drop anyone who is solely a board member, and focus on those who have executive in their title, or are labeled as a president or vice president of a department.
 I create various hospital-level characterizations based on their executive team, such as the existence of a clinically trained executive, the number of clinically trained executives, the number of total executives, and whether the hospital has a Chief Medical Officer.
-For the main analysis, I limit to hospitals who either always have a clinical executive or never have a clinical executive over the course of the sample in order to have a clean comparison.
 
@@
 
+    The purpose of this paper is to investigate hospital behavior in the midst of changes to financial incentives that change how hospitals receive payment for quality of care.
 Two programs were passed as part of the Affordable Care Act that focused on pay-for-performance incentives for hospitals: the Hospital Readmissions Reduction Program (HRRP), and the Hospital Value Based Purchasing Program (HVBP).
@@
 Beginning in October 2012, hospitals with higher readmission rates than the national average in pneumonia, heart failure, or AMI (after adjusting for demographic characteristics) receive a fixed lower reimbursement rate for all Medicare patients seen in their hospital.
-In 2015, CMS also included chronic obstructive pulmonary disease, coronary artery bypass graft surgery, and elective primary total hip arthroplasty and/or total knee arthroplasty as conditions which go into the penalty calculation (\cite{CMS}).
-
 Penalties are given in the form of a fixed rate reduction of 1-3\% for every Medicare patient regardless of the condition.
-Further, CMS does not distinguish a necessary readmission from an avoidable readmission; any repeat hospital visit is included in the penalty calculation.
 Excess readmission rates are calculated using a rolling look-back period of 3 years to determine whether the hospital is penalized.
@@
 I combine rates for pneumonia, AMI, and heart failure (the relevant HRRP conditions) into a weighted average based on the number of patients in each condition.
-Additionally, I include several relevant hospital characteristics from the AHA survey: number of beds, and whether the hospital is an academic medical center, is physician owned, or is system affiliated.
-Additionally, case mix index, a measure of overall patient complexity, comes from the CMS Case Mix Index files.
-Whether the hospital was penalized under HRRP or received payments under HVBP is found in the HCRIS data.
 
@@
 
-    The hospitals in my sample are limited to general nonprofit hospitals with executive team information in each year (2010-2014) and a link from EIN to AHA ID.
     In Table \ref{tab:sumstats_samples_stable}, I present means of relevant variables for several hospital samples used throughout the paper.
@@
 
-    Hospitals that never have a clinical executive are smaller, less likely to be an academic medical center, and less likely to be affiliated with a system.
-Further, they have smaller executive teams on average by 1-2 people, and are less likely to have anyone acting as a CMO.\footnote{Though, they may title the position differently.} Finally, in regards to exposure to pay-for-performance programs, hospitals who never have clinical executives are less likely to receive benefits from the HVBP Program, but seem more similar to hospitals with clinical leadership in terms of penalties from HRRP.
+    Comparing columns (2) and (3), hospitals that never have a clinical executive are smaller, less likely to be an academic medical center, and less likely to be affiliated with a system.
+Further, they have smaller executive teams on average by 1-2 people, and are less likely to have anyone acting as a designated CMO.
+Finally, in regards to exposure to pay-for-performance programs, hospitals who never have clinical executives are less likely to receive benefits from the HVBP Program, but seem more similar to hospitals with clinical leadership in terms of penalties from HRRP.
 
     Next, I plot weighted average readmission, mortality rates, and patient case mix index across time for different hospital types in Figure \ref{fig:outcomes_graph}.
-Pay-for-performance incentive change timing is shown as a dotted line in 2012.
-Average readmission rates trend similarly prior to 2012, and all hospital type exhibit a decrease in readmission after 2012.
-However, hospitals who consistently have a clinical executive show a slower rate of decrease in raw readmissions.
+The effective start year of pay-for-performance incentives is shown as a dotted line in 2012.
+Average readmission rates trend similarly prior to 2012, and all hospital types exhibit a decrease in raw readmission rates after 2012.
+However, there is a faster rate of decrease for hospitals who have a clinically trained executive at some point, or who never have a clinically trained executive.
 When considering mortality, there is not a drastic change in 2012 for any hospital type, but hospitals who always have a clinical executive continue on an increasing trend in mortality where the other hospitals types remain more flat.
@@
 
-    \section{Effect of Executive Clinical Training on Hospital Quality}\label{sec:clinical}
+    \section{Effect of Executive Clinical Training on Response to Quality Incentives}\label{sec:clinical}
 
@@
 That is, these hospitals do not change their propensity to hire a clinical executive during the sample period.
-Therefore, an important assumption that I make is that this sample selection is not correlated with the program enactment.
+\footnote{An important assumption for this analysis is that this sample selection is not correlated with the program enactment.
 I investigate this assumption in Section \ref{sec:changes}, and find no evidence that hospitals are more or less likely to change their propensity to hire a clinical executive because of the pay-for-performance programs.
+}
 
@@
     In the unweighted two-way-fixed-effects specification, we have reason to believe that the parallel trends assumption may not hold due to observed differences in characteristics other than the leadership team, as discussed in Section \ref{sec:data}.
-Once incorporating the weights used in the synthetic differences in differences method, we ensure parallel trends in the outcome variables prior to 2012.
+Once incorporating the weights used in the synthetic differences in differences method, we ensure parallel pre-trends in the outcome variables.
 Thus, I assume that under the weighted clinical and non-clinical hospital composition, trends would have continued to be parallel absent the incentive change in 2012.
-A component of this assumption is the implicit assumption that no other event occurred in 2012 correlated with both the outcomes and composition of leadership team.
+This includes an implicit assumption that no other event occurred in 2012 correlated with both the outcomes and composition of leadership team.
 Finally, I assume that there is no anticipation of the policy change before 2012.
@@
 In panel \ref{fig:mort_synth_clinical}, I show the estimated difference in mortality.
-While the magnitude of the difference is similar to that for readmission rates, mortality is a noisy measure, and I cannot rule out no difference in mortality response.
+While the magnitude of the difference is similar to that for readmission rates, mortality is a noisy measure, and I cannot rule out null effects for mortality.
 In Appendix \ref{app:condition}, I show that these results are primarily driven by pneumonia patients.
@@
      \begin{figure}[ht!]
-     \caption{Effect of Clinical Experience on Readmission and Mortality Rates}
+     \caption{Effect of Clinical Training on Readmission and Mortality Rates}
      \centering
@@
     Additionally, I estimate the intensive margin of clinical leadership: whether the fraction of clinical executives on the team affects quality.
-I bin the fraction of clinical executives into less than a quarter or more than a quarter and less than one half.\footnote{Less than .5\% of hospitals have greater than 50\% clinical executives on the team.} Then, I estimate:
+I categorize hospitals with clinically trained executives into two bins: above the median and below median fraction of clinically trained executives.
+Let the median fraction be $\eta$, which in this case is .2.
+Then, I estimate the synthetic difference-in-differences version of the following specifications:
 
@@
     \label{eq:clinical_continuous_onefourth}
-    y_{ht} = \beta \left(\mathbbm{1}\{0 < \rho \leq .25\}\text{ x post 2012}\right)_{ht} + \gamma_{h} + \delta_t + \epsilon_{ht}, \text{ and}
+    y_{ht} = \beta \left(\mathbbm{1}\{\rho \leq \eta\}\text{ x post 2012}\right)_{ht} + \gamma_{h} + \delta_t + \epsilon_{ht}, \text{ and}
     \end{equation}
@@
     \label{eq:clinical_continuous_onehalf}
-    y_{ht} = \beta \left(\mathbbm{1}\{.25 < \rho \leq .5\}\text{ x post 2012}\right)_{ht} + \gamma_{h} + \delta_t + \epsilon_{ht},
+    y_{ht} = \beta \left(\mathbbm{1}\{\rho > \eta\}\text{ x post 2012}\right)_{ht} + \gamma_{h} + \delta_t + \epsilon_{ht},
     \end{equation}
@@
 
-    I show the results of this estimation with readmission rates as the outcome in Figure \ref{fig:cont_clinicalsynthdid}.
-In panel \ref{fig:onefourth_read_synth_clinical}, the estimation compares executive teams with no clinical experience to those with up to 25\% of the team having clinical experience.
+    I present the results of this estimation with readmission rates as the outcome in Figure \ref{fig:cont_clinicalsynthdid}.
+In panel \ref{fig:belowmed_read_synth_clinical}, the estimation compares executive teams with no clinical training to those with up to 25\% of the team having clinical training.
 The average treatment effect is .04 and not statistically significant.
 Thus, there is no significant difference in how these types of teams respond to the incentive change.
-In panel \ref{fig:onehalf_read_synth_clinical}, I compare hospitals with no clinically trained executives over the course of the sample to hospitals with over a quarter of their team having clinical training.
-While these hospitals have parallel pre-trends, hospitals without clinically trained executives decrease readmissions by .64 (standard error .33) more than hospitals with clinically trained executives.
+In panel \ref{fig:abovemed_read_synth_clinical}, I compare hospitals with no clinically trained executives over the course of the sample to hospitals with over a quarter of their team having clinical training.
+I find that hospitals without clinically trained executives decrease readmissions by .64 (standard error .33) more than hospitals with clinically trained executives.
 Thus, the readmission rate response is fully driven by having a higher proportion of clinical executives relative to no clinical executives.
@@
          \centering
-         \caption{$0 < \rho \leq .25$}
-         \includegraphics[width=\textwidth]{Objects/cont_onefourthread_md_nomd_synth_graph.pdf}
-         \label{fig:onefourth_read_synth_clinical}
+         \caption{$\rho \leq \eta$}
+         \includegraphics[width=\textwidth]{Objects/cont_belowmedread_md_nomd_synth_graph.pdf}
+         \label{fig:belowmed_read_synth_clinical}
      \end{subfigure}
@@
          \centering
-         \caption{$.25 < \rho \leq .5$}
-         \includegraphics[width=\textwidth]{Objects/cont_onehalfread_md_nomd_synth_graph.pdf}
-         \label{fig:onehalf_read_synth_clinical}
+         \caption{$\rho > \eta$}
+         \includegraphics[width=\textwidth]{Objects/cont_abovemedread_md_nomd_synth_graph.pdf}
+         \label{fig:abovemed_read_synth_clinical}
      \end{subfigure}
@@
 
-    Any potential difference in mortality rate response also seems to be driven by executive teams with over a quarter of the executives having clinical experience, as shown in Figure \ref{fig:cont_clinicalsynthdid_mort}.
+    Any potential difference in mortality rate response also seems to be driven by executive teams with over a quarter of the executives having clinical training, as shown in Figure \ref{fig:cont_clinicalsynthdid_mort}.
 However, the standard errors of this estimation are high, potentially due to the noisiness in the mortality measure.
-The magnitude of the estimate in panel $\ref{fig:onefourth_mort_synth_clinical}$ is .07, so I can rule out differential responses by teams with less than a quarter of clinical executives.
+The magnitude of the estimate in panel $\ref{fig:belowmed_mort_synth_clinical}$ is .07, so I can rule out differential responses by teams with less than a quarter of clinical executives.
 Taking the readmission and mortality rate results together, the effect of having clinical leadership on quality after an incentive change depends on having a sufficient proportion of the team with clinical training.
@@
          \centering
-         \caption{$0 < \rho \leq .25$}
-         \includegraphics[width=\textwidth]{Objects/cont_onefourthmort_md_nomd_synth_graph.pdf}
-         \label{fig:onefourth_mort_synth_clinical}
+         \caption{$\rho \leq \eta$}
+         \includegraphics[width=\textwidth]{Objects/cont_belowmedmort_md_nomd_synth_graph.pdf}
+         \label{fig:belowmed_mort_synth_clinical}
      \end{subfigure}
@@
          \centering
-         \caption{$.25 < \rho \leq .5$}
-         \includegraphics[width=\textwidth]{Objects/cont_onehalfmort_md_nomd_synth_graph.pdf}
-         \label{fig:onehalf_mort_synth_clinical}
+         \caption{$\rho > \eta$}
+         \includegraphics[width=\textwidth]{Objects/cont_abovemedmort_md_nomd_synth_graph.pdf}
+         \label{fig:abovemed_mort_synth_clinical}
      \end{subfigure}
@@
 I lose a significant fraction of the sample with this limitation: there are 60 hospitals that consistently have an internal medicine executive or only a different specialty executive.
-The summary statistics for these samples are in Appendix
+The summary statistics for these samples are in Appendix \ref{app:sumstats}.
+I present the results from this comparison in Figure \ref{fig:specialty}.
 
@@
     \end{figure}
+
+    While I am limited in power due to the small sample size of hospitals in this analysis, I provide evidence that hospitals with internal medicine executives may change readmission rates differentially compared to teams with clinical executives of other specialty types.
+The estimated effect of having an internal medicine executive is -.38 (standard error .24), where internal medicine executives decrease readmissions at a faster rate than other specialty types.
+The results for mortality rates are similar (still not statistically significant), but smaller in magnitude and smaller relative to the main results.
+This is probably due to the specialty-specific conditions being more relevant for the readmission rate penalties than the quality rewards.
 
@@
 
-     \subsection{Estimation}
+     \subsection{Estimation and Findings}
 
@@
     While previous literature has documented correlations between executive characteristics and firm behaviors/performance, research is still limited on this topic in the health care sector.
-This paper investigates whether having executive leaders with clinical training impacts hospital behaviors and outcomes, particularly in response to a policy change that financially incentivizes quality.I collect data on the names of hospital executives from 2010-2014 from publicly available Tax Form 990s, which are completed by all nonprofits in the US.
+This paper investigates whether having executive leaders with clinical training impacts hospital behaviors and outcomes, particularly in response to a policy change that financially incentivizes quality.
+I collect data on the names of hospital executives from 2010-2014 from publicly available Tax Form 990s, which are completed by all nonprofits in the US.
 I merge this data with existing hospital and physician characteristics, and aggregate to the hospital level.
@@
 Thus, I leverage this exogenous policy change and estimate how clinically trained executives affect readmission and mortality rates in response to the change using a synthetic difference-in-differences strategy.
-Conditional on parallel trends holding, I find that having clinical experience mitigates the response to the incentive change.
+Conditional on parallel trends holding, I find that having clinical training mitigates the response to the incentive change.
 That is, while all hospitals reduce readmissions after the policies take place, non-clinical teams reduce readmissions at a fast rate than clinical teams.
 
-    I investigate several mechanisms by which this difference could occur.
+    I investigate several mechanisms by which this difference might occur.
 First, this result is not driven by non-clinical teams adjusting their patient complexity.
@@
 patient benefit that drive the difference.
+Additionally, hospitals with clinically trained executives in specialties relevant to the policies improve quality more.
 Finally, I leverage exogenous changes in executive teams to show that clinically trained executives are not simply a signal of underlying preferences, but manage the hospital differently.
@@
     These findings have significant policy implications, as it informs the way different types of hospital leadership respond to policy changes.
-While I do not directly estimate the effect of hiring a clinically trained executive, I do show evidence that clinically trained executives on average place higher value on patient care.
-Thus, another way to incentivize quality is to encourage clinical leadership.
+I provide evidence that clinical executives value quality even before quality is incentivized, and that having particular specialties in leadership might matter for quality of care.
 
@@
 \import{Tables}{change_num_table_2012.tex}
+
+\section{HRRP and HVBP Programs}\label{app:programs}
+
+In October 2011, the Center for Medicare and Medicaid Services (CMS) released a set of rules under HRRP mandating penalties for hospitals with above average readmission rates.
+The goal of HRRP is to lower readmissions through better care coordination, less initial stay complications, and better post-care instructions.
+Beginning in October 2012, hospitals with higher readmission rates than the national average in pneumonia, heart failure, or AMI (after adjusting for demographic characteristics) receive a fixed lower reimbursement rate for all Medicare patients seen in their hospital.
+In 2015, CMS also included chronic obstructive pulmonary disease, coronary artery bypass graft surgery, and elective primary total hip arthroplasty and/or total knee arthroplasty as conditions which go into the penalty calculation (\cite{CMS}).
+
+    Penalties are given in the form of a fixed rate reduction of 1-3\% for every Medicare patient regardless of the condition.
+Further, CMS does not distinguish a necessary readmission from an avoidable readmission; any repeat hospital visit is included in the penalty calculation.
+Excess readmission rates are calculated using a rolling look-back period of 3 years to determine whether the hospital is penalized.
+Therefore, hospitals had incentive to react immediately once details of the program were announced in October of 2011.
+
+    The HVBP Program instead rewards hospitals with high quality or significant improvement in quality.
+Specifically, CMS deducts Medicare payments by 2\% from all eligible hospitals, collects this sum, and divides it among the rewarded hospitals.
+Several quality and cost measures surrounding safety, efficiency, cost reductions, clinical outcomes, and community engagement are combined to create a single score metric for each hospital.
+Hospitals are then compared to the average and are rewarded for being above average quality or for showing improvement (\cite{CMS_2023}).
 

### 2024-08-15 Paper/paper2.tex

@@
 		{\small
-         Managers are often seen as drivers of firm decisions, strategies, and performance, and manager heterogeneity has been linked to employee wages, decisions of mergers and acquisitions, and financial performance.
-A significant goal in US health care is to improve quality while decreasing spending, and yet it is unclear whether hospital managers might contribute to this goal.
-In this paper, I estimate the effect of clinically trained executives on hospital quality.
-I leverage pay-for-performance initiatives that changed financial incentives on quality of care for hospitals in the US in 2012, and estimate the difference in readmission and mortality rates for nonprofit hospitals with and without clinically trained executives.
-I find that hospitals without clinically trained executives respond more to financial incentives than those without clinically trained executives.
-This difference is not driven by a change in patient complexity, but may be driven by preferences towards profit vs.
+         Managers are often seen as drivers of firm decisions, strategies, and performance.
+Further, manager characteristics have been linked to employee wages, decisions of mergers and acquisitions, and financial performance.
+Health care policies in the US often focus on improving quality of care without increasing costs, and yet it is unclear whether hospital managers might contribute to this goal.
+In this paper, I estimate the differential response to such quality incentives by different types of nonprofit hospital executives: those with and without a clinical training background.
+I find that hospitals without clinically trained executives respond more to the financial incentives than those with clinically trained executives.
+I provide evidence that this may be due to differing preferences for profit vs.
 patient welfare.
+Thus, incentivizing clinically trained executives in hospitals could be an effective method to improve value in the US health care system.
 		} 
@@
 
-    Executives are considered crucial drivers of a company's practices and performance, as they often set broad visions and implement strategies to achieve them.
-Executive heterogeneity has been linked to various decisions and outcomes of firms: characteristics such as gender, age, and background are correlated with financial performance and behaviors such as employee diversity, wages, aggressiveness in mergers and acquisitions, and other corporate strategies.\footnote{\cite{bertrand2003managing}; \cite{matsa2011chipping}; \cite{matsa2011chipping}; \cite{adams2009women}; \cite{ahern2012changing}; \cite{flabbi2019female}; \cite{benmelech2015military}; \cite{custodio2013generalists}; \cite{frydman2019rising}} However, there is still much to learn about how firm behaviors and objectives can be shaped by differences in leadership.
+    In corporate environments, executives are considered crucial drivers of a company's practices and performance, as they are heavily involved in the culture, visions, and strategies of the firm.
+Executive heterogeneity has been linked to various actions and outcomes of firms; characteristics such as gender, age, and background are correlated with financial performance and behaviors such as employee diversity, wages, aggressiveness in mergers and acquisitions, and other corporate strategies.\footnote{\cite{bertrand2003managing}; \cite{matsa2011chipping}; \cite{matsa2011chipping}; \cite{adams2009women}; \cite{ahern2012changing}; \cite{flabbi2019female}; \cite{benmelech2015military}; \cite{custodio2013generalists}; \cite{frydman2019rising}} However, there are still settings and relevant characteristics for which the relationship between management and firm behavior is not well understood.
 
-    In this paper, I study how executive background affects firm behavior in the context of US hospitals.
-In the US, health care makes up almost 20\% of spending (\cite{CMS_National_Health_Expenditure}), and a major goal of policymakers is to improve the value of health care.
-Yet, little is known about how hospital managers affect health care value, and it is unclear whether existing findings on executive characteristics can be extrapolated to the health care context.
-This is because existing literature is generally limited to large, publicly traded firms, whereas a majority of US hospitals are nonprofit.
-Relatedly, health care executives have a particularly unique trade-off between profit and patient welfare, which varies from the types of firms studied in existing literature.
+    In this paper, I study whether type of executive training affects firm behavior in the context of US hospitals.
+In the US, health care makes up almost 20\% of spending (\cite{CMS_National_Health_Expenditure}), and a major goal of policymakers is to improve the quality of health care without increasing costs.
+Yet, little is known about how hospital managers affect health care value, and it is unclear whether existing findings on executive characteristics can be extrapolated to the health care, and particularly hospital, context.
+Existing literature is generally limited to large, publicly traded firms, whereas a majority of US hospitals are nonprofit with presumably different objectives.
+In particular, health care executives have a particularly unique trade-off between profit and patient welfare, which varies from the types of firms studied in existing literature.
 
@@
 However, hiring physician executives is controversial, as physicians are not always trained in business management, which can lead to poor practices and lower quality (\cite{HarvardBusinessReview2018}).
-Thus, I estimate how having clinically trained executives in hospital leadership affects quality of care.
-I scrape executive names and titles from publicly available Tax Form 990s, filled out by nonprofit firms in the US.
-Combining this with publicly available hospital characteristics, I construct a novel data set of nonprofit hospitals from 2010-2014 containing information on both executive team characteristics and quality metrics.
-While limited to nonprofits due to data availability, this sample of hospitals is a very important driver of health care in the US.
+Thus, I estimate how having clinically trained executives in hospital leadership affects hospital outcomes.
+I scrape executive names and titles of nonprofit hospitals in the US from publicly available Tax Form 990s.
+Combining this with other public hospital data, I construct a novel data set of nonprofit hospitals from 2010-2014 containing information on both executive team characteristics and quality metrics.
+While limited to nonprofits due to data availability, this sample of hospitals is an important driver of health care in the US.
 Private nonprofit hospitals make up 50\% of all hospitals in the US, and staff, on average, 207 beds.
@@
     The non-random selection of leaders makes it difficult to interpret any direct comparison of differing executive teams as causal.
-Therefore, I leverage pay-for-performance policies enacted in the US as an exogenous shock to hospital incentives, and compare the change in readmission and mortality rates of hospitals with and without clinically trained executives over the entire sample period.
+Therefore, I leverage pay-for-performance policies enacted in the US as an exogenous shock to hospital incentives on quality, and compare the change in readmission and mortality rates of hospitals with and without clinically trained executives.
 Under the assumption that the two types of teams would have had similar trends in readmission and mortality rates absent the incentive change, I find that hospitals without clinically trained trained executives respond more to financial incentives on quality than hospitals with clinical executives.
 That is, while both types of hospital leadership teams decrease readmission rates after the policy change, those without clinical executives decrease readmissions at a faster rate than those with clinical executives.
-Interestingly, this result is driven by hospitals with between one quarter and one half of their executive team having clinical training.
-Further, executives with clinical training in fields specific to the policy change respond more than other clinically trained executives.
+I find no evidence of selective patient practices driving this result.
 
-    One potential reason for this differential response is that hospitals without clinically trained executives are more willing to participate in selective patient practices.
-To investigate this, I use patient case mix index as an outcome and estimate the difference in case mix index after the financial incentive change.
-I find no evidence that non-clinical executive teams differentially change patient complexity.
-
-    Next, I investigate whether this differential response could be due to differences in objective functions.
-I consider hospital objectives as a function of profit and patient welfare, and allow weight placed on each to vary with executive team type.
+    I investigate whether the result might be explained by differing preferences on the trade-off between profit and patient welfare, where I model these hospital objectives and allow weight placed on each to vary with executive team type.
 Hospitals that place more weight on profit will naturally respond more drastically to a policy that adds quality directly into the profit function.
 Thus, a potential reason for the difference in response is that non-clinical executive teams are more profit driven.
-I investigate this empirically by comparing both types of hospital leadership teams to for-profit hospitals, who have revealed that they are largely profit driven by their ownership type.
-This analysis reveals that non-clinical executive teams respond to the financial incentives similarly to for-profits, and hospitals with clinically trained executives act differently than for-profits.
+I investigate this empirically by comparing the response of hospitals with and without clinically trained executives to the response of for-profit hospitals, who have revealed that they are more profit driven by their ownership type.
+I find that nonprofit hospitals without clinically trained executives respond to the financial incentives on quality similarly to for-profits, while nonprofits without clinically trained executives respond differently than for-profits.
+For-profits and non-clinical executive teams change quality more in response to the incentive change than clinically trained executive teams.
 
-    Finally, the question still remains whether clinically trained executives are a signal of underlying hospital preferences, or if they actively manage the hospital differently than their non-clinically trained counterparts.
+    Finally, the question remains whether clinically trained executives are a signal of unchanging underlying hospital preferences, or if they actively manage the hospital differently, effectively changing hospital objectives.
 I analyze this by leveraging changes in hospital propensity to hire clinical executives to decompose the full effect of clinical leadership into a signaling effect (clinical executives signal underlying hospital preferences), or managing effect (clinical executives manage the hospital differently).
@@
 For the managing effect, I limit to hospitals who ever have a clinically trained executive and compare those who have a clinically trained executive when the policies were announced to those who have clinically trained executives at other times.
-I find that the difference in response is driven by a managing effect.
+I find that the difference in response is driven by clinically trained executives managing the hospital differently.
 
-    Many researchers have documented a correlation between executive characteristics and firm performance, starting with \citeauthor{bertrand2003managing} (\citeyear{bertrand2003managing}), who show that manager fixed effects are an important driver of many firm behaviors and decisions.
-Female board members are associated with better oversight and more women executives (\cite{matsa2011chipping}; \cite{adams2009women}), but are negatively correlated with firm value (\cite{ahern2012changing}).
-Having more female executives is correlated with female employee wages and corporate strategies (\cite{flabbi2019female}; \cite{matsa2013female}); young male CEOs tend to be more aggressive in mergers and acquisitions, while those with military experience are less aggressive (\cite{levi2010deal}; \cite{benmelech2015military}); CEOs with general ability tend to receive higher pay and perform better (\cite{kaplan2012ceo}; \cite{custodio2013generalists}; \cite{adams2018director}; \cite{frydman2019rising}).
+    Many researchers have documented a correlation between executive characteristics and firm actions, starting with \citeauthor{bertrand2003managing} (\citeyear{bertrand2003managing}), who show that manager fixed effects are an important driver of many firm behaviors and decisions.
+Having more female executives is correlated with female employee wages and corporate strategies (\cite{flabbi2019female}; \cite{matsa2013female}); young male CEOs tend to be more aggressive in mergers and acquisitions, while those with military experience are less aggressive (\cite{levi2010deal}; \cite{benmelech2015military}); CEOs with general ability tend to receive higher pay and perform better (\cite{kaplan2012ceo}; \cite{custodio2013generalists}; \cite{adams2018director}; \cite{frydman2019rising}); older CEOs are associated with measures of corporate culture (\cite{graham2022corporate}).
 Finally, Chief Diversity Officers are not found to have any effect on hiring more diversely in universities (\cite{bradley2022impact}).
@@
 
-    I use several data sets to analyze the relationship between hospital executives with clinical training and response to financial incentives on quality.
-I construct a novel data set of hospital executives by scraping publicly available tax forms, and merge this data to publicly available data sets containing hospital characteristics: the American Hospital Association Survey (AHA), the Centers for Medicare and Medicaid Services (CMS) Hospital Compare data and Case Mix Index files, and Healthcare Cost Report Information System (HCRIS) data.
+    I use several data sets to analyze the relationship between executives with clinical training and hospital response to financial incentives on quality.
+I construct novel information on individual hospital executives by scraping publicly available tax forms, and merge this data to publicly available data sets containing other hospital characteristics: the American Hospital Association Survey (AHA), the Centers for Medicare and Medicaid Services (CMS) Hospital Compare data and Case Mix Index files, and Healthcare Cost Report Information System (HCRIS) data.
 Ultimately, I employ this hospital-level data spanning from 2010-2014, which contains executive team characteristics, mortality and readmission rates as measures of quality, and other characteristics.
-A detailed description of the construction of this data can be found in Appendix \ref{appendixdata}.
+In this section, I describe the institutional setting and data collection process.
+A more detailed description of the construction of this data can be found in Appendix \ref{appendixdata}.
 
@@
 
-    For each hospital in the tax form data, I seek matches in the AHA data based on hospital name and location.\footnote{I assess the observable differences between matched and non-matched hospitals in Appendix \ref{app:matched}, and conclude that the samples are overall similar apart from an under-sampling of hospitals that belong to systems.} For all hospitals that can be matched to a general hospital in the AHA data, I use optimal character recognition (OCR) methods to extract the names, titles, and positions of each listed executive from its Tax Form 990.
+    For each hospital in the tax form data, I merge to hospital matches in the AHA data based on name and location.\footnote{I assess the observable differences between matched and non-matched hospitals in Appendix \ref{app:matched}, and conclude that the samples are overall similar apart from an under-sampling of hospitals that belong to systems.} For all hospitals that can be matched to a general hospital in the AHA data, I use optimal character recognition (OCR) methods to extract the names, titles, and positions of each listed executive from its Tax Form 990.
 While board members are an important aspect of hospital organization, I do not include them in this analysis as they are beyond the scope of this paper.
-While I extract text spanning 2009-2015, I retain a much larger sample of hospitals by limiting to hospitals that have text information in 2010-2014.
+I extract text spanning 2009-2015, but I retain a much larger sample of hospitals by limiting to hospitals that have text information in 2010-2014.
 This yields 852 nonprofit hospitals that are matched to AHA data, and with complete text information in each year.
 
-    Finally, I seek to (1) verify clinical training of those who are identified as doctors, and (2) obtain additional information about clinically trained executives.
-Thus, I match each name with the National Plan and Provider Enumeration System (NPPES) database of all registered physicians.
+    Finally, I verify clinical training of those who are identified as doctors, and obtain additional information about clinically trained executives by matching names to the National Plan and Provider Enumeration System (NPPES) database of all registered physicians.
 I present a table of means for various characteristics of doctor executives in the data in Table \ref{doc_sumstats}.
@@
 
-    For the purposes of this paper, I focus on executive teams as a whole instead of specific positions.
-The main reason for this is because the organization of teams varies so drastically across hospitals.
+    I focus on executive teams as a whole instead of specific positions because the organization of teams varies so drastically across hospitals.
 It is not uncommon for two hospitals to have the same position but the tasks performed are drastically different, or to have two distinct positions that essentially perform the same tasks.
@@
 
-    The purpose of this paper is to investigate hospital behavior in the midst of changes to financial incentives that change how hospitals receive payment for quality of care.
+    The purpose of this paper is to investigate hospital behavior in the midst of financial incentives that change how hospitals receive payment for quality of care.
 Two programs were passed as part of the Affordable Care Act that focused on pay-for-performance incentives for hospitals: the Hospital Readmissions Reduction Program (HRRP), and the Hospital Value Based Purchasing Program (HVBP).
@@
 
-    Next, I plot weighted average readmission, mortality rates, and patient case mix index across time for different hospital types in Figure \ref{fig:outcomes_graph}.
+    Next, I plot weighted average readmission rates, mortality rates, and patient case mix index across time for different hospital types in Figure \ref{fig:outcomes_graph}.
 The effective start year of pay-for-performance incentives is shown as a dotted line in 2012.
@@
 When considering mortality, there is not a drastic change in 2012 for any hospital type, but hospitals who always have a clinical executive continue on an increasing trend in mortality where the other hospitals types remain more flat.
-Finally, hospitals that never have clinical executives consistently have less complex patients, but there is no notable change in raw case mix index over time.
+Finally, hospitals that never have clinical executives consistently have less complex patients, but there is no observable change in raw case mix index over time.
 
@@
 
-    \section{Effect of Executive Clinical Training on Response to Quality Incentives}\label{sec:clinical}
+    \section{Effect of Clinically Trained Executives on Response to Financial Incentives on Quality}\label{sec:clinical}
 
-    First, I test empirically whether having clinical training matters in how hospitals change readmission and mortality rates in response to the pay-for-performance policies effective in 2012.
-For a clean comparison, I limit the sample to hospitals who always have clinical executives or never have clinical executives.
-That is, these hospitals do not change their propensity to hire a clinical executive during the sample period.
-\footnote{An important assumption for this analysis is that this sample selection is not correlated with the program enactment.
+    I test empirically whether having clinically trained executives matters in how hospitals respond to the pay-for-performance policies effective in 2012.
+The quality metrics targeted by the programs are readmission and mortality rates, so I use these as outcomes.
+I limit the sample to hospitals who always have clinical executives or never have clinical executives.
+That is, the hospitals in this analysis do not change their propensity to hire a clinical executive during the sample period.\footnote{An important assumption for this analysis is that this sample selection is not correlated with the program enactment.
 I investigate this assumption in Section \ref{sec:changes}, and find no evidence that hospitals are more or less likely to change their propensity to hire a clinical executive because of the pay-for-performance programs.
@@
 
-    I estimate a two-way-fixed-effects specification, where the key independent variable is the interaction between having clinically trained executives and post 2012, when the programs began.
-Importantly, I am not estimating the effect of being penalized.
+    I ultimately employ a synthetic difference-in-differences estimation strategy that mirrors two-way-fixed effects, where the key independent variable is the interaction between having clinically trained executives and being in years after pay-for-performance took effect.
 All hospitals had incentive to improve quality in response to the policy change, as penalties and payments are determined by comparing quality among hospitals.
 Thus, I leverage the timing of the policy change, rather than the policies themselves.
-However, because of potential differences in quality trends between clinical and non-clinical hospital executive teams even before the policies, I employ a synthetic difference-in-differences estimation strategy (\cite{arkhangelsky2021synthetic}).
-That is, hospitals in the control group with similar trends in readmission, mortality, and baseline case mix before the programs are more heavily weighted, as well as time periods that balance pre- and post-program outcomes for hospitals without clinical executives.
-Below is the unweighted specification for clarity, where the identification assumptions are conceptually the same for synthetic difference-in-differences:
+One may be concerned about potential differences in quality trends between clinical and non-clinical executive team hospitals leading to a violation in the parallel trends assumption.
+This is why I implement a synthetic difference-in-differences estimation strategy (\cite{arkhangelsky2021synthetic}).
+That is, non-clinical team hospitals with trends similar to clinical team hospitals in readmission, mortality, and baseline case mix before the programs are more heavily weighted, as well as time periods that balance pre- and post-program outcomes for non-clinical teams.
+I present the unweighted two-way-fixed-effects specification equation for clarity, and the identification assumptions are conceptually similar for synthetic difference-in-differences.
 
@@
     \label{eq:clinical}
-    y_{ht} = \beta \text{ (clinical training x post 2012)}_{ht} + \gamma_{h} + \delta_t + \epsilon_{ht},
+    y_{ht} = \beta \text{ (clinical training x post 2012)}_{ht} + \gamma_{h} + \delta_t + \epsilon_{ht}
     \end{equation}
 
-    \noindent where $y_{ht}$ is one of the outcome variables capturing readmission or mortality rates discussed in Section \ref{sec:hrrp}, and $\gamma_h$ and $\delta_t$ are hospital and time fixed effects, respectively.
+    \noindent The variable $y_{ht}$ is one of the outcome variables capturing readmission or mortality rates discussed in Section \ref{sec:hrrp}, and $\gamma_h$ and $\delta_t$ are hospital and time fixed effects, respectively.
 
-    In the unweighted two-way-fixed-effects specification, we have reason to believe that the parallel trends assumption may not hold due to observed differences in characteristics other than the leadership team, as discussed in Section \ref{sec:data}.
-Once incorporating the weights used in the synthetic differences in differences method, we ensure parallel pre-trends in the outcome variables.
+    Once incorporating the weights used in the synthetic differences-in-differences method, we ensure parallel pre-trends in the outcome variables.
 Thus, I assume that under the weighted clinical and non-clinical hospital composition, trends would have continued to be parallel absent the incentive change in 2012.
@@
 This is a relatively small magnitude given an average readmission rate of 21, but reveals a meaningful difference in behavior given that both hospital types significantly decrease readmissions.
+In Appendix \ref{app:condition}, I show that these results are primarily driven by pneumonia patients.
+
     In panel \ref{fig:mort_synth_clinical}, I show the estimated difference in mortality.
-While the magnitude of the difference is similar to that for readmission rates, mortality is a noisy measure, and I cannot rule out null effects for mortality.
-In Appendix \ref{app:condition}, I show that these results are primarily driven by pneumonia patients.
+While the magnitude of the difference is similar to that for readmission rates (relative to the mean), mortality is a noisy measure, and I cannot rule out null effects.
+I investigate whether the estimated difference in readmission rate response could be due to differential participation in selective patient practices in Appendix \ref{app:casemix}, and find that this is not the case.
 
@@
 
-    Additionally, I estimate the intensive margin of clinical leadership: whether the fraction of clinical executives on the team affects quality.
+    Additionally, I estimate the intensive margin of clinical leadership: whether the fraction of clinically trained executives matters for response to financial incentive changes.
 I categorize hospitals with clinically trained executives into two bins: above the median and below median fraction of clinically trained executives.
-Let the median fraction be $\eta$, which in this case is .2.
-Then, I estimate the synthetic difference-in-differences version of the following specifications:
+Again, I present the unweighted version of the synthetic difference-in-differences that I ultimately employ:
 
@@
 
-    \noindent where $\rho$ is the fraction of clinical executives on hospital $h$'s executive team, and $\gamma$ and $\delta$ are fixed effects as defined in Equation \ref{eq:clinical}.
+    \noindent where $\rho$ is the fraction of clinical executives on hospital $h$'s executive team, $\eta$ is the median fraction of clinically trained executives, which is .2 in this sample, and $\gamma$ and $\delta$ are hospital and time fixed effects.
 
     I present the results of this estimation with readmission rates as the outcome in Figure \ref{fig:cont_clinicalsynthdid}.
-In panel \ref{fig:belowmed_read_synth_clinical}, the estimation compares executive teams with no clinical training to those with up to 25\% of the team having clinical training.
-The average treatment effect is .04 and not statistically significant.
-Thus, there is no significant difference in how these types of teams respond to the incentive change.
-In panel \ref{fig:abovemed_read_synth_clinical}, I compare hospitals with no clinically trained executives over the course of the sample to hospitals with over a quarter of their team having clinical training.
-I find that hospitals without clinically trained executives decrease readmissions by .64 (standard error .33) more than hospitals with clinically trained executives.
-Thus, the readmission rate response is fully driven by having a higher proportion of clinical executives relative to no clinical executives.
+In panel \ref{fig:belowmed_read_synth_clinical}, I compare executive teams with no clinical training to those with a fraction of clinically trained executives up to the median value, 20\%.
+With an average treatment effect of .21 (standard error .31), hospitals having this range of clinically trained executives are not driving the main finding.
+However, in panel \ref{fig:abovemed_read_synth_clinical}, I compare hospitals with no clinically trained executives to hospitals with a fraction of clinically trained executives over the median of 20\%.
+I find that hospitals without clinically trained executives decrease readmissions by .6 (standard error .14) more than hospitals with clinically trained executives.
+Thus, the readmission rate response is driven by having a higher proportion of clinical executives relative to no clinical executives.
 
@@
 
-    Any potential difference in mortality rate response also seems to be driven by executive teams with over a quarter of the executives having clinical training, as shown in Figure \ref{fig:cont_clinicalsynthdid_mort}.
-However, the standard errors of this estimation are high, potentially due to the noisiness in the mortality measure.
-The magnitude of the estimate in panel $\ref{fig:belowmed_mort_synth_clinical}$ is .07, so I can rule out differential responses by teams with less than a quarter of clinical executives.
-Taking the readmission and mortality rate results together, the effect of having clinical leadership on quality after an incentive change depends on having a sufficient proportion of the team with clinical training.
+    I present the estimation results for mortality as an outcome in Figure \ref{fig:cont_clinicalsynthdid_mort}.
+Unsurprisingly given the noisiness in the main findings for mortality, it is unclear whether there is any intensive margin effect for mortality.
+Both average treatment effect magnitudes are lower than that for the main result with high standard errors.
+Thus, the intensive margin does not seem to matter as much as it does for decreasing readmission rates.
 
@@
 
-    \section{Selective Patient Practices}
-
-    A method hospitals have been shown to employ in order to decrease readmission rates is being more selective with which patients to admit to the hospital (\cite{gupta2021impacts}).
-One hypothesis behind a more drastic quality decrease by non-clinical teams is that they may be more willing to participate in such practices, while executives with clinical training prioritize care for all types of patients.
-
-    To test this, I employ case mix index as the outcome in specification \ref{eq:clinical} to determine whether one type of hospital differentially changes patient composition after pay-for-performance incentives, where the same identification assumptions hold.
-I present the estimated difference in Figure \ref{fig:main_cmi_clinical}, comparing hospitals that never have a clinical executive to hospitals that always have a clinical executive.
-Neither type of hospital drastically changes case mix index after the incentive change, and there is no estimated differential response in complexity of patients seen in the hospital.
-Thus, the estimated difference in readmission rate response between clinical and non-clinical executives is not due to differentially changing case mix.
-
-    \begin{figure}[ht!]
-        \centering
-        \caption{Effect of Clinical Executive on Cases Mix Index}
-        \includegraphics[width=0.5\textwidth]{Objects/cmi_md_nomd_synth_graph.pdf}
-        \label{fig:main_cmi_clinical}
-    \end{figure}
-
-    \section{Internal Medicine vs. Other Specialties}
-
-    While most doctors still take a hippocratic oath to provide care with integrity and compassion, doctors with training in certain specialties may be more equipped to improve care for patients with the particular conditions relevant in this context.
-Specifically, a pulmonologist typically treats pneumonia, and a cardiologist typically treats heart failure and heart attack patients, which are the three conditions considered in the HRRP penalties.
-Both of these specialties are under the umbrella of internal medicine, which is the most granular specialty type in my data.
-
-    I limit the sample of hospitals to those who always have some type of clinical executive over the sample period.
-Then, I estimate the difference in readmission and mortality rates after pay-for-performance among hospitals who do and do not have an internal medicine clinical executive.
-I lose a significant fraction of the sample with this limitation: there are 60 hospitals that consistently have an internal medicine executive or only a different specialty executive.
-The summary statistics for these samples are in Appendix \ref{app:sumstats}.
-I present the results from this comparison in Figure \ref{fig:specialty}.
-
-    \begin{figure}[ht!]
-     \caption{Effect of Internal Medicine Training on Quality}
-     \centering
-     \begin{subfigure}[b]{0.45\textwidth}
-         \centering
-         \caption{Readmission Rates}
-         \includegraphics[width=\textwidth]{Objects/read_specialty_synth_graph.pdf}
-         \label{fig:read_spec}
-     \end{subfigure}
-     \hfill
-     \begin{subfigure}[b]{0.45\textwidth}
-         \centering
-         \caption{Mortality Rates}
-         \includegraphics[width=\textwidth]{Objects/mort_specialty_synth_graph.pdf}
-         \label{fig:mort_spec}
-     \end{subfigure}
-        \label{fig:specialty}
-    \end{figure}
-
-    While I am limited in power due to the small sample size of hospitals in this analysis, I provide evidence that hospitals with internal medicine executives may change readmission rates differentially compared to teams with clinical executives of other specialty types.
-The estimated effect of having an internal medicine executive is -.38 (standard error .24), where internal medicine executives decrease readmissions at a faster rate than other specialty types.
-The results for mortality rates are similar (still not statistically significant), but smaller in magnitude and smaller relative to the main results.
-This is probably due to the specialty-specific conditions being more relevant for the readmission rate penalties than the quality rewards.
-
     \section{Are Clinical Executives Less Profit Driven?}\label{sec:forprofit}
 
-    Another hypothesis that could explain the differential response due to clinical leadership is how profit-driven a clinician is compared to other training backgrounds.
-To demonstrate this, consider hospital behavior under the enactment of pay-for-performance policies, where hospitals face a trade-off between profit and societal benefit.
+    One hypothesis that could explain the differential response due to clinical leadership is how profit-driven a clinician is compared to other training backgrounds.
+To demonstrate this, consider hospital behavior under the enactment of pay-for-performance policies, where hospitals face a trade-off between profit and societal/patient benefit.
 Pay-for-performance incentives essentially incorporate quality directly into a hospital's profit function (\cite{dranove2011health}), whereas a fee-for-service system only incentivizes quantity.
-A hospital that places all value on profit will only care about quality when it enters into the profit function, and therefore will respond more drastically to pay-for-performance than a hospital who already cared about quality through societal benefit before the incentives.\footnote{I present the theoretical model and mathematical results in Appendix \ref{sec:model}.} Therefore, observing hospital responses to pay-for-performance incentives reveals information about their underlying preferences.
+A hospital that places all value on profit will only care about quality when it enters into the profit function, and therefore will respond more drastically to pay-for-performance than a hospital who already cared about quality through societal benefit before the incentives.
 
-    I merge all general for-profit hospitals in the AHA data to the sample of nonprofit hospitals.
+    To formalize this intuition, I specify the hospital objective function as a weighted average of profit and societal (or patient) outcomes and derive characteristics of the optimal quality.
+For simplicity, I abstract away from quality having a direct effect of quantity and price.
+Without any financial incentives to increase quality, hospital quality is decided by:
+
+    $$\max_{\theta}\hspace{2mm}\alpha\pi(\theta) + (1-\alpha) u(\theta),$$
+
+    \noindent where $\pi(\theta)$ is a profit function depending on $\theta$, and the relationship between profit and $\theta$ depends on pay-for-performance incentives.
+Additionally, $u(.)$ is extra utility gained from quality, which is increasing and concave in $\theta$, and $\alpha\in[0,1]$ captures the weight placed on profit vs.
+societal benefit.
+There is an implicit stay-open condition throughout the model.
+Taking the first order condition yields
+
+    $$(1-\alpha)u'(\theta) = -\alpha \pi'(\theta),$$
+
+    \noindent where marginal benefit equals marginal cost of increasing quality. Solving for $u'(\theta)$ and differentiating with respect to $\alpha$:
+
+    $$\frac{du'(\theta)}{d\alpha} = \frac{-1}{(1-\alpha)^2}\pi'(\theta)$$
+
+    First, assume the hospital is in a strictly fee-for-service setting, where quality benefits patients, but does not directly provide profit to the hospital through reimbursements.
+Then, assuming there is a positive cost to increasing quality, $\pi(\theta)$ is decreasing in $\theta$.
+Then $\frac{du'(\theta)}{d\alpha} > 0$, revealing that in a fee-for-service setting, the more weight placed on profit, the lower quality of care chosen by the hospital.
+
+    Second, assume the hospital is in a strictly pay-for-performance setting, where quality benefits patients and directly increases profit.
+Then, assuming the marginal financial benefit of increasing quality is greater than the marginal financial cost, $\pi(\theta)$ is increasing in $\theta$.
+Then $\frac{du'(\theta)}{d\alpha} < 0$, revealing that in a pay-for-performance setting, the more weight placed on profit, the higher quality of care chosen by the hospital.
+I combine the results found in each scenario into one response function that depends on $\alpha$, where $\theta_{p4p}$ is quality under pay-for-performance incentives and $\theta_{ffs}$ is quality under fee-for-service only.
+That is,
+
+    \begin{align*}
+        \frac{d\Delta\theta}{d\alpha}&=\frac{d(\theta_{p4p}-\theta_{ffs})}{d\alpha}\\
+        &=\frac{d\theta_{p4p}}{d\alpha}-\frac{d\theta_{ffs}}{d\alpha}\\
+        &> 0.
+    \end{align*}
+
+    Hence, this model predicts that change in quality depends on how much weight the hospital places on profit vs.
+societal benefit.
+Particularly, hospitals with more weight on profit respond more than hospitals with more weight placed on societal benefit.
+In reality, $\alpha$ is unknown, except in the case of for-profits, who have revealed that they have a high $\alpha$ by their ownership status.
+Clinically trained executives could also affect this $\alpha$, as doctors commonly take a hippocratic oath to put patients first.
+
+    \subsection{Estimation}
+
+    To investigate this hypothesis empirically, I merge all general for-profit hospitals in the AHA data to the sample of nonprofit hospitals.
 Then, I compare readmission and mortality rates of always clinical executives and never clinical executives to for-profit hospitals after the incentive change.
@@
 Thus, comparing different executive team behavior to for-profit behavior helps reveal how profit driven certain leadership teams are.
-Thus, I estimate
+Thus, I estimate the weighted synthetic difference-in-differences version of
 
@@
 
-    \noindent for two subsets of data.
-First, I limit only to for-profits and nonprofits with clinically trained executives.
-Second, I limit to for-profits and nonprofits without clinically trained executives.
-Thus, $\beta$ represents the effect of being for-profit compared to being nonprofit with a certain leadership team type.
-I show summary statistics of relevant variables for for-profit hospitals in Appendix \ref{app:sumstats}.
+    \noindent where $y_{ht}$ is one of the readmission or mortality rate outcomes discussed in Section \ref{sec:hrrp}, for-profit$_{ht}$ is an indicator for having for-profit ownership status, and $\gamma$ and $\delta$ are hospital and time fixed effects.
+In my estimation, nonprofits in the comparison group are more heavily weighted if their pre-trends are more similar to for-profits, and time periods are weighted to balance pre and post trends for for-profit hospitals.
+These weights follow the synthetic difference-in-differences approach, which is more robust to the parallel trends assumption.
+
+    I create two subsets of data to compare the response of for-profits to the response of the subsets of nonprofits based on executive team.
+First, I limit only to for-profits and nonprofits that always have clinically trained executives.
+Second, I limit to for-profits and nonprofits that never have clinically trained executives.\footnote{I assume that the sample selection of stable leadership teams is not correlated with the policy changes.
+Please see Section \ref{sec:sig_man} for the analysis and discussion of this assumption.} I show summary statistics of relevant variables for for-profit hospitals in Appendix \ref{app:sumstats}.
 Again, I employ a synthetic difference-in-differences strategy to rely less heavily on a strong parallel trends assumption for different types of hospitals.
@@
     The identification assumptions I make here are largely similar to those in Section \ref{sec:clinical}, only now incorporate for-profit hospitals.
-First, I assume that, absent the change in incentives, for-profits and the relevant comparison group would have had parallel trends in outcomes.
-Second, I assume that the stability of nonprofit leadership teams is not correlated with the 2012 policy changes.
-I explore this assumption in Section \ref{sec:sig_man}.
+First, I assume that, absent the change in incentives, the weighted composition of for-profits and the relevant comparison group would have had parallel trends in outcomes.
 I also assume that there is no anticipation of the policy changes prior to 2012, which is reasonable as the rules of the policies were announced in October of 2011.
 Finally, I assume no other changes correlated with both hospital type (for-profit, nonprofit with and without clinically trained executives) and quality occurred in conjunction with the 2012 pay-for-performance policies.
+
+    The estimates and graphical representation of the effect of being for-profit is shown in Figure \ref{fig:read_synth_plot}.
+In panel \ref{fig:read_synth_plotb}, the average treatment effect shows the effect of being for-profit compared to nonprofit with clinical executives on readmission rates after pay-for-performance changes.
+I find that for-profits decrease readmission rates by .3 (standard error .12) more than nonprofits with clinically trained executives.
+That is, for-profits respond more aggressively to the incentive change than nonprofits with clinical leadership.
+This difference is nearly identical to the difference between clinical and non-clinical teams shown in Section \ref{sec:clinical}.
+So, unsurprisingly, panel \ref{fig:read_synth_plotc} shows no such difference when comparing for-profits to nonprofits without clinically trained executives.
+Not only is the estimate not statistically significant, but the magnitude indicates a precise zero.
+That is, a lack of clinically trained executives leads nonprofits to act like for-profits in terms of readmission rates when responding to this incentive change.
 
@@
 
-    The estimates and graphical representation of the effect of being for-profit is shown in Figure \ref{fig:read_synth_plot}.
-In panel \ref{fig:read_synth_plotb}, the average treatment effect shows the effect of being for-profit compared to nonprofit with clinical executives on readmission rates after pay-for-performance changes.
-While readmission rates trend similarly prior to the incentive change, for-profits decrease readmission rates by .3 more than nonprofits with clinically trained executives.
-That is, for-profits respond more aggressively to the incentive change than nonprofits with clinical leadership.
-This difference is nearly identical to the difference between clinical and non-clinical teams shown in Section \ref{sec:clinical}.
-So, unsurprisingly, panel \ref{fig:read_synth_plotc} shows no such difference when comparing for-profits to nonprofits without clinically trained executives.
-Not only is the estimate not statistically significant, but the magnitude indicates a precise zero.
-That is, a lack of clinically trained executives leads nonprofits to act like for-profits in terms of readmission rates when responding to this incentive change.
-
     Next, I present estimated differences in mortality rates between different hospital types in Figure \ref{fig:mort_synth_plot}.
@@
 While I cannot rule out a lack of differential response, the documented noisiness of mortality measures (\cite{mackenzie2016measuring}) may mean that the true effect is different from zero.
-However, the estimated difference in mortality between for-profits and nonprofits without clinically trained executives is certainly zero with a amgnitude of .03, shown in panel \ref{fig:mort_synth_plotc}.
+However, the estimated difference in mortality between for-profits and nonprofits without clinically trained executives is certainly zero with a magnitude of .03, shown in panel \ref{fig:mort_synth_plotc}.
 
@@
 
-    Given that having clinically trained executives affects behavior, an important distinction is whether the existing clinical leaders \textit{reveal} the underlying objectives of the hospital, or \textit{manage} the hospital differently.
+    Given that having clinically trained executives does affect hospital response to financial incentives on quality, an important distinction is whether the existing clinical leaders \textit{reveal} the underlying objectives of the hospital, or \textit{manage} the hospital differently.
 Up to this point, I have limited the sample of nonprofit hospitals to those that don't change their propensity to hire a clinically trained executive over the sample period, which combines any signaling and managing effects.
 However, under the assumption that hiring or firing these types of executives is not correlated with the pay-for-performance policy changes, such changes can be leveraged to decompose the effect into signaling versus managing.
-Therefore, I first investigate leadership changes in response to the policy change.
+Since this decomposition depends heavily on the assumption that clinical executive hiring is not endogenous, I first investigate leadership changes in response to the policy change.
+I provide details of the decomposition estimation in Section \ref{sec:decomp}
 
@@
 
-    I have discussed the assumptions of parallel trends, no anticipation, and no other confounding events in both Sections \ref{sec:clinical} and \ref{sec:forprofit}.
-These assumptions are also necessary here.
-Further, this decomposition analysis depends heavily on leadership team changes not endogenously occurring due to pay-for-performance policies.
-Up to this point, that assumption was important because of the way I selected the sample.
-However, this decomposition depends directly on leveraging such leadership changes.
-Thus, I now explore this assumption in detail.
+    Decomposing the observed difference in response based on timing is dependent on executive team changes not being correlated with the financial incentive changes.
 First, I present raw means over time that represent the fraction of hospitals who hire a clinical executive, fire a clinical executive, or have any change in their propensity to have a clinical executive in Figure \ref{fig:change_means}.
@@
 
-    Next, I assess whether this assumption is reasonable by analyzing how executive team changes are occurring over time for certain groups of hospitals. I estimate:
+    Next, I assess whether this assumption is reasonable by analyzing how executive team changes are occurring over time for certain groups of hospitals.
+I estimate the following two-way-fixed-effects specification:
 
@@
 And finally, I leave out the Program Exposed variable to simply see how changes are correlated with given years for all hospitals.
-The outcome change$_{ht}$ is an indicator for whether the hospital changes the number of clinical executives in a given year.
+The outcome variable change$_{ht}$ is an indicator for whether the hospital changes the number of clinical executives in a given year.
 The estimates from this analysis are presented in Figure \ref{fig:change_analysis}.
@@
 
-     \subsection{Estimation and Findings}
+     \subsection{Estimation and Findings}\label{sec:decomp}
 
-    To disentangle the two, I carefully define the ``treat" and comparison group of hospitals in the following estimating equation.
-For clarity, I present the specification details in Figure \ref{fig:decomp_spec}.
+    To disentangle clinical executives as signals of underlying hospital preferences or a different type of hospital manager, I carefully define the ``treat" and comparison group of hospitals in the following estimating equation.
+For clarity, I also present the specification details in Figure \ref{fig:decomp_spec}.
 If clinically trained leaders are simply signals of underlying objectives, it shouldn't matter when hospitals have hired a clinical executive.
@@
 
+    As in the previous analyses, I estimate the synthetic difference-in-differences version of each of these specifications.
+That is, I weight control group units more heavily when their pre-trends are similar to the treated group units, and I weight more heavily time periods that balance pre and post period trends for the control group.
+I have discussed the assumptions of parallel trends, no anticipation, and no other confounding events in both Sections \ref{sec:clinical} and \ref{sec:forprofit}.
+These assumptions are also necessary here.
+I also assume that executive team changes are not endogenous with the policy changes, which I explore in detail in Section \ref{sec:changes}.
+
     The estimates of each decomposition specification are shown in Table \ref{tab:MD_noMD_readmort_decomp_synth}.
@@
 
+    \section{Heterogeneity by Physician Specialty}
+
+    While most doctors take a hippocratic oath to provide care with integrity and compassion, doctors with training in certain specialties may be more equipped to improve care for certain patients.
+Since the pay-for-performance incentive programs directly target three conditions, physicians with knowledge of these conditions may be especially suited to respond to the policies.
+Specifically, a pulmonologist typically treats pneumonia, and a cardiologist typically treats heart failure and heart attack patients, which are the three conditions considered in the HRRP penalties.
+Both of these specialties are under the umbrella of internal medicine, which is the most granular specialty type in my data.
+
+    I limit the sample of hospitals to those who always have some type of clinical executive over the sample period.
+Then, I estimate the difference in readmission and mortality rates after pay-for-performance among hospitals who do and do not have an internal medicine clinical executive.
+I lose a significant fraction of the sample with this limitation: there are 60 hospitals that consistently have an internal medicine executive or only a different specialty executive.
+The summary statistics for these samples are in Appendix \ref{app:sumstats}.
+I present the results from this comparison in Figure \ref{fig:specialty}.
+
+    \begin{figure}[ht!]
+     \caption{Effect of Internal Medicine Training on Quality}
+     \centering
+     \begin{subfigure}[b]{0.45\textwidth}
+         \centering
+         \caption{Readmission Rates}
+         \includegraphics[width=\textwidth]{Objects/read_specialty_synth_graph.pdf}
+         \label{fig:read_spec}
+     \end{subfigure}
+     \hfill
+     \begin{subfigure}[b]{0.45\textwidth}
+         \centering
+         \caption{Mortality Rates}
+         \includegraphics[width=\textwidth]{Objects/mort_specialty_synth_graph.pdf}
+         \label{fig:mort_spec}
+     \end{subfigure}
+        \label{fig:specialty}
+    \end{figure}
+
+    While I am limited in power due to the small sample size of hospitals in this analysis, I provide evidence that hospitals with internal medicine executives may change readmission rates differentially compared to teams with clinical executives of other specialty types.
+The estimated effect of having an internal medicine executive is -.38 (standard error .24), where internal medicine executives decrease readmissions at a faster rate than other specialty types.
+The results for mortality rates are similar (still not statistically significant), but smaller in magnitude and smaller relative to the main results.
+This is probably due to the specialty-specific conditions being more relevant for the readmission rate penalties than the quality rewards.
+
     \section{Conclusion}
@@
     While previous literature has documented correlations between executive characteristics and firm behaviors/performance, research is still limited on this topic in the health care sector.
+Health care managers are tasked with maintaining a business while also prioritizing patient care, which is why existing findings on manager effects may not extend to this setting.
 This paper investigates whether having executive leaders with clinical training impacts hospital behaviors and outcomes, particularly in response to a policy change that financially incentivizes quality.
-I collect data on the names of hospital executives from 2010-2014 from publicly available Tax Form 990s, which are completed by all nonprofits in the US.
-I merge this data with existing hospital and physician characteristics, and aggregate to the hospital level.
-Thus, I utilize a unique data set of nonprofit US hospitals that captures the proportion of clinical executives, specialties of clinical executives, hospital characteristics, and hospital quality metric outcomes.
+I utilize a unique data set of nonprofit US hospitals that captures the proportion of clinical executives, specialties of clinical executives, hospital characteristics, and hospital quality metric outcomes.
 
@@
 Thus, I leverage this exogenous policy change and estimate how clinically trained executives affect readmission and mortality rates in response to the change using a synthetic difference-in-differences strategy.
-Conditional on parallel trends holding, I find that having clinical training mitigates the response to the incentive change.
-That is, while all hospitals reduce readmissions after the policies take place, non-clinical teams reduce readmissions at a fast rate than clinical teams.
+I find that hospitals without clinically trained executives improve quality more than hospitals without clinically trained executives after the incentive change.
 
-    I investigate several mechanisms by which this difference might occur.
+    I investigate several hypotheses that arise given the estimated difference in response.
 First, this result is not driven by non-clinical teams adjusting their patient complexity.
@@
 patient benefit that drive the difference.
-Additionally, hospitals with clinically trained executives in specialties relevant to the policies improve quality more.
 Finally, I leverage exogenous changes in executive teams to show that clinically trained executives are not simply a signal of underlying preferences, but manage the hospital differently.
 
-    These findings have significant policy implications, as it informs the way different types of hospital leadership respond to policy changes.
-I provide evidence that clinical executives value quality even before quality is incentivized, and that having particular specialties in leadership might matter for quality of care.
+    In US health care, a very common policy goal is to improve quality of care without increasing costs.
+However, up to this point, it was unclear whether hospital leaders might play a role in attaining this.
+In this paper, I show that having hospital managers with clinical training does affect the way hospitals respond to incentives.
+Importantly, my results point to clinically trained executives already prioritizing quality before incentives were put in place.
+My findings have implications for policy makers considering financial incentives on quality, and even point to the potential of incentivizing the presence of clinically trained hospital leaders.
 
@@
 
-\subsection{Matched vs. Unmatched Not-for-Profits}\label{app:matched}
+\subsection{Matched vs. Unmatched Non-Profits}\label{app:matched}
 
@@
 
-\section{Modeling Hospital Response to Pay-for-Performance}\label{sec:model}
+    \section{Selective Patient Practices}\label{app:casemix}
 
-    I model hospital behavior under two simplified worlds: one where quality does not directly affect profit, and one where it does.
-Pay-for-performance incentives essentially incorporate performance directly into a hospital's profit function (\cite{dranove2011health}).
-I directly model performance as quality of care, which is inversely related to readmission and mortality rates.
-While the literature speculates about the true form of NFP objective functions, we can generally think of hospitals choosing some linear combination of profit and societal benefit, without knowing how much weight the hospital places on each.
-Intuitively, a hospital that places only weight on profit (for-profit type) will only care about quality when it enters into the profit function, and therefore will respond more drastically to pay-for-performance than a hospital who already cared about quality through societal benefit before the incentives.
-Therefore, hospital responses to pay-for-performance incentives reveal information about their underlying preference on profit vs.
-societal outcomes.
+    A method hospitals have been shown to employ in order to decrease readmission rates is being more selective with which patients to admit to the hospital (\cite{gupta2021impacts}).
+One hypothesis behind a more drastic quality decrease by non-clinical teams is that they may be more willing to participate in such practices, while executives with clinical training prioritize care for all types of patients.
 
-    To formalize this intuition, I specify the hospital objective function as a weighted average of profit and societal (or patient) outcomes and derive characteristics of the optimal quality.
-For simplicity, I abstract away from quality having a direct effect of quantity and price.
-Without any financial incentives to increase quality, hospital quality is decided by:
+    To test this, I employ case mix index as the outcome in specification \ref{eq:clinical} to determine whether one type of hospital differentially changes patient composition after pay-for-performance incentives, where the same identification assumptions hold.
+I present the estimated difference in Figure \ref{fig:main_cmi_clinical}, comparing hospitals that never have a clinical executive to hospitals that always have a clinical executive.
+Neither type of hospital drastically changes case mix index after the incentive change, and there is no estimated differential response in complexity of patients seen in the hospital.
+Thus, the estimated difference in readmission rate response between clinical and non-clinical executives is not due to differentially changing case mix.
 
-    $$\max_{\theta}\hspace{2mm}\alpha\left[R - c_{\theta}(\theta) \right] + (1-\alpha) u(\theta),$$
-
-    \noindent where $R$ is net revenue, $c_{\theta}(.)$ is an increasing cost function specific to quality $\theta$, $u(.)$ is extra utility gained from quality, which is increasing and concave in $\theta$, and $\alpha\in[0,1]$ captures the weight placed on profit vs.
-societal benefit.
-There is an implicit stay-open condition throughout the model.
-Taking the first order condition yields
-
-    $$(1-\alpha)u'(\theta) = \alpha c_{\theta}'(\theta),$$
-
-    \noindent where marginal benefit equals marginal cost of increasing quality. Solving for $u'(\theta)$ and differentiating with respect to $\alpha$:
-
-    $$\frac{du'(\theta)}{d\alpha} = \frac{1}{(1-\alpha)^2}c_{\theta}'(\theta) > 0.$$
-
-    \noindent Thus, by the Implicit Function Theorem,
-
-    $$\frac{d\theta}{d\alpha} = \frac{du'(\theta)/d\alpha}{du'(\theta)/d\theta} < 0.$$
-
-    \noindent That is, the more weight on profit, the lower quality of care chosen by the hospital.
-
-    In a world with pay-for-performance incentives, which can either look like benefits from high quality (such as HVBP) or penalties for low quality (as in HRRP), quality directly affects revenue.
-Thus, $R(\theta)$ is an increasing function of $\theta$, and I assume that the marginal financial benefit of increasing quality is greater than the marginal cost ($R'(\theta)\geq c_{\theta}'(\theta))$.
-Thus, the new first order condition yields:
-
-    $$u'(\theta) = \frac{\alpha}{1-\alpha} \left[c_{\theta}'(\theta)-R(\theta)\right] \leq 0.$$
-
-    \noindent Taking the derivative of this with respect to $\alpha$,
-
-    $$\frac{du'(\theta)}{d\alpha} = \frac{1}{(1-\alpha)^2}c_{\theta}'(\theta).$$
-
-    Therefore, again using the Implicit Function Theorem,
-
-    $$\frac{d\theta}{d\alpha} = \frac{du'(\theta)/d\alpha}{du'(\theta)/d\theta}\geq0.$$
-
-    That is, all else equal, in a world with financial incentives on quality, quality is increasing with more weight placed on profit.
-I combine the results found in each scenario into one response function that depends on $\alpha$, where $\theta_2$ is quality under pay-for-performance incentives and $\theta_1$ is quality with no financial incentive to quality.
-That is,
-
-    \begin{align*}
-        \frac{d\Delta\theta}{d\alpha}&=\frac{d(\theta_2-\theta_1)}{d\alpha}\\
-        &=\frac{d\theta_2}{d\alpha}-\frac{d\theta_1}{d\alpha}\\
-        &\geq 0.
-    \end{align*}
-
-    Hence, this model predicts that change in quality depends on how much weight the hospital places on profit vs.
-societal benefit.
-Particularly, hospitals with more weight on profit respond more than hospitals with more weight placed on societal benefit.
-In reality, $\alpha$ is unknown, except in the case of for-profits, who have revealed that they have a high $\alpha$ by their ownership status.
-However, not-for-profit behavior on average does not bring much clarity about their objectives.
-An observable and economically relevant characteristic of a hospital that could affect $\alpha$ is composition of their executive leadership team.
-Thus, I compare the response of for-profits and not-for-profits with different leadership teams to bring clarity on how this might point to underlying $\alpha$.
+    \begin{figure}[ht!]
+        \centering
+        \caption{Effect of Clinical Executive on Case Mix Index}
+        \includegraphics[width=0.5\textwidth]{Objects/cmi_md_nomd_synth_graph.pdf}
+        \label{fig:main_cmi_clinical}
+    \end{figure}
 
@@
 
-\section{Synthetic Difference-in-Difference Coefficient Tables}
-
-While I present the graphical representation of results in the main text, I also present a table of coefficients and standard errors for readmission and mortality rates in Tables \ref{tab:forprofit_readmort_synth} and \ref{tab:MD_noMD_readmort_synth}.
-
-\import{Tables}{forprofit_readmort_synth.tex}
-
-\import{Tables}{MD_noMD_readmort_synth.tex}
-
 \end{document}

### 2024-08-21 Tables/condition_synthdid_table.tex
add/edit appendix tables and figures

\begin{table}[ht!]

\caption{\label{tab:results_by_condition}Effect of Clinically Trained Executive on Condition-Specific Readmission and Mortality Rates}
\centering
\begin{tabular}[t]{lccclcc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{3}{c}{Readmission Rates} & \multicolumn{3}{c}{Mortality Rates} \\
\multicolumn{1}{c}{ } & \multicolumn{1}{c}{AMI} & \multicolumn{1}{c}{HF} & \multicolumn{1}{c}{Pneum.} & \multicolumn{1}{c}{AMI} & \multicolumn{1}{c}{HF} & \multicolumn{1}{c}{Pneum.} \\
\cmidrule(l{3pt}r{3pt}){2-2} \cmidrule(l{3pt}r{3pt}){3-3} \cmidrule(l{3pt}r{3pt}){4-4} \cmidrule(l{3pt}r{3pt}){5-5} \cmidrule(l{3pt}r{3pt}){6-6} \cmidrule(l{3pt}r{3pt}){7-7}
 & (1) & (2) & (3) & (4) & (5) & (6)\\
\midrule
Ever Clinical Exec x Post Program & 0.25 & 0.14 & 0.46$^{**}$ & -0.13 & 0.03 & 0.3\\
 & (0.18) & (0.18) & (0.14) & (0.19) & (0.13) & (0.19)\\
 &  &  &  &  &  & \\
Hospital FE & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
Year FE & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
\addlinespace
Observations & 1,135 & 1,640 & 1,635 & 1,300 & 1,640 & 1,635\\
\bottomrule
\multicolumn{7}{l}{\textsuperscript{} Standard errors are clustered at the hospital level.}\\
\multicolumn{7}{l}{\textsuperscript{} Signif. codes: 0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1}\\
\end{tabular}
\end{table}

### 2024-08-21 Tables/main_twfe_match_tab.tex
add/edit appendix tables and figures

\begin{table}[ht!]

\caption{\label{tab:main_twfe_match}Effect of Clinically Trained Executive, TWFE and Matched Sample}
\centering
\begin{tabular}[t]{lcccl}
\toprule
\multicolumn{1}{c}{\underline{ }} & \multicolumn{2}{c}{\underline{TWFE}} & \multicolumn{2}{c}{\underline{Matched Sample}} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
\multicolumn{1}{c}{ } & \multicolumn{1}{c}{Read} & \multicolumn{1}{c}{Mort} & \multicolumn{1}{c}{Read} & \multicolumn{1}{c}{Mort} \\
\cmidrule(l{3pt}r{3pt}){2-2} \cmidrule(l{3pt}r{3pt}){3-3} \cmidrule(l{3pt}r{3pt}){4-4} \cmidrule(l{3pt}r{3pt}){5-5}
 & (1) & (2) & (3) & (4)\\
\midrule
Ever Clinical Exec x Post Program & 0.42$^{**}$ & 0.11 & 0.65$^{*}$ & 0.03\\
 & (0.15) & (0.15) & (0.3) & (0.4)\\
 &  &  &  & \\
Hospital FE & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
Year FE & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
\addlinespace
Observations & 2,387 & 2,381 & 1,025 & 1,023\\
\bottomrule
\multicolumn{5}{l}{\textsuperscript{} Standard errors are clustered at the hospital level.}\\
\multicolumn{5}{l}{\textsuperscript{} Signif. codes: 0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1}\\
\end{tabular}
\end{table}

### 2024-08-21 Paper/paper2.tex
Updates from Overleaf

@@
 
+    \subsection{Intensive Margin}
+
     Additionally, I estimate the intensive margin of clinical leadership: whether the fraction of clinically trained executives matters for response to financial incentive changes.
@@
 I find that hospitals without clinically trained executives decrease readmissions by .6 (standard error .14) more than hospitals with clinically trained executives.
-Thus, the readmission rate response is driven by having a higher proportion of clinical executives relative to no clinical executives.
+Thus, the readmission rate response is driven by having a higher proportion of clinical executives relative to no clinical executives.\footnote{This result holds even when limiting the sample to hospitals with a similar total executive team sizes (Appendix \ref{app:binned})}
 
@@
 
-    \subsection{Leadership Change Analysis} \label{sec:changes}
+    \subsection{Analysis of Executive Team Changes} \label{sec:changes}
 
@@
 
-     \subsection{Estimation and Findings}\label{sec:decomp}
+     \subsection{Estimation}\label{sec:decomp}
 
@@
 
+\newpage
+
  \section{Data}\label{appendixdata}
@@
 
-I use the Nonprofit Explorer API to access the archive of NFP tax form 990s.
-At the time of writing this, information on using version 2 of the API can be found at \hyperlink{https://projects.propublica.org/not-for-profits/api}{https://projects.propublica.org/not-for-profits/api}.
+I use the API of ProPublica's Nonprofit Explorer to access the archive of nonprofit Tax Form 990s.\footnote{At the time of writing this, information on using version 2 of the API can be found at \hyperlink{https://projects.propublica.org/not-for-profits/api}{https://projects.propublica.org/not-for-profits/api}.} I extract the Employee Identification Number (EIN) for all nonprofits categorized with a National Taxonomy of Exempt Entities (NTEE) code of E20 (hospital), E21 (community health system), and E22 (general hospital).
+There are 5,588 EINs total in this list.
 
-There are over 1.5 million not-for-profit entities in the US, making it crucially important to be able to filter by type of entity before analyzing any PDFs.
-The API allows this by filtering a query based on National Taxonomy of Exempt Entities (NTEE) code.
-I query only not-for-profits categorized as E20 (hospitals), E21 (community health systems), and E22 (general hospitals).
-The API has a pagination limit of 100, meaning I can only pull information on 100 hospitals at a time.
-Therefore, I filter the query further to only consider one state at a time.
-The only state that has more than 100 entities registered is California, and thus I subset the California query even further by names that include the word "hospital" and names that don't.
-I combine all of these subsets and have information on each not-for-profits Employee Identification Number.
-There are 5,588 EINs total in this list.
-This acts as a list of entities for which I can pull more information.
-
-I loop through the list of EINs found in the previous step and query more detailed information from the API on that specific EIN.
+I query more detailed information from the API on each hospital EIN.
 I save the name, secondary name, state, and zip code, all of which do not vary by year.
-I also save each year's URL link to the Tax Form 990 PDF.
+I also record URL links to the Tax Form 990 PDFs, which are unique in each year.
 For the sake of a comprehensive data set, I keep years 2006-2020 (I later limit to 2010-2014 when focusing on pay-for-performance initiatives in 2012).
 Thus, I finish this step with a panel data set of EIN characteristics and PDF locations.
-Importantly, there are multiple types of Tax Form 990s depending on the size of the not-for-profit.
-In many cases, one not-for-profit has at least two different forms filed in a given year.
-I filter out any EIN-years for which there are no PDFs on file.
-The data on PDF locations contains 4,012 EINs and 61,363 EIN-year-tax forms.
+Importantly, there are multiple types of Tax Form 990s depending on the size of the nonprofit.
+In many cases, one nonprofit has at least two different forms filed in a given year.
+I filter out any EIN-years for which there are no PDFs available.
+There are numerous foundations or auxiliary firms with the purpose of raising funds for the hospital, but do not provide services to patients.
+I filter out any not-for-profit with ``foundation" or ``auxiliary" in the name.
+I also filter out various specialty centers such as hospice or cancer centers.
 
-It is necessary to link these not-for-profits with other sources of data to recover penalties from HRRP, payments from HVBP, bed size, and outcomes of interest.
-The ideal hospital data set to match to is American Hospital Association (AHA) Survey, which contains hospital characteristics and Medicare ID number.
-However, an EIN to AHA ID crosswalk does not exist.
-Therefore, I take a conservative approach to matching EINs to American Hospital Association (AHA) ID based on hospital name and location.
-First, I will discuss limitations and cleaning of the AHA data and tax form data.
+For my analysis, I need to link these nonprofits with other sources of data to recover penalties from HRRP, payments from HVBP, bed size, and outcomes of interest.
+I match to the American Hospital Association (AHA) Survey, which contains hospital characteristics and Medicare ID number, and will easily merge to other data sets used for hospital information.
+An EIN to AHA ID crosswalk does not exist.
+Therefore, I take a conservative approach to matching EINs to AHA ID based on hospital name and location.
 
-In the AHA data, I filter only to hospitals in the contiguous US, Alaska, and Hawaii (excluding places like Puerto Rico), classified as not-for-profit or state/community, and those that are general acute care.
-I also filter out any hospitals who weren't present in the data (or change system ID) in 2009-2015, meaning they either closed or were acquired.
+In the AHA data, I keep general acute care hospitals in the contiguous US, Alaska, and Hawaii (excluding places like Puerto Rico) that are classified as nonprofit or state/community.
+I remove hospitals who weren't present in the data or change system ID in 2009-2015, meaning they either closed or were acquired.
 Due to the survey nature of this data, a hospital name may look slightly different from one year to the next.
 For example, ``Waldo County General Hospital" is also ``Waldo County General Hospital Maine Health".
-Further, zip codes may change by one or two digits, making them unreliable to match based on.
+Further, zip codes often change by one or two digits, making them unreliable to match based on.
 To deal with this, I first keep only unique AHA ID, name, zip, state, and system name combinations.
@@
 
-I consider which not-for-profit entities are not likely to be hospitals and drop them.
-There are numerous foundations or auxiliary firms with the purpose of raising funds for the hospital, but do not provide services to patients.
-I filter out any not-for-profit with "foundation" or "auxiliary" in the name.
-I also filter out various specialty centers that fell into the general hospital category, such as hospice or cancer centers.
-
-I then proceed matching based on names in multiple layers.
-I focus on exact string matches, so I remove all spaces and common characters that could cause mismatches such as \&, ', -, and inc.
-Next, I take each AHA name and look for exact matches in a not-for-profit's first or secondary name for not-for-profits in the same state as the AHA hospital.
+I proceed matching based on names in several steps.
+I focus on exact string matches, so I remove all spaces and common characters that could cause mismatches such as \&, ', -, and ``inc".
+I loop through each AHA ID, limit to nonprofits in the same state, and extract any exact name matches.
 When an exact match is found, I record the link between AHA ID and EIN.
-In this first layer of matching, 860 hospitals in the AHA data are linked to an EIN, equivalent to 31\% of AHA not-for-profit hospitals in the sample.
-
-In the next layer of matching, I remove common words such as "healthcare", "regional", "hospital", etc.
+In this first layer of matching, 860 hospitals in the AHA data are linked to an EIN, equivalent to 31\% of nonprofit hospitals in the sample.
+In the next layer of matching, I remove common words such as ``healthcare", ``regional", ``hospital", etc.
 That way if there are subtle differences in names, removing common words may allow for an exact match.
-Again, I take each AHA hospital name and look for exact matches in the not-for-profits within the same state.
+Again, I take each AHA hospital name and look for exact matches in the nonprofits in the same state.
 This adds an additional 90 hospital matches, accounting for a total of 34.5\% of AHA hospitals.
@@
 From google searches, I identify an additional 300 EIN-AHA ID matches.
+yielding a total of 1200 EINs.
+
+While it would be ideal to analyze all nonprofit hospitals, a sample size of 852 firms is relatively large in the hospital executive literature.
+I present a comparison of in-sample and out-of-sample nonprofits.
+I present means and standard deviations of these measures in Table \ref{tab:nfp_sample_compare}.
+The in-sample nonprofits are slightly larger in terms of beds and patients seen.
+They are also slightly more likely to be penalized or receive payments under the pay-for-performance incentives.
+However, average readmission and mortality rates are very similar, as well as case mix index.
+The largest difference in the two samples is that I under-represent nonprofit hospitals in systems.
+This is due to the nature of the tax form 990s, where systems often file one tax form for the entire system, making it difficult to discern the true managers of a specific hospital.
+Because of this, I drop hospitals from the sample who only have leadership information from a system tax form.
+
+\import{Tables}{NFP_sample_comparison.tex}
 
@@
 
-One tricky aspect of the NonProfit Explorer API is that, only in some cases, if two forms are present for an EIN, year, only the first one (which is typically not the one with the relevant information) is pulled.
+One aspect of the NonProfit Explorer API is that, only in some cases, if two forms are present for an EIN, year, only the first one (which is typically not the one with the relevant information) is pulled.
 Therefore, for some hospitals, a couple years will have gaps in text extraction data.
@@
 
-The form of the text data is a data frame with one column, where each line of text is saved in a different row.
+The form of the extracted text data is a data frame with one column, where each line of text is saved in a different row.
 Typically on the same page as the names and positions is a list of the highest compensated employees and their compensation.
@@
 That is, I remove all board members from the data.
-I then create hospital level indicators based on the names, titles and positions associated with the hospital: the number of MD executives, the number of total executives, and whether the hospital employs a CMO.
-In Figure \ref{fig:state_doc}, I show the percent of hospitals with a clinical executive in each state.
-This figure verifies that clinical executives are not concentrated in a small region geographically.
+I now seek to match people in this data set to existing physicians in the National Plan and Provider Enumeration System (NPPES) data, where all physician National Provider Identifiers (NPI) are recorded.
+For each executive (not just those identified as doctors), I record the number of name matches found in the NPPES data.
+There are several steps to identifying physicians with a true match.
+First, if an executive self-identifies as a doctor and has a unique match in NPPES, I record the unique NPI number for that executive.
+Second, if an executive self-identifies as a doctor and has multiple name matches in NPPES, I manually search online for information on the executive, and find the unique NPI most likely associated with the executive.
+Finally, for any executives who do not self-identify as a physician, but have name matches in the NPPES data, I manually research them and record the unique NPI number if they are a physician.
 
-\begin{figure}[ht!]
+The data is now an executive-year level data set with information on clinical training and specialty when relevant.
+I then create hospital level indicators based on the names, titles and positions associated with the hospital: the number of clinically trained executives, the number of executives with particular specialties, the number of total executives, and whether the hospital employs a CMO.
+
+\subsection{Executive Team Changes}\label{app:changes}
+
+For the main analysis, I limit to hospitals who either always have a clinically trained executive or never have a clinically trained executive.
+In a supplemental analyses, I leverage changes to hospital propensity to hire a clinically trained executive.
+I explore the exogeneity of such changes in Section \ref{sec:changes}.
+Additionally, I present the number of hospitals who change their propensity to hire a clinically trained executive at different times in Table \ref{tab:change_timing}.
+I break the timing up into pre-2012, 2012, and post-2012.
+I show the number of hospitals who have clinically trained executives in different time combinations.
+The number of hospitals that only have an MD executive in 2012, but at no other time, is only 6.
+The majority of hospitals (187) who ever have a clinical executive have one in every time period.
+
+\import{Tables}{change_num_table_2012.tex}
+
+\subsection{Merging to Other Hospital Data}
+
+Using the hospital Medicare ID number found in the AHA survey, I merge the executive information to various publicly available hospital data sets.
+For information on penalties or payments given by the pay-for-performance policies, I use the Hospital Cost Report Information System (HCRIS) data.
+For 30-day readmission rates, 30-day mortality rates, and the number of patients in the relevant conditions, I merge in the CMS Hospital Compare data.
+Finally, to investigate the role of selective patient practices I use a case mix index variable, which comes from the CMS Case Mix Index files.
+
+\subsection{Additional Summary Statistics}\label{app:sumstats}
+
+\begin{figure}[p!]
     \centering
@@
     \label{fig:state_doc}
+    \caption*{\footnotesize{\textit{Notes:} This graph shows the percent of hospitals with a clinically trained executive in each state.}}
 \end{figure}
 
-\section{Additional Summary Statistics}\label{app:sumstats}
+\newpage
 
@@
 
+\newpage
+
 \import{Tables}{specialty_sumstats.tex}
 
-\subsection{Matched vs. Unmatched Non-Profits}\label{app:matched}
-
-The sample of all NFP hospitals that I use for analysis consists of 852 hospitals.
-In the AHA data, there are 2,258 additional NFP hospitals that either were not matched to tax forms, did not exist in the tax form data, or did not have executive names in the tax forms.
-While it would be ideal to analyze all NFPs, a sample size of 852 firms is relatively large in the hospital executive literature.
-
-In this section, I compare the in-sample NFPs to the out-of-sample NFPs along observable measures.
-I present means and standard deviations of these measures in Table \ref{tab:nfp_sample_compare}.
-The in-sample NFPs are slightly larger in terms of beds and patients seen.
-They are also slightly more likely to be penalized or receive payments under the pay-for-performance incentives.
-However, average readmission and mortality rates are very similar, as well as case mix index.
-The largest difference in the two samples is that I under-represent NFP hospitals in systems.
-This is due to the nature of the tax form 990s, where systems often file one tax form for the entire system, making it difficult to discern the true managers of a specific hospital.
-Because of this, I drop hospitals from the sample who only have leadership information from a system tax form.
-
-\import{Tables}{NFP_sample_comparison.tex}
-
-\subsection{Executive Team Changes}\label{app:changes}
-
-I present the number of hospitals who change their propensity to hire a clinically trained executive at different times in Table \ref{tab:change_timing}.
-I break the timing up into pre-2012, 2012, and post-2012.
-I show the number of hospitals who have MDs in different time combinations.
-The number of hospitals that only have an MD executive in 2012, but at no other time, is only 6.
-The majority of hospitals (187) who ever have a clinical executive have one in every time period.
-
-\import{Tables}{change_num_table_2012.tex}
+\newpage
 
@@
 
-    \section{Selective Patient Practices}\label{app:casemix}
+    \section{Supplemental Analyses}
+
+    \subsection{Binned Treatment}\label{app:binned}
+
+    In Section \ref{sec:clinical}, I present intensive margin results by binning hospitals according to the fraction of their executive team with clinical experience.
+There, I find that the intensive margin does matter for how hospitals change readmissions after the incentive change.
+One concern could be that this result is driven by a difference in the total size of hospital executive teams, the denominator of that fraction.
+In this section, I use the same estimating equation, but I limit to only nonprofit hospitals with 7 or less executives in total, removing outliers of executive team size.
+The difference in readmission rate response is still driven by teams with a fraction of clinical executives above the median.
+
+        \begin{figure}[ht!]
+     \caption{Effect of Clinical Training on Readmission Rates, Binned Treatment}
+     \centering
+     \begin{subfigure}[b]{0.45\textwidth}
+         \centering
+         \caption{$\rho \leq \eta$}
+         \includegraphics[width=\textwidth]{Objects/cont_belowmedread_md_nomd_size_synth_graph.pdf}
+         \label{fig:belowmed_read_synth_clinical}
+     \end{subfigure}
+     \hfill
+     \begin{subfigure}[b]{0.45\textwidth}
+         \centering
+         \caption{$\rho > \eta$}
+         \includegraphics[width=\textwidth]{Objects/cont_abovemedread_md_nomd_size_synth_graph.pdf}
+         \label{fig:abovemed_read_synth_clinical}
+     \end{subfigure}
+        \label{fig:cont_clinicalsynthdid}
+    \end{figure}
+
+            \begin{figure}[ht!]
+     \caption{Effect of Clinical Training on Mortality Rates, Binned Treatment}
+     \centering
+     \begin{subfigure}[b]{0.45\textwidth}
+         \centering
+         \caption{$\rho \leq \eta$}
+         \includegraphics[width=\textwidth]{Objects/cont_belowmedmort_md_nomd_size_synth_graph.pdf}
+         \label{fig:belowmed_read_synth_clinical}
+     \end{subfigure}
+     \hfill
+     \begin{subfigure}[b]{0.45\textwidth}
+         \centering
+         \caption{$\rho > \eta$}
+         \includegraphics[width=\textwidth]{Objects/cont_abovemedmort_md_nomd_size_synth_graph.pdf}
+         \label{fig:abovemed_read_synth_clinical}
+     \end{subfigure}
+        \label{fig:cont_clinicalsynthdid}
+    \end{figure}
+
+    \subsection{Selective Patient Practices}\label{app:casemix}
 
@@
 
-\section{Results By Condition}\label{app:condition}
+\subsection{Results By Condition}\label{app:condition}
 
@@
 
-\section{TWFE Regression and Event Study Results}\label{app:fullsample}
+\subsection{TWFE Regression and Event Study Results}\label{app:fullsample}
 

### 2024-08-21 Tables/MD_noMD_readmort_condition_synth.tex
Updates from Overleaf

@@
 \toprule
-\multicolumn{1}{c}{\underline{ }} & \multicolumn{3}{c}{\underline{Readmission Rates}} & \multicolumn{3}{c}{\underline{Mortality Rates}} \\
+\multicolumn{1}{c}{} & \multicolumn{3}{c}{\underline{Readmission Rates}} & \multicolumn{3}{c}{\underline{Mortality Rates}} \\
  & AMI & HF & Pneum. & AMI & HF & Pneum.\\

### 2024-08-21 Paper/paper2.tex
Updates from Overleaf

@@
 
-\import{Tables}{MD_noMD_readmort_condition_synth.tex}
+\import{Tables}{condition_synthdid_table.tex}
 
@@
 
-While I ultimately employ a synthetic difference-in-differences framework, I also present the coefficient estimates and event study tables under a typical TWFE model specification.
-That is, there is no weighting of the control group or time periods.
-Table \ref{tab:forprofit_readmort_fullsample} shows coefficient estimates for the three specifications under Equation \ref{eq:forprofit}.
-Columns (1)-(3) show coefficient estimates for the weighted average readmission rate outcome, and columns (4)-(6) show coefficient estimates from the weighted average mortality outcome.
-Generally, these results are qualitatively identical to the synthetic difference-in-difference specification results.
-There is a statistically significant difference in readmission rate response between for-profits and NFPs with a clinical executive.
-Specifically, for-profits decrease their readmission rates by .4ppts more than NFPs with a clinical executive.
-There is no difference (statistically or in magnitude) between readmission rate responses of for-profits compared to all NFPs and NFPs without clincal executives.
+\subsubsection{Effect of Clinically Trained Executives on Response to Financial Incentive on Quality}
 
-Similarly to the synthetic diff-in-diff estimates, there is no statistical difference in mortality between any of the hospital types, but the magnitude difference between for-profits and NFPs with clinical executives is much larger.
-This oculd be due to the noisiness of mortality as a measure.
-
-\import{Tables}{forprofit_readmort_fullsample.tex}
-
-Further, I present coefficient estimates in Table \ref{tab:MD_noMD_readmort_fullsample} directly comparing NFPs with and without clinical executives for both readmission and mortality rates.
-Again, these results are very similar to the main results of the paper, with only slightly larger magnitudes.
-NFPs without a clinical executive respond to the pay-for-performance incentives by decreasing readmission and mortality rates more aggressively than NFPs with clinical executives.
-
-\import{Tables}{MD_noMD_readmort_fullsample.tex}
-
-Finally, I present these results in typical event study form in Figure \ref{fig:es_plot}, which show no evidence of significant pre-trends driving the findings.
+\import{Tables}{main_twfe_match_tab.tex}
 
 \begin{figure}[ht!]
-     \caption{TWFE Event Study Results}
     \centering
-          \begin{subfigure}[b]{0.45\textwidth}
-         \centering
-         \caption{}
-         \includegraphics[width=\textwidth]{Objects/read_forprofit_es_graph.pdf}
-         \label{fig:es_plota}
-     \end{subfigure}%
-     \hfill
-     \begin{subfigure}[b]{0.45\textwidth}
-         \centering
-         \caption{}
-         \includegraphics[width=\textwidth]{Objects/read_MD_es_graph.pdf}
-         \label{fig:es_plotb}
-     \end{subfigure}%
-     \vspace{5mm}
-     \hfill
-     \begin{subfigure}[b]{0.45\textwidth}
-         \centering
-         \caption{}
-         \includegraphics[width=\textwidth]{Objects/mort_forprofit_es_graph.pdf}
-         \label{fig:es_plotc}
-     \end{subfigure}
-     \hfill
-     \begin{subfigure}[b]{0.45\textwidth}
-         \centering
-         \caption{}
-         \includegraphics[width=\textwidth]{Objects/mort_MD_es_graph.pdf}
-         \label{fig:es_plotd}
-     \end{subfigure}
-        \label{fig:es_plot}
+    \includegraphics[width=.99\linewidth]{Objects/main_twfe_match_eventstudy.pdf}
+    \caption{Effect of Clinical Executives; Event Study Results}
+    \label{fig:main_twfe_match_es}
 \end{figure}

### 2024-08-21 Tables/condition_synthdid_table.tex
Updates from Overleaf

@@
 \centering
-\begin{tabular}[t]{lccclcc}
+\begin{tabular}[t]{lcccccc}
 \toprule
 \multicolumn{1}{c}{ } & \multicolumn{3}{c}{Readmission Rates} & \multicolumn{3}{c}{Mortality Rates} \\
+\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-7}
+\addlinespace[0.3em]
 \multicolumn{1}{c}{ } & \multicolumn{1}{c}{AMI} & \multicolumn{1}{c}{HF} & \multicolumn{1}{c}{Pneum.} & \multicolumn{1}{c}{AMI} & \multicolumn{1}{c}{HF} & \multicolumn{1}{c}{Pneum.} \\
@@
 \midrule
-Ever Clinical Exec x Post Program & 0.25 & 0.14 & 0.46$^{**}$ & -0.13 & 0.03 & 0.3\\
+Always Clinical Exec x Post Program & 0.25 & 0.14 & 0.46$^{**}$ & -0.13 & 0.03 & 0.3\\
  & (0.18) & (0.18) & (0.14) & (0.19) & (0.13) & (0.19)\\

### 2024-08-21 Tables/main_twfe_match_tab.tex
Updates from Overleaf

@@
 
-\caption{\label{tab:main_twfe_match}Effect of Clinically Trained Executive, TWFE and Matched Sample}
+\caption{\label{tab:main_twfe_match}Effect of Clinically Trained Executive; TWFE and Matched Sample}
 \centering
-\begin{tabular}[t]{lcccl}
+\begin{tabular}[t]{lcccc}
 \toprule
-\multicolumn{1}{c}{\underline{ }} & \multicolumn{2}{c}{\underline{TWFE}} & \multicolumn{2}{c}{\underline{Matched Sample}} \\
+\multicolumn{1}{c}{\underline{ }} & \multicolumn{2}{c}{TWFE} & \multicolumn{2}{c}{Matched Sample} \\
 \cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
-\multicolumn{1}{c}{ } & \multicolumn{1}{c}{Read} & \multicolumn{1}{c}{Mort} & \multicolumn{1}{c}{Read} & \multicolumn{1}{c}{Mort} \\
+\addlinespace[0.3em]
+\multicolumn{1}{c}{ } & \multicolumn{1}{c}{Readmission} & \multicolumn{1}{c}{Mortality} & \multicolumn{1}{c}{Readmission} & \multicolumn{1}{c}{Mortality} \\
 \cmidrule(l{3pt}r{3pt}){2-2} \cmidrule(l{3pt}r{3pt}){3-3} \cmidrule(l{3pt}r{3pt}){4-4} \cmidrule(l{3pt}r{3pt}){5-5}
@@
 \midrule
-Ever Clinical Exec x Post Program & 0.42$^{**}$ & 0.11 & 0.65$^{*}$ & 0.03\\
+Always Clinical Exec x Post Program & 0.42$^{**}$ & 0.11 & 0.65$^{*}$ & 0.03\\
  & (0.15) & (0.15) & (0.3) & (0.4)\\

### 2024-08-22 Paper/paper2.tex
Updates from Overleaf

@@
     In Section \ref{sec:clinical}, I present intensive margin results by binning hospitals according to the fraction of their executive team with clinical experience.
-There, I find that the intensive margin does matter for how hospitals change readmissions after the incentive change.
+There, when comparing hospitals with greater than the median (20\%) to less than the median fraction of clinical executives, I find that the difference in response is driven by hospitals with a fraction greater than the median.
 One concern could be that this result is driven by a difference in the total size of hospital executive teams, the denominator of that fraction.
 In this section, I use the same estimating equation, but I limit to only nonprofit hospitals with 7 or less executives in total, removing outliers of executive team size.
-The difference in readmission rate response is still driven by teams with a fraction of clinical executives above the median.
+The results are shown in Figure \ref{fig:cont_clinicalsynthdid_read_size}, and show that the response is still driven by hospitals with a fraction of clinical executives above the median.
 
@@
      \end{subfigure}
-        \label{fig:cont_clinicalsynthdid}
+        \label{fig:cont_clinicalsynthdid_read_size}
     \end{figure}
+
+    I also present results with mortality as the dependent variable, shown in Figure \ref{fig:cont_clinicalsynthdid_size}.
+These results do not differ from the main finding in that there is no statistical difference in response for hospitals with different types of leadership teams with above or below the median clinically trained executives.
 
@@
      \end{subfigure}
-        \label{fig:cont_clinicalsynthdid}
+        \label{fig:cont_clinicalsynthdid_size}
     \end{figure}
@@
 
-    A method hospitals have been shown to employ in order to decrease readmission rates is being more selective with which patients to admit to the hospital (\cite{gupta2021impacts}).
-One hypothesis behind a more drastic quality decrease by non-clinical teams is that they may be more willing to participate in such practices, while executives with clinical training prioritize care for all types of patients.
+    A method hospitals have been shown to employ in order to decrease readmission rates is being more selective with which patients to admit to the hospital (\cite{gupta2021impacts}), as only admitting healthy patients mechanically decreases readmission rates.
+One hypothesis behind a more drastic quality decrease by non-clinical teams is that they may be more willing to participate in such practices, while executives with clinical training are less likely to turn patients away.
 
@@
 
+In the main body of the paper, the main dependent variables I consider are weighted averages of readmission and mortality rates for several conditions relevant to the policy changes, weighted by the number of patients seen with each condition.
+Here, I investigate whether the findings are driven by any particular condition.
+The conditions I consider are heart attack (AMI), heart failure, and pneumonia, which are the conditions that go into the calculation for HRRP penalties.
+Thus, I consider the effect of clinically trained executives on readmission and mortality rates for each condition individually.
+I estimate equation \ref{eq:clinical} with these outcome variables.
+
+I present the resulting estimates and standard errors in Table \ref{tab:results_by_condition}.
+Columns (1)-(3) show results for readmission rate outcomes in each condition.
+The main finding, that clinically trained executives lead hospitals to have higher readmission rates after pay-for-performance policies, is primarily driven by patients with pneumonia.
+While magnitudes are positive for all conditions, pneumonia has the largest magnitude and is the only statistically significant estimate.
+Columns (4)-(6) present results for mortality rates of each condition.
+Similarly to the results for weighted average mortality, there are no statistically significant results here.
+However, pneumonia has the largest positive magnitude, just like pneumonia readmissions.
+
 \import{Tables}{condition_synthdid_table.tex}
 
-\subsection{TWFE Regression and Event Study Results}\label{app:fullsample}
+\subsection{Robustness to Other Estimations}\label{app:fullsample}
+
+In analyses presented in the main body of the paper, I use a synthetic difference-in-differences approach to identify the effect of clinically trained executives.
+I choose this as the main estimation strategy due to potential unobserved differences correlated with executive choice, where synthetic DiD assures parallel trends by choosing the optimal composition of control and treatment hospitals.
+In this section, I present results under three different estimation techniques for robustness: classic two way fixed effects, two-way fixed effects with manual weights that balance case mix index and penalties of hospitals with different leadership teams, and synthetic control.
 
@@
 
+In Section \ref{sec:clinical}, I first establish the effect of clinically trained executives on response to change in incentives on quality using a synthetic difference-in-differences approach.
+Now, I present results using various estimation techniques in Table \ref{tab:main_twfe_match} in order to assess to robustness of this finding.
+
 \import{Tables}{main_twfe_match_tab.tex}
+
+Additionally, I estimate an event study version of the two way fixed effects specifications.
+The estimates and 95\% confidence intervals are shown in Figure \ref{fig:main_twfe_match_es}.
+These confirm the main findings with slightly noisier estimates, and show that the difference in response in readmission rates occurs in 2013 and 2014, one and two years after the policy changes.
+The mortality results confirm the main findings, that there is no difference in response in mortality rates.
 
@@
 
+\subsubsection{Are Clinical Executives Less Profit Driven?}
+
+\subsubsection{Signaling vs. Managing Decomposition}
+
 \end{document}

### 2024-08-22 Tables/main_twfe_match_tab.tex
Updates from Overleaf
update tables

--- 2024-08-21 Tables/main_twfe_match_tab.tex
@@
 
-\caption{\label{tab:main_twfe_match}Effect of Clinically Trained Executive, TWFE and Matched Sample}
+\caption{\label{tab:main_twfe_match}Effect of Clinically Trained Executive, Various Estimations}
 \centering
-\begin{tabular}[t]{lcccl}
+\begin{tabular}[t]{lcccccl}
 \toprule
-\multicolumn{1}{c}{\underline{ }} & \multicolumn{2}{c}{\underline{TWFE}} & \multicolumn{2}{c}{\underline{Matched Sample}} \\
-\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
-\multicolumn{1}{c}{ } & \multicolumn{1}{c}{Read} & \multicolumn{1}{c}{Mort} & \multicolumn{1}{c}{Read} & \multicolumn{1}{c}{Mort} \\
-\cmidrule(l{3pt}r{3pt}){2-2} \cmidrule(l{3pt}r{3pt}){3-3} \cmidrule(l{3pt}r{3pt}){4-4} \cmidrule(l{3pt}r{3pt}){5-5}
- & (1) & (2) & (3) & (4)\\
+\multicolumn{1}{c}{\underline{ }} & \multicolumn{3}{c}{\underline{Readmission Rate}} & \multicolumn{3}{c}{\underline{Mortality Rate}} \\
+\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-7}
+\multicolumn{1}{c}{ } & \multicolumn{1}{c}{TWFE} & \multicolumn{1}{c}{Matched} & \multicolumn{1}{c}{SC} & \multicolumn{1}{c}{TWFE} & \multicolumn{1}{c}{Matched} & \multicolumn{1}{c}{SC} \\
+\cmidrule(l{3pt}r{3pt}){2-2} \cmidrule(l{3pt}r{3pt}){3-3} \cmidrule(l{3pt}r{3pt}){4-4} \cmidrule(l{3pt}r{3pt}){5-5} \cmidrule(l{3pt}r{3pt}){6-6} \cmidrule(l{3pt}r{3pt}){7-7}
+ & (1) & (2) & (3) & (4) & (5) & (6)\\
 \midrule
-Ever Clinical Exec x Post Program & 0.42$^{**}$ & 0.11 & 0.65$^{*}$ & 0.03\\
- & (0.15) & (0.15) & (0.3) & (0.4)\\
- &  &  &  & \\
-Hospital FE & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
-Year FE & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
+Ever Clinical Exec x Post Program & 0.42$^{**}$ & 0.65$^{*}$ & 0.25 & 0.11 & 0.03 & 0.06\\
+ & (0.15) & (0.3) & (0.14) & (0.15) & (0.4) & (0.12)\\
+ &  &  &  &  &  & \\
+Hospital FE & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
+Year FE & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
 \addlinespace
-Observations & 2,387 & 2,381 & 1,025 & 1,023\\
+Observations & 2,387 & 1,025 & 1,640 & 2,381 & 1,023 & 1,640\\
 \bottomrule
-\multicolumn{5}{l}{\textsuperscript{} Standard errors are clustered at the hospital level.}\\
-\multicolumn{5}{l}{\textsuperscript{} Signif. codes: 0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1}\\
+\multicolumn{7}{l}{\textsuperscript{} Standard errors are clustered at the hospital level.}\\
+\multicolumn{7}{l}{\textsuperscript{} Signif. codes: 0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1}\\
 \end{tabular}

### 2024-08-23 Tables/fp_always_estimators_mort_tab.tex
add tables for appendix, various estimators

--- 2024-08-22 Tables/main_twfe_match_tab.tex
@@
 
-\caption{\label{tab:main_twfe_match}Effect of Clinically Trained Executive, Various Estimations}
+\caption{\label{tab:forprofit_mort_estimators}Compare For-Profit to Types of Leadership (Mortality Rates), Various Estimations}
 \centering
@@
 \toprule
-\multicolumn{1}{c}{\underline{ }} & \multicolumn{3}{c}{\underline{Readmission Rate}} & \multicolumn{3}{c}{\underline{Mortality Rate}} \\
+\multicolumn{1}{c}{ } & \multicolumn{3}{c}{Compare to Always Clinical Exec} & \multicolumn{3}{c}{Compare to Never Clinical Exec} \\
 \cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-7}
@@
 \midrule
-Ever Clinical Exec x Post Program & 0.42$^{**}$ & 0.65$^{*}$ & 0.25 & 0.11 & 0.03 & 0.06\\
- & (0.15) & (0.3) & (0.14) & (0.15) & (0.4) & (0.12)\\
+For-Profit x Post Program & 0.03 & 0.09 & -0.16 & 0.06 & 0.05 & 0.03\\
+ & (0.14) & (0.45) & (0.2) & (0.07) & (0.11) & (0.1)\\
  &  &  &  &  &  & \\
@@
 \addlinespace
-Observations & 2,387 & 1,025 & 1,640 & 2,381 & 1,023 & 1,640\\
+Observations & 2,660 & 660 & 2,660 & 3,820 & 1,790 & 3,820\\
 \bottomrule

### 2024-08-23 Tables/fp_always_estimators_read_tab.tex
add tables for appendix, various estimators

--- 2024-08-23 Tables/fp_always_estimators_mort_tab.tex
@@
 
-\caption{\label{tab:forprofit_mort_estimators}Compare For-Profit to Types of Leadership (Mortality Rates), Various Estimations}
+\caption{\label{tab:forprofit_estimators}Compare For-Profit to Types of Leadership (Readmission Rates), Various Estimations}
 \centering
@@
 \midrule
-For-Profit x Post Program & 0.03 & 0.09 & -0.16 & 0.06 & 0.05 & 0.03\\
- & (0.14) & (0.45) & (0.2) & (0.07) & (0.11) & (0.1)\\
+For-Profit x Post Program & -0.37$^{*}$ & -0.28 & -0.43$^{*}$ & -0.03 & -0.08 & 0.09\\
+ & (0.15) & (0.44) & (0.19) & (0.07) & (0.12) & (0.1)\\
  &  &  &  &  &  & \\

### 2024-08-23 Tables/signal_manage_estimators_mort_tab.tex
add tables for appendix, various estimators

--- 2024-08-22 Tables/main_twfe_match_tab.tex
@@
 
-\caption{\label{tab:main_twfe_match}Effect of Clinically Trained Executive, Various Estimations}
+\caption{\label{tab:signal_manage_mort_estimators}Signaling vs. Managing Decomp. (mortmission Rates), Various Estimations}
 \centering
@@
 \toprule
-\multicolumn{1}{c}{\underline{ }} & \multicolumn{3}{c}{\underline{Readmission Rate}} & \multicolumn{3}{c}{\underline{Mortality Rate}} \\
+\multicolumn{1}{c}{ } & \multicolumn{3}{c}{Signaling Effect} & \multicolumn{3}{c}{Managing Effect} \\
 \cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-7}
@@
 \midrule
-Ever Clinical Exec x Post Program & 0.42$^{**}$ & 0.65$^{*}$ & 0.25 & 0.11 & 0.03 & 0.06\\
- & (0.15) & (0.3) & (0.14) & (0.15) & (0.4) & (0.12)\\
+Estimate & -0.01 & -0.23 & 0.02 & 0.14 & -0.05 & 0.11\\
+ & (0.08) & (0.18) & (0.06) & (0.11) & (0.28) & (0.08)\\
  &  &  &  &  &  & \\
@@
 \addlinespace
-Observations & 2,387 & 1,025 & 1,640 & 2,381 & 1,023 & 1,640\\
+Observations & 3,005 & 670 & 3,005 & 1,595 & 320 & 2,040\\
 \bottomrule

### 2024-08-23 Paper/paper2.tex

@@
 I present the unweighted two-way-fixed-effects specification equation for clarity, and the identification assumptions are conceptually similar for synthetic difference-in-differences.
+In Appendix \ref{app:otherestimations}, I show that the findings are robust to using two way fixed effects, matching, or synthetic control.
 
@@
 
+    The hypothesis driven from this theory is that clinically trained executives put more weight on quality of care before the financial incentives on quality began.
 To investigate this hypothesis empirically, I merge all general for-profit hospitals in the AHA data to the sample of nonprofit hospitals.
 Then, I compare readmission and mortality rates of always clinical executives and never clinical executives to for-profit hospitals after the incentive change.
-Ideally, I would also have leadership heterogeneity information for for-profit hospitals, but this data does not exist publicly.
-However, on average, for-profit hospitals have revealed through their ownership status that they more heavily weigh profit.
+While I do not have information on leadership teams of for-profit hospitals, on average, they have revealed through their ownership status that they more heavily weigh profit.
 Thus, comparing different executive team behavior to for-profit behavior helps reveal how profit driven certain leadership teams are.
@@
 
-\subsection{Robustness to Other Estimations}\label{app:fullsample}
+\subsection{Robustness to Other Estimations}\label{app:otherestimations}
 
@@
 I choose this as the main estimation strategy due to potential unobserved differences correlated with executive choice, where synthetic DiD assures parallel trends by choosing the optimal composition of control and treatment hospitals.
-In this section, I present results under three different estimation techniques for robustness: classic two way fixed effects, two-way fixed effects with manual weights that balance case mix index and penalties of hospitals with different leadership teams, and synthetic control.
+As a robustness check, I now present results under three different estimation techniques for robustness: classic two way fixed effects, two-way fixed effects with manual weights that balance case mix index and penalties of hospitals with different leadership teams, and synthetic control.
+
+For the manual weighting strategy, I apply the Coarsened Exact Matching (CEM) procedure as described by \citeauthor{azoulay2010superstar} (\citeyear{azoulay2010superstar}).
+This approach involves grouping hospitals based on observable characteristics such as bed size, the number of patients in each relevant condition, and eventual HRRP penalty status in each relevant condition.
+Hospitals are then placed into bins according to these characteristics, and matches are made exactly between hospitals within the same bin.
+Hospitals that do not find an exact match in this process are excluded from the sample.
 
@@
 Now, I present results using various estimation techniques in Table \ref{tab:main_twfe_match} in order to assess to robustness of this finding.
+Columns (1)-(3) present estimates and standard errors for the three estimations: TWFE, TWFE with matched sample, and synthetic control, respectively.
+For both TWFE and TWFE with matching, the magnitude is larger than that for synthetic DiD (0.27) and statistically significant.
+The estimate and standard error for synthetic control are very similar to those for synthetic DiD.
+In columns (4)-(6), I present the results for mortality rates, which confirm no differential response.
+This exercise supports the finding that clinically trained executives affect the readmission rate response of hospitals to change in incentives on quality.
 

### 2024-08-23 Tables/MD_noMD_readmort_decomp_synth.tex
update tables

@@
 \toprule
-\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Weighted Avg. Readmission Rate} & \multicolumn{2}{c}{Weighted Avg. Mortality Rate} \\
- & (1) & (2) & (1) & (2)\\
+\multicolumn{1}{c}{ } & \multicolumn{2}{c}{\underline{Weighted Avg. Readmission Rate}} & \multicolumn{2}{c}{\underline{Weighted Avg. Mortality Rate}} \\
+ & (1) & (2) & (3) & (4)\\
 \midrule
-Ever MD Exec x Post Programs & -0.01 &  & 0.03 & \\
+Signaling Effect & -0.01 &  & 0.03 & \\
  & (0.07) &  & (0.06) & \\
-2012 MD Exec x Post Programs &  & 0.17 &  & 0.12\\
+Managing Effect &  & 0.17 &  & 0.12\\
  &  & (0.1) &  & (0.09)\\
  &  &  &  & \\
-\addlinespace
-\textbf{Comparison Group:} &  &  &  & \\
-Never MD & $\checkmark$ &  & $\checkmark$ & \\
-Has MD not in 2012 &  & $\checkmark$ &  & $\checkmark$\\
- &  &  &  & \\
-Observations & 3005 & 1595 & 3005 & 1595\\
+Observations & 3,005 & 1,595 & 3,005 & 1,595\\
 \bottomrule

### 2024-09-03 Paper/paper2.tex
Updates from Overleaf

@@
 Thus, I leverage the timing of the policy change, rather than the policies themselves.
+
     One may be concerned about potential differences in quality trends between clinical and non-clinical executive team hospitals leading to a violation in the parallel trends assumption.
 This is why I implement a synthetic difference-in-differences estimation strategy (\cite{arkhangelsky2021synthetic}).
-That is, non-clinical team hospitals with trends similar to clinical team hospitals in readmission, mortality, and baseline case mix before the programs are more heavily weighted, as well as time periods that balance pre- and post-program outcomes for non-clinical teams.
+In particular, I weight more heavily any non-clinical team hospitals with pre-trends similar to clinical team hospitals in readmission, mortality, and baseline case mix.
+Further, I more heavily weight time periods that balance pre- and post-program outcomes for non-clinical teams.
+[description of weight distribution here]
+
     I present the unweighted two-way-fixed-effects specification equation for clarity, and the identification assumptions are conceptually similar for synthetic difference-in-differences.
@@
 
+In Section \ref{sec:forprofit}, I test empirically whether nonprofits with different leadership teams respond differently to the incentive change than for-profit hospitals using a synthetic difference-in-differences methodology.
+Here, I present results from various other estimation strategies.
+In Table \ref{tab:forprofit_estimators}, I present estimates and standard errors with weighted average readmission rate as the outcome.
+These results confirm what is shown in the main body of the paper, that nonprofits without clinically trained executives act similarly to for-profits, but nonprofits with clinically trained executives do not.
+
+\import{Tables}{fp_always_estimators_read_tab.tex}
+
+While the results in the main body of the paper for mortality rates as the dependent variable do not indicate differences in response, I still present the estimates from other estimations in Table \ref{tab:forprofit_mort_estimators}.
+These results do not contradict what I show in the main body of the paper, as I still see small and statistically insiginificant estimates.
+
+\import{Tables}{fp_always_estimators_mort_tab.tex}
+
 \subsubsection{Signaling vs. Managing Decomposition}
 
+In Section \ref{sec:sig_man}, I test empirically whether clinical executives make a difference because of a signaling or managing effect.
+That is, I leverage the timing of hiring clinically trained executives.
+Here, I present results from the same specification, but various estimation strategies.
+In Table \ref{tab:signal_manage_read_estimators}, I present estimates and standard errors with weighted average readmission rate as the outcome.
+These results confirm what is shown in the main body of the paper, that the managing effect is driving the findings.
+
+\import{Tables}{signal_manage_estimators_read_tab.tex}
+
+While the results in the main body of the paper for mortality rates as the dependent variable do not indicate differences in response, I still present the estimates from other estimation strategies in Table \ref{tab:signal_manage_mort_estimators}.
+These results do not contradict what I show in the main body of the paper, as I still see small and statistically insiginificant estimates.
+Therefore, there is no opposing signaling vs.
+managing effect.
+
+\import{Tables}{signal_manage_estimators_mort_tab.tex}
+
 \end{document}

### 2024-09-03 RScripts/analysis.R
update scripts

@@ -1,2125 +1,0 @@
-library(readr)
-library(dplyr)
-library(fixest)
-library(ggplot2)
-library(ggpubr)
-library(tidyr)
-library(synthdid)
-library(kableExtra)
-options(knitr.kable.NA=" ")
-
-# Comparisons I need: 
-# 1. FP - NFP
-# 2. FP - has MD
-# 3. FP - no MD
-# 4. has MD - no MD
-
-# Methodologies:
-# 1. full sample difference in difference
-# 2. Matched difference in difference
-# 3. Synthetic difference in difference
-
-# Outcomes:
-# 1. Readmission rates
-# 2. Mortality Rates
-# 3. Uncompensated Care and Case Mix Index
-
-# Sample limitations:
-# 1. dropping hospitals with leadership changes
-# 2. not dropping hospitals with leadership changes
-
-# limit to 2010 to 2014
-hospital_data <- readRDS(paste0(created_data_path, "all_hospital_data.rds")) %>%
-  filter(year>=2010 & year<=2014)
-
-# drop hospitals with less than 10 beds
-hospital_data <- hospital_data %>%
-  group_by(ID) %>%
-  mutate(maxbeds=max(beds, na.rm=T)) %>%
-  ungroup() %>%
-  filter(maxbeds>=10) %>%
-  select(-maxbeds)
-
-hospital_data$uncomp_care <- abs(hospital_data$uncomp_care)
-hospital_data <- hospital_data %>%
-  mutate(uncomp_care=uncomp_care/1000000) 
-
-# fill ID
-hospital_data <- hospital_data %>%
-  group_by(MCRNUM) %>%
-  fill(ID, .direction="downup") %>%
-  ungroup() %>%
-  filter(!is.na(ID)) 
-
-# Remove hospitals that change from nonprofit to for-profit or vice versa
-hospital_data <- hospital_data %>%
-  mutate(post_2012=ifelse(year>2012, 1, 0),
-         NFP = ifelse(profit_status=="nonprofit", 1, 0),
-         FP = ifelse(profit_status=="forprofit",1,0)) %>%
-  group_by(ID) %>%
-  arrange(year) %>%
-  mutate(lag_NFP = dplyr::lag(NFP)) %>%
-  ungroup() %>%
-  filter(NFP==lag_NFP | year==2010)
-
-comp <- c("FP - NFP", "FP - has MD", "FP - no MD", "has MD - no MD")
-method <- c("full", "matched", "synth")
-
[…]

### 2024-09-03 RScripts/appendix_tables_and_figures.R

@@ -1,0 +1,1958 @@
+library(dplyr)
+library(readr)
+library(knitr)
+library(kableExtra)
+library(ggplot2)
+library(ggpubr)
+library(RColorBrewer)
+library(synthdid)
+library(abind)
+library(reshape2)
+library(data.table)
+library(tidyr)
+library(maps)
+library(cdlTools)
+library(MatchIt)
+library(fixest)
+
+options(knitr.kable.NA=" ")
+
+# read in data
+hospital_data <- read_rds(paste0(created_data_path, "all_hospital_data3.rds"))
+doc_exec_data <- read_rds(paste0(created_data_path, "doc_exec_data.rds"))
+
+hospital_data <- hospital_data %>%
+  mutate(ever_pen_ha_only=ifelse(ever_pen_ha>=1 & ever_pen_hf==0 & ever_pen_pnem==0,1,0), 
+         ever_pen_hf_only=ifelse(ever_pen_hf>=1 & ever_pen_ha==0 & ever_pen_pnem==0,1,0), 
+         ever_pen_pnem_only=ifelse(ever_pen_pnem>=1 & ever_pen_hf==0 & ever_pen_ha==0,1,0), 
+         ever_pen_ami_hf = ifelse(ever_pen_hf>=1 & ever_pen_ha>=1 & ever_pen_pnem==0,1,0),
+         ever_pen_ami_pneum = ifelse(ever_pen_hf==0 & ever_pen_pnem>=1 & ever_pen_ha==1,1,0),
+         ever_pen_hf_pneum = ifelse(ever_pen_hf>=1 & ever_pen_pnem>=1 & ever_pen_ha>=0,1,0),
+         ever_pen_all = ifelse(ever_pen_hf>=1 & ever_pen_ha>=1 & ever_pen_pnem>=1,1,0)) %>%
+  mutate(frac_doc_exec = num_doctors/num_execs) %>%
+  mutate(rate_exec=num_doctors/num_execs) %>%
+  mutate(rate_int_exec=num_int_med_doctors/num_execs) %>%
+  mutate(ever_has_md = ifelse(is.na(ever_has_md),0,ever_has_md)) 
+
+# TABLE 4: compare in-sample and out-of-sample nonprofit hospitals ####
+in_sample_NFP_data <- hospital_data %>%
+  filter(year!=2009 & year<=2014) %>%
+  filter(!is.na(no_num_md_change_2010_2014)) %>%
+  summarise_at(c("Number Beds"="beds",
[…]

### 2024-09-03 RScripts/data10_create_leadership_variables.R

@@ -2,10 +2,12 @@ library(readr)
 library(dplyr)
 library(stringr)
 library(tidyr)
+library(readxl)
 
-
-# read in names data
+# read in data
 cleaned_text <- read_rds(paste0(created_data_path, "cleaned_text.rds"))
+hospital_pdf_locations <- read_rds(paste0(created_data_path,"hospital_pdf_locations.rds"))
+npidata <- readRDS(paste0(created_data_path, "npidata_names.rds"))
 
 # First objective is to identify whether someone is part of the board or part of the actual day to day decision making
 # if the only position is physician, drop them as they are just a high compensated employee
@@ -137,7 +139,669 @@ executive_data <- executive_data_filled %>%
 
 rm(executive_data_filled, name_data, data, cleaned_text)
 
-# goal 2: identify whether changes occur from year to year ####
+# goal 2: match physicians to NPI numbers ####
+
+executive_names <- executive_data %>% ungroup() %>%
+  mutate(doctor = ifelse(title %in% c("md", "dr", "do"), 1, NA))
+
+# Filter only to physicians
+npidata <- npidata %>%
+  filter(t_group=="Allopathic & Osteopathic Physicians"|t_group=="Other Service Providers")
+
+# get rid of people who don't have names (I don't know why so many of them have NA for name??)
+npidata <- npidata %>%
+  filter(!(is.na(firstname) & is.na(lastname)))
+# still have a million observations (lost 300k)
+
+# change enumeration date to enumeration year
+npidata <- npidata %>%
+  mutate(enum_date = as.Date(enum_date,'%m/%d/%Y')) %>%
+  mutate(enum_year = as.numeric(format(enum_date,'%Y'))) %>%
+  select(-enum_date)
+
+# I don't need group or tax code anymore, just specialty
+npidata <- npidata %>%
+  select(-t_code, -t_group)
+
+# convert first name and last name to one name line and make sure both are lowercase
+npidata <- npidata %>%
+  mutate(name = paste0(firstname, " ", lastname)) %>%
+  mutate(name = tolower(name))
+
+# create a data set that will be used to store matches 
+names_data <- executive_names %>%
+  distinct(ein, name, title, doctor) %>%
+  mutate(num_matches=NA)
+
[…]

### 2024-09-03 RScripts/data11_combine_hosp_level_data.R

@@ -22,7 +22,7 @@ hc_readm_2013 <- read_csv(paste0(raw_data_path, "/Hospital Compare/vwhqi_readm_r
 hc_readm_2014 <- read_csv(paste0(raw_data_path, "/Hospital Compare/vwhqi_readm_reduction_2014.csv"))
 hc_readm_2015 <- read_csv(paste0(raw_data_path, "/Hospital Compare/vwhqi_readm_reduction_2015.csv"))
 # Data that I created on leadership teams from the tax forms
-ein_leadership_data <- read_rds(paste0(created_data_path, "ein_leadership_changes_data.rds"))
+ein_leadership_data <- read_rds(paste0(created_data_path, "ein_leadership_changes_data4.rds"))
 # Data that I created on outcomes from Hospital Compare
 outcomes_data <- read_rds(paste0(created_data_path, "hosp_outcomes.rds"))
 # read in case mix index
@@ -68,7 +68,7 @@ AHA <- AHA %>%
   select(-count, -sum)
   # 3,766
 
-AHA <- complete(AHA, MCRNUM, YEAR=2009:2015) %>%
+AHA <- complete(AHA, MCRNUM, YEAR=2009:2014) %>%
   group_by(MCRNUM) %>%
   fill(CNTRL, .direction="downup") %>%
   ungroup()
@@ -77,11 +77,11 @@ AHA <- complete(AHA, MCRNUM, YEAR=2009:2015) %>%
 matches <- rbind(AHA_ein_matches, manual_matched_eins)
 
 matches <- matches %>%
-  mutate(year=2008) %>%
+  mutate(year=2010) %>%
   filter(!is.na(ein_hosp)) %>%
   select(-ein_sys)
 
-matches <- complete(matches,ID, year=2008:2015) %>%
+matches <- complete(matches,ID, year=2010:2014) %>%
   group_by(ID) %>%
   fill(ein_hosp, .direction="downup") %>%
   ungroup()
@@ -97,7 +97,7 @@ observe <- matches %>%
 
 # only keep rows that had values in that data set
 matches <- matches %>%
-  mutate(drop=ifelse(year %in% c(2010,2011,2012,2013,2014) & is.na(num_execs),1,0)) %>%
+  mutate(drop=ifelse(is.na(num_execs),1,0)) %>%
   group_by(ein_hosp) %>%
   mutate(drop=sum(drop, na.rm=T)) %>%
   ungroup() %>%
[…]

### 2024-09-03 RScripts/physician_name_matching.R

@@ -2,16 +2,17 @@ library(readr)
 library(dplyr)
 library(stringr)
 library(tidyr)
-
-source("paths.R")
+library(readxl)
 
 # read in names data
-names_data <- read_rds(paste0(created_data_path, "names_data.rds"))
+executive_names <- read_rds(paste0(created_data_path, "executive_names.rds"))
+hospital_pdf_locations <- read_rds(paste0(created_data_path,"hospital_pdf_locations.rds"))
 
 # create an indicator for MD
-names_data <- names_data %>%
-  mutate(doctor = ifelse(title %in% c("md", "dr", "do"), 1, NA))
-  # 3.6% of people have a doctor title
+executive_names <- executive_names %>%
+  mutate(doctor = ifelse(title %in% c("md", "dr", "do"), 1, NA)) %>%
+  ungroup()
+  # 11% of people have a doctor title
 
 
 # read in npi data of physician names
@@ -41,7 +42,7 @@ npidata <- npidata %>%
   mutate(name = paste0(firstname, " ", lastname)) %>%
   mutate(name = tolower(name))
 
-# look for the max number of one name occuring in npi data
+# look for the max number of one name occurring in npi data
 npidata <- npidata %>%
   mutate(count=1) %>%
   group_by(name) %>%
@@ -51,8 +52,8 @@ max <- max(npidata$sum)
   # 136
 
 # create a data set that will be used to store matches 
-names_data <- names_data %>%
-  distinct(ein, name, title, doctor, years) %>%
+names_data <- executive_names %>%
+  distinct(ein, name, title, doctor) %>%
   mutate(num_matches=NA)
 
 # create list of names in the NPPES data
@@ -82,7 +83,6 @@ for (i in 1:dim(names_data)[1]){
 sure_matches <- names_data %>%
   filter(doctor==1 & num_matches==1) %>%
   select(name, ein)
-  # 76 people
 
 npidata <- npidata %>%
   select(npi, name, t_class)
@@ -90,8 +90,19 @@ npidata <- npidata %>%
 sure_matches <- sure_matches %>%
   left_join(npidata, by=c("name"))
 
-# join back to cleaned_text
-names_data <- names_data %>%
+# duplicates?
+dups <- sure_matches %>%
+  mutate(count=1) %>%
+  group_by(name, ein) %>%
+  mutate(sum=sum(count)) %>%
+  ungroup() %>%
+  filter(sum>1)
+
+sure_matches <- sure_matches %>%
+  distinct()
+
[…]

### 2024-09-03 RScripts/prelim_regressions.R

@@ -1,347 +1,0 @@
-library(readr)
-library(dplyr)
-library(fixest)
-library(ggplot2)
-library(ggpubr)
-library(tidyr)
-
-# Read in hospital_data created in "create_hospital_data.R"
-hospital_data <- readRDS(paste0(created_data_path, "all_hospital_data.rds")) %>%
-  mutate(yr2009 = ifelse(year==2009,1,0),
-         yr2010 = ifelse(year==2010,1,0),
-         yr2011 = ifelse(year==2011,1,0),
-         yr2012 = ifelse(year==2012,1,0),
-         yr2013 = ifelse(year==2013,1,0),
-         yr2014 = ifelse(year==2014,1,0),
-         yr2015 = ifelse(year==2015,1,0)) %>%
-  rename(heartattack_mortality = rate_heartattack_mortality,
-         heartattack_readmission = rate_heartattack_readmission,
-         heartfailure_mortality = rate_heartfailure_mortality,
-         heartfailure_readmission = rate_heartfailure_readmission,
-         pneum_mortality = rate_pneum_mortality,
-         pneum_readmission = rate_pneum_readmission) %>%
-  mutate(post_2012 = ifelse(year>=2012, 1, 0)) %>%
-  mutate(post_2011 = ifelse(year>=2011, 1, 0)) %>%
-  mutate(has_any_md = ifelse(num_doctors>0, 1, 0)) %>%
-  mutate(did = has_any_md*post_2012) %>%
-  mutate(did_2011 = has_any_md*post_2011) %>%
-  mutate(es2009 = yr2009*has_any_md,
-         es2010 = yr2010*has_any_md,
-         es2011 = yr2011*has_any_md,
-         es2012 = yr2012*has_any_md,
-         es2013 = yr2013*has_any_md,
-         es2014 = yr2014*has_any_md,
-         es2015 = yr2015*has_any_md)
-
-# create min year penalized variable
-minyr_penalized <- hospital_data %>%
-  mutate(year_penalized = ifelse(penalized_HC==1, year, NA)) %>%
-  filter(!is.na(year_penalized)) %>%
-  group_by(ID) %>%
-  mutate(minyr_pen = min(year)) %>%
-  ungroup() %>%
-  distinct(ID, minyr_pen)
-
-hospital_data <- hospital_data %>%
-  left_join(minyr_penalized, by="ID")
-
-# create sample of no md changes for all hospitals
-noMDchg_all_sample <- hospital_data %>%
[…]

### 2024-09-03 RScripts/summary_stats.R

@@ -1,750 +1,0 @@
-library(readr)
-library(dplyr)
-library(tidyr)
-library(knitr)
-library(kableExtra)
-library(weights)
-library(ggplot2)
-library(maps)
-library(mapdata)
-library(cdlTools)
-library(tibble)
-
-# read in hospital data
-hospital_data <- read_rds(paste0(created_data_path, "all_hospital_data.rds"))
-
-# put uncompensated care in millions
-hospital_data$uncomp_care <- abs(hospital_data$uncomp_care)
-hospital_data <- hospital_data %>%
-  mutate(uncomp_care=uncomp_care/1000000) 
-
-# drop hospitals with less than 10 beds
-hospital_data <- hospital_data %>%
-  group_by(ID) %>%
-  mutate(maxbeds=max(beds, na.rm=T)) %>%
-  ungroup() %>%
-  filter(maxbeds>=10) %>%
-  select(-maxbeds)
-
-
-# List of Figures:
-# 1. Graph outcomes over time for forprofit and nonprofit
-# 2. Graph outcomes over time for MD and non-MD
-
-
-# List of tables:
-# Table 1: mean, SD, min, max for the whole sample
-# Table 2: means for different sub-samples
-  
-
-
-## FIGURE 1: outcomes for forprofit and nonprofit over time ####
-forprofit_avgs <- hospital_data %>%
-  filter(profit_status=="forprofit") %>%
-  group_by(year) %>%
-  summarise_at(c("Weighted Avg. Readmission Rate"="weightedavg_read",
-                 "Weighted Avg. Mortality Rate"="weightedavg_mort"), list(mean), na.rm=T) %>%
-  mutate(Group="For Profit")
-nonprofit_avgs <- hospital_data %>%
-  filter(profit_status=="nonprofit") %>%
-  group_by(year) %>%
-  summarise_at(c("Weighted Avg. Readmission Rate"="weightedavg_read",
-                 "Weighted Avg. Mortality Rate"="weightedavg_mort"), list(mean), na.rm=T) %>%
-  mutate(Group="All Nonprofits")
-NP_md_avgs <- hospital_data %>%
-  filter(ever_has_md==1 & no_num_md_change_2010_2014==1) %>%
-  group_by(year) %>%
-  summarise_at(c("Weighted Avg. Readmission Rate"="weightedavg_read",
[…]

### 2024-09-03 RScripts/sumstats_profitstatus.R

@@ -1,73 +1,0 @@
-library(readr)
-library(dplyr)
-library(tidyr)
-library(knitr)
-library(kableExtra)
-library(weights)
-library(ggplot2)
-
-# create graphs exploring differences in outcomes between nonprofit and for-profit ownership status
-
-hospital_data <- read_rds(paste0(created_data_path, "all_hospital_data.rds")) %>%
-  filter(year!=2016 & year!=2017)
-
-# create "ever has MD" variable
-hospital_data <- hospital_data %>%
-  group_by(ID) %>%
-  mutate(ever_has_md = ifelse(sum(has_any_md, na.rm=T)>0,1,0)) %>%
-  ungroup() 
-
-# create variable for "has leadership info"
-hospital_data <- hospital_data %>%
-  group_by(ID) %>%
-  mutate(has_leader_info = ifelse(sum(num_execs,na.rm=T)>0,1,0)) %>%
-  ungroup()
-
-# compare weighted avg readmissions for: for profit, NFP w/ leadership info, NFP w/out leadership info
-hospital_data <- hospital_data %>%
-  mutate(profit_leadership_status = ifelse(profit_status=="forprofit","forprofit",NA)) %>%
-  mutate(profit_leadership_status = ifelse(profit_status=="nonprofit" & has_leader_info==1,"nonprofit w/ leader info",profit_leadership_status)) %>%
-  mutate(profit_leadership_status = ifelse(profit_status=="nonprofit" & has_leader_info==0,"nonprofit w/out leader info",profit_leadership_status)) 
-
-sum <- hospital_data %>%
-  group_by(year,profit_leadership_status) %>%
-  summarise_at(c("weightedavg_read"), list(mean), na.rm=T) %>%
-  filter(year!=2016 & year!=2017)
-
-ggplot(sum, aes(x=year, y=weightedavg_read, color=profit_leadership_status)) + geom_line()
-
-# compare weighted avg readmissions for: for profit, NFP w/ stable team, NFP w/out stable team
-hospital_data <- hospital_data %>%
-  group_by(MCRNUM) %>%
-  fill(no_md_change_2010_2014, .direction="downup") %>%
-  ungroup() %>%
-  mutate(profit_stability_status = ifelse(profit_status=="forprofit","forprofit",NA)) %>%
[…]

### 2024-09-03 RScripts/tables_and_figures.R

@@ -1,0 +1,1763 @@
+library(dplyr)
+library(readr)
+library(knitr)
+library(kableExtra)
+library(ggplot2)
+library(ggpubr)
+library(RColorBrewer)
+library(synthdid)
+library(abind)
+library(reshape2)
+library(data.table)
+
+options(knitr.kable.NA=" ")
+group.colors <- c("For-profit" = "#D65828", "Never Clinical Exec" = "#C2DEF2", "Ever Clinical Exec" ="#6BAED6", "Always Clinical Exec" = "#2C6B8E")
+ 
+# Produce all tables and figures for the body of the paper #########
+
+# read in data
+hospital_data <- read_rds(paste0(created_data_path, "all_hospital_data3.rds"))
+doc_exec_data <- read_rds(paste0(created_data_path, "doc_exec_data.rds"))
+
+hospital_data <- hospital_data %>%
+  mutate(ever_pen_ha_only=ifelse(ever_pen_ha>=1 & ever_pen_hf==0 & ever_pen_pnem==0,1,0), 
+         ever_pen_hf_only=ifelse(ever_pen_hf>=1 & ever_pen_ha==0 & ever_pen_pnem==0,1,0), 
+         ever_pen_pnem_only=ifelse(ever_pen_pnem>=1 & ever_pen_hf==0 & ever_pen_ha==0,1,0), 
+         ever_pen_ami_hf = ifelse(ever_pen_hf>=1 & ever_pen_ha>=1 & ever_pen_pnem==0,1,0),
+         ever_pen_ami_pneum = ifelse(ever_pen_hf==0 & ever_pen_pnem>=1 & ever_pen_ha==1,1,0),
+         ever_pen_hf_pneum = ifelse(ever_pen_hf>=1 & ever_pen_pnem>=1 & ever_pen_ha>=0,1,0),
+         ever_pen_all = ifelse(ever_pen_hf>=1 & ever_pen_ha>=1 & ever_pen_pnem>=1,1,0)) %>%
+  mutate(frac_doc_exec = num_doctors/num_execs)
+
+
+# TABLE 1: summary statistics of individual executives who are clinically trained ####
+doc_exec_data <- doc_exec_data %>%
+  mutate(surgery_doc = ifelse(t_class == "Surgery", 1, 0),
+         anest_doc = ifelse(t_class == "Anesthesiology", 1, 0),
+         family_doc = ifelse(t_class == "Family Medicine",1,0),
[…]

### 2024-10-25 Slides/JobTalk.tex
Updates from Overleaf

\documentclass[notes,11pt, aspectratio=169]{beamer}

\usepackage{pgfpages}
% These slides also contain speaker notes. You can print just the slides,
% just the notes, or both, depending on the setting below. Comment out the want
% you want.
\setbeameroption{hide notes} % Only slide
%\setbeameroption{show only notes} % Only notes
%\setbeameroption{show notes on second screen=right} % Both

\usepackage{helvet}
\usepackage[default]{lato}
\usepackage{array}

\usepackage[backend=biber,style=authoryear,
sorting=ynt,citestyle=authoryear]{biblatex}
\addbibresource{Paper/papercitations.bib}

\usepackage{tikz}
\usepackage{verbatim}
\setbeamertemplate{note page}{\pagecolor{yellow!5}\insertnote}
\usetikzlibrary{positioning}
\usetikzlibrary{snakes}
\usetikzlibrary{calc}
\usetikzlibrary{arrows}
\usetikzlibrary{decorations.markings}
\usetikzlibrary{shapes.misc}
\usetikzlibrary{matrix,shapes,arrows,fit,tikzmark}
\usepackage{amsmath}
\usepackage{mathpazo}
\usepackage{hyperref}
\usepackage{lipsum}
\usepackage{multimedia}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{graphicx}
\usepackage{dcolumn}
\usepackage{bbm}
\newcolumntype{d}[0]{D{.}{.}{5}}
\usepackage{subfigure}
\usepackage{import}

\usepackage{changepage}
\usepackage{appendixnumberbeamer}
\newcommand{\beginbackup}{
   \newcounter{framenumbervorappendix}
   \setcounter{framenumbervorappendix}{\value{framenumber}}
   \setbeamertemplate{footline}
   {
     \leavevmode%
     \hline
     box{%
       \begin{beamercolorbox}[wd=\paperwidth,ht=2.25ex,dp=1ex,right]{footlinecolor}%
%         \insertframenumber  \hspace*{2ex} 
       \end{beamercolorbox}}%
     \vskip0pt%
   }
 }
\newcommand{\backupend}{
   \addtocounter{framenumbervorappendix}{-\value{framenumber}}
   \addtocounter{framenumber}{\value{framenumbervorappendix}} 
}

\setbeamertemplate{blocks}[rounded]
\setbeamercolor{block title}{bg=eggplant!70, fg=white}
\setbeamercolor{block body}{bg=eggplant!20}

\usepackage{graphicx}
\usepackage[space]{grffile}
\usepackage{booktabs}

% These are my colors -- there are many like them, but these ones are mine.
\definecolor{sage}{RGB}{102,153,102}
\definecolor{cambridgeblue}{RGB}{104, 166, 145}
\definecolor{yellow}{RGB}{255,173,1}
\definecolor{purple}{RGB}{153,102,153}
\definecolor{ashgray}{RGB}{191, 211, 193}
\definecolor{eggplant}{RGB}{105, 79, 93}
\definecolor{silver}{RGB}{192, 197, 193}
\definecolor{darkcyan}{RGB}{78, 135, 140}
\definecolor{lightcyan}{RGB}{201, 228, 231}
\definecolor{columbiablue}{RGB}{187, 213, 237}
\definecolor{lightblue}{RGB}{184, 219, 217}
\definecolor{earthyellow}{RGB}{234, 180, 100}

\setbeamercolor{mycolor}{fg=white,bg=sage}

% colors for diagrams
\definecolor{diagramtan}{RGB}{225,190,106}
\definecolor{diagramteal}{RGB}{64, 176, 166}
\definecolor{diagrampurple}{RGB}{170, 131, 239}
\definecolor{diagramred}{RGB}{146, 79, 79}

\hypersetup{
  colorlinks=false,
  linkbordercolor = {white},
  linkcolor = {sage}
}

\newcommand{\btVFill}{\vskip0pt plus 1filll}

%% I use a beige off white for my background
\definecolor{MyBackground}{RGB}{255,253,218}

%% Uncomment this if you want to change the background color to something else
%\setbeamercolor{background canvas}{bg=MyBackground}

%% Change the bg color to adjust your transition slide background color!
\newenvironment{transitionframe}{
  \setbeamercolor{background canvas}{bg=lightcyan}
  \begin{frame}[plain,noframenumbering]}{
    \end{frame}
}

\setbeamercolor{frametitle}{fg=black, bg=lightcyan}
\setbeamercolor{title}{fg=black}
\setbeamertemplate{footline}{%
  \raisebox{5pt}{\makebox[\paperwidth]{\hfill\makebox[15pt]{\scriptsize\insertframenumber}}}}
\setbeamertemplate{navigation symbols}{} 
\setbeamertemplate{itemize items}{$\bullet$}
\setbeamercolor{itemize item}{fg=lightcyan}
\setbeamercolor{itemize subitem}{fg=lightcyan}
\setbeamercolor{enumerate item}{fg=black}
\setbeamercolor{enumerate subitem}{fg=black}
\setbeamercolor{button}{bg=eggplant!20,fg=black}

\setbeamertemplate{frametitle}
{
    \nointerlineskip
    \begin{beamercolorbox}[sep=0.3cm,ht=2em,wd=\paperwidth]{frametitle}
        \vbox{}\vskip-2ex%
        \strut\insertframetitle\strut
        \vskip-0.8ex%
    \end{beamercolorbox}
}

% If you like road maps, rather than having clutter at the top, have a roadmap show up at the end of each section 
% (and after your introduction)
% Uncomment this is if you want the roadmap!
% \AtBeginSection[]
% {
%    \begin{frame}
%        \frametitle{Roadmap of Talk}
%        \tableofcontents[currentsection]
%    \end{frame}
% }

\setbeamercolor{section in toc}{fg=sage}
\setbeamercolor{subsection in toc}{fg=sage}
\setbeamersize{text margin left=1em,text margin right=1em}

\newenvironment{wideitemize}{\itemize\addtolength{\itemsep}{10pt}}{\enditemize}

\usepackage{environ}
\NewEnviron{videoframe}[1]{
  \begin{frame}
    \vspace{-8pt}
    \begin{columns}[onlytextwidth, T] % align columns
      \begin{column}{.58\textwidth}
        \begin{minipage}[t][\textheight][t]
          {\dimexpr\textwidth}
          \vspace{8pt}
          \hspace{4pt} {\Large \sc \textcolor{blue}{#1}}
          \vspace{8pt}

          \BODY
        \end{minipage}
      \end{column}%
      \hfill%
      \begin{column}{.42\textwidth}
        \colorbox{green!20}{\begin{minipage}[t][1.2\textheight][t]
            {\dimexpr\textwidth}
            Face goes here
          \end{minipage}}
      \end{column}%
    \end{columns}
  \end{frame}
}

\title[]{\textcolor{sage}{Does Hospital Leadership Matter? \newline
Evidence from Pay-for-Performance}}

\author[]{Hanna Glenn}
\date{}

\begin{document}

%%% TIKZ STUFF
\tikzset{   
        every picture/.style={remember picture,baseline},
        every node/.style={anchor=base,align=center,outer sep=1.5pt},
        every path/.style={thick},
        }
\newcommand\marktopleft[1]{%
    \tikz[overlay,remember picture] 
        \node (marker-#1-a) at (-.3em,.3em) {};%
}
\newcommand\markbottomright[2]{%
    \tikz[overlay,remember picture] 
        \node (marker-#1-b) at (0em,0em) {};%
}
\tikzstyle{every picture}+=[remember picture] 
\tikzstyle{mybox} =[draw=black, very thick, rectangle, inner sep=10pt, inner ysep=20pt]
\tikzstyle{fancytitle} =[draw=black,fill=red, text=white]
%%%% END TIKZ STUFF

% Title Slide
\begin{frame}[plain]
% Background image
\begin{tikzpicture}[remember picture,overlay]
    \node
[
    above=0.5cm,
    align=center,
    fill=lightcyan,
    rounded corners,
    inner xsep=15pt,
    inner ysep=10pt, 
    minimum width=0.9\textwidth,
    text width=0.6\textwidth
] (title) at (current page.center)
{
    \LARGE Does Hospital Leadership Matter?  \\[5pt]
    \normalsize Evidence from Pay-for-Performance
};
% Author 
\node[below=1cm] (author) at (title.south){Hanna Glenn, Emory University};
% Date
\node[below=0.25cm] (date) at (author.south){\today};
\end{tikzpicture}

\end{frame}

\section{Intro}

\begin{frame}{Firm Leadership}
\centering
    \begin{tikzpicture}
        % First image (top-left)
        \node[anchor=north west] (img1) at (0, 3) {\includegraphics[width=0.55\textwidth]{Graphics/theroleleadershiphasincompanyculture.png}};
        \draw[line width=.45mm, lightcyan] (img1.north west) rectangle (img1.south east);

        % Second image (slightly down and right of the first image)
        \node[anchor=north west] (img2) at (0.28, 0.58) {\includegraphics[width=0.5\textwidth]{Graphics/linkingbusinessstrategyandleadershipforbetteroutcomes.png}};
        \draw[line width=.45mm, lightcyan] (img2.north west) rectangle (img2.south east);

        % Fourth image (bottom-right)
        \node[anchor=north west] (img4) at (0, -2) {\includegraphics[width=0.55\textwidth]{Graphics/theroleofleadershipinorganizationalsuccess.png}};
        \draw[line width=.45mm, lightcyan] (img4.north west) rectangle (img4.south east);

    \end{tikzpicture}
\end{frame}

\begin{frame}{Leadership Characteristics}

\vspace{1mm}

\textbf{Gender}
\begin{itemize}
    \item \small More female board members $\implies$ lower firm value \scriptsize(\cite{ahern2012changing}) \small
    \item \small More female board members $\implies$ less employment reduction \scriptsize (\cite{matsa2013female}) \small
    \item female executives affect the female wage distribution \scriptsize (\cite{flabbi2019female}) \small
\end{itemize}

\vspace{2mm}

\textbf{Background}
\begin{itemize}
    \item \small Firms pay more for general background CEOs \scriptsize (\cite{custodio2013generalists}) \small
    \item \small CEOs w/ prior experience do better in M\&A \scriptsize(\cite{custodio2013ceos}) \small
    \item \small Military experience CEOs are less aggressive in R\&D and investing \scriptsize (\cite{benmelech2015military})
\end{itemize}

\vspace{2mm}

\textbf{Age}
\begin{itemize}
    \item \small Young male CEOs more aggressive in M\&A \scriptsize (\cite{levi2010deal})
\end{itemize}

\btVFill\pause

\begin{block}{\normalsize Gaps in Literature}
    \small 
    \begin{enumerate}
        \item Limited to large, publicly traded firms: reason to believe this does not extend to nonprofit setting
        \item Direct responses to policy changes are not well-understood
    \end{enumerate}
\end{block}
\end{frame}

\begin{frame}{Hospitals \& Health Care Policy Goals}

\begin{block}{This Paper}
    Examine the role of leadership in nonprofit hospitals responding to policy incentives 
\end{block}

\vspace{5mm}

\textbf{Hospital care makes up a third of health care spending} \scriptsize (\cite{cms2024nhe}) \normalsize
\begin{itemize}
    \item A majority of hospitals are nonprofit and their objectives are not as clear as classic for-profit firms
\end{itemize}

\vspace{5mm}

\textbf{Policies often aim to improve the value of health care delivery}

\vspace{3mm}

    \begin{wideitemize}
        \item It is unclear how hospital leaders view value
        \item Or whether they might play a role in achieving this goal
    \end{wideitemize}
\end{frame}

\begin{frame}{Research Question}
    \centering
    \large
    Does having executives with clinical training affect response to pay-for-performance incentives of nonprofit hospitals?
\end{frame}

\begin{frame}{Why Clinical Training?}

\textbf{Physician executives could be good or bad}

\vspace{2mm}

\begin{wideitemize}
    \item Physicians can bring a unique combination of clinical and administration expertise that can potentially lead to better quality \scriptsize (\cite{Stajduhar_2023}, \cite{Ahmed_2022}) \normalsize

    \vspace{2mm}

    \item Not always trained in business management, which can lead to poor practices and lower quality \scriptsize (\cite{HarvardBusinessReview2018})
\end{wideitemize}

\vspace{10mm}

\textbf{Clinical training is potentially correlated with how the leader values quality}
\end{frame}

\begin{frame}{Preview of Findings}

\textbf{Non-clinical executive teams respond more drastically to financial incentives on quality compared to clinical executive teams}

\vspace{2mm}

\begin{wideitemize}
    \item Non-clinical reduce readmissions by blank more than clinical
    \item Non-clinical executive teams behave like profit-driven hospitals
\end{wideitemize}
\end{frame}

\begin{frame}{Contribution}
    \textbf{Hospital Executives/Board of Directors}
    \begin{itemize}
        \item Physician board members $\implies$ lower public donations in nonprofit hospitals \scriptsize (\cite{brickley2010board}) \normalsize
        \item No correlation between CEOs and hospital production in England \scriptsize (\cite{janke2019impact})\normalsize
        \item Public hospitals in Chile improved when older physicians were removed as CEO \scriptsize (\cite{otero2022managers}) \normalsize
    \end{itemize}

    \vspace{5mm}

    \textbf{Hospital Response to Pay-for-Performance}
    \begin{itemize}
        \item Hospitals decreased readmissions by 40\% on average, partially due to selective patient practices \scriptsize (\cite{gupta2021impacts}) \normalsize
        \item Market concentration and systems affect response \scriptsize (\cite{kunz2024assessing}) \normalsize
        \item Reward program had little impact on quality \scriptsize (\cite{us2015hospital}; \cite{norton2018moneyball}; \cite{friedson2019so})
    \end{itemize}
\end{frame}

\begin{frame}{Rest of Talk}
    Theoretical Motivation

    \vspace{4mm}

    Setting and Data Collection

    \vspace{4mm}

    About Executive Teams

    \vspace{4mm}

    Effect of Clinical Leadership

    \vspace{4mm}

    Comparison to For-Profits

    \vspace{4mm}

    Signaling vs. Managing
\end{frame}

\section{Theoretical Framework}

\begin{transitionframe}
\centering \LARGE
    \textcolor{black}{Theoretical Framework}
\end{transitionframe}

\begin{frame}{Hospital Objectives}\label{hosp_obj}
\large
    \begin{equation*}
        \max_{\theta}\hspace{2mm}\alpha\pi(\theta) + (1-\alpha) u(\theta)
    \end{equation*}

    \vspace{5mm} \normalsize

    \begin{wideitemize}
        \item Profit $\pi$ and patient welfare $u$; depend on quality $\theta$
        \item Profit-driven hospitals have high $\alpha$
        \item Abstract away from quality impacting demand for care, price \hyperlink{complex_model}{\beamergotobutton{more complex model}} 
    \end{wideitemize}
\end{frame}

\begin{frame}{Choice of Quality}
Take FOC, rearrange and take derivative wrt $\alpha$:

\vspace{5mm}

\begin{equation*}
    \frac{du'(\theta)}{d\alpha} = \frac{-1}{(1-\alpha)^2}\pi'(\theta)
\end{equation*}

\vspace{5mm}

$\pi'(\theta)$:

\vspace{3mm}

\begin{wideitemize}
    \item In fee-for-service setting: negative $\implies \frac{du'(\theta)}{d\alpha} > 0$
    \item In pay-for-performance setting: positive $\implies \frac{du'(\theta)}{d\alpha} < 0$
\end{wideitemize}

\end{frame}

\begin{frame}{FFS $\implies$ P4P}
    \large
    \begin{align*}
        \frac{d\Delta\theta}{d\alpha}&=\frac{d\theta_{p4p}}{d\alpha}-\frac{d\theta_{ffs}}{d\alpha}\\ 
        &\\
        &=\frac{\frac{du'(\theta)}{d\alpha}}{\frac{du'(\theta)}{d\theta_{p4p}}} - \frac{\frac{du'(\theta)}{d\alpha}}{\frac{du'(\theta)}{d\theta_{ffs}}}\\
        &\\
        &> 0
    \end{align*}

    \btVFill \normalsize

    \begin{block}{Takeaway}
        Hospitals with higher weight on profit will respond more to the financial incentives on quality
    \end{block}
\end{frame}

\begin{frame}{Testable Predictions}
    \begin{enumerate}
        \item If non-clinical executives are more profit-driven than clinical executive hospitals, they will reduce quality by more

        \vspace{5mm}
        \item If non-clinical executives are more profit-driven, they will act more like for-profit hospitals

        \vspace{5mm}
        \item Executive teams could reflect an underlying $\alpha$ or they could effectively change $\alpha$
    \end{enumerate}
\end{frame}

\section{Setting and Data Collection}

\begin{transitionframe}
\centering \LARGE
    \textcolor{black}{Setting and Data Collection}
\end{transitionframe}

\begin{frame}{Nonprofit Hospital Executives}\label{nonprofithospexec}
\textbf{What do executives do?}
\begin{wideitemize}
    \item Board of directors selects executives
    \item Executives handle the day-to-day operations of the hospital \hyperlink{executivejob}{\beamergotobutton{executive job}} 
\end{wideitemize}

\vspace{7mm}

\textbf{Executive teams}
\begin{wideitemize}
    \item Lots of variation in what executive teams look like
    \item Typically CEO and CFO, sometimes CMO 
\end{wideitemize}

\vspace{7mm}

Usable data on executives is rare
\end{frame}

\begin{frame}{}
    \begin{tikzpicture}
        % First image in the top-left
        \node[anchor=north west] at (0, 4) {\includegraphics[width=0.7\textwidth]{Graphics/990_snip_frontpage.PNG}};
        % Second image in the middle
        \node[anchor=north west] (img2) at (2.5, 1.5) {\includegraphics[width=0.75\textwidth]{Graphics/Screenshot 2024-10-03 160258.png}};
        \draw[line width=1mm, eggplant] (img2.north west) rectangle (img2.south east);
    \end{tikzpicture}
\end{frame}

\begin{frame}{Data Collection Process}
    \begin{enumerate}
        \item Get list of Employee Identification Numbers (EIN) that represent nonprofit hospitals

        \vspace{5mm}
        \item Download tax forms from 2009-2015 from Nonprofit Explorer API

        \vspace{5mm}
        \item Use Optimal Character Recognition to extract text from all of these documents

        \vspace{5mm}
        \item Write string cleaning algorithms to extract the correct name, title, position of each entry
    \end{enumerate}
\end{frame}

\begin{frame}{Merge to Other Hospital Data}
    I create links between EIN and ID number from the American Hospital Association Survey (AHA) based on hospital name and location

    \vspace{7mm}
    Merge to:

    \vspace{2mm}
    \begin{itemize}
        \item AHA data for hospital demographics
        \item Hospital Compare data for readmission and mortality rates
        \item Center for Medicare and Medicaid Services (CMS) Case Mix Index file for patient complexity
        \item Health Care Cost Report Information System (HCRIS) for information on pay-for-performance rewards or penalties
    \end{itemize}
\end{frame}

\begin{frame}{Merge to Physician Data}
    I use executives that self-identify as MD

    \vspace{8mm}

    To verify, I match all executive names to database of all registered physicians and match NPI by hand
\end{frame}

\begin{frame}{About Pay-for-Performance Policies}
    Policies enacted in 2012 that either penalized for low quality or rewarded for high quality

    \vspace{3mm}

    \textbf{Hospital Readmissions Reduction Program}
    \begin{itemize}
        \item If hospitals have above average readmission rates they get 1-3\% lower reimbursement rates
        \item Pneumonia, Heart Attack (AMI), Heart Failure
    \end{itemize}

    \vspace{3mm}

    \textbf{Hospital Value-Based Purchasing Program}
    \begin{itemize}
        \item safety, efficiency, cost reductions, clinical outcomes, and community engagement, improvement combined into single score metric
        \item ``good" hospitals rewarded
    \end{itemize}

    \pause

    \begin{block}{\normalsize Takeaway}
        \normalsize All hospitals in the middle of the distribution had incentive to improve quality to try to avoid penalties, which were not perfectly predictable
    \end{block}
\end{frame}

\section{About Executive Teams}

\begin{transitionframe}
\centering \LARGE
    \textcolor{black}{Descriptive Statistics}
\end{transitionframe}

\begin{frame}{Executive Teams}
    \begin{table}[ht!]
        \centering
        \begin{tabular}[t]{lccc}
        \toprule
         & Ever & Always & Never\\
         & Clinical Exec & Clinical Exec &  Clinical Exec \\
        \midrule
        Number Executives & 4.85 & 5.25 & 3.11\\
        Frac Clinical Execs & 0.20 & 0.26 & 0\\
        Frac Int. Medicine Execs & 0.06 & 0.10 & 0.0\\
        Has a CMO & 0.53 & 0.45 & 0.15\\
        Has Clinical CEO & 0.25 & 0.11 & 0.0\\
        \addlinespace[0.6em]
        Num. Hospitals & 391 & 56 & 445\\
        \bottomrule
        \end{tabular}
    \end{table}
\end{frame}

\begin{frame}{Doctors on Executive Teams (N=813)}
    \begin{table}[ht!]
    \centering
    \begin{tabular}[t]{p{6cm} c}
    \toprule
     & Mean\\
    \midrule
    Age & 52.5\\
    Female & 0.11\\
    CEO & 0.16\\
    CMO & 0.31\\
    \addlinespace[0.3em]
    \multicolumn{2}{l}{\textbf{Specialty}}\\
    \hspace{1em}Internal Medicine & 0.32\\
    \hspace{1em}Family Medicine & 0.17\\
    \hspace{1em}Emergency Medicine & 0.11\\
    \hspace{1em}Surgery & 0.09\\
    \hspace{1em}Anesthesiologist & 0.04\\
    \hspace{1em}Other & 0.28\\
    \bottomrule
    \end{tabular}
    \end{table}
\end{frame}

\begin{frame}{Average Team Changes Over Time}
\centering 
    \includegraphics[width=.8\textwidth]{Objects/change_means.pdf}
\end{frame}

\begin{frame}{Changes are not correlated with policy exposure}
\centering \small
\begin{equation*}\label{eq:change2}
    \text{change}_{ht} = \sum_{j=2011}^{2014}\beta_j(\mathbf{1}\{t=j\}\times \text{Program Exposed})_{ht} + \alpha_h + \epsilon_{ht},
    \end{equation*}

    \btVFill

    \includegraphics[width=.6\textwidth]{Objects/change_analysis_plot.pdf}

\end{frame}

\begin{frame}{Hospital Statistics}
\vspace{-2mm}
    \begin{table}[ht!]
    \centering
    \begin{tabular}[t]{lccc}
     & Ever & Always & Never\\
      & Clinical Exec & Clinical Exec & Clinical Exec\\
    \midrule
    \addlinespace[0.3em]
    \multicolumn{4}{l}{\textbf{Hospital Characteristics}}\\
    \hspace{1em}Academic Med. Center & 0.41 & 0.52 & 0.24\\
    \hspace{1em}Number Beds & 231 & 222 & 116\\
    \hspace{1em}Physician Owned & 0.01 & 0.02 & 0.02\\
    \hspace{1em}System Affiliated & 0.58 & 0.52 & 0.44\\
    \addlinespace[0.3em]
    \multicolumn{4}{l}{\textbf{Penalty/Payment Variables}}\\
    \hspace{1em}Ever Received HVBP Incentive & 0.79 & 0.75 & 0.62\\
    \hspace{1em}Penalized for One Condition & 0.18 & 0.20 & 0.13\\
    \hspace{1em}Penalized for AMI + HF/Pneum. & 0.09 & 0.09 & 0.06\\
    \hspace{1em}Penalized for HF + Pneumonia & 0.41 & 0.39 & 0.34\\
    \hspace{1em}Penalized for All Conditions & 0.30 & 0.27 & 0.22\\
    Num. Hospitals & 391 & 56 & 445\\
    \bottomrule
    \end{tabular}
    \end{table}
\end{frame}

\begin{frame}{}
    put graph of outcomes here
\end{frame}

\section{Effect of Clinical Leadership on Response}

\begin{transitionframe}
\centering \LARGE
    \textcolor{black}{Effect of Clinical Leadership on Response}
\end{transitionframe}

\begin{frame}{Estimation}\label{est_equation}

    \vspace{5mm}\large
    \begin{equation*}
    y_{ht} = \beta \text{ (clinical training x post 2012)}_{ht} + \gamma_{h} + \delta_t + \epsilon_{ht}
    \end{equation*}
    \begin{itemize}
        \item Compare always to never clinical executives
        \item Leverage the timing of the programs
    \end{itemize}

    \vspace{8mm}

    Identifies the effect of having a clinically trained executive on readmission and mortality rates after pay-for-performance

    \vspace{2mm}

    \begin{itemize}
        \item Assume that hospitals would have had parallel trends absent clinical executives
    \end{itemize}

    \btVFill
    \normalsize
    \hyperlink{synthdid_equation}{\beamergotobutton{weighted}} \hyperlink{match_equation}{\beamergotobutton{matching}}
\end{frame}

\begin{frame}{Readmission Rates}\label{main:readrates}
\centering
    \includegraphics[width=.45\textwidth]{Objects/read_md_nomd_synth_graph.pdf}

    \btVFill

    \hyperlink{int_margin}{\beamergotobutton{intensive margin}}
\end{frame}

\begin{frame}{Mortality Rates}

\end{frame}

\section{Comparison to For-Profit Hospitals}

\begin{transitionframe}
\centering \LARGE
    \textcolor{black}{Comparison to For-Profit Hospitals}
\end{transitionframe}

\begin{frame}{Estimation}

    \vspace{5mm}\large
    \begin{equation*}
    y_{ht} = \beta \text{ (for-profit x post 2012)}_{ht} + \gamma_{h} + \delta_t + \epsilon_{ht}
    \end{equation*}
    \begin{itemize}
        \item Two comparison groups: clinical and non-clinical executive teams
    \end{itemize}

\end{frame}

\begin{frame}{Results}

\end{frame}

\section{Signaling vs. Managing}

\begin{transitionframe}
\centering \LARGE
    \textcolor{black}{Signaling vs. Managing}
\end{transitionframe}

\begin{frame}{Estimation}

\end{frame}

\begin{frame}{Results}

\end{frame}

\begin{frame}{Conclusion}

\end{frame}

\begin{frame}{Future Work}

\end{frame}

\begin{transitionframe}
\centering \Large
    \textcolor{black}{hkagele@emory.edu}
\end{transitionframe}

\section{Appendix}

\begin{transitionframe}
\centering \LARGE
    \textcolor{black}{Appendix Slides}
\end{transitionframe}

\appendix

\begin{frame}{More Complex Model}\label{complex_model}

\btVFill

\hyperlink{hosp_obj}{\beamerreturnbutton{hospital objectives}}

\end{frame}

\begin{frame}{Executive Job}\label{executivejob}

\btVFill

\hyperlink{nonprofithospexec}{\beamerreturnbutton{Nonprofit Hosp Execs}}

\end{frame}

\begin{frame}{Using synthetic DiD weights}\label{synthdid_equation}

\btVFill

\hyperlink{est_equation}{\beamerreturnbutton{main equation}}

\end{frame}

\begin{frame}{Matching}\label{match_equation}

\btVFill

\hyperlink{est_equation}{\beamerreturnbutton{main equation}}

\end{frame}

\begin{frame}{Intensive Margin}\label{int_margin}

    \btVFill

    \hyperlink{main:readrates}{\beamerreturnbutton{main results}}
\end{frame}

\end{document}

### 2024-10-25 Paper/paperdraft_Sep2024.tex

--- 2024-09-03 Paper/paper2.tex
@@
 Further, I more heavily weight time periods that balance pre- and post-program outcomes for non-clinical teams.
-[description of weight distribution here]
+The control group consists of approximately 240 hospitals, and I show the fairly uniform distribution of their weights in Appendix \ref{app:sumstats}.
 
@@
 
-    The hypothesis driven from this theory is that clinically trained executives put more weight on quality of care before the financial incentives on quality began.
-To investigate this hypothesis empirically, I merge all general for-profit hospitals in the AHA data to the sample of nonprofit hospitals.
+    The hypothesis driven from the theory described in Section \ref{sec:forprofit} is that clinically trained executives put more weight on quality of care before the financial incentives on quality began.
+To investigate this hypothesis empirically, I merge all general for-profit hospitals in the AHA survey to my sample of nonprofit hospitals.
 Then, I compare readmission and mortality rates of always clinical executives and never clinical executives to for-profit hospitals after the incentive change.
@@
 
+\begin{figure}
+    \centering
+    \caption{Control Group Weights, Readmission and Baseline Case Mix}
+    \includegraphics[width=0.8\linewidth]{Objects/main_read_control_weights.pdf}
+    \label{fig:main_read_control_weights}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \caption{Time Weights, Readmission and Baseline Case Mix}
+    \includegraphics[width=0.8\linewidth]{Objects/main_read_time_weights.pdf}
+    \label{fig:main_read_time_weights}
+\end{figure}
+
+\newpage
+
+\begin{figure}
+    \centering
+    \caption{Control Group Weights, Mortality and Baseline Case Mix}
+    \includegraphics[width=0.8\linewidth]{Objects/main_mort_control_weights.pdf}
+    \label{fig:main_mort_control_weights}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \caption{Time Weights, Mortality and Baseline Case Mix}
+    \includegraphics[width=0.8\linewidth]{Objects/main_mort_time_weights.pdf}
+    \label{fig:main_mort_time_weights}
+\end{figure}
+
 \section{HRRP and HVBP Programs}\label{app:programs}

### 2024-10-25 Paper/paper.tex

@@
 
-    \import{Tables}{leader_sumstats.tex}
+    %\import{Tables}{leader_sumstats.tex}
 
@@
 
-    \begin{figure}[ht!]
-    \centering
-        \caption{Outcomes Over Time}
-        \includegraphics[width=\textwidth]{Objects/weighted_read_mort_adjusted_graph.pdf}
-        \label{fig:weighted_read_mort_graph}
-    \end{figure}
+    %\begin{figure}[ht!]
+    %\centering
+        %\caption{Outcomes Over Time}
+        %\includegraphics[width=\textwidth]{Objects/weighted_read_mort_adjusted_graph.pdf}
+        %\label{fig:weighted_read_mort_graph}
+    %\end{figure}
 
@@
 
-     \begin{figure}
-     \caption{Readmission Rate Synthetic Difference in Differences Results}
-     \centering
-          \begin{subfigure}[b]{0.45\textwidth}
-         \centering
-         \caption{For-Profit and All NFP}
-         \includegraphics[width=\textwidth]{Objects/read_fp_nfp_synth_graph.pdf}
-         \label{fig:read_synth_plota}
-     \end{subfigure}%
-     \hfill
-     \begin{subfigure}[b]{0.45\textwidth}
-         \centering
-         \caption{For-Profit and NFP w/ MD Exec}
-         \includegraphics[width=\textwidth]{Objects/read_fp_md_synth_graph.pdf}
-         \label{fig:read_synth_plotb}
-     \end{subfigure}%
-     \vspace{5mm}
-     \hfill
-     \begin{subfigure}[b]{0.45\textwidth}
-         \centering
-         \caption{For-Profit and NFP w/out MD Exec}
-         \includegraphics[width=\textwidth]{Objects/read_fp_nomd_synth_graph.pdf}
-         \label{fig:read_synth_plotc}
-     \end{subfigure}
-     \hfill
-     \begin{subfigure}[b]{0.45\textwidth}
-         \centering
-         \caption{NFP w/ and w/out MD Exec}
-         \includegraphics[width=\textwidth]{Objects/read_md_nomd_synth_graph.pdf}
-         \label{fig:read_synth_plotd}
-     \end{subfigure}
-        \label{fig:read_synth_plot}
-    \end{figure}
+    %  \begin{figure}
+    %  \caption{Readmission Rate Synthetic Difference in Differences Results}
+    %  \centering
+    %       \begin{subfigure}[b]{0.45\textwidth}
+    %      \centering
+    %      \caption{For-Profit and All NFP}
+    %      \includegraphics[width=\textwidth]{Objects/read_fp_nfp_synth_graph.pdf}
+    %      \label{fig:read_synth_plota}
+    %  \end{subfigure}%
+    %  \hfill
+    %  \begin{subfigure}[b]{0.45\textwidth}
+    %      \centering
+    %      \caption{For-Profit and NFP w/ MD Exec}
+    %      \includegraphics[width=\textwidth]{Objects/read_fp_md_synth_graph.pdf}
+    %      \label{fig:read_synth_plotb}
+    %  \end{subfigure}%
+    %  \vspace{5mm}
+    %  \hfill
+    %  \begin{subfigure}[b]{0.45\textwidth}
+    %      \centering
+    %      \caption{For-Profit and NFP w/out MD Exec}
+    %      \includegraphics[width=\textwidth]{Objects/read_fp_nomd_synth_graph.pdf}
+    %      \label{fig:read_synth_plotc}
+    %  \end{subfigure}
+    %  \hfill
+    %  \begin{subfigure}[b]{0.45\textwidth}
+    %      \centering
+    %      \caption{NFP w/ and w/out MD Exec}
+    %      \includegraphics[width=\textwidth]{Objects/read_md_nomd_synth_graph.pdf}
+    %      \label{fig:read_synth_plotd}
+    %  \end{subfigure}
+    %     \label{fig:read_synth_plot}
+    % \end{figure}
 
@@
 
-     \begin{figure}
-     \caption{Mortality Rate Synthetic Difference in Differences Results}
-     \centering
-          \begin{subfigure}[b]{0.45\textwidth}
-         \centering
-         \caption{For-Profit and All NFP}
-         \includegraphics[width=\textwidth]{Objects/mort_fp_nfp_synth_graph.pdf}
-         \label{fig:mort_synth_plota}
-     \end{subfigure}%
-     \hfill
-     \begin{subfigure}[b]{0.45\textwidth}
-         \centering
-         \caption{For-Profit and NFP w/ MD Exec}
-         \includegraphics[width=\textwidth]{Objects/mort_fp_md_synth_graph.pdf}
-         \label{fig:mort_synth_plotb}
-     \end{subfigure}%
-     \hfill
-     \begin{subfigure}[b]{0.45\textwidth}
-         \centering
-         \caption{For-Profit and NFP w/out MD Exec}
-         \includegraphics[width=\textwidth]{Objects/mort_fp_nomd_synth_graph.pdf}
-         \label{fig:mort_synth_plotc}
-     \end{subfigure}
-     \hfill
-     \begin{subfigure}[b]{0.45\textwidth}
-         \centering
-         \caption{NFP w/ and w/out MD Exec}
-         \includegraphics[width=\textwidth]{Objects/mort_md_nomd_synth_graph.pdf}
-         \label{fig:mort_synth_plotd}
-     \end{subfigure}
-        \label{fig:mort_synth_plot}
-    \end{figure}
+    %  \begin{figure}
+    %  \caption{Mortality Rate Synthetic Difference in Differences Results}
+    %  \centering
+    %       \begin{subfigure}[b]{0.45\textwidth}
+    %      \centering
+    %      \caption{For-Profit and All NFP}
+    %      \includegraphics[width=\textwidth]{Objects/mort_fp_nfp_synth_graph.pdf}
+    %      \label{fig:mort_synth_plota}
+    %  \end{subfigure}%
+    %  \hfill
+    %  \begin{subfigure}[b]{0.45\textwidth}
+    %      \centering
+    %      \caption{For-Profit and NFP w/ MD Exec}
+    %      \includegraphics[width=\textwidth]{Objects/mort_fp_md_synth_graph.pdf}
+    %      \label{fig:mort_synth_plotb}
+    %  \end{subfigure}%
+    %  \hfill
+    %  \begin{subfigure}[b]{0.45\textwidth}
+    %      \centering
+    %      \caption{For-Profit and NFP w/out MD Exec}
+    %      \includegraphics[width=\textwidth]{Objects/mort_fp_nomd_synth_graph.pdf}
+    %      \label{fig:mort_synth_plotc}
+    %  \end{subfigure}
+    %  \hfill
+    %  \begin{subfigure}[b]{0.45\textwidth}
+    %      \centering
+    %      \caption{NFP w/ and w/out MD Exec}
+    %      \includegraphics[width=\textwidth]{Objects/mort_md_nomd_synth_graph.pdf}
+    %      \label{fig:mort_synth_plotd}
+    %  \end{subfigure}
+    %     \label{fig:mort_synth_plot}
+    % \end{figure}
 
@@
 
-     \import{Tables}{change_analysis.tex}
+     % \import{Tables}{change_analysis.tex}
 
@@
 
-\begin{figure}[ht!]
-    \centering
-    \caption{Executive Team Changes Over Time}
-    \includegraphics[width=\textwidth]{Objects/change_graph.pdf}
-    \label{fig:change_graph}
-\end{figure}
+% \begin{figure}[ht!]
+%     \centering
+%     \caption{Executive Team Changes Over Time}
+%     \includegraphics[width=\textwidth]{Objects/change_graph.pdf}
+%     \label{fig:change_graph}
+% \end{figure}
 
@@
 
-\import{Tables}{forprofit_readmort_fullsample.tex}
+%\import{Tables}{forprofit_readmort_fullsample.tex}
 
@@
 
-\import{Tables}{MD_noMD_readmort_fullsample.tex}
+%\import{Tables}{MD_noMD_readmort_fullsample.tex}
 
@@
 
-\begin{figure}[ht!]
-     \caption{TWFE Event Study Results}
-     \centering
-          \begin{subfigure}[b]{0.45\textwidth}
-         \centering
-         \caption{}
-         \includegraphics[width=\textwidth]{Objects/read_forprofit_es_graph.pdf}
-         \label{fig:es_plota}
-     \end{subfigure}%
-     \hfill
-     \begin{subfigure}[b]{0.45\textwidth}
-         \centering
-         \caption{}
-         \includegraphics[width=\textwidth]{Objects/read_MD_es_graph.pdf}
-         \label{fig:es_plotb}
-     \end{subfigure}%
-     \vspace{5mm}
-     \hfill
-     \begin{subfigure}[b]{0.45\textwidth}
-         \centering
-         \caption{}
-         \includegraphics[width=\textwidth]{Objects/mort_forprofit_es_graph.pdf}
-         \label{fig:es_plotc}
-     \end{subfigure}
-     \hfill
-     \begin{subfigure}[b]{0.45\textwidth}
-         \centering
-         \caption{}
-         \includegraphics[width=\textwidth]{Objects/mort_MD_es_graph.pdf}
-         \label{fig:es_plotd}
-     \end{subfigure}
-        \label{fig:es_plot}
-    \end{figure}
+% \begin{figure}[ht!]
+%      \caption{TWFE Event Study Results}
+%      \centering
+%           \begin{subfigure}[b]{0.45\textwidth}
+%          \centering
+%          \caption{}
+%          \includegraphics[width=\textwidth]{Objects/read_forprofit_es_graph.pdf}
+%          \label{fig:es_plota}
+%      \end{subfigure}%
+%      \hfill
+%      \begin{subfigure}[b]{0.45\textwidth}
+%          \centering
+%          \caption{}
+%          \includegraphics[width=\textwidth]{Objects/read_MD_es_graph.pdf}
+%          \label{fig:es_plotb}
+%      \end{subfigure}%
+%      \vspace{5mm}
+%      \hfill
+%      \begin{subfigure}[b]{0.45\textwidth}
+%          \centering
+%          \caption{}
+%          \includegraphics[width=\textwidth]{Objects/mort_forprofit_es_graph.pdf}
+%          \label{fig:es_plotc}
+%      \end{subfigure}
+%      \hfill
+%      \begin{subfigure}[b]{0.45\textwidth}
+%          \centering
+%          \caption{}
+%          \includegraphics[width=\textwidth]{Objects/mort_MD_es_graph.pdf}
+%          \label{fig:es_plotd}
+%      \end{subfigure}
+%         \label{fig:es_plot}
+%     \end{figure}
 
@@
 
-\import{Tables}{forprofit_readmort_synth.tex}
+%\import{Tables}{forprofit_readmort_synth.tex}
 
-\import{Tables}{MD_noMD_readmort_synth.tex}
+%\import{Tables}{MD_noMD_readmort_synth.tex}
 

### 2024-10-25 Paper/paperdraft_Oct2024.tex

--- 2024-10-25 Paper/paperdraft_Sep2024.tex
@@
     In corporate environments, executives are considered crucial drivers of a company's practices and performance, as they are heavily involved in the culture, visions, and strategies of the firm.
-Executive heterogeneity has been linked to various actions and outcomes of firms; characteristics such as gender, age, and background are correlated with financial performance and behaviors such as employee diversity, wages, aggressiveness in mergers and acquisitions, and other corporate strategies.\footnote{\cite{bertrand2003managing}; \cite{matsa2011chipping}; \cite{matsa2011chipping}; \cite{adams2009women}; \cite{ahern2012changing}; \cite{flabbi2019female}; \cite{benmelech2015military}; \cite{custodio2013generalists}; \cite{frydman2019rising}} However, there are still settings and relevant characteristics for which the relationship between management and firm behavior is not well understood.
+Executive heterogeneity has been linked to various actions and outcomes of firms; characteristics such as gender, age, and background are correlated with financial performance and behaviors such as employee diversity, wages, aggressiveness in mergers and acquisitions, and other corporate strategies.\footnote{\cite{bertrand2003managing}; \cite{matsa2011chipping}; \cite{matsa2011chipping}; \cite{adams2009women}; \cite{ahern2012changing}; \cite{flabbi2019female}; \cite{benmelech2015military}; \cite{custodio2013generalists}; \cite{frydman2019rising}} However, one aspect of firm leadership that is not well understood is how different leaders may interact with changing policy incentives differently.
 
     In this paper, I study whether type of executive training affects firm behavior in the context of US hospitals.
-In the US, health care makes up almost 20\% of spending (\cite{CMS_National_Health_Expenditure}), and a major goal of policymakers is to improve the quality of health care without increasing costs.
+In the US, health care makes up almost 20\% of spending (\cite{CMSNationalHealthExpenditure}), and a major goal of policymakers is to improve the quality of health care without increasing costs.
 Yet, little is known about how hospital managers affect health care value, and it is unclear whether existing findings on executive characteristics can be extrapolated to the health care, and particularly hospital, context.
-Existing literature is generally limited to large, publicly traded firms, whereas a majority of US hospitals are nonprofit with presumably different objectives.
-In particular, health care executives have a particularly unique trade-off between profit and patient welfare, which varies from the types of firms studied in existing literature.
+Existing literature is generally limited to large, publicly traded firms, whereas a majority of US hospitals are nonprofit with presumably different objectives and policy incentives.
 
@@
 However, hiring physician executives is controversial, as physicians are not always trained in business management, which can lead to poor practices and lower quality (\cite{HarvardBusinessReview2018}).
-Thus, I estimate how having clinically trained executives in hospital leadership affects hospital outcomes.
+Thus, I estimate how clinically trained executives in hospital leadership affects hospital outcomes.
 I scrape executive names and titles of nonprofit hospitals in the US from publicly available Tax Form 990s.
@@
 
+    \section{Theoretical Framework}
+
+    Consider a simplified model of hospital objectives, in which hospitals choose a level of quality of care that maximizes an objective function capturing the trade-off between profit and patient welfare.
+That is,
+
+    $$\max_{\theta}\hspace{2mm}\alpha\pi(\theta) + (1-\alpha) u(\theta),$$
+
+    \noindent where $\pi(\theta)$ is a profit function depending on quality $\theta$, $u(.)$ is extra utility gained from quality, which is increasing and concave in $\theta$, and $\alpha\in[0,1]$ captures the weight placed on profit vs.
+societal benefit.
+There is an implicit stay-open condition throughout the model.
+
+    In this paper, I consider hospital choice of quality under pay-for-performance incentives, which essentially incorporate quality directly into a hospital's profit function (\cite{dranove2011health}).
+Alternatively, a fee-for-service system only incentivizes quantity.
+A hospital that places all value on profit will only care about quality when it enters into the profit function, and therefore will respond more drastically to pay-for-performance than a hospital who already cared about quality through societal benefit before the incentives.
+I will now formalize this intuition mathematically.
+For simplicity, I abstract away from quality having a direct impact on quantity of patients or price.\footnote{As long as the relationship between quality and quantity/prove are not dependent on clinical experience on the leadership team, this does not change the conclusion of the model.} Taking the first order condition yields:
+
+    $$(1-\alpha)u'(\theta) = -\alpha \pi'(\theta),$$
+
+    \noindent where marginal benefit equals marginal cost of increasing quality. Solving for $u'(\theta)$ and differentiating with respect to $\alpha$,
+
+    $$\frac{du'(\theta)}{d\alpha} = \frac{-1}{(1-\alpha)^2}\pi'(\theta).$$
+
+    Now, consider incentives when the hospital is in a strictly fee-for-service environment.
+That is, quality benefits patients, but hospitals do not directly derive profit from having a high quality.
+Then, assuming there is a positive cost to increasing quality, $\pi(\theta)$ is decreasing in $\theta$.
+Then $\frac{du'(\theta)}{d\alpha} > 0$.
+Using the Implicit Function, I can write $\frac{d\theta}{d\alpha}$ as
+    $\frac{du'(\theta)/d\alpha}{du'(\theta)/d\theta}<0$, revealing that in a fee-for-service setting, the more weight the hospital places on profit, the lower quality of care chosen.
+
+    Next, assume the hospital is in a strictly pay-for-performance setting, where quality benefits patients and directly increases profit.
+Then, assuming the marginal financial benefit of increasing quality is greater than the marginal financial cost, $\pi(\theta)$ is increasing in $\theta$.
+Then $\frac{du'(\theta)}{d\alpha} < 0$, revealing that in a pay-for-performance setting, the more weight placed on profit, the higher quality of care chosen by the hospital.
+I combine the results found in each scenario into one response function that depends on $\alpha$, where $\theta_{p4p}$ is quality under pay-for-performance incentives and $\theta_{ffs}$ is quality under fee-for-service only.
+That is,
+
+    \begin{align*}
+        \frac{d\Delta\theta}{d\alpha}&=\frac{d(\theta_{p4p}-\theta_{ffs})}{d\alpha}\\
+        &=\frac{d\theta_{p4p}}{d\alpha}-\frac{d\theta_{ffs}}{d\alpha}\\
+        &> 0.
+    \end{align*}
+
+    Hence, this model predicts that change in quality depends on how much weight the hospital places on profit vs.
+societal benefit.
+Particularly, hospitals with more weight on profit respond more than hospitals with more weight placed on societal benefit.
+Therefore, for leadership characteristics that are plausibly correlated with how much weight is placed on patient welfare versus quality, such as clinical training, we expect to see a differential response to pay-for-performance incentives being enacted.
+
     \section{Setting and Data}
@@
     \section{Are Clinical Executives Less Profit Driven?}\label{sec:forprofit}
-
-    One hypothesis that could explain the differential response due to clinical leadership is how profit-driven a clinician is compared to other training backgrounds.
-To demonstrate this, consider hospital behavior under the enactment of pay-for-performance policies, where hospitals face a trade-off between profit and societal/patient benefit.
-Pay-for-performance incentives essentially incorporate quality directly into a hospital's profit function (\cite{dranove2011health}), whereas a fee-for-service system only incentivizes quantity.
-A hospital that places all value on profit will only care about quality when it enters into the profit function, and therefore will respond more drastically to pay-for-performance than a hospital who already cared about quality through societal benefit before the incentives.
-
-    To formalize this intuition, I specify the hospital objective function as a weighted average of profit and societal (or patient) outcomes and derive characteristics of the optimal quality.
-For simplicity, I abstract away from quality having a direct effect of quantity and price.
-Without any financial incentives to increase quality, hospital quality is decided by:
-
-    $$\max_{\theta}\hspace{2mm}\alpha\pi(\theta) + (1-\alpha) u(\theta),$$
-
-    \noindent where $\pi(\theta)$ is a profit function depending on $\theta$, and the relationship between profit and $\theta$ depends on pay-for-performance incentives.
-Additionally, $u(.)$ is extra utility gained from quality, which is increasing and concave in $\theta$, and $\alpha\in[0,1]$ captures the weight placed on profit vs.
-societal benefit.
-There is an implicit stay-open condition throughout the model.
-Taking the first order condition yields
-
-    $$(1-\alpha)u'(\theta) = -\alpha \pi'(\theta),$$
-
-    \noindent where marginal benefit equals marginal cost of increasing quality. Solving for $u'(\theta)$ and differentiating with respect to $\alpha$:
-
-    $$\frac{du'(\theta)}{d\alpha} = \frac{-1}{(1-\alpha)^2}\pi'(\theta)$$
-
-    First, assume the hospital is in a strictly fee-for-service setting, where quality benefits patients, but does not directly provide profit to the hospital through reimbursements.
-Then, assuming there is a positive cost to increasing quality, $\pi(\theta)$ is decreasing in $\theta$.
-Then $\frac{du'(\theta)}{d\alpha} > 0$, revealing that in a fee-for-service setting, the more weight placed on profit, the lower quality of care chosen by the hospital.
-
-    Second, assume the hospital is in a strictly pay-for-performance setting, where quality benefits patients and directly increases profit.
-Then, assuming the marginal financial benefit of increasing quality is greater than the marginal financial cost, $\pi(\theta)$ is increasing in $\theta$.
-Then $\frac{du'(\theta)}{d\alpha} < 0$, revealing that in a pay-for-performance setting, the more weight placed on profit, the higher quality of care chosen by the hospital.
-I combine the results found in each scenario into one response function that depends on $\alpha$, where $\theta_{p4p}$ is quality under pay-for-performance incentives and $\theta_{ffs}$ is quality under fee-for-service only.
-That is,
-
-    \begin{align*}
-        \frac{d\Delta\theta}{d\alpha}&=\frac{d(\theta_{p4p}-\theta_{ffs})}{d\alpha}\\
-        &=\frac{d\theta_{p4p}}{d\alpha}-\frac{d\theta_{ffs}}{d\alpha}\\
-        &> 0.
-    \end{align*}
-
-    Hence, this model predicts that change in quality depends on how much weight the hospital places on profit vs.
-societal benefit.
-Particularly, hospitals with more weight on profit respond more than hospitals with more weight placed on societal benefit.
-In reality, $\alpha$ is unknown, except in the case of for-profits, who have revealed that they have a high $\alpha$ by their ownership status.
-Clinically trained executives could also affect this $\alpha$, as doctors commonly take a hippocratic oath to put patients first.
-
-    \subsection{Estimation}
 

### 2024-11-06 Slides/JobTalk.tex
Updates from Overleaf

@@
     \begin{enumerate}
-        \item Limited to large, publicly traded firms: reason to believe this does not extend to nonprofit setting
-        \item Direct responses to policy changes are not well-understood
+        \item Nonprofit firms: reason to believe these findings might not inform leadership in other settings
+        \item Direct responses to policy changes
     \end{enumerate}
@@
 
-\textbf{Hospital care makes up a third of health care spending} \scriptsize (\cite{cms2024nhe}) \normalsize
+\begin{enumerate}
+    \item \textbf{Hospital care makes up a third of health care spending} \scriptsize (\cite{cms2024nhe}) \normalsize
+
+\vspace{2mm}
+
     \begin{itemize}
-    \item A majority of hospitals are nonprofit and their objectives are not as clear as classic for-profit firms
+        \item Nonprofits drive the hospital industry \scriptsize (\cite{ASPE_2023}) \normalsize
     \end{itemize}
@@
 
-\textbf{Policies often aim to improve the value of health care delivery}
+    \item \textbf{Policies often aim to improve the value of health care delivery}
 
-\vspace{3mm}
+\vspace{2mm}
 
@@
         \item It is unclear how hospital leaders view value
-        \item Or whether they might play a role in achieving this goal
+        \item Or whether they might play a role in achieving better value
     \end{wideitemize}
+
+    \end{enumerate}
 \end{frame}
@@
     \large
-    Does having executives with clinical training affect response to pay-for-performance incentives of nonprofit hospitals?
+    Does having executives with clinical training affect response to pay-for-performance incentives of nonprofit hospitals in the US?
 \end{frame}
@@
 
-\textbf{Physician executives could be good or bad}
+\textbf{Do physicians make good firm leaders?}
 
@@
 \begin{wideitemize}
-    \item Non-clinical reduce readmissions by blank more than clinical
+    \item Non-clinical reduce readmission by 2\% more than clinical
     \item Non-clinical executive teams behave like profit-driven hospitals
@@
 \begin{frame}{Rest of Talk}
-    Theoretical Motivation
-
-    \vspace{4mm}
 
@@
     About Executive Teams
+
+    \vspace{4mm}
+
+    Theoretical Motivation
 
@@
     Signaling vs. Managing
+\end{frame}
+
+\section{Setting and Data Collection}
+
+\begin{transitionframe}
+\centering \LARGE
+    \textcolor{black}{Setting and Data Collection}
+\end{transitionframe}
+
+\begin{frame}{Nonprofit Hospital Executives}\label{nonprofithospexec}
+\textbf{What do executives do?}
+
+\vspace{1mm}
+
+\begin{wideitemize}
+    \item Board of directors selects executives
+    \item Executives handle the day-to-day operations of the hospital \hyperlink{executivejob}{\beamergotobutton{executive job}} 
+\end{wideitemize}
+
+\vspace{7mm}
+
+\textbf{Executive teams}
+
+\vspace{1mm}
+
+\begin{wideitemize}
+    \item Lots of variation in what executive teams look like
+    \item Typically CEO and CFO, sometimes CMO 
+    \item I focus on the team as a whole instead of specific executive roles
+\end{wideitemize}
+
+\end{frame}
+
+\begin{frame}{}
+    \begin{tikzpicture}
+        % First image in the top-left
+        \node[anchor=north west] at (0, 4) {\includegraphics[width=0.7\textwidth]{Graphics/990_snip_frontpage.PNG}};
+        % Second image in the middle
+        \node[anchor=north west] (img2) at (2.5, 1.5) {\includegraphics[width=0.75\textwidth]{Graphics/Screenshot 2024-10-03 160258.png}};
+        \draw[line width=1mm, eggplant] (img2.north west) rectangle (img2.south east);
+    \end{tikzpicture}
+\end{frame}
+
+\begin{frame}{Data Collection Process}
+    \begin{enumerate}
+        \item Get list of Employee Identification Numbers (EIN) that represent nonprofit hospitals
+
+        \vspace{5mm}
+        \item Download tax forms from 2009-2015 from Nonprofit Explorer API
+
+        \vspace{5mm}
+        \item Use Optimal Character Recognition to extract text from all of these documents
+
+        \vspace{5mm}
+        \item Write string cleaning algorithms to extract the correct name, title, position of each entry
+    \end{enumerate}
+\end{frame}
+
+\begin{frame}{Merge to Other Data}
+    \textbf{Hospital:} link EIN and ID number from the American Hospital Association (AHA) Survey based on name and location
+
+    \vspace{1mm}
+    \begin{itemize}
+        \item \underline{AHA} data for hospital demographics
+        \item \underline{Hospital Compare} data for readmission and mortality rates
+        \item Center for Medicare and Medicaid Services (CMS) \underline{Case Mix Index file} for patient complexity
+        \item \underline{Health Care Cost Report Information System (HCRIS)} for information on pay-for-performance rewards or penalties
+    \end{itemize}
+
+    \vspace{7mm}
+
+    \textbf{Physician:} I use executives that self-identify as MD in the tax form data
+
+    \vspace{1mm}
+
+    \begin{itemize}
+        \item Manually match physician executives to \underline{NPPES} NPI number
+    \end{itemize}
+\end{frame}
+
+\begin{frame}{About Pay-for-Performance Policies}
+    Policies enacted in 2012 that either penalized for low quality or rewarded for high quality
+
+    \vspace{3mm}
+
+    \textbf{Hospital Readmissions Reduction Program}
+    \begin{itemize}
+        \item If hospitals have above average readmission rates they get 1-3\% lower reimbursement rates
+        \item Pneumonia, Heart Attack (AMI), Heart Failure
+    \end{itemize}
+
+    \vspace{3mm}
+
+    \textbf{Hospital Value-Based Purchasing Program}
+    \begin{itemize}
+        \item safety, efficiency, cost reductions, clinical outcomes, and community engagement, improvement combined into single score metric
+        \item ``good" hospitals rewarded
+    \end{itemize}
+
+    \begin{block}{\normalsize Takeaway}
+        \normalsize Most hospitals had incentive to improve quality to try to avoid meaningful penalties, which were not perfectly predictable
+    \end{block}
+\end{frame}
+
+\section{About Executive Teams}
+
+\begin{transitionframe}
+\centering \LARGE
+    \textcolor{black}{Descriptive Statistics}
+\end{transitionframe}
+
+\begin{frame}{Executive Teams}
+    \begin{table}[ht!]
+        \centering
+        \begin{tabular}[t]{lccc}
+        \toprule
+         & Ever & Always & Never\\
+         & Clinical Exec & Clinical Exec &  Clinical Exec \\
+        \midrule
+        Number of Executives & 4.85 & 5.25 & 3.11\\
+        \addlinespace[0.2em]
+        Fraction Clinical Execs & 0.20 & 0.26 & 0.0\\
+        \addlinespace[0.2em]
+        Has Clinical CEO & 0.25 & 0.11 & 0.0\\
+        \addlinespace[0.2em]
+        Has CMO & 0.53 & 0.45 & 0.15\\
+        \addlinespace[0.6em]
+        Number of Hospitals & 391 & 56 & 445\\
+        \bottomrule
+        \end{tabular}
+    \end{table}
+\end{frame}
+
+\begin{frame}{Doctors on Executive Teams (N=813)}
+    \begin{table}[ht!]
+    \centering
+    \begin{tabular}[t]{p{6cm} c}
+    \toprule
+     & Mean\\
+    \midrule
+    Age & 52.5\\
+    Female & 0.11\\
+    CEO & 0.16\\
+    CMO & 0.31\\
+    \addlinespace[0.3em]
+    \multicolumn{2}{l}{\textbf{Specialty}}\\
+    \hspace{1em}Internal Medicine & 0.32\\
+    \hspace{1em}Family Medicine & 0.17\\
+    \hspace{1em}Emergency Medicine & 0.11\\
+    \hspace{1em}Surgery & 0.09\\
+    \hspace{1em}Anesthesiologist & 0.04\\
+    \hspace{1em}Other & 0.28\\
+    \bottomrule
+    \end{tabular}
+    \end{table}
+\end{frame}
+
+\begin{frame}{Average Team Changes Over Time}
+\centering 
+    \includegraphics[width=.8\textwidth]{Objects/change_means.pdf}
+\end{frame}
+
+\begin{frame}{Changes are not correlated with policy exposure}
+\centering \small
+\begin{equation*}\label{eq:change2}
+    \text{change}_{ht} = \sum_{j=2011}^{2014}\beta_j(\mathbf{1}\{t=j\}\times \text{Program Exposed})_{ht} + \alpha_h + \epsilon_{ht},
+    \end{equation*}
+
+    \btVFill
+
+    \includegraphics[width=.7\textwidth]{Objects/change_analysis_plot.pdf}
+
+\end{frame}
+
+\begin{frame}{Hospital Statistics}
+\vspace{-2mm}
+    \begin{table}[ht!]
+    \centering
+    \begin{tabular}[t]{lccc}
+     & Ever & Always & Never\\
+      & Clinical Exec & Clinical Exec & Clinical Exec\\
+    \midrule
+    \addlinespace[0.3em]
+    \multicolumn{4}{l}{\textbf{Hospital Characteristics}}\\
+    \hspace{1em}Academic Med. Center & 0.41 & 0.52 & 0.24\\
+    \hspace{1em}Number Beds & 231 & 222 & 116\\
+    \hspace{1em}Physician Owned & 0.01 & 0.02 & 0.02\\
+    \hspace{1em}System Affiliated & 0.58 & 0.52 & 0.44\\
+    \addlinespace[0.3em]
+    \multicolumn{4}{l}{\textbf{Penalty/Payment Variables}}\\
+    \hspace{1em}Ever Received HVBP Incentive & 0.79 & 0.75 & 0.62\\
+    \hspace{1em}Penalized for One Condition & 0.18 & 0.20 & 0.13\\
+    \hspace{1em}Penalized for AMI + HF/Pneum. & 0.09 & 0.09 & 0.06\\
+    \hspace{1em}Penalized for HF + Pneumonia & 0.41 & 0.39 & 0.34\\
+    \hspace{1em}Penalized for All Conditions & 0.30 & 0.27 & 0.22\\
+    Num. Hospitals & 391 & 56 & 445\\
+    \bottomrule
+    \end{tabular}
+    \end{table}
+\end{frame}
+
+\begin{frame}{Raw Outcomes, controlling for baseline case mix index}
+\centering
+    \includegraphics[width=.9\textwidth]{Objects/outcomes_graph_presentation.pdf}
 \end{frame}
@@
 
-\section{Setting and Data Collection}
+\section{Effect of Clinical Leadership on Response}
 
@@
 \centering \LARGE
-    \textcolor{black}{Setting and Data Collection}
+    \textcolor{black}{Effect of Clinical Leadership on Response to P4P}
 \end{transitionframe}
 
-\begin{frame}{Nonprofit Hospital Executives}\label{nonprofithospexec}
-\textbf{What do executives do?}
+\begin{frame}{Estimation}\label{est_equation}
+    \textbf{Compare always clinical to never clinical hospital executive teams:}
+
+    \vspace{3mm}\large
+    \begin{equation*}
+    y_{ht} = \beta \text{ (clinical training x post 2012)}_{ht} + \gamma_{h} + \delta_t + \epsilon_{ht}
+    \end{equation*}
+
+    \vspace{5mm}
+    \normalsize
     \begin{wideitemize}
-    \item Board of directors selects executives
-    \item Executives handle the day-to-day operations of the hospital \hyperlink{executivejob}{\beamergotobutton{executive job}} 
+        \item $y_{ht}$: weighted average readmission/ mortality rates
+        \item Leverage the \textit{timing} of the programs
+    \end{wideitemize}
+
+    \vspace{5mm}
+
+    Different methods:
+    \begin{itemize}
+        \item No weights
+        \item \textbf{``Synthetic DiD" type weights} \hyperlink{synthdid_equation}{\beamergotobutton{weighted}}
+        \item Coarsened exact matching \hyperlink{match_equation}{\beamergotobutton{CEM}}
+    \end{itemize}
+\end{frame}
+
+\begin{frame}{Identification}
+\begin{wideitemize}
+    \item Trends would have continued in a parallel manner
+    \item No other 2012 event correlated with the composition of leadership team and the outcomes
+    \item No other hospital characteristics that are perfectly correlated with clinical leadership are driving the results
+    \item No anticipation of the incentive changes
+\end{wideitemize}
+\end{frame}
+
+\begin{frame}{Readmission Rates}\label{main:readrates}
+    \begin{columns}
+    \column{0.5\textwidth}
+        \centering
+        \includegraphics[width=.9\textwidth]{Objects/read_md_nomd_synth_graph.pdf}
+    \column{0.5\textwidth}
+        \begin{wideitemize}
+            \item Non-clinical teams reduce readmissions by .3 ppts more than clinical teams 
+            \item Small relative to average readmission rate of 21
+            \item Large difference relative to the average effect of these programs (1ppt)
         \end{wideitemize}
@@
 
-\textbf{Executive teams}
-\begin{wideitemize}
-    \item Lots of variation in what executive teams look like
-    \item Typically CEO and CFO, sometimes CMO 
-\end{wideitemize}
+        \hyperlink{int_margin}{\beamergotobutton{intensive margin}}
 
-\vspace{7mm}
-
-Usable data on executives is rare
-\end{frame}
-
-\begin{frame}{}
-    \begin{tikzpicture}
-        % First image in the top-left
-        \node[anchor=north west] at (0, 4) {\includegraphics[width=0.7\textwidth]{Graphics/990_snip_frontpage.PNG}};
-        % Second image in the middle
-        \node[anchor=north west] (img2) at (2.5, 1.5) {\includegraphics[width=0.75\textwidth]{Graphics/Screenshot 2024-10-03 160258.png}};
-        \draw[line width=1mm, eggplant] (img2.north west) rectangle (img2.south east);
-    \end{tikzpicture}
-\end{frame}
-
-\begin{frame}{Data Collection Process}
-    \begin{enumerate}
-        \item Get list of Employee Identification Numbers (EIN) that represent nonprofit hospitals
-
-        \vspace{5mm}
-        \item Download tax forms from 2009-2015 from Nonprofit Explorer API
-
-        \vspace{5mm}
-        \item Use Optimal Character Recognition to extract text from all of these documents
-
-        \vspace{5mm}
-        \item Write string cleaning algorithms to extract the correct name, title, position of each entry
-    \end{enumerate}
-\end{frame}
-
-\begin{frame}{Merge to Other Hospital Data}
-    I create links between EIN and ID number from the American Hospital Association Survey (AHA) based on hospital name and location
-
-    \vspace{7mm}
-    Merge to:
-
-    \vspace{2mm}
-    \begin{itemize}
-        \item AHA data for hospital demographics
-        \item Hospital Compare data for readmission and mortality rates
-        \item Center for Medicare and Medicaid Services (CMS) Case Mix Index file for patient complexity
-        \item Health Care Cost Report Information System (HCRIS) for information on pay-for-performance rewards or penalties
-    \end{itemize}
-\end{frame}
-
-\begin{frame}{Merge to Physician Data}
-    I use executives that self-identify as MD
-
-    \vspace{8mm}
-
-    To verify, I match all executive names to database of all registered physicians and match NPI by hand
-\end{frame}
-
-\begin{frame}{About Pay-for-Performance Policies}
-    Policies enacted in 2012 that either penalized for low quality or rewarded for high quality
-
-    \vspace{3mm}
-
-    \textbf{Hospital Readmissions Reduction Program}
-    \begin{itemize}
-        \item If hospitals have above average readmission rates they get 1-3\% lower reimbursement rates
-        \item Pneumonia, Heart Attack (AMI), Heart Failure
-    \end{itemize}
-
-    \vspace{3mm}
-
-    \textbf{Hospital Value-Based Purchasing Program}
-    \begin{itemize}
-        \item safety, efficiency, cost reductions, clinical outcomes, and community engagement, improvement combined into single score metric
-        \item ``good" hospitals rewarded
-    \end{itemize}
-
-    \pause
-
-    \begin{block}{\normalsize Takeaway}
-        \normalsize All hospitals in the middle of the distribution had incentive to improve quality to try to avoid penalties, which were not perfectly predictable
-    \end{block}
-\end{frame}
-
-\section{About Executive Teams}
-
-\begin{transitionframe}
-\centering \LARGE
-    \textcolor{black}{Descriptive Statistics}
-\end{transitionframe}
-
-\begin{frame}{Executive Teams}
-    \begin{table}[ht!]
-        \centering
-        \begin{tabular}[t]{lccc}
-        \toprule
-         & Ever & Always & Never\\
-         & Clinical Exec & Clinical Exec &  Clinical Exec \\
-        \midrule
-        Number Executives & 4.85 & 5.25 & 3.11\\
-        Frac Clinical Execs & 0.20 & 0.26 & 0\\
-        Frac Int. Medicine Execs & 0.06 & 0.10 & 0.0\\
-        Has a CMO & 0.53 & 0.45 & 0.15\\
-        Has Clinical CEO & 0.25 & 0.11 & 0.0\\
-        \addlinespace[0.6em]
-        Num. Hospitals & 391 & 56 & 445\\
-        \bottomrule
-        \end{tabular}
-    \end{table}
-\end{frame}
-
-\begin{frame}{Doctors on Executive Teams (N=813)}
-    \begin{table}[ht!]
-    \centering
-    \begin{tabular}[t]{p{6cm} c}
-    \toprule
-     & Mean\\
-    \midrule
-    Age & 52.5\\
-    Female & 0.11\\
-    CEO & 0.16\\
-    CMO & 0.31\\
-    \addlinespace[0.3em]
-    \multicolumn{2}{l}{\textbf{Specialty}}\\
-    \hspace{1em}Internal Medicine & 0.32\\
-    \hspace{1em}Family Medicine & 0.17\\
-    \hspace{1em}Emergency Medicine & 0.11\\
-    \hspace{1em}Surgery & 0.09\\
-    \hspace{1em}Anesthesiologist & 0.04\\
-    \hspace{1em}Other & 0.28\\
-    \bottomrule
-    \end{tabular}
-    \end{table}
-\end{frame}
-
-\begin{frame}{Average Team Changes Over Time}
-\centering 
-    \includegraphics[width=.8\textwidth]{Objects/change_means.pdf}
-\end{frame}
-
-\begin{frame}{Changes are not correlated with policy exposure}
-\centering \small
-\begin{equation*}\label{eq:change2}
-    \text{change}_{ht} = \sum_{j=2011}^{2014}\beta_j(\mathbf{1}\{t=j\}\times \text{Program Exposed})_{ht} + \alpha_h + \epsilon_{ht},
-    \end{equation*}
-
-    \btVFill
-
-    \includegraphics[width=.6\textwidth]{Objects/change_analysis_plot.pdf}
-
-\end{frame}
-
-\begin{frame}{Hospital Statistics}
-\vspace{-2mm}
-    \begin{table}[ht!]
-    \centering
-    \begin{tabular}[t]{lccc}
-     & Ever & Always & Never\\
-      & Clinical Exec & Clinical Exec & Clinical Exec\\
-    \midrule
-    \addlinespace[0.3em]
-    \multicolumn{4}{l}{\textbf{Hospital Characteristics}}\\
-    \hspace{1em}Academic Med. Center & 0.41 & 0.52 & 0.24\\
-    \hspace{1em}Number Beds & 231 & 222 & 116\\
-    \hspace{1em}Physician Owned & 0.01 & 0.02 & 0.02\\
-    \hspace{1em}System Affiliated & 0.58 & 0.52 & 0.44\\
-    \addlinespace[0.3em]
-    \multicolumn{4}{l}{\textbf{Penalty/Payment Variables}}\\
-    \hspace{1em}Ever Received HVBP Incentive & 0.79 & 0.75 & 0.62\\
-    \hspace{1em}Penalized for One Condition & 0.18 & 0.20 & 0.13\\
-    \hspace{1em}Penalized for AMI + HF/Pneum. & 0.09 & 0.09 & 0.06\\
-    \hspace{1em}Penalized for HF + Pneumonia & 0.41 & 0.39 & 0.34\\
-    \hspace{1em}Penalized for All Conditions & 0.30 & 0.27 & 0.22\\
-    Num. Hospitals & 391 & 56 & 445\\
-    \bottomrule
-    \end{tabular}
-    \end{table}
-\end{frame}
-
-\begin{frame}{}
-    put graph of outcomes here
-\end{frame}
-
-\section{Effect of Clinical Leadership on Response}
-
-\begin{transitionframe}
-\centering \LARGE
-    \textcolor{black}{Effect of Clinical Leadership on Response}
-\end{transitionframe}
-
-\begin{frame}{Estimation}\label{est_equation}
-
-    \vspace{5mm}\large
-    \begin{equation*}
-    y_{ht} = \beta \text{ (clinical training x post 2012)}_{ht} + \gamma_{h} + \delta_t + \epsilon_{ht}
-    \end{equation*}
-    \begin{itemize}
-        \item Compare always to never clinical executives
-        \item Leverage the timing of the programs
-    \end{itemize}
-
-    \vspace{8mm}
-
-    Identifies the effect of having a clinically trained executive on readmission and mortality rates after pay-for-performance
-
-    \vspace{2mm}
-
-    \begin{itemize}
-        \item Assume that hospitals would have had parallel trends absent clinical executives
-    \end{itemize}
-
-    \btVFill
-    \normalsize
-    \hyperlink{synthdid_equation}{\beamergotobutton{weighted}} \hyperlink{match_equation}{\beamergotobutton{matching}}
-\end{frame}
-
-\begin{frame}{Readmission Rates}\label{main:readrates}
-\centering
-    \includegraphics[width=.45\textwidth]{Objects/read_md_nomd_synth_graph.pdf}
-
-    \btVFill
-
-    \hyperlink{int_margin}{\beamergotobutton{intensive margin}}
+    \end{columns}
 \end{frame}
@@
 \begin{frame}{Mortality Rates}
-
+    \begin{columns}
+    \column{0.5\textwidth}
+        \centering
+        \includegraphics[width=.9\textwidth]{Objects/mort_md_nomd_synth_graph.pdf}
+    \column{0.5\textwidth}
+        \begin{wideitemize}
+            \item No statistical difference in mortality between clinical and non-clinical teams
+        \end{wideitemize}
+    \end{columns}
 \end{frame}
@@
 
-    \vspace{5mm}\large
+    \textbf{Compare both nonprofit leadership team types to the response of for-profits:}
+
+    \vspace{3mm}\large
     \begin{equation*}
@@
     \end{equation*}
+
+    \vspace{5mm}
+    \normalsize
+
+    Comparison group:
     \begin{itemize}
-        \item Two comparison groups: clinical and non-clinical executive teams
+        \item nonprofit clinical; nonprofit non-clinical
     \end{itemize}
+
+    \vspace{5mm}
+
+    Synthetic diff-in-diff type weights
 
@@
 
-\begin{frame}{Results}
+\begin{frame}{Readmission Rate}
+    \begin{columns}
+    \column{0.5\textwidth}
+        \centering
+        \includegraphics[width=.9\textwidth]{Objects/fp_read_md_synth_graph.pdf}
+    \column{0.5\textwidth}
+        \centering
+        \includegraphics[width=.9\textwidth]{Objects/fp_read_nomd_synth_graph.pdf}
+    \end{columns}
+\end{frame}
 
+\begin{frame}{Mortality Rate}
+    \begin{columns}
+    \column{0.5\textwidth}
+        \centering
+        \includegraphics[width=.9\textwidth]{Objects/fp_mort_md_synth_graph.pdf}
+    \column{0.5\textwidth}
+        \centering
+        \includegraphics[width=.9\textwidth]{Objects/fp_mort_nomd_synth_graph.pdf}
+    \end{columns}
 \end{frame}
@@
 
+\begin{frame}{Hypotheses}
+    \textbf{Clinical leaders represent underlying $\alpha$}
+
+\vspace{2mm}
+
+    \begin{wideitemize}
+        \item Timing of the clinical leader doesn't matter \item Clinical leader represents underlying objectives of the hospital
+    \end{wideitemize}
+
+    \vspace{10mm}
+
+    \textbf{Clinical leaders effectively change $\alpha$}
+
+\vspace{2mm}
+
+    \begin{wideitemize}
+        \item Timing of the clinical leader does matter
+        \item Clinical leaders manage the hospital differently
+    \end{wideitemize}
+\end{frame}
+
 \begin{frame}{Estimation}
+    \begin{figure}[ht!]
+\begin{center}
+
+ \begin{tabular}{| m{17em} |}
+ \hline
+ Signaling:\\ [0.5ex]
+ \hline\hline 
+ \vspace{2mm}
+ Treat Group:  \hspace{15mm} Never MD \\
+ \vspace{2mm}
+ Comparison Group: \hspace{3mm} Ever MD  \\
+ [1ex]
+ \hline
+ \end{tabular}
+\hfil   %<---
+ \begin{tabular}{|m{19em}|}
+ \hline
+ Managing:\\ [0.5ex]
+ \hline\hline
+ \vspace{2mm}
+ Treat Group: \hspace{11mm} MD in 2012 \\
+ \vspace{2mm}
+ Comparison Group:  Ever MD (not in 2012)  \\
+ [1ex]
+ \hline
+ \end{tabular}
+
+\end{center}
+ \end{figure}
 
@@
 \begin{frame}{Results}
-
+    \begin{table}[ht!]
+\centering
+\begin{tabular}[t]{lcccc}
+\toprule
+\multicolumn{1}{c}{ } & \multicolumn{2}{c}{\underline{Weighted Avg. Readmission Rate}} & \multicolumn{2}{c}{\underline{Weighted Avg. Mortality Rate}} \\
+ & (1) & (2) & (3) & (4)\\
+\midrule
+Signaling Effect & -0.01 &  & 0.03 & \\
+ & (0.07) &  & (0.06) & \\
+Managing Effect &  & 0.17 &  & 0.12\\
+ &  & (0.1) &  & (0.09)\\
+ &  &  &  & \\
+Observations & 3,005 & 1,595 & 3,005 & 1,595\\
+\bottomrule
+\end{tabular}
+\end{table}
 \end{frame}
@@
 \begin{frame}{Conclusion}
+    \textbf{We know very little about whether leaders matter in health care}
 
-\end{frame}
+    \vspace{2mm}
 
-\begin{frame}{Future Work}
+    \begin{wideitemize}
+        \item In this paper, I show that non-clinically trained hospital executives internalize financial incentives on quality more than clinically trained executives
+        \item Nonprofit objectives are not all created equal, and financial incentives might not be the only way to improve value in health care
+    \end{wideitemize}
 
+    \vspace{5mm}
+
+    \textbf{Future Work}
+
+    \vspace{2mm}
+
+    \begin{itemize}
+        \item Effects of mandated gender and racial diversity on hospital behavior
+        \item Networks of executives over time 
+        \item Are leaders correlated with hospital fraudulent behavior?
+    \end{itemize}
 \end{frame}
@@
 \centering \Large
+    Thank you!
+
+    \vspace{3mm}
+
     \textcolor{black}{hkagele@emory.edu}

### 2024-11-13 Slides/JobTalk.tex

@@
 \usepackage{import}
+\usepackage{fontawesome5}
 
@@
 \setbeamercolor{block title}{bg=eggplant!70, fg=white}
-\setbeamercolor{block body}{bg=eggplant!20}
+\setbeamercolor{block body}{bg=eggplant!30}
 
@@
 \setbeamercolor{enumerate subitem}{fg=black}
-\setbeamercolor{button}{bg=eggplant!20,fg=black}
+\setbeamercolor{button}{bg=eggplant!80,fg=white}
 
@@
     \LARGE Does Hospital Leadership Matter?  \\[5pt]
-    \normalsize Evidence from Pay-for-Performance
+    \large Evidence from Pay-for-Performance
 };
 % Author 
-\node[below=1cm] (author) at (title.south){Hanna Glenn, Emory University};
+\node[below=1cm] (author) at (title.south){\large Hanna Glenn, Emory University};
 % Date
@@
 
-\begin{frame}{Firm Leadership}
+\begin{frame}{Leadership Matters}
 \centering
@@
         % First image (top-left)
-        \node[anchor=north west] (img1) at (0, 3) {\includegraphics[width=0.55\textwidth]{Graphics/theroleleadershiphasincompanyculture.png}};
+        \node[anchor=north west] (img1) at (0, 3) {\includegraphics[width=0.75\textwidth]{Graphics/theroleleadershiphasincompanyculture.png}};
         \draw[line width=.45mm, lightcyan] (img1.north west) rectangle (img1.south east);
@@
         % Second image (slightly down and right of the first image)
-        \node[anchor=north west] (img2) at (0.28, 0.58) {\includegraphics[width=0.5\textwidth]{Graphics/linkingbusinessstrategyandleadershipforbetteroutcomes.png}};
+        \node[anchor=north west] (img2) at (0, -.25) {\includegraphics[width=0.75\textwidth]{Graphics/linkingbusinessstrategyandleadershipforbetteroutcomes.png}};
         \draw[line width=.45mm, lightcyan] (img2.north west) rectangle (img2.south east);
@@
         % Fourth image (bottom-right)
-        \node[anchor=north west] (img4) at (0, -2) {\includegraphics[width=0.55\textwidth]{Graphics/theroleofleadershipinorganizationalsuccess.png}};
-        \draw[line width=.45mm, lightcyan] (img4.north west) rectangle (img4.south east);
+        %\node[anchor=north west] (img4) at (0, -2) {\includegraphics[width=0.55\textwidth]{Graphics/theroleofleadershipinorganizationalsuccess.png}};
+        %\draw[line width=.45mm, lightcyan] (img4.north west) rectangle (img4.south east);
 
@@
 
-\begin{frame}{Leadership Characteristics}
-
-\vspace{1mm}
+\begin{frame}{Leadership Characteristics and Firm Behavior/Performance}
+
+\large
 
@@
 \begin{itemize}
-    \item \small More female board members $\implies$ lower firm value \scriptsize(\cite{ahern2012changing}) \small
-    \item \small More female board members $\implies$ less employment reduction \scriptsize (\cite{matsa2013female}) \small
-    \item female executives affect the female wage distribution \scriptsize (\cite{flabbi2019female}) \small
+    \item \normalsize Female board members $\implies$ lower firm value, less employment reduction \\ \scriptsize(\cite{ahern2012changing}; \cite{matsa2013female}) \normalsize
+    \item Female executives $\implies$ female wage distribution \scriptsize (\cite{flabbi2019female}) \normalsize
 \end{itemize}
 
-\vspace{2mm}
-
-\textbf{Background}
+\vspace{4mm}
+
+\textbf{Prior Experience}
 \begin{itemize}
-    \item \small Firms pay more for general background CEOs \scriptsize (\cite{custodio2013generalists}) \small
-    \item \small CEOs w/ prior experience do better in M\&A \scriptsize(\cite{custodio2013ceos}) \small
-    \item \small Military experience CEOs are less aggressive in R\&D and investing \scriptsize (\cite{benmelech2015military})
+    \item \normalsize General background CEOs paid more \scriptsize (\cite{custodio2013generalists}) \normalsize
+    \item Background matters for acquisitions \scriptsize(\cite{custodio2013ceos}) \normalsize
+
+\vspace{1mm}
+
+    \item Military experience CEOs \scriptsize (\cite{benmelech2015military})
 \end{itemize}
 
-\vspace{2mm}
+\vspace{4mm}
 
@@
 \begin{itemize}
-    \item \small Young male CEOs more aggressive in M\&A \scriptsize (\cite{levi2010deal})
+    \item \normalsize Young male CEOs more aggressive in M\&A \scriptsize (\cite{levi2010deal})
 \end{itemize}
 
-\btVFill\pause
-
-\begin{block}{\normalsize Gaps in Literature}
-    \small 
+\end{frame}
+
+\begin{frame}{Gaps in Literature}
+    \large
     \begin{enumerate}
-        \item Nonprofit firms: reason to believe these findings might not inform leadership in other settings
-        \item Direct responses to policy changes
+        \item Findings are limited to large, publicly traded firms
+
+\vspace{2mm}
+
+        \begin{itemize}
+            \item \normalsize Different objectives/incentives in other types of firms (nonprofits)
+        \end{itemize}
+
+        \vspace{10mm}
+
+        \item Unclear whether leadership matters when policies change incentives
     \end{enumerate}
+\end{frame}
+
+\begin{frame}{Addressing the Gaps}
+
+    \Large
+    \begin{block}{}
+    \begin{center}
+    I examine the role of leadership in nonprofit hospitals
+
+    responding to policy incentives
+    \end{center}
     \end{block}
-\end{frame}
-
-\begin{frame}{Hospitals \& Health Care Policy Goals}
-
-\begin{block}{This Paper}
-    Examine the role of leadership in nonprofit hospitals responding to policy incentives 
-\end{block}
-
-\vspace{5mm}
+
+\vspace{5mm} \large \pause
 
@@
     \begin{itemize}
-        \item Nonprofits drive the hospital industry \scriptsize (\cite{ASPE_2023}) \normalsize
+        \item \normalsize Nonprofits drive the hospital industry \scriptsize (\cite{ASPE_2023}) \normalsize
     \end{itemize}
 
-\vspace{5mm}
-
-    \item \textbf{Policies often aim to improve the value of health care delivery}
-
-\vspace{2mm}
-
-    \begin{wideitemize}
-        \item It is unclear how hospital leaders view value
-        \item Or whether they might play a role in achieving better value
+\vspace{8mm}
+
+    \item \large \textbf{Policies often aim to improve the value of health care delivery}
+
+\vspace{2mm}
+
+    \begin{wideitemize}
+        \item \normalsize Unclear how leaders interact with this goal
     \end{wideitemize}
@@
     \centering
-    \large
-    Does having executives with clinical training affect response to pay-for-performance incentives of nonprofit hospitals in the US?
+    \Large
+    Do executives with clinical training
+
+    \vspace{1mm}
+
+    affect response to pay-for-performance incentives
+
+    \vspace{1mm}
+
+    of nonprofit hospitals in the US?
 \end{frame}
@@
 
+\large
+
 \textbf{Do physicians make good firm leaders?}
@@
 \begin{wideitemize}
-    \item Physicians can bring a unique combination of clinical and administration expertise that can potentially lead to better quality \scriptsize (\cite{Stajduhar_2023}, \cite{Ahmed_2022}) \normalsize
-
-    \vspace{2mm}
-
-    \item Not always trained in business management, which can lead to poor practices and lower quality \scriptsize (\cite{HarvardBusinessReview2018})
-\end{wideitemize}
-
-\vspace{10mm}
-
-\textbf{Clinical training is potentially correlated with how the leader values quality}
-\end{frame}
-
-\begin{frame}{Preview of Findings}
-
-\textbf{Non-clinical executive teams respond more drastically to financial incentives on quality compared to clinical executive teams}
-
-\vspace{2mm}
-
-\begin{wideitemize}
-    \item Non-clinical reduce readmission by 2\% more than clinical
+    \item Relevant expertise \scriptsize (\cite{Stajduhar_2023}, \cite{Ahmed_2022}) \large
+
+    \item Lack of business training \scriptsize (\cite{HarvardBusinessReview2018})
+\end{wideitemize}
+
+\vspace{12mm}
+
+\textbf{Potentially correlated with how the leader values quality versus profit}
+\end{frame}
+
+\begin{frame}{Pay-for-Performance Incentives}
+    \large
+    \begin{block}{\Large Main Takeaway:}
+        \large 
+        \centering
+        \vspace{1mm}
+        Hospitals now have financial incentives to improve quality
+
+        \vspace{1mm}
+    \end{block}
+
+    \vspace{6mm}
+
+    \textbf{Quality: 30-day readmissions \& 30-day mortality}
+
+    \vspace{5mm}
+
+    Hospital Readmissions Reduction Program
+    \begin{itemize}
+        \item Penalized high readmissions
+        \item Pneumonia, Heart Attack (AMI), Heart Failure
+    \end{itemize}
+
+    \vspace{5mm}
+
+    Hospital Value-Based Purchasing Program
+    \begin{itemize}
+        \item Score metric determines rewards
+        \item Rewards high metric
+    \end{itemize}
+\end{frame}
+
+\begin{frame}{Preview of Main Findings}
+
+\large
+
+\textbf{Non-clinical executive teams respond more to financial incentives on quality}
+
+\vspace{3mm}
+
+\begin{wideitemize}
+    \item Through readmission rates
     \item Non-clinical executive teams behave like profit-driven hospitals
@@
 \begin{frame}{Contribution}
+\large
     \textbf{Hospital Executives/Board of Directors}
-    \begin{itemize}
-        \item Physician board members $\implies$ lower public donations in nonprofit hospitals \scriptsize (\cite{brickley2010board}) \normalsize
-        \item No correlation between CEOs and hospital production in England \scriptsize (\cite{janke2019impact})\normalsize
-        \item Public hospitals in Chile improved when older physicians were removed as CEO \scriptsize (\cite{otero2022managers}) \normalsize
-    \end{itemize}
-
-    \vspace{5mm}
+
+\vspace{2mm}
+
+    \begin{wideitemize}
+        \item \normalsize Physician board members $\implies$ lower public donations \scriptsize (\cite{brickley2010board}) \normalsize
+        \item No correlation b/w CEOs and hospital productivity in England \scriptsize (\cite{janke2019impact})\normalsize
+        \item Public hospitals in Chile improved when older physicians left CEO \scriptsize (\cite{otero2022managers}) \normalsize
+    \end{wideitemize}
+
+    \vspace{5mm}\pause
 
     \textbf{Hospital Response to Pay-for-Performance}
-    \begin{itemize}
-        \item Hospitals decreased readmissions by 40\% on average, partially due to selective patient practices \scriptsize (\cite{gupta2021impacts}) \normalsize
+
+\vspace{2mm}
+
+    \begin{wideitemize}
+        \item \normalsize Hospitals decreased readmissions \scriptsize (\cite{gupta2021impacts}) \normalsize
         \item Market concentration and systems affect response \scriptsize (\cite{kunz2024assessing}) \normalsize
-        \item Reward program had little impact on quality \scriptsize (\cite{us2015hospital}; \cite{norton2018moneyball}; \cite{friedson2019so})
-    \end{itemize}
+        \item Rewards had little impact on quality \scriptsize (\cite{norton2018moneyball}; \cite{friedson2019so})
+    \end{wideitemize}
 \end{frame}
@@
 
+\large
+
     Setting and Data Collection
 
-    \vspace{4mm}
+    \vspace{5mm}
 
@@
 
-    \vspace{4mm}
+    \vspace{5mm}
 
@@
 
-\begin{frame}{Nonprofit Hospital Executives}\label{nonprofithospexec}
-\textbf{What do executives do?}
-
-\vspace{1mm}
+\begin{frame}{About the Executives}\label{nonprofithospexec}
+
+\large
+\textbf{What do they do?}
+
+\vspace{2mm}
 
@@
     \item Board of directors selects executives
-    \item Executives handle the day-to-day operations of the hospital \hyperlink{executivejob}{\beamergotobutton{executive job}} 
-\end{wideitemize}
-
-\vspace{7mm}
+    \item Executives handle the day-to-day operations of the hospital 
+    \item Quality often improved from the top down \scriptsize (\cite{silow2011reducing})
+\end{wideitemize}
+
+\vspace{4mm} \pause
 
@@
 
-\vspace{1mm}
-
-\begin{wideitemize}
-    \item Lots of variation in what executive teams look like
-    \item Typically CEO and CFO, sometimes CMO 
+\vspace{2mm}
+
+\begin{wideitemize}
+    \item Lot of variation in team organization
+    \item CEO and CFO
     \item I focus on the team as a whole instead of specific executive roles
 \end{wideitemize}
+
+\btVFill
+
+\hyperlink{executivejob}{\beamergotobutton{executive job}}
 
@@
 \begin{frame}{Data Collection Process}
-    \begin{enumerate}
-        \item Get list of Employee Identification Numbers (EIN) that represent nonprofit hospitals
-
-        \vspace{5mm}
-        \item Download tax forms from 2009-2015 from Nonprofit Explorer API
-
-        \vspace{5mm}
-        \item Use Optimal Character Recognition to extract text from all of these documents
-
-        \vspace{5mm}
-        \item Write string cleaning algorithms to extract the correct name, title, position of each entry
-    \end{enumerate}
+\begin{center}
+\begin{tikzpicture}[
+    node distance=3.1cm and 1.8cm,
+    every node/.style={align=center, font=\normalsize},
+    process/.style={rectangle, draw, rounded corners, text centered, minimum width=2.5cm, minimum height=2cm, fill=purple!30},
+    output/.style={rectangle, draw, text centered, minimum width=2.7cm, minimum height=2.5cm, fill=purple},
+    extract/.style={rectangle, draw, rounded corners, text centered, minimum width=2.5cm, minimum height=2cm, fill=sage!30},
+    arrow/.style={thick, ->, >=stealth}
+]
+
+% Nodes
+\node (step1) [process] {Hospital \\EINs};\pause
+\node (step2) [process, right of=step1] {Download \\PDFs};
+\draw[arrow] (step1) -- (step2); \pause
+\node (step3) [process, right of=step2] {Extract\\text};
+\draw[arrow] (step2) -- (step3); 
+% New green box below step3
+\node (extract) [extract, below of=step3, yshift=0.5cm] {Optical\\Character\\Recognition}; 
+\draw[arrow] (extract) -- (step3);\pause
+\node (step4) [process, right of=step3] {String\\cleaning\\algorithms};
+\draw[arrow] (step3) -- (step4); \pause
+\node (output) [output, right of=step4, text=white] {Executive\\Name\\Position\\Title};
+\draw[arrow] (step4) -- (output);
+\end{tikzpicture}
+\end{center}
 \end{frame}
@@
 \begin{frame}{Merge to Other Data}
-    \textbf{Hospital:} link EIN and ID number from the American Hospital Association (AHA) Survey based on name and location
-
-    \vspace{1mm}
+
+\large
+    \textbf{Hospital:}
+
+    \vspace{2mm}
+    \begin{wideitemize}
+        \item AHA
+        \item Hospital Compare
+        \item CMS Case Mix Index
+        \item Health Care Cost Report Information System (HCRIS)
+    \end{wideitemize}
+
+    \vspace{7mm}
+
+    \textbf{Physician:}
+
+    \vspace{2mm}
+
     \begin{itemize}
-        \item \underline{AHA} data for hospital demographics
-        \item \underline{Hospital Compare} data for readmission and mortality rates
-        \item Center for Medicare and Medicaid Services (CMS) \underline{Case Mix Index file} for patient complexity
-        \item \underline{Health Care Cost Report Information System (HCRIS)} for information on pay-for-performance rewards or penalties
+        \item Manually match physician executives to NPPES NPI number
     \end{itemize}
-
-    \vspace{7mm}
-
-    \textbf{Physician:} I use executives that self-identify as MD in the tax form data
-
-    \vspace{1mm}
-
-    \begin{itemize}
-        \item Manually match physician executives to \underline{NPPES} NPI number
-    \end{itemize}
-\end{frame}
-
-\begin{frame}{About Pay-for-Performance Policies}
-    Policies enacted in 2012 that either penalized for low quality or rewarded for high quality
-
-    \vspace{3mm}
-
-    \textbf{Hospital Readmissions Reduction Program}
-    \begin{itemize}
-        \item If hospitals have above average readmission rates they get 1-3\% lower reimbursement rates
-        \item Pneumonia, Heart Attack (AMI), Heart Failure
-    \end{itemize}
-
-    \vspace{3mm}
-
-    \textbf{Hospital Value-Based Purchasing Program}
-    \begin{itemize}
-        \item safety, efficiency, cost reductions, clinical outcomes, and community engagement, improvement combined into single score metric
-        \item ``good" hospitals rewarded
-    \end{itemize}
-
-    \begin{block}{\normalsize Takeaway}
-        \normalsize Most hospitals had incentive to improve quality to try to avoid meaningful penalties, which were not perfectly predictable
-    \end{block}
 \end{frame}
@@
  & Ever & Always & Never\\
-         & Clinical Exec & Clinical Exec &  Clinical Exec \\
+  & Clinical Exec & Clinical Exec & Clinical Exec\\
 \midrule
-        Number of Executives & 4.85 & 5.25 & 3.11\\
-        \addlinespace[0.2em]
-        Fraction Clinical Execs & 0.20 & 0.26 & 0.0\\
-        \addlinespace[0.2em]
-        Has Clinical CEO & 0.25 & 0.11 & 0.0\\
-        \addlinespace[0.2em]
-        Has CMO & 0.53 & 0.45 & 0.15\\
+\addlinespace[0.3em]
+Number Executives & 4.82 & 5.19 & 2.86\\
+Fraction Clinical Execs & 0.20 & 0.26 & 0.00\\
+Fraction Int. Medicine Execs & 0.06 & 0.08 & 0.00\\
+Has a CMO & 0.55 & 0.52 & 0.07\\
+Has Clinical CEO & 0.25 & 0.10 & 0.00\\
 \addlinespace[0.6em]
-        Number of Hospitals & 391 & 56 & 445\\
+Num. Hospitals & 452 & 62 & 385\\
 \bottomrule
@@
 
-\begin{frame}{Doctors on Executive Teams (N=813)}
+\begin{frame}{Doctors on Executive Teams (N=941)}
 \begin{table}[ht!]
 \centering
-    \begin{tabular}[t]{p{6cm} c}
+\begin{tabular}[t]{lc}
 \toprule
@@
 \midrule
-    Age & 52.5\\
+Age & 52.56\\
 Female & 0.11\\
-    CEO & 0.16\\
-    CMO & 0.31\\
+CEO & 0.15\\
+CMO & 0.36\\
+CQO & 0.09\\
 \addlinespace[0.3em]
@@
 \hspace{1em}Internal Medicine & 0.32\\
-    \hspace{1em}Family Medicine & 0.17\\
-    \hspace{1em}Emergency Medicine & 0.11\\
-    \hspace{1em}Surgery & 0.09\\
-    \hspace{1em}Anesthesiologist & 0.04\\
-    \hspace{1em}Other & 0.28\\
+\hspace{1em}Family Medicine & 0.18\\
+\hspace{1em}Surgery & 0.14\\
+\hspace{1em}Emergency Medicine \hspace{4mm} & 0.11\\
+\hspace{1em}Pediatrics & 0.06\\
+\hspace{1em}Other & 0.16\\
 \bottomrule
@@
 \end{table}
+\end{frame}
+
+\begin{frame}{Hospital Statistics}\label{sumstats}
+\vspace{-3mm}
+\begin{table}[ht!]
+\centering
+\begin{tabular}[t]{lccc}
+ & Ever & Always & Never\\
+  & Clinical Exec &Clinical Exec &Clinical Exec\\
+\midrule
+\addlinespace[0.3em]
+\multicolumn{4}{l}{\textbf{Hospital Characteristics}}\\
+\hspace{1em}Academic Med. Center & 0.41 & 0.51 & 0.20\\
+\hspace{1em}Number Beds & 227 & 218 & 102\\
+\hspace{1em}Physician Owned & 0.01 & 0.02 & 0.02\\
+\hspace{1em}System Affiliated & 0.57 & 0.56 & 0.42\\
+\addlinespace[0.3em]
+\multicolumn{4}{l}{\textbf{Penalty/Payment Variables}}\\
+\hspace{1em}Ever Received HVBP Incentive & 0.79 & 0.77 & 0.59\\
+\hspace{1em}Penalized for one condition & 0.19 & 0.25 & 0.13\\
+\hspace{1em}Penalized for AMI + HF & 0.07 & 0.05 & 0.03\\
+\hspace{1em}Penalized for AMI + Pneumonia & 0.02 & 0.02 & 0.02\\
+\hspace{1em}Penalized for HF + Pneumonia & 0.41 & 0.40 & 0.33\\
+\hspace{1em}Penalized for All Conditions & 0.31 & 0.31 & 0.20\\
+\bottomrule
+\end{tabular}
+\end{table}
+\hyperlink{forprofitstats}{\beamergotobutton{for-profit stats}}
+\end{frame}
+
+\begin{frame}{Raw Outcomes, controlling for baseline case mix index}\label{outcomes}
+\begin{center}
+    \includegraphics[width=.92\textwidth]{Objects/outcomes_graph_presentation.pdf}
+\end{center}
+    \btVFill
+    \hyperlink{fp_outcomes}{\beamergotobutton{with for-profit}}
+\end{frame}
+
+\section{Theoretical Framework}
+
+\begin{transitionframe}
+\centering \LARGE
+    \textcolor{black}{Theoretical Framework}
+\end{transitionframe}
+
+\begin{frame}{Hospital Objectives: Profit versus Patient Welfare}
+    \Large
+    \centering
+    \begin{equation*}
+        \max_{\theta}\hspace{2mm} \alpha\pi(\theta) + (1-\alpha)u(\theta)
+    \end{equation*}
+\end{frame}
+
+\begin{frame}[noframenumbering]{Hospital Objectives: Profit versus Patient Welfare} \Large
+    \begin{equation*}
+\max_{\theta}\hspace{2mm}\alpha
+\tikz[baseline]{
+    \node[draw=sage,rounded corners,anchor=base] (m1)
+    {$\pi(\theta)$};
+    \node[above of=m1, text=sage, yshift=5mm] (l1) {profit};
+    \draw[-,sage] (l1) -- (m1);
+}
++ (1-\alpha)\hspace{-11mm}
+\tikz[baseline]{
+    \node[draw=sage,rounded corners,anchor=base] (m2)
+    {$u(\theta)$};
+    \node[above of=m2, text=sage, yshift=5mm] (l2) {patient welfare};
+    \draw[-,sage] (l2) -- (m2);
+}
+\end{equation*}
+\end{frame}
+
+\begin{frame}[noframenumbering]{Hospital Objectives: Profit versus Patient Welfare} \Large
+    \begin{equation*}
+\max_{\theta}\hspace{2mm}\alpha\pi(\hspace{-5mm}
+\tikz[baseline]{
+    \node[draw=purple,rounded corners,anchor=base] (m1)
+    {$\theta$};
+    \node[above of=m1, text=purple, yshift=5mm] (l1) {quality};
+    \draw[-,purple] (l1) -- (m1);
+}
+\hspace{-5mm}) + (1-\alpha)u(\hspace{-5mm}
+\tikz[baseline]{
+    \node[draw=purple,rounded corners,anchor=base] (m2)
+    {$\theta$};
+    \node[above of=m2, text=purple, yshift=5mm] (l2) {quality};
+    \draw[-,purple] (l2) -- (m2);
+}
+\hspace{-5mm})
+\end{equation*}
+\end{frame}
+
+\begin{frame}[noframenumbering]{Hospital Objectives: Profit versus Patient Welfare} \Large
+    \begin{equation*}
+\max_{\theta}\hspace{-4mm}
+\tikz[baseline]{
+    \node[draw=earthyellow,rounded corners,anchor=base] (m1)
+    {$\alpha$};
+    \node[above of=m1, text=earthyellow, yshift=5mm] (l1) {weight on\\profit};
+    \draw[-,earthyellow] (l1) -- (m1);
+}
+\hspace{-8mm}\pi(\theta) + \hspace{-7mm}
+\tikz[baseline]{
+    \node[draw=earthyellow,rounded corners,anchor=base] (m2)
+    {$(1-\alpha)$};
+    \node[above of=m2, text=earthyellow, yshift=5mm] (l2) {weight on\\patient welfare};
+    \draw[-,earthyellow] (l2) -- (m2);
+}
+\hspace{-7mm}u(\theta)
+\end{equation*}
+\end{frame}
+
+\begin{frame}{Choice of Quality}\label{choiceofquality}
+
+\vspace{9mm}
+
+\Large
+
+    \begin{columns}
+    \begin{column}{0.5\textwidth}
+    \begin{center}
+        \underline{Fee-for-Service}
+
+        \vspace{3mm}
+
+            $\pi$ $\downarrow$ in $\theta$
+    \end{center}
+    \end{column}
+    \begin{column}{0.5\textwidth}  %%<--- here
+    \begin{center}
+        \underline{Pay-for-Performance}
+
+        \vspace{3mm}
+
+            $\pi$ $\uparrow$ in $\theta$
+    \end{center}
+    \end{column}
+\end{columns}
+
+\btVFill
+
+\begin{block}{\Large Fee-for-service to Pay-for-performance}
+    \centering
+    \vspace{3mm}
+    Higher $\alpha$ $\implies$ larger response
+
+    \vspace{3mm}
+\end{block}
+
+\hyperlink{choiceofqualityapp}{\beamergotobutton{more details}}
+\end{frame}
+
+\begin{frame}{Informing the Empirical Analysis}
+
+\large
+    \begin{enumerate}
+        \item If clinicians affect $\alpha$, they affect quality change after P4P
+
+        \vspace{8mm}\pause
+        \item If non-clinicians have higher $\alpha$, they act more like for-profits
+
+        \vspace{8mm}\pause
+        \item Mechanism could be revealing objectives or changing objectives
+    \end{enumerate}
+\end{frame}
+
+\section{Effect of Clinical Leadership on Response}
+
+\begin{transitionframe}
+\centering \LARGE
+    \textcolor{black}{Effect of Clinical Leadership on Response to P4P}
+\end{transitionframe}
+
+\begin{frame}{Estimation}\label{est_equation}
+
+\large
+    \textbf{Compare always clinical to never clinical hospital executive teams:}
+
+    \vspace{3mm}\large
+    \begin{equation*}
+    y_{ht} = \beta \text{ (clinical training x post 2012)}_{ht} + \gamma_{h} + \delta_t + \epsilon_{ht}
+    \end{equation*}
+
+    \vspace{3mm}
+    \begin{wideitemize}
+        \item $y_{ht}$: weighted average readmission/ mortality rates
+        \item Leverage \textit{timing} of the programs
+    \end{wideitemize}
+
+    \vspace{8mm}\pause
+
+    Different methods:
+    \begin{itemize}
+        \item No weights
+        \item \textbf{``Synthetic DiD" type weights} 
+        \item Coarsened exact matching 
+    \end{itemize}
+\end{frame}
+
+\begin{frame}{Identification Assumptions}
+
+\large
+\begin{wideitemize}
+    \item Parallel trends
+    \item No other 2012 event correlated with executive team and outcomes
+    \item No hospital characteristics perfectly correlated with executive team
+    \item No anticipation of incentive changes
+    \item Stability in clinical executives not correlated with incentive change
+\end{wideitemize}
 \end{frame}
@@
 
-\begin{frame}{Hospital Statistics}
-\vspace{-2mm}
-    \begin{table}[ht!]
-    \centering
-    \begin{tabular}[t]{lccc}
-     & Ever & Always & Never\\
-      & Clinical Exec & Clinical Exec & Clinical Exec\\
-    \midrule
-    \addlinespace[0.3em]
-    \multicolumn{4}{l}{\textbf{Hospital Characteristics}}\\
-    \hspace{1em}Academic Med. Center & 0.41 & 0.52 & 0.24\\
-    \hspace{1em}Number Beds & 231 & 222 & 116\\
-    \hspace{1em}Physician Owned & 0.01 & 0.02 & 0.02\\
-    \hspace{1em}System Affiliated & 0.58 & 0.52 & 0.44\\
-    \addlinespace[0.3em]
-    \multicolumn{4}{l}{\textbf{Penalty/Payment Variables}}\\
-    \hspace{1em}Ever Received HVBP Incentive & 0.79 & 0.75 & 0.62\\
-    \hspace{1em}Penalized for One Condition & 0.18 & 0.20 & 0.13\\
-    \hspace{1em}Penalized for AMI + HF/Pneum. & 0.09 & 0.09 & 0.06\\
-    \hspace{1em}Penalized for HF + Pneumonia & 0.41 & 0.39 & 0.34\\
-    \hspace{1em}Penalized for All Conditions & 0.30 & 0.27 & 0.22\\
-    Num. Hospitals & 391 & 56 & 445\\
-    \bottomrule
-    \end{tabular}
-    \end{table}
-\end{frame}
-
-\begin{frame}{Raw Outcomes, controlling for baseline case mix index}
-\centering
-    \includegraphics[width=.9\textwidth]{Objects/outcomes_graph_presentation.pdf}
-\end{frame}
-
-\section{Theoretical Framework}
-
-\begin{transitionframe}
-\centering \LARGE
-    \textcolor{black}{Theoretical Framework}
-\end{transitionframe}
-
-\begin{frame}{Hospital Objectives}\label{hosp_obj}
-\large
-    \begin{equation*}
-        \max_{\theta}\hspace{2mm}\alpha\pi(\theta) + (1-\alpha) u(\theta)
-    \end{equation*}
-
-    \vspace{5mm} \normalsize
-
-    \begin{wideitemize}
-        \item Profit $\pi$ and patient welfare $u$; depend on quality $\theta$
-        \item Profit-driven hospitals have high $\alpha$
-        \item Abstract away from quality impacting demand for care, price \hyperlink{complex_model}{\beamergotobutton{more complex model}} 
-    \end{wideitemize}
-\end{frame}
-
-\begin{frame}{Choice of Quality}
-Take FOC, rearrange and take derivative wrt $\alpha$:
-
-\vspace{5mm}
-
-\begin{equation*}
-    \frac{du'(\theta)}{d\alpha} = \frac{-1}{(1-\alpha)^2}\pi'(\theta)
-\end{equation*}
-
-\vspace{5mm}
-
-$\pi'(\theta)$:
-
-\vspace{3mm}
-
-\begin{wideitemize}
-    \item In fee-for-service setting: negative $\implies \frac{du'(\theta)}{d\alpha} > 0$
-    \item In pay-for-performance setting: positive $\implies \frac{du'(\theta)}{d\alpha} < 0$
-\end{wideitemize}
-
-\end{frame}
-
-\begin{frame}{FFS $\implies$ P4P}
-    \large
-    \begin{align*}
-        \frac{d\Delta\theta}{d\alpha}&=\frac{d\theta_{p4p}}{d\alpha}-\frac{d\theta_{ffs}}{d\alpha}\\ 
-        &\\
-        &=\frac{\frac{du'(\theta)}{d\alpha}}{\frac{du'(\theta)}{d\theta_{p4p}}} - \frac{\frac{du'(\theta)}{d\alpha}}{\frac{du'(\theta)}{d\theta_{ffs}}}\\
-        &\\
-        &> 0
-    \end{align*}
-
-    \btVFill \normalsize
-
-    \begin{block}{Takeaway}
-        Hospitals with higher weight on profit will respond more to the financial incentives on quality
-    \end{block}
-\end{frame}
-
-\begin{frame}{Testable Predictions}
-    \begin{enumerate}
-        \item If non-clinical executives are more profit-driven than clinical executive hospitals, they will reduce quality by more
-
-        \vspace{5mm}
-        \item If non-clinical executives are more profit-driven, they will act more like for-profit hospitals
-
-        \vspace{5mm}
-        \item Executive teams could reflect an underlying $\alpha$ or they could effectively change $\alpha$
-    \end{enumerate}
-\end{frame}
-
-\section{Effect of Clinical Leadership on Response}
-
-\begin{transitionframe}
-\centering \LARGE
-    \textcolor{black}{Effect of Clinical Leadership on Response to P4P}
-\end{transitionframe}
-
-\begin{frame}{Estimation}\label{est_equation}
-    \textbf{Compare always clinical to never clinical hospital executive teams:}
-
-    \vspace{3mm}\large
-    \begin{equation*}
-    y_{ht} = \beta \text{ (clinical training x post 2012)}_{ht} + \gamma_{h} + \delta_t + \epsilon_{ht}
-    \end{equation*}
-
-    \vspace{5mm}
-    \normalsize
-    \begin{wideitemize}
-        \item $y_{ht}$: weighted average readmission/ mortality rates
-        \item Leverage the \textit{timing} of the programs
-    \end{wideitemize}
-
-    \vspace{5mm}
-
-    Different methods:
-    \begin{itemize}
-        \item No weights
-        \item \textbf{``Synthetic DiD" type weights} \hyperlink{synthdid_equation}{\beamergotobutton{weighted}}
-        \item Coarsened exact matching \hyperlink{match_equation}{\beamergotobutton{CEM}}
-    \end{itemize}
-\end{frame}
-
-\begin{frame}{Identification}
-\begin{wideitemize}
-    \item Trends would have continued in a parallel manner
-    \item No other 2012 event correlated with the composition of leadership team and the outcomes
-    \item No other hospital characteristics that are perfectly correlated with clinical leadership are driving the results
-    \item No anticipation of the incentive changes
-\end{wideitemize}
-\end{frame}
-
 \begin{frame}{Readmission Rates}\label{main:readrates}
+\large
     \begin{columns}
@@
         \begin{wideitemize}
-            \item Non-clinical teams reduce readmissions by .3 ppts more than clinical teams 
-            \item Small relative to average readmission rate of 21
-            \item Large difference relative to the average effect of these programs (1ppt)
-        \end{wideitemize}
-
-        \vspace{7mm}
+            \item Magnitude: 0.3ppt difference
+
+\vspace{5mm}
+
+            \item Large relative to effect of HRRP (1ppt)
+        \end{wideitemize}
+
+        \vspace{20mm}
 
         \hyperlink{int_margin}{\beamergotobutton{intensive margin}}
+        \hyperlink{cmi}{\beamergotobutton{due to patient selection?}}
+        \hyperlink{specifications}{\beamergotobutton{specifications}}
+        \hyperlink{specialty}{\beamergotobutton{specialty}}
 
@@
 \begin{frame}{Mortality Rates}
+\large
     \begin{columns}
@@
         \begin{wideitemize}
-            \item No statistical difference in mortality between clinical and non-clinical teams
+            \item Magnitude: 0.1ppt
+
+            \vspace{5mm}
+
+            \item Mortality is noisy measure 
         \end{wideitemize}
@@
 \begin{frame}{Estimation}
-
+\large
     \textbf{Compare both nonprofit leadership team types to the response of for-profits:}
 
-    \vspace{3mm}\large
+    \vspace{5mm}\large
     \begin{equation*}
@@
 
-    \vspace{5mm}
-    \normalsize
+    \vspace{10mm}
 
@@
 
-    \vspace{5mm}
-
-    Synthetic diff-in-diff type weights
-
 \end{frame}
@@
 \begin{frame}{Hypotheses}
-    \textbf{Clinical leaders represent underlying $\alpha$}
-
-\vspace{2mm}
-
-    \begin{wideitemize}
-        \item Timing of the clinical leader doesn't matter \item Clinical leader represents underlying objectives of the hospital
-    \end{wideitemize}
-
-    \vspace{10mm}
-
-    \textbf{Clinical leaders effectively change $\alpha$}
+
+\large
+    \textbf{Clinical leaders represent underlying objectives ($\alpha$)}
+
+\vspace{2mm}
+
+    \begin{wideitemize}
+        \item Timing of the clinical leader doesn't matter 
+    \end{wideitemize}
+
+    \vspace{15mm}
+
+    \textbf{Clinical leaders effectively change objectives ($\alpha$)}
 
@@
         \item Timing of the clinical leader does matter
-        \item Clinical leaders manage the hospital differently
     \end{wideitemize}
@@
 \begin{frame}{Estimation}
+\large
     \begin{figure}[ht!]
@@
 
- \begin{tabular}{| m{17em} |}
- \hline
- Signaling:\\ [0.5ex]
+ \begin{tabular}{| m{15em} |}
+ \hline \vspace{1mm}
+ \textbf{Signaling:}\\ [0.5ex]
  \hline\hline 
@@
 \hfil   %<---
- \begin{tabular}{|m{19em}|}
- \hline
- Managing:\\ [0.5ex]
+ \begin{tabular}{|m{18em}|}
+ \hline \vspace{1mm}
+ \textbf{Managing:}\\ [0.5ex]
  \hline\hline
@@
  \vspace{2mm}
- Comparison Group:  Ever MD (not in 2012)  \\
+ Comparison Group:  MD not in 2012  \\
  [1ex]
@@
 
-\begin{frame}{Conclusion}
+\begin{frame}{Summary}
+\large
     \textbf{We know very little about whether leaders matter in health care}
 
-    \vspace{2mm}
-
-    \begin{wideitemize}
-        \item In this paper, I show that non-clinically trained hospital executives internalize financial incentives on quality more than clinically trained executives
-        \item Nonprofit objectives are not all created equal, and financial incentives might not be the only way to improve value in health care
-    \end{wideitemize}
+    \begin{wideitemize}
+        \item This paper: non-clinical executive teams internalize financial incentives more
+    \end{wideitemize}
+
+    \vspace{5mm}
+
+    \textbf{Implications}
+    \begin{itemize}
+        \item Not all nonprofits have the same objectives
+        \item More clinicians in leadership could shift objectives towards quality
+    \end{itemize}
 
@@
     \textbf{Future Work}
-
-    \vspace{2mm}
 
@@
 \begin{transitionframe}
-\centering \Large
+\centering 
+%\includegraphics[width=.2\textwidth]{Graphics/HannaGlenn_headshot.jpg}
+
+\vspace{5mm}
+
+\Large
     Thank you!
 
-    \vspace{3mm}
-
-    \textcolor{black}{hkagele@emory.edu}
+    \vspace{15mm}
+
+    \textcolor{black}{\faEnvelope \hspace{2mm} hanna.glenn@emory.edu}
+
+    \vspace{2mm}
+
+    \textcolor{black}{\faTwitter \hspace{2mm} @hannalynnglenn}
 \end{transitionframe}
@@
 
-\begin{frame}{More Complex Model}\label{complex_model}
+\begin{frame}{Choice of Quality}\label{choiceofqualityapp}
+Take FOC, rearrange and take derivative wrt $\alpha$:
+
+\vspace{5mm}
+
+\begin{equation*}
+    \frac{du'(\theta)}{d\alpha} = \frac{-1}{(1-\alpha)^2}\pi'(\theta)
+\end{equation*}
+
+\vspace{5mm}
+
+$\pi'(\theta)$:
+
+\vspace{3mm}
+
+\begin{wideitemize}
+    \item In fee-for-service setting: negative $\implies \frac{du'(\theta)}{d\alpha} > 0$
+    \item In pay-for-performance setting: positive $\implies \frac{du'(\theta)}{d\alpha} < 0$
+\end{wideitemize}
+
+\end{frame}
+
+\begin{frame}{FFS $\implies$ P4P}
+
+\vspace{3mm{}}
+    \large
+
+        \begin{block}{}
+        \centering \Large
+        Hospitals with higher weight on profit will respond more
+
+        to the financial incentives on quality
+    \end{block}
+
+    \begin{align*}
+        \frac{d\Delta\theta}{d\alpha}&=\frac{d\theta_{p4p}}{d\alpha}-\frac{d\theta_{ffs}}{d\alpha}\\ 
+        &\\
+        &=\frac{\frac{du'(\theta)}{d\alpha}}{\frac{du'(\theta)}{d\theta_{p4p}}} - \frac{\frac{du'(\theta)}{d\alpha}}{\frac{du'(\theta)}{d\theta_{ffs}}}\\
+        &\\
+        &> 0
+    \end{align*}
+
+    \btVFill \normalsize
+    \hyperlink{choiceofquality}{\beamerreturnbutton{hospital choice of quality}}
+\end{frame}
+
+\begin{frame}{How Leaders Affect Quality}\label{executivejob}
+\large \vspace{5mm}
+\textbf{Ways to Improve Quality/ Decrease Readmissions}
+
+\vspace{2mm}
+
+\begin{wideitemize}
+    \item Improving care transition from inpatient to outpatient
+    \item Patient education
+    \item Improve care coordination
+    \item Follow-up appointments
+\end{wideitemize}
 
@@
 
-\hyperlink{hosp_obj}{\beamerreturnbutton{hospital objectives}}
-
-\end{frame}
-
-\begin{frame}{Executive Job}\label{executivejob}
-
+\scriptsize (\cite{silow2011reducing}) \hspace{4mm}
+\hyperlink{nonprofithospexec}{\beamerreturnbutton{Nonprofit Hosp Execs}}
+
+\end{frame}
+
+\begin{frame}{For-Profit and Nonprofit Statistics}\label{forprofitstats}
+\vspace{-3mm}
+    \begin{table}[ht!]
+\centering
+\begin{tabular}[t]{lcc}
+ & Nonprofit & For-Profit\\
+\midrule
+\addlinespace[0.3em]
+\multicolumn{3}{l}{\textbf{Hospital Characteristics}}\\
+\hspace{1em}Academic Med. Center & 0.27 & 0.17\\
+\hspace{1em}Number Beds & 151 & 157\\
+\hspace{1em}Physician Owned & 0.01 & 0.19\\
+\hspace{1em}System Affiliated & 0.59 & 0.89\\
+\addlinespace[0.3em]
+\multicolumn{3}{l}{\textbf{Penalty/Payment Variables}}\\
+\hspace{1em}Ever Received HVBP Incentive & 0.63 & 0.87\\
+\hspace{1em}Penalized for one condition & 0.14 & 0.15\\
+\hspace{1em}Penalized for AMI + HF & 0.04 & 0.06\\
+\hspace{1em}Penalized for AMI + Pneumonia & 0.01 & 0.01\\
+\hspace{1em}Penalized for HF + Pneumonia & 0.34 & 0.55\\
+\hspace{1em}Penalized for All Conditions & 0.22 & 0.36\\
+\addlinespace[0.3em]
+Num. Hospitals & 3059 & 680\\
+\bottomrule
+\end{tabular}
+\end{table}
+\hyperlink{sumstats}{\beamerreturnbutton{summary stats}}
+\end{frame}
+
+\begin{frame}{Raw Outcomes with For-Profit}\label{fp_outcomes}
+\begin{center}
+    \includegraphics[width=\textwidth]{}
+\end{center}
     \btVFill
-
-\hyperlink{nonprofithospexec}{\beamerreturnbutton{Nonprofit Hosp Execs}}
-
-\end{frame}
-
-\begin{frame}{Using synthetic DiD weights}\label{synthdid_equation}
-
+    \hyperlink{outcomes}{\beamerreturnbutton{outcomes graph}}
+\end{frame}
+
+\begin{frame}{Intensive Margin}\label{int_margin}
+\large
+    \begin{equation*}
+    y_{ht} = \beta \left(\mathbbm{1}\{\rho \leq \eta\}\text{ x post 2012}\right)_{ht} + \gamma_{h} + \delta_t + \epsilon_{ht}
+    \end{equation*}
+
+\vspace{5mm}
+
+    \begin{equation*}
+    y_{ht} = \beta \left(\mathbbm{1}\{\rho > \eta\}\text{ x post 2012}\right)_{ht} + \gamma_{h} + \delta_t + \epsilon_{ht}
+    \end{equation*}
+\end{frame}
+
+\begin{frame}{Intensive Margin, Readmissions}
+    \begin{columns}
+    \column{0.5\textwidth}
+        \centering
+        below median
+        \includegraphics[width=.8\textwidth]{Objects/cont_belowmedread_md_nomd_synth_graph.pdf}
+    \column{0.5\textwidth}
+        \centering
+        above median
+        \includegraphics[width=.8\textwidth]{Objects/cont_abovemedread_md_nomd_synth_graph.pdf}
+    \end{columns}
     \btVFill
 
-\hyperlink{est_equation}{\beamerreturnbutton{main equation}}
-
-\end{frame}
-
-\begin{frame}{Matching}\label{match_equation}
-
+    \hyperlink{main:readrates}{\beamerreturnbutton{main results}}
+\end{frame}
+
+\begin{frame}{Intensive Margin, Mortality}
+    \begin{columns}
+    \column{0.5\textwidth}
+        \centering
+        below median
+        \includegraphics[width=.8\textwidth]{Objects/cont_belowmedmort_md_nomd_synth_graph.pdf}
+    \column{0.5\textwidth}
+        \centering
+        above median
+        \includegraphics[width=.8\textwidth]{Objects/cont_abovemedmort_md_nomd_synth_graph.pdf}
+    \end{columns}
     \btVFill
 
-\hyperlink{est_equation}{\beamerreturnbutton{main equation}}
-
-\end{frame}
-
-\begin{frame}{Intensive Margin}\label{int_margin}
+    \hyperlink{main:readrates}{\beamerreturnbutton{main results}}
+\end{frame}
+
+\begin{frame}{Due to patient selection practices?}\label{cmi}
+\begin{center}
+    \includegraphics[width=.4\textwidth]{Objects/cmi_md_nomd_synth_graph.pdf}
+\end{center}
 
@@
 \hyperlink{main:readrates}{\beamerreturnbutton{main results}}
+
+\end{frame}
+
+\begin{frame}{Results of other specifications}\label{specifications}
+    \begin{table}[ht!]
+\centering
+\begin{tabular}[t]{lcccccl}
+\toprule
+\multicolumn{1}{c}{} & \multicolumn{3}{c}{Readmission Rate} & \multicolumn{3}{c}{Mortality Rate} \\
+\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-7}
+\multicolumn{1}{c}{ } & \multicolumn{1}{c}{TWFE} & \multicolumn{1}{c}{Matched} & \multicolumn{1}{c}{SC} & \multicolumn{1}{c}{TWFE} & \multicolumn{1}{c}{Matched} & \multicolumn{1}{c}{SC} \\
+\cmidrule(l{3pt}r{3pt}){2-2} \cmidrule(l{3pt}r{3pt}){3-3} \cmidrule(l{3pt}r{3pt}){4-4} \cmidrule(l{3pt}r{3pt}){5-5} \cmidrule(l{3pt}r{3pt}){6-6} \cmidrule(l{3pt}r{3pt}){7-7}
+ & (1) & (2) & (3) & (4) & (5) & (6)\\
+\midrule
+Ever Clinical Exec x Post 2012 & 0.42$^{**}$ & 0.65$^{*}$ & 0.25 & 0.11 & 0.03 & 0.06\\
+ & (0.15) & (0.3) & (0.14) & (0.15) & (0.4) & (0.12)\\
+ &  &  &  &  &  & \\
+Hospital FE & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
+Year FE & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
+\addlinespace
+Observations & 2,387 & 1,025 & 1,640 & 2,381 & 1,023 & 1,640\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\btVFill
+
+\hyperlink{main:readrates}{\beamerreturnbutton{main results}}
+\end{frame}
+
+\begin{frame}{Results by Specialty}\label{specialty}
+\begin{center}
+    \includegraphics[width=.4\textwidth]{Objects/read_specialty_synth_graph.pdf}
+\end{center}
+
+    \btVFill
+
+\hyperlink{main:readrates}{\beamerreturnbutton{main results}}
 \end{frame}

### 2025-10-30 Paper/paperdraft_Sep2024.tex
Updates from Overleaf

@@
          Managers are often seen as drivers of firm decisions, strategies, and performance.
-Further, manager characteristics have been linked to employee wages, decisions of mergers and acquisitions, and financial performance.
-Health care policies in the US often focus on improving quality of care without increasing costs, and yet it is unclear whether hospital managers might contribute to this goal.
-In this paper, I estimate the differential response to such quality incentives by different types of nonprofit hospital executives: those with and without a clinical training background.
-I find that hospitals without clinically trained executives respond more to the financial incentives than those with clinically trained executives.
-I provide evidence that this may be due to differing preferences for profit vs.
-patient welfare.
-Thus, incentivizing clinically trained executives in hospitals could be an effective method to improve value in the US health care system.
+In addition, characteristics of managers have been related to employee wages, mergers and acquisition decisions, and financial performance.
+Health care policies in the US often focus on improving quality of care without increasing costs, but it is unclear how different types of leaders might interact with this goal.
+In this paper, I estimate the differential response to these quality incentives by different types of nonprofit hospital executives: those with and without a clinical training background.
+Using a synthetic difference-in-differences methodology, I show that nonclinical executive teams internalize financial incentives on quality more than clinical executive teams.
+Further, nonclinical executive teams respond to the incentives more similarly to for-profit hospitals, indicating that this may be due to differing preferences for profit versus patient welfare.
 		} 

### 2025-10-30 Paper/paperdraft_fall2025.tex
Updates from Overleaf

--- 2024-10-25 Paper/paperdraft_Oct2024.tex
@@
 
-	\author{\vspace{10mm}Hanna Glenn\footnote{Department of Economics, Emory University, 1602 Fishburne Drive, Atlanta, GA 30322, hanna.glenn@emory.edu.} }
+	\author{\vspace{10mm}Hanna Glenn\footnote{School of Economics, University of Queensland, hanna.glenn@uq.edu.au.} }
 
@@
 		{\small
-         Managers are often seen as drivers of firm decisions, strategies, and performance.
-Further, manager characteristics have been linked to employee wages, decisions of mergers and acquisitions, and financial performance.
-Health care policies in the US often focus on improving quality of care without increasing costs, and yet it is unclear whether hospital managers might contribute to this goal.
-In this paper, I estimate the differential response to such quality incentives by different types of nonprofit hospital executives: those with and without a clinical training background.
+         Firm managers are often seen as key drivers of firm decisions, strategies, and performance.
+While extensive research links managerial characteristics to firm outcomes, less is known about how hospital executives influence responses to health care incentives.
+Health care policies in the US often focus on improving quality of care without increasing costs, yet it is unclear whether hospital leaders might contribute to this goal.
+Leveraging pay-for-performance incentives, I estimate the differential response to financial incentives on quality by different types of nonprofit hospital executives: those with and without a clinical training background.
 I find that hospitals without clinically trained executives respond more to the financial incentives than those with clinically trained executives.
-I provide evidence that this may be due to differing preferences for profit vs.
-patient welfare.
-Thus, incentivizing clinically trained executives in hospitals could be an effective method to improve value in the US health care system.
+This can be explained by a model of hospital objectives in which clinical training affects hospital preferences towards patient welfare.
+These findings highlight the importance of considering leadership as an important factor in hospital decision making when adjusting incentives.
 		} 
@@
 
-    In corporate environments, executives are considered crucial drivers of a company's practices and performance, as they are heavily involved in the culture, visions, and strategies of the firm.
-Executive heterogeneity has been linked to various actions and outcomes of firms; characteristics such as gender, age, and background are correlated with financial performance and behaviors such as employee diversity, wages, aggressiveness in mergers and acquisitions, and other corporate strategies.\footnote{\cite{bertrand2003managing}; \cite{matsa2011chipping}; \cite{matsa2011chipping}; \cite{adams2009women}; \cite{ahern2012changing}; \cite{flabbi2019female}; \cite{benmelech2015military}; \cite{custodio2013generalists}; \cite{frydman2019rising}} However, one aspect of firm leadership that is not well understood is how different leaders may interact with changing policy incentives differently.
-
-    In this paper, I study whether type of executive training affects firm behavior in the context of US hospitals.
-In the US, health care makes up almost 20\% of spending (\cite{CMSNationalHealthExpenditure}), and a major goal of policymakers is to improve the quality of health care without increasing costs.
-Yet, little is known about how hospital managers affect health care value, and it is unclear whether existing findings on executive characteristics can be extrapolated to the health care, and particularly hospital, context.
-Existing literature is generally limited to large, publicly traded firms, whereas a majority of US hospitals are nonprofit with presumably different objectives and policy incentives.
-
-    A characteristic relevant to this context is whether or not hospital executives have clinical training.
-Physicians can bring a unique combination of clinical and administration expertise, and they have insights that can potentially benefit their employees and lead to better quality (\cite{Stajduhar_2023}, \cite{Ahmed_2022}).
+    In corporate environments, executives are considered crucial drivers of a company's practices and performance, as they are the highest level of management that affects day-to-day operations of the organization.
+Executive heterogeneity has been linked to various actions and outcomes of firms; characteristics such as gender, age, and background are correlated with financial performance and behaviors such as employee diversity, wages, aggressiveness in mergers and acquisitions, and other corporate strategies.\footnote{\cite{bertrand2003managing}; \cite{matsa2011chipping}; \cite{matsa2011chipping}; \cite{adams2009women}; \cite{ahern2012changing}; \cite{flabbi2019female}; \cite{benmelech2015military}; \cite{custodio2013generalists}; \cite{frydman2019rising}} However, our understanding is typically limited to executives working for large, publicly traded corporations.
+Additionally, while correlations are an important first step to understanding the effects of leadership, little work has been done to uncover whether such correlations occur due to executive characteristics interacting with the objectives of firms.
+
+    In this paper, I study the effect of specialized executive training on firm behavior when the firm experiences an external shock to their objective function.
+I do this specifically in the context of US hospitals for two reasons.
+First, in the US, health care makes up almost 20\% of spending (\cite{CMSNationalHealthExpenditure}), and a major goal of policymakers is to improve the quality of health care without increasing costs, as the US preforms poorly in these metrics compared to other similar countries.
+Yet, little is known about how leaders affect health care value in the hospitals they run.
+Second, nonprofit hospitals are particularly important in this context, and it is unclear whether existing findings on executive characteristics can be extrapolated to firms with presumably different incentives and objectives than those which have been previously studied.
+
+    An executive characteristic relevant to this context is clinical training.
+Physicians can bring a unique combination of clinical and administration expertise to leadership, and they have insights that can potentially benefit their employees and improve quality (\cite{Stajduhar_2023}, \cite{Ahmed_2022}).
 However, hiring physician executives is controversial, as physicians are not always trained in business management, which can lead to poor practices and lower quality (\cite{HarvardBusinessReview2018}).
-Thus, I estimate how clinically trained executives in hospital leadership affects hospital outcomes.
+Thus, I estimate how the presence of clinically trained executives on a hospital leadership team affects how the hospital makes decisions.
 I scrape executive names and titles of nonprofit hospitals in the US from publicly available Tax Form 990s.
@@
 
-    I investigate whether the result might be explained by differing preferences on the trade-off between profit and patient welfare, where I model these hospital objectives and allow weight placed on each to vary with executive team type.
+    I use a theoretical model to investigate whether these results might be explained by differing preferences on the trade-off between profit and patient welfare, where I model these hospital objectives and allow the weight placed on each to vary with executive team type.
 Hospitals that place more weight on profit will naturally respond more drastically to a policy that adds quality directly into the profit function.
@@
 
-    Finally, the question remains whether clinically trained executives are a signal of unchanging underlying hospital preferences, or if they actively manage the hospital differently, effectively changing hospital objectives.
+    Finally, the question remains as to whether clinically trained executives are simply a signal of unchanging underlying hospital preferences (of, for example, the board of directors), or if they actively manage the hospital differently, effectively changing hospital objectives.
 I analyze this by leveraging changes in hospital propensity to hire clinical executives to decompose the full effect of clinical leadership into a signaling effect (clinical executives signal underlying hospital preferences), or managing effect (clinical executives manage the hospital differently).
 First, I show that hospital executive changes are not correlated with the change to financial incentives.
-For the signaling effect, I compare hospitals who ever have a clinically trained executive to hospitals who never have a clinically trained executive.
+To measure the signaling effect, I compare hospitals who ever have a clinically trained executive to hospitals who never have a clinically trained executive.
 For the managing effect, I limit to hospitals who ever have a clinically trained executive and compare those who have a clinically trained executive when the policies were announced to those who have clinically trained executives at other times.
@@
 
-    \section{Theoretical Framework}
+    \section{Theoretical Framework}\label{sec:theory}
 
@@
 
-    \noindent where $\pi(\theta)$ is a profit function depending on quality $\theta$, $u(.)$ is extra utility gained from quality, which is increasing and concave in $\theta$, and $\alpha\in[0,1]$ captures the weight placed on profit vs.
-societal benefit.
+    \noindent where $\pi(\theta)$ is a profit function depending on quality $\theta$, $u(.)$ is extra utility gained from patient welfare, which is increasing and concave in quality $\theta$, and $\alpha\in[0,1]$ captures the weight placed on profit vs.
+patient welfare.
 There is an implicit stay-open condition throughout the model.
 
-    In this paper, I consider hospital choice of quality under pay-for-performance incentives, which essentially incorporate quality directly into a hospital's profit function (\cite{dranove2011health}).
-Alternatively, a fee-for-service system only incentivizes quantity.
-A hospital that places all value on profit will only care about quality when it enters into the profit function, and therefore will respond more drastically to pay-for-performance than a hospital who already cared about quality through societal benefit before the incentives.
+    I consider hospital choice of quality under pay-for-performance.
+In a world where performance is not incentivized through financial gain (as in traditional fee-for-service settings), hospital profit $\pi$ is maximized when the quantity of patients is maximized.
+However, pay-for-performance incentives incorporate quality into the profit function (\cite{dranove2011health}).
+Therefore, hospitals that only value financial profit will only consider quality if it is an input to the profit function, and therefore will respond more drastically to pay-for-performance than a hospital who values quality as an input to patient welfare before the incentives.
+
     I will now formalize this intuition mathematically.
-For simplicity, I abstract away from quality having a direct impact on quantity of patients or price.\footnote{As long as the relationship between quality and quantity/prove are not dependent on clinical experience on the leadership team, this does not change the conclusion of the model.} Taking the first order condition yields:
+For simplicity, I abstract away from quality having a direct impact on quantity of patients or price.\footnote{As long as the relationship between quality and quantity/price are not correlated with $\alpha$, this does not change the conclusion of the model.} Taking the first order condition yields:
 
@@
 Then $\frac{du'(\theta)}{d\alpha} > 0$.
-Using the Implicit Function, I can write $\frac{d\theta}{d\alpha}$ as
-    $\frac{du'(\theta)/d\alpha}{du'(\theta)/d\theta}<0$, revealing that in a fee-for-service setting, the more weight the hospital places on profit, the lower quality of care chosen.
-
-    Next, assume the hospital is in a strictly pay-for-performance setting, where quality benefits patients and directly increases profit.
+Using the Implicit Function Theorem, I can write $\frac{d\theta}{d\alpha}$ as
+    $\frac{du'(\theta)/d\alpha}{du'(\theta)/d\theta}<0$, revealing that in a fee-for-service setting, the more weight the hospital places on profit, the lower quality of care chosen at baseline.
+
+    Next, assume the hospital is in a pay-for-performance setting, where quality benefits patients and directly increases profit.
 Then, assuming the marginal financial benefit of increasing quality is greater than the marginal financial cost, $\pi(\theta)$ is increasing in $\theta$.
@@
 
-    Hence, this model predicts that change in quality depends on how much weight the hospital places on profit vs.
-societal benefit.
-Particularly, hospitals with more weight on profit respond more than hospitals with more weight placed on societal benefit.
-Therefore, for leadership characteristics that are plausibly correlated with how much weight is placed on patient welfare versus quality, such as clinical training, we expect to see a differential response to pay-for-performance incentives being enacted.
+    The first prediction of the model is that, all else held equal, hospitals that value profit will have a lower quality than those who value patient welfare in a fee-for-service setting.
+Second, a hospital's change in quality once pay-for-performance is enacted depends on how much weight the hospital places on profit vs.
+patient welfare.
+Particularly, hospitals with higher weight on profit respond more than hospitals with higher weight on patient welfare.
+Therefore, factors that affect weight $\alpha$ will be correlated with choice of quality, and with how quality changes under new pay-for-performance incentives.
+In this paper, I test the hypothesis that clinical experience on a hospital's leadership team is a factor that is correlated with weight $\alpha$.
 
@@
 
-    For each hospital in the tax form data, I merge to hospital matches in the AHA data based on name and location.\footnote{I assess the observable differences between matched and non-matched hospitals in Appendix \ref{app:matched}, and conclude that the samples are overall similar apart from an under-sampling of hospitals that belong to systems.} For all hospitals that can be matched to a general hospital in the AHA data, I use optimal character recognition (OCR) methods to extract the names, titles, and positions of each listed executive from its Tax Form 990.
+    For each hospital in the tax form data, I merge to hospital matches in the AHA data based on name and location.\footnote{I assess the observable differences between matched and non-matched hospitals in Appendix \ref{app:matched}, and conclude that the samples are overall similar apart from an under-sampling of hospitals that belong to systems.} For all hospitals that can be matched to a general hospital in the AHA data, I use optical character recognition (OCR) methods to extract the names, titles, and positions of each listed executive from its Tax Form 990.
 While board members are an important aspect of hospital organization, I do not include them in this analysis as they are beyond the scope of this paper.
@@
 
-    The hypothesis driven from the theory described in Section \ref{sec:forprofit} is that clinically trained executives put more weight on quality of care before the financial incentives on quality began.
+    The hypothesis driven from the theory described in Section \ref{sec:theory} is that clinically trained executives put more weight on quality of care before the financial incentives on quality began.
 To investigate this hypothesis empirically, I merge all general for-profit hospitals in the AHA survey to my sample of nonprofit hospitals.
@@
  \vspace{2mm}
- Treat Group: \hspace{11mm} No MD in 2012 \\
+ Treat Group: \hspace{11mm} MD in 2012 \\
  \vspace{2mm}

### 2025-10-30 Slides/ASHE2025.tex
Updates from Overleaf

--- 2024-11-13 Slides/JobTalk.tex
@@
 % Author 
-\node[below=1cm] (author) at (title.south){\large Hanna Glenn, Emory University};
+\node[below=1cm] (author) at (title.south){\large Hanna Glenn};
+% Affiliation
+\node[below=0.15cm] (affiliation) at (author.south){\large The University of Queensland};
 % Date
-\node[below=0.25cm] (date) at (author.south){\today};
+\node[below=0.5cm] (date) at (affiliation.south){\today};
 \end{tikzpicture}
@@
 \begin{frame}{Leadership Characteristics and Firm Behavior/Performance}
-
-\large
-
-\textbf{Gender}
+\large
+\begin{columns}[T]
+    \begin{column}{0.45\textwidth}
+    \textbf{Characteristics:}
         \begin{itemize}
-    \item \normalsize Female board members $\implies$ lower firm value, less employment reduction \\ \scriptsize(\cite{ahern2012changing}; \cite{matsa2013female}) \normalsize
-    \item Female executives $\implies$ female wage distribution \scriptsize (\cite{flabbi2019female}) \normalsize
+            \item \tikz[baseline]{\node[anchor=base] (L1) {Gender};}
+            \item \tikz[baseline]{\node[anchor=base] (L2) {Experience};}
+            \item \tikz[baseline]{\node[anchor=base] (L3) {Age};}
         \end{itemize}
-
-\vspace{4mm}
-
-\textbf{Prior Experience}
+    \end{column}
+    \begin{column}{0.45\textwidth}
+    \textbf{Firm Outcomes:}
         \begin{itemize}
-    \item \normalsize General background CEOs paid more \scriptsize (\cite{custodio2013generalists}) \normalsize
-    \item Background matters for acquisitions \scriptsize(\cite{custodio2013ceos}) \normalsize
-
-\vspace{1mm}
-
-    \item Military experience CEOs \scriptsize (\cite{benmelech2015military})
+            \item \tikz[baseline]{\node[anchor=base] (R1) {Value};}
+            \item \tikz[baseline]{\node[anchor=base] (R2) {Employees};}
+            \item \tikz[baseline]{\node[anchor=base] (R3) {M\&A};}
+            \item \tikz[baseline]{\node[anchor=base] (R4) {Leader Salaries};}
         \end{itemize}
-
-\vspace{4mm}
-
-\textbf{Age}
-\begin{itemize}
-    \item \normalsize Young male CEOs more aggressive in M\&A \scriptsize (\cite{levi2010deal})
-\end{itemize}
-
-\end{frame}
-
-\begin{frame}{Gaps in Literature}
-    \large
+    \end{column}
+\end{columns}
+
+% Now draw arrows between the nodes
+\begin{tikzpicture}[overlay, remember picture]
+    \draw[->, thick, shorten >=15pt] (L1.east) to [out=0, in=180] (R1.west);
+    \draw[->, thick, shorten >=15pt] (L1.east) to [out=0, in=180] (R2.west);
+    \draw[->, thick, shorten >=15pt] (L2.east) to [out=0, in=180] (R3.west);
+    \draw[->, thick, shorten >=15pt] (L2.east) to [out=0, in=180] (R4.west);
+    \draw[->, thick, shorten >=15pt] (L3.east) to [out=0, in=180] (R3.west);
+\end{tikzpicture}
+
+\vspace{1.5em}\pause
+
+\textbf{Gaps in Literature:}
+\vspace{2mm}
 \begin{enumerate}
     \item Findings are limited to large, publicly traded firms
-
-\vspace{2mm}
-
-        \begin{itemize}
-            \item \normalsize Different objectives/incentives in other types of firms (nonprofits)
-        \end{itemize}
-
-        \vspace{10mm}
-
-        \item Unclear whether leadership matters when policies change incentives
+    \item Unclear in the context of exogenous policy shocks
 \end{enumerate}
+
 \end{frame}
@@
 
-\vspace{5mm} \large \pause
+\vspace{10mm} \large
 
 \begin{enumerate}
-    \item \textbf{Hospital care makes up a third of health care spending} \scriptsize (\cite{cms2024nhe}) \normalsize
-
-\vspace{2mm}
-
-    \begin{itemize}
-        \item \normalsize Nonprofits drive the hospital industry \scriptsize (\cite{ASPE_2023}) \normalsize
-    \end{itemize}
-
-\vspace{8mm}
-
-    \item \large \textbf{Policies often aim to improve the value of health care delivery}
-
-\vspace{2mm}
-
-    \begin{wideitemize}
-        \item \normalsize Unclear how leaders interact with this goal
-    \end{wideitemize}
+    \item Nonprofits drive the hospital industry \scriptsize (\cite{ASPE_2023}) \large
+
+    \vspace{3mm}
+
+    \item Policies often aim to improve the value of health care delivery
+
+\vspace{2mm}
 
@@
     \Large
-    Do executives with clinical training
+    Does having clinically trained executives
 
@@
 
-    affect response to pay-for-performance incentives
+    affect the response to pay-for-performance incentives
 
@@
 
-\begin{frame}{Why Clinical Training?}
-
-\large
-
-\textbf{Do physicians make good firm leaders?}
-
-\vspace{2mm}
-
-\begin{wideitemize}
-    \item Relevant expertise \scriptsize (\cite{Stajduhar_2023}, \cite{Ahmed_2022}) \large
-
-    \item Lack of business training \scriptsize (\cite{HarvardBusinessReview2018})
-\end{wideitemize}
-
-\vspace{12mm}
-
-\textbf{Potentially correlated with how the leader values quality versus profit}
-\end{frame}
-
-\begin{frame}{Pay-for-Performance Incentives}
-    \large
-    \begin{block}{\Large Main Takeaway:}
-        \large 
-        \centering
-        \vspace{1mm}
-        Hospitals now have financial incentives to improve quality
-
-        \vspace{1mm}
-    \end{block}
-
-    \vspace{6mm}
-
-    \textbf{Quality: 30-day readmissions \& 30-day mortality}
-
-    \vspace{5mm}
-
-    Hospital Readmissions Reduction Program
-    \begin{itemize}
-        \item Penalized high readmissions
-        \item Pneumonia, Heart Attack (AMI), Heart Failure
-    \end{itemize}
-
-    \vspace{5mm}
-
-    Hospital Value-Based Purchasing Program
-    \begin{itemize}
-        \item Score metric determines rewards
-        \item Rewards high metric
-    \end{itemize}
-\end{frame}
-
-\begin{frame}{Preview of Main Findings}
-
-\large
-
-\textbf{Non-clinical executive teams respond more to financial incentives on quality}
-
-\vspace{3mm}
-
-\begin{wideitemize}
-    \item Through readmission rates
-    \item Non-clinical executive teams behave like profit-driven hospitals
-\end{wideitemize}
-\end{frame}
-
-\begin{frame}{Contribution}
-\large
-    \textbf{Hospital Executives/Board of Directors}
-
-\vspace{2mm}
-
-    \begin{wideitemize}
-        \item \normalsize Physician board members $\implies$ lower public donations \scriptsize (\cite{brickley2010board}) \normalsize
-        \item No correlation b/w CEOs and hospital productivity in England \scriptsize (\cite{janke2019impact})\normalsize
-        \item Public hospitals in Chile improved when older physicians left CEO \scriptsize (\cite{otero2022managers}) \normalsize
-    \end{wideitemize}
-
-    \vspace{5mm}\pause
-
-    \textbf{Hospital Response to Pay-for-Performance}
-
-\vspace{2mm}
-
-    \begin{wideitemize}
-        \item \normalsize Hospitals decreased readmissions \scriptsize (\cite{gupta2021impacts}) \normalsize
-        \item Market concentration and systems affect response \scriptsize (\cite{kunz2024assessing}) \normalsize
-        \item Rewards had little impact on quality \scriptsize (\cite{norton2018moneyball}; \cite{friedson2019so})
-    \end{wideitemize}
-\end{frame}
-
-\begin{frame}{Rest of Talk}
-
-\large
-
-    Setting and Data Collection
-
-    \vspace{5mm}
-
-    About Executive Teams
-
-    \vspace{5mm}
-
-    Theoretical Motivation
-
-    \vspace{5mm}
-
-    Effect of Clinical Leadership
-
-    \vspace{5mm}
-
-    Comparison to For-Profits
-
-    \vspace{5mm}
-
-    Signaling vs. Managing
+\begin{frame}{More Details}
+    \large
+    I leverage several 2012 pay-for-performance initiatives
+
+    \vspace{2mm}
+    \begin{wideitemize}
+        \item Penalty (reward) for below (above) average \hspace{-8mm} $\underbrace{\text{quality}}_{\text{30-day read \& mort}}$
+        \item \textit{Changes the relationship between revenue and quality in objective function}
+    \end{wideitemize}
+
+    \vspace{10mm}
+
+    Why clinical training?
+
+    \vspace{2mm}
+    \begin{wideitemize}
+        \item Potentially correlated with hospital's preferences towards revenue versus patient welfare
+    \end{wideitemize}
+
 \end{frame}
@@
 \section{Setting and Data Collection}
-
-\begin{transitionframe}
-\centering \LARGE
-    \textcolor{black}{Setting and Data Collection}
-\end{transitionframe}
-
-\begin{frame}{About the Executives}\label{nonprofithospexec}
-
-\large
-\textbf{What do they do?}
-
-\vspace{2mm}
-
-\begin{wideitemize}
-    \item Board of directors selects executives
-    \item Executives handle the day-to-day operations of the hospital 
-    \item Quality often improved from the top down \scriptsize (\cite{silow2011reducing})
-\end{wideitemize}
-
-\vspace{4mm} \pause
-
-\textbf{Executive teams}
-
-\vspace{2mm}
-
-\begin{wideitemize}
-    \item Lot of variation in team organization
-    \item CEO and CFO
-    \item I focus on the team as a whole instead of specific executive roles
-\end{wideitemize}
-
-\btVFill
-
-\hyperlink{executivejob}{\beamergotobutton{executive job}}
-
-\end{frame}
 
@@
 
-\begin{frame}{Data Collection Process}
-\begin{center}
-\begin{tikzpicture}[
-    node distance=3.1cm and 1.8cm,
-    every node/.style={align=center, font=\normalsize},
-    process/.style={rectangle, draw, rounded corners, text centered, minimum width=2.5cm, minimum height=2cm, fill=purple!30},
-    output/.style={rectangle, draw, text centered, minimum width=2.7cm, minimum height=2.5cm, fill=purple},
-    extract/.style={rectangle, draw, rounded corners, text centered, minimum width=2.5cm, minimum height=2cm, fill=sage!30},
-    arrow/.style={thick, ->, >=stealth}
-]
-
-% Nodes
-\node (step1) [process] {Hospital \\EINs};\pause
-\node (step2) [process, right of=step1] {Download \\PDFs};
-\draw[arrow] (step1) -- (step2); \pause
-\node (step3) [process, right of=step2] {Extract\\text};
-\draw[arrow] (step2) -- (step3); 
-% New green box below step3
-\node (extract) [extract, below of=step3, yshift=0.5cm] {Optical\\Character\\Recognition}; 
-\draw[arrow] (extract) -- (step3);\pause
-\node (step4) [process, right of=step3] {String\\cleaning\\algorithms};
-\draw[arrow] (step3) -- (step4); \pause
-\node (output) [output, right of=step4, text=white] {Executive\\Name\\Position\\Title};
-\draw[arrow] (step4) -- (output);
-\end{tikzpicture}
-\end{center}
-\end{frame}
-
-\begin{frame}{Merge to Other Data}
-
-\large
+\begin{frame}{Final Data Set}
+
+\large
+
+    \textbf{Executives:}
+
+    \vspace{1mm}
+
+    \begin{itemize}
+        \item Top level C-suite employees scraped from 990s
+    \end{itemize}
+
+    \vspace{4mm}
+
     \textbf{Hospital:}
 
-    \vspace{2mm}
-    \begin{wideitemize}
-        \item AHA
-        \item Hospital Compare
-        \item CMS Case Mix Index
-        \item Health Care Cost Report Information System (HCRIS)
-    \end{wideitemize}
-
-    \vspace{7mm}
+    \vspace{1mm}
+    \begin{wideitemize}
+        \item AHA, Hospital Compare, CMS Case Mix Index, HCRIS
+    \end{wideitemize}
+
+    \vspace{4mm}
 
@@
 
-    \vspace{2mm}
+    \vspace{1mm}
 
@@
     \end{itemize}
+
+    \vspace{8mm}
+
+    \textbf{Years: 2010-2014}
 \end{frame}
@@
 \section{About Executive Teams}
-
-\begin{transitionframe}
-\centering \LARGE
-    \textcolor{black}{Descriptive Statistics}
-\end{transitionframe}
 
@@
 Fraction Clinical Execs & 0.20 & 0.26 & 0.00\\
-Fraction Int. Medicine Execs & 0.06 & 0.08 & 0.00\\
 Has a CMO & 0.55 & 0.52 & 0.07\\
@@
 \end{table}
+\end{frame}
+
+\begin{frame}{Raw Outcomes, controlling for baseline case mix index}\label{outcomes}
+\begin{center}
+    \includegraphics[width=.92\textwidth]{Objects/outcomes_graph_presentation.pdf}
+\end{center}
+    \btVFill
+    \hyperlink{fp_outcomes}{\beamergotobutton{with for-profit}}
+\end{frame}
+
+\section{Effect of Clinical Leadership on Response}
+
+\begin{transitionframe}
+\centering \LARGE
+    \textcolor{black}{Effect of Clinical Leadership on Response to P4P}
+\end{transitionframe}
+
+\begin{frame}{Estimation}\label{est_equation}
+
+\large
+    \textbf{Compare always clinical to never clinical hospital executive teams:}
+
+    \vspace{3mm}\large
+    \begin{equation*}
+    y_{ht} = \beta \text{ (clinical training x post 2012)}_{ht} + \gamma_{h} + \delta_t + \epsilon_{ht}
+    \end{equation*}
+
+    \vspace{3mm}
+    \begin{wideitemize}
+        \item $y_{ht}$: weighted average readmission/ mortality rates
+        \item Leverage \textit{timing} of the programs
+    \end{wideitemize}
+
+    \vspace{8mm}\pause
+
+    Different methods:
+    \begin{itemize}
+        \item No weights
+        \item \textbf{Weights based on outcomes pre-2012} 
+        \item Coarsened exact matching 
+    \end{itemize}
+\end{frame}
+
+\begin{frame}{Identification Assumptions}\label{ident}
+
+\large
+\begin{wideitemize}
+    \item Parallel trends
+    \item No other 2012 event correlated with executive team and outcomes
+    \item No hospital characteristics perfectly correlated with executive team
+    \item Stability in clinical executives not correlated with incentive change \hyperlink{changes}{\beamergotobutton{analysis}}
+    \begin{itemize}
+        \item No correlation between 2012 and likelihood of changing propensity to hire clinical executive
+        \item No correlation between exposure to policies and likelihood of changing propensity
+    \end{itemize}
+\end{wideitemize}
+\end{frame}
+
+\begin{frame}{Readmission Rates}\label{main:readrates}
+\large
+    \begin{columns}
+    \column{0.5\textwidth}
+        \centering
+        \includegraphics[width=.9\textwidth]{Objects/read_md_nomd_synth_graph.pdf}
+    \column{0.5\textwidth}
+        \begin{wideitemize}
+            \item Magnitude: 0.3ppt difference
+
+\vspace{5mm}
+
+            \item Large relative to effect of HRRP (1ppt)
+        \end{wideitemize}
+
+        \vspace{20mm}
+
+        \hyperlink{int_margin}{\beamergotobutton{intensive margin}}
+        \hyperlink{cmi}{\beamergotobutton{due to patient selection?}}
+        \hyperlink{specifications}{\beamergotobutton{specifications}}
+        \hyperlink{specialty}{\beamergotobutton{specialty}}
+
+    \end{columns}
+\end{frame}
+
+\begin{frame}{Mortality Rates}
+\large
+    \begin{columns}
+    \column{0.5\textwidth}
+        \centering
+        \includegraphics[width=.9\textwidth]{Objects/mort_md_nomd_synth_graph.pdf}
+    \column{0.5\textwidth}
+        \begin{wideitemize}
+            \item Magnitude: 0.1ppt
+
+            \vspace{5mm}
+
+            \item Mortality is noisy measure 
+        \end{wideitemize}
+    \end{columns}
+\end{frame}
+
+\section{Comparison to For-Profit Hospitals}
+
+\begin{transitionframe}
+\centering \LARGE
+    \textcolor{black}{Comparison to For-Profit Hospitals}
+\end{transitionframe}
+
+\begin{frame}{Estimation}
+\large
+    \textbf{Compare both nonprofit leadership team types to for-profits:}
+
+    \vspace{5mm}\large
+    \begin{equation*}
+    y_{ht} = \beta \text{ (for-profit x post 2012)}_{ht} + \gamma_{h} + \delta_t + \epsilon_{ht}
+    \end{equation*}
+
+    \vspace{10mm}
+
+    Comparison groups:
+    \begin{itemize}
+        \item nonprofit clinical; nonprofit non-clinical
+    \end{itemize}
+
+\end{frame}
+
+\begin{frame}{Readmission Rate}
+    \begin{columns}
+    \column{0.5\textwidth}
+        \centering
+        \includegraphics[width=.9\textwidth]{Objects/fp_read_md_synth_graph.pdf}
+    \column{0.5\textwidth}
+        \centering
+        \includegraphics[width=.9\textwidth]{Objects/fp_read_nomd_synth_graph.pdf}
+    \end{columns}
+\end{frame}
+
+\begin{frame}{Mortality Rate}\label{forprofmort}
+    \begin{columns}
+    \column{0.45\textwidth}
+        \centering
+        \includegraphics[width=.9\textwidth]{Objects/fp_mort_md_synth_graph.pdf}
+    \column{0.45\textwidth}
+        \centering
+        \includegraphics[width=.9\textwidth]{Objects/fp_mort_nomd_synth_graph.pdf}
+    \end{columns}
+
+    \hyperlink{unobserved}{\beamergotobutton{unobserved}}
+\end{frame}
+
+\section{Unobserved}
+
+\begin{frame}{Summary}
+\large
+    \textbf{We know very little about whether leaders matter in health care}
+
+    \vspace{2mm}
+
+    \begin{wideitemize}
+        \item This paper: non-clinical executive teams internalize financial incentives more
+    \end{wideitemize}
+
+    \vspace{8mm}
+
+    \textbf{Implications}
+
+    \vspace{2mm}
+    \begin{wideitemize}
+        \item Despite all being nonprofits, hospitals with clinician leaders act more like we would expect nonprofits to
+        \item In the absence of clinical leadership, nonprofits just act like for-profits
+    \end{wideitemize}
+
+    \vspace{5mm}
+
+    \begin{center}
+        \Large Thank you! 
+    \end{center}
+\end{frame}
+
+\section{Appendix}
+
+\begin{transitionframe}
+\centering \LARGE
+    \textcolor{black}{Appendix Slides}
+\end{transitionframe}
+
+\appendix
+
+\begin{frame}{Choice of Quality}\label{choiceofqualityapp}
+Take FOC, rearrange and take derivative wrt $\alpha$:
+
+\vspace{5mm}
+
+\begin{equation*}
+    \frac{du'(\theta)}{d\alpha} = \frac{-1}{(1-\alpha)^2}\pi'(\theta)
+\end{equation*}
+
+\vspace{5mm}
+
+$\pi'(\theta)$:
+
+\vspace{3mm}
+
+\begin{wideitemize}
+    \item In fee-for-service setting: negative $\implies \frac{du'(\theta)}{d\alpha} > 0$
+    \item In pay-for-performance setting: positive $\implies \frac{du'(\theta)}{d\alpha} < 0$
+\end{wideitemize}
+
+\end{frame}
+
+\begin{frame}{FFS $\implies$ P4P}
+
+\vspace{3mm{}}
+    \large
+
+        \begin{block}{}
+        \centering \Large
+        Hospitals with higher weight on profit will respond more
+
+        to the financial incentives on quality
+    \end{block}
+
+    \begin{align*}
+        \frac{d\Delta\theta}{d\alpha}&=\frac{d\theta_{p4p}}{d\alpha}-\frac{d\theta_{ffs}}{d\alpha}\\ 
+        &\\
+        &=\frac{\frac{du'(\theta)}{d\alpha}}{\frac{du'(\theta)}{d\theta_{p4p}}} - \frac{\frac{du'(\theta)}{d\alpha}}{\frac{du'(\theta)}{d\theta_{ffs}}}\\
+        &\\
+        &> 0
+    \end{align*}
+
+    \btVFill \normalsize
+    \hyperlink{choiceofquality}{\beamerreturnbutton{hospital choice of quality}}
+\end{frame}
+
+\begin{frame}{How Leaders Affect Quality}\label{executivejob}
+\large \vspace{5mm}
+\textbf{Ways to Improve Quality/ Decrease Readmissions}
+
+\vspace{2mm}
+
+\begin{wideitemize}
+    \item Care transition from inpatient to outpatient
+    \item Patient education
+    \item Improve care coordination
+    \item Follow-up appointments
+\end{wideitemize}
+
+\btVFill
+
+\scriptsize (\cite{silow2011reducing}) \hspace{4mm}
+\hyperlink{nonprofithospexec}{\beamerreturnbutton{Nonprofit Hosp Execs}}
+
 \end{frame}
@@
 \hspace{1em}Penalized for one condition & 0.19 & 0.25 & 0.13\\
-\hspace{1em}Penalized for AMI + HF & 0.07 & 0.05 & 0.03\\
-\hspace{1em}Penalized for AMI + Pneumonia & 0.02 & 0.02 & 0.02\\
-\hspace{1em}Penalized for HF + Pneumonia & 0.41 & 0.40 & 0.33\\
-\hspace{1em}Penalized for All Conditions & 0.31 & 0.31 & 0.20\\
+\hspace{1em}Penalized for two conditions & 0.50 & 0.47 & 0.38\\
+\hspace{1em}Penalized for all Conditions & 0.31 & 0.31 & 0.20\\
 \bottomrule
@@
 \hyperlink{forprofitstats}{\beamergotobutton{for-profit stats}}
-\end{frame}
-
-\begin{frame}{Raw Outcomes, controlling for baseline case mix index}\label{outcomes}
-\begin{center}
-    \includegraphics[width=.92\textwidth]{Objects/outcomes_graph_presentation.pdf}
-\end{center}
-    \btVFill
-    \hyperlink{fp_outcomes}{\beamergotobutton{with for-profit}}
-\end{frame}
-
-\section{Theoretical Framework}
-
-\begin{transitionframe}
-\centering \LARGE
-    \textcolor{black}{Theoretical Framework}
-\end{transitionframe}
-
-\begin{frame}{Hospital Objectives: Profit versus Patient Welfare}
-    \Large
-    \centering
-    \begin{equation*}
-        \max_{\theta}\hspace{2mm} \alpha\pi(\theta) + (1-\alpha)u(\theta)
-    \end{equation*}
-\end{frame}
-
-\begin{frame}[noframenumbering]{Hospital Objectives: Profit versus Patient Welfare} \Large
-    \begin{equation*}
-\max_{\theta}\hspace{2mm}\alpha
-\tikz[baseline]{
-    \node[draw=sage,rounded corners,anchor=base] (m1)
-    {$\pi(\theta)$};
-    \node[above of=m1, text=sage, yshift=5mm] (l1) {profit};
-    \draw[-,sage] (l1) -- (m1);
-}
-+ (1-\alpha)\hspace{-11mm}
-\tikz[baseline]{
-    \node[draw=sage,rounded corners,anchor=base] (m2)
-    {$u(\theta)$};
-    \node[above of=m2, text=sage, yshift=5mm] (l2) {patient welfare};
-    \draw[-,sage] (l2) -- (m2);
-}
-\end{equation*}
-\end{frame}
-
-\begin{frame}[noframenumbering]{Hospital Objectives: Profit versus Patient Welfare} \Large
-    \begin{equation*}
-\max_{\theta}\hspace{2mm}\alpha\pi(\hspace{-5mm}
-\tikz[baseline]{
-    \node[draw=purple,rounded corners,anchor=base] (m1)
-    {$\theta$};
-    \node[above of=m1, text=purple, yshift=5mm] (l1) {quality};
-    \draw[-,purple] (l1) -- (m1);
-}
-\hspace{-5mm}) + (1-\alpha)u(\hspace{-5mm}
-\tikz[baseline]{
-    \node[draw=purple,rounded corners,anchor=base] (m2)
-    {$\theta$};
-    \node[above of=m2, text=purple, yshift=5mm] (l2) {quality};
-    \draw[-,purple] (l2) -- (m2);
-}
-\hspace{-5mm})
-\end{equation*}
-\end{frame}
-
-\begin{frame}[noframenumbering]{Hospital Objectives: Profit versus Patient Welfare} \Large
-    \begin{equation*}
-\max_{\theta}\hspace{-4mm}
-\tikz[baseline]{
-    \node[draw=earthyellow,rounded corners,anchor=base] (m1)
-    {$\alpha$};
-    \node[above of=m1, text=earthyellow, yshift=5mm] (l1) {weight on\\profit};
-    \draw[-,earthyellow] (l1) -- (m1);
-}
-\hspace{-8mm}\pi(\theta) + \hspace{-7mm}
-\tikz[baseline]{
-    \node[draw=earthyellow,rounded corners,anchor=base] (m2)
-    {$(1-\alpha)$};
-    \node[above of=m2, text=earthyellow, yshift=5mm] (l2) {weight on\\patient welfare};
-    \draw[-,earthyellow] (l2) -- (m2);
-}
-\hspace{-7mm}u(\theta)
-\end{equation*}
-\end{frame}
-
-\begin{frame}{Choice of Quality}\label{choiceofquality}
-
-\vspace{9mm}
-
-\Large
-
-    \begin{columns}
-    \begin{column}{0.5\textwidth}
-    \begin{center}
-        \underline{Fee-for-Service}
-
-        \vspace{3mm}
-
-            $\pi$ $\downarrow$ in $\theta$
-    \end{center}
-    \end{column}
-    \begin{column}{0.5\textwidth}  %%<--- here
-    \begin{center}
-        \underline{Pay-for-Performance}
-
-        \vspace{3mm}
-
-            $\pi$ $\uparrow$ in $\theta$
-    \end{center}
-    \end{column}
-\end{columns}
-
-\btVFill
-
-\begin{block}{\Large Fee-for-service to Pay-for-performance}
-    \centering
-    \vspace{3mm}
-    Higher $\alpha$ $\implies$ larger response
-
-    \vspace{3mm}
-\end{block}
-
-\hyperlink{choiceofqualityapp}{\beamergotobutton{more details}}
-\end{frame}
-
-\begin{frame}{Informing the Empirical Analysis}
-
-\large
-    \begin{enumerate}
-        \item If clinicians affect $\alpha$, they affect quality change after P4P
-
-        \vspace{8mm}\pause
-        \item If non-clinicians have higher $\alpha$, they act more like for-profits
-
-        \vspace{8mm}\pause
-        \item Mechanism could be revealing objectives or changing objectives
-    \end{enumerate}
-\end{frame}
-
-\section{Effect of Clinical Leadership on Response}
-
-\begin{transitionframe}
-\centering \LARGE
-    \textcolor{black}{Effect of Clinical Leadership on Response to P4P}
-\end{transitionframe}
-
-\begin{frame}{Estimation}\label{est_equation}
-
-\large
-    \textbf{Compare always clinical to never clinical hospital executive teams:}
-
-    \vspace{3mm}\large
-    \begin{equation*}
-    y_{ht} = \beta \text{ (clinical training x post 2012)}_{ht} + \gamma_{h} + \delta_t + \epsilon_{ht}
-    \end{equation*}
-
-    \vspace{3mm}
-    \begin{wideitemize}
-        \item $y_{ht}$: weighted average readmission/ mortality rates
-        \item Leverage \textit{timing} of the programs
-    \end{wideitemize}
-
-    \vspace{8mm}\pause
-
-    Different methods:
-    \begin{itemize}
-        \item No weights
-        \item \textbf{``Synthetic DiD" type weights} 
-        \item Coarsened exact matching 
-    \end{itemize}
-\end{frame}
-
-\begin{frame}{Identification Assumptions}
-
-\large
-\begin{wideitemize}
-    \item Parallel trends
-    \item No other 2012 event correlated with executive team and outcomes
-    \item No hospital characteristics perfectly correlated with executive team
-    \item No anticipation of incentive changes
-    \item Stability in clinical executives not correlated with incentive change
-\end{wideitemize}
-\end{frame}
-
-\begin{frame}{Average Team Changes Over Time}
-\centering 
-    \includegraphics[width=.8\textwidth]{Objects/change_means.pdf}
-\end{frame}
-
-\begin{frame}{Changes are not correlated with policy exposure}
-\centering \small
-\begin{equation*}\label{eq:change2}
-    \text{change}_{ht} = \sum_{j=2011}^{2014}\beta_j(\mathbf{1}\{t=j\}\times \text{Program Exposed})_{ht} + \alpha_h + \epsilon_{ht},
-    \end{equation*}
-
-    \btVFill
-
-    \includegraphics[width=.7\textwidth]{Objects/change_analysis_plot.pdf}
-
-\end{frame}
-
-\begin{frame}{Readmission Rates}\label{main:readrates}
-\large
-    \begin{columns}
-    \column{0.5\textwidth}
-        \centering
-        \includegraphics[width=.9\textwidth]{Objects/read_md_nomd_synth_graph.pdf}
-    \column{0.5\textwidth}
-        \begin{wideitemize}
-            \item Magnitude: 0.3ppt difference
-
-\vspace{5mm}
-
-            \item Large relative to effect of HRRP (1ppt)
-        \end{wideitemize}
-
-        \vspace{20mm}
-
-        \hyperlink{int_margin}{\beamergotobutton{intensive margin}}
-        \hyperlink{cmi}{\beamergotobutton{due to patient selection?}}
-        \hyperlink{specifications}{\beamergotobutton{specifications}}
-        \hyperlink{specialty}{\beamergotobutton{specialty}}
-
-    \end{columns}
-\end{frame}
-
-\begin{frame}{Mortality Rates}
-\large
-    \begin{columns}
-    \column{0.5\textwidth}
-        \centering
-        \includegraphics[width=.9\textwidth]{Objects/mort_md_nomd_synth_graph.pdf}
-    \column{0.5\textwidth}
-        \begin{wideitemize}
-            \item Magnitude: 0.1ppt
-
-            \vspace{5mm}
-
-            \item Mortality is noisy measure 
-        \end{wideitemize}
-    \end{columns}
-\end{frame}
-
-\section{Comparison to For-Profit Hospitals}
-
-\begin{transitionframe}
-\centering \LARGE
-    \textcolor{black}{Comparison to For-Profit Hospitals}
-\end{transitionframe}
-
-\begin{frame}{Estimation}
-\large
-    \textbf{Compare both nonprofit leadership team types to the response of for-profits:}
-
-    \vspace{5mm}\large
-    \begin{equation*}
-    y_{ht} = \beta \text{ (for-profit x post 2012)}_{ht} + \gamma_{h} + \delta_t + \epsilon_{ht}
-    \end{equation*}
-
-    \vspace{10mm}
-
-    Comparison group:
-    \begin{itemize}
-        \item nonprofit clinical; nonprofit non-clinical
-    \end{itemize}
-
-\end{frame}
-
-\begin{frame}{Readmission Rate}
-    \begin{columns}
-    \column{0.5\textwidth}
-        \centering
-        \includegraphics[width=.9\textwidth]{Objects/fp_read_md_synth_graph.pdf}
-    \column{0.5\textwidth}
-        \centering
-        \includegraphics[width=.9\textwidth]{Objects/fp_read_nomd_synth_graph.pdf}
-    \end{columns}
-\end{frame}
-
-\begin{frame}{Mortality Rate}
-    \begin{columns}
-    \column{0.5\textwidth}
-        \centering
-        \includegraphics[width=.9\textwidth]{Objects/fp_mort_md_synth_graph.pdf}
-    \column{0.5\textwidth}
-        \centering
-        \includegraphics[width=.9\textwidth]{Objects/fp_mort_nomd_synth_graph.pdf}
-    \end{columns}
-\end{frame}
-
-\section{Signaling vs. Managing}
-
-\begin{transitionframe}
-\centering \LARGE
-    \textcolor{black}{Signaling vs. Managing}
-\end{transitionframe}
-
-\begin{frame}{Hypotheses}
-
-\large
-    \textbf{Clinical leaders represent underlying objectives ($\alpha$)}
-
-\vspace{2mm}
-
-    \begin{wideitemize}
-        \item Timing of the clinical leader doesn't matter 
-    \end{wideitemize}
-
-    \vspace{15mm}
-
-    \textbf{Clinical leaders effectively change objectives ($\alpha$)}
-
-\vspace{2mm}
-
-    \begin{wideitemize}
-        \item Timing of the clinical leader does matter
-    \end{wideitemize}
-\end{frame}
-
-\begin{frame}{Estimation}
-\large
-    \begin{figure}[ht!]
-\begin{center}
-
- \begin{tabular}{| m{15em} |}
- \hline \vspace{1mm}
- \textbf{Signaling:}\\ [0.5ex]
- \hline\hline 
- \vspace{2mm}
- Treat Group:  \hspace{15mm} Never MD \\
- \vspace{2mm}
- Comparison Group: \hspace{3mm} Ever MD  \\
- [1ex]
- \hline
- \end{tabular}
-\hfil   %<---
- \begin{tabular}{|m{18em}|}
- \hline \vspace{1mm}
- \textbf{Managing:}\\ [0.5ex]
- \hline\hline
- \vspace{2mm}
- Treat Group: \hspace{11mm} MD in 2012 \\
- \vspace{2mm}
- Comparison Group:  MD not in 2012  \\
- [1ex]
- \hline
- \end{tabular}
-
-\end{center}
- \end{figure}
-
-\end{frame}
-
-\begin{frame}{Results}
-    \begin{table}[ht!]
-\centering
-\begin{tabular}[t]{lcccc}
-\toprule
-\multicolumn{1}{c}{ } & \multicolumn{2}{c}{\underline{Weighted Avg. Readmission Rate}} & \multicolumn{2}{c}{\underline{Weighted Avg. Mortality Rate}} \\
- & (1) & (2) & (3) & (4)\\
-\midrule
-Signaling Effect & -0.01 &  & 0.03 & \\
- & (0.07) &  & (0.06) & \\
-Managing Effect &  & 0.17 &  & 0.12\\
- &  & (0.1) &  & (0.09)\\
- &  &  &  & \\
-Observations & 3,005 & 1,595 & 3,005 & 1,595\\
-\bottomrule
-\end{tabular}
-\end{table}
-\end{frame}
-
-\begin{frame}{Summary}
-\large
-    \textbf{We know very little about whether leaders matter in health care}
-
-    \begin{wideitemize}
-        \item This paper: non-clinical executive teams internalize financial incentives more
-    \end{wideitemize}
-
-    \vspace{5mm}
-
-    \textbf{Implications}
-    \begin{itemize}
-        \item Not all nonprofits have the same objectives
-        \item More clinicians in leadership could shift objectives towards quality
-    \end{itemize}
-
-    \vspace{5mm}
-
-    \textbf{Future Work}
-
-    \begin{itemize}
-        \item Effects of mandated gender and racial diversity on hospital behavior
-        \item Networks of executives over time 
-        \item Are leaders correlated with hospital fraudulent behavior?
-    \end{itemize}
-\end{frame}
-
-\begin{transitionframe}
-\centering 
-%\includegraphics[width=.2\textwidth]{Graphics/HannaGlenn_headshot.jpg}
-
-\vspace{5mm}
-
-\Large
-    Thank you!
-
-    \vspace{15mm}
-
-    \textcolor{black}{\faEnvelope \hspace{2mm} hanna.glenn@emory.edu}
-
-    \vspace{2mm}
-
-    \textcolor{black}{\faTwitter \hspace{2mm} @hannalynnglenn}
-\end{transitionframe}
-
-\section{Appendix}
-
-\begin{transitionframe}
-\centering \LARGE
-    \textcolor{black}{Appendix Slides}
-\end{transitionframe}
-
-\appendix
-
-\begin{frame}{Choice of Quality}\label{choiceofqualityapp}
-Take FOC, rearrange and take derivative wrt $\alpha$:
-
-\vspace{5mm}
-
-\begin{equation*}
-    \frac{du'(\theta)}{d\alpha} = \frac{-1}{(1-\alpha)^2}\pi'(\theta)
-\end{equation*}
-
-\vspace{5mm}
-
-$\pi'(\theta)$:
-
-\vspace{3mm}
-
-\begin{wideitemize}
-    \item In fee-for-service setting: negative $\implies \frac{du'(\theta)}{d\alpha} > 0$
-    \item In pay-for-performance setting: positive $\implies \frac{du'(\theta)}{d\alpha} < 0$
-\end{wideitemize}
-
-\end{frame}
-
-\begin{frame}{FFS $\implies$ P4P}
-
-\vspace{3mm{}}
-    \large
-
-        \begin{block}{}
-        \centering \Large
-        Hospitals with higher weight on profit will respond more
-
-        to the financial incentives on quality
-    \end{block}
-
-    \begin{align*}
-        \frac{d\Delta\theta}{d\alpha}&=\frac{d\theta_{p4p}}{d\alpha}-\frac{d\theta_{ffs}}{d\alpha}\\ 
-        &\\
-        &=\frac{\frac{du'(\theta)}{d\alpha}}{\frac{du'(\theta)}{d\theta_{p4p}}} - \frac{\frac{du'(\theta)}{d\alpha}}{\frac{du'(\theta)}{d\theta_{ffs}}}\\
-        &\\
-        &> 0
-    \end{align*}
-
-    \btVFill \normalsize
-    \hyperlink{choiceofquality}{\beamerreturnbutton{hospital choice of quality}}
-\end{frame}
-
-\begin{frame}{How Leaders Affect Quality}\label{executivejob}
-\large \vspace{5mm}
-\textbf{Ways to Improve Quality/ Decrease Readmissions}
-
-\vspace{2mm}
-
-\begin{wideitemize}
-    \item Improving care transition from inpatient to outpatient
-    \item Patient education
-    \item Improve care coordination
-    \item Follow-up appointments
-\end{wideitemize}
-
-\btVFill
-
-\scriptsize (\cite{silow2011reducing}) \hspace{4mm}
-\hyperlink{nonprofithospexec}{\beamerreturnbutton{Nonprofit Hosp Execs}}
-
 \end{frame}
@@
 \begin{center}
-    \includegraphics[width=\textwidth]{}
+    \includegraphics[width=\textwidth]{Objects/outcomes_graph_fp.pdf}
 \end{center}
@@
 
+\begin{frame}{Results by Specialty}\label{specialty}
+\begin{center}
+    \includegraphics[width=.4\textwidth]{Objects/read_nospecialty_nomd_synth_graph.pdf}
+\end{center}
+
+    \btVFill
+
+\hyperlink{main:readrates}{\beamerreturnbutton{main results}}
+\end{frame}
+
+\begin{frame}{Average Team Changes Over Time}\label{changes}
+\centering 
+    \includegraphics[width=.8\textwidth]{Objects/change_means.pdf}
+\end{frame}
+
+\begin{frame}{Changes are not correlated with policy exposure}
+\centering \small
+\begin{equation*}\label{eq:change2}
+    \text{change}_{ht} = \sum_{j=2011}^{2014}\beta_j(\mathbf{1}\{t=j\}\times \text{Program Exposed})_{ht} + \alpha_h + \epsilon_{ht},
+    \end{equation*}
+
+    \btVFill
+
+    \includegraphics[width=.7\textwidth]{Objects/change_analysis_plot.pdf}
+
+    \hyperlink{ident}{\beamerreturnbutton{back}}
+
+\end{frame}
+
+\begin{frame}{Driven by Unobserved Compositional Differences?}\label{unobserved}
+\large
+
+\textbf{Concern:} presence of clinicians may not matter but is correlated with some unobserved hospital `type', possibly time varying
+
+\vspace{3mm}
+\begin{wideitemize}
+    \item if true, we should see a difference in response no matter the timing of having a clinical executive
+\end{wideitemize}
+\end{frame}
+
+\begin{frame}{Robustness Check}
+\large
+    \begin{figure}[ht!]
+\begin{center}
+
+ \begin{tabular}{| m{15em} |}
+ \hline \vspace{1mm}
+ \textbf{Signaling Effect:}\\ [0.5ex]
+ \hline\hline 
+ \vspace{2mm}
+ Treat Group:  \hspace{15mm} Never MD \\
+ \vspace{2mm}
+ Comparison Group: \hspace{3mm} Ever MD  \\
+ [1ex]
+ \hline
+ \end{tabular}
+\hfil   %<---
+ \begin{tabular}{|m{18em}|}
+ \hline \vspace{1mm}
+ \textbf{Clinician Effect:}\\ [0.5ex]
+ \hline\hline
+ \vspace{2mm}
+ Treat Group: \hspace{11mm} MD in 2012 \\
+ \vspace{2mm}
+ Comparison Group:  MD not in 2012  \\
+ [1ex]
+ \hline
+ \end{tabular}
+
+\end{center}
+ \end{figure}
+
+\end{frame}
+
+\begin{frame}{Results}
+    \begin{table}[ht!]
+\centering
+\begin{tabular}[t]{lcccc}
+\toprule
+\multicolumn{1}{c}{ } & \multicolumn{2}{c}{\underline{Weighted Avg. Readmission Rate}} & \multicolumn{2}{c}{\underline{Weighted Avg. Mortality Rate}} \\
+ & (1) & (2) & (3) & (4)\\
+\midrule
+Signaling Effect & -0.01 &  & 0.03 & \\
+ & (0.07) &  & (0.06) & \\
+Clinician Effect &  & 0.17 &  & 0.12\\
+ &  & (0.1) &  & (0.09)\\
+ &  &  &  & \\
+Observations & 3,005 & 1,595 & 3,005 & 1,595\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\hyperlink{forprofmort}{\beamerreturnbutton{back}}
+\end{frame}
+
+\begin{frame}{Leadership Characteristics and Firm Behavior/Performance}
+
+\large
+
+\textbf{Gender}
+\begin{itemize}
+    \item \normalsize Female board members $\implies$ lower firm value, less employment reduction \\ \scriptsize(\cite{ahern2012changing}; \cite{matsa2013female}) \normalsize
+    \item Female executives $\implies$ female wage distribution \scriptsize (\cite{flabbi2019female}) \normalsize
+\end{itemize}
+
+\vspace{4mm}
+
+\textbf{Prior Experience}
+\begin{itemize}
+    \item \normalsize General background CEOs paid more \scriptsize (\cite{custodio2013generalists}) \normalsize
+    \item Background matters for acquisitions \scriptsize(\cite{custodio2013ceos}) \normalsize
+
+\vspace{1mm}
+
+    \item Military experience CEOs \scriptsize (\cite{benmelech2015military})
+\end{itemize}
+
+\vspace{4mm}
+
+\textbf{Age}
+\begin{itemize}
+    \item \normalsize Young male CEOs more aggressive in M\&A \scriptsize (\cite{levi2010deal})
+\end{itemize}
+
+\end{frame}
+
 \end{document}

### 2025-10-30 Slides/JobTalk(30minute).tex
Updates from Overleaf

--- 2024-11-13 Slides/JobTalk.tex
@@
 
-        \item Unclear whether leadership matters when policies change incentives
+        \item Unclear whether leadership matters when policies cause exogenous
+
+        change in incentives
     \end{enumerate}
@@
 
-\vspace{5mm} \large \pause
+\vspace{10mm} \large \pause
 
 \begin{enumerate}
-    \item \textbf{Hospital care makes up a third of health care spending} \scriptsize (\cite{cms2024nhe}) \normalsize
+    \item \textbf{Nonprofits drive the hospital industry} \scriptsize (\cite{ASPE_2023})
 
@@
 
-    \begin{itemize}
-        \item \normalsize Nonprofits drive the hospital industry \scriptsize (\cite{ASPE_2023}) \normalsize
-    \end{itemize}
-
-\vspace{8mm}
-
+    \begin{wideitemize}
+        \item Mixed evidence about nonprofit behavior
+    \end{wideitemize}
+    \end{enumerate}
+\end{frame}
+
+%\begin{frame}{Addressing the Gaps}
+    %\centering
+    %\begin{tikzpicture}
+        % First image (top-left)
+        %\node[anchor=north west] (img1) at (0, 3) {\includegraphics[width=0.75\textwidth]{Graphics/nonprofit_screenshot1.png}};
+        %\draw[line width=.45mm, lightcyan] (img1.north west) rectangle (img1.south east);
+
+        % Second image (slightly down and right of the first image)
+        %\node[anchor=north west] (img2) at (0, 0) {\includegraphics[width=0.75\textwidth]{Graphics/nonprofit_screenshot2.png}};
+        %\draw[line width=.45mm, lightcyan] (img2.north west) rectangle (img2.south east);
+
+    %\end{tikzpicture}
+%\end{frame}
+
+\begin{frame}{Addressing the Gaps}
+
+    \Large
+    \begin{block}{}
+    \begin{center}
+    I examine the role of leadership in nonprofit hospitals
+
+    responding to policy incentives
+    \end{center}
+    \end{block}
+
+\vspace{18mm} \large
+
+\begin{enumerate}
+    \setcounter{enumi}{1}
     \item  \large \textbf{Policies often aim to improve the value of health care delivery}
@@
     \large
-    \begin{block}{\Large Main Takeaway:}
+    \begin{block}{\Large Main Takeaway of P4P:}
         \large 
@@
 
-\begin{frame}{Preview of Main Findings}
-
-\large
-
-\textbf{Non-clinical executive teams respond more to financial incentives on quality}
-
-\vspace{3mm}
-
-\begin{wideitemize}
-    \item Through readmission rates
-    \item Non-clinical executive teams behave like profit-driven hospitals
-\end{wideitemize}
-\end{frame}
-
-\begin{frame}{Contribution}
-\large
-    \textbf{Hospital Executives/Board of Directors}
-
-\vspace{2mm}
-
-    \begin{wideitemize}
-        \item \normalsize Physician board members $\implies$ lower public donations \scriptsize (\cite{brickley2010board}) \normalsize
-        \item No correlation b/w CEOs and hospital productivity in England \scriptsize (\cite{janke2019impact})\normalsize
-        \item Public hospitals in Chile improved when older physicians left CEO \scriptsize (\cite{otero2022managers}) \normalsize
-    \end{wideitemize}
-
-    \vspace{5mm}\pause
-
-    \textbf{Hospital Response to Pay-for-Performance}
-
-\vspace{2mm}
-
-    \begin{wideitemize}
-        \item \normalsize Hospitals decreased readmissions \scriptsize (\cite{gupta2021impacts}) \normalsize
-        \item Market concentration and systems affect response \scriptsize (\cite{kunz2024assessing}) \normalsize
-        \item Rewards had little impact on quality \scriptsize (\cite{norton2018moneyball}; \cite{friedson2019so})
-    \end{wideitemize}
-\end{frame}
-
-\begin{frame}{Rest of Talk}
-
-\large
-
-    Setting and Data Collection
-
-    \vspace{5mm}
-
-    About Executive Teams
-
-    \vspace{5mm}
-
-    Theoretical Motivation
-
-    \vspace{5mm}
-
-    Effect of Clinical Leadership
-
-    \vspace{5mm}
-
-    Comparison to For-Profits
-
-    \vspace{5mm}
-
-    Signaling vs. Managing
-\end{frame}
+% \begin{frame}{Contribution}
+% \large
+%     \textbf{Hospital Executives/Board of Directors}
+
+% \vspace{2mm}
+
+%     \begin{wideitemize}
+%         \normalsize \item Physician board members $\implies$ lower public donations \scriptsize (\cite{brickley2010board}) \normalsize
+%         \normalsize \item No correlation b/w CEOs and hospital productivity in England \scriptsize (\cite{janke2019impact})\normalsize
+%         \normalsize \item Public hospitals in Chile improved when older physicians left CEO \scriptsize (\cite{otero2022managers}) \normalsize
+%     \end{wideitemize}
+
+%     \vspace{5mm}\pause
+
+%     \textbf{Hospital Response to Pay-for-Performance}
+
+% \vspace{2mm}
+
+%     \begin{wideitemize}
+%         \normalsize \item Hospitals decreased readmissions \scriptsize (\cite{gupta2021impacts})
+
+%         \vspace{-.5mm}
+
+%         \normalsize \item Market concentration and systems affect response 
+%         \scriptsize (\cite{kunz2024assessing})
+
+%         \normalsize\item Rewards had little impact on quality \scriptsize (\cite{norton2018moneyball}; \cite{friedson2019so})
+%     \end{wideitemize}
+% \end{frame}
 
@@
 
-\vspace{4mm} \pause
+\vspace{3mm} \pause
 
@@
 
-\hyperlink{executivejob}{\beamergotobutton{executive job}}
+\hyperlink{executivejob}{\beamergotobutton{improving quality}}
 
@@
 
-\begin{frame}{Data Collection Process}
-\begin{center}
-\begin{tikzpicture}[
-    node distance=3.1cm and 1.8cm,
-    every node/.style={align=center, font=\normalsize},
-    process/.style={rectangle, draw, rounded corners, text centered, minimum width=2.5cm, minimum height=2cm, fill=purple!30},
-    output/.style={rectangle, draw, text centered, minimum width=2.7cm, minimum height=2.5cm, fill=purple},
-    extract/.style={rectangle, draw, rounded corners, text centered, minimum width=2.5cm, minimum height=2cm, fill=sage!30},
-    arrow/.style={thick, ->, >=stealth}
-]
-
-% Nodes
-\node (step1) [process] {Hospital \\EINs};\pause
-\node (step2) [process, right of=step1] {Download \\PDFs};
-\draw[arrow] (step1) -- (step2); \pause
-\node (step3) [process, right of=step2] {Extract\\text};
-\draw[arrow] (step2) -- (step3); 
-% New green box below step3
-\node (extract) [extract, below of=step3, yshift=0.5cm] {Optical\\Character\\Recognition}; 
-\draw[arrow] (extract) -- (step3);\pause
-\node (step4) [process, right of=step3] {String\\cleaning\\algorithms};
-\draw[arrow] (step3) -- (step4); \pause
-\node (output) [output, right of=step4, text=white] {Executive\\Name\\Position\\Title};
-\draw[arrow] (step4) -- (output);
-\end{tikzpicture}
-\end{center}
-\end{frame}
-
 \begin{frame}{Merge to Other Data}
@@
 
-    \vspace{2mm}
+    \vspace{1mm}
     \begin{wideitemize}
@@
 
-    \vspace{7mm}
+    \vspace{4mm}
 
@@
     \end{itemize}
+
+    \vspace{4mm}
+
+    \textbf{Years: 2009-2015}
 \end{frame}
@@
 \end{transitionframe}
+
+\begin{frame}{Sample Definitions}
+\centering
+    \includegraphics[scale=.75]{Graphics/2010.png}
+\end{frame}
 
@@
     \centering
+
+    \vspace{2mm}
+
     \begin{equation*}
@@
     \end{equation*}
+
 \end{frame}
@@
 
-\begin{frame}{Choice of Quality}\label{choiceofquality}
+\begin{frame}{Choice of Quality ($\theta$)}\label{choiceofquality}
 
@@
 \large
-    \begin{enumerate}
-        \item If clinicians affect $\alpha$, they affect quality change after P4P
-
-        \vspace{8mm}\pause
-        \item If non-clinicians have higher $\alpha$, they act more like for-profits
-
-        \vspace{8mm}\pause
-        \item Mechanism could be revealing objectives or changing objectives
-    \end{enumerate}
+
+        \textbf{Response to P4P is tied to underlying objectives $\alpha$}
+        \vspace{3mm}
+        \begin{wideitemize}
+            \item Ownership status is important to underlying objectives
+            \item It's reasonable that priorities of leaders are also an important factor
+        \end{wideitemize}
+
 \end{frame}
@@
         \item No weights
-        \item \textbf{``Synthetic DiD" type weights} 
+        \item \textbf{Weights based on outcomes pre-2012} 
         \item Coarsened exact matching 
@@
 
-\begin{frame}{Identification Assumptions}
+\begin{frame}{Identification Assumptions}\label{ident}
 
@@
     \item No hospital characteristics perfectly correlated with executive team
-    \item No anticipation of incentive changes
-    \item Stability in clinical executives not correlated with incentive change
-\end{wideitemize}
-\end{frame}
-
-\begin{frame}{Average Team Changes Over Time}
-\centering 
-    \includegraphics[width=.8\textwidth]{Objects/change_means.pdf}
-\end{frame}
-
-\begin{frame}{Changes are not correlated with policy exposure}
-\centering \small
-\begin{equation*}\label{eq:change2}
-    \text{change}_{ht} = \sum_{j=2011}^{2014}\beta_j(\mathbf{1}\{t=j\}\times \text{Program Exposed})_{ht} + \alpha_h + \epsilon_{ht},
-    \end{equation*}
-
-    \btVFill
-
-    \includegraphics[width=.7\textwidth]{Objects/change_analysis_plot.pdf}
-
+    \item Stability in clinical executives not correlated with incentive change \hyperlink{changes}{\beamergotobutton{analysis}}
+\end{wideitemize}
 \end{frame}
@@
 \large
-    \textbf{Compare both nonprofit leadership team types to the response of for-profits:}
+    \textbf{Compare both nonprofit leadership team types to for-profits:}
 
@@
 
-    Comparison group:
+    Comparison groups:
     \begin{itemize}
@@
 
-\begin{frame}{Mortality Rate}
+\begin{frame}{Mortality Rate}\label{forprofmort}
     \begin{columns}
-    \column{0.5\textwidth}
+    \column{0.45\textwidth}
         \centering
         \includegraphics[width=.9\textwidth]{Objects/fp_mort_md_synth_graph.pdf}
-    \column{0.5\textwidth}
+    \column{0.45\textwidth}
         \centering
@@
     \end{columns}
-\end{frame}
-
-\section{Signaling vs. Managing}
-
-\begin{transitionframe}
-\centering \LARGE
-    \textcolor{black}{Signaling vs. Managing}
-\end{transitionframe}
-
-\begin{frame}{Hypotheses}
-
-\large
-    \textbf{Clinical leaders represent underlying objectives ($\alpha$)}
-
-\vspace{2mm}
-
-    \begin{wideitemize}
-        \item Timing of the clinical leader doesn't matter 
-    \end{wideitemize}
-
-    \vspace{15mm}
-
-    \textbf{Clinical leaders effectively change objectives ($\alpha$)}
-
-\vspace{2mm}
-
-    \begin{wideitemize}
-        \item Timing of the clinical leader does matter
-    \end{wideitemize}
-\end{frame}
-
-\begin{frame}{Estimation}
-\large
-    \begin{figure}[ht!]
-\begin{center}
-
- \begin{tabular}{| m{15em} |}
- \hline \vspace{1mm}
- \textbf{Signaling:}\\ [0.5ex]
- \hline\hline 
- \vspace{2mm}
- Treat Group:  \hspace{15mm} Never MD \\
- \vspace{2mm}
- Comparison Group: \hspace{3mm} Ever MD  \\
- [1ex]
- \hline
- \end{tabular}
-\hfil   %<---
- \begin{tabular}{|m{18em}|}
- \hline \vspace{1mm}
- \textbf{Managing:}\\ [0.5ex]
- \hline\hline
- \vspace{2mm}
- Treat Group: \hspace{11mm} MD in 2012 \\
- \vspace{2mm}
- Comparison Group:  MD not in 2012  \\
- [1ex]
- \hline
- \end{tabular}
-
-\end{center}
- \end{figure}
-
-\end{frame}
-
-\begin{frame}{Results}
-    \begin{table}[ht!]
-\centering
-\begin{tabular}[t]{lcccc}
-\toprule
-\multicolumn{1}{c}{ } & \multicolumn{2}{c}{\underline{Weighted Avg. Readmission Rate}} & \multicolumn{2}{c}{\underline{Weighted Avg. Mortality Rate}} \\
- & (1) & (2) & (3) & (4)\\
-\midrule
-Signaling Effect & -0.01 &  & 0.03 & \\
- & (0.07) &  & (0.06) & \\
-Managing Effect &  & 0.17 &  & 0.12\\
- &  & (0.1) &  & (0.09)\\
- &  &  &  & \\
-Observations & 3,005 & 1,595 & 3,005 & 1,595\\
-\bottomrule
-\end{tabular}
-\end{table}
-\end{frame}
+
+    \hyperlink{unobserved}{\beamergotobutton{unobserved}}
+\end{frame}
+
+\section{Unobserved}
 
@@
     \begin{itemize}
-        \item Not all nonprofits have the same objectives
-        \item More clinicians in leadership could shift objectives towards quality
+        \item Despite all being nonprofits, hospitals with clinician leaders act more like we would expect nonprofits to
+        \item In the absence of clinical leadership, nonprofits just act like for-profits
     \end{itemize}
@@
 \begin{wideitemize}
-    \item Improving care transition from inpatient to outpatient
+    \item Care transition from inpatient to outpatient
     \item Patient education
@@
 
+\begin{frame}{Results by Specialty}\label{specialty}
+\begin{center}
+    \includegraphics[width=.4\textwidth]{Objects/read_nospecialty_nomd_synth_graph.pdf}
+\end{center}
+
+    \btVFill
+
+\hyperlink{main:readrates}{\beamerreturnbutton{main results}}
+\end{frame}
+
+\begin{frame}{Average Team Changes Over Time}\label{changes}
+\centering 
+    \includegraphics[width=.8\textwidth]{Objects/change_means.pdf}
+\end{frame}
+
+\begin{frame}{Changes are not correlated with policy exposure}
+\centering \small
+\begin{equation*}\label{eq:change2}
+    \text{change}_{ht} = \sum_{j=2011}^{2014}\beta_j(\mathbf{1}\{t=j\}\times \text{Program Exposed})_{ht} + \alpha_h + \epsilon_{ht},
+    \end{equation*}
+
+    \btVFill
+
+    \includegraphics[width=.7\textwidth]{Objects/change_analysis_plot.pdf}
+
+    \hyperlink{ident}{\beamerreturnbutton{back}}
+
+\end{frame}
+
+\begin{frame}{Driven by Unobserved Compositional Differences?}\label{unobserved}
+\large
+
+\textbf{Concern:} presence of clinicians may not matter but is correlated with some unobserved hospital `type', possibly time varying
+
+\vspace{3mm}
+\begin{wideitemize}
+    \item if true, we should see a difference in response no matter the timing of having a clinical executive
+\end{wideitemize}
+\end{frame}
+
+\begin{frame}{Robustness Check}
+\large
+    \begin{figure}[ht!]
+\begin{center}
+
+ \begin{tabular}{| m{15em} |}
+ \hline \vspace{1mm}
+ \textbf{Signaling Effect:}\\ [0.5ex]
+ \hline\hline 
+ \vspace{2mm}
+ Treat Group:  \hspace{15mm} Never MD \\
+ \vspace{2mm}
+ Comparison Group: \hspace{3mm} Ever MD  \\
+ [1ex]
+ \hline
+ \end{tabular}
+\hfil   %<---
+ \begin{tabular}{|m{18em}|}
+ \hline \vspace{1mm}
+ \textbf{Clinician Effect:}\\ [0.5ex]
+ \hline\hline
+ \vspace{2mm}
+ Treat Group: \hspace{11mm} MD in 2012 \\
+ \vspace{2mm}
+ Comparison Group:  MD not in 2012  \\
+ [1ex]
+ \hline
+ \end{tabular}
+
+\end{center}
+ \end{figure}
+
+\end{frame}
+
+\begin{frame}{Results}
+    \begin{table}[ht!]
+\centering
+\begin{tabular}[t]{lcccc}
+\toprule
+\multicolumn{1}{c}{ } & \multicolumn{2}{c}{\underline{Weighted Avg. Readmission Rate}} & \multicolumn{2}{c}{\underline{Weighted Avg. Mortality Rate}} \\
+ & (1) & (2) & (3) & (4)\\
+\midrule
+Signaling Effect & -0.01 &  & 0.03 & \\
+ & (0.07) &  & (0.06) & \\
+Clinician Effect &  & 0.17 &  & 0.12\\
+ &  & (0.1) &  & (0.09)\\
+ &  &  &  & \\
+Observations & 3,005 & 1,595 & 3,005 & 1,595\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\hyperlink{forprofmort}{\beamerreturnbutton{back}}
+\end{frame}
+
 \end{document}

### 2025-10-30 Slides/JobTalk(40minute).tex
Updates from Overleaf

--- 2025-10-30 Slides/JobTalk(30minute).tex
@@
     \begin{wideitemize}
-        \item Mixed evidence about nonprofit behavior
+        \item \large Mixed evidence on nonprofit hospital behavior
     \end{wideitemize}
@@
     \begin{wideitemize}
-        \item \normalsize Unclear how leaders interact with this goal
+        \item \large Unclear how leaders interact with this goal
     \end{wideitemize}
@@
 
-% \begin{frame}{Contribution}
-% \large
-%     \textbf{Hospital Executives/Board of Directors}
-
-% \vspace{2mm}
-
-%     \begin{wideitemize}
-%         \normalsize \item Physician board members $\implies$ lower public donations \scriptsize (\cite{brickley2010board}) \normalsize
-%         \normalsize \item No correlation b/w CEOs and hospital productivity in England \scriptsize (\cite{janke2019impact})\normalsize
-%         \normalsize \item Public hospitals in Chile improved when older physicians left CEO \scriptsize (\cite{otero2022managers}) \normalsize
-%     \end{wideitemize}
-
-%     \vspace{5mm}\pause
-
-%     \textbf{Hospital Response to Pay-for-Performance}
-
-% \vspace{2mm}
-
-%     \begin{wideitemize}
-%         \normalsize \item Hospitals decreased readmissions \scriptsize (\cite{gupta2021impacts})
-
-%         \vspace{-.5mm}
-
-%         \normalsize \item Market concentration and systems affect response 
-%         \scriptsize (\cite{kunz2024assessing})
-
-%         \normalsize\item Rewards had little impact on quality \scriptsize (\cite{norton2018moneyball}; \cite{friedson2019so})
-%     \end{wideitemize}
-% \end{frame}
+\begin{frame}{Preview of Main Findings}
+
+\large
+
+\textbf{Non-clinical executive teams respond more to financial incentives on quality}
+
+\vspace{3mm}
+
+\begin{wideitemize}
+    \item Through readmission rates
+    \item Non-clinical executive teams behave like profit-driven hospitals
+\end{wideitemize}
+\end{frame}
+
+\begin{frame}{Contribution}
+\large
+    \textbf{Hospital Executives/Board of Directors}
+
+\vspace{2mm}
+
+    \begin{wideitemize}
+        \normalsize \item Physician board members $\implies$ lower public donations \scriptsize (\cite{brickley2010board}) \normalsize
+        \normalsize \item No correlation b/w CEOs and hospital productivity in England \scriptsize (\cite{janke2019impact})\normalsize
+        \normalsize \item Public hospitals in Chile improved when older physicians left CEO \scriptsize (\cite{otero2022managers}) \normalsize
+    \end{wideitemize}
+
+    \vspace{5mm}\pause
+
+    \textbf{Hospital Response to Pay-for-Performance}
+
+\vspace{2mm}
+
+    \begin{wideitemize}
+        \normalsize \item Hospitals decreased readmissions \scriptsize (\cite{gupta2021impacts})
+
+        \vspace{-.5mm}
+
+        \normalsize \item Market concentration and systems affect response 
+        \scriptsize (\cite{kunz2024assessing})
+
+        \normalsize\item Rewards had little impact on quality \scriptsize (\cite{norton2018moneyball}; \cite{friedson2019so})
+    \end{wideitemize}
+\end{frame}
 
@@
 
-    \vspace{1mm}
+    \vspace{2mm}
     \begin{wideitemize}
@@
 
-    \vspace{4mm}
+    \vspace{7mm}
 
@@
 
-    \vspace{1mm}
+    \vspace{2mm}
 
@@
     \end{itemize}
-
-    \vspace{4mm}
-
-    \textbf{Years: 2009-2015}
 \end{frame}
@@
 \end{transitionframe}
-
-\begin{frame}{Sample Definitions}
-\centering
-    \includegraphics[scale=.75]{Graphics/2010.png}
-\end{frame}
 
@@
 Fraction Clinical Execs & 0.20 & 0.26 & 0.00\\
+Fraction Int. Medicine Execs & 0.06 & 0.08 & 0.00\\
 Has a CMO & 0.55 & 0.52 & 0.07\\
@@
 \hspace{1em}Penalized for one condition & 0.19 & 0.25 & 0.13\\
-\hspace{1em}Penalized for two conditions & 0.50 & 0.47 & 0.38\\
-\hspace{1em}Penalized for all Conditions & 0.31 & 0.31 & 0.20\\
+\hspace{1em}Penalized for AMI + HF & 0.07 & 0.05 & 0.03\\
+\hspace{1em}Penalized for AMI + Pneumonia & 0.02 & 0.02 & 0.02\\
+\hspace{1em}Penalized for HF + Pneumonia & 0.41 & 0.40 & 0.33\\
+\hspace{1em}Penalized for All Conditions & 0.31 & 0.31 & 0.20\\
 \bottomrule
@@
         \item No weights
-        \item \textbf{Weights based on outcomes pre-2012} 
+        \item \textbf{``Synthetic DiD" type weights} 
         \item Coarsened exact matching 
@@
 
-\begin{frame}{Identification Assumptions}\label{ident}
+\begin{frame}{Identification Assumptions}
 
@@
     \item No hospital characteristics perfectly correlated with executive team
-    \item Stability in clinical executives not correlated with incentive change \hyperlink{changes}{\beamergotobutton{analysis}}
-\end{wideitemize}
+    \item No anticipation of incentive changes
+    \item Stability in clinical executives not correlated with incentive change
+\end{wideitemize}
+\end{frame}
+
+\begin{frame}{Average Team Changes Over Time}
+\centering 
+    \includegraphics[width=.8\textwidth]{Objects/change_means.pdf}
+\end{frame}
+
+\begin{frame}{Changes are not correlated with policy exposure}
+\centering \small
+\begin{equation*}\label{eq:change2}
+    \text{change}_{ht} = \sum_{j=2011}^{2014}\beta_j(\mathbf{1}\{t=j\}\times \text{Program Exposed})_{ht} + \alpha_h + \epsilon_{ht},
+    \end{equation*}
+
+    \btVFill
+
+    \includegraphics[width=.7\textwidth]{Objects/change_analysis_plot.pdf}
+
 \end{frame}
@@
 
-\begin{frame}{Mortality Rate}\label{forprofmort}
+\begin{frame}{Mortality Rate}
     \begin{columns}
-    \column{0.45\textwidth}
+    \column{0.5\textwidth}
         \centering
         \includegraphics[width=.9\textwidth]{Objects/fp_mort_md_synth_graph.pdf}
-    \column{0.45\textwidth}
+    \column{0.5\textwidth}
         \centering
@@
     \end{columns}
-
-    \hyperlink{unobserved}{\beamergotobutton{unobserved}}
 \end{frame}
@@
 \section{Unobserved}
+
+\begin{transitionframe}
+\centering \LARGE
+    \textcolor{black}{Driven by Unobserved Compositional Differences?}
+\end{transitionframe}
+
+\begin{frame}{Driven by Unobserved Compositional Differences?}
+\large
+
+\textbf{Concern:} presence of clinicians may not matter but is correlated with some unobserved hospital `type', possibly time varying
+
+\vspace{3mm}
+\begin{wideitemize}
+    \item if true, difference in response should hold when timing of clinician changes
+\end{wideitemize}
+\end{frame}
+
+\begin{frame}{Driven by Unobserved Compositional Differences?}
+\large
+    \begin{figure}[ht!]
+\begin{center}
+
+ \begin{tabular}{| m{15em} |}
+ \hline \vspace{1mm}
+ \textbf{Signaling Effect:}\\ [0.5ex]
+ \hline\hline 
+ \vspace{2mm}
+ Treat Group:  \hspace{15mm} Never MD \\
+ \vspace{2mm}
+ Comparison Group: \hspace{3mm} Ever MD  \\
+ [1ex]
+ \hline
+ \end{tabular}
+\hfil   %<---
+ \begin{tabular}{|m{18em}|}
+ \hline \vspace{1mm}
+ \textbf{Clinician Effect:}\\ [0.5ex]
+ \hline\hline
+ \vspace{2mm}
+ Treat Group: \hspace{11mm} MD in 2012 \\
+ \vspace{2mm}
+ Comparison Group:  MD not in 2012  \\
+ [1ex]
+ \hline
+ \end{tabular}
+
+\end{center}
+ \end{figure}
+
+\end{frame}
+
+\begin{frame}{Results}
+    \begin{table}[ht!]
+\centering
+\begin{tabular}[t]{lcccc}
+\toprule
+\multicolumn{1}{c}{ } & \multicolumn{2}{c}{\underline{Weighted Avg. Readmission Rate}} & \multicolumn{2}{c}{\underline{Weighted Avg. Mortality Rate}} \\
+ & (1) & (2) & (3) & (4)\\
+\midrule
+Signaling Effect & -0.01 &  & 0.03 & \\
+ & (0.07) &  & (0.06) & \\
+Clinician Effect &  & 0.17 &  & 0.12\\
+ &  & (0.1) &  & (0.09)\\
+ &  &  &  & \\
+Observations & 3,005 & 1,595 & 3,005 & 1,595\\
+\bottomrule
+\end{tabular}
+\end{table}
+\end{frame}
 
@@
 
-\begin{frame}{Average Team Changes Over Time}\label{changes}
-\centering 
-    \includegraphics[width=.8\textwidth]{Objects/change_means.pdf}
-\end{frame}
-
-\begin{frame}{Changes are not correlated with policy exposure}
-\centering \small
-\begin{equation*}\label{eq:change2}
-    \text{change}_{ht} = \sum_{j=2011}^{2014}\beta_j(\mathbf{1}\{t=j\}\times \text{Program Exposed})_{ht} + \alpha_h + \epsilon_{ht},
-    \end{equation*}
-
-    \btVFill
-
-    \includegraphics[width=.7\textwidth]{Objects/change_analysis_plot.pdf}
-
-    \hyperlink{ident}{\beamerreturnbutton{back}}
-
-\end{frame}
-
-\begin{frame}{Driven by Unobserved Compositional Differences?}\label{unobserved}
-\large
-
-\textbf{Concern:} presence of clinicians may not matter but is correlated with some unobserved hospital `type', possibly time varying
-
-\vspace{3mm}
-\begin{wideitemize}
-    \item if true, we should see a difference in response no matter the timing of having a clinical executive
-\end{wideitemize}
-\end{frame}
-
-\begin{frame}{Robustness Check}
-\large
-    \begin{figure}[ht!]
-\begin{center}
-
- \begin{tabular}{| m{15em} |}
- \hline \vspace{1mm}
- \textbf{Signaling Effect:}\\ [0.5ex]
- \hline\hline 
- \vspace{2mm}
- Treat Group:  \hspace{15mm} Never MD \\
- \vspace{2mm}
- Comparison Group: \hspace{3mm} Ever MD  \\
- [1ex]
- \hline
- \end{tabular}
-\hfil   %<---
- \begin{tabular}{|m{18em}|}
- \hline \vspace{1mm}
- \textbf{Clinician Effect:}\\ [0.5ex]
- \hline\hline
- \vspace{2mm}
- Treat Group: \hspace{11mm} MD in 2012 \\
- \vspace{2mm}
- Comparison Group:  MD not in 2012  \\
- [1ex]
- \hline
- \end{tabular}
-
-\end{center}
- \end{figure}
-
-\end{frame}
-
-\begin{frame}{Results}
-    \begin{table}[ht!]
-\centering
-\begin{tabular}[t]{lcccc}
-\toprule
-\multicolumn{1}{c}{ } & \multicolumn{2}{c}{\underline{Weighted Avg. Readmission Rate}} & \multicolumn{2}{c}{\underline{Weighted Avg. Mortality Rate}} \\
- & (1) & (2) & (3) & (4)\\
-\midrule
-Signaling Effect & -0.01 &  & 0.03 & \\
- & (0.07) &  & (0.06) & \\
-Clinician Effect &  & 0.17 &  & 0.12\\
- &  & (0.1) &  & (0.09)\\
- &  &  &  & \\
-Observations & 3,005 & 1,595 & 3,005 & 1,595\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-\hyperlink{forprofmort}{\beamerreturnbutton{back}}
-\end{frame}
-
 \end{document}

### 2025-11-04 Paper/paperdraft_fall2025.tex
Updates from Overleaf

@@
 
-    In this paper, I study the effect of specialized executive training on firm behavior when the firm experiences an external shock to their objective function.
-I do this specifically in the context of US hospitals for two reasons.
+    In this paper, I study the effect of specialized executive training on firm behavior when the firm experiences an external shock to their objective function, specifically in the context of US hospitals.
+I leverage this setting for two reasons.
 First, in the US, health care makes up almost 20\% of spending (\cite{CMSNationalHealthExpenditure}), and a major goal of policymakers is to improve the quality of health care without increasing costs, as the US preforms poorly in these metrics compared to other similar countries.
@@
 Therefore, factors that affect weight $\alpha$ will be correlated with choice of quality, and with how quality changes under new pay-for-performance incentives.
-In this paper, I test the hypothesis that clinical experience on a hospital's leadership team is a factor that is correlated with weight $\alpha$.
+In this paper, I test the hypothesis that clinical experience on a hospital's leadership team is a factor that is correlated with weight $\alpha$, and thus affects response to pay-for-performance incentives.
 
@@
 
-    I use several data sets to analyze the relationship between executives with clinical training and hospital response to financial incentives on quality.
-I construct novel information on individual hospital executives by scraping publicly available tax forms, and merge this data to publicly available data sets containing other hospital characteristics: the American Hospital Association Survey (AHA), the Centers for Medicare and Medicaid Services (CMS) Hospital Compare data and Case Mix Index files, and Healthcare Cost Report Information System (HCRIS) data.
+    I construct novel information on individual hospital executives by scraping publicly available tax forms, and merge this to publicly available data sets containing other hospital characteristics: the American Hospital Association Survey (AHA), the Centers for Medicare and Medicaid Services (CMS) Hospital Compare data and Case Mix Index files, and Healthcare Cost Report Information System (HCRIS) data.
 Ultimately, I employ this hospital-level data spanning from 2010-2014, which contains executive team characteristics, mortality and readmission rates as measures of quality, and other characteristics.
 In this section, I describe the institutional setting and data collection process.
-A more detailed description of the construction of this data can be found in Appendix \ref{appendixdata}.
+I include more details on the data construction in \ref{appendixdata}.
 
@@
 
-    Nonprofit hospitals, as with most nonprofits, are typically governed by a board of directors, whose role is to set broad goals and strategies, and provide general oversight.
+    Nonprofit hospitals, as with most nonprofits, are typically governed by a board of directors, whose role is complex, but generally is to set broad goals and provide oversight \cite{council_board_roles_2025}.
 The board selects executives, the highest level of management of a firm, to carry out day-to-day operations.
@@
 I extract text spanning 2009-2015, but I retain a much larger sample of hospitals by limiting to hospitals that have text information in 2010-2014.
-This yields 852 nonprofit hospitals that are matched to AHA data, and with complete text information in each year.
-
-    Finally, I verify clinical training of those who are identified as doctors, and obtain additional information about clinically trained executives by matching names to the National Plan and Provider Enumeration System (NPPES) database of all registered physicians.
+This yields 852 nonprofit hospitals that are (1) matched to AHA data, and (2) contain complete leadership information in each year from 2010-2014.
+
+    Clinical training need not be limited to physicians.
+However, those are the most common and verifiable clinicians in the data.
+For the main analysis, I verify clinical training of those who are identified as doctors, and obtain additional information about clinically trained executives by matching names to the National Plan and Provider Enumeration System (NPPES) database of all registered physicians.
 I present a table of means for various characteristics of doctor executives in the data in Table \ref{doc_sumstats}.
@@
 The most common specialty is internal medicine.
+While I cannot perfectly verify nurse experience in the data, in Appendix \ref{app:nurses}, I include executives who are marked as registered nurses in the tax forms.
+The results are similar to those of the main analysis, which is not surprising given that only 0.5\% of executives have a nursing title.
 
@@
 
-    I focus on executive teams as a whole instead of specific positions because the organization of teams varies so drastically across hospitals.
-It is not uncommon for two hospitals to have the same position but the tasks performed are drastically different, or to have two distinct positions that essentially perform the same tasks.
-I create various hospital-level characterizations based on their executive team, such as the existence of a clinically trained executive, the number of clinically trained executives, the number of total executives, and whether the hospital has a Chief Medical Officer.
-
-    \subsection{Pay-for-Performance Policies Targeting Readmission and Mortality Rates}\label{sec:hrrp}
-
-    The purpose of this paper is to investigate hospital behavior in the midst of financial incentives that change how hospitals receive payment for quality of care.
+    As the organization of specific c-suite positions can vary drastically across hospitals, I focus on executive teams as a whole instead of specific positions.
+It is not uncommon for two hospitals to have the same position but the tasks performed be drastically different, or to have two distinct positions that essentially perform the same tasks.
+I utilize various hospital-level characterizations based on their executive team, such as the existence of a clinically trained executive, the number of clinically trained executives, the number of total executives, and whether the hospital has a Chief Medical Officer.
+
+    \subsection{Policies Targeting Readmission and Mortality Rates}\label{sec:hrrp}
+
+    The purpose of this paper is to investigate hospital behavior in the midst of financial incentives that change how hospitals receive payment depending on quality.
 Two programs were passed as part of the Affordable Care Act that focused on pay-for-performance incentives for hospitals: the Hospital Readmissions Reduction Program (HRRP), and the Hospital Value Based Purchasing Program (HVBP).
@@
     In Table \ref{tab:sumstats_samples_stable}, I present means of relevant variables for several hospital samples used throughout the paper.
-First, in column (1), are means for hospitals who ever have a clinical executive at any point in the sample.
-In column (2) are means for hospitals who have a clinical executive for the entire sample period, and column (3) shows means for hospitals who never have a clinical executive in the sample period.
+First, in column (1), are means for hospitals who have a clinical executive at any point in time.
+Column (2) contains means for hospitals who have a clinical executive for the entire sample period, and column (3) shows means for hospitals who never have a clinical executive in the sample period.
 The main analysis compares hospitals in columns (2) and (3), and supplemental analyses leverage hospitals in column (1).
@@
 When considering mortality, there is not a drastic change in 2012 for any hospital type, but hospitals who always have a clinical executive continue on an increasing trend in mortality where the other hospitals types remain more flat.
-Finally, hospitals that never have clinical executives consistently have less complex patients, but there is no observable change in raw case mix index over time.
-
-    \begin{figure}[ht!]
-    \centering
-        \caption{Outcomes Over Time}
+
+    \begin{figure}[ht!]
+    \centering
+        \caption{Main Outcomes Over Time}
         \includegraphics[width=.8\textwidth]{Objects/outcomes_graph.pdf}
@@
     I test empirically whether having clinically trained executives matters in how hospitals respond to the pay-for-performance policies effective in 2012.
-The quality metrics targeted by the programs are readmission and mortality rates, so I use these as outcomes.
-I limit the sample to hospitals who always have clinical executives or never have clinical executives.
+The quality metrics targeted by the programs are readmission and mortality rates, so I focus on these measures as outcomes.
+To first analyze a clear control and treatment group, I limit the sample to hospitals who always have clinical executives or never have clinical executives.
 That is, the hospitals in this analysis do not change their propensity to hire a clinical executive during the sample period.\footnote{An important assumption for this analysis is that this sample selection is not correlated with the program enactment.
-I investigate this assumption in Section \ref{sec:changes}, and find no evidence that hospitals are more or less likely to change their propensity to hire a clinical executive because of the pay-for-performance programs.
-}
-
-    I ultimately employ a synthetic difference-in-differences estimation strategy that mirrors two-way-fixed effects, where the key independent variable is the interaction between having clinically trained executives and being in years after pay-for-performance took effect.
+I investigate this assumption in Section \ref{sec:changes}, and find no evidence that hospitals are more or less likely to change their propensity to hire a clinical executive because of the pay-for-performance programs.} I employ a two-way-fixed-effects strategy as follows:
+
+    \begin{equation}
+    \label{eq:clinical}
+    y_{ht} = \beta \text{ (clinical training x post 2012)}_{ht} + \gamma_{h} + \delta_t + \epsilon_{ht},
+    \end{equation}
+
+    \noindent where the variable $y_{ht}$ is one of the outcome variables capturing readmission or mortality rates discussed in Section \ref{sec:hrrp}, and $\gamma_h$ and $\delta_t$ are hospital and time fixed effects, respectively.
+The coefficient of interest is $\beta$, capturing the effect of having a clinician on change in quality after 2012, when pay-for-performance policies were rolled out.
 All hospitals had incentive to improve quality in response to the policy change, as penalties and payments are determined by comparing quality among hospitals.
@@
     One may be concerned about potential differences in quality trends between clinical and non-clinical executive team hospitals leading to a violation in the parallel trends assumption.
-This is why I implement a synthetic difference-in-differences estimation strategy (\cite{arkhangelsky2021synthetic}).
-In particular, I weight more heavily any non-clinical team hospitals with pre-trends similar to clinical team hospitals in readmission, mortality, and baseline case mix.
-Further, I more heavily weight time periods that balance pre- and post-program outcomes for non-clinical teams.
-The control group consists of approximately 240 hospitals, and I show the fairly uniform distribution of their weights in Appendix \ref{app:sumstats}.
-
-    I present the unweighted two-way-fixed-effects specification equation for clarity, and the identification assumptions are conceptually similar for synthetic difference-in-differences.
-In Appendix \ref{app:otherestimations}, I show that the findings are robust to using two way fixed effects, matching, or synthetic control.
-
-    \begin{equation}
-    \label{eq:clinical}
-    y_{ht} = \beta \text{ (clinical training x post 2012)}_{ht} + \gamma_{h} + \delta_t + \epsilon_{ht}
-    \end{equation}
-
-    \noindent The variable $y_{ht}$ is one of the outcome variables capturing readmission or mortality rates discussed in Section \ref{sec:hrrp}, and $\gamma_h$ and $\delta_t$ are hospital and time fixed effects, respectively.
-
-    Once incorporating the weights used in the synthetic differences-in-differences method, we ensure parallel pre-trends in the outcome variables.
-Thus, I assume that under the weighted clinical and non-clinical hospital composition, trends would have continued to be parallel absent the incentive change in 2012.
+To account for this, I ultimately implement a synthetic difference-in-differences strategy (\cite{arkhangelsky2021synthetic}), which weights more heavily any non-clinical team hospitals with pre-trends similar to clinical team hospitals in the outcome being considered, essentially guaranteeing no differential pre-trends between always and never clinical teams.
+Using this strategy, the never-clinical group consists of approximately 240 hospitals, and I show the fairly uniform distribution of their weights in Appendix \ref{app:sumstats}.
+As a robustness check, I also pre-select the sample of always and never clinical hospitals using the Coarsened Exact Matching (CEM) procedure as described by \citeauthor{azoulay2010superstar} (\citeyear{azoulay2010superstar}).
+I match based on observable characteristics such as hospital size, number of patients in each condition, and eventual penalty status.
+For brevity, I only present the synthetic difference-in-differences results here.
+In Appendix \ref{app:otherestimations}, I show that the results are robust to using standard two-way-fixed-effects, matching, or synthetic control.
+
+    For identification, I assume that under the weighted clinical and non-clinical hospital composition, trends would have continued to be parallel absent the incentive change in 2012.
 This includes an implicit assumption that no other event occurred in 2012 correlated with both the outcomes and composition of leadership team.
@@
 Panel \ref{fig:read_synth_clinical} shows the estimated difference in readmission rates, and reveals that non-clinical teams decrease readmissions by .3 (standard error .13) more than clinical teams.
-This is a relatively small magnitude given an average readmission rate of 21, but reveals a meaningful difference in behavior given that both hospital types significantly decrease readmissions.
-In Appendix \ref{app:condition}, I show that these results are primarily driven by pneumonia patients.
-
-    In panel \ref{fig:mort_synth_clinical}, I show the estimated difference in mortality.
+This is a relatively small magnitude given an average readmission rate of 21.
+However, \citeauthor{gupta2021impacts} (\citeyear{gupta2021impacts}) finds that hospitals overall decrease readmissions by around 3 percentage points due to the program, so this reveals a meaningful difference in behavior relative to the overall behavior change.
+In panel \ref{fig:mort_synth_clinical}, I show the estimated difference in mortality for always clinical and never clinical leadership teams after 2012.
 While the magnitude of the difference is similar to that for readmission rates (relative to the mean), mortality is a noisy measure, and I cannot rule out null effects.
-I investigate whether the estimated difference in readmission rate response could be due to differential participation in selective patient practices in Appendix \ref{app:casemix}, and find that this is not the case.
 
@@
 I find that hospitals without clinically trained executives decrease readmissions by .6 (standard error .14) more than hospitals with clinically trained executives.
-Thus, the readmission rate response is driven by having a higher proportion of clinical executives relative to no clinical executives.\footnote{This result holds even when limiting the sample to hospitals with a similar total executive team sizes (Appendix \ref{app:binned})}
+Thus, while the entire readmission rate response is not driven by highly clinical teams, the effect is exacerbated for these teams that have more clinicians than the median hospital.\footnote{This result holds even when limiting the sample to hospitals with a similar total executive team sizes (Appendix \ref{app:binned})}
 
@@
 
-    \section{Are Clinical Executives Less Profit Driven?}\label{sec:forprofit}
-
-    The hypothesis driven from the theory described in Section \ref{sec:theory} is that clinically trained executives put more weight on quality of care before the financial incentives on quality began.
+    \section{Potential Mechanisms}
+
+    The specific aspects of leadership that link executives to employees to patient outcomes are nearly impossible to observe.
+As the data on this topic progresses, these granular details of leadership will become more accessible to analyze.
+For the purpose of this paper, it is important for policymakers to understand why there is a behavioral difference in clinician versus non-clinician leaders, as this might affect the types of incentives they will respond to.
+
+    There are several hypotheses that I test given the data available.
+First, as the theoretical predictions in Section \ref{sec:theory} state, it is possible that clinicians are less financially driven than non-clinicians, and thus respond less drastically to financial incentives.
+I test this empirically in Section \ref{sec:forprofit}, comparing the response of nonprofits with different leadership teams to the response of for-profit hospitals on average.
+Second, I investigate whether clinicians are actively managing hospitals differently or whether they are a signal of underlying preferences of the hospital by leveraging the timing of clinical leaders relative to the incentives in Section \ref{sec:sig_man}.
+Third, in Section \ref{sec:selection}, I consider whether different types of executives might have different propensities to engage in "gaming" through selective patient practices.
+In Section \ref{sec:conditions}, I dig deeper into the specific conditions driving the main result, and whether these can help uncover the types of leadership changes made.
+Similarly, I analyze whether different executives change investment behavior in response to the incentives in Section \ref{sec:investment}.
+Finally, I consider the altruism versus expertise argument using a heterogeneity analysis of different physician specialties in Section \ref{sec:expertise}.
+
+    \subsection{Financial vs. Patient Driven}\label{sec:forprofit}
+
+    The typical theoretical prediction surrounding quality and pay-for-performance given different valuations of finances vs.
+patient welfare, shown in Section \ref{sec:theory}, certainly support the idea that, if all else is equal, clinician leaders value patients at baseline more than non-clinician leaders.
 To investigate this hypothesis empirically, I merge all general for-profit hospitals in the AHA survey to my sample of nonprofit hospitals.
-Then, I compare readmission and mortality rates of always clinical executives and never clinical executives to for-profit hospitals after the incentive change.
+I then compare readmission and mortality rates of always clinical executives and never clinical executives to for-profit hospitals after the incentive change.
+
     While I do not have information on leadership teams of for-profit hospitals, on average, they have revealed through their ownership status that they more heavily weigh profit.
@@
 
-\section{Signaling vs. Managing Decomposition}\label{sec:sig_man}
+    \subsection{Signaling vs. Managing}\label{sec:sig_man}
 
@@
 
-    \subsection{Analysis of Executive Team Changes} \label{sec:changes}
+    \subsubsection{Analysis of Executive Team Changes} \label{sec:changes}
 
@@
 
-     \subsection{Estimation}\label{sec:decomp}
+     \subsubsection{Estimation}\label{sec:decomp}
 
@@
 
-    \section{Heterogeneity by Physician Specialty}
+    \subsection{Patient Selection}\label{sec:selection}
+
+    \subsection{Specific Conditions (rename later)}\label{sec:conditions}
+
+    \subsection{Changes in Investment}\label{sec:investment}
+
+    \subsection{Altruism vs. Expertise}\label{sec:expertise}
 
@@
 
+    \subsection{Including Nurses in the Sample}\label{app:nurses}
+
     \subsection{Binned Treatment}\label{app:binned}
@@
          \includegraphics[width=\textwidth]{Objects/cont_belowmedread_md_nomd_size_synth_graph.pdf}
-         \label{fig:belowmed_read_synth_clinical}
+
      \end{subfigure}
@@
          \includegraphics[width=\textwidth]{Objects/cont_abovemedread_md_nomd_size_synth_graph.pdf}
-         \label{fig:abovemed_read_synth_clinical}
-     \end{subfigure}
-        \label{fig:cont_clinicalsynthdid_read_size}
+
+     \end{subfigure}
+
     \end{figure}
@@
          \includegraphics[width=\textwidth]{Objects/cont_belowmedmort_md_nomd_size_synth_graph.pdf}
-         \label{fig:belowmed_read_synth_clinical}
+
      \end{subfigure}
@@
          \includegraphics[width=\textwidth]{Objects/cont_abovemedmort_md_nomd_size_synth_graph.pdf}
-         \label{fig:abovemed_read_synth_clinical}
-     \end{subfigure}
-        \label{fig:cont_clinicalsynthdid_size}
+
+     \end{subfigure}
+
     \end{figure}

### 2025-11-04 Slides/JobTalk.tex
Updates from Overleaf

@@
 % Author 
-\node[below=1cm] (author) at (title.south){\large Hanna Glenn, Emory University};
+\node[below=1cm] (author) at (title.south){\large Hanna Glenn};
 % Date
@@
 
-\vspace{5mm} \large \pause
+\vspace{10mm} \large \pause
 
 \begin{enumerate}
-    \item \textbf{Hospital care makes up a third of health care spending} \scriptsize (\cite{cms2024nhe}) \normalsize
-
-\vspace{2mm}
-
-    \begin{itemize}
-        \item \normalsize Nonprofits drive the hospital industry \scriptsize (\cite{ASPE_2023}) \normalsize
-    \end{itemize}
-
-\vspace{8mm}
-
+    \item \textbf{Nonprofits drive the hospital industry} \scriptsize (\cite{ASPE_2023})
+
+\vspace{2mm}
+
+    \begin{wideitemize}
+        \item Should nonprofits receive the benefits that they do? 
+        \item Push to make nonprofits act more like nonprofits
+    \end{wideitemize}
+    \end{enumerate}
+\end{frame}
+
+%\begin{frame}{Addressing the Gaps}
+    %\centering
+    %\begin{tikzpicture}
+        % First image (top-left)
+        %\node[anchor=north west] (img1) at (0, 3) {\includegraphics[width=0.75\textwidth]{Graphics/nonprofit_screenshot1.png}};
+        %\draw[line width=.45mm, lightcyan] (img1.north west) rectangle (img1.south east);
+
+        % Second image (slightly down and right of the first image)
+        %\node[anchor=north west] (img2) at (0, 0) {\includegraphics[width=0.75\textwidth]{Graphics/nonprofit_screenshot2.png}};
+        %\draw[line width=.45mm, lightcyan] (img2.north west) rectangle (img2.south east);
+
+    %\end{tikzpicture}
+%\end{frame}
+
+\begin{frame}{Addressing the Gaps}
+
+    \Large
+    \begin{block}{}
+    \begin{center}
+    I examine the role of leadership in nonprofit hospitals
+
+    responding to policy incentives
+    \end{center}
+    \end{block}
+
+\vspace{18mm} \large
+
+\begin{enumerate}
+    \setcounter{enumi}{1}
     \item  \large \textbf{Policies often aim to improve the value of health care delivery}
@@
         \item Score metric determines rewards
-        \item Rewards high metric
     \end{itemize}
@@
     \item Through readmission rates
-    \item Non-clinical executive teams behave like profit-driven hospitals
 \end{wideitemize}
@@
     \begin{wideitemize}
-        \item \normalsize Physician board members $\implies$ lower public donations \scriptsize (\cite{brickley2010board}) \normalsize
-        \item No correlation b/w CEOs and hospital productivity in England \scriptsize (\cite{janke2019impact})\normalsize
-        \item Public hospitals in Chile improved when older physicians left CEO \scriptsize (\cite{otero2022managers}) \normalsize
+        \normalsize \item Physician board members $\implies$ lower public donations \scriptsize (\cite{brickley2010board}) \normalsize
+        \normalsize \item No correlation b/w CEOs and hospital productivity in England \scriptsize (\cite{janke2019impact})\normalsize
+        \normalsize \item Public hospitals in Chile improved when older physicians left CEO \scriptsize (\cite{otero2022managers}) \normalsize
     \end{wideitemize}
@@
     \begin{wideitemize}
-        \item \normalsize Hospitals decreased readmissions \scriptsize (\cite{gupta2021impacts}) \normalsize
-        \item Market concentration and systems affect response \scriptsize (\cite{kunz2024assessing}) \normalsize
-        \item Rewards had little impact on quality \scriptsize (\cite{norton2018moneyball}; \cite{friedson2019so})
-    \end{wideitemize}
-\end{frame}
-
-\begin{frame}{Rest of Talk}
-
-\large
-
-    Setting and Data Collection
-
-    \vspace{5mm}
-
-    About Executive Teams
-
-    \vspace{5mm}
-
-    Theoretical Motivation
-
-    \vspace{5mm}
-
-    Effect of Clinical Leadership
-
-    \vspace{5mm}
-
-    Comparison to For-Profits
-
-    \vspace{5mm}
-
-    Signaling vs. Managing
+        \normalsize \item Hospitals decreased readmissions \scriptsize (\cite{gupta2021impacts})
+
+        \vspace{-.5mm}
+
+        \normalsize \item Market concentration and systems affect response 
+        \scriptsize (\cite{kunz2024assessing})
+
+        \normalsize\item Rewards had little impact on quality \scriptsize (\cite{norton2018moneyball}; \cite{friedson2019so})
+    \end{wideitemize}
 \end{frame}
@@
     \end{itemize}
+
+    \vspace{5mm}
+
+    \textbf{Years 2010-2014}
 \end{frame}
@@
 
-\section{Signaling vs. Managing}
+\section{Unobserved}
 
@@
 \centering \LARGE
-    \textcolor{black}{Signaling vs. Managing}
+    \textcolor{black}{Driven by Unobserved Compositional Differences?}
 \end{transitionframe}
 
-\begin{frame}{Hypotheses}
-
-\large
-    \textbf{Clinical leaders represent underlying objectives ($\alpha$)}
-
-\vspace{2mm}
-
-    \begin{wideitemize}
-        \item Timing of the clinical leader doesn't matter 
-    \end{wideitemize}
-
-    \vspace{15mm}
-
-    \textbf{Clinical leaders effectively change objectives ($\alpha$)}
-
-\vspace{2mm}
-
-    \begin{wideitemize}
-        \item Timing of the clinical leader does matter
-    \end{wideitemize}
-\end{frame}
-
-\begin{frame}{Estimation}
+\begin{frame}{Driven by Unobserved Compositional Differences?}
+\large
+
+\textbf{Concern:} presence of clinicians may not matter but is correlated with some unobserved hospital `type', possibly time varying
+
+\vspace{3mm}
+\begin{wideitemize}
+    \item if true, we should see a difference in response no matter the timing of having a clinical executive
+\end{wideitemize}
+\end{frame}
+
+\begin{frame}{Robustness Check}
 \large
@@
  \hline \vspace{1mm}
- \textbf{Signaling:}\\ [0.5ex]
+ \textbf{Signaling Effect:}\\ [0.5ex]
  \hline\hline 
@@
  \hline \vspace{1mm}
- \textbf{Managing:}\\ [0.5ex]
+ \textbf{Clinician Effect:}\\ [0.5ex]
  \hline\hline
@@
  & (0.07) &  & (0.06) & \\
-Managing Effect &  & 0.17 &  & 0.12\\
+Clinician Effect &  & 0.17 &  & 0.12\\
  &  & (0.1) &  & (0.09)\\
@@
 
+\begin{frame}{Results by Specialty}\label{specialty}
+\begin{center}
+    \includegraphics[width=.4\textwidth]{Objects/read_nospecialty_nomd_synth_graph.pdf}
+\end{center}
+
+    \btVFill
+
+\hyperlink{main:readrates}{\beamerreturnbutton{main results}}
+\end{frame}
+
 \end{document}

### 2025-11-10 Tables/investment_table.tex
add HCRIS investment variables

\begin{table}[ht!]

\caption{\label{tab:investment}Effect of Clinically Trained Executive on Hospital-Level Expenses}
\centering
\begin{tabular}[t]{lccclc}
\toprule
 & Moveable Equip. & Fixed Equipment & Labor Costs & Building Purchases & Total Expenses\\
\midrule
Clinical Exec x Post 2012 & -1 & 6.95 & -1.37 & 0.56 & 1.83\\
 & (2.05) & (6.14) & (2.22) & (7.28) & (8.75)\\
 &  &  &  &  & \\
Observations & 1,532 & 660 & 1,670 & 1,435 & 2,485\\
\bottomrule
\multicolumn{6}{l}{\textsuperscript{} Standard errors are clustered at the hospital level.}\\
\multicolumn{6}{l}{\textsuperscript{} Signif. codes: 0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1}\\
\end{tabular}
\end{table}

### 2025-11-10 Paper/paperdraft_fall2025.tex

@@
 
-    \section{Theoretical Framework}\label{sec:theory}
-
-    Consider a simplified model of hospital objectives, in which hospitals choose a level of quality of care that maximizes an objective function capturing the trade-off between profit and patient welfare.
-That is,
-
-    $$\max_{\theta}\hspace{2mm}\alpha\pi(\theta) + (1-\alpha) u(\theta),$$
-
-    \noindent where $\pi(\theta)$ is a profit function depending on quality $\theta$, $u(.)$ is extra utility gained from patient welfare, which is increasing and concave in quality $\theta$, and $\alpha\in[0,1]$ captures the weight placed on profit vs.
-patient welfare.
-There is an implicit stay-open condition throughout the model.
-
-    I consider hospital choice of quality under pay-for-performance.
-In a world where performance is not incentivized through financial gain (as in traditional fee-for-service settings), hospital profit $\pi$ is maximized when the quantity of patients is maximized.
-However, pay-for-performance incentives incorporate quality into the profit function (\cite{dranove2011health}).
-Therefore, hospitals that only value financial profit will only consider quality if it is an input to the profit function, and therefore will respond more drastically to pay-for-performance than a hospital who values quality as an input to patient welfare before the incentives.
-
-    I will now formalize this intuition mathematically.
-For simplicity, I abstract away from quality having a direct impact on quantity of patients or price.\footnote{As long as the relationship between quality and quantity/price are not correlated with $\alpha$, this does not change the conclusion of the model.} Taking the first order condition yields:
-
-    $$(1-\alpha)u'(\theta) = -\alpha \pi'(\theta),$$
-
-    \noindent where marginal benefit equals marginal cost of increasing quality. Solving for $u'(\theta)$ and differentiating with respect to $\alpha$,
-
-    $$\frac{du'(\theta)}{d\alpha} = \frac{-1}{(1-\alpha)^2}\pi'(\theta).$$
-
-    Now, consider incentives when the hospital is in a strictly fee-for-service environment.
-That is, quality benefits patients, but hospitals do not directly derive profit from having a high quality.
-Then, assuming there is a positive cost to increasing quality, $\pi(\theta)$ is decreasing in $\theta$.
-Then $\frac{du'(\theta)}{d\alpha} > 0$.
-Using the Implicit Function Theorem, I can write $\frac{d\theta}{d\alpha}$ as
-    $\frac{du'(\theta)/d\alpha}{du'(\theta)/d\theta}<0$, revealing that in a fee-for-service setting, the more weight the hospital places on profit, the lower quality of care chosen at baseline.
-
-    Next, assume the hospital is in a pay-for-performance setting, where quality benefits patients and directly increases profit.
-Then, assuming the marginal financial benefit of increasing quality is greater than the marginal financial cost, $\pi(\theta)$ is increasing in $\theta$.
-Then $\frac{du'(\theta)}{d\alpha} < 0$, revealing that in a pay-for-performance setting, the more weight placed on profit, the higher quality of care chosen by the hospital.
-I combine the results found in each scenario into one response function that depends on $\alpha$, where $\theta_{p4p}$ is quality under pay-for-performance incentives and $\theta_{ffs}$ is quality under fee-for-service only.
-That is,
-
-    \begin{align*}
-        \frac{d\Delta\theta}{d\alpha}&=\frac{d(\theta_{p4p}-\theta_{ffs})}{d\alpha}\\
-        &=\frac{d\theta_{p4p}}{d\alpha}-\frac{d\theta_{ffs}}{d\alpha}\\
-        &> 0.
-    \end{align*}
-
-    The first prediction of the model is that, all else held equal, hospitals that value profit will have a lower quality than those who value patient welfare in a fee-for-service setting.
-Second, a hospital's change in quality once pay-for-performance is enacted depends on how much weight the hospital places on profit vs.
-patient welfare.
-Particularly, hospitals with higher weight on profit respond more than hospitals with higher weight on patient welfare.
-Therefore, factors that affect weight $\alpha$ will be correlated with choice of quality, and with how quality changes under new pay-for-performance incentives.
-In this paper, I test the hypothesis that clinical experience on a hospital's leadership team is a factor that is correlated with weight $\alpha$, and thus affects response to pay-for-performance incentives.
-
     \section{Setting and Data}
@@
 
-    The specific aspects of leadership that link executives to employees to patient outcomes are nearly impossible to observe.
-As the data on this topic progresses, these granular details of leadership will become more accessible to analyze.
-For the purpose of this paper, it is important for policymakers to understand why there is a behavioral difference in clinician versus non-clinician leaders, as this might affect the types of incentives they will respond to.
-
-    There are several hypotheses that I test given the data available.
-First, as the theoretical predictions in Section \ref{sec:theory} state, it is possible that clinicians are less financially driven than non-clinicians, and thus respond less drastically to financial incentives.
-I test this empirically in Section \ref{sec:forprofit}, comparing the response of nonprofits with different leadership teams to the response of for-profit hospitals on average.
-Second, I investigate whether clinicians are actively managing hospitals differently or whether they are a signal of underlying preferences of the hospital by leveraging the timing of clinical leaders relative to the incentives in Section \ref{sec:sig_man}.
-Third, in Section \ref{sec:selection}, I consider whether different types of executives might have different propensities to engage in "gaming" through selective patient practices.
-In Section \ref{sec:conditions}, I dig deeper into the specific conditions driving the main result, and whether these can help uncover the types of leadership changes made.
-Similarly, I analyze whether different executives change investment behavior in response to the incentives in Section \ref{sec:investment}.
-Finally, I consider the altruism versus expertise argument using a heterogeneity analysis of different physician specialties in Section \ref{sec:expertise}.
-
-    \subsection{Financial vs. Patient Driven}\label{sec:forprofit}
-
-    The typical theoretical prediction surrounding quality and pay-for-performance given different valuations of finances vs.
-patient welfare, shown in Section \ref{sec:theory}, certainly support the idea that, if all else is equal, clinician leaders value patients at baseline more than non-clinician leaders.
-To investigate this hypothesis empirically, I merge all general for-profit hospitals in the AHA survey to my sample of nonprofit hospitals.
-I then compare readmission and mortality rates of always clinical executives and never clinical executives to for-profit hospitals after the incentive change.
-
-    While I do not have information on leadership teams of for-profit hospitals, on average, they have revealed through their ownership status that they more heavily weigh profit.
-Thus, comparing different executive team behavior to for-profit behavior helps reveal how profit driven certain leadership teams are.
-Thus, I estimate the weighted synthetic difference-in-differences version of
-
-    \begin{equation}
-    \label{eq:forprofit}
-    y_{ht} = \beta \text{ (for-profit x post 2012)}_{ht} + \gamma_{h} + \delta_t + \epsilon_{ht}
-    \end{equation}
-
-    \noindent where $y_{ht}$ is one of the readmission or mortality rate outcomes discussed in Section \ref{sec:hrrp}, for-profit$_{ht}$ is an indicator for having for-profit ownership status, and $\gamma$ and $\delta$ are hospital and time fixed effects.
-In my estimation, nonprofits in the comparison group are more heavily weighted if their pre-trends are more similar to for-profits, and time periods are weighted to balance pre and post trends for for-profit hospitals.
-These weights follow the synthetic difference-in-differences approach, which is more robust to the parallel trends assumption.
-
-    I create two subsets of data to compare the response of for-profits to the response of the subsets of nonprofits based on executive team.
-First, I limit only to for-profits and nonprofits that always have clinically trained executives.
-Second, I limit to for-profits and nonprofits that never have clinically trained executives.\footnote{I assume that the sample selection of stable leadership teams is not correlated with the policy changes.
-Please see Section \ref{sec:sig_man} for the analysis and discussion of this assumption.} I show summary statistics of relevant variables for for-profit hospitals in Appendix \ref{app:sumstats}.
-Again, I employ a synthetic difference-in-differences strategy to rely less heavily on a strong parallel trends assumption for different types of hospitals.
-
-    The identification assumptions I make here are largely similar to those in Section \ref{sec:clinical}, only now incorporate for-profit hospitals.
-First, I assume that, absent the change in incentives, the weighted composition of for-profits and the relevant comparison group would have had parallel trends in outcomes.
-I also assume that there is no anticipation of the policy changes prior to 2012, which is reasonable as the rules of the policies were announced in October of 2011.
-Finally, I assume no other changes correlated with both hospital type (for-profit, nonprofit with and without clinically trained executives) and quality occurred in conjunction with the 2012 pay-for-performance policies.
-
-    The estimates and graphical representation of the effect of being for-profit is shown in Figure \ref{fig:read_synth_plot}.
-In panel \ref{fig:read_synth_plotb}, the average treatment effect shows the effect of being for-profit compared to nonprofit with clinical executives on readmission rates after pay-for-performance changes.
-I find that for-profits decrease readmission rates by .3 (standard error .12) more than nonprofits with clinically trained executives.
-That is, for-profits respond more aggressively to the incentive change than nonprofits with clinical leadership.
-This difference is nearly identical to the difference between clinical and non-clinical teams shown in Section \ref{sec:clinical}.
-So, unsurprisingly, panel \ref{fig:read_synth_plotc} shows no such difference when comparing for-profits to nonprofits without clinically trained executives.
-Not only is the estimate not statistically significant, but the magnitude indicates a precise zero.
-That is, a lack of clinically trained executives leads nonprofits to act like for-profits in terms of readmission rates when responding to this incentive change.
-
-         \begin{figure}[ht!]
-     \caption{Comparison to For-Profit: Readmission}
-     \centering
-     \begin{subfigure}[b]{0.45\textwidth}
-         \centering
-         \caption{For-Profit and Always Clinical Exec}
-         \includegraphics[width=\textwidth]{Objects/fp_read_md_synth_graph.pdf}
-         \label{fig:read_synth_plotb}
-     \end{subfigure}%
-     \vspace{5mm}
-     \hfill
-     \begin{subfigure}[b]{0.45\textwidth}
-         \centering
-         \caption{For-Profit and Never Clinical Exec}
-         \includegraphics[width=\textwidth]{Objects/fp_read_nomd_synth_graph.pdf}
-         \label{fig:read_synth_plotc}
-     \end{subfigure}
-        \label{fig:read_synth_plot}
-    \end{figure}
-
-    Next, I present estimated differences in mortality rates between different hospital types in Figure \ref{fig:mort_synth_plot}.
-In panel \ref{fig:mort_synth_plotb}, the estimated difference is between for-profits and nonprofits with clinically trained executives.
-The magnitude of the difference in mortality rates is .08, where clinical team hospitals continue on an increasing trend and for-profits remain stable in mortality rate.
-While I cannot rule out a lack of differential response, the documented noisiness of mortality measures (\cite{mackenzie2016measuring}) may mean that the true effect is different from zero.
-However, the estimated difference in mortality between for-profits and nonprofits without clinically trained executives is certainly zero with a magnitude of .03, shown in panel \ref{fig:mort_synth_plotc}.
-
-    \begin{figure}[ht!]
-     \caption{Comparison to For-Profit: Mortality}
-     \centering
-     \begin{subfigure}[b]{0.45\textwidth}
-         \centering
-         \caption{For-Profit and Always Clinical Exec}
-         \includegraphics[width=\textwidth]{Objects/fp_mort_md_synth_graph.pdf}
-         \label{fig:mort_synth_plotb}
-     \end{subfigure}%
-     \hfill
-     \begin{subfigure}[b]{0.45\textwidth}
-         \centering
-         \caption{For-Profit and Never Clinical Exec}
-         \includegraphics[width=\textwidth]{Objects/fp_mort_nomd_synth_graph.pdf}
-         \label{fig:mort_synth_plotc}
-     \end{subfigure}
-        \label{fig:mort_synth_plot}
-    \end{figure}
-
-    \subsection{Signaling vs. Managing}\label{sec:sig_man}
+    While uncovering that there is some foundational difference in clinical and non-clinical hospital leaders is important in its own right, it is especially relevant to understand why this difference might come about.
+In this Section, I present several theoretical predictions consistent with the observed difference in quality after pay-for-performance incentives, and  test the predictions empirically.
+First, I investigate whether clinicians are actively managing hospitals differently or whether they are a signal of underlying preferences of the hospital by leveraging the timing of clinical leaders relative to the incentives in Section \ref{sec:sig_man}.
+Second, the difference could be driven by the altruism, or patient-centered focus, of clinicians.
+I investigate this in Section \ref{sec:forprofit}, comparing the response of nonprofits with different leadership teams to the response of for-profit hospitals on average.
+Third, I consider the altruism versus expertise argument using a heterogeneity analysis of different physician specialties in Section \ref{sec:expertise}.
+In Section \ref{sec:selection}, I consider whether different types of executives might have different propensities to engage in "gaming" through selective patient practices.
+Finally, I analyze whether different executives change investment behavior in response to the incentives in Section \ref{sec:investment}.
+
+    \subsection{Signaling versus Managing}\label{sec:sig_man}
 
     Given that having clinically trained executives does affect hospital response to financial incentives on quality, an important distinction is whether the existing clinical leaders \textit{reveal} the underlying objectives of the hospital, or \textit{manage} the hospital differently.
-Up to this point, I have limited the sample of nonprofit hospitals to those that don't change their propensity to hire a clinically trained executive over the sample period, which combines any signaling and managing effects.
+In the main analysis, I limited the sample of nonprofit hospitals to those that don't change their propensity to hire a clinically trained executive over the sample period, which combines any signaling and managing effects.
 However, under the assumption that hiring or firing these types of executives is not correlated with the pay-for-performance policy changes, such changes can be leveraged to decompose the effect into signaling versus managing.
+Specifically, if the difference is driven by underlying preferences of the hospital, the observed difference in quality will still hold whether the hospital had a clinical executive at any point in time, even if they were not a leader in 2012 when the policies took effect.
+Alternatively, if the observed difference is driven by managing, then the difference should only be observed when there is a clinician in leadership during the policy enactment.
 Since this decomposition depends heavily on the assumption that clinical executive hiring is not endogenous, I first investigate leadership changes in response to the policy change.
@@
 
-    \subsection{Patient Selection}\label{sec:selection}
-
-    \subsection{Specific Conditions (rename later)}\label{sec:conditions}
-
-    \subsection{Changes in Investment}\label{sec:investment}
-
-    \subsection{Altruism vs. Expertise}\label{sec:expertise}
-
+    \subsection{Financially Driven versus Patient Driven}\label{sec:forprofit}
+
+    Consider a simplified model of hospital objectives, in which hospitals choose a level of quality of care that maximizes an objective function capturing the trade-off between profit and patient welfare.
+That is,
+
+    $$\max_{\theta}\hspace{2mm}\alpha\pi(\theta) + (1-\alpha) u(\theta),$$
+
+    \noindent where $\pi(\theta)$ is a profit function depending on quality $\theta$, $u(.)$ is extra utility gained from patient welfare, which is increasing and concave in quality $\theta$, and $\alpha\in[0,1]$ captures the weight placed on profit vs.
+patient welfare.
+There is an implicit stay-open condition throughout the model.
+
+    I consider hospital choice of quality under pay-for-performance.
+In a world where performance is not incentivized through financial gain (as in traditional fee-for-service settings), hospital profit $\pi$ is maximized when the quantity of patients is maximized.
+However, pay-for-performance incentives incorporate quality into the profit function (\cite{dranove2011health}).
+Therefore, hospitals that only value financial profit will only consider quality if it is an input to the profit function, and therefore will respond more drastically to pay-for-performance than a hospital who values quality as an input to patient welfare before the incentives.
+
+    I will now formalize this intuition mathematically.
+For simplicity, I abstract away from quality having a direct impact on quantity of patients or price.\footnote{As long as the relationship between quality and quantity/price are not correlated with $\alpha$, this does not change the conclusion of the model.} Taking the first order condition yields:
+
+    $$(1-\alpha)u'(\theta) = -\alpha \pi'(\theta),$$
+
+    \noindent where marginal benefit equals marginal cost of increasing quality. Solving for $u'(\theta)$ and differentiating with respect to $\alpha$,
+
+    $$\frac{du'(\theta)}{d\alpha} = \frac{-1}{(1-\alpha)^2}\pi'(\theta).$$
+
+    Now, consider incentives when the hospital is in a strictly fee-for-service environment.
+That is, quality benefits patients, but hospitals do not directly derive profit from having a high quality.
+Then, assuming there is a positive cost to increasing quality, $\pi(\theta)$ is decreasing in $\theta$.
+Then $\frac{du'(\theta)}{d\alpha} > 0$.
+Using the Implicit Function Theorem, I can write $\frac{d\theta}{d\alpha}$ as
+    $\frac{du'(\theta)/d\alpha}{du'(\theta)/d\theta}<0$, revealing that in a fee-for-service setting, the more weight the hospital places on profit, the lower quality of care chosen at baseline.
+
+    Next, assume the hospital is in a pay-for-performance setting, where quality benefits patients and directly increases profit.
+Then, assuming the marginal financial benefit of increasing quality is greater than the marginal financial cost, $\pi(\theta)$ is increasing in $\theta$.
+Then $\frac{du'(\theta)}{d\alpha} < 0$, revealing that in a pay-for-performance setting, the more weight placed on profit, the higher quality of care chosen by the hospital.
+I combine the results found in each scenario into one response function that depends on $\alpha$, where $\theta_{p4p}$ is quality under pay-for-performance incentives and $\theta_{ffs}$ is quality under fee-for-service only.
+That is,
+
+    \begin{align*}
+        \frac{d\Delta\theta}{d\alpha}&=\frac{d(\theta_{p4p}-\theta_{ffs})}{d\alpha}\\
+        &=\frac{d\theta_{p4p}}{d\alpha}-\frac{d\theta_{ffs}}{d\alpha}\\
+        &> 0.
+    \end{align*}
+
+    The first prediction of the model is that, all else held equal, hospitals that value profit will have a lower quality than those who value patient welfare in a fee-for-service setting.
+Second, a hospital's change in quality once pay-for-performance is enacted depends on how much weight the hospital places on profit vs.
+patient welfare.
+Particularly, hospitals with higher weight on profit respond more than hospitals with higher weight on patient welfare.
+Therefore, factors that affect weight $\alpha$ will be correlated with choice of quality, and with how quality changes under new pay-for-performance incentives.
+In this paper, I test the hypothesis that clinical experience on a hospital's leadership team is a factor that is correlated with weight $\alpha$, and thus affects response to pay-for-performance incentives.
+To investigate this hypothesis empirically, I merge all general for-profit hospitals in the AHA survey to my sample of nonprofit hospitals.
+I then compare readmission and mortality rates of always clinical executives and never clinical executives to for-profit hospitals after the incentive change.
+
+    While I do not have information on leadership teams of for-profit hospitals, on average, they have revealed through their ownership status that they more heavily weigh profit, and therefore have an $\alpha$ close to 1.
+This is not inconsistent with previous work positing that nonprofits and for-profits act similarly on average (\cite{duggan2002hospital}).
+In fact, my data is consistent with this idea when nonprofits are not broken down by leadership teams.
+However, comparing different executive team behavior to for-profit behavior helps reveal how profit driven certain leadership teams are, which could actually help us understand why there are similarities between nonprofits and for-profits on average.
+I estimate the weighted synthetic difference-in-differences version of
+
+    \begin{equation}
+    \label{eq:forprofit}
+    y_{ht} = \beta \text{ (for-profit x post 2012)}_{ht} + \gamma_{h} + \delta_t + \epsilon_{ht}
+    \end{equation}
+
+    \noindent where $y_{ht}$ is one of the readmission or mortality rate outcomes discussed in Section \ref{sec:hrrp}, for-profit$_{ht}$ is an indicator for having for-profit ownership status, and $\gamma$ and $\delta$ are hospital and time fixed effects.
+In my estimation, nonprofits in the comparison group are more heavily weighted if their pre-trends are more similar to for-profits, and time periods are weighted to balance pre and post trends for for-profit hospitals.
+These weights follow the synthetic difference-in-differences approach, which is more robust to the parallel trends assumption.
+
+    I create two subsets of data to compare the response of for-profits to the response of the subsets of nonprofits based on executive team.
+First, I limit only to for-profits and nonprofits that always have clinically trained executives.
+Second, I limit to for-profits and nonprofits that never have clinically trained executives.\footnote{I assume that the sample selection of stable leadership teams is not correlated with the policy changes.
+Please see Section \ref{sec:sig_man} for the analysis and discussion of this assumption.} I show summary statistics of relevant variables for for-profit hospitals in Appendix \ref{app:sumstats}.
+Again, I employ a synthetic difference-in-differences strategy to rely less heavily on a strong parallel trends assumption for different types of hospitals.
+
+    The identification assumptions I make here are largely similar to those in Section \ref{sec:clinical}, only now incorporate for-profit hospitals.
+First, I assume that, absent the change in incentives, the weighted composition of for-profits and the relevant comparison group would have had parallel trends in outcomes.
+I also assume that there is no anticipation of the policy changes prior to 2012, which is reasonable as the rules of the policies were announced in October of 2011.
+Finally, I assume no other changes correlated with both hospital type (for-profit, nonprofit with and without clinically trained executives) and quality occurred in conjunction with the 2012 pay-for-performance policies.
+
+    The estimates and graphical representation of the effect of being for-profit is shown in Figure \ref{fig:read_synth_plot}.
+In panel \ref{fig:read_synth_plotb}, the average treatment effect shows the effect of being for-profit compared to nonprofit with clinical executives on readmission rates after pay-for-performance changes.
+I find that for-profits decrease readmission rates by .3 (standard error .12) more than nonprofits with clinically trained executives.
+That is, for-profits respond more aggressively to the incentive change than nonprofits with clinical leadership.
+This difference is nearly identical to the difference between clinical and non-clinical teams shown in Section \ref{sec:clinical}.
+So, unsurprisingly, panel \ref{fig:read_synth_plotc} shows no such difference when comparing for-profits to nonprofits without clinically trained executives.
+Not only is the estimate not statistically significant, but the magnitude indicates a precise zero.
+That is, a lack of clinically trained executives leads nonprofits to act like for-profits in terms of readmission rates when responding to this incentive change.
+
+         \begin{figure}[ht!]
+     \caption{Comparison to For-Profit: Readmission}
+     \centering
+     \begin{subfigure}[b]{0.45\textwidth}
+         \centering
+         \caption{For-Profit and Always Clinical Exec}
+         \includegraphics[width=\textwidth]{Objects/fp_read_md_synth_graph.pdf}
+         \label{fig:read_synth_plotb}
+     \end{subfigure}%
+     \vspace{5mm}
+     \hfill
+     \begin{subfigure}[b]{0.45\textwidth}
+         \centering
+         \caption{For-Profit and Never Clinical Exec}
+         \includegraphics[width=\textwidth]{Objects/fp_read_nomd_synth_graph.pdf}
+         \label{fig:read_synth_plotc}
+     \end{subfigure}
+        \label{fig:read_synth_plot}
+    \end{figure}
+
+    Next, I present estimated differences in mortality rates between different hospital types in Figure \ref{fig:mort_synth_plot}.
+In panel \ref{fig:mort_synth_plotb}, the estimated difference is between for-profits and nonprofits with clinically trained executives.
+The magnitude of the difference in mortality rates is .08, where clinical team hospitals continue on an increasing trend and for-profits remain stable in mortality rate.
+While I cannot rule out a lack of differential response, the documented noisiness of mortality measures (\cite{mackenzie2016measuring}) may mean that the true effect is different from zero.
+However, the estimated difference in mortality between for-profits and nonprofits without clinically trained executives is certainly zero with a magnitude of .03, shown in panel \ref{fig:mort_synth_plotc}.
+
+    \begin{figure}[ht!]
+     \caption{Comparison to For-Profit: Mortality}
+     \centering
+     \begin{subfigure}[b]{0.45\textwidth}
+         \centering
+         \caption{For-Profit and Always Clinical Exec}
+         \includegraphics[width=\textwidth]{Objects/fp_mort_md_synth_graph.pdf}
+         \label{fig:mort_synth_plotb}
+     \end{subfigure}%
+     \hfill
+     \begin{subfigure}[b]{0.45\textwidth}
+         \centering
+         \caption{For-Profit and Never Clinical Exec}
+         \includegraphics[width=\textwidth]{Objects/fp_mort_nomd_synth_graph.pdf}
+         \label{fig:mort_synth_plotc}
+     \end{subfigure}
+        \label{fig:mort_synth_plot}
+    \end{figure}
+
+    \subsection{Driven by Expertise}\label{sec:expertise}
+
+    It may be that the observed difference in quality after pay-for-performance could be correlated with the expertise of leadership teams.
 While most doctors take a hippocratic oath to provide care with integrity and compassion, doctors with training in certain specialties may be more equipped to improve care for certain patients.
@@
 
-    I limit the sample of hospitals to those who always have some type of clinical executive over the sample period.
+    Consider an extension to the model discussed in Section \ref{sec:forprofit}, where I now introduce a measure of expertise $\lambda$.
+To capture the role of hospital leadership expertise in shaping quality decisions, we introduce a binary variable $\lambda \in \{0,1\}$, where $\lambda=1$ indicates high expertise and $\lambda=0$ indicates low expertise.
+Expertise affects the marginal cost of quality improvements: hospitals with high expertise can implement quality enhancements more efficiently, reducing the effective cost of quality.
+
+    The hospital’s objective function becomes:
+    \begin{equation}
+        \max_{\theta} \alpha \big[R(\theta) - \frac{c(\theta)}{1+\lambda}\big] + (1 - \alpha) u(\theta),
+    \end{equation}
+
+    \noindent where $R(\theta)$ is revenue, which depends on quality,
+    $c(\theta)(1 + \lambda)$ is the cost of quality, increasing and convex in $\theta$, scaled by expertise, $u(\theta)$ is patient welfare, increasing and concave in $\theta$, $\alpha \in [0,1]$ is the weight on profit relative to patient welfare.
+The hospital's first-order condition under fee-for-service and pay-for-performance is then:
+
+    \[
+    F_{ffs}(\theta,\alpha,\lambda) = (1-\alpha)u'(\theta) - \alpha\frac{c'(\theta)}{1+\lambda} = 0,
+    \]
+
+    \[
+    F_{p4p}(\theta,\alpha,\lambda) = (1-\alpha)u'(\theta) - \alpha\Bigg[\frac{c'(\theta)}{1+\lambda} - R'(\theta)\Bigg] = 0.
+    \]
+
+    \noindent I make the following assumptions on patient welfare, cost, and revenue: $u'(\theta) > 0,\; u''(\theta) < 0,\; c'(\theta) > 0,\; c''(\theta) > 0,\; R'(\theta) > 0,\; R''(\theta) \approx 0$.
+By the Implicit Function Theorem.
+    \begin{align}
+    \frac{d\theta}{d\lambda} &= -\frac{\partial F/\partial \lambda}{\partial F/\partial \theta}\\
+    &= -\frac{\alpha c'(\theta)/(1+\lambda)^2}{(1-\alpha)u''(\theta)-\alpha\big[c''(\theta)/(1+\lambda)-R''(\theta)\big]}\\
+    &>0. 
+    \end{align}
+
+    I now want to compare the change in $\theta$ when going from fee-for-service to pay-for-performance depending on different levels of $\lambda$. 
+    \begin{align}
+        \frac{\partial \Delta \theta}{\partial \lambda} &= \frac{\partial\theta_{p4p}}{\partial \lambda} - \frac{\partial\theta_{ffs}}{\partial \lambda}\\
+        &= \frac{-\alpha c'(\theta)/(1+\lambda)^2}{(1-\alpha)u''(\theta)-\alpha\big[c''(\theta)/(1+\lambda)-R''(\theta)\big]} - \frac{-\alpha c'(\theta)/(1+\lambda)^2}{(1-\alpha)u''(\theta)-\alpha\big[c''(\theta)/(1+\lambda)\big]}.
+    \end{align}
+
+    Since $\theta_{p4p}>\theta_{ffs}$, $u(\theta_{p4p}>u(\theta_{ffs}$ (that is, less negative).
+Therefore, the pay-for-performance denominator is less negative than the fee-for-service denominator.
+Hence, $\frac{\partial\theta_{p4p}}{\partial \lambda} > \frac{\partial\theta_{ffs}}{\partial \lambda}$.
+This implies that for higher expertise, the response to pay-for-performance policies will be greater.
+
+    To test this empirically, I limit the sample of hospitals to those who always have some type of clinical executive over the sample period.
 Then, I estimate the difference in readmission and mortality rates after pay-for-performance among hospitals who do and do not have an internal medicine clinical executive.
@@
 This is probably due to the specialty-specific conditions being more relevant for the readmission rate penalties than the quality rewards.
+
+    \subsection{Patient Selection}\label{sec:selection}
+
+    \citeauthor{gupta2021impacts} (\citeyear{gupta2021impacts}) finds that a large portion of hospital readmission reductions are due to hospitals being less willing to readmit patients from their own hospital, while they will still readmit patients coming from other hospitals.
+Given the differential response to the HRRP by clinical and non-clinical hospital leaders, the natural question arises as to whether there is a differential propensity to game the system, as considered in previous work (\cite{gupta2021impacts}).
+While I do not have patient level data, I can test empirically whether the overall patient population of hospitals changes differentially for clinician or non-clinician leaders.
+
+    I use three outcomes to test this hypothesis.
+First, the CMS Case Mix Index files provide an aggregate measure of patient complexity for each hospital in each year.
+A higher index implies the patient population in the hospital is more complex in their treatment requirement.
+Additionally, I consider the population of patients treated in the hospital with public insurance, either Medicare or Medicaid.
+These outcomes will indicate if certain types of leadership teams attempt to cost shift relative to others after the penalties or rewards took effect.
+
+    I estimate the same synthetic difference-in-differences equation discussed in Section \ref{sec:clinical}.
+The results of these estimations are presented in Figure \ref{fig:selection}.
+All of the treatment effects in these estimations are insignificant and null in magnitude, suggesting that there is no evidence of the overall patient population of hospitals changing differentially based on leadership team.
+This does not mean that there is no change in specific readmission behavior as found in \citeauthor{gupta2021impacts} (\citeyear{gupta2021impacts}), but a limitation of this data only allows me to look at the aggregate patient population as opposed to specifically readmissions.
+
+    \begin{figure}[ht!]
+     \caption{Patient Population as Outcomes}
+     \centering
+     \begin{subfigure}[b]{0.45\textwidth}
+         \centering
+         \caption{Case Mix Index}
+         \includegraphics[width=\textwidth]{Objects/cmi_md_nomd_synth_graph.pdf}
+         \label{fig:cmi}
+     \end{subfigure}
+     \hfill
+     \begin{subfigure}[b]{0.45\textwidth}
+         \centering
+         \caption{Medicare Percent Discharges}
+         \includegraphics[width=\textwidth]{Objects/mcare_md_nomd_synth_graph.pdf}
+         \label{fig:mcare}
+     \end{subfigure}%
+        \begin{subfigure}[b]{0.45\textwidth}
+         \centering
+         \caption{Medicaid Percent Discharges}
+         \includegraphics[width=\textwidth]{Objects/mcaid_md_nomd_synth_graph.pdf}
+         \label{fig:mcaid}
+     \end{subfigure}
+        \label{fig:selection}
+    \end{figure}
+
+    \subsection{Changes in Investment}\label{sec:investment}
 

### 2025-11-10 Slides/JobTalk.tex

@@
 % Author 
-\node[below=1cm] (author) at (title.south){\large Hanna Glenn};
+\node[below=1cm] (author) at (title.south){\large Hanna Glenn \\
+University of Queensland};
 % Date
@@
 
-\begin{frame}{Data Collection Process}
-\begin{center}
-\begin{tikzpicture}[
-    node distance=3.1cm and 1.8cm,
-    every node/.style={align=center, font=\normalsize},
-    process/.style={rectangle, draw, rounded corners, text centered, minimum width=2.5cm, minimum height=2cm, fill=purple!30},
-    output/.style={rectangle, draw, text centered, minimum width=2.7cm, minimum height=2.5cm, fill=purple},
-    extract/.style={rectangle, draw, rounded corners, text centered, minimum width=2.5cm, minimum height=2cm, fill=sage!30},
-    arrow/.style={thick, ->, >=stealth}
-]
+%\begin{frame}{Data Collection Process}
+%\begin{center}
+%\begin{tikzpicture}[
+    %node distance=3.1cm and 1.8cm,
+    %every node/.style={align=center, font=\normalsize},
+    %process/.style={rectangle, draw, rounded corners, %text centered, minimum width=2.5cm, minimum height=2cm, fill=purple!30},
+    %output/.style={rectangle, draw, text centered, minimum width=2.7cm, minimum height=2.5cm, fill=purple},
+    %extract/.style={rectangle, draw, rounded corners, text centered, minimum width=2.5cm, minimum height=2cm, fill=sage!30},
+    %arrow/.style={thick, ->, >=stealth}
+%]
 
 % Nodes
-\node (step1) [process] {Hospital \\EINs};\pause
-\node (step2) [process, right of=step1] {Download \\PDFs};
-\draw[arrow] (step1) -- (step2); \pause
-\node (step3) [process, right of=step2] {Extract\\text};
-\draw[arrow] (step2) -- (step3); 
+%\node (step1) [process] {Hospital \\EINs};\pause
+%\node (step2) [process, right of=step1] {Download \\PDFs};
+%\draw[arrow] (step1) -- (step2); \pause
+%\node (step3) [process, right of=step2] {Extract\\text};
+%\draw[arrow] (step2) -- (step3); 
 % New green box below step3
-\node (extract) [extract, below of=step3, yshift=0.5cm] {Optical\\Character\\Recognition}; 
-\draw[arrow] (extract) -- (step3);\pause
-\node (step4) [process, right of=step3] {String\\cleaning\\algorithms};
-\draw[arrow] (step3) -- (step4); \pause
-\node (output) [output, right of=step4, text=white] {Executive\\Name\\Position\\Title};
-\draw[arrow] (step4) -- (output);
-\end{tikzpicture}
-\end{center}
-\end{frame}
+%\node (extract) [extract, below of=step3, yshift=0.5cm] {Optical\\Character\\Recognition}; 
+%\draw[arrow] (extract) -- (step3);\pause
+%\node (step4) [process, right of=step3] {String\\cleaning\\algorithms};
+%\draw[arrow] (step3) -- (step4); \pause
+%\node (output) [output, right of=step4, text=white] {Executive\\Name\\Position\\Title};
+%\draw[arrow] (step4) -- (output);
+%\end{tikzpicture}
+%\end{center}
+%\end{frame}
 
@@
     \begin{wideitemize}
-        \item AHA
+        \item American Hospital Association
         \item Hospital Compare
@@
 
-        \vspace{20mm}
-
-        \hyperlink{int_margin}{\beamergotobutton{intensive margin}}
-        \hyperlink{cmi}{\beamergotobutton{due to patient selection?}}
-        \hyperlink{specifications}{\beamergotobutton{specifications}}
-        \hyperlink{specialty}{\beamergotobutton{specialty}}
-
     \end{columns}
@@
         \end{wideitemize}
+    \end{columns}
+\end{frame}
+
+\begin{frame}{Intensive Margin}
+\large
+    \begin{equation*}
+    y_{ht} = \beta \left(\mathbb{1}\{\rho \leq \eta\}\text{ x post 2012}\right)_{ht} + \gamma_{h} + \delta_t + \epsilon_{ht}
+    \end{equation*}
+
+    \begin{equation*}
+    y_{ht} = \beta \left(\mathbb{1}\{\rho > \eta\}\text{ x post 2012}\right)_{ht} + \gamma_{h} + \delta_t + \epsilon_{ht}
+    \end{equation*}
+
+    \vspace{10mm}
+
+    \begin{itemize}
+        \item $\rho$: fraction of clinical executives
+        \item $\eta$: median fraction of clinically trained executives (.2)
+    \end{itemize}
+\end{frame}
+
+\begin{frame}{Intensive Margin: Readmission Rates}
+    \begin{columns}
+    \column{0.5\textwidth}
+        \centering
+        \includegraphics[width=.9\textwidth]{Objects/cont_belowmedread_md_nomd_synth_graph.pdf}
+    \column{0.5\textwidth}
+        \centering
+        \includegraphics[width=.9\textwidth]{Objects/cont_abovemedread_md_nomd_synth_graph.pdf}
     \end{columns}
@@
 
+\begin{frame}{Other Mechanisms}
+\large
+    \begin{wideitemize}
+        \item Not driven by changes in patient population
+        \item Driven primarily by pneumonia
+    \end{wideitemize}
+\end{frame}
+
 \begin{frame}{Summary}
@@
 
-    \vspace{5mm}
+    \vspace{10mm}
 
@@
 
-    \vspace{5mm}
-
-    \textbf{Future Work}
-
-    \begin{itemize}
-        \item Effects of mandated gender and racial diversity on hospital behavior
-        \item Networks of executives over time 
-        \item Are leaders correlated with hospital fraudulent behavior?
-    \end{itemize}
 \end{frame}
@@
 
-    \textcolor{black}{\faEnvelope \hspace{2mm} hanna.glenn@emory.edu}
-
-    \vspace{2mm}
-
-    \textcolor{black}{\faTwitter \hspace{2mm} @hannalynnglenn}
+    \textcolor{black}{\faEnvelope \hspace{2mm} hanna.glenn@uq.edu.au}
+
 \end{transitionframe}

### 2025-11-12 Tables/investment_table.tex
fix investment table
add CEO analysis graphs

@@
 \centering
-\begin{tabular}[t]{lccclc}
+\begin{tabular}[t]{lcccccc}
 \toprule
- & Moveable Equip. & Fixed Equipment & Labor Costs & Building Purchases & Total Expenses\\
+ & Nurses & Moveable Equip & Fixed Equip & Labor & Building & Total\\
 \midrule
-Clinical Exec x Post 2012 & -1 & 6.95 & -1.37 & 0.56 & 1.83\\
- & (2.05) & (6.14) & (2.22) & (7.28) & (8.75)\\
- &  &  &  &  & \\
-Observations & 1,532 & 660 & 1,670 & 1,435 & 2,485\\
+Clinical Exec x Post 2012 & 9.08 & -1 & 6.95 & -1.37 & 0.56 & 1.83\\
+ & (8.1) & (1.83) & (6.78) & (2.38) & (7.42) & (8.18)\\
+ &  &  &  &  &  & \\
+Observations & 2,485 & 1,532 & 660 & 1,670 & 1,435 & 2,485\\
 \bottomrule
-\multicolumn{6}{l}{\textsuperscript{} Standard errors are clustered at the hospital level.}\\
-\multicolumn{6}{l}{\textsuperscript{} Signif. codes: 0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1}\\
+\multicolumn{7}{l}{\textsuperscript{} Standard errors are clustered at the hospital level.}\\
+\multicolumn{7}{l}{\textsuperscript{} Signif. codes: 0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1}\\
 \end{tabular}

### 2025-11-12 Paper/paperdraft_fall2025.tex

@@
 Health care policies in the US often focus on improving quality of care without increasing costs, yet it is unclear whether hospital leaders might contribute to this goal.
-Leveraging pay-for-performance incentives, I estimate the differential response to financial incentives on quality by different types of nonprofit hospital executives: those with and without a clinical training background.
+Leveraging pay-for-performance incentives, I estimate the differential response to financial incentives on quality by different types of nonprofit hospital executives: those with and without a clinical training background as a medical doctor.
 I find that hospitals without clinically trained executives respond more to the financial incentives than those with clinically trained executives.
-This can be explained by a model of hospital objectives in which clinical training affects hospital preferences towards patient welfare.
+I find that this difference can be explained by patient-centered focus and by expertise lowering the cost of quality improvement.
 These findings highlight the importance of considering leadership as an important factor in hospital decision making when adjusting incentives.
@@
 Further, a major contributor of this quality improvement is the exodus of older physicians as CEOs (\cite{otero2022managers}).
-I contribute to our knowledge of how executive team characteristics affect firm behaviors by (1) considering the important setting of nonprofit hospitals, which serve the majority of patients in the US, (2) considering a unique and relevant executive characteristic, and (3) observing firm behavior in the presence of an exogenous incentive change.
+Finally, while not in the context of hospitals, \citeauthor{la2023impact} (\citeyear{la2023impact}) finds that OB-GYNs acquired by financially driven Physician Practice Management Companies (PPMCs) increase C-Sections compared to those acquired by clinically driven PPMCs (\cite{la2023impact}).
+Thus, the motivations of the management of health care facilities matters for provider behavior.
+I contribute to our knowledge of how executive team characteristics affect firm behaviors by (1) considering the important setting of nonprofit hospitals, which serve the majority of patients in the US, (2) considering a unique and relevant executive characteristic of the entire executive team, and (3) observing firm behavior in the presence of an exogenous incentive change.
 
@@
 I extract text spanning 2009-2015, but I retain a much larger sample of hospitals by limiting to hospitals that have text information in 2010-2014.
-This yields 852 nonprofit hospitals that are (1) matched to AHA data, and (2) contain complete leadership information in each year from 2010-2014.
+This yields 847 nonprofit hospitals that are (1) matched to AHA data, and (2) contain complete leadership information in each year from 2010-2014.
+For clarity, I include a table outlining the number of observations in each sample limitation in Table \ref{tab:sample}.
+
+    \begin{table}[ht!]
+    \centering
+    \caption{\label{tab:sample}Observations in Sample Limitations}
+    \centering
+    \begin{tabular}[t]{lc}
+    \toprule
+    Sample & Num. Observations\\
+    \toprule
+    Hospital EINs & 5,588\\
+    \addlinespace
+    Hospital EINs Matched to AHA Hospital or System \hspace{2mm} & 1,340\\
+    \addlinespace
+    Hospital EINs Match to AHA Hospital Only & 1,152\\
+    \addlinespace
+    Hospitals with Leadership Info Years 2010-2014 & 847\\
+    \bottomrule
+    \end{tabular}
+    \end{table}
 
@@
         \caption{Main Outcomes Over Time}
-        \includegraphics[width=.8\textwidth]{Objects/outcomes_graph.pdf}
+        \includegraphics[width=.9\textwidth]{Objects/outcomes_graph.pdf}
         \label{fig:outcomes_graph}
@@
     \end{figure}
+
+    Additionally, I consider the same comparison of clinical and non-clinical leaders, but only for CEOs.
+That is, do the results hold when considering only the most top manager instead of the executive team as a whole?
+I compare hospitals with a CEO trained as a doctor to hospitals without a CEO trained as a doctor using the same synthetic difference-in-differences strategy.
+I present the average treatment effect and standard errors of this estimation in Table .
+There is no difference in response when only considering clinical CEOs as the treatment since the effects are statistically insignificant or smaller in magnitude.
+This highlights the importance of considering executive teams as a whole as opposed to just upper managers.
 
@@
 
+    Lastly, I examine whether clinician versus non-clinician leaders respond differentially in terms of their investment strategy after the policies took effect.
+Ideally, I would have more granular information within the hospital to observe differences in management.
+That is a limitation of the publicly available hospital-level data that I utilize here.
+I use the same synthetic difference-in-differences estimation strategy to compare clinical leaders with non-clinical leaders before and after 2012.
+The variables I use as outcomes are the number of full time nurses hired (from AHA), movable equipment purchases, fixed equipment purchases, labor costs, building purchases, and total operating expenses (all from HCRIS).
+For several of these variables, over 90\% of observations are missing in 2010.
+Therefore, I remove this year for the relevant variables.
+The results of these estimations are shown in Table \ref{tab:investment}, and do not present any evidence that hospitals with different leadership teams are changing their investment strategies differentially as a result of the policies.
+
+    \import{Tables}{investment_table.tex}
+
     \section{Conclusion}

### 2025-11-12 Paper/paperdraft_fall2025.tex
Updates from Overleaf

@@
 However, \citeauthor{gupta2021impacts} (\citeyear{gupta2021impacts}) finds that hospitals overall decrease readmissions by around 3 percentage points due to the program, so this reveals a meaningful difference in behavior relative to the overall behavior change.
+Using estimates from \citeauthor{gupta2021impacts} (\citeyear{gupta2021impacts}), a 1 percentage point reduction in readmissions across three penalized conditions corresponds to approximately \$110 million in avoided Medicare reimbursements annually.
+Our results indicate that hospitals led by non-clinical executive teams achieve an additional 0.3 percentage point reduction in readmissions compared to those led by clinical teams.
+Under the same assumptions, this differential accounts for roughly \$33 million in annual avoided spending—equivalent to 30\% of the savings associated with a 1 percentage point reduction for these conditions.
+This back-of-the-envelope calculation reflects the incremental contribution of leadership composition to payer-side savings.
+
     In panel \ref{fig:mort_synth_clinical}, I show the estimated difference in mortality for always clinical and never clinical leadership teams after 2012.
@@
     \end{figure}
-
-    Additionally, I consider the same comparison of clinical and non-clinical leaders, but only for CEOs.
-That is, do the results hold when considering only the most top manager instead of the executive team as a whole?
-I compare hospitals with a CEO trained as a doctor to hospitals without a CEO trained as a doctor using the same synthetic difference-in-differences strategy.
-I present the average treatment effect and standard errors of this estimation in Table .
-There is no difference in response when only considering clinical CEOs as the treatment since the effects are statistically insignificant or smaller in magnitude.
-This highlights the importance of considering executive teams as a whole as opposed to just upper managers.
 
@@
         \label{fig:cont_clinicalsynthdid_mort}
+    \end{figure}
+
+    \subsection{Clinical CEO}
+
+        Additionally, I consider the same comparison of clinical and non-clinical leaders, but only for CEOs.
+That is, do the results hold when considering only the most top manager instead of the executive team as a whole?
+I compare hospitals with a CEO trained as a doctor to hospitals without a CEO trained as a doctor using the same synthetic difference-in-differences strategy.
+I present the average treatment effect and standard errors of this estimation in Table .
+There is no difference in response when only considering clinical CEOs as the treatment since the effects are statistically insignificant or smaller in magnitude.
+This highlights the importance of considering executive teams as a whole as opposed to just upper managers.
+
+    \begin{figure}[ht!]
+     \caption{Effect of CEO Clinical Training on Readmission and Mortality Rates}
+     \centering
+     \begin{subfigure}[b]{0.45\textwidth}
+         \centering
+         \caption{Readmission Rate}
+         \includegraphics[width=\textwidth]{Objects/ceo_read_md_nomd_synth_graph.pdf}
+         \label{fig:ceo_read_synth_clinical}
+     \end{subfigure}
+     \hfill
+     \begin{subfigure}[b]{0.45\textwidth}
+         \centering
+         \caption{Mortality Rate}
+         \includegraphics[width=\textwidth]{Objects/ceo_mort_md_nomd_synth_graph.pdf}
+         \label{fig:ceo_mort_synth_clinical}
+     \end{subfigure}
+        \label{fig:ceo_clinicalsynthdid}
     \end{figure}

### 2025-11-13 Tables/ceo_table.tex
add table of CEO results

\begin{table}[ht!]

\caption{\label{tab:mainanalysis_ceo}Effect of Clinically Trained CEO on Readmission and Mortality Rates}
\centering
\begin{tabular}[t]{lcc}
\toprule
 & Readmission Rate & Mortality Rate\\
\midrule
Clinical CEO x Post 2012 & -0.75 & -0.4\\
 & (0.45) & (0.43)\\
 &  & \\
Observations & 1,630 & 1,630\\
\bottomrule
\multicolumn{3}{l}{\textsuperscript{} Standard errors are clustered at the}\\
\multicolumn{3}{l}{hospital level.}\\
\multicolumn{3}{l}{\textsuperscript{} Signif. codes: 0 '***' 0.001 '**' 0.01 '*'}\\
\multicolumn{3}{l}{0.05 '.' 0.1}\\
\end{tabular}
\end{table}

### 2025-11-13 Paper/paperdraft_fall2025.tex
Updates from Overleaf

@@
 
-    I construct novel information on individual hospital executives by scraping publicly available tax forms, and merge this to publicly available data sets containing other hospital characteristics: the American Hospital Association Survey (AHA), the Centers for Medicare and Medicaid Services (CMS) Hospital Compare data and Case Mix Index files, and Healthcare Cost Report Information System (HCRIS) data.
-Ultimately, I employ this hospital-level data spanning from 2010-2014, which contains executive team characteristics, mortality and readmission rates as measures of quality, and other characteristics.
+    I compile novel data on individual hospital executives by scraping publicly available tax forms, and merge this with publicly available data sets containing other hospital characteristics: the American Hospital Association Survey (AHA), the Centers for Medicare and Medicaid Services (CMS) Hospital Compare data and Case Mix Index files, and Healthcare Cost Report Information System (HCRIS) data.
+Ultimately, I employ hospital-level data spanning 2010-2014 containing executive team characteristics and mortality and readmission rates as measures of quality.
 In this section, I describe the institutional setting and data collection process.
-I include more details on the data construction in \ref{appendixdata}.
+I include more details on the data construction in Appendix \ref{appendixdata}.
 
@@
 
-    Nonprofit hospitals, as with most nonprofits, are typically governed by a board of directors, whose role is complex, but generally is to set broad goals and provide oversight \cite{council_board_roles_2025}.
-The board selects executives, the highest level of management of a firm, to carry out day-to-day operations.
-An executive team usually consists of at least a Chief Executive Officer (CEO) and Chief Financial Officer (CFO), but there is variation in how firms organize executive teams.
-Some hospital executives specialize in health care settings by earning a degree in health care management, or an MBA specific to health care.
-There are executive positions that are often filled by someone with clinical training, such as a Chief Medical Officer (CMO), but medical doctors can also fill other executive positions.
-While some doctors earn additional degrees before stepping into an executive role, this is not a necessary condition to becoming a physician executive.
-
-    Hospital executive teams are understudied partly due to the inaccessibility of granular data.
-I construct a novel data set of nonprofit hospitals in the US which contains names, titles, and positions of all board members and executives tied to the hospital from 2009-2015.
-I gather this data from publicly available Tax Form 990s, which every sufficiently large nonprofit files each year in the US.
-To my knowledge, this is the first large scale gathering of executive names from these forms.\footnote{\citeauthor{brickley2010board} (\citeyear{brickley2010board}) collects compensation data for a select number of hospitals.} All tax-exempt organizations in the US are required to file a Form 990 with Internal Revenue Services (IRS) each year.
+    Nonprofit hospitals, as with most nonprofits, are typically governed by a board of directors, whose role is complex but generally involves setting broad goals and providing oversight (\cite{council_board_roles_2025}).
+The board selects executives, the highest level of management, to carry out day-to-day operations.
+An executive team usually consists of at least a Chief Executive Officer (CEO) and Chief Financial Officer (CFO), but there is variation in how firms organize these teams.
+Some hospital executives specialize in health care administration by earning a degree in health care management, or an MBA specific to health care.
+There are executive positions that are often filled by someone with clinical training (ex.
+Chief Medical Officer), but medical doctors can also fill other executive positions.
+While some doctors earn additional degrees before stepping into an executive role, this is not a necessary condition for becoming a physician executive.
+
+    Hospital executive teams are understudied in part because of the inaccessibility of granular data.
+I compile a novel data set of nonprofit hospitals in the US which contains names, titles, and positions of all board members and executives associated with each hospital in 2009-2015.
+I gather this data from publicly available Tax Form 990s.
+To my knowledge, this is the first large-scale gathering of executive names from these forms.\footnote{\citeauthor{brickley2010board} (\citeyear{brickley2010board}) collect compensation data for a select number of hospitals.} All tax-exempt organizations in the US are required to file a Form 990 with Internal Revenue Services (IRS) each year.
 There are different types of forms, but any organization grossing over \$200,000 must file the most extensive Form 990.
-Sections of this form include a statement of revenue, statement of functional expenses, a balance sheet, and, as used in this project, a list of all key employees, executives, and board members.
-
-    For each hospital in the tax form data, I merge to hospital matches in the AHA data based on name and location.\footnote{I assess the observable differences between matched and non-matched hospitals in Appendix \ref{app:matched}, and conclude that the samples are overall similar apart from an under-sampling of hospitals that belong to systems.} For all hospitals that can be matched to a general hospital in the AHA data, I use optical character recognition (OCR) methods to extract the names, titles, and positions of each listed executive from its Tax Form 990.
+Sections of this form include a statement of revenue, functional expenses, a balance sheet, and, as used in this project, a list of key employees, executives, and board members.
+
+    For each hospital in the tax forms, I merge with hospital records in the AHA data that match based on name and location.\footnote{I assess the observable differences between matched and non-matched hospitals in Appendix \ref{app:matched}, and conclude that the samples are overall similar apart from an under-sampling of hospitals that belong to systems.} For matched hospitals, I use optical character recognition (OCR) methods to extract the names, titles, and positions of each individual listed in the tax form.
 While board members are an important aspect of hospital organization, I do not include them in this analysis as they are beyond the scope of this paper.
-I extract text spanning 2009-2015, but I retain a much larger sample of hospitals by limiting to hospitals that have text information in 2010-2014.
-This yields 847 nonprofit hospitals that are (1) matched to AHA data, and (2) contain complete leadership information in each year from 2010-2014.
-For clarity, I include a table outlining the number of observations in each sample limitation in Table \ref{tab:sample}.
+I extract text spanning 2009-2015, but I retain a much larger sample of hospitals by restricting the sample to hospitals with text information in 2010-2014.
+This yields 847 nonprofit hospitals that are matched to the AHA survey and contain complete leadership information in each year from 2010-2014.
+For clarity, I include a table outlining the number of observations in each sample restriction in Table \ref{tab:sample}.
 
@@
     \centering
-    \caption{\label{tab:sample}Observations in Sample Limitations}
+    \caption{\label{tab:sample}Observations in Sample Restrictions}
     \centering
@@
     \toprule
-    Sample & Num. Observations\\
+    Sample & Observations\\
     \toprule
@@
     \addlinespace
-    Hospital EINs Match to AHA Hospital Only & 1,152\\
+    Hospital EINs Matched to AHA Hospital Only & 1,152\\
     \addlinespace
-    Hospitals with Leadership Info Years 2010-2014 & 847\\
+    Hospitals with Leadership Info, 2010-2014 & 847\\
     \bottomrule
@@
     Clinical training need not be limited to physicians.
-However, those are the most common and verifiable clinicians in the data.
-For the main analysis, I verify clinical training of those who are identified as doctors, and obtain additional information about clinically trained executives by matching names to the National Plan and Provider Enumeration System (NPPES) database of all registered physicians.
-I present a table of means for various characteristics of doctor executives in the data in Table \ref{doc_sumstats}.
-In total, there are 813 clinically trained executives.
+However, physicians are the most common and verifiable clinicians in the data.
+I verify clinical training for individuals identified as doctors, and obtain additional information about clinically trained executives by matching names to the National Plan and Provider Enumeration System (NPPES) database of all registered physicians.
+I present a table of means for various characteristics of doctor executives in Table \ref{doc_sumstats}.
+In total, there are 954 clinically trained executives.
 On average, they are 52 years old, and 11\% of clinically trained executives are female.
-While it is more likely they are a Chief Medical Officer than a Chief Executive Officer, still less than a third are CMOs.
+While it is more likely they are a Chief Medical Officer than a Chief Executive Officer, less than one-third hold CMO positions.
 The most common specialty is internal medicine.
-While I cannot perfectly verify nurse experience in the data, in Appendix \ref{app:nurses}, I include executives who are marked as registered nurses in the tax forms.
-The results are similar to those of the main analysis, which is not surprising given that only 0.5\% of executives have a nursing title.
+While I cannot perfectly verify nurse experience in the data, in Appendix \ref{app:nurses}, I include executives who are labeled as registered nurses in the tax forms.
+The results are similar to those of the main analysis, which is unsurprising given that only 0.5\% of executives have a nursing title.
 
@@
 
-    As the organization of specific c-suite positions can vary drastically across hospitals, I focus on executive teams as a whole instead of specific positions.
-It is not uncommon for two hospitals to have the same position but the tasks performed be drastically different, or to have two distinct positions that essentially perform the same tasks.
-I utilize various hospital-level characterizations based on their executive team, such as the existence of a clinically trained executive, the number of clinically trained executives, the number of total executives, and whether the hospital has a Chief Medical Officer.
+    As the organization of specific C-suite positions can vary drastically across hospitals, I focus on executive teams as a whole rather than individual positions.
+It is not uncommon for two hospitals to have the same position but the tasks performed to be drastically different, or to have two distinct positions that essentially perform the same tasks.
+I use various hospital-level characterizations based on their executive team, such as the existence of a clinically trained executive, the number of clinically trained executives, the number of total executives, and whether the hospital has a Chief Medical Officer.
 
@@
 
-    The purpose of this paper is to investigate hospital behavior in the midst of financial incentives that change how hospitals receive payment depending on quality.
-Two programs were passed as part of the Affordable Care Act that focused on pay-for-performance incentives for hospitals: the Hospital Readmissions Reduction Program (HRRP), and the Hospital Value Based Purchasing Program (HVBP).
-HRRP focused on penalizing hospitals with poor quality, and HVBP focused on rewarding hospitals with high quality or show improvement in quality.
-
-    In October 2011, the Center for Medicare and Medicaid Services (CMS) released a set of rules under HRRP mandating penalties for hospitals with above average readmission rates.
+    The purpose of this paper is to investigate hospital behavior under financial incentives that change how hospital quality affects payment.
+Two programs under the Affordable Care Act that focused on pay-for-performance incentives for hospitals: the Hospital Readmissions Reduction Program (HRRP) and the Hospital Value-Based Purchasing Program (HVBP).
+HRRP focused on penalizing hospitals with poor quality, and HVBP focused on rewarding hospitals with higher quality.
+
+    In October 2011, the Centers for Medicare \& Medicaid Services (CMS) released a set of rules under HRRP mandating penalties for hospitals with above-average readmission rates.
 The goal of HRRP is to lower readmissions through better care coordination, less initial stay complications, and better post-care instructions.
-Beginning in October 2012, hospitals with higher readmission rates than the national average in pneumonia, heart failure, or AMI (after adjusting for demographic characteristics) receive a fixed lower reimbursement rate for all Medicare patients seen in their hospital.
-Penalties are given in the form of a fixed rate reduction of 1-3\% for every Medicare patient regardless of the condition.
-Excess readmission rates are calculated using a rolling look-back period of 3 years to determine whether the hospital is penalized.
-Therefore, hospitals had incentive to react immediately once details of the program were announced in October of 2011.
+Beginning in October 2012, hospitals with higher readmission rates than the national average in pneumonia, heart failure, or acute myocardial infarction (AMI) received a fixed lower reimbursement rate for all Medicare patients seen in their hospital.
+Penalties are given in the form of a fixed-rate reduction of 1-3\% for every Medicare patient regardless of the condition.
+Excess readmission rates are calculated using a rolling, 3 year look-back period to determine whether the hospital is penalized.
+Therefore, hospitals had an incentive to react immediately once details of the program were announced in October of 2011.
 
     The HVBP Program instead rewards hospitals with high quality or significant improvement in quality.
-Specifically, CMS deducts Medicare payments by 2\% from all eligible hospitals, collects this sum, and divides it among the rewarded hospitals.
-Several quality and cost measures surrounding safety, efficiency, cost reductions, clinical outcomes, and community engagement are combined to create a single score metric for each hospital.
-Hospitals are then compared to the average and are rewarded for being above average quality or for showing improvement (\cite{CMS_2023}).
-
-    I focus on outcomes of readmission and mortality rates, which are publicly available at the hospital-condition level in CMS Hospital Compare.
-I combine rates for pneumonia, AMI, and heart failure (the relevant HRRP conditions) into a weighted average based on the number of patients in each condition.
+Specifically, CMS reduces Medicare payments by 2\% for all eligible hospitals, collects this sum, and redistributes it among the rewarded hospitals.
+Several quality and cost measures surrounding safety, efficiency, cost reductions, clinical outcomes, and community engagement are combined to create a single metric for each hospital.
+Hospitals are then compared to the average and are rewarded for being above average quality or showing improvement (\cite{CMS_2023}).
+
+    I focus on outcomes of readmission and mortality, which are publicly available at the hospital-condition-level in CMS Hospital Compare.
+I combine rates for pneumonia, AMI, and heart failure (the relevant HRRP conditions), weighted by the number of patients in each condition.
 
@@
 
-    In Table \ref{tab:sumstats_samples_stable}, I present means of relevant variables for several hospital samples used throughout the paper.
-First, in column (1), are means for hospitals who have a clinical executive at any point in time.
-Column (2) contains means for hospitals who have a clinical executive for the entire sample period, and column (3) shows means for hospitals who never have a clinical executive in the sample period.
-The main analysis compares hospitals in columns (2) and (3), and supplemental analyses leverage hospitals in column (1).
+    In Table \ref{tab:sumstats_samples_stable}, I present mean values of relevant variables for several hospital samples used throughout the paper.
+First, in column (1), are means for hospitals that have a clinical executive at any point in time.
+Column (2) contains means for hospitals that have a clinical executive for the entire sample period, and column (3) shows means for hospitals that never had a clinical executive in the sample period.
+The main analysis compares hospitals in columns (2) and (3), and supplemental analyses draw on hospitals in column (1).
 
@@
 
-    Comparing columns (2) and (3), hospitals that never have a clinical executive are smaller, less likely to be an academic medical center, and less likely to be affiliated with a system.
-Further, they have smaller executive teams on average by 1-2 people, and are less likely to have anyone acting as a designated CMO.
-Finally, in regards to exposure to pay-for-performance programs, hospitals who never have clinical executives are less likely to receive benefits from the HVBP Program, but seem more similar to hospitals with clinical leadership in terms of penalties from HRRP.
+    Comparing columns (2) and (3), hospitals that never had a clinical executive are smaller, less likely to be an academic medical center, and less likely to be affiliated with a system.
+Further, they have smaller executive teams by an average of 1-2 people, and are less likely to have a designated CMO.
+Finally, with regard to exposure to pay-for-performance programs, hospitals that never had clinical executives are less likely to receive benefits from the HVBP Program, but seem more similar to hospitals with clinical leadership in terms of penalties from HRRP.
 
     Next, I plot weighted average readmission rates, mortality rates, and patient case mix index across time for different hospital types in Figure \ref{fig:outcomes_graph}.
-The effective start year of pay-for-performance incentives is shown as a dotted line in 2012.
-Average readmission rates trend similarly prior to 2012, and all hospital types exhibit a decrease in raw readmission rates after 2012.
-However, there is a faster rate of decrease for hospitals who have a clinically trained executive at some point, or who never have a clinically trained executive.
-When considering mortality, there is not a drastic change in 2012 for any hospital type, but hospitals who always have a clinical executive continue on an increasing trend in mortality where the other hospitals types remain more flat.
+The effective start year of pay-for-performance incentives is indicated as a dotted line in 2012.
+Average readmission rates follow similar trends prior to 2012, and all hospital types exhibit a decrease in raw readmission rates after 2012.
+However, there is a faster rate of decrease for hospitals that have a clinically trained executive at some point.
+When considering mortality, there is not a drastic change in 2012 for any hospital type, but hospitals that always have a clinical executive continue on an increasing trend in mortality while the other hospitals types remain relatively flat.
 
@@
 
-    \section{Effect of Clinically Trained Executives on Response to Financial Incentives on Quality}\label{sec:clinical}
-
-    I test empirically whether having clinically trained executives matters in how hospitals respond to the pay-for-performance policies effective in 2012.
-The quality metrics targeted by the programs are readmission and mortality rates, so I focus on these measures as outcomes.
-To first analyze a clear control and treatment group, I limit the sample to hospitals who always have clinical executives or never have clinical executives.
-That is, the hospitals in this analysis do not change their propensity to hire a clinical executive during the sample period.\footnote{An important assumption for this analysis is that this sample selection is not correlated with the program enactment.
-I investigate this assumption in Section \ref{sec:changes}, and find no evidence that hospitals are more or less likely to change their propensity to hire a clinical executive because of the pay-for-performance programs.} I employ a two-way-fixed-effects strategy as follows:
+    \section{Effect of Clinical Leadership on Quality Incentive Response}\label{sec:clinical}
+
+    I test empirically whether having clinically trained executives affects how hospitals respond to the pay-for-performance policies effective in 2012.
+The quality metrics targeted by the programs are readmission and mortality rates, and therefore I focus on these measures as outcomes.
+To establish a clear control and treatment group, I limit the sample to hospitals that always have clinical executives or never have clinical executives.
+That is, the hospitals in this analysis do not change their propensity to hire a clinical executive during the sample period.\footnote{An important assumption underlying this analysis is that sample selection is not correlated with the program enactment.
+I investigate this assumption in Section \ref{sec:changes}, and find no evidence that hospitals are more or less likely to change their propensity to hire a clinical executive because of the pay-for-performance programs.} I employ a two-way fixed effects strategy as follows:
 
@@
 
-    \noindent where the variable $y_{ht}$ is one of the outcome variables capturing readmission or mortality rates discussed in Section \ref{sec:hrrp}, and $\gamma_h$ and $\delta_t$ are hospital and time fixed effects, respectively.
-The coefficient of interest is $\beta$, capturing the effect of having a clinician on change in quality after 2012, when pay-for-performance policies were rolled out.
-All hospitals had incentive to improve quality in response to the policy change, as penalties and payments are determined by comparing quality among hospitals.
-Thus, I leverage the timing of the policy change, rather than the policies themselves.
+    \noindent where the variable $y_{ht}$ denotes outcome variables (readmission or mortality rates) discussed in Section \ref{sec:hrrp}, and $\gamma_h$ and $\delta_t$ are hospital and time fixed effects, respectively.
+The coefficient of interest is $\beta$, capturing the effect of having a clinician on changes in quality after 2012, following the implementation of pay-for-performance policies.
+All hospitals had an incentive to improve quality in response to the policy change, as penalties and payments are determined by comparing quality among hospitals.
+Thus, I exploit the timing of the policy change rather than variation in policy design.
 
     One may be concerned about potential differences in quality trends between clinical and non-clinical executive team hospitals leading to a violation in the parallel trends assumption.
-To account for this, I ultimately implement a synthetic difference-in-differences strategy (\cite{arkhangelsky2021synthetic}), which weights more heavily any non-clinical team hospitals with pre-trends similar to clinical team hospitals in the outcome being considered, essentially guaranteeing no differential pre-trends between always and never clinical teams.
-Using this strategy, the never-clinical group consists of approximately 240 hospitals, and I show the fairly uniform distribution of their weights in Appendix \ref{app:sumstats}.
+To account for this, I ultimately implement a synthetic difference-in-differences strategy (\cite{arkhangelsky2021synthetic}), which assigns greater weight to non-clinical team hospitals with pre-trends similar to clinical team hospitals in the outcome being considered, effectively ensuring no differential pre-trends between always and never clinical teams.
+Using this strategy, the never-clinical group consists of approximately 240 hospitals, and I show the relatively uniform distribution of their weights in Appendix \ref{app:sumstats}.
 As a robustness check, I also pre-select the sample of always and never clinical hospitals using the Coarsened Exact Matching (CEM) procedure as described by \citeauthor{azoulay2010superstar} (\citeyear{azoulay2010superstar}).
@@
 For brevity, I only present the synthetic difference-in-differences results here.
-In Appendix \ref{app:otherestimations}, I show that the results are robust to using standard two-way-fixed-effects, matching, or synthetic control.
-
-    For identification, I assume that under the weighted clinical and non-clinical hospital composition, trends would have continued to be parallel absent the incentive change in 2012.
-This includes an implicit assumption that no other event occurred in 2012 correlated with both the outcomes and composition of leadership team.
-Finally, I assume that there is no anticipation of the policy change before 2012.
-The rules of the policies were announced in October of 2011, so 2012 was the first full year that hospitals could meaningfully respond to the policies.
-
-    The estimated difference in readmission and mortality rates between hospitals with and without clinical executives are shown in Figure \ref{fig:clinicalsynthdid}.
-Estimate $\beta$ is represented in the difference in slope after 2012 when pay-for-performance incentives began.
-Panel \ref{fig:read_synth_clinical} shows the estimated difference in readmission rates, and reveals that non-clinical teams decrease readmissions by .3 (standard error .13) more than clinical teams.
-This is a relatively small magnitude given an average readmission rate of 21.
+In Appendix \ref{app:otherestimations}, I show that the results are robust to using standard two-way fixed effects, matching, and synthetic control.
+
+    For identification, I assume that given the weighted clinical and non-clinical hospital composition, trends would have continued in a parallel manner in the absence of the incentive change in 2012.
+This includes an implicit assumption that no other event occurred in 2012 correlated with both the outcomes and composition of the leadership team.
+Finally, I assume that there is no anticipation of the policy change prior to 2012.
+The rules of the policies were announced in October 2011, so 2012 was the first full year that hospitals could meaningfully respond to the policies.
+
+    The estimated difference in readmission and mortality rates between hospitals with and without clinical executives is shown in Figure \ref{fig:clinicalsynthdid}.
+Estimate $\beta$ is represented by the difference in slope after pay-for-performance incentives began.
+Panel \ref{fig:read_synth_clinical} shows the estimated difference in readmission rates, and reveals that non-clinical teams decrease readmissions by 0.3 (standard error 0.13) more than clinical teams.
+This is a relatively small magnitude given an average readmission rate of 21\%.
 However, \citeauthor{gupta2021impacts} (\citeyear{gupta2021impacts}) finds that hospitals overall decrease readmissions by around 3 percentage points due to the program, so this reveals a meaningful difference in behavior relative to the overall behavior change.
@@
 Our results indicate that hospitals led by non-clinical executive teams achieve an additional 0.3 percentage point reduction in readmissions compared to those led by clinical teams.
-Under the same assumptions, this differential accounts for roughly \$33 million in annual avoided spending—equivalent to 30\% of the savings associated with a 1 percentage point reduction for these conditions.
+Under the same assumptions, this differential accounts for approximately \$33 million in annual avoided spending—equivalent to 30\% of the savings associated with a 1 percentage point reduction for these conditions.
 This back-of-the-envelope calculation reflects the incremental contribution of leadership composition to payer-side savings.
@@
 
-    \subsection{Including Nurses in the Sample}\label{app:nurses}
+    \subsection{Including Nurses as Clinical Executives}\label{app:nurses}
+
+    Everything in this section is identical to the main analysis, apart from the definition of clinical executive.
+Here, I include executives labeled as a registered nurse in the tax forms as a clinical executives.
+These are relatively rare, as only of executives say they are a registered nurse.
+Unlike doctors, there is no publicly available database of all nurses, so the cost of verifying this type of clinical training is much higher.
+Therefore, in the main body of the paper I only consider doctor executives.
+However, the results when including nurses are identical, as I show in Figure \ref{fig:nurse_results}.
+
+    \begin{figure}[ht!]
+     \caption{\label{fig:nurse_results}Effect of Clinical Training on Readmission/Mortality Rates, Including Nurses}
+     \centering
+     \begin{subfigure}[b]{0.45\textwidth}
+         \centering
+         \caption{Readmission Rates}
+         \includegraphics[width=\textwidth]{Objects/nurses_read_md_nomd_synth_graph.pdf}
+
+     \end{subfigure}
+     \hfill
+     \begin{subfigure}[b]{0.45\textwidth}
+         \centering
+         \caption{Mortality Rates}
+         \includegraphics[width=\textwidth]{Objects/nurses_mort_md_nomd_synth_graph.pdf}
+
+     \end{subfigure}
+     \end{figure}
 

### 2025-11-13 Paper/paperdraft_fall2025.tex

@@
 
-    In panel \ref{fig:mort_synth_clinical}, I show the estimated difference in mortality for always clinical and never clinical leadership teams after 2012.
-While the magnitude of the difference is similar to that for readmission rates (relative to the mean), mortality is a noisy measure, and I cannot rule out null effects.
+    In Figure \ref{fig:mort_synth_clinical}, I show the estimated difference in mortality for always-clinical and never-clinical leadership teams after 2012.
+While the magnitude of the difference is similar to that of readmission rates (relative to the mean), mortality is a noisy measure, and I cannot rule out the possibility of a null effect.
 
@@
 
-    Additionally, I estimate the intensive margin of clinical leadership: whether the fraction of clinically trained executives matters for response to financial incentive changes.
-I categorize hospitals with clinically trained executives into two bins: above the median and below median fraction of clinically trained executives.
+    Additionally, I estimate the intensive margin of clinical leadership: whether the fraction of clinically trained executives affects the response to financial incentive changes.
+I categorize hospitals into two bins: above the median and below the median fraction of clinically trained executives.
 Again, I present the unweighted version of the synthetic difference-in-differences that I ultimately employ:
@@
 
-    \noindent where $\rho$ is the fraction of clinical executives on hospital $h$'s executive team, $\eta$ is the median fraction of clinically trained executives, which is .2 in this sample, and $\gamma$ and $\delta$ are hospital and time fixed effects.
-
-    I present the results of this estimation with readmission rates as the outcome in Figure \ref{fig:cont_clinicalsynthdid}.
-In panel \ref{fig:belowmed_read_synth_clinical}, I compare executive teams with no clinical training to those with a fraction of clinically trained executives up to the median value, 20\%.
-With an average treatment effect of .21 (standard error .31), hospitals having this range of clinically trained executives are not driving the main finding.
+    \noindent where $\rho$ is the fraction of clinical executives on hospital $h$'s executive team, $\eta$ is the median fraction of clinically trained executives, which is 0.2 in this sample, and $\gamma$ and $\delta$ are hospital and time fixed effects.
+
+    I present the estimation results for readmission rates as the outcome in Figure \ref{fig:cont_clinicalsynthdid}.
+In Figure \ref{fig:belowmed_read_synth_clinical}, I compare executive teams with no clinical training to those with a fraction of clinically trained executives up to the median value (20\%).
+With an average treatment effect of 0.21 (standard error 0.31), hospitals having this range of clinically trained executives do not appear to drive the main finding.
 However, in panel \ref{fig:abovemed_read_synth_clinical}, I compare hospitals with no clinically trained executives to hospitals with a fraction of clinically trained executives over the median of 20\%.
-I find that hospitals without clinically trained executives decrease readmissions by .6 (standard error .14) more than hospitals with clinically trained executives.
-Thus, while the entire readmission rate response is not driven by highly clinical teams, the effect is exacerbated for these teams that have more clinicians than the median hospital.\footnote{This result holds even when limiting the sample to hospitals with a similar total executive team sizes (Appendix \ref{app:binned})}
+I find that hospitals without clinically trained executives decrease readmissions by 0.6 (standard error 0.14) more than hospitals with clinically trained executives.
+Thus, while the entire readmission rate response is not driven by teams with a higher proportion of clinically trained executives, the effect is exacerbated for hospitals that have a higher proportion of clinicians.\footnote{This result holds even when limiting the sample to hospitals with a similar total executive team size (Appendix \ref{app:binned})}
 
@@
 
-    I present the estimation results for mortality as an outcome in Figure \ref{fig:cont_clinicalsynthdid_mort}.
-Unsurprisingly given the noisiness in the main findings for mortality, it is unclear whether there is any intensive margin effect for mortality.
-Both average treatment effect magnitudes are lower than that for the main result with high standard errors.
-Thus, the intensive margin does not seem to matter as much as it does for decreasing readmission rates.
+    I present the estimation results for mortality in Figure \ref{fig:cont_clinicalsynthdid_mort}.
+Unsurprisingly given the noisiness in the main findings for mortality, it is unclear whether there is any intensive margin effect.
+Both average treatment effect estimates are smaller than those in the primary analysis and are associated with large standard errors.
+Thus, the intensive margin appears less influential for mortality than for reducing readmission rates.
 
@@
         Additionally, I consider the same comparison of clinical and non-clinical leaders, but only for CEOs.
-That is, do the results hold when considering only the most top manager instead of the executive team as a whole?
-I compare hospitals with a CEO trained as a doctor to hospitals without a CEO trained as a doctor using the same synthetic difference-in-differences strategy.
-I present the average treatment effect and standard errors of this estimation in Table .
+Specifically, I examine whether the results hold when focusing solely on the CEO rather than the entire executive team.
+I compare hospitals led by a clinically trained CEO to those led by a non-clinical CEO using the same synthetic difference-in-differences strategy.
+I present the average treatment effect and standard errors of this estimation in Table \ref{tab:mainanalysis_ceo}.
 There is no difference in response when only considering clinical CEOs as the treatment since the effects are statistically insignificant or smaller in magnitude.
@@
 
-    \begin{figure}[ht!]
-     \caption{Effect of CEO Clinical Training on Readmission and Mortality Rates}
-     \centering
-     \begin{subfigure}[b]{0.45\textwidth}
-         \centering
-         \caption{Readmission Rate}
-         \includegraphics[width=\textwidth]{Objects/ceo_read_md_nomd_synth_graph.pdf}
-         \label{fig:ceo_read_synth_clinical}
-     \end{subfigure}
-     \hfill
-     \begin{subfigure}[b]{0.45\textwidth}
-         \centering
-         \caption{Mortality Rate}
-         \includegraphics[width=\textwidth]{Objects/ceo_mort_md_nomd_synth_graph.pdf}
-         \label{fig:ceo_mort_synth_clinical}
-     \end{subfigure}
-        \label{fig:ceo_clinicalsynthdid}
-    \end{figure}
+   \import{Tables}{ceo_table.tex}
 

### 2025-11-14 Tables/selection_table.tex
add table for patient population instead of graphs

\begin{table}[ht!]

\caption{\label{tab:selection}Effect of Clinically Trained Executive on Patient Population}
\centering
\begin{tabular}[t]{lccc}
\toprule
 & CMI & Frac. Medicare & Frac. Medicaid\\
\midrule
Clinical Exec x Post 2012 & -0.03$^{***}$ & 0 & 0\\
 & (0.01) & (0) & (0.01)\\
 &  &  & \\
Observations & 1,645 & 2,490 & 2,390\\
\bottomrule
\multicolumn{4}{l}{\textsuperscript{} Standard errors are clustered at the hospital level.}\\
\multicolumn{4}{l}{\textsuperscript{} Signif. codes: 0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1}\\
\end{tabular}
\end{table}

### 2025-11-14 Paper/paperdraft_fall2025.tex

@@
 
-        Additionally, I consider the same comparison of clinical and non-clinical leaders, but only for CEOs.
-Specifically, I examine whether the results hold when focusing solely on the CEO rather than the entire executive team.
-I compare hospitals led by a clinically trained CEO to those led by a non-clinical CEO using the same synthetic difference-in-differences strategy.
-I present the average treatment effect and standard errors of this estimation in Table \ref{tab:mainanalysis_ceo}.
-There is no difference in response when only considering clinical CEOs as the treatment since the effects are statistically insignificant or smaller in magnitude.
-This highlights the importance of considering executive teams as a whole as opposed to just upper managers.
+        Additionally, I consider the same comparison of clinical and non-clinical leaders, restricted to CEOs.
+Specifically, I examine whether the results hold when focusing exclusively on the CEO rather than the entire executive team.
+This analysis is motivated by the literature, which often emphasizes the CEO as the central figure in organizational leadership.
+Therefore, this analysis serves to assess whether team-level analysis offers different insights compared to CEO-focused analysis.
+
+        I compare hospitals led by a clinically trained CEO to those led by a non-clinical CEO by applying the same synthetic difference-in-differences strategy described in Section \ref{sec:clinical}.
+I present the average treatment effect and standard errors in Table \ref{tab:mainanalysis_ceo}.
+For both readmission and mortality rates, the estimated effect of having a clinical CEO after 2012 is statistically insignificant.
+Point estimates suggest clinical CEOs may improve quality slightly more than non-clinical CEOs, though the results lack statistical precision.
+This highlights the importance of considering executive teams as a whole rather than just upper managers.
 
@@
 
-    While uncovering that there is some foundational difference in clinical and non-clinical hospital leaders is important in its own right, it is especially relevant to understand why this difference might come about.
-In this Section, I present several theoretical predictions consistent with the observed difference in quality after pay-for-performance incentives, and  test the predictions empirically.
-First, I investigate whether clinicians are actively managing hospitals differently or whether they are a signal of underlying preferences of the hospital by leveraging the timing of clinical leaders relative to the incentives in Section \ref{sec:sig_man}.
-Second, the difference could be driven by the altruism, or patient-centered focus, of clinicians.
-I investigate this in Section \ref{sec:forprofit}, comparing the response of nonprofits with different leadership teams to the response of for-profit hospitals on average.
-Third, I consider the altruism versus expertise argument using a heterogeneity analysis of different physician specialties in Section \ref{sec:expertise}.
-In Section \ref{sec:selection}, I consider whether different types of executives might have different propensities to engage in "gaming" through selective patient practices.
-Finally, I analyze whether different executives change investment behavior in response to the incentives in Section \ref{sec:investment}.
+    While identifying foundational differences between clinical and non-clinical hospital leaders is important in its own right, understanding why these differences arise is especially relevant.
+In this section, I present several theoretical predictions consistent with the observed differences in quality following pay-for-performance incentives and test these predictions empirically.
+First, I investigate whether clinicians actively manage hospitals differently or simply signal underlying hospital preferences.
+I do this by by leveraging the timing of clinical leaders relative to the incentives in Section \ref{sec:sig_man}.
+Second, the difference could be driven by altruism, or patient-centered focus of clinicians.
+I investigate this in Section \ref{sec:forprofit}by comparing nonprofit hospitals with varying leadership teams to for-profit hospitals.
+Third, I consider whether differences reflect altruism or expertise through a heterogeneity analysis of physician specialties in Section \ref{sec:expertise}.
+In Section \ref{sec:selection}, I consider whether different types of executives may differ in their propensity to engage in ‘gaming’ through selective patient practices.
+Finally, I analyze whether whether executives alter aggregate investment behavior in response to incentives in Section \ref{sec:investment}.
 
@@
 
-    Given that having clinically trained executives does affect hospital response to financial incentives on quality, an important distinction is whether the existing clinical leaders \textit{reveal} the underlying objectives of the hospital, or \textit{manage} the hospital differently.
-In the main analysis, I limited the sample of nonprofit hospitals to those that don't change their propensity to hire a clinically trained executive over the sample period, which combines any signaling and managing effects.
+    Given that having clinically trained executives affects hospitals' response to financial incentives on quality, an important distinction is whether the existing clinical leaders \textit{reveal} the underlying objectives of the hospital or \textit{manage} the hospital differently.
+In the main analysis, I limited the sample of nonprofit hospitals to those that do not alter their propensity to hire a clinically trained executive over the sample period, which combines signaling and managing effects.
 However, under the assumption that hiring or firing these types of executives is not correlated with the pay-for-performance policy changes, such changes can be leveraged to decompose the effect into signaling versus managing.
-Specifically, if the difference is driven by underlying preferences of the hospital, the observed difference in quality will still hold whether the hospital had a clinical executive at any point in time, even if they were not a leader in 2012 when the policies took effect.
-Alternatively, if the observed difference is driven by managing, then the difference should only be observed when there is a clinician in leadership during the policy enactment.
-Since this decomposition depends heavily on the assumption that clinical executive hiring is not endogenous, I first investigate leadership changes in response to the policy change.
-I provide details of the decomposition estimation in Section \ref{sec:decomp}
+If the difference is driven by underlying preferences of the hospital, the observed difference in quality will still hold whether the hospital ever had a clinical executive, even if they were not in leadership in 2012 when the policies took effect.
+Alternatively, if the observed difference is driven by managing, then the difference should only appear when a clinician is in leadership during policy enactment.
+Because this decomposition relies heavily on the assumption that clinical executive hiring is exogenous, I first examine leadership changes in response to the policy.
+I provide details of the decomposition estimation in Section \ref{sec:decomp}.
 
@@
 
-    Decomposing the observed difference in response based on timing is dependent on executive team changes not being correlated with the financial incentive changes.
-First, I present raw means over time that represent the fraction of hospitals who hire a clinical executive, fire a clinical executive, or have any change in their propensity to have a clinical executive in Figure \ref{fig:change_means}.
-There do not seem to be any drastic changes in 2011 or 2012 when the policies were becoming relevant to hospitals.
+    Decomposing the observed difference in response based on timing is dependent on executive team changes not being correlated with the financial incentive.
+First, I present raw means over time showing the fraction of hospitals that hire a clinical executive, fire a clinical executive, or change their propensity to have a clinical executive in Figure \ref{fig:change_means}.
+The proportion of hospitals that make changes to their leadership team is small, around 15\% on average.
+Additionally, the data show no drastic changes in 2011 or 2012 when the policies became relevant to hospitals, especially when aggregating any change.
 
@@
 
-    Next, I assess whether this assumption is reasonable by analyzing how executive team changes are occurring over time for certain groups of hospitals.
-I estimate the following two-way-fixed-effects specification:
+    Next, I assess whether this assumption holds by analyzing executive team changes over time for specific hospital groups. I estimate the following two-way fixed effects specification:
 
@@
 
-    where the variable $\text{Program Exposed}_{h}$ has three definitions.
-First, an indicator for whether the hospital was ever penalized under HRRP.
-Second, an indicator for whether the hospital ever received payments under HVBP.
-And finally, I leave out the Program Exposed variable to simply see how changes are correlated with given years for all hospitals.
-The outcome variable change$_{ht}$ is an indicator for whether the hospital changes the number of clinical executives in a given year.
-The estimates from this analysis are presented in Figure \ref{fig:change_analysis}.
-All estimates are not statistically different from zero, indicating that expectations of program exposure do not lead to changes in propensity to hire a clinical executive.
-Thus, it seems unlikely that endogenous team formation is biasing the estimates.
+    \noindent where the variable $\text{Program Exposed}_{h}$ takes three forms.
+First, I define it as an indicator for whether the hospital ever faced penalties under HRRP.
+Second, I define it as an indicator for whether the hospital ever received payments under HVBP.
+Finally, I omit the Program Exposed variable to examine how changes correlate with calendar years for all hospitals.
+The outcome variable change$_{ht}$ indicates whether the hospital changes the number of clinical executives in a given year.
+Figure \ref{fig:change_analysis} shows the estimates from this analysis.
+All estimates remain statistically indistinguishable from zero, indicating that expectations of program exposure do not drive changes in the propensity to hire a clinical executive.
+These results suggest that endogenous team changes are unlikely to bias the estimates.
 
@@
 
-    To disentangle clinical executives as signals of underlying hospital preferences or a different type of hospital manager, I carefully define the ``treat" and comparison group of hospitals in the following estimating equation.
+    To disentangle clinical executives as signals of underlying hospital preferences or a different type of hospital manager, I carefully define the treatment and comparison group of hospitals in the following estimating equation.
 For clarity, I also present the specification details in Figure \ref{fig:decomp_spec}.
-If clinically trained leaders are simply signals of underlying objectives, it shouldn't matter when hospitals have hired a clinical executive.
-For example, take two hypothetical hospitals.
-Hospital A has a clinically trained executive from 2010-2011.
-Hospital B has a clinical executive from 2010-2013.
-If clinically trained executives are a signal of underlying objectives, these two hospitals should respond similarly to pay-for-performance objectives despite only one of them having a clinical executive in 2012.
-However, if clinically trained leaders manage the hospital in a unique way, these two hospitals would respond differently since only one of them has clinical leadership in 2012 when the policy changes occurred.
 
@@
 
-    Thus, in one specification I compare hospitals who ever have a clinically trained executive to hospitals who never have a clinically trained executive.
-This captures the signaling effect.
-In another specification, I remove all hospitals who never had a clinically trained executive, and compare hospitals who had a clinically trained executive in 2012 to those who had a clinically trained executive in any year except 2012, capturing a managing effect.
-
     \begin{figure}[ht!]
@@
     \caption{\label{fig:decomp_spec}Decomposition Model Specification Details}
-
     \begin{tabular}{| m{18em} |}
@@
      \end{tabular}
-
     \end{center}
@@
 
-    As in the previous analyses, I estimate the synthetic difference-in-differences version of each of these specifications.
-That is, I weight control group units more heavily when their pre-trends are similar to the treated group units, and I weight more heavily time periods that balance pre and post period trends for the control group.
-I have discussed the assumptions of parallel trends, no anticipation, and no other confounding events in both Sections \ref{sec:clinical} and \ref{sec:forprofit}.
-These assumptions are also necessary here.
-I also assume that executive team changes are not endogenous with the policy changes, which I explore in detail in Section \ref{sec:changes}.
-
-    The estimates of each decomposition specification are shown in Table \ref{tab:MD_noMD_readmort_decomp_synth}.
-The readmission rate estimates are in columns (1) and (2), and mortality rate estimates are in columns (3) and (4).
-The first row compares ever clinical executive hospitals to never clinical executive hospitals, and the second row compares hospitals with a clinical executive in 2012 to hospitals who have a clinical executive in a year other than 2012.
-While there is no statistically significant estimate, the coefficients of the managing effect are very similar to the full effect found in Section \ref{sec:clinical}.
-The coefficients for the signaling effect are both statistically insignificant and close to zero in magnitude.
-Thus, I can rule out that clinical executives are simply a signal of underlying hospital objectives.
+      If clinically trained leaders only signal underlying objectives, the timing of hiring a clinical executive should not matter.
+For example, consider two hospitals: Hospital A employs a clinically trained executive from 2010 to 2011, while Hospital B employs one from 2010 to 2013.
+If clinically trained executives signal underlying objectives, both hospitals should respond similarly to pay-for-performance policies, even though only one has a clinical executive in 2012.
+However, if clinically trained leaders manage hospitals differently, the hospitals should respond differently because only one has clinical leadership in 2012 when the policy changes occurred.
+
+    Thus, in one specification I compare hospitals that ever have a clinically trained executive to hospitals that never have one.
+This specification captures the signaling effect.
+In another specification, I exclude hospitals that never had a clinically trained executive and compare hospitals with a clinical executive in 2012 to those with a clinical executive in any other year, capturing the managing effect.
+As in the previous analyses, I apply the synthetic difference-in-differences approach to each specification.
+That is, I assign greater weight to control units with pre-trends similar to treated units and to time periods that balance pre- and post-policy trends for the control group.
+I discuss the assumptions of parallel trends, no anticipation, and no confounding events in Sections \ref{sec:clinical} and \ref{sec:forprofit}, and these assumptions also apply here.
+I also assume executive team changes are exogenous to policy changes, an assumption I examine in Section \ref{sec:changes}.
+
+    Table \ref{tab:MD_noMD_readmort_decomp_synth} reports the estimates for each decomposition specification.
+Columns (1) and (2) report readmission rate estimates, and columns (3) and (4) report mortality rate estimates.
+The first row compares hospitals that ever had a clinical executive to those that never had one (signaling), and the second row compares hospitals with a clinical executive in 2012 to those with a clinical executive in any other year (managing).
+Although none of the estimates are statistically significant, the managing effect coefficients closely match the full effect reported in Section \ref{sec:clinical}.
+Signaling effect coefficients are statistically insignificant and near zero.
+These results rule out the possibility that clinical executives merely signal underlying hospital objectives and provide evidence that it is likely a difference in management.
 
@@
 
-    Consider a simplified model of hospital objectives, in which hospitals choose a level of quality of care that maximizes an objective function capturing the trade-off between profit and patient welfare.
-That is,
+    \subsubsection{Theoretical Model}
+
+    I now examine a potential mechanism based on financial and altruistic motivations.
+Consider a simplified model of hospital objectives, in which hospitals choose an optimal level of quality of care that maximizes an objective function capturing the trade-off between profit and patient welfare.
+Formally,
 
@@
 
-    \noindent where $\pi(\theta)$ is a profit function depending on quality $\theta$, $u(.)$ is extra utility gained from patient welfare, which is increasing and concave in quality $\theta$, and $\alpha\in[0,1]$ captures the weight placed on profit vs.
-patient welfare.
-There is an implicit stay-open condition throughout the model.
-
-    I consider hospital choice of quality under pay-for-performance.
-In a world where performance is not incentivized through financial gain (as in traditional fee-for-service settings), hospital profit $\pi$ is maximized when the quantity of patients is maximized.
+    \noindent where $\pi(\theta)$ is a profit function depending on quality $\theta$, $u(.)$ represents additional utility from patient welfare, which increases with quality $\theta$ and is concave, and $\alpha\in[0,1]$ captures the relative weight on profit versus patient welfare.
+The model assumes an implicit stay-open condition.
+
+    Under traditional fee-for-service settings, where performance is not financially incentivized, hospitals maximize profit $\pi$ by maximizing patient volume.
 However, pay-for-performance incentives incorporate quality into the profit function (\cite{dranove2011health}).
-Therefore, hospitals that only value financial profit will only consider quality if it is an input to the profit function, and therefore will respond more drastically to pay-for-performance than a hospital who values quality as an input to patient welfare before the incentives.
-
-    I will now formalize this intuition mathematically.
-For simplicity, I abstract away from quality having a direct impact on quantity of patients or price.\footnote{As long as the relationship between quality and quantity/price are not correlated with $\alpha$, this does not change the conclusion of the model.} Taking the first order condition yields:
+Hospitals that prioritize financial profit consider quality only when it enters the profit function, and thus respond more strongly to pay-for-performance than hospitals that value quality for patient welfare before incentives.
+
+    I now formalize this intuition mathematically.
+For simplicity, I abstract away from quality having a direct impact on quantity of patients or price.\footnote{As long as the relationship between quality and quantity/price is not correlated with $\alpha$, this does not change the conclusion of the model.} Taking the first order condition yields:
 
@@
 That is, quality benefits patients, but hospitals do not directly derive profit from having a high quality.
-Then, assuming there is a positive cost to increasing quality, $\pi(\theta)$ is decreasing in $\theta$.
+Assuming a positive cost to increasing quality, $\pi(\theta)$ is decreasing in $\theta$.
 Then $\frac{du'(\theta)}{d\alpha} > 0$.
@@
 
-    Next, assume the hospital is in a pay-for-performance setting, where quality benefits patients and directly increases profit.
-Then, assuming the marginal financial benefit of increasing quality is greater than the marginal financial cost, $\pi(\theta)$ is increasing in $\theta$.
-Then $\frac{du'(\theta)}{d\alpha} < 0$, revealing that in a pay-for-performance setting, the more weight placed on profit, the higher quality of care chosen by the hospital.
-I combine the results found in each scenario into one response function that depends on $\alpha$, where $\theta_{p4p}$ is quality under pay-for-performance incentives and $\theta_{ffs}$ is quality under fee-for-service only.
-That is,
+    Alternatively, assume the hospital is in a pay-for-performance setting, where quality benefits patients and directly increases profit.
+If the marginal financial benefit of increasing quality is greater than the marginal financial cost, $\pi(\theta)$ is increasing in $\theta$.
+Then $\frac{du'(\theta)}{d\alpha} < 0$, revealing that in a pay-for-performance setting, the more weight placed on profit, the higher quality of care chosen.
+I combine both scenarios into a single response function that depends on $\alpha$, where $\theta_{p4p}$ denotes quality under pay-for-performance and $\theta_{ffs}$ denotes quality under fee-for-service:
 
@@
 
-    The first prediction of the model is that, all else held equal, hospitals that value profit will have a lower quality than those who value patient welfare in a fee-for-service setting.
-Second, a hospital's change in quality once pay-for-performance is enacted depends on how much weight the hospital places on profit vs.
-patient welfare.
-Particularly, hospitals with higher weight on profit respond more than hospitals with higher weight on patient welfare.
-Therefore, factors that affect weight $\alpha$ will be correlated with choice of quality, and with how quality changes under new pay-for-performance incentives.
-In this paper, I test the hypothesis that clinical experience on a hospital's leadership team is a factor that is correlated with weight $\alpha$, and thus affects response to pay-for-performance incentives.
-To investigate this hypothesis empirically, I merge all general for-profit hospitals in the AHA survey to my sample of nonprofit hospitals.
+    Therefore, the response to pay-for-performance incentives depends on $\alpha$.
+Hospitals that prioritize profit respond more strongly to financial incentives, while hospitals that prioritize patient welfare respond less to financial incentives on quality.
+
+    \subsubsection{Empirical Estimation}
+
+    The first prediction of the model is that, in a fee-for-service setting with all else held equal, hospitals that value profit have a lower quality than those who value patient welfare.
+Second, the change in quality after pay-for-performance depends on how much weight the hospital places on profit versus patient welfare.
+Particularly, hospitals that place greater weight on profit respond more strongly than those that emphasize patient welfare.
+To investigate this hypothesis empirically, I merge general for-profit hospitals from the AHA survey to my sample of nonprofit hospitals.
 I then compare readmission and mortality rates of always clinical executives and never clinical executives to for-profit hospitals after the incentive change.
 
-    While I do not have information on leadership teams of for-profit hospitals, on average, they have revealed through their ownership status that they more heavily weigh profit, and therefore have an $\alpha$ close to 1.
-This is not inconsistent with previous work positing that nonprofits and for-profits act similarly on average (\cite{duggan2002hospital}).
-In fact, my data is consistent with this idea when nonprofits are not broken down by leadership teams.
-However, comparing different executive team behavior to for-profit behavior helps reveal how profit driven certain leadership teams are, which could actually help us understand why there are similarities between nonprofits and for-profits on average.
+    Although I lack data on leadership teams of for-profit hospitals, their ownership status indicates they prioritize profit and likely have an $\alpha$ close to 1.
+This is not inconsistent with previous work suggesting that nonprofits and for-profits act similarly on average (\cite{duggan2002hospital}).
+In fact, my data is consistent with this idea when nonprofits are not disaggregated by leadership teams.
+However, comparing different executive team behavior to for-profit behavior helps reveal how profit driven certain leadership teams are, which could reveal why there are similarities between nonprofits and for-profits on average.
 I estimate the weighted synthetic difference-in-differences version of
@@
 
-    \noindent where $y_{ht}$ is one of the readmission or mortality rate outcomes discussed in Section \ref{sec:hrrp}, for-profit$_{ht}$ is an indicator for having for-profit ownership status, and $\gamma$ and $\delta$ are hospital and time fixed effects.
-In my estimation, nonprofits in the comparison group are more heavily weighted if their pre-trends are more similar to for-profits, and time periods are weighted to balance pre and post trends for for-profit hospitals.
-These weights follow the synthetic difference-in-differences approach, which is more robust to the parallel trends assumption.
-
-    I create two subsets of data to compare the response of for-profits to the response of the subsets of nonprofits based on executive team.
-First, I limit only to for-profits and nonprofits that always have clinically trained executives.
-Second, I limit to for-profits and nonprofits that never have clinically trained executives.\footnote{I assume that the sample selection of stable leadership teams is not correlated with the policy changes.
-Please see Section \ref{sec:sig_man} for the analysis and discussion of this assumption.} I show summary statistics of relevant variables for for-profit hospitals in Appendix \ref{app:sumstats}.
-Again, I employ a synthetic difference-in-differences strategy to rely less heavily on a strong parallel trends assumption for different types of hospitals.
-
-    The identification assumptions I make here are largely similar to those in Section \ref{sec:clinical}, only now incorporate for-profit hospitals.
-First, I assume that, absent the change in incentives, the weighted composition of for-profits and the relevant comparison group would have had parallel trends in outcomes.
-I also assume that there is no anticipation of the policy changes prior to 2012, which is reasonable as the rules of the policies were announced in October of 2011.
-Finally, I assume no other changes correlated with both hospital type (for-profit, nonprofit with and without clinically trained executives) and quality occurred in conjunction with the 2012 pay-for-performance policies.
-
-    The estimates and graphical representation of the effect of being for-profit is shown in Figure \ref{fig:read_synth_plot}.
-In panel \ref{fig:read_synth_plotb}, the average treatment effect shows the effect of being for-profit compared to nonprofit with clinical executives on readmission rates after pay-for-performance changes.
-I find that for-profits decrease readmission rates by .3 (standard error .12) more than nonprofits with clinically trained executives.
+    \noindent where $y_{ht}$ is readmission or mortality rate, discussed in Section \ref{sec:hrrp}, for-profit$_{ht}$ is an indicator for having for-profit ownership status, and $\gamma$ and $\delta$ are hospital and time fixed effects.
+In my estimation, nonprofits in the comparison group are more heavily weighted if their pre-trends are more similar to for-profits, and time periods are weighted to balance pre- and post-trends of for-profit hospitals.
+These weights follow the synthetic difference-in-differences approach, which is more robust to violations in the parallel trends assumption.
+
+    I create two data subsets to compare for-profit responses with nonprofit responses by executive team type.
+First, I include for-profits and nonprofits that always have clinically trained executives.
+Second, I include for-profits and nonprofits that never have clinically trained executives.\footnote{I assume sample selection of stable leadership teams does not correlate with policy changes.
+See Section \ref{sec:sig_man} for analysis and discussion of this assumption.} Appendix \ref{app:sumstats} reports summary statistics for relevant variables for for-profit hospitals.
+I again use a synthetic difference-in-differences strategy to reduce reliance on a strong parallel trends assumption across hospital types.
+
+    The identification assumptions here mirror those in Section \ref{sec:clinical}, but now include for-profit hospitals.
+First, I assume that without the incentive change, the weighted composition of for-profits and the comparison group would follow parallel outcome trends.
+Second, I assume no anticipation of policy changes before 2012, which is reasonable because the rules were announced in October 2011.
+Finally, I assume no other changes correlated with hospital type (for-profit, nonprofit with or without clinically trained executives) and quality occurred alongside the 2012 pay-for-performance policies.
+
+    The estimates and graphical representation of the effect of being for-profit are shown in Figure \ref{fig:read_synth_plot}.
+In Figure \ref{fig:read_synth_plotb}, I present the effect of being for-profit compared to nonprofit with clinical executives on readmission rates after pay-for-performance changes.
+I find that for-profits decrease readmission rates by 0.3 (standard error 0.12) more than nonprofits with clinically trained executives.
 That is, for-profits respond more aggressively to the incentive change than nonprofits with clinical leadership.
 This difference is nearly identical to the difference between clinical and non-clinical teams shown in Section \ref{sec:clinical}.
-So, unsurprisingly, panel \ref{fig:read_synth_plotc} shows no such difference when comparing for-profits to nonprofits without clinically trained executives.
+Figure \ref{fig:read_synth_plotc} shows no such difference when comparing for-profits to nonprofits without clinically trained executives.
 Not only is the estimate not statistically significant, but the magnitude indicates a precise zero.
@@
 
-    Next, I present estimated differences in mortality rates between different hospital types in Figure \ref{fig:mort_synth_plot}.
-In panel \ref{fig:mort_synth_plotb}, the estimated difference is between for-profits and nonprofits with clinically trained executives.
-The magnitude of the difference in mortality rates is .08, where clinical team hospitals continue on an increasing trend and for-profits remain stable in mortality rate.
-While I cannot rule out a lack of differential response, the documented noisiness of mortality measures (\cite{mackenzie2016measuring}) may mean that the true effect is different from zero.
-However, the estimated difference in mortality between for-profits and nonprofits without clinically trained executives is certainly zero with a magnitude of .03, shown in panel \ref{fig:mort_synth_plotc}.
+    Next, I present estimated differences in mortality rates across hospital types in Figure \ref{fig:mort_synth_plot}.
+Figure \ref{fig:mort_synth_plotb} compares for-profits to nonprofits with clinically trained executives.
+The difference in mortality rates is 0.08: clinical team hospitals continue an upward trend, while for-profits remain stable.
+Although I cannot rule out no differential response, the documented noisiness of mortality measures (\cite{mackenzie2016measuring}) suggests the true effect may differ from zero.
+However, the estimated difference in mortality between for-profits and nonprofits without clinically trained executives is certainly zero with a magnitude of 0.03, shown in Figure \ref{fig:mort_synth_plotc}.
 
@@
 
-    It may be that the observed difference in quality after pay-for-performance could be correlated with the expertise of leadership teams.
-While most doctors take a hippocratic oath to provide care with integrity and compassion, doctors with training in certain specialties may be more equipped to improve care for certain patients.
-Since the pay-for-performance incentive programs directly target three conditions, physicians with knowledge of these conditions may be especially suited to respond to the policies.
-Specifically, a pulmonologist typically treats pneumonia, and a cardiologist typically treats heart failure and heart attack patients, which are the three conditions considered in the HRRP penalties.
-Both of these specialties are under the umbrella of internal medicine, which is the most granular specialty type in my data.
-
-    Consider an extension to the model discussed in Section \ref{sec:forprofit}, where I now introduce a measure of expertise $\lambda$.
+    \subsubsection{Theoretical Model}
+
+    The observed difference in quality after pay-for-performance may be correlated with the expertise of leadership teams.
+While most doctors take a Hippocratic Oath to deliver care with integrity and compassion, doctors with training in certain specialties could be more equipped to improve care for specific patients.
+Because the pay-for-performance incentive programs directly target three conditions, physicians with knowledge of these conditions may be especially suited to respond to these policies.
+Specifically, a pulmonologist typically treats pneumonia, and a cardiologist typically treats heart failure and patients with heart attacks, the three conditions included in HRRP penalties.
+Both of these specialties are under the umbrella of internal medicine, which is the most granular specialty type in the data.
+
+    I extend the model discussed in Section \ref{sec:forprofit} by introducing a measure of expertise, $\lambda$.
 To capture the role of hospital leadership expertise in shaping quality decisions, we introduce a binary variable $\lambda \in \{0,1\}$, where $\lambda=1$ indicates high expertise and $\lambda=0$ indicates low expertise.
 Expertise affects the marginal cost of quality improvements: hospitals with high expertise can implement quality enhancements more efficiently, reducing the effective cost of quality.
-
 The hospital’s objective function becomes:
@@
     \noindent where $R(\theta)$ is revenue, which depends on quality,
-    $c(\theta)(1 + \lambda)$ is the cost of quality, increasing and convex in $\theta$, scaled by expertise, $u(\theta)$ is patient welfare, increasing and concave in $\theta$, $\alpha \in [0,1]$ is the weight on profit relative to patient welfare.
-The hospital's first-order condition under fee-for-service and pay-for-performance is then:
+    $c(\theta)(1 + \lambda)$ is the cost of quality, increasing and convex in $\theta$, and scaled by expertise, $u(\theta)$ is patient welfare, increasing and concave in $\theta$, $\alpha \in [0,1]$ is the weight on profit relative to patient welfare.
+The hospital's first-order conditions under fee-for-service and pay-for-performance are then:
 
@@
     \noindent I make the following assumptions on patient welfare, cost, and revenue: $u'(\theta) > 0,\; u''(\theta) < 0,\; c'(\theta) > 0,\; c''(\theta) > 0,\; R'(\theta) > 0,\; R''(\theta) \approx 0$.
-By the Implicit Function Theorem.
+By the Implicit Function Theorem,
     \begin{align}
@@
     I now want to compare the change in $\theta$ when going from fee-for-service to pay-for-performance depending on different levels of $\lambda$.
+The partial derivative of quality response with respect to expertise is defined as
     \begin{align}
@@
 
-    Since $\theta_{p4p}>\theta_{ffs}$, $u(\theta_{p4p}>u(\theta_{ffs}$ (that is, less negative).
-Therefore, the pay-for-performance denominator is less negative than the fee-for-service denominator.
+    Since $\theta_{p4p}>\theta_{ffs}$, $u(\theta_{p4p})>u(\theta_{ffs})$, which means $u(\theta_{p4p})$ is less negative than $u(\theta_{ffs})$.
+Therefore, the denominator under pay-for-performance is less negative than under fee-for-service.
 Hence, $\frac{\partial\theta_{p4p}}{\partial \lambda} > \frac{\partial\theta_{ffs}}{\partial \lambda}$.
-This implies that for higher expertise, the response to pay-for-performance policies will be greater.
-
-    To test this empirically, I limit the sample of hospitals to those who always have some type of clinical executive over the sample period.
-Then, I estimate the difference in readmission and mortality rates after pay-for-performance among hospitals who do and do not have an internal medicine clinical executive.
-I lose a significant fraction of the sample with this limitation: there are 60 hospitals that consistently have an internal medicine executive or only a different specialty executive.
+This implies that as expertise increases, the response to pay-for-performance policies becomes stronger.
+
+    \subsubsection{Estimation}
+
+    To test this empirically, I limit the sample to hospitals that consistently have a clinical executive throughout the sample period.
+Then, I estimate the difference in readmission and mortality rates after pay-for-performance among hospitals that do and do not have an internal medicine clinical executive.
+This restriction reduces the sample size substantially: 60 hospitals consistently have either an internal medicine executive or an executive from a different specialty.
 The summary statistics for these samples are in Appendix \ref{app:sumstats}.
@@
 
-    While I am limited in power due to the small sample size of hospitals in this analysis, I provide evidence that hospitals with internal medicine executives may change readmission rates differentially compared to teams with clinical executives of other specialty types.
-The estimated effect of having an internal medicine executive is -.38 (standard error .24), where internal medicine executives decrease readmissions at a faster rate than other specialty types.
-The results for mortality rates are similar (still not statistically significant), but smaller in magnitude and smaller relative to the main results.
-This is probably due to the specialty-specific conditions being more relevant for the readmission rate penalties than the quality rewards.
+    Although statistical power is limited due to the small sample size, I provide evidence that hospitals with internal medicine executives may change readmission rates differently compared to teams led by executives from other specialties.
+The estimated effect of having an internal medicine executive is -0.38 (standard error 0.24), indicating that hospitals with internal medicine executives reduce readmissions at a faster rate than those with executives from other specialties.
+Results for mortality rates are similar, though still not statistically significant, but smaller in magnitude and relative to the main findings.
+This likely reflects the greater relevance of specialty-specific conditions to readmission penalties compared to quality rewards.
+Although the small sample size limits statistical power, the results point to a potential role for leadership expertise in shaping hospital responses to policy incentives.
+Future research with larger samples could clarify whether these patterns hold more broadly.
 
@@
 
-    \citeauthor{gupta2021impacts} (\citeyear{gupta2021impacts}) finds that a large portion of hospital readmission reductions are due to hospitals being less willing to readmit patients from their own hospital, while they will still readmit patients coming from other hospitals.
-Given the differential response to the HRRP by clinical and non-clinical hospital leaders, the natural question arises as to whether there is a differential propensity to game the system, as considered in previous work (\cite{gupta2021impacts}).
-While I do not have patient level data, I can test empirically whether the overall patient population of hospitals changes differentially for clinician or non-clinician leaders.
-
-    I use three outcomes to test this hypothesis.
-First, the CMS Case Mix Index files provide an aggregate measure of patient complexity for each hospital in each year.
-A higher index implies the patient population in the hospital is more complex in their treatment requirement.
-Additionally, I consider the population of patients treated in the hospital with public insurance, either Medicare or Medicaid.
-These outcomes will indicate if certain types of leadership teams attempt to cost shift relative to others after the penalties or rewards took effect.
-
-    I estimate the same synthetic difference-in-differences equation discussed in Section \ref{sec:clinical}.
-The results of these estimations are presented in Figure \ref{fig:selection}.
-All of the treatment effects in these estimations are insignificant and null in magnitude, suggesting that there is no evidence of the overall patient population of hospitals changing differentially based on leadership team.
-This does not mean that there is no change in specific readmission behavior as found in \citeauthor{gupta2021impacts} (\citeyear{gupta2021impacts}), but a limitation of this data only allows me to look at the aggregate patient population as opposed to specifically readmissions.
-
-    \begin{figure}[ht!]
-     \caption{Patient Population as Outcomes}
-     \centering
-     \begin{subfigure}[b]{0.45\textwidth}
-         \centering
-         \caption{Case Mix Index}
-         \includegraphics[width=\textwidth]{Objects/cmi_md_nomd_synth_graph.pdf}
-         \label{fig:cmi}
-     \end{subfigure}
-     \hfill
-     \begin{subfigure}[b]{0.45\textwidth}
-         \centering
-         \caption{Medicare Percent Discharges}
-         \includegraphics[width=\textwidth]{Objects/mcare_md_nomd_synth_graph.pdf}
-         \label{fig:mcare}
-     \end{subfigure}%
-        \begin{subfigure}[b]{0.45\textwidth}
-         \centering
-         \caption{Medicaid Percent Discharges}
-         \includegraphics[width=\textwidth]{Objects/mcaid_md_nomd_synth_graph.pdf}
-         \label{fig:mcaid}
-     \end{subfigure}
-        \label{fig:selection}
-    \end{figure}
+    \citeauthor{gupta2021impacts} (\citeyear{gupta2021impacts}) finds that a large portion of hospital readmission reductions result from hospitals being less willing to readmit their own patients, while continuing to readmit patients from other hospitals.
+Given the differential response to HRRP by clinical and non-clinical leaders, a natural question is whether these groups differ in their propensity to game the system, as explored in prior work (\cite{gupta2021impacts}).
+Although patient-level data are unobserved, I can test whether the overall hospital patient population changes differentially under clinician versus non-clinician leadership.
+
+    I examine three outcomes to test this hypothesis.First, I use the CMS Case Mix Index, which provides an aggregate measure of patient complexity for each hospital and year.
+A higher index indicates that the hospital’s patient population requires more complex treatment.
+Additionally, I examine the share of patients treated in the hospital who have public insurance, either Medicare or Medicaid.
+These outcomes help assess whether certain leadership types attempt to shift costs relative to others after penalties or rewards take effect.
+
+    I estimate the same synthetic difference-in-differences specification described in Section \ref{sec:clinical}.
+The results are presented in Table \ref{tab:selection}.
+All estimated treatment effects are statistically insignificant and near zero, indicating no evidence that overall hospital patient populations change differentially by leadership type.
+This does not rule out changes in specific readmission behavior, as documented by \citeauthor{gupta2021impacts} (\citeyear{gupta2021impacts}).
+However, a limitation of this dataset is that it only permits analysis of aggregate patient populations rather than individual readmissions.
+
+    \import{Tables}{selection_table.tex}
 
@@
 
-    Lastly, I examine whether clinician versus non-clinician leaders respond differentially in terms of their investment strategy after the policies took effect.
-Ideally, I would have more granular information within the hospital to observe differences in management.
-That is a limitation of the publicly available hospital-level data that I utilize here.
-I use the same synthetic difference-in-differences estimation strategy to compare clinical leaders with non-clinical leaders before and after 2012.
-The variables I use as outcomes are the number of full time nurses hired (from AHA), movable equipment purchases, fixed equipment purchases, labor costs, building purchases, and total operating expenses (all from HCRIS).
-For several of these variables, over 90\% of observations are missing in 2010.
-Therefore, I remove this year for the relevant variables.
-The results of these estimations are shown in Table \ref{tab:investment}, and do not present any evidence that hospitals with different leadership teams are changing their investment strategies differentially as a result of the policies.
+    Lastly, I examine whether clinician and non-clinician leaders differ in their investment strategies after the policies took effect.
+Ideally, more granular hospital-level data would allow observation of management differences, which is a limitation of the publicly available hospital-level data used in this analysis.
+I apply the same synthetic difference-in-differences estimation strategy to compare clinical leaders with non-clinical leaders before and after 2012.
+The variables I use as outcomes are the number of full-time nurses hired (AHA), movable equipment purchases, fixed equipment purchases, labor costs, building purchases, and total operating expenses (HCRIS).
+Because over 90\% of observations for several HCRIS variables are missing in 2010, I exclude this year from the analysis.
+Table \ref{tab:investment} presents the results, which show no evidence that hospitals with different leadership teams changed their investment strategies in response to the policies.
 
@@
 
-    While previous literature has documented correlations between executive characteristics and firm behaviors/performance, research is still limited on this topic in the health care sector.
+    While previous literature has documented correlations between executive characteristics and firm behaviors/performance, research on this topic remains limited in the health care sector.
 Health care managers are tasked with maintaining a business while also prioritizing patient care, which is why existing findings on manager effects may not extend to this setting.

### 2025-11-14 Paper/paperdraft_fall2025.tex
Updates from Overleaf

@@
 
-	\linespread{1.2}\title{\vspace{-0.5in} Does Hospital Leadership Matter?\\ \large Evidence from Pay-for-Performance}
+	\linespread{1.2}\title{\vspace{-0.5in} DOES HOSPITAL LEADERSHIP MATTER?\\ \large EVIDENCE FROM PAY-FOR-PERFORMANCE\thanks{I thank Aliyah Bagro and Vidyah Vedula for their excellent research assistance.
+I express my deepest thanks to Ian McCarthy for his advice and ongoing helpful feedback.
+Further, I thank Sara Markowitz, Matt Grennan, and Evan Saltzman for their helpful comments.
+I am grateful to the audiences at EuHEA and ASHEcon conferences, as well as at the University of Queensland, the University of Melbourne, the University of Miami, and the University of Iowa for their insightful feedback.}}
 
@@
 
-	\author{\vspace{10mm}Hanna Glenn\footnote{School of Economics, University of Queensland, hanna.glenn@uq.edu.au.} }
+	\author{\vspace{10mm}Hanna Glenn\footnote{School of Economics, University of Queensland, hanna.glenn@uq.edu.au.}
+    }
 
@@
          Firm managers are often seen as key drivers of firm decisions, strategies, and performance.
-While extensive research links managerial characteristics to firm outcomes, less is known about how hospital executives influence responses to health care incentives.
+Extensive research links managerial traits to firm outcomes, yet little is known about how hospital executives shape responses to healthcare incentives.
 Health care policies in the US often focus on improving quality of care without increasing costs, yet it is unclear whether hospital leaders might contribute to this goal.
-Leveraging pay-for-performance incentives, I estimate the differential response to financial incentives on quality by different types of nonprofit hospital executives: those with and without a clinical training background as a medical doctor.
-I find that hospitals without clinically trained executives respond more to the financial incentives than those with clinically trained executives.
-I find that this difference can be explained by patient-centered focus and by expertise lowering the cost of quality improvement.
-These findings highlight the importance of considering leadership as an important factor in hospital decision making when adjusting incentives.
+Leveraging pay-for-performance incentives, I estimate the differential response to financial incentives on quality by different types of nonprofit hospital executives: those with and without training as a medical doctor.
+I find that hospitals led by non-clinical executives respond more strongly to financial incentives than those led by clinically trained executives.
+The difference is not a signal of underlying hospital objectives, but a difference in active management.
+Further, the observed difference reflects two mechanisms: clinically trained leaders prioritize patient-centered care and possess expertise that reduces the cost of quality improvements.
+These findings highlight the role of leadership in firm decision-making and the need to account for executive characteristics when designing incentives.
 		} 
@@
 
-    In corporate environments, executives are considered crucial drivers of a company's practices and performance, as they are the highest level of management that affects day-to-day operations of the organization.
-Executive heterogeneity has been linked to various actions and outcomes of firms; characteristics such as gender, age, and background are correlated with financial performance and behaviors such as employee diversity, wages, aggressiveness in mergers and acquisitions, and other corporate strategies.\footnote{\cite{bertrand2003managing}; \cite{matsa2011chipping}; \cite{matsa2011chipping}; \cite{adams2009women}; \cite{ahern2012changing}; \cite{flabbi2019female}; \cite{benmelech2015military}; \cite{custodio2013generalists}; \cite{frydman2019rising}} However, our understanding is typically limited to executives working for large, publicly traded corporations.
-Additionally, while correlations are an important first step to understanding the effects of leadership, little work has been done to uncover whether such correlations occur due to executive characteristics interacting with the objectives of firms.
-
-    In this paper, I study the effect of specialized executive training on firm behavior when the firm experiences an external shock to their objective function, specifically in the context of US hospitals.
-I leverage this setting for two reasons.
-First, in the US, health care makes up almost 20\% of spending (\cite{CMSNationalHealthExpenditure}), and a major goal of policymakers is to improve the quality of health care without increasing costs, as the US preforms poorly in these metrics compared to other similar countries.
-Yet, little is known about how leaders affect health care value in the hospitals they run.
-Second, nonprofit hospitals are particularly important in this context, and it is unclear whether existing findings on executive characteristics can be extrapolated to firms with presumably different incentives and objectives than those which have been previously studied.
-
-    An executive characteristic relevant to this context is clinical training.
-Physicians can bring a unique combination of clinical and administration expertise to leadership, and they have insights that can potentially benefit their employees and improve quality (\cite{Stajduhar_2023}, \cite{Ahmed_2022}).
-However, hiring physician executives is controversial, as physicians are not always trained in business management, which can lead to poor practices and lower quality (\cite{HarvardBusinessReview2018}).
-Thus, I estimate how the presence of clinically trained executives on a hospital leadership team affects how the hospital makes decisions.
+    In corporate environments, executives are key drivers of company practices and performance, shaping day-to-day operations at the highest level of management.
+Executive heterogeneity (based on gender, age, and professional background) has been linked to firm outcomes, including financial performance and behaviors such as workforce diversity, wage setting, and merger strategies.\footnote{\cite{bertrand2003managing}; \cite{matsa2011chipping}; \cite{matsa2011chipping}; \cite{adams2009women}; \cite{ahern2012changing}; \cite{flabbi2019female}; \cite{benmelech2015military}; \cite{custodio2013generalists}; \cite{frydman2019rising}} However, most research focuses on executives in large, publicly traded corporations.
+Moreover, while correlations provide an initial understanding of leadership effects, few studies examine whether these patterns arise from interactions between executive characteristics and firm objectives.
+
+    In this paper, I study the effect of specialized executive training on firm behavior when an external shock alters the objective function, specifically in the context of US hospitals.
+I focus on this setting for two reasons.
+First, in the US, health care makes up almost 20\% of spending (\cite{CMSNationalHealthExpenditure}), and a major goal of policymakers is to improve the quality of health care without increasing costs, as the US performs poorly in these metrics compared to other high-income countries.
+Yet, little evidence exists on how hospital leaders affect health care value.
+Second, nonprofit hospitals are particularly important, and it is unclear whether existing findings on executive characteristics from other sectors can be extrapolated to firms with presumably different incentives and objectives.
+
+    A key executive characteristic in context is clinical training.
+Physicians combine clinical and administrative expertise, offering insights that may benefit staff and improve care quality (\cite{Stajduhar_2023}, \cite{Ahmed_2022}).
+However, hiring physician executives is controversial because many lack formal business training, which can result in inefficient management and reduced quality (\cite{HarvardBusinessReview2018}).
+Thus, I estimate how the presence of clinically trained executives on a hospital leadership team influences hospital decision making.
 I scrape executive names and titles of nonprofit hospitals in the US from publicly available Tax Form 990s.
@@
 While limited to nonprofits due to data availability, this sample of hospitals is an important driver of health care in the US.
-Private nonprofit hospitals make up 50\% of all hospitals in the US, and staff, on average, 207 beds.
-For comparison, for-profits make up 36\% of hospitals and staff 107 beds on average (\cite{ASPE_2023}).
-
-    The non-random selection of leaders makes it difficult to interpret any direct comparison of differing executive teams as causal.
-Therefore, I leverage pay-for-performance policies enacted in the US as an exogenous shock to hospital incentives on quality, and compare the change in readmission and mortality rates of hospitals with and without clinically trained executives.
-Under the assumption that the two types of teams would have had similar trends in readmission and mortality rates absent the incentive change, I find that hospitals without clinically trained trained executives respond more to financial incentives on quality than hospitals with clinical executives.
-That is, while both types of hospital leadership teams decrease readmission rates after the policy change, those without clinical executives decrease readmissions at a faster rate than those with clinical executives.
-I find no evidence of selective patient practices driving this result.
-
-    I use a theoretical model to investigate whether these results might be explained by differing preferences on the trade-off between profit and patient welfare, where I model these hospital objectives and allow the weight placed on each to vary with executive team type.
-Hospitals that place more weight on profit will naturally respond more drastically to a policy that adds quality directly into the profit function.
-Thus, a potential reason for the difference in response is that non-clinical executive teams are more profit driven.
-I investigate this empirically by comparing the response of hospitals with and without clinically trained executives to the response of for-profit hospitals, who have revealed that they are more profit driven by their ownership type.
-I find that nonprofit hospitals without clinically trained executives respond to the financial incentives on quality similarly to for-profits, while nonprofits without clinically trained executives respond differently than for-profits.
-For-profits and non-clinical executive teams change quality more in response to the incentive change than clinically trained executive teams.
-
-    Finally, the question remains as to whether clinically trained executives are simply a signal of unchanging underlying hospital preferences (of, for example, the board of directors), or if they actively manage the hospital differently, effectively changing hospital objectives.
-I analyze this by leveraging changes in hospital propensity to hire clinical executives to decompose the full effect of clinical leadership into a signaling effect (clinical executives signal underlying hospital preferences), or managing effect (clinical executives manage the hospital differently).
-First, I show that hospital executive changes are not correlated with the change to financial incentives.
-To measure the signaling effect, I compare hospitals who ever have a clinically trained executive to hospitals who never have a clinically trained executive.
-For the managing effect, I limit to hospitals who ever have a clinically trained executive and compare those who have a clinically trained executive when the policies were announced to those who have clinically trained executives at other times.
-I find that the difference in response is driven by clinically trained executives managing the hospital differently.
-
-    Many researchers have documented a correlation between executive characteristics and firm actions, starting with \citeauthor{bertrand2003managing} (\citeyear{bertrand2003managing}), who show that manager fixed effects are an important driver of many firm behaviors and decisions.
-Having more female executives is correlated with female employee wages and corporate strategies (\cite{flabbi2019female}; \cite{matsa2013female}); young male CEOs tend to be more aggressive in mergers and acquisitions, while those with military experience are less aggressive (\cite{levi2010deal}; \cite{benmelech2015military}); CEOs with general ability tend to receive higher pay and perform better (\cite{kaplan2012ceo}; \cite{custodio2013generalists}; \cite{adams2018director}; \cite{frydman2019rising}); older CEOs are associated with measures of corporate culture (\cite{graham2022corporate}).
-Finally, Chief Diversity Officers are not found to have any effect on hiring more diversely in universities (\cite{bradley2022impact}).
-
-    Due to data limitations, many of these studies focus on large, publicly traded firms in the US.
-\citeauthor{brickley2010board} (\citeyear{brickley2010board}) have the only study to my knowledge in the US nonprofit context considering hospital board of directors.
-Using exogenous variation in expected Medicare profits, the authors find that having an internal board of directors increases CEO compensation, and having physician board members decreases public donations (\cite{brickley2010board}).
-Additionally, two studies focus on hospital performance in other countries.
+Private nonprofit hospitals account for 50\% of all US hospitals and average 207 beds, compared to for-profit hospitals, which comprise 36\% and average 107 beds (\cite{ASPE_2023}).
+
+    The non-random selection of leaders complicates causal interpretation of direct comparisons between executive teams.
+Therefore, I leverage pay-for-performance policies enacted in the US as an exogenous shock to hospital incentives on quality and compare changes in readmission and mortality rates between hospitals with and without clinically trained executives using a synthetic difference in differences estimation strategy.
+Assuming the two types of teams would have followed similar trends absent the policy change, I find that hospitals without clinically trained executives respond more strongly to financial incentives than those with clinical executives.
+In other words, both types of leadership teams reduce readmission rates after the policy change, but hospitals without clinical executives do so at a faster rate.
+
+    I investigate several mechanisms to explain the observed difference.
+First, I leverage the timing of clinically trained executives to show that clinicians actively manage hospitals differently, rather than serving only as a signal of underlying preferences.
+Next, I develop a theoretical model to illustrate how these results may stem from differing preferences over the trade-off between profit and patient welfare, allowing the weight on each objective to vary by executive team type.
+Hospitals that prioritize profit respond more strongly to policies that incorporate quality into the profit function.
+Thus, one potential explanation is that non-clinical executive teams are more profit-driven.
+I test this empirically by comparing responses of hospitals with and without clinically trained executives to those of for-profit hospitals, which are known to be more profit-driven by ownership structure.
+I find that nonprofit hospitals without clinically trained executives respond to quality incentives similarly to for-profits, whereas those with clinically trained executives respond differently.
+Both for-profits and non-clinical executive teams increase quality more in response to the incentive change than clinically trained teams.
+Similarly, I show theoretically that leadership expertise may drive the observed difference and investigate this empirically by comparing physician specialties.
+This analysis suggests that expertise is a likely underlying mechanism.
+Finally, I find that changes in aggregate patient populations and investment decisions do not explain the observed difference.
+
+    Many researchers have documented a correlation between executive characteristics and firm actions, beginning with \citeauthor{bertrand2003managing} (\citeyear{bertrand2003managing}), who shows that manager fixed effects are an important factor in many firm behaviors and decisions.
+Greater female representation among executives correlates with higher female employee wages and shifts in corporate strategy (\cite{flabbi2019female}; \cite{matsa2013female}).
+Young male CEOs tend to be more aggressive in mergers and acquisitions, whereas those with military experience are less aggressive (\cite{levi2010deal}; \cite{benmelech2015military}).
+CEOs with general managerial ability receive higher pay and achieve better performance (\cite{kaplan2012ceo}; \cite{custodio2013generalists}; \cite{adams2018director}; \cite{frydman2019rising}).
+Older CEOs are associated with differences in corporate culture (\cite{graham2022corporate}).
+Finally, Chief Diversity Officers have not been found to influence diversity in university hiring (\cite{bradley2022impact}).
+
+    Because of data limitations, many of these studies focus on large, publicly traded firms.
+To my knowledge, \citeauthor{brickley2010board} (\citeyear{brickley2010board}) provide the only U.S.
+nonprofit study examining hospital boards of directors.
+Using exogenous variation in expected Medicare profits, the authors find that internal boards increase CEO compensation, while physician board members reduce public donations (\cite{brickley2010board}).
+Two additional studies examine hospital performance in other countries.
 \citeauthor{janke2019impact} (\citeyear{janke2019impact}) use data from England to study whether CEOs affect hospital production, and find no association (\cite{janke2019impact}).
-However, \citeauthor{otero2022managers} (\citeyear{otero2022managers}) investigate the role of CEOs in public hospitals in Chile, and finds an 8\% decrease in mortality rates for hospitals with top managers.
-Further, a major contributor of this quality improvement is the exodus of older physicians as CEOs (\cite{otero2022managers}).
-Finally, while not in the context of hospitals, \citeauthor{la2023impact} (\citeyear{la2023impact}) finds that OB-GYNs acquired by financially driven Physician Practice Management Companies (PPMCs) increase C-Sections compared to those acquired by clinically driven PPMCs (\cite{la2023impact}).
-Thus, the motivations of the management of health care facilities matters for provider behavior.
-I contribute to our knowledge of how executive team characteristics affect firm behaviors by (1) considering the important setting of nonprofit hospitals, which serve the majority of patients in the US, (2) considering a unique and relevant executive characteristic of the entire executive team, and (3) observing firm behavior in the presence of an exogenous incentive change.
-
-    Additionally, I contribute to our understanding of how providers respond to pay-for-performance incentives in health care.
-The most in depth study of how hospitals respond to the Hospital Readmissions Reduction Program, one of the two large pay-for-performance initiatives, is \citeauthor{gupta2021impacts} (\citeyear{gupta2021impacts}).
+However, \citeauthor{munoz2025managers} (\citeyear{munoz2025managers}) investigate the role of CEOs in public hospitals in Chile, and find an 8\% decrease in mortality rates for hospitals with top managers.
+Further, a major contributor of this quality improvement is the exodus of older physicians as CEOs (\cite{munoz2025managers}).
+Finally, although not in a hospital context, \citeauthor{la2023impact} (\citeyear{la2023impact}) find that OB-GYNs acquired by financially driven Physician Practice Management Companies (PPMCs) perform more C-sections than those acquired by clinically driven PPMCs (\cite{la2023impact}).
+Thus, the motivations of the management of health care facilities matter for provider behavior.
+This paper contributes to understanding how executive team characteristics affect firm behavior by (1) examining the critical setting of nonprofit hospitals, which serve a large share of US patients, (2) analyzing a unique and relevant characteristic of the entire executive team, and (3) observing behavior under an exogenous incentive change.
+
+    Additionally, this paper contributes to understanding how providers respond to pay-for-performance incentives in health care.
+The most in-depth study of hospital responses to the Hospital Readmissions Reduction Program (HRRP), one of the two major pay-for-performance initiatives, is by \citeauthor{gupta2021impacts} (\citeyear{gupta2021impacts}).
 This study finds that hospitals decreased readmissions by 5\% and mortality rates by 2\% on average as a result of the program, confirming prior studies (\cite{mellor2017does}; \cite{ziedan2018essays}; \cite{ody2019decreases}; \cite{gupta2021impacts}).
@@
 Further, market concentration and hospital systems play a large role in how hospitals respond to the program (\cite{kunz2024assessing}).
-Research on the other large pay-for-performance initiative, the Hospital Value Based Purchasing Program, generally finds that the program had no effect on underlying hospital quality (\cite{us2015hospital}; \cite{norton2018moneyball}; \cite{friedson2019so}).
-This paper contributes to our understanding of how characteristics of hospitals affect response to pay-for-performance incentives.
+Research on the other large pay-for-performance initiative, the Hospital Value Based Purchasing Program, generally concludes that the program had no effect on underlying hospital quality (\cite{us2015hospital}; \cite{norton2018moneyball}; \cite{friedson2019so}).
+This paper extends this literature by examining how hospital leadership characteristics shape responses to pay-for-performance incentives.
 
@@
 
-    I examine three outcomes to test this hypothesis.First, I use the CMS Case Mix Index, which provides an aggregate measure of patient complexity for each hospital and year.
+    I examine three outcomes to test this hypothesis.
+First, I use the CMS Case Mix Index, which provides an aggregate measure of patient complexity for each hospital and year.
 A higher index indicates that the hospital’s patient population requires more complex treatment.
-Additionally, I examine the share of patients treated in the hospital who have public insurance, either Medicare or Medicaid.
+Additionally, I examine the share of patients discharged from the hospital who have public insurance, either Medicare or Medicaid.
 These outcomes help assess whether certain leadership types attempt to shift costs relative to others after penalties or rewards take effect.
@@
 The results are presented in Table \ref{tab:selection}.
-All estimated treatment effects are statistically insignificant and near zero, indicating no evidence that overall hospital patient populations change differentially by leadership type.
+For Case Mix Index, clinically the estimate of -0.03 is statistically significant, but this is such a small magnitude (relative to a mean of 1.5) that this is a precise null effect.
+The estimated treatment effects for both Medicare and Medicaid populations are statistically insignificant and near zero, indicating no evidence that overall hospital patient populations change differentially by leadership type.
 This does not rule out changes in specific readmission behavior, as documented by \citeauthor{gupta2021impacts} (\citeyear{gupta2021impacts}).
@@
     While previous literature has documented correlations between executive characteristics and firm behaviors/performance, research on this topic remains limited in the health care sector.
-Health care managers are tasked with maintaining a business while also prioritizing patient care, which is why existing findings on manager effects may not extend to this setting.
-This paper investigates whether having executive leaders with clinical training impacts hospital behaviors and outcomes, particularly in response to a policy change that financially incentivizes quality.
-I utilize a unique data set of nonprofit US hospitals that captures the proportion of clinical executives, specialties of clinical executives, hospital characteristics, and hospital quality metric outcomes.
-
-    In 2012, two pay-for-performance policies went into effect that either penalized hospitals with high readmission rates, or gave payments to hospitals with low readmission/mortality rates.
-These policies were relevant to hospital finances, and gave tangible incentive for hospitals to increase quality.
-Thus, I leverage this exogenous policy change and estimate how clinically trained executives affect readmission and mortality rates in response to the change using a synthetic difference-in-differences strategy.
-I find that hospitals without clinically trained executives improve quality more than hospitals without clinically trained executives after the incentive change.
-
-    I investigate several hypotheses that arise given the estimated difference in response.
-First, this result is not driven by non-clinical teams adjusting their patient complexity.
-However, comparing non-clinical teams to for-profit hospitals reveals that it could be underlying preferences for profit vs.
-patient benefit that drive the difference.
-Finally, I leverage exogenous changes in executive teams to show that clinically trained executives are not simply a signal of underlying preferences, but manage the hospital differently.
-
-    In US health care, a very common policy goal is to improve quality of care without increasing costs.
-However, up to this point, it was unclear whether hospital leaders might play a role in attaining this.
-In this paper, I show that having hospital managers with clinical training does affect the way hospitals respond to incentives.
-Importantly, my results point to clinically trained executives already prioritizing quality before incentives were put in place.
-My findings have implications for policy makers considering financial incentives on quality, and even point to the potential of incentivizing the presence of clinically trained hospital leaders.
+Health care managers must balance business operations with patient care, so existing findings on manager effects may not fully apply to this setting.
+This paper examines whether executive leaders with clinical training influence hospital behaviors and outcomes, particularly in response to a policy change that financially incentivizes quality.
+I use a unique dataset of US nonprofit hospitals that includes the proportion and specialties of clinical executives, hospital characteristics, and quality outcomes.
+
+    In 2012, two pay-for-performance policies went into effect: one penalized hospitals with high readmission rates, and the other rewarded hospitals with low readmission or mortality rates.
+These policies significantly affected hospital finances and created strong incentives to improve quality.
+I leverage this exogenous policy change to estimate how clinically trained executives influence readmission and mortality rates using a synthetic difference-in-differences approach.
+I find that hospitals without clinically trained executives improve quality more than hospitals with clinically trained executives after the incentive change.
+
+    I explore several hypotheses to explain this differential response.
+First, I leverage the timing of clinically trained executive appointments to decompose the effect into signaling versus management.
+This decomposition shows that having a clinically trained executive in a leadership role during policy implementation is crucial, suggesting the difference reflects management practices rather than underlying hospital preferences.
+Next, I compare the two types of nonprofits to for-profit hospitals on average, revealing that non-clinical executive teams behave similarly to for-profits.
+I also examine whether the observed difference reflects the expertise of specific physician specialties by comparing internal medicine executives to others, and find evidence that this may be a contributing mechanism.
+Finally, I show that the difference is not explained by changes in aggregate patient populations or investment strategies.
+
+    In US health care, a central policy goal is to improve quality of care without increasing costs.
+However, up to this point, it has been unclear whether hospital leaders might play a role in attaining this.
+The implications of the findings in this paper are two-fold.
+First, the background of hospital leaders in the US is an important factor for hospital behaviors, particularly whether the extent to which they respond to financial incentives.
+Policymakers should consider whether trends in clinician-led hospitals are shifting, as this will influence responses to policy incentives.
+Second, although the nonprofit versus for-profit debate remains important, leadership composition may be a more critical factor.
+Ultimately, understanding how leadership expertise shapes organizational responses to policy can inform more effective strategies for aligning financial incentives with quality goals.
 
@@
 \appendix
+\setcounter{table}{0}
+\renewcommand{\thetable}{A.\arabic{table}}
+\renewcommand{\thefigure}{A.\arabic{figure}}
 
@@
 
-I use the API of ProPublica's Nonprofit Explorer to access the archive of nonprofit Tax Form 990s.\footnote{At the time of writing this, information on using version 2 of the API can be found at \hyperlink{https://projects.propublica.org/not-for-profits/api}{https://projects.propublica.org/not-for-profits/api}.} I extract the Employee Identification Number (EIN) for all nonprofits categorized with a National Taxonomy of Exempt Entities (NTEE) code of E20 (hospital), E21 (community health system), and E22 (general hospital).
-There are 5,588 EINs total in this list.
+I use the API of ProPublica's Nonprofit Explorer to access the archive of nonprofit Tax Form 990s.\footnote{At the time of writing this, information on using version 2 of the API can be found at \href{https://projects.propublica.org/not-for-profits/api}{https://projects.propublica.org/not-for-profits/api}.} I extract the Employee Identification Number (EIN) for all nonprofits categorized under a National Taxonomy of Exempt Entities (NTEE) code of E20 (hospital), E21 (community health system), and E22 (general hospital).
+There are 5,588 EINs in total in this list.
 
 I query more detailed information from the API on each hospital EIN.
-I save the name, secondary name, state, and zip code, all of which do not vary by year.
+I save the name, secondary name, state, and zip code, none of which vary by year.
 I also record URL links to the Tax Form 990 PDFs, which are unique in each year.
@@
 Finally, I manually search through unmatched EINs to identify any matches.
-From google searches, I identify an additional 300 EIN-AHA ID matches.
-yielding a total of 1200 EINs.
-
-While it would be ideal to analyze all nonprofit hospitals, a sample size of 852 firms is relatively large in the hospital executive literature.
+From Google searches, I identify an additional 300 EIN-AHA ID matches.
+yielding a total of 1,200 EINs.
+
+Although analyzing all nonprofit hospitals would be ideal, a sample size of 852 firms is considered relatively large in the executive literature.
 I present a comparison of in-sample and out-of-sample nonprofits.
@@
 
-I then extract the names of board members and executives from the Tax Form 990 PDFs of matched hospitals.
+I then extract the names of board members and executives from the Tax Form 990 PDFs for matched hospitals.
 In the data set of hospital PDF URLs that I collected earlier, I limit to the hospitals with solid matches described above.
@@
 
-One aspect of the NonProfit Explorer API is that, only in some cases, if two forms are present for an EIN, year, only the first one (which is typically not the one with the relevant information) is pulled.
+One aspect of the NonProfit Explorer API is that in some cases, if two forms are present for an EIN-year, only the first one is pulled.
 Therefore, for some hospitals, a couple years will have gaps in text extraction data.
-I locate EIN, years where this problem is occurring, and a team of RAs locates and downloads the correct forms manually.
+I locate EIN-years where this problem is occurring, and a team of RAs locates and downloads the correct forms manually.
 I extract text from these manually downloaded forms in the same manner as above.
@@
 
-The data is now an executive-year level data set with information on clinical training and specialty when relevant.
-I then create hospital level indicators based on the names, titles and positions associated with the hospital: the number of clinically trained executives, the number of executives with particular specialties, the number of total executives, and whether the hospital employs a CMO.
+The data is now an executive-year-level data set with information on clinical training and specialty when relevant.
+I then create hospital-level indicators based on the names, titles and positions associated with the hospital: the number of clinically trained executives, the number of executives with particular specialties, the number of total executives, and whether the hospital employs a CMO.
 
@@
 I explore the exogeneity of such changes in Section \ref{sec:changes}.
-Additionally, I present the number of hospitals who change their propensity to hire a clinically trained executive at different times in Table \ref{tab:change_timing}.
+Additionally, I present the frequency of hospitals who change their propensity to hire a clinically trained executive at different times in Table \ref{tab:change_timing}.
 I break the timing up into pre-2012, 2012, and post-2012.
@@
 The number of hospitals that only have an MD executive in 2012, but at no other time, is only 6.
-The majority of hospitals (187) who ever have a clinical executive have one in every time period.
+The majority of hospitals (197) who ever have a clinical executive have one in every time period.
 
@@
 \subsection{Additional Summary Statistics}\label{app:sumstats}
+
+In this section, I present several additional figures and tables providing more information on the data.
+In Figure \ref{fig:state_doc}, I show that clinical executives are not concentrated in one region geographically.
+In Table \ref{tab:fp_samples_stable}, I provide means of all relevant variables for for-profit hospitals.
+In Table \ref{tab:specialty}, I show means of all relevant variables for hospital with and without internal medicine executives.
+In Figures \ref{fig:main_read_control_weights} and \ref{fig:main_read_time_weights}, I provide the time and control weights from the synthetic difference in differences estimation with readmission rate as the outcome.
+In Figures \ref{fig:main_mort_control_weights} and \ref{fig:main_mort_time_weights}, I provide the time and control weights from the synthetic difference in differences estimation with mortality rate as the outcome.
 
@@
 
-    \subsection{Binned Treatment}\label{app:binned}
+    \subsection{Intensive Margin, Controlling for Num. Executives}\label{app:binned}
 
@@
 
-    \end{figure}
-
-    \subsection{Selective Patient Practices}\label{app:casemix}
-
-    A method hospitals have been shown to employ in order to decrease readmission rates is being more selective with which patients to admit to the hospital (\cite{gupta2021impacts}), as only admitting healthy patients mechanically decreases readmission rates.
-One hypothesis behind a more drastic quality decrease by non-clinical teams is that they may be more willing to participate in such practices, while executives with clinical training are less likely to turn patients away.
-
-    To test this, I employ case mix index as the outcome in specification \ref{eq:clinical} to determine whether one type of hospital differentially changes patient composition after pay-for-performance incentives, where the same identification assumptions hold.
-I present the estimated difference in Figure \ref{fig:main_cmi_clinical}, comparing hospitals that never have a clinical executive to hospitals that always have a clinical executive.
-Neither type of hospital drastically changes case mix index after the incentive change, and there is no estimated differential response in complexity of patients seen in the hospital.
-Thus, the estimated difference in readmission rate response between clinical and non-clinical executives is not due to differentially changing case mix.
-
-    \begin{figure}[ht!]
-        \centering
-        \caption{Effect of Clinical Executive on Case Mix Index}
-        \includegraphics[width=0.5\textwidth]{Objects/cmi_md_nomd_synth_graph.pdf}
-        \label{fig:main_cmi_clinical}
     \end{figure}

### 2025-11-14 RScripts/appendix_tables_and_figures.R
update scripts. project is now up to date at time of first submission Nov 2025

@@ -822,57 +822,6 @@ cont_abovemed_mort_plot_size <- synthdid_plot(cont_abovemed_mort_did_size, facet
 ggsave(cont_abovemed_mort_plot_size, filename=paste0(objects_path,"cont_abovemedmort_md_nomd_size_synth_graph.pdf"), width=5.5, height=6, units="in")
 
 
-# FIGURE 16: case mix index as outcome ####
-main_analysis_data <- hospital_data %>%
-  filter(no_num_md_change_2010_2014==1 & !is.na(ever_has_md)) %>%
-  mutate(never_has_md=ifelse(ever_has_md==1,0,1),
-         post_2012 = ifelse(year>2012,1,0)) %>%
-  mutate(did=post_2012*ever_has_md) %>%
-  mutate(count=1) %>%
-  group_by(ID, year) %>%
-  mutate(sum=sum(count)) %>%
-  ungroup() %>%
-  filter(!(sum==2 & is.na(weightedavg_read))) %>%
-  select(-count, -sum)
-
-cmi_data <- main_analysis_data %>%
-  filter(year!=2015) %>%
-  select(ID, year, cmiv, did) %>%
-  na.omit() %>%
-  mutate(count=1) %>%
-  group_by(ID) %>%
-  mutate(min=min(year[did==1])) %>%
-  mutate(sum=sum(count)) %>%
-  ungroup() %>%
-  filter(sum==5) %>%
-  filter(min!=2014) %>%
-  select(-sum, -count, -min) %>%
-  mutate(ID=as.factor(ID))
-cmi_panel <- panel.matrices(as.data.frame(cmi_data))
-
-cmi_did <- synthdid_estimate(cmi_panel$Y, cmi_panel$N0, cmi_panel$T0)
-
-cmi_se = sqrt(vcov(cmi_did))
-
-main_cmi_plot <- synthdid_plot(cmi_did, facet.vertical=FALSE,
-                               control.name='Never Clinical Exec', treated.name='Always Clinical Exec',
-                               lambda.comparable=TRUE, se.method = 'none',
-                               lambda.plot.scale = 0,
-                               trajectory.linetype = 1, line.width=.8, effect.curvature=-.4,
-                               trajectory.alpha=1, effect.alpha=0,
-                               diagram.alpha=0, onset.alpha=1) +
-  theme(legend.position=c(.5,-.095), legend.direction='horizontal',
-        legend.key=element_blank(), legend.background=element_blank(),
[…]

### 2025-11-14 RScripts/data10_create_leadership_variables.R

@@ -81,7 +81,7 @@ cleaned_text <- cleaned_text %>%
 cleaned_text <- cleaned_text %>%
   mutate(cmo = ifelse(position1=="cmo" | position2 == "cmo" | position1 == "chief medical officer" | position2 == "chief medical officer", 1,0)) %>%
   mutate(cmo = ifelse(is.na(cmo),0,cmo))
-  
+
 executive_data <- cleaned_text %>%
   mutate(vp = ifelse(pos1_vp==1 | pos2_vp==1, 1, 0)) %>%
   filter(position == "executive") %>%
@@ -107,7 +107,7 @@ for (i in seq_along(eins)){
   data <- complete(data, year=2009:2016, id)
   
   max_id <- max(data$id)
-
+  
   for (j in 1:max_id){
     name_data <- data %>%
       filter(id==j) %>%
@@ -278,8 +278,6 @@ names_data <- names_data %>%
 
 # manually add EINs of docs I can find ####
 executive_names <- executive_names %>%
-  mutate(npi=npi.x, t_class=t_class.x) %>%
-  select(-npi.x, -npi.y, -t_class.x, -t_class.y) %>%
   mutate(npi=ifelse(ein=="10198331" & name=="katheryn resinbrink", 1124067194, npi),
          npi=ifelse(ein=="10211494" & name=="richard goldstein", 1811967698, npi),
          npi=ifelse(ein=="10211503" & name=="david hyde", 1144288341, npi),
@@ -784,7 +782,10 @@ executive_names <- executive_names %>%
 # change doctor of NPIs that didn't match
 executive_names <- executive_names %>%
   mutate(doctor=ifelse(doctor==1 & is.na(npi),0,doctor)) %>%
-  mutate(doctor=ifelse(is.na(doctor),0,doctor))
+  mutate(doctor=ifelse(is.na(doctor),0,doctor)) %>%
+  mutate(doc_nurse = ifelse(title=="rn",1,NA)) %>%
+  mutate(doc_nurse = ifelse(is.na(doc_nurse) & doctor==1,1,doc_nurse)) %>%
+  mutate(doc_nurse = ifelse(is.na(doc_nurse),0,doc_nurse))
 
 # join specialty to columns that don't already have it
 npidata <- npidata %>%
@@ -795,7 +796,7 @@ executive_names <- executive_names %>%
 executive_names <- executive_names %>%
   mutate(t_class=ifelse(is.na(t_class), t_class2, t_class)) %>%
   select(-t_class2)
[…]

### 2025-11-14 RScripts/data11_combine_hosp_level_data.R

@@ -13,7 +13,7 @@ library(MatchIt)
 AHA_ein_matches <- read_rds(paste0(created_data_path,"AHA_ein_matches.rds"))
 manual_matched_eins <- read_rds(paste0(created_data_path, "manual_matched_eins.rds"))
 # Data from AHA 
-AHA <- read_csv(paste0(raw_data_path, "/allAHAvariables_20092017.csv"))
+AHA <- read_csv(paste0(raw_data_path, "/AHAdata_20052023.csv"))
 # HCRIS data that has HRRP penalty information (and beds)
 HCRIS <- read_csv(paste0(raw_data_path, "/final_HCRIS_data.csv"))
 # Hospital Compare data that also has penalty information
@@ -22,7 +22,7 @@ hc_readm_2013 <- read_csv(paste0(raw_data_path, "/Hospital Compare/vwhqi_readm_r
 hc_readm_2014 <- read_csv(paste0(raw_data_path, "/Hospital Compare/vwhqi_readm_reduction_2014.csv"))
 hc_readm_2015 <- read_csv(paste0(raw_data_path, "/Hospital Compare/vwhqi_readm_reduction_2015.csv"))
 # Data that I created on leadership teams from the tax forms
-ein_leadership_data <- read_rds(paste0(created_data_path, "ein_leadership_changes_data4.rds"))
+ein_leadership_data <- read_rds(paste0(created_data_path, "ein_leadership_changes_data.rds"))
 # Data that I created on outcomes from Hospital Compare
 outcomes_data <- read_rds(paste0(created_data_path, "hosp_outcomes.rds"))
 # read in case mix index
@@ -53,10 +53,10 @@ impact2015 <- read_csv(paste0(raw_data_path, "/HospCaseMix/impact2015.csv"))%>%
 # start with AHA hospitals to include both for-profit and not-for-profit
 AHA <- AHA %>%
   filter(FSTCD<=56 & (CNTRL==23 | (CNTRL>=12 & CNTRL<=16) | (CNTRL>=31 & CNTRL<=33)) & SERV==10)
-  # 4,475 hospitals
+# 4,475 hospitals
 
 AHA <- AHA %>%
-  select(ID, MCRNUM, YEAR, CNTRL, MSTATE, MAPP5, MAPP8, FTMT, PHYGP, FTRNTF, SUBS, SYSID, EHLTH, MNGT)
[…]

### 2025-11-14 RScripts/paths.R

@@ -1,0 +1,3 @@
+created_data_path <- "C:\\Users\\uqhglen2\\Documents\\ResearchProjects\\Executives_P4P\\CreatedData\\"
+raw_data_path <- "C:\\Users\\uqhglen2\\Documents\\ResearchProjects\\RawData\\"
+objects_path <- "C:\\Users\\uqhglen2\\Documents\\ResearchProjects\\Executives_P4P\\Objects\\"

### 2025-11-14 RScripts/tables_and_figures.R

@@ -9,14 +9,20 @@ library(synthdid)
 library(abind)
 library(reshape2)
 library(data.table)
+library(stringr)
+library(tidyr)
+library(fixest)
+
+# don't show scientific notation
+options(scipen=999)
 
 options(knitr.kable.NA=" ")
 group.colors <- c("For-profit" = "#D65828", "Never Clinical Exec" = "#C2DEF2", "Ever Clinical Exec" ="#6BAED6", "Always Clinical Exec" = "#2C6B8E")
- 
+
 # Produce all tables and figures for the body of the paper #########
 
 # read in data
-hospital_data <- read_rds(paste0(created_data_path, "all_hospital_data3.rds"))
+hospital_data <- read_rds(paste0(created_data_path, "all_hospital_data.rds"))
 doc_exec_data <- read_rds(paste0(created_data_path, "doc_exec_data.rds"))
 
 hospital_data <- hospital_data %>%
@@ -29,14 +35,32 @@ hospital_data <- hospital_data %>%
          ever_pen_all = ifelse(ever_pen_hf>=1 & ever_pen_ha>=1 & ever_pen_pnem>=1,1,0)) %>%
   mutate(frac_doc_exec = num_doctors/num_execs)
 
+# create variable for penalized in 2012
+hospital_data <- hospital_data %>%
+  mutate(penalized_2012 = ifelse(year==2012 & (pen_ha==1 | pen_hf==1 | pen_pnem==1),1,NA)) %>%
+  group_by(ID) %>%
+  tidyr::fill(penalized_2012, .direction="downup") %>%
+  ungroup() %>%
+  mutate(penalized_2012 = ifelse(is.na(penalized_2012),0,penalized_2012))
+
+hospital_data <- hospital_data %>%
+  mutate(baseline_cmi = ifelse(year==2010,cmiv,NA)) %>%
+  group_by(ID) %>%
+  fill(baseline_cmi, .direction="downup") %>%
+  ungroup()
+
+
 
 # TABLE 1: summary statistics of individual executives who are clinically trained ####
 doc_exec_data <- doc_exec_data %>%
-  mutate(surgery_doc = ifelse(t_class == "Surgery", 1, 0),
+  mutate(surgery_doc = ifelse(stringr::str_detect(t_class,"Surgery"), 1, 0),
          anest_doc = ifelse(t_class == "Anesthesiology", 1, 0),
          family_doc = ifelse(t_class == "Family Medicine",1,0),
[…]
